### 2024-11-28 paper/main.tex
add very basic versions of initial paper draft

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[hidelinks]{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsmath,amssymb,amsthm}
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{mathrsfs}
\usepackage{onimage}
\usepackage{graphicx}
\usepackage{doi}
\usepackage{acronym}
\usepackage{listings}
\usepackage{tikz}

\usetikzlibrary{trees}

\title{
    A mechanical Model of Rod-Shaped Bacteria
}

%\date{September 9, 1985}	% Here you can change the date presented in the paper title
%\date{} 					% Or removing it

\author{
    \href{https://orcid.org/0009-0001-0613-7978}{
        \includegraphics[scale=0.06]{orcid.pdf}
        \hspace{1mm}Jonas Pleyer
    }
    \thanks{
        \href{https://jonas.pleyer.org}{jonas.pleyer.org},
        \href{https://cellular-raza.com}{cellular-raza.com}
    }\\
	Freiburg Center for Data-Analysis and Modeling\\
	University of Freiburg\\
	\texttt{jonas.pleyer@fdm.uni-freiburg.de} \\
	%% examples of more authors
	\And
	\href{https://orcid.org/0000-0002-6371-4495}{
        \includegraphics[scale=0.06]{orcid.pdf}
        \hspace{1mm}Christian Fleck
    }\\
	Freiburg Center for Data-Analysis and Modeling\\
	University of Freiburg
}

% Uncomment to remove the date
%\date{}

% Uncomment to override  the `A preprint' in the header
\renewcommand{\headeright}{Preprint}
%\renewcommand{\undertitle}{Technical Report}
\renewcommand{\shorttitle}{A mechanical Model of Rod-Shaped Bacteria}

\usepackage{enumitem}
\setlist{nolistsep}

%%% Add PDF metadata to help others organize their library
%%% Once the PDF is generated, you can check the metadata with
%%% $ pdfinfo template.pdf
\hypersetup{
pdftitle={% TODO
},
pdfsubject={q-bio.NC, q-bio.QM},
pdfauthor={Jonas Pleyer, Christian Fleck},
pdfkeywords={},
}

% Change numbering of equations
% \numberwithin{equation}{section}

% MAKE TITLES IN THEOREMS BOLD
\makeatletter
\def\th@plain{%
  \thm@notefont{}% same as heading font
  \itshape % body font
}
\def\th@definition{%
  \thm@notefont{}% same as heading font
  \normalfont % body font
}
\makeatother

\begin{document}
\maketitle

%###################################################################################################
\begin{abstract}
    This paper develops the formal notions of cellular \dots
\end{abstract}

% keywords can be removed
% \keywords{}

%###################################################################################################
\section{Introduction}
\label{section:introduction}

%###################################################################################################
\section{A mechanical Model of Rod-Shaped Bacteria}
\label{section:mechanical-model-rod-shaped-bacteria}

Bacteria exist in many physical shapes such as spheroidal, rod-shaped or spiral
\cite{Zapun2008,Young2006}.

To model the spatial mechanics of elongated bacteria \cite{Billaudeau2017}, we represent them as a
collection of auxiliary vertices $\{\vec{v}_i\}$ which are connected by springs in
ascending order.
Furthermore, we assume that the cells are flexible described by their rigidity.
A force $\vec{F}$ interacting between cellular agents determines the radius (thickness) of the
rods and an attractive component can model adhesion between cells.

% --------------------------------------------------------------------------------------------------
\subsection{Mechanics}

%TODO use these citations
\cite{Chatterjee1988},\cite{Ursell2014}

In principle we can assign individual lengths $\{l_i\}$ and strengths $\{\gamma\}_i$ to each spring.
The internal force acting on vertex $\vec{v}_i$ can be divided into 2 contributions coming
from the 2 springs pulling on it.
In the case when $i=0,N_\text{vertices}$, this is reduced to only one internal component.
We denote with $\vec{c}_{i}$ the connection between two vertices

\begin{equation}
    \vec{c}_i = \vec{v}_{i}-\vec{v}_{i-1}
\end{equation}

and can write down the resulting force

\begin{align}
    \vec{F}_{i,\text{springs}} =
        &-\gamma_i\left(1 - \frac{l_i}{\left|\vec{c}_i\right|}\right)\vec{c}_i\\
        &+ \gamma_{i+1}\left(1 - \frac{l_{i+1}}{\left|\vec{c}_{i+1}\right|}\right)\vec{c}_{i+1}
    \label{force-springs}
\end{align}

In addition to springs between individual vertices $\vec{v}_i$, we assume that each angle at
a vertex between two other is subject to a curvature force.
Assuming that $\alpha_i$ is the angle between the connections and
$\vec{d}_i=\vec{c}_i/|\vec{c}_i|$ is the normalized connection,
we can write down the forces acting on vertices $\vec{v}_i,\vec{v}_{i-1},\vec{v}_{i+1}$

\begin{align}
    \vec{F}_{i,\text{curvature}} &= \eta_i\tan\left(\pi-\alpha_i\right)
        \frac{\vec{d}_i - \vec{d}_{i+1}}{|\vec{d}_i-\vec{d}_{i+1}|}\\
    \vec{F}_{i-1,\text{curvature}} &= -\frac{1}{2}\vec{F}_{i,\text{curvature}}\\
    \vec{F}_{i+1,\text{curvature}} &= -\frac{1}{2}\vec{F}_{i,\text{curvature}}
    \label{force-curvature}
\end{align}

where $\eta_i$ is the curvature at vertex $\vec{v}_i$.
We can see that the curvature force does not move the overall center of the cell in space.

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]
        {../docs/source/_static/mechanics.png}
    \caption{Example of a polygonal representation of a rod-shaped bacterium.}
    \label{fig:mechanics-bacterium}
\end{figure}

The total force is the sum of external and interal forces.

\begin{equation}
    \vec{F}_{i,\text{total}} = \vec{F}_{i,\text{springs}}+ \vec{F}_{i,\text{curvature}} + \vec{F}_{i,\text{external}}
   \label{force-total}
\end{equation}

and are integrated via

\begin{align}
    \partial_t^2 \vec{x} &= \partial\vec{x} + D\vec{\xi}\\
    \partial_t\vec{x} &= \vec{F}_\text{total}
    \label{equations-of-motion}
\end{align}

where $D$ is the diffusion constant and  $\vec{\xi}$ is the wiener process (compare with
brownian motion such as given by the
% Brownian3D <https://cellular-raza.com/docs/cellular_raza_building_blocks/struct.Brownian3D.html>_
% struct of cellular-raza <https://cellular-raza.com>.

% --------------------------------------------------------------------------------------------------
\subsection{Interaction}

When calculating forces acting between the cells, we can use a simplified model to circumvent the
numerically expensive integration over the complete length of the rod.
Given a vertex $\vec{v}_i$ on one cell, we calculate the closest point $\vec{p}$ on the polygonal
line given by the vertices $\{\vec{w}_j\}$ of the interacting cell.
Furthermore we determine the value $q\in[0,1]$ such that

\begin{equation}
    \vec{p} = (1-q)\vec{w}_j + q\vec{w}_{j+1}
   \label{connection}
\end{equation}

for some specific $j$.
The force is then calculated between the points $\vec{v}_i$ and $\vec{p}_i$ and acts on
the vertex $\vec{w}_i,\vec{w}_{i+1}$ with relative strength $(1-q)$ and $q$.

\begin{equation}
    \vec{F}_{i,\text{External}} = \vec{F}(\vec{v}_i,\vec{p})
    \label{force-external}
\end{equation}

% .. _ TODO MORSEPOTENTIAL

% .. _ TODO insert table with all parameters

% --------------------------------------------------------------------------------------------------
\subsection{Cycle}

To simulate proliferation, we introduce a growth term for the spring lengths $l_i$

\begin{equation}
    \partial_t l_i = \mu
    \label{growth-ode}
\end{equation}

which will increase the length of the cell indefenitely unless we introduce a
% `division event <https://cellular-raza.com/internals/concepts/cell/cycle>`_.
We define a threshold (in our case double of the original length) for the total length of the
cell at which it divides.
To construct a new cell, we cannot simply copy the existing one twice, but we also need to adjust
internal parameters in the process.
The following actions need to be taken for the old and new agent.

% .. _ TODO

1. Assign a new growth rate (pick randomly from uniform distribution in $[0.8\mu_0,1.2\mu_0]$
   where $\mu_0$ is some fixed value)
2. Assign new positions
    1. Calculate new spring lengths
        $\tilde{l}_i = l_i\left(\frac{1}{2} - \frac{r}{\sum\limits_i l_i}\right)$
    2. Calculate middle of old cell
        $\vec{m} = \frac{1}{N_\text{vertices}}\sum\limits_i\vec{v}_i$
    3. Calculate positions of new vertices $\vec{w}_i$

\begin{align}
    q_i &= \frac{i}{N_\text{vertices}}\\
    \vec{w}_{i,\text{new},\pm} &= (1-q_i)\vec{m} + q_i(\vec{v}_{\pm\text{start}} - \vec{m})
\end{align}

%###################################################################################################
\section{Generating Masks and Images}
\label{section:generating-masks-and-images}

\begin{figure}
    \centering
    \begin{tikzonimage}[width=0.3\textwidth]
        {../docs/source/_static/09395645494836445480/raw_pv/000000400.png}
        \node at (0.1, 0.9)[anchor=north west, rectangle, draw, white, minimum width=20pt, minimum height=20pt]{a)};
    \end{tikzonimage}
    \includegraphics[width=0.3\textwidth]
        {../docs/source/_static/09395645494836445480/images/000000400.png}
    \includegraphics[width=0.3\textwidth]
        {../docs/source/_static/09395645494836445480/masks/000000400.png}
    \caption{Snapshots at \dots}
    \label{fig:progression-image-generation}
\end{figure}

%###################################################################################################
\section{Extract Positions from Masks}
\label{section:fitting-and-comparing-masks}

\begin{figure}
    \centering
    \includegraphics[width=0.3\textwidth]
        {../docs/source/_static/fitting-methods/progressions-1.png}
    \includegraphics[width=0.3\textwidth]
        {../docs/source/_static/fitting-methods/progressions-2.png}
    \linebreak
    \includegraphics[width=0.3\textwidth]
        {../docs/source/_static/fitting-methods/progressions-3.png}
    \includegraphics[width=0.3\textwidth]
        {../docs/source/_static/fitting-methods/progressions-4.png}
    \caption{1}
    \label{fig:mask-difference-metric}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.32\textwidth]
        {../docs/source/_static/fitting-methods/extract_positions-000400.png}
    \includegraphics[width=0.32\textwidth]
        {../docs/source/_static/fitting-methods/extract_positions-000600.png}
    \includegraphics[width=0.32\textwidth]
        {../docs/source/_static/fitting-methods/extract_positions-001200.png}
    \caption{Progression of generated masks with estimated and exact positions.}
    \label{fig:position-extraction}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]
        {../docs/source/_static/fitting-methods/penalty-time-flow.png}%
    \includegraphics[width=0.5\textwidth]
        {../docs/source/_static/fitting-methods/displacement-calculations.png}
    \caption{Comparison of masks across generations. Compare estimated with exact rod lengths.}
    \label{fig:benchmarking-fitting-algorithm}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.32\textwidth]
        {../docs/source/_static/fitting-methods/estimate-parameters1/Growth Rate.png}%
    \includegraphics[width=0.32\textwidth]
        {../docs/source/_static/fitting-methods/estimate-parameters1/Potential Stiffness.png}%
    \includegraphics[width=0.32\textwidth]
        {../docs/source/_static/fitting-methods/estimate-parameters1/Radius.png}\\%
    \includegraphics[width=0.32\textwidth]
        {../docs/source/_static/fitting-methods/estimate-parameters1/Rigidity.png}%
    \includegraphics[width=0.32\textwidth]
        {../docs/source/_static/fitting-methods/estimate-parameters1/Strength.png}%
    \caption{Snapshots at \dots}
    \label{fig:parameter-estimates-single-step}
\end{figure}

%###################################################################################################
\section{Estimating Parameters}
\label{section:parameter-estimation}

\bibliographystyle{IEEEtran}
\bibliography{../docs/source/references}

\end{document}

### 2025-01-29 plots/param_estim_single_step.py
print timestamps between every major step

@@ -3,7 +3,8 @@ import cr_mech_coli as crm
 import numpy as np
 import matplotlib.pyplot as plt
 import scipy as sp
-from tqdm import tqdm
+import time
+
 
 def predict(
     # Parameters
@@ -98,6 +99,7 @@ def predict_flatten(
 
 
 if __name__ == "__main__":
+    interval = time.time()
     # markers = np.fromfile("./data/growth-2-marked/image001042-markers.tif").reshape(576, 768)
     mask1 = np.loadtxt("data/growth-2-marked/image001042-markers.csv", delimiter=",")
     img1 = imread("data/growth-2/image001042.png")
@@ -112,6 +114,9 @@ if __name__ == "__main__":
     ax[0, 1].imshow(img2)
     ax[1, 0].imshow(mask1)
     ax[1, 1].imshow(mask2)
+
+    print("[x] +[{:8.3}] Generated initial plots".format(time.time() - interval))
+    interval = time.time()
 
     domain_size = np.max(mask1.shape)
     n_agents = len(pos1)
@@ -146,6 +151,10 @@ if __name__ == "__main__":
         recombination=0.3,
         popsize=128,
     )
+    print(
+        "[x] +[{:8.4}s] Finished Parameter Optimization".format(time.time() - interval)
+    )
+    interval = time.time()
 
     param_infos = [
         ("Growth Rate", "\\mu m\\text{min}^{-1}"),
@@ -178,6 +187,9 @@ if __name__ == "__main__":
         )
         plt.close(fig2)
 
+    print("[x] +[{:8.3}] Plotted Profiles".format(time.time() - interval))
+    interval = time.time()
+
     cell_container = predict_flatten(
         res.x,
         *args,
@@ -206,3 +218,5 @@ if __name__ == "__main__":
         ax[0, 1].plot(p[:,0], p[:,1], color="white")
     fig.tight_layout()
     plt.show()
+
+    print("[x] +[{:8.3}] Rendered Masks".format(time.time() - interval))

### 2025-01-29 plots/param_estim_single_step.py
initialize via new interaction field of agent

@@ -39,13 +39,15 @@ def predict(
             vel=np.zeros(
                 (positions.shape[1], positions.shape[2] + 1), dtype=np.float32
             ),
+            interaction=crm.MorsePotentialF32(
+                radius=radius,
+                potential_stiffness=potential_stiffness,
+                cutoff=cutoff,
+                strength=strength,
+            ),
             growth_rate=growth_rate,
-            radius=radius,
             spring_length=pos_to_spring_length(positions[i]),
             spring_length_threshold=1000,
-            strength=strength,
-            potential_stiffness=potential_stiffness,
-            cutoff=cutoff,
             rigidity=rigidity,
         )
         for i in range(n_agents)

### 2025-01-29 plots/param_estim_single_step.py
create individual plots and not one big shared one

@@ -111,11 +111,15 @@ if __name__ == "__main__":
     pos1 = np.array(crm.extract_positions(mask1, n_vertices))
     pos2 = np.array(crm.extract_positions(mask2, n_vertices))
 
-    fig, ax = plt.subplots(3, 2, figsize=(12, 12))
-    ax[0, 0].imshow(img1)
-    ax[0, 1].imshow(img2)
-    ax[1, 0].imshow(mask1)
-    ax[1, 1].imshow(mask2)
+    figs_axs = [plt.subplots() for _ in range(4)]
+    figs_axs[0][1].imshow(img1)
+    figs_axs[0][1].set_axis_off()
+    figs_axs[1][1].imshow(img2)
+    figs_axs[1][1].set_axis_off()
+    figs_axs[2][1].imshow(mask1)
+    figs_axs[2][1].set_axis_off()
+    figs_axs[3][1].imshow(mask2)
+    figs_axs[3][1].set_axis_off()
 
     print("[x] +[{:8.3}] Generated initial plots".format(time.time() - interval))
     interval = time.time()
@@ -208,17 +212,19 @@ if __name__ == "__main__":
     mask_gen2 = crm.render_mask(
         agents_predicted, cell_container.cell_to_color, domain_size
     )
-    ax[2, 0].imshow(mask_gen1[::-1])
-    ax[2, 1].imshow(mask_gen2[::-1])
 
     for p in pos1:
-        ax[0, 0].plot(p[:,0], p[:,1], color="white")
-    # for p in pos2:
-    #     ax[0, 1].plot(p[:,0], p[:,1], color="white")
+        figs_axs[0][1].plot(p[:, 0], p[:, 1], color="white")
     for agent, _ in agents_predicted.values():
         p = agent.pos
-        ax[0, 1].plot(p[:,0], p[:,1], color="white")
-    fig.tight_layout()
-    plt.show()
+        figs_axs[1][1].plot(p[:, 0], p[:, 1], color="white")
+
+    for i, (fig, _) in enumerate(figs_axs):
+        fig.tight_layout()
+        fig.savefig(
+            "docs/source/_static/fitting-methods/estimate-parameters1/microscopic-images-{}.png".format(
+                i
+            )
+        )
 
     print("[x] +[{:8.3}] Rendered Masks".format(time.time() - interval))

### 2025-01-29 plots/param_estim_single_step.py
do not use n_agents as arg

@@ -125,9 +125,8 @@ if __name__ == "__main__":
     interval = time.time()
 
     domain_size = np.max(mask1.shape)
-    n_agents = len(pos1)
-    cutoff=30.0
-    args=(n_agents, cutoff, domain_size, pos1, pos2)
+    cutoff = 30.0
+    args = (cutoff, domain_size, pos1, pos2)
 
     growth_rate = 0.03
     radius = 8.0

### 2025-01-29 plots/param_estim_single_step.py
adjust parameter and optimization values to achieve very nice result

@@ -131,17 +131,17 @@ if __name__ == "__main__":
     growth_rate = 0.03
     radius = 8.0
     strength = 0.1
-    potential_stiffness = 0.2
+    potential_stiffness = 0.4
     rigidity = 0.8
     parameters = (growth_rate, radius, strength, potential_stiffness, rigidity)
 
     # Optimize values
     bounds = [
-        [0.00, 0.05],# Growth Rate
-        [4.0, 10.0],# Radius
-        [0.1, 0.4],# Strength
-        [0.2, 0.6],# Potential Stiffness
-        [0.4, 1.0],# Rigidity
+        [0.00, 0.05],  # Growth Rate
+        [4.0, 10.0],  # Radius
+        [0.1, 0.4],  # Strength
+        [0.25, 0.55],  # Potential Stiffness
+        [0.4, 1.0],  # Rigidity
     ]
     res = sp.optimize.differential_evolution(
         predict_flatten,
@@ -149,10 +149,10 @@ if __name__ == "__main__":
         x0=parameters,
         args=args,
         workers=-1,
-        updating='deferred',
-        maxiter=20,
+        updating="deferred",
+        maxiter=100,
         disp=True,
-        tol=1e-3,
+        tol=1e-4,
         recombination=0.3,
         popsize=128,
     )

### 2025-01-29 plots/param_estim_single_step.py
use interaction directly instead of going through individual parameters for prediction method

@@ -9,11 +9,8 @@ import time
 def predict(
     # Parameters
     growth_rate: float,  # Shape (N)
-    radius: float,
-    strength: float,
-    cutoff: float,
-    potential_stiffness: float,
     rigidity: float,
+    interaction,
     # Constants
     positions: np.ndarray,  # Shape (N, n_vertices, 3)
     domain_size: float,
@@ -39,12 +36,7 @@ def predict(
             vel=np.zeros(
                 (positions.shape[1], positions.shape[2] + 1), dtype=np.float32
             ),
-            interaction=crm.MorsePotentialF32(
-                radius=radius,
-                potential_stiffness=potential_stiffness,
-                cutoff=cutoff,
-                strength=strength,
-            ),
+            interaction=interaction,
             growth_rate=growth_rate,
             spring_length=pos_to_spring_length(positions[i]),
             spring_length_threshold=1000,
@@ -61,6 +53,17 @@ def predict(
         return np.array([agent.pos for agent, _ in agents_predicted.values()])
 
 
+def reconstruct_morse_potential(parameters):
+    (growth_rate, rigidity, radius, strength, potential_stiffness) = parameters
+    interaction = crm.MorsePotentialF32(
+        radius=radius,
+        potential_stiffness=potential_stiffness,
+        cutoff=cutoff,
+        strength=strength,
+    )
+    return (growth_rate, rigidity, interaction)
+
+
 def predict_flatten(
     parameters,
     cutoff,
@@ -69,14 +72,11 @@ def predict_flatten(
     pos_final,
     return_cells: bool = False,
 ):
-    (growth_rate, radius, strength, potential_stiffness, rigidity) = parameters
+    growth_rate, rigidity, interaction = reconstruct_morse_potential(parameters)
     pos_predicted = predict(
         growth_rate,
-        radius,
-        strength,
-        cutoff,
-        potential_stiffness,
         rigidity,
+        interaction,
         pos_initial,
         domain_size,
         return_cells=return_cells,
@@ -133,15 +133,15 @@ if __name__ == "__main__":
     strength = 0.1
     potential_stiffness = 0.4
     rigidity = 0.8
-    parameters = (growth_rate, radius, strength, potential_stiffness, rigidity)
+    parameters = (growth_rate, rigidity, radius, strength, potential_stiffness)
 
     # Optimize values
     bounds = [
[…]

### 2025-01-29 cr_mech_coli/fitting.py
use function correctly when shape of array is (xdim, ydim)

@@ -143,7 +143,10 @@ def extract_positions(mask: np.ndarray, n_vertices: int = 8) -> list[np.ndarray]
         )
     colors = filter(lambda x: np.sum(x) != 0, np.unique(m, axis=0))
 
-    cell_masks = [np.all(mask == c, axis=2) for c in colors]
+    if len(mask.shape) > 2:
+        cell_masks = [np.all(mask == c, axis=2) for c in colors]
+    else:
+        cell_masks = [mask == c for c in colors]
     skeleton_points = [
         _sort_points(sk.morphology.skeletonize(m, method="lee")) for m in cell_masks
     ]

### 2025-01-29 paper/main.tex
use tikzonimage to refine figures

@@
         {../docs/source/_static/09395645494836445480/raw_pv/000000400.png}
-        \node at (0.1, 0.9)[anchor=north west, rectangle, draw, white, minimum width=20pt, minimum height=20pt]{a)};
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{a}};
     \end{tikzonimage}
-    \includegraphics[width=0.3\textwidth]
+    \begin{tikzonimage}[width=0.3\textwidth]
         {../docs/source/_static/09395645494836445480/images/000000400.png}
-    \includegraphics[width=0.3\textwidth]
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{b}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.3\textwidth]
         {../docs/source/_static/09395645494836445480/masks/000000400.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{c}};
+    \end{tikzonimage}
     \caption{Snapshots at \dots}
@@
     \centering
-    \includegraphics[width=0.3\textwidth]
-        {../docs/source/_static/fitting-methods/progressions-1.png}
-    \includegraphics[width=0.3\textwidth]
-        {../docs/source/_static/fitting-methods/progressions-2.png}
-    \linebreak
-    \includegraphics[width=0.3\textwidth]
-        {../docs/source/_static/fitting-methods/progressions-3.png}
-    \includegraphics[width=0.3\textwidth]
-        {../docs/source/_static/fitting-methods/progressions-4.png}
-    \caption{1}
-    \label{fig:mask-difference-metric}
+    \begin{tikzonimage}[width=0.32\textwidth]
+        {../docs/source/_static/fitting-methods/algorithm/image001042.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{a}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.32\textwidth]
+        {../docs/source/_static/fitting-methods/algorithm/mask-zoom.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{b}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.32\textwidth]
+        {../docs/source/_static/fitting-methods/algorithm/mask-zoom.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{c}};
+    \end{tikzonimage}
+    \caption{}
+    \label{fig:position-extraction-algorithm}
 \end{figure}
@@
     \centering
-    \includegraphics[width=0.32\textwidth]
+    \begin{tikzonimage}[width=0.32\textwidth]
         {../docs/source/_static/fitting-methods/extract_positions-000400.png}
-    \includegraphics[width=0.32\textwidth]
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{a}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.32\textwidth]
         {../docs/source/_static/fitting-methods/extract_positions-000600.png}
-    \includegraphics[width=0.32\textwidth]
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{b}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.32\textwidth]
         {../docs/source/_static/fitting-methods/extract_positions-001200.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{c}};
+    \end{tikzonimage}
     \caption{Progression of generated masks with estimated and exact positions.}

### 2025-01-29 paper/main.tex
insert as individual figure later
add section for parameter estimation and profiles of parameters

@@
     \includegraphics[width=0.5\textwidth]
-        {../docs/source/_static/fitting-methods/penalty-time-flow.png}%
-    \includegraphics[width=0.5\textwidth]
         {../docs/source/_static/fitting-methods/displacement-calculations.png}
@@
 
+%###################################################################################################
+\section{Estimating Parameters}
+\label{section:parameter-estimation}
+
 \begin{figure}
     \centering
-    \includegraphics[width=0.32\textwidth]
-        {../docs/source/_static/fitting-methods/estimate-parameters1/Growth Rate.png}%
-    \includegraphics[width=0.32\textwidth]
-        {../docs/source/_static/fitting-methods/estimate-parameters1/Potential Stiffness.png}%
-    \includegraphics[width=0.32\textwidth]
-        {../docs/source/_static/fitting-methods/estimate-parameters1/Radius.png}\\%
-    \includegraphics[width=0.32\textwidth]
+    \includegraphics[width=0.33\textwidth]
+        {../docs/source/_static/fitting-methods/estimate-parameters1/microscopic-images-0.png}%
+    \includegraphics[width=0.33\textwidth]
+        {../docs/source/_static/fitting-methods/estimate-parameters1/microscopic-images-1.png}%
+    \includegraphics[width=0.33\textwidth]
+        {../docs/source/_static/fitting-methods/estimate-parameters1/Growth Rate.png}\\%
+    \includegraphics[width=0.33\textwidth]
+        {../docs/source/_static/fitting-methods/estimate-parameters1/microscopic-images-2.png}%
+    \includegraphics[width=0.33\textwidth]
+        {../docs/source/_static/fitting-methods/estimate-parameters1/microscopic-images-3.png}%
+    \includegraphics[width=0.33\textwidth]
+        {../docs/source/_static/fitting-methods/estimate-parameters1/Potential Stiffness.png}\\
+    \includegraphics[width=0.33\textwidth]
+        {../docs/source/_static/fitting-methods/estimate-parameters1/Radius.png}%
+    \includegraphics[width=0.33\textwidth]
         {../docs/source/_static/fitting-methods/estimate-parameters1/Rigidity.png}%
-    \includegraphics[width=0.32\textwidth]
+    \includegraphics[width=0.33\textwidth]
         {../docs/source/_static/fitting-methods/estimate-parameters1/Strength.png}%
-    \caption{Snapshots at \dots}
+    \caption{Profiles of the Parameters}
     \label{fig:parameter-estimates-single-step}
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]
+        {../docs/source/_static/fitting-methods/penalty-time-flow.png}%
+    \caption{Comparison of masks across generations.}
+    \label{fig:penalty-calculation-cell-division}
+\end{figure}
+
 \bibliographystyle{IEEEtran}

### 2025-01-29 paper/main.tex
add subsection for comparing across cell-divisions

@@
 
-%###################################################################################################
-\section{Estimating Parameters}
-\label{section:parameter-estimation}
+% --------------------------------------------------------------------------------------------------
+\subsection{Comparing Results Across Generations}
+
+\begin{figure}
+    \centering
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {../docs/source/_static/fitting-methods/progressions-1.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{a}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {../docs/source/_static/fitting-methods/progressions-2.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{b}};
+    \end{tikzonimage}
+    \linebreak
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {../docs/source/_static/fitting-methods/progressions-3.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{c}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {../docs/source/_static/fitting-methods/progressions-4.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{d}};
+    \end{tikzonimage}
+    \caption{TODO}
+    \label{fig:mask-difference-metric}
+\end{figure}
 

### 2025-05-08 src/agent.rs
implement random distribution of growth rate

@@ -20,6 +20,9 @@ pub struct RodAgent {
     /// Rate with which the cell grows in units `1/MIN`.
     #[pyo3(set, get)]
     pub growth_rate: f32,
+    /// Determines the mean and width of distribution for sampling new values of growth rate.
+    #[pyo3(set, get)]
+    pub growth_rate_distr: (f32, f32),
     /// Threshold at which the cell will divide in units `MICROMETRE`.
     #[pyo3(set, get)]
     pub spring_length_threshold: f32,
@@ -208,6 +211,7 @@ impl RodAgent {
         spring_length=3.0,
         damping=1.0,
         growth_rate=0.1,
+        growth_rate_distr=(0.1, 0.0),
         spring_length_threshold=6.0,
         neighbor_reduction=None,
     ))]
@@ -223,6 +227,7 @@ impl RodAgent {
         spring_length: f32,
         damping: f32,
         growth_rate: f32,
+        growth_rate_distr: (f32, f32),
         spring_length_threshold: f32,
         neighbor_reduction: Option<(usize, f32)>,
     ) -> pyo3::PyResult<Self> {
@@ -244,6 +249,7 @@ impl RodAgent {
             },
             interaction: RodInteraction(interaction),
             growth_rate,
+            growth_rate_distr,
             spring_length_threshold,
             neighbor_reduction,
         })
@@ -322,9 +328,17 @@ impl Cycle<RodAgent, f32> for RodAgent {
         }
     }
 
-    fn divide(_rng: &mut rand_chacha::ChaCha8Rng, cell: &mut Self) -> Result<Self, DivisionError> {
+    fn divide(rng: &mut rand_chacha::ChaCha8Rng, cell: &mut Self) -> Result<Self, DivisionError> {
+        use rand_distr::Distribution;
         let c2_mechanics = cell.mechanics.divide(cell.radius())?;
         let mut c2 = cell.clone();
+        // Pick new growth parameters
+        let distr = rand_distr::Normal::new(cell.growth_rate_distr.0, cell.growth_rate_distr.1)
+            .map_err(|e| DivisionError(format!("{e}")))?;
+        let g1 = distr.sample(rng);
+        let g2 = distr.sample(rng);
+        cell.growth_rate = g1;
+        c2.growth_rate = g2;
         c2.mechanics = c2_mechanics;
         Ok(c2)
     }

### 2025-05-08 src/bin/crm_bench.rs

@@ -63,6 +63,7 @@ fn main() {
                 mechanics: mechanics.clone(),
                 interaction: interaction.clone(),
                 growth_rate: 0.1,
+                growth_rate_distr: (0.1, 0.),
                 spring_length_threshold: 6.0,
                 neighbor_reduction: None,
             };

### 2025-05-08 src/crm_fit/mod.rs

@@ -732,6 +732,7 @@ impl Settings {
                     },
                     interaction: interaction[n].clone(),
                     growth_rate: growth_rate[n],
+                    growth_rate_distr: (growth_rate[n], 0.),
                     spring_length_threshold: f32::INFINITY,
                     neighbor_reduction: None,
                 }

### 2025-05-08 src/crm_multilayer.rs

@@ -126,6 +126,10 @@ impl MultilayerConfig {
                             ),
                         )?,
                         growth_rate: 0.1 * MICRO_METRE / MINUTE,
+                        growth_rate_distr: (
+                            0.1 * MICRO_METRE / MINUTE,
+                            0.02 * MICRO_METRE / MINUTE,
+                        ),
                         spring_length_threshold: 6. * MICRO_METRE,
                         neighbor_reduction: Some((16, 1.)),
                     },

### 2025-05-08 src/simulation.rs

@@ -131,6 +131,9 @@ pub struct AgentSettings {
     pub interaction: Py<PhysicalInteraction>,
     /// Rate with which the length of the bacterium grows
     pub growth_rate: f32,
+    /// See :class:`RodAgent`
+    #[approx(epsilon_map = |x| (x, x))]
+    pub growth_rate_distr: (f32, f32),
     /// Threshold when the bacterium divides
     pub spring_length_threshold: f32,
     /// Reduces the growth rate with multiplier $((max - N)/max)^q $
@@ -145,6 +148,7 @@ impl PartialEq for AgentSettings {
             mechanics,
             interaction,
             growth_rate,
+            growth_rate_distr,
             spring_length_threshold,
             neighbor_reduction,
         } = &self;
@@ -156,6 +160,7 @@ impl PartialEq for AgentSettings {
                     .deref()
                     .eq(&other.interaction.borrow(py))
                 && growth_rate.eq(&other.growth_rate)
+                && growth_rate_distr.eq(&other.growth_rate_distr)
                 && spring_length_threshold.eq(&other.spring_length_threshold)
                 && neighbor_reduction.eq(&other.neighbor_reduction)
         })
@@ -191,6 +196,7 @@ impl AgentSettings {
                     ),
                 )?,
                 growth_rate: 0.1,
+                growth_rate_distr: (0.1, 0.),
                 spring_length_threshold: 6.0,
                 neighbor_reduction: None,
             },
@@ -243,6 +249,7 @@ impl AgentSettings {
             mechanics,
             interaction,
             growth_rate,
+            growth_rate_distr,
             spring_length_threshold,
             neighbor_reduction,
         } = self;
@@ -263,6 +270,15 @@ impl AgentSettings {
                     .into_pyobject_or_pyerr(py)?
                     .into_any()
                     .unbind(),
+            ),
+            (
+                "growth_rate_distr",
+                pyo3::types::PyTuple::new(
+                    py,
+                    [growth_rate_distr.0 as f64, growth_rate_distr.1 as f64],
+                )?
+                .into_any()
+                .unbind(),
             ),
             (
                 "growth_rate",

### 2025-05-08 src/agent.rs
enable getting and setting of neighbor reduction

@@ -27,6 +27,7 @@ pub struct RodAgent {
     #[pyo3(set, get)]
     pub spring_length_threshold: f32,
     /// Reduces the growth rate with multiplier $((max - N)/max)^q $
+    #[pyo3(set, get)]
     pub neighbor_reduction: Option<(usize, f32)>,
 }
 

### 2025-05-08 src/crm_fit/mod.rs
add storage_suffix field in order to distinguish same simulations started at identical times

@@ -439,6 +439,7 @@ impl Settings {
             surface_friction_distance: 1.,
             storage_options: vec![StorageOption::Memory],
             storage_location: std::path::PathBuf::new(),
+            storage_suffix: None,
         })
     }
 

### 2025-05-08 src/datatypes.rs

@@ -1,5 +1,5 @@
 use cellular_raza::prelude::{
-    CellBox, CellIdentifier, SimulationError, StorageInterfaceLoad, StorageOption,
+    CellBox, CellIdentifier, SimulationError, StorageBuilder, StorageInterfaceLoad,
 };
 use pyo3::prelude::*;
 use serde::{Deserialize, Serialize};
@@ -309,18 +309,22 @@ impl CellContainer {
     #[staticmethod]
     pub fn load_from_storage(
         config: Configuration,
-        date: std::path::PathBuf,
+        path: std::path::PathBuf,
     ) -> Result<Self, SimulationError> {
-        let mut new_config = config.clone();
-        new_config.storage_options = config
+        let mut config = config.clone();
+        config.storage_options = config
             .storage_options
             .clone()
             .into_iter()
             .filter(|x| x != &cellular_raza::prelude::StorageOption::Memory)
             .collect();
-        let builder = crate::simulation::new_storage_builder(&new_config)
-            .suffix(std::path::PathBuf::from("cells"))
-            .init_with_date(&date);
+
+        let builder = StorageBuilder::new()
+            .priority(config.storage_options.clone())
+            .location(&path)
+            .add_date(false)
+            .suffix("cells")
+            .init();
         let cells_storage = cellular_raza::prelude::StorageManager::<
             CellIdentifier,
             (CellBox<RodAgent>, serde::de::IgnoredAny),
@@ -338,15 +342,6 @@ impl CellContainer {
                 )
             })
             .collect();
-        let path = if config.storage_options.contains(&StorageOption::SerdeJson) {
-            cells_storage
-                .extract_builder()
-                .get_full_path()
-                .parent()
-                .map(|x| x.to_path_buf())
-        } else {
-            None
-        };
-        Ok(CellContainer::new(cells, path))
+        Ok(CellContainer::new(cells, Some(path)))
     }
 }

### 2025-05-08 src/simulation.rs

@@ -350,6 +350,9 @@ pub struct Configuration {
     /// utilized.
     #[approx(skip)]
     pub storage_location: std::path::PathBuf,
+    /// Appends a suffix to the date which is generated automatically.
+    #[approx(skip)]
+    pub storage_suffix: Option<std::path::PathBuf>,
 }
 
 impl Default for Configuration {
@@ -370,6 +373,7 @@ impl Default for Configuration {
             surface_friction_distance: 1.,
             storage_options: vec![StorageOption::Memory],
             storage_location: std::path::PathBuf::from("out"),
+            storage_suffix: None,
         }
     }
 }
@@ -673,9 +677,14 @@ fn backwards_compat_generate_positions_old() -> PyResult<()> {
 }
 
 pub(crate) fn new_storage_builder(config: &Configuration) -> StorageBuilder {
-    StorageBuilder::new()
+    let builder = StorageBuilder::new()
         .priority(config.storage_options.clone())
-        .location(&config.storage_location)
+        .location(&config.storage_location);
+    if let Some(suffix) = &config.storage_suffix {
+        builder.suffix(suffix)
+    } else {
+        builder
+    }
 }
 
 /// Executes a simulation given a :class:`Configuration` and a list of :class:`RodAgent`.

### 2025-05-08 paper/main.tex
include table with citations for simulation aspects

@@
 \usepackage{tikz}
+\usepackage{tabularx}
 
@@
 rods and an attractive component can model adhesion between cells.
+\begin{table}
+    \centering
+    \def\arraystretch{1.3}
+    \begin{tabularx}{\textwidth}{c l X}
+        &\textbf{Aspect} & \textbf{Description}\\
+        \toprule
+        &\textbf{(C) Cellular}\\
+        \midrule
+        (1) & Rod-Shaped Mechanics &
+            Rod-shaped bacteria are flexible rods which are able to freely move around (off-lattice
+            approach) \cite{Takeuchi2005,Ursell2014,Amir2014_2}.\\
+        (2) & Growth &
+            Cells grow exponentially by inserting new material along the circular part of the rod
+            \cite{Robert2014,Takeuchi2005}.\\
+        (3) & Differentiation &
+            \textit{B.subtilis} is known to differentiate into matrix-producing and
+            surfactin-producing cells \cite{vanGestel2015,Lpez2010}.\\
+        (4) & Division &
+            The formation and continuation of van Gogh bundles is driven by cell division
+            \cite{vanGestel2015}.
+            Bacteria can form multilayers during their growth phase \cite{Duvernoy2018}.\\
+        (5) & Variable Parameters &
+            Parameters for individual cells are not fixed values but rather taken from a
+            distribution \cite{Koutsoumanis2013}.\\
+        &\textbf{(CC) Cell-Cell Interactions}\\
+        \midrule
+        (6) & Adhesion &
+            Bacteria adhere to each other at longer distances and attach when in close contact
+            \cite{Verwey1947,Trejo2013}.
+            The interaction between rods can be polarized \cite{Duvernoy2018}.\\
+        (7) & Friction &
+            Friction between cells \cite{Grant2014} can be asymmetrical \cite{Doumic2020}.\\
+        &\textbf{(DC) Domain-Cell Interactions}\\
+        \midrule
+        (8) & External Forces &
+            Bacteria tend to stick to surfaces \cite{vanLoosdrecht1989}.
+            The extracellular gel exerts a force onto the cells \cite{Grant2014}.\\
+        (9) & Extracellular Reactions &
+            Bacteria can take up nutrients or possible secrete/take up signalling molecules
+            \cite{Li2025}.\\
+        \bottomrule
+    \end{tabularx}
+    \label{table:simulation-aspects}
+    \caption{TODO Insert caption}
+\end{table}
 

### 2025-05-08 src/config.rs
create new submodule for configs

@@ -1,0 +1,516 @@
+use approx::AbsDiffEq;
+use std::{hash::Hasher, num::NonZeroUsize};
+
+use numpy::PyUntypedArrayMethods;
+use pyo3::IntoPyObjectExt;
+use pyo3::{prelude::*, types::PyString};
+use serde::{Deserialize, Serialize};
+
+use crate::agent::{PhysInt, PhysicalInteraction};
+
+fn serialize_matrixxx3<S>(m: &nalgebra::MatrixXx3<f32>, s: S) -> Result<S::Ok, S::Error>
+where
+    S: serde::Serializer,
+{
+    let n = nalgebra::DMatrix::<f32>::from_iterator(m.nrows(), m.ncols(), m.iter().copied());
+    serde::Serialize::serialize(&n, s)
+}
+
+fn deserialize_matrixxx3<'de, D>(de: D) -> Result<nalgebra::MatrixXx3<f32>, D::Error>
+where
+    D: serde::Deserializer<'de>,
+{
+    let n: nalgebra::DMatrix<f32> = serde::Deserialize::deserialize(de)?;
+    let m = nalgebra::MatrixXx3::from_iterator(n.nrows(), n.iter().copied());
+    Ok(m)
+}
+
+/// Contains all settings required to construct :class:`RodMechanics`
+#[pyclass]
+#[derive(Clone, Debug, Deserialize, Serialize, PartialEq, AbsDiffEq)]
+#[approx(epsilon_type = f32)]
+pub struct RodMechanicsSettings {
+    /// The current position
+    #[approx(into_iter)]
+    #[serde(serialize_with = "serialize_matrixxx3")]
+    #[serde(deserialize_with = "deserialize_matrixxx3")]
+    pub pos: nalgebra::MatrixXx3<f32>,
+    /// The current velocity
+    #[approx(into_iter)]
+    #[serde(serialize_with = "serialize_matrixxx3")]
+    #[serde(deserialize_with = "deserialize_matrixxx3")]
+    pub vel: nalgebra::MatrixXx3<f32>,
+    /// Controls magnitude of32 stochastic motion
+    #[pyo3(get, set)]
+    pub diffusion_constant: f32,
+    /// Spring tension between individual vertices
+    #[pyo3(get, set)]
+    pub spring_tension: f32,
+    /// Stif32fness at each joint connecting two edges
+    #[pyo3(get, set)]
+    pub rigidity: f32,
+    /// Target spring length
[…]

### 2025-05-08 src/lib.rs

@@ -5,6 +5,7 @@
 
 mod agent;
 mod cell_container;
+mod config;
 mod crm_fit;
 mod crm_multilayer;
 mod fitting;
@@ -16,6 +17,7 @@ pub use agent::*;
 pub use cell_container::*;
 pub use cellular_raza::prelude::{CellIdentifier, VoxelPlainIndex};
 use cellular_raza::prelude::{MiePotentialF32, MorsePotentialF32, StorageOption};
+pub use config::*;
 pub use crm_fit::*;
 pub use fitting::*;
 pub use imaging::*;

### 2025-05-08 src/simulation.rs

@@ -1,524 +1,13 @@
-use approx::AbsDiffEq;
-use std::{hash::Hasher, num::NonZeroUsize};
-
 use backend::chili::SimulationError;
 use cellular_raza::prelude::*;
-use numpy::{PyUntypedArrayMethods, ToPyArray};
-use pyo3::IntoPyObjectExt;
-use pyo3::{prelude::*, types::PyString};
+use numpy::ToPyArray;
+use pyo3::prelude::*;
 use serde::{Deserialize, Serialize};
-use time::FixedStepsize;
 
 use crate::cell_container::CellContainer;
 
 use crate::agent::*;
-
-fn serialize_matrixxx3<S>(m: &nalgebra::MatrixXx3<f32>, s: S) -> Result<S::Ok, S::Error>
-where
-    S: serde::Serializer,
-{
-    let n = nalgebra::DMatrix::<f32>::from_iterator(m.nrows(), m.ncols(), m.iter().copied());
-    serde::Serialize::serialize(&n, s)
-}
-
-fn deserialize_matrixxx3<'de, D>(de: D) -> Result<nalgebra::MatrixXx3<f32>, D::Error>
-where
-    D: serde::Deserializer<'de>,
-{
-    let n: nalgebra::DMatrix<f32> = serde::Deserialize::deserialize(de)?;
-    let m = nalgebra::MatrixXx3::from_iterator(n.nrows(), n.iter().copied());
-    Ok(m)
-}
-
-/// Contains all settings required to construct :class:`RodMechanics`
-#[pyclass]
-#[derive(Clone, Debug, Deserialize, Serialize, PartialEq, AbsDiffEq)]
-#[approx(epsilon_type = f32)]
-pub struct RodMechanicsSettings {
-    /// The current position
-    #[approx(into_iter)]
-    #[serde(serialize_with = "serialize_matrixxx3")]
-    #[serde(deserialize_with = "deserialize_matrixxx3")]
-    pub pos: nalgebra::MatrixXx3<f32>,
-    /// The current velocity
-    #[approx(into_iter)]
-    #[serde(serialize_with = "serialize_matrixxx3")]
-    #[serde(deserialize_with = "deserialize_matrixxx3")]
-    pub vel: nalgebra::MatrixXx3<f32>,
-    /// Controls magnitude of32 stochastic motion
-    #[pyo3(get, set)]
-    pub diffusion_constant: f32,
-    /// Spring tension between individual vertices
[…]

### 2025-05-08 src/config.rs
add conversions

@@ -6,7 +6,9 @@ use pyo3::IntoPyObjectExt;
 use pyo3::{prelude::*, types::PyString};
 use serde::{Deserialize, Serialize};
 
-use crate::agent::{PhysInt, PhysicalInteraction};
+use crate::agent::{PhysInt, PhysicalInteraction, RodAgent};
+
+use cellular_raza::prelude::{RodInteraction, RodMechanics, StorageOption};
 
 fn serialize_matrixxx3<S>(m: &nalgebra::MatrixXx3<f32>, s: S) -> Result<S::Ok, S::Error>
 where
@@ -55,6 +57,29 @@ pub struct RodMechanicsSettings {
     /// Damping constant
     #[pyo3(get, set)]
     pub damping: f32,
+}
+
+impl From<RodMechanicsSettings> for RodMechanics<f32, 3> {
+    fn from(value: RodMechanicsSettings) -> Self {
+        let RodMechanicsSettings {
+            pos,
+            vel,
+            diffusion_constant,
+            spring_tension,
+            rigidity,
+            spring_length,
+            damping,
+        } = value;
+        RodMechanics {
+            pos,
+            vel,
+            diffusion_constant,
+            spring_tension,
+            rigidity,
+            spring_length,
+            damping,
+        }
+    }
 }
 
 #[pymethods]
@@ -158,6 +183,31 @@ impl PartialEq for AgentSettings {
                 && growth_rate_distr.eq(&other.growth_rate_distr)
                 && spring_length_threshold.eq(&other.spring_length_threshold)
                 && neighbor_reduction.eq(&other.neighbor_reduction)
+        })
+    }
+}
+
+impl From<AgentSettings> for RodAgent {
+    fn from(value: AgentSettings) -> Self {
+        Python::with_gil(|py| {
+            let AgentSettings {
+                mechanics,
+                interaction,
+                growth_rate,
+                growth_rate_distr,
+                spring_length_threshold,
+                neighbor_reduction,
+            } = value;
+            let mechanics = mechanics.borrow(py).clone().into();
+            let interaction = RodInteraction(interaction.borrow(py).clone());
+            RodAgent {
+                mechanics,
+                interaction,
+                growth_rate,
+                growth_rate_distr,
+                spring_length_threshold,
+                neighbor_reduction,
+            }
         })
     }
 }

### 2025-05-08 src/lib.rs
add new function to generate agents directly instead of just positions

@@ -45,6 +45,7 @@ fn cr_mech_coli(py: Python, m: &Bound<'_, PyModule>) -> PyResult<()> {
     )?;
 
     m.add_function(wrap_pyfunction!(generate_positions, m)?)?;
+    m.add_function(wrap_pyfunction!(generate_agents, m)?)?;
     m.add_function(wrap_pyfunction!(run_simulation_with_agents, m)?)?;
     m.add_function(wrap_pyfunction!(sort_cellular_identifiers, m)?)?;
     m.add_class::<CellIdentifier>()?;

### 2025-05-08 src/simulation.rs

@@ -50,6 +50,55 @@ pub fn generate_positions<'py>(
     .collect())
 }
 
+/// Uses the :func:`generate_positions_old` function to generate positions and then sets parameters
+/// of the generated agents from the supplied :class:`AgentSettings`.
+#[allow(clippy::too_many_arguments)]
+#[pyfunction]
+#[pyo3(signature = (
+    n_agents,
+    agent_settings,
+    config,
+    rng_seed = 0,
+    dx = [0.0, 0.0],
+    randomize_positions = 0.0,
+    n_vertices = 8,
+))]
+pub fn generate_agents(
+    py: Python,
+    n_agents: usize,
+    agent_settings: &AgentSettings,
+    config: &Configuration,
+    rng_seed: u64,
+    dx: [f32; 2],
+    randomize_positions: f32,
+    n_vertices: usize,
+) -> Vec<crate::RodAgent> {
+    use core::ops::Deref;
+    let positions = _generate_positions_old(
+        n_agents,
+        agent_settings.mechanics.borrow(py).deref(),
+        config,
+        rng_seed,
+        dx,
+        randomize_positions,
+        n_vertices,
+    );
+    positions
+        .into_iter()
+        .map(|pos| crate::RodAgent {
+            mechanics: RodMechanics {
+                pos,
+                ..agent_settings.mechanics.borrow(py).clone().into()
+            },
+            interaction: RodInteraction(agent_settings.interaction.borrow(py).clone()),
+            growth_rate: agent_settings.growth_rate,
+            growth_rate_distr: agent_settings.growth_rate_distr,
+            spring_length_threshold: agent_settings.spring_length_threshold,
+            neighbor_reduction: agent_settings.neighbor_reduction,
+        })
+        .collect()
+}
+
 /// Backend functionality to use within rust-specific code for [generate_positions_old]
 pub fn _generate_positions_old(
     n_agents: usize,

### 2025-05-08 src/lib.rs
cleanup

@@ -10,7 +10,6 @@ mod crm_fit;
 mod crm_multilayer;
 mod fitting;
 mod imaging;
-mod sampling;
 mod simulation;
 
 pub use agent::*;
@@ -19,10 +18,9 @@ pub use cellular_raza::prelude::{CellIdentifier, VoxelPlainIndex};
 use cellular_raza::prelude::{MiePotentialF32, MorsePotentialF32, StorageOption};
 pub use config::*;
 pub use crm_fit::*;
+pub use crm_multilayer::*;
 pub use fitting::*;
 pub use imaging::*;
-// pub use sampling::*;
-pub use crm_multilayer::*;
 pub use simulation::*;
 
 use pyo3::prelude::*;

### 2025-05-08 src/crm_fit/mod.rs
add documentation for Parameter of crm_fit

@@ -38,7 +38,11 @@ impl SampledFloat {
     }
 }
 
-/// TODO
+/// This enum has 3 variants:
+///
+/// - :class:`SampledFloat` Samples the value in the given range
+/// - :class:`float` Fixes it to the given value
+/// - :class:`list` Fixes it on a per-agent basis to the given values.
 #[pyclass(get_all, set_all, module = "cr_mech_coli.crm_fit")]
 #[derive(Clone, Debug, Serialize, Deserialize, AbsDiffEq, PartialEq)]
 #[approx(epsilon_type = f32)]

### 2025-05-08 paper/main.tex
formatting
use new images

@@
 
-\title{
-    A mechanical Model of Rod-Shaped Bacteria
-}
+\title{A mechanical Model of Rod-Shaped Bacteria}
 
@@
 
-\begin{figure}
+\begin{figure}[!h]
     \centering
@@
     \begin{tikzonimage}[width=0.32\textwidth]
-        {../docs/source/_static/fitting-methods/extract_positions-000400.png}
+        {../docs/source/_static/fitting-methods/extract_positions-004000.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{a}};
@@
     \begin{tikzonimage}[width=0.32\textwidth]
-        {../docs/source/_static/fitting-methods/extract_positions-000600.png}
+        {../docs/source/_static/fitting-methods/extract_positions-006000.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{b}};
@@
     \begin{tikzonimage}[width=0.32\textwidth]
-        {../docs/source/_static/fitting-methods/extract_positions-001200.png}
+        {../docs/source/_static/fitting-methods/extract_positions-008000.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{c}};

### 2025-05-08 plots/fitting_extract_positions.py
use tuples instead of list for correct types

@@ -40,7 +40,7 @@ def calculate_lengths_distances(
     lengths_exact = []
     for p in positions:
         color = mask[int(np.round(p[0][0])), int(np.round(p[0][1]))]
-        ident = cell_container.get_cell_from_color([*color])
+        ident = cell_container.get_cell_from_color((color[0], color[1], color[2]))
         cell = cells_at_iteration[ident][0]
         q = cell.pos[:, :2]
         p = crm.convert_pixel_to_position(p, config.domain_size, mask.shape[:2])
@@ -134,7 +134,6 @@ if __name__ == "__main__":
 
         for iteration, mask in tqdm(iter_masks):
             # This generates extracted positions in pixel units
-            cv.imwrite("temp.png", mask)
             positions = crm.extract_positions(
                 mask,
                 n_vertices=pyargs.n_vertices,
@@ -148,7 +147,9 @@ if __name__ == "__main__":
             for n, p0 in enumerate(positions):
                 # Get color
                 color = mask[p0[0][0], p0[0][1]]
-                ident = cell_container.get_cell_from_color([*color])
+                ident = cell_container.get_cell_from_color(
+                    (color[0], color[1], color[2])
+                )
                 cell = all_cells[iteration][ident][0]
                 p1 = crm.convert_cell_pos_to_pixels(
                     cell.pos[:, :2], config.domain_size, mask.shape[:2]

### 2025-05-08 plots/fitting_extract_positions.py
fixup! rename functions; remove _old suffix

@@ -79,7 +79,7 @@ def create_simulation_result(n_vertices: int, rng_seed: int = 3):
     agent_settings.mechanics.rigidity = 8.0
     config.domain_height = 0.2
 
-    positions = crm.generate_positions_old(
+    positions = crm.generate_positions(
         n_agents,
         agent_settings,
         config,

### 2025-05-08 plots/fitting_extract_positions.py
also set distribution parameters of growth_rate

@@ -75,7 +75,7 @@ def create_simulation_result(n_vertices: int, rng_seed: int = 3):
         n_saves=49,
         domain_size=np.array([150, 150]),
     )
-    agent_settings = crm.AgentSettings(growth_rate=0.05)
+    agent_settings = crm.AgentSettings(growth_rate=0.05, growth_rate_distr=(0.05, 0.01))
     agent_settings.mechanics.rigidity = 8.0
     config.domain_height = 0.2
 

### 2025-05-08 plots/fitting_extract_positions.py
ensure that we only store in Memory

@@ -75,6 +75,7 @@ def create_simulation_result(n_vertices: int, rng_seed: int = 3):
         n_saves=49,
         domain_size=np.array([150, 150]),
     )
+    config.storage_options = [crm.StorageOption.Memory]
     agent_settings = crm.AgentSettings(growth_rate=0.05, growth_rate_distr=(0.05, 0.01))
     agent_settings.mechanics.rigidity = 8.0
     config.domain_height = 0.2

### 2025-05-08 cr_mech_coli/crm_multilayer/main.py
cleanup

@@ -363,10 +363,6 @@ def plot_colony_height_versus_gel_pressure():
         np.column_stack([np.mean(t, axis=0), np.mean(d, axis=0) / radius])
         for t, d in zip(data_times, data_y95th)
     ]
-    # y_collection = [
-    #     np.column_stack([np.linspace(0, 1.2), np.linspace(0, 1.2) ** i])
-    #     for i in range(5)
-    # ]
 
     line_collection = mpl.collections.LineCollection(
         y_collection, array=gel_pressures, cmap=cmap

### 2025-06-19 paper/main.tex
HUGE update for paper; this could be an initial draft from where we start iterating

@@
 \usepackage{amsfonts}       % blackboard math symbols
+\usepackage{float}
 \usepackage{nicefrac}       % compact symbols for 1/2, etc.
@@
 \usepackage{tabularx}
+\usepackage{siunitx}
+\usepackage{cite}
+\usepackage{acronym}
 
@@
 }
+
+\newacro{abm}[ABM]{Agent-Based Model}
+\newacro{ode}[ODE]{Ordinary Differential Equation}
+\newacro{pbr}[PBR]{Physically-Based Rendering}
 
@@
 \begin{abstract}
-    This paper develops the formal notions of cellular \dots
+    This paper establishes the basis for a novel individual-based mechanical model of rod-shaped
+    bacteria.
+    It describes methods of how to obtain data from microscopic images such that they can be used
+    for parameter estimation purposes.
+    Furthermore, we show initial results for parameter estimates and an application of the model to
+    the formation of multilayers within bacterial colonies.
 \end{abstract}
@@
 
+Bacteria exist in many physical shapes such as spheroidal, rod-shaped or spiral
+\cite{Zapun2008,Young2006}.
+The last decades have fundamentally challenged and changed our understanding of how bacteria retain
+their shape.
+Rod-shaped bacteria can grow by extending their circular part, or by inserting new material at the
+tip of the rod.
+Species such as E.coli and B.subtilis~\cite{Errington2020} fall under the first category while
+S.pombe represents the latter.
+The material is inserted in small bursts on the nanoscale in forms of patches, bands or
+hoops~\cite{DePedro2003}.
+Despite their differences in wall-size, B.subtilis and E.coli follow growth rules which are
+comparable with respect to the extension of the rod~\cite{Chang2014}.
+
+\begin{itemize}
+    \item This publication contains 5 things:
+    \begin{enumerate}
+        \item Initial (incomplete) model of rod-shaped bacteria
+        \item Methods of how to obtain data from microscopic images and perform parameter
+            estimations; $\rightarrow$ also possibly across multiple generations
+        \item Showcase the usefulness of the methods and the model by applying it to parameter
+            estimation via microscopic images and multilayer formation
+        \item Generate 3D Images (masks)
+        \item A python package which can do all of that and can be reused and extended
+    \end{enumerate}
+    \item Make clear that this is mainly a proof of the methods
+    \item For complete model, more work has to be done
+    \begin{itemize}
+        \item More aspects (see Review; put in discussion)
+        \item Note that we only care about \textit{E.Coli} and \textit{B.Subtilits}! This is
+            important.
+    \end{itemize}
+\end{itemize}
+
+% TODO write more text here
+% Use some parts of the proposal text
+
 %###################################################################################################
@@
 
-Bacteria exist in many physical shapes such as spheroidal, rod-shaped or spiral
-\cite{Zapun2008,Young2006}.
+In order to construct a mechanical model of rod-shaped bacteria, we have to take care of a number of
+different aspects.
+The mechanical behaviour of rod-shaped bacteria is governed by a plethora of effects.
+Although we set out to construct a model which is able to describe as many of these effects as
+possible, we will fall short in this paper and restrict ourselves to the aspects presented
+in~\ref{table:simulation-aspects}.
+We will discuss, how this model can be extended in order to capture more effects and how we can
+validate our approach given the existing literaturesection:supplement-more-simulation-aspects
+(see section~\ref{section:discussion}, and~\ref{section:supplement-more-simulation-aspects}).
+
+\paragraph{(C) Cellular}
+The key component of this model consists of the cellular representation.
+
+It was long believed that bacteria lack cytoskeletal filaments but with the end of the last century,
+it was discovered that the MreB protein fills this role since it “[..] forms an actin-like
+cytoskeleton in bacteria [..]”~\cite{Erickson2001}.
+It can polymerize and form filaments which behave similar to actin microfilaments~\cite{Dersch2020}.
+The effect of MreB on the shape of cells during growth was shown in more detail
+by~\cite{Ursell2014}.
+They found that MreB localizes specifically at regions of negative curvature and thus introduces
+heterogeneous cell growth which results in a bending force during the growth of cells.
+It also means that the cells are able to sense the curvature of the cell wall through cytoskeleteal
+MreB filaments during growth.
+\cite{Bratton2018} showed that the trans-membrane protein RodZ can modulate MreB density and
+facilitates local MreB assembly.
+This feedback mechanism will result in straight rod-shaped bacteria if not interfered from the
+external environment.
+These results were also confirmed by the studies of~\cite{Wang2010}.
+In their experiments, a single bacterium was fixed at the bottom pole while an optically trapped
+bead was placed on the upper pole.
+They could now exert a force on the free part and thus study the response of the cell under
+different conditions.
+When treating cells with A22, an antibiotic which disassembles
+MreB~\cite{IWAI2002,Gitai2005,Karczmarek2007,Bean2009}, a higher displacement with identical force
+could be measured, thus indicating that MreB increases mechanical rigidity of the cell.
+
+While some rod-shaped bacteria grow by inserting new material at the ends, \textit{E.Coli} and
+\textit{B.Subtilis} utilize a method where new material is inserted along the cylindrical part of
+the rod.
+The details of the division event (i.e. Z-Ring formation) are very intricate and are heavily
+studied as a subject on their own (TODO citation).
+On a higher level view, the rod is split apart at a point close to the center of the cell, thus
+creating two new daughter cells.
+We know that the best predictor for when the division event occurs, is the length of the rod as
+opposed to the age of bacterium~\cite{Robert2014}.
+Furthermore, only the growth of the whole colony but also the individual growth of the bacteria is
+exponential~\cite{Amir2014,Takeuchi2005}.
+
+\paragraph{(CC) Cell-Cell Interactions \& (DC) Domain-Cell Interactions}
+A colony of bacteria is known to form biofilms (Dunne 2002). This strategy is is beneficial for the
+survival of the collective as it allows it to harness growth factors and migrate when necessary and
+also plays an important role in bacterial infections in the biomedical sector~\cite{Ong1999}.
+The formation of biofilms is only possible due to the attractive nature of bacterial interactions
+towards each other and surfaces (Berne et al. 2018). To analyze bacterial adhesion to surfaces,
+\cite{vanLoosdrecht1989} investigated the interactions of various bacteria with negatively charged
+polystyrene.
+They found that the interaction is characterized by a low Gibbs energy of $(2 - 3)kT$ per cell
+which can be described by a secondary minimum of the DLVO-theory~\cite{Verwey1947,Derjaguin1993}
+which is relevant for distances $\geq\SI{1}{\nano\meter}$.
+At these distances, the DLVO potential acts as a generic adhesive potential.
+Particles which stay in the secondary minimum are held together but even weak interactions are
+enough to redisperse them which means that this state is reversible.
+When approaching closer distances, more intricate mechanisms come into play in which the
+heterogeneity of the materials in question needs to be considered.
+\cite{Hori2010} described bacterial adhesion as a “two-phase process including an initial,
+instantaneous and reversible physical phase (phase one) and a time-dependent and irreversible
+molecular and cellular phase (phase two)”.
+
+All these considerations become even more intricate when not considering interactions of bacteria
+with surfaces but bacteria with each other.
+Gram-negative bacteria (eg. E.Coli) express outer membrane proteins (adhesins) which allow them to
+attach to various hosts (Beachey 1981, Vaca et al. 2019).
+In addition, the flagella can play an important role in the direct adhesion between
+cells~\cite{Haiko2013}.
+Thus it is not correct to blindly apply the results for adhesion to surfaces to the interactions
+between bacteria.
+
+\begin{table}
+    \centering
+    \def\arraystretch{1.3}
+    \begin{tabularx}{\textwidth}{c l X}
+        &\textbf{Aspect} & \textbf{Description}\\
+        \toprule
+        &\textbf{(C) Cellular}\\
+        \midrule
+        (1) & Rod-Shaped Mechanics &
+            Rod-shaped bacteria are flexible rods which are able to freely move around (off-lattice
+            approach) \cite{Takeuchi2005,Ursell2014,Amir2014_2}.\\
+        (2) & Growth &
+            Cells grow exponentially by inserting new material along the circular part of the rod
+            \cite{Robert2014,Takeuchi2005}.\\
+        % (3) & Differentiation &
+        %     \textit{B.subtilis} is known to differentiate into matrix-producing and
+        %     surfactin-producing cells \cite{vanGestel2015,Lpez2010}.\\
+        (3) & Division &
+            Rod-Shaped bacteria divide in the middle of the rod, thus forming two new
+            agents.
+            Bacteria can form multilayers during their growth phase \cite{Duvernoy2018}.\\
+        (4) & Variable Parameters &
+            Parameters for individual cells are not fixed values but rather taken from a
+            distribution \cite{Koutsoumanis2013}.\\
+        &\textbf{(CC) Cell-Cell Interactions}\\
+        \midrule
+        (5) & Adhesion &
+            Bacteria adhere to each other at longer distances and attach when in close contact
+            \cite{Verwey1947,Trejo2013}.\\
+            % The interaction between rods can be polarized \cite{Duvernoy2018}.\\
+        % (7) & Friction &
+        %     Friction between cells \cite{Grant2014} can be asymmetrical \cite{Doumic2020}.\\
+        &\textbf{(DC) Domain-Cell Interactions}\\
+        \midrule
+        (6) & External Forces &
+            Bacteria which rub against surfaces tend to stick to them after some
+            time~\cite{vanLoosdrecht1989}.
+            The extracellular gel which surrounds the bacteria exerts a force onto the cells
+            \cite{Grant2014}.\\
+        % (9) & Extracellular Reactions &
+        %     Bacteria can take up nutrients or possible secrete/take up signalling molecules
+        %     \cite{Li2025}.\\
+        \bottomrule
+    \end{tabularx}
+    \label{table:simulation-aspects}
+    \caption{TODO Insert caption}
+\end{table}
 
 To model the spatial mechanics of elongated bacteria \cite{Billaudeau2017}, we represent them as a
-collection of auxiliary vertices $\{\vec{v}_i\}$ which are connected by springs in
+collection of auxiliary vertices $\{\vec{x}_i\}$ which are connected by springs in
 ascending order.
@@
 rods and an attractive component can model adhesion between cells.
-\begin{table}
+
+% --------------------------------------------------------------------------------------------------
+\subsection{Mechanics}
+\label{subsection:mechanical-model-mechanics}
+
+We use principles from discrete differential geometry~\cite{bobenko2008discrete} to describe the
+stiffness and flexibility of the bacterial rod.
+As explained previously, MReB will counteract the bending of the rod, resulting in a stiffening
+force.
+In order to construct a model which can be parametrized and by method of Occam's Razor, we chose to
+model the bacterium as a bending beam, thus utilizing the simplest solution which can still produce
+the desired effects.
+Such a beam can be represented by a collection of vertices $\vec{x}_i$ which are connected by a
+springs of lengths $l_i$.
+We use the notation $\vec{c}_i:=\vec{x}_i-\vec{x}_{i-1}$ with which the resulting force acting
+between two vertices if given by
+
+\begin{align}
+    \vec{F}_{i,\text{springs}} =
+        &-\gamma\left(1 - \frac{l}{\left|\vec{c}_i\right|}\right)
+        \vec{c}_i\\
+        &+ \gamma\left(1 - \frac{l}{\left|\vec{c}_{i+1}\right|}\right)
+        \vec{c}_{i+1}.
+\end{align}
+
+In addition to springs between individual vertices $\vec{x}_i$, we assume that each angle at a
+vertex between edges is subject to a force indiced by curvature.
+% We assume that we can model the mechanical properties of the bacterium as an elastic rod.
+Given the angle $\alpha_i = \sphericalangle(\vec{c}_{i-1},\vec{c}_i)$ between adjacent edges,
+the bending force is proportional to the curvature $\kappa_i$ at each vertex
+$\kappa_i = 2\tan(\alpha_i/2)$.
+The resulting force acts along the angle bisector which can be calculated from the edge vectors.
+Figure~\ref{fig:mechanics-bacterium} shows an elongated rod represented by $6$ vertices.
+The forces acting on vertices $\vec{x}_i,\vec{x}_{i-1},\vec{x}_{i+1}$ are given by
+
+\begin{align}
+    \vec{F}_{i,\text{curvature}} &= \eta\kappa_i
+        \frac{\vec{c}_i - \vec{c}_{i+1}}{|\vec{c}_i-\vec{c}_{i+1}|}\\
+    \vec{F}_{i-1,\text{curvature}} &= -\frac{1}{2}\vec{F}_{i,\text{curvature}}\\
+    \vec{F}_{i+1,\text{curvature}} &= -\frac{1}{2}\vec{F}_{i,\text{curvature}}
+\end{align}
+
+where $\eta_i$ is the rigidity at vertex $\vec{x}_i$ (see Figure~\ref{fig:mechanics-bacterium}).
+We can see that the curvature force does not move the overall center of the rod in space.
+The total force is the sum of external and interal forces.
+
+\begin{equation}
+    \vec{F}_{i,\text{total}} = \vec{F}_{i,\text{springs}}+ \vec{F}_{i,\text{curvature}}
+        + \vec{F}_{i,\text{external}}
+\end{equation}
+
+and are integrated via
+
+\begin{align}
+    \partial_t^2 \vec{x}_i &= \partial_t\vec{x}_i + \sqrt{2D}\vec{\xi}_i\\
+    \partial_t\vec{x}_i &= \vec{F}_{i,\text{total}} - \lambda \partial_t\vec{x}_i
+\end{align}
+
+where $D$ is the diffusion constant and $\vec{\xi}$ is the wiener process.
+Together they determine the stochastic motion of the individual vertices which can arise from
+recurring collisions with small particles of the surrounding medium or other minor fluctuations.
+
+%TODO use these citations
+\cite{Chatterjee1988},\cite{Ursell2014}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{../docs/source/_static/mechanics.png}
+    \caption{
+        Example of an elonged bacterial rod represented by a collection of vertices which are
+        connected by springs.
+        The angles $\alpha_i$ have been exaggerated such that they can be identified more easily.
+        A bacterial agent which is not interacting with any external objects will tend to the
+        equilibrium state of a straight rod with $0$ curvature.
+    }
+    \label{fig:mechanics-bacterium}
+\end{figure}
+
+% --------------------------------------------------------------------------------------------------
+\subsection{Interactions}
+
+When calculating forces acting between the cells, we can use a simplified model to circumvent the
+numerically expensive integration over the complete length of the rod.
+Given a vertex $\vec{x}_i$ on one cell, we calculate the closest point $\vec{p}$ on the polygonal
+line given by the vertices $\{\vec{w}_j\}$ of the interacting cell.
+Furthermore we determine the value $q\in[0,1]$ such that $\vec{p} = (1-q)\vec{w}_j + q\vec{w}_{j+1}$
+for some specific $j$.
+The force is then calculated between the points $\vec{x}_i$ and $\vec{p}_i$ and acts on
+the vertex $\vec{w}_i,\vec{w}_{i+1}$ with relative strength $(1-q)$ and $q$.
+
+\begin{equation}
+    \vec{F}_{i,\text{External}} = \vec{F}(\vec{x}_i,\vec{p})
+    \label{eq:cell-force-external}
+\end{equation}
+
+The specific choice of interaction potential plays of course a crucial role for the dynamics of the
+agents.
+But we cannot expect to simply know the shape and parameters of this potential.
+Thus, we would like to experiment with various interaction potentials in order to quantify their
+ability to produce accurate results.
+We utilize two distinct potentials that are both able to capture the essence of our interactions by
+providing a short-range repulsive part and a long-ranged attractive component.
+
+\paragraph{Morse Potential}
+The morse potential~\cite{Morse1929} is a common example of numerically inexpensive interaction
+potential.
+It was originally designed to describe the interaction of diatomic molecules but has since found use
+in other areas~\cite{Breitwieser2021}.
+It contains a repulsive and attractive part and is bound from above which makes it particularly
+suitable for numerical simulations since unbound potentials can lead to large numerical deviations
+thus compromising the prediction of the model.
+The potential is of the form
+\begin{equation}
+    V(r) = V_0\left(1-e^{-\lambda(r-(R_1+R_2))}\right)^2
+\end{equation}
+and has 3 parameters.
+The overall potential strength $V_0$, the radii of the particles $R_1,R_2$ and the potential
+stiffness $\lambda$ which controls the rate of change as one travels along the radial axis.
+We employ a spatial cutoff which sets the interaction strength to $0$ when exceeding a distance
+$r\geq\xi$.
+
+\paragraph{Mie Potential}
+The Mie~\cite{Mie1903} potential is the generalization of the well-known
+Lennard-Jones~\cite{Jones1924} potential.
+It was also designed to describe the interactions of molecules but due to the flexible nature of its
+parameters, can be easily adapted to other situations.
+It is able to describe a wider range of shapes compared to the Morse potential but is also stiffer
+and not bound from above.
+In our simulation, we enforce an upper bound to guarantee numerical regularity.
+It has 4 parameters and is given by
+\begin{align}
+    U(r) &= C\epsilon\left[ \left(\frac{\sigma}{r}\right)^n -
+        \left(\frac{\sigma}{r}\right)^m\right]\\
+    C &= \frac{n}{n-m}\left(\frac{n}{m}\right)^{\frac{n}{n-m}}\\
+    V(r) &= \min(U(r), \beta)\theta(r-\zeta)
+\end{align}
+where $C$ is a constant that depends on the two exponents $n,m$ that describe the overall shape of
+the potential.
+The strength is given by $\epsilon$ and the radius is encoded in $\sigma =
+(m/n)^{1/(n-m)}(R_1+R_2)$ where $R_1,R_2$ are the radii of the two interacting particles.
+
+% --------------------------------------------------------------------------------------------------
+\subsection{Cycle}
+
+To simulate proliferation, we introduce a growth term for the spring lengths $l_i$.
+In most of our simulations, we choose to use identical lengths for all segments (i.e. $l_i=l$).
+However, in order to generalize our approach we assume that individual lengths of segments could
+vary.
+We reduce the growth rate with increasing number of neighbors $N$.
+This behaviour can be toggled on or off thus removing the additional term in brackets of
+equation~\ref{eq:growth-rod-lengths}.
+
+\begin{equation}
+    \partial_t l_i = \mu\left(1-\frac{\text{min}(n,N)}{N}\right)l_i
+    \label{eq:growth-rod-lengths}
+\end{equation}
+
+Note that the growth rate is proportional to the rod length $l$ which leads to exponential growth.
+This process will increase the length of the cell indefenitely unless we impose a condition for the
+division event.
+We define a fixed length threshold for the total length of the cell at which it divides.
+To construct a new cell, we cannot simply copy the existing one twice, but we also need to adjust
+internal parameters in the process.
+The following actions need to be taken for the old and new agent.
+We first calculate the new spring lengths of the two resulting rods.
+\begin{equation}
+    \tilde{l} = l_i\left(\frac{1}{2} - \frac{r}{\sum\limits_i l_i}\right)
+    \label{eq:growth-rod-new-lengths}
+\end{equation}
+Afterwards, we start at one endpoint $\vec{v}_0$ and travel along the polygon defined by the
+bacterial rod.
+We stop when we have reached a length that matches the length of the new spring length $\tilde{l}$.
+From here, we continue and iterate this procedure until we have found all new vertices $\vec{w}_i$.
+Equation~\ref{eq:growth-rod-new-lengths} ensures that that the tips of the two new rods are touching
+right after the division event.
+
+% --------------------------------------------------------------------------------------------------
+\subsection{External Forces}
+TODO
+\begin{itemize}
+    \item Friction force against surface. perpendicular to motion; proportional to velocity.
+    \item gel pressure coming from experiment (simply downwards force $F_z$)
+\end{itemize}
+
+% --------------------------------------------------------------------------------------------------
+\subsection{Numerical Implementation}
+
+We implemented the described model in \texttt{cellular\_raza}~\cite{Pleyer2025}.
+Doing so has many benefits.
+Not only, was a more simpler model already available which could be adapted to our needs.
+But furthermore, due to the nature of \texttt{cellular\_raza} in that it does not assume any
+particular cellular representation, we are free to change the underlying model at any time, even
+making fundamental changes.
+It is an \ac{abm} which is able to natively track cells through the course of a numerical simulation
+and also keeps track of parent/daughter relationships.
+This allows us to to experiment with different model implementations and still retain performance.
+By leveraging the strong safety guarantees provided by the Rust programming language, we eliminate a
+whole class of problems from the start.
+% TODO CITATION
+In addition, we are using the \texttt{pyo3} and \texttt{maturin} crates which allow us to easily
+generate python bindings and thus interoperate between the two ecosystems.
+While this particular model is more complex due to the nature of non-trivial (asymmetric)
+interactions, we were still able to properly utilize the speedup from executing simulations in
+parallalel using the built-in parallelization methods of \texttt{cellular\_raza} (see
+Supplementary Material, Figure~\ref{fig:performance}).
+
+%###################################################################################################
+\section{Generating Masks and Images}
+\label{section:generating-masks-and-images}
+\begin{itemize}
+    \item use pyvista (VTK) for Visualization
+    \item always do 3D render. This way, we can have overalp between agents and produce more
+        realistic masks.
+    \item Assign unique color values for idividual agents.
+        We have a 1:1 mapping between agent identifier and color! (for each simulation different
+        though)
+    \item use parallel projection; this should be most close to what microscopes measure due to the
+        relatively large distance between object and lens.
+    \item use some noise to smear out the rods; this is already similar to microscopic images; use
+        more like this and deal with real-world defects to make it more realistic (future, not now)
+    \item Figure~\ref{fig:progression-image-generation} shows steps in generating image.
+    \item Reflectivity, placement of light etc. can already introduce some heterogeneity in the
+        visual of the rod.
+    \item Plan to implement noise on the mesh itself; make it have rough surface (like real rods as
+        seen under electron microscopy)
+\end{itemize}
+
+\paragraph{Algorithm}
+\begin{enumerate}
+    \item Get position of cell
+    \item Insert spheres at each vertex
+    \item Insert cylinder along edges
+    \item Combine individual objects into one big mesh
+    \item Use \ac{pbr}
+\end{enumerate}
+
+\begin{figure}[H]
+    \centering
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {../docs/source/_static/09395645494836445480/raw_pv/000000400.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {../docs/source/_static/09395645494836445480/images/000000400.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {../docs/source/_static/09395645494836445480/masks/000000400.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+    \end{tikzonimage}
+    \caption{
+        (A) shows the raw result from combining sphere and cylinder meshes to obtain the shape of a
+        bacterial rod.
+        This render contains lighting coming from 2 light sources which thus produces a small glow
+        on the upper part of the meshes.
+        Subfigure (B) takes this image and applies noise to it in order to make it look more
+        like an actual microscopic image.
+        (C) The render is projected along the z-axis and flat numerical values are assigned to the
+        bacteria.
+        The resuling render can have overlaps in which case the color of the rod is displayed that
+        is higher along the z-direction.
+    }
+    \label{fig:progression-image-generation}
+\end{figure}
+
+%###################################################################################################
+\section{Parameter Estimation}
+
+In order to effectively estimate the parameters of our model, we need to set up a cost function
+which can be minimized.
+In doing so, we hope to identify all parameters, given relevant experimental data.
+We use microscopic images and construct a novel algorithm that can extract the position of the agent
+from a provided cell-mask.
+These cell-masks are easily obtainable by either using existing segmentation tools or manually
+annottating images.
+Following this, we can define a cost-function and an optimization scheme to minimize said function.
+
+% --------------------------------------------------------------------------------------------------
+\subsection{Extract Positions from Masks}
+\label{section:fitting-and-comparing-masks}
+
+This section will explain how we can determine the collection of vertices $\vec{v}_i$ introduced in
+subsection~\ref{subsection:mechanical-model-mechanics} from microscopic images
+(Figure~\ref{fig:position-extraction-algorithm} (A)).
+They will serve as initial, final and intermediate values with which we can compare our predictions.
+The algorithms which we devised requires the input of cell-masks which assign a unique color for
+every cell in the image.
+Such masks can be generated by cell-segmentation tools or obtained by annottating microscopic images
+by hand (Figure~\ref{fig:position-extraction-algorithm} (B)).
+Afterwards, we proceed with the following 5 steps.
+\begin{enumerate}
+    \item Obtain submasks for each cell individually.
+    \item Skeletonize submask with algorithm from Lee~\cite{Lee1994}
+    \item Determine endpoints $\vec{q}_0,\vec{q}_1$.
+        Assert that there are only 2 endpoints.
+    \item Sort points such that they form a polygon.
+    \item Interpolate vertices $\vec{v}_i$ from obtained polygon.
+\end{enumerate}
+In the first step, the image will be split into a collection of subimages that only contain
+information for one single cell.
+From there, we apply a skeletonization algorithm~\cite{Lee1994}
+(Figure~\ref{fig:position-extraction-algorithm} (C)).
+It has to be noted that there are multiple variants of skeletonization and in general, these classes
+of algorithms can produce results which have intersection points and multiple endpoints.
+This described behaviour is undesirable in our case.
+We have found that for most practical examples, the algorithm devised by~\cite{Lee1994} has worked
+well.\\
+The skeletonization algorithm supplies us with a collection of unordered lattice points $\vec{p}_i$
+(pixel coordinates) where the skeleton lives.
+We identify the endpoints $\vec{q}_0,\vec{q}_1$ by counting the number of neighbors in the von
+Neumann neighborhood for each lattice point.
+Any point with exactly one neighboring point must be and endpoint.
+If we encounter more or less than two points which fit this criterion, the routine can not commence
+and is aborted.
+Next, we start with one of the endpoints and determine the next neighbor which was not yet picked.
+We list the points in this order, thus forming a sorted polygon of pixel coordinates.\\
+In the last step, we take the sorted, extracted points $\vec{p}_i$ and interpolate $n$ values for
+the desired number of vertices $\vec{v}_i$.
+
+\begin{figure}
+    \centering
+    \begin{tikzonimage}[width=0.32\textwidth]
+        {../docs/source/_static/fitting-methods/algorithm/image001042.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.32\textwidth]
+        {../docs/source/_static/fitting-methods/algorithm/mask-zoom.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.32\textwidth]
+        {../docs/source/_static/fitting-methods/algorithm/mask-zoom.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+    \end{tikzonimage}
+    \caption{
+        TODO change 2nd image!\\
+        (A) An example of a microscopic image.
+        (B) The associated mask which assigns individual color values to each bacterium.
+        (C) The result of the  skeletonization algorithm when applying it to a single cell.
+    }
+    \label{fig:position-extraction-algorithm}
+\end{figure}
+
+% --------------------------------------------------------------------------------------------------
+\subsection{Benchmarking the Extraction Algorithm}
+\label{subsection:benchmarking-extraction-algorithm}
+
+In order to determine if our extraction algorithm provides reliable results, we have to take
+advantage of a functionality which will be introduced later in
+section~\ref{section:generating-masks-and-images}.
+Our methodology makes use of the fact that we have already set up our model and solving scheme and
+are able to produce cell masks from these numerical results.
+In doing so, we have exact knowledge about the position of our agents since we can simply obtain
+this information from our model.
+But on the other hand we can also try to extract this position from the generated cell-masks.
+By comparing the differences between them, we get a good estimate of how well our extraction
+algorithm can determine the true position of the bacteria.
+
+In figure~\ref{fig:benchmarking-extraction-algorithm} (A-C) we see snapshots of a time series of
+masks.
+This particular simulation includes division events between images (B) and (C).
+For all images (A-C), we have drawn the exact position of the agents with white lines and indicated
+the extracted positions with crosses.
+They align very well visually.
+It may be hard to spot with the bare eye, but the agreement between extracted and exact positions is
+visually worse at the endpoints of the rods.
+To quantify this disparity, we calculate the distance between these individual vertices.
+Figure~\ref{fig:benchmarking-extraction-algorithm} (D) shows the estimated length of the rods from
+the extracted and exact vertices which is surrounded by the average vertex distance.
+Division events lead to steps in our rod-length function since the bacterial length is halved.
+We can clearly see that the agreement is very good in the beginning but right after every division
+event, the uncertainty becomes large.
+To illustrate this, we can see that in figure~\ref{fig:benchmarking-extraction-algorithm} (C) some
+of the rods are partially overlapping which hides parts of the individual cell-masks and thus
+changes the extraction result.
+We also plotted the distribution of individual vertex distances in
+figure~\ref{fig:benchmarking-extraction-algorithm} (E).
+Darker areas in the plot correspond to distances at time-points earlier in the simulatin time.
+We can see that brighter areas tend to shift towards the right, thus increasing the uncertainty of
+the extraction.
+However, the overall distribution shows small values of distances for a large number of vertices and
+is not dominated by outliers.
+
+\begin{figure}
+    \centering
+    \begin{tikzonimage}[width=0.32\textwidth]
+        {../docs/source/_static/fitting-methods/extract_positions-004000.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.32\textwidth]
+        {../docs/source/_static/fitting-methods/extract_positions-006000.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.32\textwidth]
+        {../docs/source/_static/fitting-methods/extract_positions-008000.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.5\textwidth]
+        {../docs/source/_static/fitting-methods/displacement-calculations.png}%
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.5\textwidth]
+        {../docs/source/_static/fitting-methods/displacement-distribution.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+    \end{tikzonimage}
+    \caption{
+        Progression of generated masks with estimated and exact positions.
+        Comparison of masks across generations. Compare estimated with exact rod lengths.
+    }
+    \label{fig:benchmarking-extraction-algorithm}
+\end{figure}
+
+% --------------------------------------------------------------------------------------------------
+\subsection{Constructing a Cost Function}
+
+As within any optimization scheme, it can be beneficial to construct custom cost functions
+which are tailored to the problem at hand.
+With the methods that we have already estalished, we need to compare a collection of agents which
+were numerically calcualted through our \ac{abm} modeling approach whith microscopic images.
+Utilizing the extraction algorithm, we can simply compare the numerically obtained positions with
+the extracted positions of the data.
+We choose a typical least-squares approach for the comparison of the individual vertices.
+However, this procedure is only valid for the case where cell-division events do not play a role.
+In the case where we either the numerical model or the experimental data contain more or less cells
+than the other result, we are not able to compare them with this approach.
+
+\paragraph{Comparison across Generations}
+\label{paragraph:comparison-across-generations}
+
+In order to compare a variable amount of cells, we can not utilize the positions of these cells
+directly.
+In this approach, we propose a method which is similar to already existing routines (TODO mention
+machine learning, some vision algorithms use this).
+From the numerical results of the \ac{abm}, we can again generate an image of cell-masks.
+We then compare this image together to the experimental cell-mask pixel by pixel with each other.
+Whenever two pixel colors are not matching, a penalty value is assigned.
+The sum of all those penalties is then the total cost.
+This algorithm is an implementation of a signed area measure.
+It is very important to take into account the parental relationship between the cells, checking if
+if either one of the cells is a daughter of the other cell.
+If a cell has just divided, the resulting two daughter cells will cover precisely the space that the
+preceding agent had taken up.
+Let's assume as an example that one result has more cells than the other
+(Figure~\ref{fig:mask-difference-metric} (A,B)).
+When we compare the pixels of both images, we can relate the given color of the pixel to a certain
+cell.
+Whenever the colors of two distinct cells are in a parental relationship, we reduce the penalty
+which would be assigned or set it to zero.
+
+\begin{figure}
+    \centering
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {../docs/source/_static/fitting-methods/progressions-1.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {../docs/source/_static/fitting-methods/progressions-2.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+    \end{tikzonimage}
+    \linebreak
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {../docs/source/_static/fitting-methods/progressions-3.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {../docs/source/_static/fitting-methods/progressions-4.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+    \end{tikzonimage}
+    \caption{
+        The figure shows two masks (A,B) which will be compared against each other.
+        As we can see, image (B) contains more cells which has resulted from a division process that
+        is not yet occurred in image (A).
+        In (C) we can see the "naive" comparison, simply identifying non-matching colors.
+        By assigning penalties to all highlighted pixels, we would not take into consideration the
+        parental relationship.
+        This approach is displayed in (D) where pixels of cells that are related have been colored
+        in gray.
+    }
+    \label{fig:mask-difference-metric}
+\end{figure}
+
+We can use this procedure to calculate the cost between successive iteration steps
+$t_n$ with $t_{n+1}$.
+We expect to obtain an exponential growth in the penalty.
+This can be explained by to the fact that the growth of the total area is exponential in time.
+Furthremore, the temporal area difference can be viewed as a derivative but the derivative of an
+exponential is again an exponential.
+Figure~\ref{fig:parameter-estimates-single-step} shows the total penalty for this approach.
+We can see that by accounting for the parental relationship, we get the overall desired effect.
+This method allows us to compare results between states that have a different number of cells
+inside as long as their origin stems from the a shared initial state.
+However, since this algorithm requires us to calculate a cell-mask for each comparison and then also
+process this mask as well as the comparator, it is much more computationally demanding.
+In cases where the simpler approach of comparing positions of the agents is reasonable, this method
+should be preferred.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]
+        {../docs/source/_static/fitting-methods/penalty-time-flow.png}%
+    \caption{
+        Total cost between successive iteration steps $t_n,t_{n+1}$.
+        The overall shape is expected to have the slope of an exponential function.
+        We can clearly see that division events introduce jumps within this relationship.
+        In the case where we account for the parental relationship, these jumps are mitigated.
+        The penalty was set to $0$ for pixel-comparisons which belong to cells that are in a
+        parental relationship.
+    }
+    \label{fig:penalty-calculation-cell-division}
+\end{figure}
+
+% --------------------------------------------------------------------------------------------------
+\subsection{Estimating Parameters}
+\label{section:parameter-estimation}
+
+We will now focus on using the presented methods for parameter estimation via microscopic images.
+Our initial state is given by TODO~\cite{}.% TODO
+We use our extraction algorithm to obtain the initial positions of our cells.
+From there, we supply the model with parameters such that we can make a prediction for the final
+positions of the bacteria.
+We chose a time-interval which does not contain any division events.
+By doing so, we can utilize the simpler and more performant position comparison scheme for our cost
+function.
+Another important detail is the individuality of the agents.
+In our case, this means that each agent might have different parameters associated to it.
+For this dataset, we assumed that all interaction parameters which specify the interaction
+potential are identical across all agents but growth rates and the physical radii of the cells are
+individual and we also optimize them.
+We chose to optimize for 3 interaction parameters of the Mie potential $n,m,V_0$ and two parameters
+which were individual for each cell, namely the radii $r_i$ and growth rates $\mu_i$.
+We also included the damping constant $\lambda$ and removed any stochastic motion contributions
+($D=0$).
+This leads to a total of $4+2\times 6=16$ parameters.
+To estimate these parameters we used a latin-hypercube sweep across the parameter space and
+performed a local minimization with the Nelder-Mead method (TODO citations).% CITATIONS
+
+Figure~\ref{fig:parameter-estimates-single-step} shows the cost function of multiple simulations.
+The curves show the value at the found optimium and how it changes when varying the value of the
+parameter in question.
+We have exemplified the values for radii $r_i$ and growth rates $\mu_i$ with values of the first
+cell.
+The remaining plots can be found in the supplementary material in
+section~\ref{sec:supplement-parameter-profiles}.
+
+This depiction is similar to the method of profile likelihood~\cite{Raue2009}.
+We can see that almost all parameters can be estimated by our optimization approach.
+However, the damping constant exhibits the well-known phenomenon of
+unidentifiability~\cite{Raue2009}.
+This is of course problematic since the real value of $\lambda$ can not be idenified in this case.
+But it also shows that the approach which we used is similar in its ability to detect these defects
+in the first place.
+As we continue to develop these methods, we expect to be able to reuse many of the existing
+theoretical approaches which have already been developed for less complex systems such as \acp{ode}.
+Fortunately, \texttt{cellular\_raza} was chosen to construct our model such that we are flexible in
+changing the underlying representation if should find that such unidentifiabilities are structural
+(model-based) rather than practical (due to lack of data).
+This allows us to iterate and find the model formulation which suits the data best but still retain
+the mechanistic methodology of our approach.
+
+\begin{figure}[H]
+    \centering
+    \begin{tikzonimage}[width=0.33\textwidth]
+        {figures/crm_fit/profiles/damping.pdf}%
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.33\textwidth]
+        {figures/crm_fit/profiles/exponent-m.pdf}%
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.33\textwidth]
+        {figures/crm_fit/profiles/exponent-n.pdf}%
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+    \end{tikzonimage}\\
+    \begin{tikzonimage}[width=0.33\textwidth]
+        {figures/crm_fit/profiles/strength.pdf}%
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.33\textwidth]
+        {figures/crm_fit/profiles/growth-rate-0.pdf}%
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.33\textwidth]
+        {figures/crm_fit/profiles/radius-0.pdf}%
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
+    \end{tikzonimage}
+    \caption{
+        The curves show the value of the cost function at the optimium as determined by our
+        otimization scheme (blue dot).
+        We have chosen a single parameter and calculated the resulting cost function across a range
+        of values and plotted it.
+        In the cases of (B-F), we can see that a minimum was achieved and thus the parameters could
+        be identified.
+        However, for the case of the damping constant $\lambda$ in (A), a non-identifiability
+        arises.
+    }
+    \label{fig:parameter-estimates-single-step}
+\end{figure}
+
+%###################################################################################################
+\section{Multilayer Predictions}
+% TODO
+\begin{itemize}
+    \item Do we need this in this publication?
+    \item Maybe this is out of scope as this is a more methodological paper.
+    \item If we want to omit it, we might also opt to change the model description in the beginning
+        and remove some parts of the external forces discussion.
+\end{itemize}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.49\textwidth]
+        {../docs/source/_static/scripts/crm_multilayer/multilayer-time-evolution.pdf}\\
+    \includegraphics[width=0.49\textwidth]
+        {../docs/source/_static/scripts/crm_multilayer/colony-height-vs-gel_pressure.pdf}
+    \includegraphics[width=0.49\textwidth]
+        {../docs/source/_static/scripts/crm_multilayer/colony-height-vs-time.pdf}
+    \caption{TODO}
+    \label{fig:multilayer}
+\end{figure}
+
+%###################################################################################################
+\section{Discussion}
+\label{section:discussion}
+
+\begin{itemize}
+    \item Model does not capture every aspect of rod-shaped bacteria. Needs to extended (see our
+        review).
+    \begin{enumerate}
+        \item polar interactions
+        \item differentiation for van Gogh bundles; surfactin-producing, matrix-producing cells.
+        \item Proper friction; this might have a huge implication on the estimation of
+            parameters~\cite{Grant2014}.
+            This paper has already implemented this~\cite{Doumic2020}.
+            We can build on top of that.
+        \item Extracellular Reactions~\cite{Li2025}
+        \item see supplement for full list \ref{table:simulation-aspects-supplement}
+    \end{enumerate}
+    \item Image generation; Produce high-quality near-realistic microscopic images with cell-masks
+        from scratch with mechanistic model; train cell-segmentation and more importantly
+        cell-tracking algorithms with that.
+\end{itemize}
+
+\section{Conclusion}
+\label{section:conclusion}
+
+\begin{itemize}
+    \item Highlight methodological advancements
+    \item Constructed Model which already combines many effects; extend this even further
+    \item Proof of concept: Able to fit individual parameters of the agents; can be extend this?
+    \item Provide software package which combines all functionality and is readily available
+\end{itemize}
+
+%###################################################################################################
+% EXCLUDE THIS FOR NOW
+% \section{Plastic Deformations}
+% 
+% \begin{figure}[H]
+%     \centering
+%     \includegraphics[width=0.49\textwidth]
+%         {../docs/source/_static/scripts/crm_amir/0000000023.png}\\
+%     \includegraphics[width=0.49\textwidth]
+%         {../docs/source/_static/scripts/crm_amir/angles.pdf}
+%     \includegraphics[width=0.49\textwidth]
+%         {../docs/source/_static/scripts/crm_amir/endpoints.pdf}
+%     \caption{TODO}
+%     \label{fig:amir}
+% \end{figure}
+
+\bibliographystyle{IEEEtran}
+\bibliography{../docs/source/references}
+
+%###################################################################################################
+\section*{Supplementary Material}
+
+\renewcommand{\thesection}{S\arabic{section}}
+\setcounter{section}{0}
+
+%###################################################################################################
+\section{Parameter Estimation}
+\label{sec:supplement-parameter-profiles}
+%TODO {explain profiles}
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.33\textwidth]
+        {figures/crm_fit/profiles/growth-rate-0.pdf}%
+    \includegraphics[width=0.33\textwidth]
+        {figures/crm_fit/profiles/growth-rate-1.pdf}%
+    \includegraphics[width=0.33\textwidth]
+        {figures/crm_fit/profiles/growth-rate-2.pdf}\\%
+    \includegraphics[width=0.33\textwidth]
+        {figures/crm_fit/profiles/growth-rate-3.pdf}%
+    \includegraphics[width=0.33\textwidth]
+        {figures/crm_fit/profiles/growth-rate-4.pdf}%
+    \includegraphics[width=0.33\textwidth]
+        {figures/crm_fit/profiles/growth-rate-5.pdf}\\%
+    \includegraphics[width=0.33\textwidth]
+        {figures/crm_fit/profiles/radius-0.pdf}%
+    \includegraphics[width=0.33\textwidth]
+        {figures/crm_fit/profiles/radius-1.pdf}%
+    \includegraphics[width=0.33\textwidth]
+        {figures/crm_fit/profiles/radius-2.pdf}\\%
+    \includegraphics[width=0.33\textwidth]
+        {figures/crm_fit/profiles/radius-3.pdf}%
+    \includegraphics[width=0.33\textwidth]
+        {figures/crm_fit/profiles/radius-4.pdf}%
+    \includegraphics[width=0.33\textwidth]
+        {figures/crm_fit/profiles/radius-5.pdf}%
+        \caption{}%TODO insert caption}
+    \label{fig:parameter-estimates-supplement}
+\end{figure}
+
+%###################################################################################################
+\section{More Simulation Aspects}
+\label{section:supplement-more-simulation-aspects}
+%TODO insert this into the discussion later
+\begin{table}[H]
     \centering
@@
     \end{tabularx}
-    \label{table:simulation-aspects}
+    \label{table:simulation-aspects-supplement}
     \caption{TODO Insert caption}
@@
 
-% --------------------------------------------------------------------------------------------------
-\subsection{Mechanics}
+%###################################################################################################
+\section{Performance}
 
-%TODO use these citations
-\cite{Chatterjee1988},\cite{Ursell2014}
+\begin{itemize}
+    \item Quadratic Scaling comes from more dense simulation
+    \item Overall number of agents which can be simulated is satisfactory for current purposes
+    \item Improvement from $1 \Rightarrow 2$ and $2\Rightarrow4$ threads; first is major; second
+        only minor
+    \item Flamegraph still shows much cycles running against hurdles::Barrier::wait $\Rightarrow$
+        this leaves room for improvement
+    \item Flamegraph also shows that $\approx 75\%$ of all cycles is related to actual work (not
+        just waiting).
+        In particular to update\_mechanics functions.
+\end{itemize}
 
-In principle we can assign individual lengths $\{l_i\}$ and strengths $\{\gamma\}_i$ to each spring.
-The internal force acting on vertex $\vec{v}_i$ can be divided into 2 contributions coming
-from the 2 springs pulling on it.
-In the case when $i=0,N_\text{vertices}$, this is reduced to only one internal component.
-We denote with $\vec{c}_{i}$ the connection between two vertices
-
-\begin{equation}
-    \vec{c}_i = \vec{v}_{i}-\vec{v}_{i-1}
-\end{equation}
-
-and can write down the resulting force
-
-\begin{align}
-    \vec{F}_{i,\text{springs}} =
-        &-\gamma_i\left(1 - \frac{l_i}{\left|\vec{c}_i\right|}\right)\vec{c}_i\\
-        &+ \gamma_{i+1}\left(1 - \frac{l_{i+1}}{\left|\vec{c}_{i+1}\right|}\right)\vec{c}_{i+1}
-    \label{force-springs}
-\end{align}
-
-In addition to springs between individual vertices $\vec{v}_i$, we assume that each angle at
-a vertex between two other is subject to a curvature force.
-Assuming that $\alpha_i$ is the angle between the connections and
-$\vec{d}_i=\vec{c}_i/|\vec{c}_i|$ is the normalized connection,
-we can write down the forces acting on vertices $\vec{v}_i,\vec{v}_{i-1},\vec{v}_{i+1}$
-
-\begin{align}
-    \vec{F}_{i,\text{curvature}} &= \eta_i\tan\left(\pi-\alpha_i\right)
-        \frac{\vec{d}_i - \vec{d}_{i+1}}{|\vec{d}_i-\vec{d}_{i+1}|}\\
-    \vec{F}_{i-1,\text{curvature}} &= -\frac{1}{2}\vec{F}_{i,\text{curvature}}\\
-    \vec{F}_{i+1,\text{curvature}} &= -\frac{1}{2}\vec{F}_{i,\text{curvature}}
-    \label{force-curvature}
-\end{align}
-
-where $\eta_i$ is the curvature at vertex $\vec{v}_i$.
-We can see that the curvature force does not move the overall center of the cell in space.
-
-\begin{figure}
+\begin{figure}[H]
     \centering
-    \includegraphics[width=0.5\textwidth]
-        {../docs/source/_static/mechanics.png}
-    \caption{Example of a polygonal representation of a rod-shaped bacterium.}
-    \label{fig:mechanics-bacterium}
+    \includegraphics[width=0.49\textwidth]
+        {../docs/source/_static/performance/computation-time-with-initial-agents.pdf}
+    \includegraphics[width=0.49\textwidth]{../docs/source/_static/performance/flamegraph.pdf}
+    \caption{TODO}
+    \label{fig:performance}
 \end{figure}
 
-The total force is the sum of external and interal forces.
-
-\begin{equation}
-    \vec{F}_{i,\text{total}} = \vec{F}_{i,\text{springs}}+ \vec{F}_{i,\text{curvature}} + \vec{F}_{i,\text{external}}
-   \label{force-total}
-\end{equation}
-
-and are integrated via
-
-\begin{align}
-    \partial_t^2 \vec{x} &= \partial\vec{x} + D\vec{\xi}\\
-    \partial_t\vec{x} &= \vec{F}_\text{total}
-    \label{equations-of-motion}
-\end{align}
-
-where $D$ is the diffusion constant and  $\vec{\xi}$ is the wiener process (compare with
-brownian motion such as given by the
-% Brownian3D <https://cellular-raza.com/docs/cellular_raza_building_blocks/struct.Brownian3D.html>_
-% struct of cellular-raza <https://cellular-raza.com>.
-
-% --------------------------------------------------------------------------------------------------
-\subsection{Interaction}
-
-When calculating forces acting between the cells, we can use a simplified model to circumvent the
-numerically expensive integration over the complete length of the rod.
-Given a vertex $\vec{v}_i$ on one cell, we calculate the closest point $\vec{p}$ on the polygonal
-line given by the vertices $\{\vec{w}_j\}$ of the interacting cell.
-Furthermore we determine the value $q\in[0,1]$ such that
-
-\begin{equation}
-    \vec{p} = (1-q)\vec{w}_j + q\vec{w}_{j+1}
-   \label{connection}
-\end{equation}
-
-for some specific $j$.
-The force is then calculated between the points $\vec{v}_i$ and $\vec{p}_i$ and acts on
-the vertex $\vec{w}_i,\vec{w}_{i+1}$ with relative strength $(1-q)$ and $q$.
-
-\begin{equation}
-    \vec{F}_{i,\text{External}} = \vec{F}(\vec{v}_i,\vec{p})
-    \label{force-external}
-\end{equation}
-
-% .. _ TODO MORSEPOTENTIAL
-
-% .. _ TODO insert table with all parameters
-
-% --------------------------------------------------------------------------------------------------
-\subsection{Cycle}
-
-To simulate proliferation, we introduce a growth term for the spring lengths $l_i$
-
-\begin{equation}
-    \partial_t l_i = \mu
-    \label{growth-ode}
-\end{equation}
-
-which will increase the length of the cell indefenitely unless we introduce a
-% `division event <https://cellular-raza.com/internals/concepts/cell/cycle>`_.
-We define a threshold (in our case double of the original length) for the total length of the
-cell at which it divides.
-To construct a new cell, we cannot simply copy the existing one twice, but we also need to adjust
-internal parameters in the process.
-The following actions need to be taken for the old and new agent.
-
-% .. _ TODO
-
-1. Assign a new growth rate (pick randomly from uniform distribution in $[0.8\mu_0,1.2\mu_0]$
-   where $\mu_0$ is some fixed value)
-2. Assign new positions
-    1. Calculate new spring lengths
-        $\tilde{l}_i = l_i\left(\frac{1}{2} - \frac{r}{\sum\limits_i l_i}\right)$
-    2. Calculate middle of old cell
-        $\vec{m} = \frac{1}{N_\text{vertices}}\sum\limits_i\vec{v}_i$
-    3. Calculate positions of new vertices $\vec{w}_i$
-
-\begin{align}
-    q_i &= \frac{i}{N_\text{vertices}}\\
-    \vec{w}_{i,\text{new},\pm} &= (1-q_i)\vec{m} + q_i(\vec{v}_{\pm\text{start}} - \vec{m})
-\end{align}
-
-%###################################################################################################
-\section{Generating Masks and Images}
-\label{section:generating-masks-and-images}
-
-\begin{figure}[!h]
-    \centering
-    \begin{tikzonimage}[width=0.3\textwidth]
-        {../docs/source/_static/09395645494836445480/raw_pv/000000400.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{a}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.3\textwidth]
-        {../docs/source/_static/09395645494836445480/images/000000400.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{b}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.3\textwidth]
-        {../docs/source/_static/09395645494836445480/masks/000000400.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{c}};
-    \end{tikzonimage}
-    \caption{Snapshots at \dots}
-    \label{fig:progression-image-generation}
-\end{figure}
-
-%###################################################################################################
-\section{Extract Positions from Masks}
-\label{section:fitting-and-comparing-masks}
-
-\begin{figure}
-    \centering
-    \begin{tikzonimage}[width=0.32\textwidth]
-        {../docs/source/_static/fitting-methods/algorithm/image001042.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{a}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.32\textwidth]
-        {../docs/source/_static/fitting-methods/algorithm/mask-zoom.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{b}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.32\textwidth]
-        {../docs/source/_static/fitting-methods/algorithm/mask-zoom.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{c}};
-    \end{tikzonimage}
-    \caption{}
-    \label{fig:position-extraction-algorithm}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \begin{tikzonimage}[width=0.32\textwidth]
-        {../docs/source/_static/fitting-methods/extract_positions-004000.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{a}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.32\textwidth]
-        {../docs/source/_static/fitting-methods/extract_positions-006000.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{b}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.32\textwidth]
-        {../docs/source/_static/fitting-methods/extract_positions-008000.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{c}};
-    \end{tikzonimage}
-    \caption{Progression of generated masks with estimated and exact positions.}
-    \label{fig:position-extraction}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.5\textwidth]
-        {../docs/source/_static/fitting-methods/displacement-calculations.png}
-    \caption{Comparison of masks across generations. Compare estimated with exact rod lengths.}
-    \label{fig:benchmarking-fitting-algorithm}
-\end{figure}
-
-%###################################################################################################
-\section{Estimating Parameters}
-\label{section:parameter-estimation}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.33\textwidth]
-        {../docs/source/_static/fitting-methods/estimate-parameters1/microscopic-images-0.png}%
-    \includegraphics[width=0.33\textwidth]
-        {../docs/source/_static/fitting-methods/estimate-parameters1/microscopic-images-1.png}%
-    \includegraphics[width=0.33\textwidth]
-        {../docs/source/_static/fitting-methods/estimate-parameters1/Growth Rate.png}\\%
-    \includegraphics[width=0.33\textwidth]
-        {../docs/source/_static/fitting-methods/estimate-parameters1/microscopic-images-2.png}%
-    \includegraphics[width=0.33\textwidth]
-        {../docs/source/_static/fitting-methods/estimate-parameters1/microscopic-images-3.png}%
-    \includegraphics[width=0.33\textwidth]
-        {../docs/source/_static/fitting-methods/estimate-parameters1/Potential Stiffness.png}\\
-    \includegraphics[width=0.33\textwidth]
-        {../docs/source/_static/fitting-methods/estimate-parameters1/Radius.png}%
-    \includegraphics[width=0.33\textwidth]
-        {../docs/source/_static/fitting-methods/estimate-parameters1/Rigidity.png}%
-    \includegraphics[width=0.33\textwidth]
-        {../docs/source/_static/fitting-methods/estimate-parameters1/Strength.png}%
-    \caption{Profiles of the Parameters}
-    \label{fig:parameter-estimates-single-step}
-\end{figure}
-
-% --------------------------------------------------------------------------------------------------
-\subsection{Comparing Results Across Generations}
-
-\begin{figure}
-    \centering
-    \begin{tikzonimage}[width=0.3\textwidth]
-        {../docs/source/_static/fitting-methods/progressions-1.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{a}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.3\textwidth]
-        {../docs/source/_static/fitting-methods/progressions-2.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{b}};
-    \end{tikzonimage}
-    \linebreak
-    \begin{tikzonimage}[width=0.3\textwidth]
-        {../docs/source/_static/fitting-methods/progressions-3.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{c}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.3\textwidth]
-        {../docs/source/_static/fitting-methods/progressions-4.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{d}};
-    \end{tikzonimage}
-    \caption{TODO}
-    \label{fig:mask-difference-metric}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.5\textwidth]
-        {../docs/source/_static/fitting-methods/penalty-time-flow.png}%
-    \caption{Comparison of masks across generations.}
-    \label{fig:penalty-calculation-cell-division}
-\end{figure}
-
-\bibliographystyle{IEEEtran}
-\bibliography{../docs/source/references}
-
 \end{document}

### 2025-06-19 src/crm_fit/optimize.rs
refactor optimization; use nelder-mead after latin-hypercube global sampling

@@ -89,6 +89,58 @@ fn lhs_optimization(
     Ok(result)
 }
 
+fn lhs_optimization_iter(
+    py: Python,
+    n_points: usize,
+    n_steps: usize,
+    bounds: &numpy::ndarray::Array2<f32>,
+    iterations_images: &[usize],
+    positions_all: numpy::ndarray::ArrayView4<f32>,
+    settings: &Settings,
+    relative_reduction: f32,
+) -> PyResult<Option<(Vec<f32>, f32)>> {
+    // Do lowering via repeated applications of LatinHypercube algorithm
+    let mut bounds = bounds.clone();
+    let mut result = None;
+
+    for n in 0..n_steps {
+        let new_result = lhs_optimization(
+            py,
+            n_points,
+            n,
+            &bounds,
+            &iterations_images,
+            positions_all,
+            settings,
+        )?;
+        // If the newly calculated results are better, we update. Otherwise do nothing.
+        // This will still decrease the overall size
+        if result.is_none() && new_result.is_some() {
+            result = new_result;
+        } else if let (Some((_, c0)), Some((p1, c1))) = (&result, new_result) {
+            if c1 < *c0 {
+                result = Some((p1, c1));
+            }
+        }
+
+        if let Some((params, _)) = &result {
+            let dists = &bounds.column(1) - &bounds.column(0);
+            for i in 0..params.len() {
+                let d = dists[i];
+                let q: f32 = d * relative_reduction / 2.0;
+                let b0 = params[i] - q;
+                let b1 = params[i] + q;
+                bounds[(i, 0)] = bounds[(i, 0)].max(b0);
+                bounds[(i, 1)] = bounds[(i, 1)].min(b1);
+            }
+        } else {
+            break;
+        }
+    }
+
+    Ok(result)
+}
+
 #[pyfunction]
 #[pyo3(signature = (iterations_images, positions_all, settings, n_workers=-1))]
 pub fn run_optimizer(
@@ -197,43 +249,16 @@ res = sp.optimize.differential_evolution(
                 .build_global()
                 .map_err(|e| pyo3::exceptions::PyValueError::new_err(format!("{e}")))?;
 
-            // Do lowering via repeated applications of LatinHypercube algorithm
-            let mut bounds = bounds;
-            let mut result = None;
-            for n_iteration in 0..*n_steps {
-                let new_result = lhs_optimization(
-                    py,
-                    *n_points,
-                    n_iteration,
-                    &bounds,
[…]

### 2025-06-19 src/crm_fit/settings.rs

@@ -242,10 +242,13 @@ pub struct LatinHypercube {
 #[pyclass(get_all, set_all, module = "cr_mech_coli.crm_fit")]
 #[derive(Clone, Debug, Serialize, Deserialize, AbsDiffEq, PartialEq)]
 #[approx(epsilon_type = f32)]
-pub struct NelderMead {
+pub struct LHSNelderMead {
     /// Maximum number of iterations
     #[approx(equal)]
     pub max_iter: usize,
+    /// Performs a LatinHypercube sweep before applying the Nelder-Mead method
+    #[serde(rename = "latin_hypercube")]
+    pub latin_hypercube: Option<LatinHypercube>,
 }
 
 /// Other settings which are not related to the outcome of the simulation
@@ -348,8 +351,8 @@ pub enum OptimizationMethod {
     #[serde(rename = "latin_hypercube")]
     LatinHypercube(LatinHypercube),
     /// Uses [egobox_ego] to optimize
-    #[serde(rename = "ego")]
-    NelderMead(NelderMead),
+    #[serde(rename = "lhs_nelder_mead")]
+    LHSNelderMead(LHSNelderMead),
 }
 
 /// Return type of the :meth:`Settings.generate_optimization_infos` method.

### 2025-06-19 src/crm_fit/settings.rs
do not produce subscript

@@ -567,7 +567,7 @@ impl Settings {
                                     .map(|i| (
                                         format!("{} {}", $var_name, i),
                                         $units.to_string(),
-                                        format!("{}_{{{}}}", $symbol, i),
+                                        format!("{}", $symbol),
                                     ))
                             );
                             initial_values.extend(vec![initial.clone(); n_agents]);

### 2025-06-19 paper/main.tex
move references to this directory
change path; use local

@@
     \centering
-    \includegraphics[width=0.5\textwidth]{../docs/source/_static/mechanics.png}
+    \includegraphics[width=0.5\textwidth]{docs/source/_static/mechanics.png}
     \caption{
@@
     \begin{tikzonimage}[width=0.3\textwidth]
-        {../docs/source/_static/09395645494836445480/raw_pv/000000400.png}
+        {docs/source/_static/09395645494836445480/raw_pv/000000400.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
@@
     \begin{tikzonimage}[width=0.3\textwidth]
-        {../docs/source/_static/09395645494836445480/images/000000400.png}
+        {docs/source/_static/09395645494836445480/images/000000400.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
@@
     \begin{tikzonimage}[width=0.3\textwidth]
-        {../docs/source/_static/09395645494836445480/masks/000000400.png}
+        {docs/source/_static/09395645494836445480/masks/000000400.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
@@
     \begin{tikzonimage}[width=0.32\textwidth]
-        {../docs/source/_static/fitting-methods/algorithm/image001042.png}
+        {docs/source/_static/fitting-methods/algorithm/image001042.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
@@
     \begin{tikzonimage}[width=0.32\textwidth]
-        {../docs/source/_static/fitting-methods/algorithm/mask-zoom.png}
+        {docs/source/_static/fitting-methods/algorithm/mask-zoom.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
@@
     \begin{tikzonimage}[width=0.32\textwidth]
-        {../docs/source/_static/fitting-methods/algorithm/mask-zoom.png}
+        {docs/source/_static/fitting-methods/algorithm/mask-zoom.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
@@
     \begin{tikzonimage}[width=0.32\textwidth]
-        {../docs/source/_static/fitting-methods/extract_positions-004000.png}
+        {docs/source/_static/fitting-methods/extract_positions-004000.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
@@
     \begin{tikzonimage}[width=0.32\textwidth]
-        {../docs/source/_static/fitting-methods/extract_positions-006000.png}
+        {docs/source/_static/fitting-methods/extract_positions-006000.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
@@
     \begin{tikzonimage}[width=0.32\textwidth]
-        {../docs/source/_static/fitting-methods/extract_positions-008000.png}
+        {docs/source/_static/fitting-methods/extract_positions-008000.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
@@
     \begin{tikzonimage}[width=0.5\textwidth]
-        {../docs/source/_static/fitting-methods/displacement-calculations.png}%
+        {docs/source/_static/fitting-methods/displacement-calculations.png}%
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
@@
     \begin{tikzonimage}[width=0.5\textwidth]
-        {../docs/source/_static/fitting-methods/displacement-distribution.png}
+        {docs/source/_static/fitting-methods/displacement-distribution.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
@@
     \begin{tikzonimage}[width=0.3\textwidth]
-        {../docs/source/_static/fitting-methods/progressions-1.png}
+        {docs/source/_static/fitting-methods/progressions-1.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
@@
     \begin{tikzonimage}[width=0.3\textwidth]
-        {../docs/source/_static/fitting-methods/progressions-2.png}
+        {docs/source/_static/fitting-methods/progressions-2.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
@@
     \begin{tikzonimage}[width=0.3\textwidth]
-        {../docs/source/_static/fitting-methods/progressions-3.png}
+        {docs/source/_static/fitting-methods/progressions-3.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
@@
     \begin{tikzonimage}[width=0.3\textwidth]
-        {../docs/source/_static/fitting-methods/progressions-4.png}
+        {docs/source/_static/fitting-methods/progressions-4.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{D}};
@@
     \includegraphics[width=0.5\textwidth]
-        {../docs/source/_static/fitting-methods/penalty-time-flow.png}%
+        {docs/source/_static/fitting-methods/penalty-time-flow.png}%
     \caption{
@@
     \includegraphics[width=0.49\textwidth]
-        {../docs/source/_static/scripts/crm_multilayer/multilayer-time-evolution.pdf}\\
+        {docs/source/_static/scripts/crm_multilayer/multilayer-time-evolution.pdf}\\
     \includegraphics[width=0.49\textwidth]
-        {../docs/source/_static/scripts/crm_multilayer/colony-height-vs-gel_pressure.pdf}
+        {docs/source/_static/scripts/crm_multilayer/colony-height-vs-gel_pressure.pdf}
     \includegraphics[width=0.49\textwidth]
-        {../docs/source/_static/scripts/crm_multilayer/colony-height-vs-time.pdf}
+        {docs/source/_static/scripts/crm_multilayer/colony-height-vs-time.pdf}
     \caption{TODO}
@@
 %     \includegraphics[width=0.49\textwidth]
-%         {../docs/source/_static/scripts/crm_amir/0000000023.png}\\
+%         {docs/source/_static/scripts/crm_amir/0000000023.png}\\
 %     \includegraphics[width=0.49\textwidth]
-%         {../docs/source/_static/scripts/crm_amir/angles.pdf}
+%         {docs/source/_static/scripts/crm_amir/angles.pdf}
 %     \includegraphics[width=0.49\textwidth]
-%         {../docs/source/_static/scripts/crm_amir/endpoints.pdf}
+%         {docs/source/_static/scripts/crm_amir/endpoints.pdf}
 %     \caption{TODO}
@@
 \bibliographystyle{IEEEtran}
-\bibliography{../docs/source/references}
+\bibliography{references}
 
@@
     \includegraphics[width=0.49\textwidth]
-        {../docs/source/_static/performance/computation-time-with-initial-agents.pdf}
-    \includegraphics[width=0.49\textwidth]{../docs/source/_static/performance/flamegraph.pdf}
+        {docs/source/_static/performance/computation-time-with-initial-agents.pdf}
+    \includegraphics[width=0.49\textwidth]{docs/source/_static/performance/flamegraph.pdf}
     \caption{TODO}

### 2025-06-23 paper/main.tex
add plot for overview of project

@@
 \usepackage{acronym}
+\usepackage{xcolor}
 
 \usetikzlibrary{trees}
+\usetikzlibrary{shapes.geometric, arrows}
+
+\definecolor{color1}{HTML}{6bd2db}
+\definecolor{color2}{HTML}{0ea7b5}
+\definecolor{color3}{HTML}{0c457d}
+\definecolor{color4}{HTML}{ffbe4f}
+\definecolor{color5}{HTML}{e8702a}
 
@@
 \end{itemize}
+
+\begin{figure}
+    \centering
+    \tikzstyle{rect} = [
+        rectangle,
+        % rounded corners,
+        minimum width=4cm,
+        minimum height=1cm,
+        text centered,
+        draw=black,
+        fill=color4!70,
+    ]
+    \tikzstyle{frect} = [rect, fill=color3!40]
+    \tikzstyle{arrow} = [thick,->,>=stealth]
+    \tikzstyle{darrow} = [arrow, dashed]
+    \begin{tikzpicture}
+        \node (model) [rect, align=center] at (0,0) {Mathematical \&\\ Computational Model};
+        \node (extraction) [rect] at (6, 0) {Data Extraction};
+        \node (params) [rect] at (6, -2) {Parameter Estimation};
+        \node (visualization) [rect] at (0, -2) {Visualization};
+
+        \draw[arrow] (model.east) -- (extraction.west);
+        \draw[darrow] (params.north) -- node[anchor=south, rotate=-10] {validate} (model.south);
+        \draw[arrow] (model.south) -- (visualization.north);
+        \draw[arrow] (visualization.east) -- (params.west);
+        \draw[arrow] (extraction.south) -- (params.north);
+
+        % Future stuff
+        \node (realimages) [frect] at (0, -4) {Data Generation};
+        \node (segmentation) [frect] at (6, -3.5) {Cell Segmentation};
+        \node (tracking) [frect] at (6, -4.5) {Cell Tracking};
+
+        \draw[arrow] (visualization.south) -- (realimages.north);
+        \draw[arrow] (realimages) -- (segmentation.west);
+        \draw[arrow] (realimages) -- (tracking.west);
+    \end{tikzpicture}
+    \caption{
+        The project can be divided into 4 subparts (yellow) that interact with and depend on each
+        other.
+        Additional blocks (blue-gray) indicate possible future extensions of this project.
+        Arrows with solid lines indicate dependencies between the subparts while dashed arrows carry
+        special meaning, which is specifically annotated.
+        The mathematical models provides the theoretical framework and is thus the parent of all
+        following work.
+        "Data Extraction" allows us to read variable and parameter values directly from microscopic
+        images.
+        The obtained data can then be used to initialize a numerical simulations and thus estimate
+        parameters.
+        On the other hand, results produced by our model can be visualized as 3D renders and
+        projected to cell-masks.
+        These algorithms can be exploited to construct cost-functions which can be used in the
+        parameter estimation process.
+        In future work, this visualization scheme could be extended to generate realistic
+        microscopic images with cell-masks and thus train cell-segmentation and cell-tracking
+        algorithms.
+    }
+    \label{fig:flowchart-project-structure}
+\end{figure}
 

### 2025-07-07 paper/main.tex
change title
change title of paper
rework placing of figures

@@
 
-\title{A mechanical Model of Rod-Shaped Bacteria}
+\title{Methods for Analyzing Models of Rod-Shaped Bacteria}
 
@@
 %\renewcommand{\undertitle}{Technical Report}
-\renewcommand{\shorttitle}{A mechanical Model of Rod-Shaped Bacteria}
+\renewcommand{\shorttitle}{Methods for Analyzing Models of Rod-Shaped Bacteria}
 
@@
         \node (model) [rect, align=center] at (0,0) {Mathematical \&\\ Computational Model};
-        \node (extraction) [rect] at (6, 0) {Data Extraction};
+        \node (extraction) [rect] at (0, -2) {Data Extraction};
         \node (params) [rect] at (6, -2) {Parameter Estimation};
-        \node (visualization) [rect] at (0, -2) {Visualization};
-
-        \draw[arrow] (model.east) -- (extraction.west);
-        \draw[darrow] (params.north) -- node[anchor=south, rotate=-10] {validate} (model.south);
-        \draw[arrow] (model.south) -- (visualization.north);
-        \draw[arrow] (visualization.east) -- (params.west);
-        \draw[arrow] (extraction.south) -- (params.north);
+        \node (visualization) [rect] at (6, 0) {Visualization};
+
+        \draw[arrow] (model) -- (extraction);
+        \draw[darrow] (params) -- node[anchor=south, rotate=-18.43] {validate} (model);
+        \draw[arrow] (model) -- (visualization);
+        \draw[arrow] (visualization) -- (params);
+        \draw[arrow] (extraction) -- (params);
 
         % Future stuff
-        \node (realimages) [frect] at (0, -4) {Data Generation};
-        \node (segmentation) [frect] at (6, -3.5) {Cell Segmentation};
-        \node (tracking) [frect] at (6, -4.5) {Cell Tracking};
-
-        \draw[arrow] (visualization.south) -- (realimages.north);
-        \draw[arrow] (realimages) -- (segmentation.west);
-        \draw[arrow] (realimages) -- (tracking.west);
+        \node (realimages) [frect] at (12, 0) {Data Generation};
+        \node (segmentation) [frect] at (12, -2) {Cell Segmentation};
+        \node (tracking) [frect] at (12, -3) {Cell Tracking};
+        \node (analysis) [frect] at (0, -4) {Analysis Methods};
+
+        \draw[arrow] (visualization) -- (realimages);
+        \draw[arrow] (realimages) -- (segmentation);
+        \draw[arrow] (extraction) -- (analysis);
+        \draw[arrow] (params) -- (realimages);
     \end{tikzpicture}

### 2025-07-07 paper/main.tex
split into two separate figures

@@
     \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/growth-rate-5.pdf}\\%
+        {figures/crm_fit/profiles/growth-rate-5.pdf}
+    \caption{Parameter estimation of growth rates for individual cells}%TODO extend caption}
+    \label{fig:parameter-estimates-supplement}
+\end{figure}
+\begin{figure}[H]
+    \centering
     \includegraphics[width=0.33\textwidth]
@@
         {figures/crm_fit/profiles/radius-5.pdf}%
-        \caption{}%TODO insert caption}
+    \caption{Parameter estimation of radii for individual cells}%TODO extend caption}
     \label{fig:parameter-estimates-supplement}

### 2025-07-07 plots/index_simulation.py
rework script

@@ -1,18 +1,27 @@
 import cr_mech_coli as crm
 
 if __name__ == "__main__":
-    config = crm.Configuration(
-        n_agents=4,
-    )
+    config = crm.Configuration()
     agent_settings = crm.AgentSettings(
         growth_rate=0.078947365,
     )
     config.t0 = 0.0
     config.dt = 0.1
     config.t_max = 100.0
-    config.save_interval = 20.0
+    config.n_saves = 5
+    config.domain_size = (150, 150)
 
-    sim_result = crm.run_simulation(config, agent_settings)
+    positions = crm.generate_positions(
+        n_agents=4,
+        agent_settings=agent_settings,
+        config=config,
+        rng_seed=3,
+        dx=(config.domain_size[0] * 0.1, config.domain_size[1] * 0.1),
+    )
+    agent_dict = agent_settings.to_rod_agent_dict()
+    agents = [crm.RodAgent(p, 0.0 * p, **agent_dict) for p in positions]
+
+    sim_result = crm.run_simulation_with_agents(config, agents)
     render_settings = crm.RenderSettings()
     render_settings.noise = 50
     render_settings.kernel_size = 30

### 2025-07-10 paper/main.tex
adjust formulation
add idea for discussion

@@
 S.pombe represents the latter.
-The material is inserted in small bursts on the nanoscale in forms of patches, bands or
+New material is inserted in small bursts on the nanoscale in forms of patches, bands or
 hoops~\cite{DePedro2003}.
@@
     \end{enumerate}
+    \item Data Extraction: Skeletonization Algorithm ultimately not good enough for some scenarios
+        (i.e. overlapping of cells, endpoints can sometimes be "too short")
+    \begin{itemize}
+        \item Possibly use AI for new algorithm
+    \end{itemize}
     \item Image generation; Produce high-quality near-realistic microscopic images with cell-masks

### 2025-07-11 paper/main.tex
completely restructure paper; focus on methods; use estimation results only as an example

@@
     \begin{enumerate}
-        \item Initial (incomplete) model of rod-shaped bacteria
         \item Methods of how to obtain data from microscopic images and perform parameter
@@
             estimation via microscopic images and multilayer formation
+        \item Initial (incomplete) model of rod-shaped bacteria
         \item Generate 3D Images (masks)
@@
     \end{enumerate}
-    \item Make clear that this is mainly a proof of the methods
+    \item Overall structure
+    \begin{enumerate}
+        \item users bring model + data
+        \item this paper provides methods to do the parameter estimation + visualization
+    \end{enumerate}
     \item For complete model, more work has to be done
@@
 
-\begin{figure}
+\begin{figure}[H]
     \centering
@@
     ]
+    \tikzstyle{brect} = [rect, fill=color5!50]
     \tikzstyle{frect} = [rect, fill=color3!40]
@@
     \begin{tikzpicture}
-        \node (model) [rect, align=center] at (0,0) {Mathematical \&\\ Computational Model};
+        \node (model) [brect, align=center] at (0,0) {Mathematical \&\\ Computational Model};
         \node (extraction) [rect] at (0, -2) {Data Extraction};
-        \node (params) [rect] at (6, -2) {Parameter Estimation};
+        \node (params) [brect] at (6, -2) {Parameter Estimation};
         \node (visualization) [rect] at (6, 0) {Visualization};
@@
     \caption{
-        The project can be divided into 4 subparts (yellow) that interact with and depend on each
-        other.
+        The project can be divided into 4 subparts (orange + yellow) that interact with and depend
+        on each other.
+        The orange blocks require input by the user by specifying the model and supplying input
+        data.
         Additional blocks (blue-gray) indicate possible future extensions of this project.
@@
 %###################################################################################################
-\section{A mechanical Model of Rod-Shaped Bacteria}
-\label{section:mechanical-model-rod-shaped-bacteria}
+\section{Modeling Prerequesites}
+\begin{enumerate}
+    \item represent bacteria as polygon with vertices $\vec{v}_i$
+    \item Assumes that these vertices are not fundamental but rather a discretization of the actual
+        problem
+\end{enumerate}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.5\textwidth]{docs/source/_static/mechanics.png}
+    \caption{
+        Example of an elonged bacterial rod represented by a collection of vertices which are
+        connected by springs.
+        The angles $\alpha_i$ have been exaggerated such that they can be identified more easily.
+        A bacterial agent which is not interacting with any external objects will tend to the
+        equilibrium state of a straight rod with $0$ curvature.
+    }
+    \label{fig:mechanics-bacterium}
+\end{figure}
+
+%###################################################################################################
+\section{Data Extraction}
+\label{section:fitting-and-comparing-masks}
+
+This section will explain how we can determine the collection of vertices $\vec{v}_i$ introduced in
+subsection~\ref{subsection:mechanical-model-mechanics} from microscopic images
+(Figure~\ref{fig:position-extraction-algorithm} (A)).
+They will serve as initial, final and intermediate values with which we can compare our predictions.
+The algorithms which we devised requires the input of cell-masks which assign a unique color for
+every cell in the image.
+Such masks can be generated by cell-segmentation tools or obtained by annottating microscopic images
+by hand (Figure~\ref{fig:position-extraction-algorithm} (B)).
+Afterwards, we proceed with the following 5 steps.
+\begin{enumerate}
+    \item Obtain submasks for each cell individually.
+    \item Skeletonize submask with algorithm from Lee~\cite{Lee1994}
+    \item Determine endpoints $\vec{q}_0,\vec{q}_1$.
+        Assert that there are only 2 endpoints.
+    \item Sort points such that they form a polygon.
+    \item Interpolate vertices $\vec{v}_i$ from obtained polygon.
+\end{enumerate}
+In the first step, the image will be split into a collection of subimages that only contain
+information for one single cell.
+From there, we apply a skeletonization algorithm~\cite{Lee1994}
+(Figure~\ref{fig:position-extraction-algorithm} (C)).
+It has to be noted that there are multiple variants of skeletonization and in general, these classes
+of algorithms can produce results which have intersection points and multiple endpoints.
+This described behaviour is undesirable in our case.
+We have found that for most practical examples, the algorithm devised by~\cite{Lee1994} has worked
+well.\\
+The skeletonization algorithm supplies us with a collection of unordered lattice points $\vec{p}_i$
+(pixel coordinates) where the skeleton lives.
+We identify the endpoints $\vec{q}_0,\vec{q}_1$ by counting the number of neighbors in the von
+Neumann neighborhood for each lattice point.
+Any point with exactly one neighboring point must be and endpoint.
+If we encounter more or less than two points which fit this criterion, the routine can not commence
+and is aborted.
+Next, we start with one of the endpoints and determine the next neighbor which was not yet picked.
+We list the points in this order, thus forming a sorted polygon of pixel coordinates.\\
+In the last step, we take the sorted, extracted points $\vec{p}_i$ and interpolate $n$ values for
+the desired number of vertices $\vec{v}_i$.
+
+\begin{figure}
+    \centering
+    \begin{tikzonimage}[width=0.32\textwidth]
+        {docs/source/_static/fitting-methods/algorithm/image001042.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.32\textwidth]
+        {docs/source/_static/fitting-methods/algorithm/mask-zoom.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.32\textwidth]
+        {docs/source/_static/fitting-methods/algorithm/mask-zoom.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+    \end{tikzonimage}
+    \caption{
+        TODO change 2nd image!\\
+        (A) An example of a microscopic image.
+        (B) The associated mask which assigns individual color values to each bacterium.
+        (C) The result of the  skeletonization algorithm when applying it to a single cell.
+    }
+    \label{fig:position-extraction-algorithm}
+\end{figure}
+
+% --------------------------------------------------------------------------------------------------
+\paragraph{Benchmarking the Extraction Algorithm}
+\label{subsection:benchmarking-extraction-algorithm}
+
+In order to determine if our extraction algorithm provides reliable results, we have to take
+advantage of a functionality which will be introduced later in
+section~\ref{section:generating-masks-and-images}.
+Our methodology makes use of the fact that we have already set up our model and solving scheme and
+are able to produce cell masks from these numerical results.
+In doing so, we have exact knowledge about the position of our agents since we can simply obtain
+this information from our model.
+But on the other hand we can also try to extract this position from the generated cell-masks.
+By comparing the differences between them, we get a good estimate of how well our extraction
+algorithm can determine the true position of the bacteria.
+
+In figure~\ref{fig:benchmarking-extraction-algorithm} (A-C) we see snapshots of a time series of
+masks.
+This particular simulation includes division events between images (B) and (C).
+For all images (A-C), we have drawn the exact position of the agents with white lines and indicated
+the extracted positions with crosses.
+They align very well visually.
+It may be hard to spot with the bare eye, but the agreement between extracted and exact positions is
+visually worse at the endpoints of the rods.
+To quantify this disparity, we calculate the distance between these individual vertices.
+Figure~\ref{fig:benchmarking-extraction-algorithm} (D) shows the estimated length of the rods from
+the extracted and exact vertices which is surrounded by the average vertex distance.
+Division events lead to steps in our rod-length function since the bacterial length is halved.
+We can clearly see that the agreement is very good in the beginning but right after every division
+event, the uncertainty becomes large.
+To illustrate this, we can see that in figure~\ref{fig:benchmarking-extraction-algorithm} (C) some
+of the rods are partially overlapping which hides parts of the individual cell-masks and thus
+changes the extraction result.
+We also plotted the distribution of individual vertex distances in
+figure~\ref{fig:benchmarking-extraction-algorithm} (E).
+Darker areas in the plot correspond to distances at time-points earlier in the simulatin time.
+We can see that brighter areas tend to shift towards the right, thus increasing the uncertainty of
+the extraction.
+However, the overall distribution shows small values of distances for a large number of vertices and
+is not dominated by outliers.
+
+\begin{figure}
+    \centering
+    \begin{tikzonimage}[width=0.32\textwidth]
+        {docs/source/_static/fitting-methods/extract_positions-004000.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.32\textwidth]
+        {docs/source/_static/fitting-methods/extract_positions-006000.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.32\textwidth]
+        {docs/source/_static/fitting-methods/extract_positions-008000.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.5\textwidth]
+        {docs/source/_static/fitting-methods/displacement-calculations.png}%
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.5\textwidth]
+        {docs/source/_static/fitting-methods/displacement-distribution.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+    \end{tikzonimage}
+    \caption{
+        Progression of generated masks with estimated and exact positions.
+        Comparison of masks across generations. Compare estimated with exact rod lengths.
+    }
+    \label{fig:benchmarking-extraction-algorithm}
+\end{figure}
+
+%###################################################################################################
+\section{Visualization}
+\label{section:generating-masks-and-images}
+\begin{itemize}
+    \item use pyvista (VTK) for Visualization
+    \item always do 3D render. This way, we can have overalp between agents and produce more
+        realistic masks.
+    \item Assign unique color values for idividual agents.
+        We have a 1:1 mapping between agent identifier and color! (for each simulation different
+        though)
+    \item use parallel projection; this should be most close to what microscopes measure due to the
+        relatively large distance between object and lens.
+    \item use some noise to smear out the rods; this is already similar to microscopic images; use
+        more like this and deal with real-world defects to make it more realistic (future, not now)
+    \item Figure~\ref{fig:progression-image-generation} shows steps in generating image.
+    \item Reflectivity, placement of light etc. can already introduce some heterogeneity in the
+        visual of the rod.
+    \item Plan to implement noise on the mesh itself; make it have rough surface (like real rods as
+        seen under electron microscopy)
+\end{itemize}
+
+\paragraph{Algorithm}
+\begin{enumerate}
+    \item Get position of cell
+    \item Insert spheres at each vertex
+    \item Insert cylinder along edges
+    \item Combine individual objects into one big mesh
+    \item Use \ac{pbr}
+\end{enumerate}
+
+\begin{figure}[H]
+    \centering
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {docs/source/_static/09395645494836445480/raw_pv/000000400.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {docs/source/_static/09395645494836445480/images/000000400.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {docs/source/_static/09395645494836445480/masks/000000400.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+    \end{tikzonimage}
+    \caption{
+        (A) shows the raw result from combining sphere and cylinder meshes to obtain the shape of a
+        bacterial rod.
+        This render contains lighting coming from 2 light sources which thus produces a small glow
+        on the upper part of the meshes.
+        Subfigure (B) takes this image and applies noise to it in order to make it look more
+        like an actual microscopic image.
+        (C) The render is projected along the z-axis and flat numerical values are assigned to the
+        bacteria.
+        The resuling render can have overlaps in which case the color of the rod is displayed that
+        is higher along the z-direction.
+    }
+    \label{fig:progression-image-generation}
+\end{figure}
+
+%###################################################################################################
+\section{Parameter Estimation}
+In order to effectively estimate the parameters of our model, we need to set up a cost function
+which can be minimized.
+In doing so, we hope to identify all parameters, given relevant experimental data.
+We use microscopic images and construct a novel algorithm that can extract the position of the agent
+from a provided cell-mask.
+These cell-masks are easily obtainable by either using existing segmentation tools or manually
+annottating images.
+Following this, we can define a cost-function and an optimization scheme to minimize said function.
+
+% --------------------------------------------------------------------------------------------------
+\subsection{Constructing a Cost Function}
+
+As within any optimization scheme, it can be beneficial to construct custom cost functions
+which are tailored to the problem at hand.
+With the methods that we have already estalished, we need to compare a collection of agents which
+were numerically calcualted through our \ac{abm} modeling approach whith microscopic images.
+Utilizing the extraction algorithm, we can simply compare the numerically obtained positions with
+the extracted positions of the data.
+We choose a typical least-squares approach for the comparison of the individual vertices.
+However, this procedure is only valid for the case where cell-division events do not play a role.
+In the case where we either the numerical model or the experimental data contain more or less cells
+than the other result, we are not able to compare them with this approach.
+
+\paragraph{Comparison across Generations}
+\label{paragraph:comparison-across-generations}
+
+In order to compare a variable amount of cells, we can not utilize the positions of these cells
+directly.
+In this approach, we propose a method which is similar to already existing routines (TODO mention
+machine learning, some vision algorithms use this).
+From the numerical results of the \ac{abm}, we can again generate an image of cell-masks.
+We then compare this image together to the experimental cell-mask pixel by pixel with each other.
+Whenever two pixel colors are not matching, a penalty value is assigned.
+The sum of all those penalties is then the total cost.
+This algorithm is an implementation of a signed area measure.
+It is very important to take into account the parental relationship between the cells, checking if
+if either one of the cells is a daughter of the other cell.
+If a cell has just divided, the resulting two daughter cells will cover precisely the space that the
+preceding agent had taken up.
+Let's assume as an example that one result has more cells than the other
+(Figure~\ref{fig:mask-difference-metric} (A,B)).
+When we compare the pixels of both images, we can relate the given color of the pixel to a certain
+cell.
+Whenever the colors of two distinct cells are in a parental relationship, we reduce the penalty
+which would be assigned or set it to zero.
+
+\begin{figure}
+    \centering
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {docs/source/_static/fitting-methods/progressions-1.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {docs/source/_static/fitting-methods/progressions-2.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+    \end{tikzonimage}
+    \linebreak
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {docs/source/_static/fitting-methods/progressions-3.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {docs/source/_static/fitting-methods/progressions-4.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+    \end{tikzonimage}
+    \caption{
+        The figure shows two masks (A,B) which will be compared against each other.
+        As we can see, image (B) contains more cells which has resulted from a division process that
+        is not yet occurred in image (A).
+        In (C) we can see the "naive" comparison, simply identifying non-matching colors.
+        By assigning penalties to all highlighted pixels, we would not take into consideration the
+        parental relationship.
+        This approach is displayed in (D) where pixels of cells that are related have been colored
+        in gray.
+    }
+    \label{fig:mask-difference-metric}
+\end{figure}
+
+We can use this procedure to calculate the cost between successive iteration steps
+$t_n$ with $t_{n+1}$.
+We expect to obtain an exponential growth in the penalty.
+This can be explained by to the fact that the growth of the total area is exponential in time.
+Furthremore, the temporal area difference can be viewed as a derivative but the derivative of an
+exponential is again an exponential.
+Figure~\ref{fig:parameter-estimates-single-step} shows the total penalty for this approach.
+We can see that by accounting for the parental relationship, we get the overall desired effect.
+This method allows us to compare results between states that have a different number of cells
+inside as long as their origin stems from the a shared initial state.
+However, since this algorithm requires us to calculate a cell-mask for each comparison and then also
+process this mask as well as the comparator, it is much more computationally demanding.
+In cases where the simpler approach of comparing positions of the agents is reasonable, this method
+should be preferred.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]
+        {docs/source/_static/fitting-methods/penalty-time-flow.png}%
+    \caption{
+        Total cost between successive iteration steps $t_n,t_{n+1}$.
+        The overall shape is expected to have the slope of an exponential function.
+        We can clearly see that division events introduce jumps within this relationship.
+        In the case where we account for the parental relationship, these jumps are mitigated.
+        The penalty was set to $0$ for pixel-comparisons which belong to cells that are in a
+        parental relationship.
+    }
+    \label{fig:penalty-calculation-cell-division}
+\end{figure}
+
+\section{Applications}
+% --------------------------------------------------------------------------------------------------
+\subsection{A Mechanical Model of Rod-Shaped Bacteria}
+\label{subsubsection:mechanical-model-rod-shaped-bacteria}
+
+%###################################################################################################
 
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.5\textwidth]{docs/source/_static/mechanics.png}
-    \caption{
-        Example of an elonged bacterial rod represented by a collection of vertices which are
-        connected by springs.
-        The angles $\alpha_i$ have been exaggerated such that they can be identified more easily.
-        A bacterial agent which is not interacting with any external objects will tend to the
-        equilibrium state of a straight rod with $0$ curvature.
-    }
-    \label{fig:mechanics-bacterium}
-\end{figure}
-
 % --------------------------------------------------------------------------------------------------
 \subsection{Interactions}
+\label{subsection:mechanical-model-interactions}
 
@@
 \subsection{Cycle}
+\label{subsection:mechanical-model-cycle}
 
@@
 \subsection{External Forces}
+\label{subsection:mechanical-model-external-forces}
 TODO
@@
 
-%###################################################################################################
-\section{Generating Masks and Images}
-\label{section:generating-masks-and-images}
-\begin{itemize}
-    \item use pyvista (VTK) for Visualization
-    \item always do 3D render. This way, we can have overalp between agents and produce more
-        realistic masks.
-    \item Assign unique color values for idividual agents.
-        We have a 1:1 mapping between agent identifier and color! (for each simulation different
-        though)
-    \item use parallel projection; this should be most close to what microscopes measure due to the
-        relatively large distance between object and lens.
-    \item use some noise to smear out the rods; this is already similar to microscopic images; use
-        more like this and deal with real-world defects to make it more realistic (future, not now)
-    \item Figure~\ref{fig:progression-image-generation} shows steps in generating image.
-    \item Reflectivity, placement of light etc. can already introduce some heterogeneity in the
-        visual of the rod.
-    \item Plan to implement noise on the mesh itself; make it have rough surface (like real rods as
-        seen under electron microscopy)
-\end{itemize}
-
-\paragraph{Algorithm}
-\begin{enumerate}
-    \item Get position of cell
-    \item Insert spheres at each vertex
-    \item Insert cylinder along edges
-    \item Combine individual objects into one big mesh
-    \item Use \ac{pbr}
-\end{enumerate}
-
-\begin{figure}[H]
-    \centering
-    \begin{tikzonimage}[width=0.3\textwidth]
-        {docs/source/_static/09395645494836445480/raw_pv/000000400.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.3\textwidth]
-        {docs/source/_static/09395645494836445480/images/000000400.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.3\textwidth]
-        {docs/source/_static/09395645494836445480/masks/000000400.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
-    \end{tikzonimage}
-    \caption{
-        (A) shows the raw result from combining sphere and cylinder meshes to obtain the shape of a
-        bacterial rod.
-        This render contains lighting coming from 2 light sources which thus produces a small glow
-        on the upper part of the meshes.
-        Subfigure (B) takes this image and applies noise to it in order to make it look more
-        like an actual microscopic image.
-        (C) The render is projected along the z-axis and flat numerical values are assigned to the
-        bacteria.
-        The resuling render can have overlaps in which case the color of the rod is displayed that
-        is higher along the z-direction.
-    }
-    \label{fig:progression-image-generation}
-\end{figure}
-
-%###################################################################################################
-\section{Parameter Estimation}
-
-In order to effectively estimate the parameters of our model, we need to set up a cost function
-which can be minimized.
-In doing so, we hope to identify all parameters, given relevant experimental data.
-We use microscopic images and construct a novel algorithm that can extract the position of the agent
-from a provided cell-mask.
-These cell-masks are easily obtainable by either using existing segmentation tools or manually
-annottating images.
-Following this, we can define a cost-function and an optimization scheme to minimize said function.
-
 % --------------------------------------------------------------------------------------------------
-\subsection{Extract Positions from Masks}
-\label{section:fitting-and-comparing-masks}
-
-This section will explain how we can determine the collection of vertices $\vec{v}_i$ introduced in
-subsection~\ref{subsection:mechanical-model-mechanics} from microscopic images
-(Figure~\ref{fig:position-extraction-algorithm} (A)).
-They will serve as initial, final and intermediate values with which we can compare our predictions.
-The algorithms which we devised requires the input of cell-masks which assign a unique color for
-every cell in the image.
-Such masks can be generated by cell-segmentation tools or obtained by annottating microscopic images
-by hand (Figure~\ref{fig:position-extraction-algorithm} (B)).
-Afterwards, we proceed with the following 5 steps.
-\begin{enumerate}
-    \item Obtain submasks for each cell individually.
-    \item Skeletonize submask with algorithm from Lee~\cite{Lee1994}
-    \item Determine endpoints $\vec{q}_0,\vec{q}_1$.
-        Assert that there are only 2 endpoints.
-    \item Sort points such that they form a polygon.
-    \item Interpolate vertices $\vec{v}_i$ from obtained polygon.
-\end{enumerate}
-In the first step, the image will be split into a collection of subimages that only contain
-information for one single cell.
-From there, we apply a skeletonization algorithm~\cite{Lee1994}
-(Figure~\ref{fig:position-extraction-algorithm} (C)).
-It has to be noted that there are multiple variants of skeletonization and in general, these classes
-of algorithms can produce results which have intersection points and multiple endpoints.
-This described behaviour is undesirable in our case.
-We have found that for most practical examples, the algorithm devised by~\cite{Lee1994} has worked
-well.\\
-The skeletonization algorithm supplies us with a collection of unordered lattice points $\vec{p}_i$
-(pixel coordinates) where the skeleton lives.
-We identify the endpoints $\vec{q}_0,\vec{q}_1$ by counting the number of neighbors in the von
-Neumann neighborhood for each lattice point.
-Any point with exactly one neighboring point must be and endpoint.
-If we encounter more or less than two points which fit this criterion, the routine can not commence
-and is aborted.
-Next, we start with one of the endpoints and determine the next neighbor which was not yet picked.
-We list the points in this order, thus forming a sorted polygon of pixel coordinates.\\
-In the last step, we take the sorted, extracted points $\vec{p}_i$ and interpolate $n$ values for
-the desired number of vertices $\vec{v}_i$.
-
-\begin{figure}
-    \centering
-    \begin{tikzonimage}[width=0.32\textwidth]
-        {docs/source/_static/fitting-methods/algorithm/image001042.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.32\textwidth]
-        {docs/source/_static/fitting-methods/algorithm/mask-zoom.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.32\textwidth]
-        {docs/source/_static/fitting-methods/algorithm/mask-zoom.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
-    \end{tikzonimage}
-    \caption{
-        TODO change 2nd image!\\
-        (A) An example of a microscopic image.
-        (B) The associated mask which assigns individual color values to each bacterium.
-        (C) The result of the  skeletonization algorithm when applying it to a single cell.
-    }
-    \label{fig:position-extraction-algorithm}
-\end{figure}
-
-% --------------------------------------------------------------------------------------------------
-\subsection{Benchmarking the Extraction Algorithm}
-\label{subsection:benchmarking-extraction-algorithm}
-
-In order to determine if our extraction algorithm provides reliable results, we have to take
-advantage of a functionality which will be introduced later in
-section~\ref{section:generating-masks-and-images}.
-Our methodology makes use of the fact that we have already set up our model and solving scheme and
-are able to produce cell masks from these numerical results.
-In doing so, we have exact knowledge about the position of our agents since we can simply obtain
-this information from our model.
-But on the other hand we can also try to extract this position from the generated cell-masks.
-By comparing the differences between them, we get a good estimate of how well our extraction
-algorithm can determine the true position of the bacteria.
-
-In figure~\ref{fig:benchmarking-extraction-algorithm} (A-C) we see snapshots of a time series of
-masks.
-This particular simulation includes division events between images (B) and (C).
-For all images (A-C), we have drawn the exact position of the agents with white lines and indicated
-the extracted positions with crosses.
-They align very well visually.
-It may be hard to spot with the bare eye, but the agreement between extracted and exact positions is
-visually worse at the endpoints of the rods.
-To quantify this disparity, we calculate the distance between these individual vertices.
-Figure~\ref{fig:benchmarking-extraction-algorithm} (D) shows the estimated length of the rods from
-the extracted and exact vertices which is surrounded by the average vertex distance.
-Division events lead to steps in our rod-length function since the bacterial length is halved.
-We can clearly see that the agreement is very good in the beginning but right after every division
-event, the uncertainty becomes large.
-To illustrate this, we can see that in figure~\ref{fig:benchmarking-extraction-algorithm} (C) some
-of the rods are partially overlapping which hides parts of the individual cell-masks and thus
-changes the extraction result.
-We also plotted the distribution of individual vertex distances in
-figure~\ref{fig:benchmarking-extraction-algorithm} (E).
-Darker areas in the plot correspond to distances at time-points earlier in the simulatin time.
-We can see that brighter areas tend to shift towards the right, thus increasing the uncertainty of
-the extraction.
-However, the overall distribution shows small values of distances for a large number of vertices and
-is not dominated by outliers.
-
-\begin{figure}
-    \centering
-    \begin{tikzonimage}[width=0.32\textwidth]
-        {docs/source/_static/fitting-methods/extract_positions-004000.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.32\textwidth]
-        {docs/source/_static/fitting-methods/extract_positions-006000.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.32\textwidth]
-        {docs/source/_static/fitting-methods/extract_positions-008000.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.5\textwidth]
-        {docs/source/_static/fitting-methods/displacement-calculations.png}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
-    \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.5\textwidth]
-        {docs/source/_static/fitting-methods/displacement-distribution.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
-    \end{tikzonimage}
-    \caption{
-        Progression of generated masks with estimated and exact positions.
-        Comparison of masks across generations. Compare estimated with exact rod lengths.
-    }
-    \label{fig:benchmarking-extraction-algorithm}
-\end{figure}
-
-% --------------------------------------------------------------------------------------------------
-\subsection{Constructing a Cost Function}
-
-As within any optimization scheme, it can be beneficial to construct custom cost functions
-which are tailored to the problem at hand.
-With the methods that we have already estalished, we need to compare a collection of agents which
-were numerically calcualted through our \ac{abm} modeling approach whith microscopic images.
-Utilizing the extraction algorithm, we can simply compare the numerically obtained positions with
-the extracted positions of the data.
-We choose a typical least-squares approach for the comparison of the individual vertices.
-However, this procedure is only valid for the case where cell-division events do not play a role.
-In the case where we either the numerical model or the experimental data contain more or less cells
-than the other result, we are not able to compare them with this approach.
-
-\paragraph{Comparison across Generations}
-\label{paragraph:comparison-across-generations}
-
-In order to compare a variable amount of cells, we can not utilize the positions of these cells
-directly.
-In this approach, we propose a method which is similar to already existing routines (TODO mention
-machine learning, some vision algorithms use this).
-From the numerical results of the \ac{abm}, we can again generate an image of cell-masks.
-We then compare this image together to the experimental cell-mask pixel by pixel with each other.
-Whenever two pixel colors are not matching, a penalty value is assigned.
-The sum of all those penalties is then the total cost.
-This algorithm is an implementation of a signed area measure.
-It is very important to take into account the parental relationship between the cells, checking if
-if either one of the cells is a daughter of the other cell.
-If a cell has just divided, the resulting two daughter cells will cover precisely the space that the
-preceding agent had taken up.
-Let's assume as an example that one result has more cells than the other
-(Figure~\ref{fig:mask-difference-metric} (A,B)).
-When we compare the pixels of both images, we can relate the given color of the pixel to a certain
-cell.
-Whenever the colors of two distinct cells are in a parental relationship, we reduce the penalty
-which would be assigned or set it to zero.
-
-\begin{figure}
-    \centering
-    \begin{tikzonimage}[width=0.3\textwidth]
-        {docs/source/_static/fitting-methods/progressions-1.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.3\textwidth]
-        {docs/source/_static/fitting-methods/progressions-2.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
-    \end{tikzonimage}
-    \linebreak
-    \begin{tikzonimage}[width=0.3\textwidth]
-        {docs/source/_static/fitting-methods/progressions-3.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.3\textwidth]
-        {docs/source/_static/fitting-methods/progressions-4.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{D}};
-    \end{tikzonimage}
-    \caption{
-        The figure shows two masks (A,B) which will be compared against each other.
-        As we can see, image (B) contains more cells which has resulted from a division process that
-        is not yet occurred in image (A).
-        In (C) we can see the "naive" comparison, simply identifying non-matching colors.
-        By assigning penalties to all highlighted pixels, we would not take into consideration the
-        parental relationship.
-        This approach is displayed in (D) where pixels of cells that are related have been colored
-        in gray.
-    }
-    \label{fig:mask-difference-metric}
-\end{figure}
-
-We can use this procedure to calculate the cost between successive iteration steps
-$t_n$ with $t_{n+1}$.
-We expect to obtain an exponential growth in the penalty.
-This can be explained by to the fact that the growth of the total area is exponential in time.
-Furthremore, the temporal area difference can be viewed as a derivative but the derivative of an
-exponential is again an exponential.
-Figure~\ref{fig:parameter-estimates-single-step} shows the total penalty for this approach.
-We can see that by accounting for the parental relationship, we get the overall desired effect.
-This method allows us to compare results between states that have a different number of cells
-inside as long as their origin stems from the a shared initial state.
-However, since this algorithm requires us to calculate a cell-mask for each comparison and then also
-process this mask as well as the comparator, it is much more computationally demanding.
-In cases where the simpler approach of comparing positions of the agents is reasonable, this method
-should be preferred.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.5\textwidth]
-        {docs/source/_static/fitting-methods/penalty-time-flow.png}%
-    \caption{
-        Total cost between successive iteration steps $t_n,t_{n+1}$.
-        The overall shape is expected to have the slope of an exponential function.
-        We can clearly see that division events introduce jumps within this relationship.
-        In the case where we account for the parental relationship, these jumps are mitigated.
-        The penalty was set to $0$ for pixel-comparisons which belong to cells that are in a
-        parental relationship.
-    }
-    \label{fig:penalty-calculation-cell-division}
-\end{figure}
-
-% --------------------------------------------------------------------------------------------------
-\subsection{Estimating Parameters}
-\label{section:parameter-estimation}
-
+\subsection{Estimating Mechanical Parameters of Growing \textit{E.Coli.} on Agar}
 We will now focus on using the presented methods for parameter estimation via microscopic images.
@@
 
-%###################################################################################################
-\section{Multilayer Predictions}
-% TODO
-\begin{itemize}
-    \item Do we need this in this publication?
-    \item Maybe this is out of scope as this is a more methodological paper.
-    \item If we want to omit it, we might also opt to change the model description in the beginning
-        and remove some parts of the external forces discussion.
-\end{itemize}
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.49\textwidth]
-        {docs/source/_static/scripts/crm_multilayer/multilayer-time-evolution.pdf}\\
-    \includegraphics[width=0.49\textwidth]
-        {docs/source/_static/scripts/crm_multilayer/colony-height-vs-gel_pressure.pdf}
-    \includegraphics[width=0.49\textwidth]
-        {docs/source/_static/scripts/crm_multilayer/colony-height-vs-time.pdf}
-    \caption{TODO}
-    \label{fig:multilayer}
-\end{figure}
+% --------------------------------------------------------------------------------------------------
+\subsection{Data Generation}
 
@@
 
+%###################################################################################################
+\section{Multilayer Predictions}
+% TODO
+\begin{itemize}
+    \item Do we need this in this publication?
+    \item Maybe this is out of scope as this is a more methodological paper.
+    \item If we want to omit it, we might also opt to change the model description in the beginning
+        and remove some parts of the external forces discussion.
+\end{itemize}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.49\textwidth]
+        {docs/source/_static/scripts/crm_multilayer/multilayer-time-evolution.pdf}\\
+    \includegraphics[width=0.49\textwidth]
+        {docs/source/_static/scripts/crm_multilayer/colony-height-vs-gel_pressure.pdf}
+    \includegraphics[width=0.49\textwidth]
+        {docs/source/_static/scripts/crm_multilayer/colony-height-vs-time.pdf}
+    \caption{TODO}
+    \label{fig:multilayer}
+\end{figure}
+
 \end{document}

### 2025-07-11 plots/fitting_extract_positions.py
update visual of plots
- use new simulation methods
- new plotting configuration
- colors from cr_mech_coli

@@ -7,16 +7,7 @@ import matplotlib as mpl
 from tqdm import tqdm
 import argparse
 import time
-
-mpl.use("pgf")
-plt.rcParams.update(
-    {
-        "font.family": "serif",  # use serif/main font for text elements
-        "text.usetex": True,  # use inline math for ticks
-        "pgf.rcfonts": False,  # don't setup fonts from rc parameters
-        "pgf.preamble": "\\usepackage{siunitx}",  # load additional packages
-    }
-)
+import scipy as sp
 
 
 def calculate_lengths_distances(
@@ -76,7 +67,11 @@ def create_simulation_result(n_vertices: int, rng_seed: int = 3):
         domain_size=np.array([150, 150]),
     )
     config.storage_options = [crm.StorageOption.Memory]
-    agent_settings = crm.AgentSettings(growth_rate=0.05, growth_rate_distr=(0.05, 0.01))
+    config.show_progressbar = True
+    agent_settings = crm.AgentSettings(
+        growth_rate=0.012,
+        growth_rate_distr=(0.012, 0.000),
+    )
     agent_settings.mechanics.rigidity = 8.0
     config.domain_height = 0.2
 
@@ -188,6 +183,7 @@ if __name__ == "__main__":
     ]
 
     if not pyargs.skip_graph or not pyargs.skip_distribution:
+        crm.plotting.set_mpl_rc_params()
         if pyargs.workers < 0:
             import multiprocessing as mp
 
@@ -208,54 +204,86 @@ if __name__ == "__main__":
         exit()
 
     if not pyargs.skip_distribution:
-        fig, ax = plt.subplots()
+        fig, ax = plt.subplots(figsize=(8, 8))
+        crm.plotting.configure_ax(ax)
 
-        t = 0.7
         q = len(distances_vertices)
-        colors = [(t * i / q, t * i / q, t * i / q) for i in range(q)]
-        ax.set_title("Distribution of distances between individual vertices")
+        c1 = np.array(mpl.colors.to_rgba(crm.plotting.COLOR3))
+        c2 = np.array(mpl.colors.to_rgba(crm.plotting.COLOR1))
+        colors = [c2 * i / q + (1 - i / q) * c1 for i in range(q)]
+        # ax.set_title("Distribution of distances between individual vertices")
         n_bins = 50
         logbins = np.logspace(
             np.log10(np.min([np.min(d) for d in distances_vertices])),
             np.log10(np.max([np.max(d) for d in distances_vertices])),
             n_bins + 1,
         )
[…]

### 2025-07-11 plots/fitting_penalty_over_time.py

@@ -10,16 +10,6 @@ import argparse
 from fitting_extract_positions import create_simulation_result
 
 
-plt.rcParams.update(
-    {
-        "font.family": "serif",  # use serif/main font for text elements
-        "text.usetex": True,  # use inline math for ticks
-        "pgf.rcfonts": False,  # don't setup fonts from rc parameters
-        "pgf.preamble": "\\usepackage{siunitx}",  # load additional packages
-    }
-)
-
-
 def render_single_mask(n_iter: int, cell_container, domain_size, render_settings):
     cell_container = crm.CellContainer.deserialize(cell_container)
     cells_at_iter = cell_container.get_cells_at_iteration(n_iter)
@@ -43,14 +33,15 @@ if __name__ == "__main__":
     interval = time.time()
     pool = mp.Pool()
 
-    rs = crm.RenderSettings(resolution=800)
+    rs = crm.RenderSettings()
     args = [(i, cell_container.serialize(), config.domain_size, rs) for i in iterations]
     masks = pool.starmap(render_single_mask, args)
     print(f"{time.time() - interval:8.4} Calculated Masks:")
     interval = time.time()
 
+    save_interval = config.t_max / (config.n_saves + 1)
     penalties_area_diff = [
-        crm.penalty_area_diff(masks[i - 1], masks[i]) / config.save_interval
+        crm.penalty_area_diff(masks[i - 1], masks[i]) / save_interval
         for i in range(1, len(iterations))
     ]
     print(f"{time.time() - interval:8.4} Calculated Penalties without parents:")
@@ -58,14 +49,14 @@ if __name__ == "__main__":
 
     penalties_parents = [
         crm.penalty_area_diff_account_parents(masks[i - 1], masks[i], cell_container, 0)
-        / config.save_interval
+        / save_interval
         for i in range(1, len(iterations))
     ]
     print(f"{time.time() - interval:8.4} Calculated Penalties with parents:")
     interval = time.time()
 
     n_cells = [len(cell_container.get_cells_at_iteration(i)) for i in iterations]
-    x = np.array([i * config.save_interval for i in range(len(iterations))])
+    x = np.array([i * save_interval for i in range(len(iterations))])
 
     # Fit exponential function to penalties with parents
[…]

### 2025-07-14 plots/fitting_extract_positions.py
use randomized initial growth rates

@@ -64,13 +64,13 @@ def create_simulation_result(n_vertices: int, rng_seed: int = 3):
         dt=0.02,
         t_max=200.0,
         n_saves=49,
-        domain_size=np.array([150, 150]),
+        domain_size=np.array([200, 200]),
     )
     config.storage_options = [crm.StorageOption.Memory]
     config.show_progressbar = True
     agent_settings = crm.AgentSettings(
         growth_rate=0.012,
-        growth_rate_distr=(0.012, 0.000),
+        growth_rate_distr=(0.012, 0.002),
     )
     agent_settings.mechanics.rigidity = 8.0
     config.domain_height = 0.2
@@ -86,6 +86,9 @@ def create_simulation_result(n_vertices: int, rng_seed: int = 3):
     )
     rod_args = agent_settings.to_rod_agent_dict()
     agents = [crm.RodAgent(pos=p, vel=p * 0.0, **rod_args) for p in positions]
+    rng = np.random.default_rng(rng_seed)
+    for a in agents:
+        a.growth_rate += 0.002 * rng.random(1)
     res = crm.run_simulation_with_agents(config, agents)
     print(f"{time.time() - interval:8.4} Created Simulation Result:")
     return config, res

### 2025-07-14 plots/fitting_extract_positions.py
remove fitting of curve for distribution

@@ -7,7 +7,6 @@ import matplotlib as mpl
 from tqdm import tqdm
 import argparse
 import time
-import scipy as sp
 
 
 def calculate_lengths_distances(
@@ -228,27 +227,6 @@ if __name__ == "__main__":
             color=colors,
             label="Vertex Diff",
         )
-
-        # Fit exponential decay
-        # def fit_func(t, a, b, t0):
-        #     z1 = a * np.exp(-b * (t - t0) ** 2)
-        #     z2 = a * np.exp(-b * (t + t0) ** 2)
-        #     return z1 + z2
-        #
-        # y_fit = np.max(values, axis=0)
-        # x_fit = 0.5 * (np.array(bins)[1:] + np.array(bins)[:-1])
-        # popt, pcov = sp.optimize.curve_fit(
-        #     fit_func,
-        #     x_fit,
-        #     y_fit,
-        #     p0=(np.max(y_fit), 100, np.mean(x_fit)),
-        #     bounds=(0, np.inf),
-        # )
-        # ax.plot(
-        #     x_fit, fit_func(x_fit, *popt), c=crm.plotting.COLOR5, label="Log-normal Fit"
-        # )
-        # ymax = np.max(values)
-        # ax.set_ylim(1, 2 * ymax)
 
         ax.set_xscale("log")
         ax.set_yscale("log")

### 2025-07-14 plots/fitting_extract_positions.py
use tqdm to show progrssion

@@ -179,10 +179,13 @@ if __name__ == "__main__":
             )
 
     ccs = cell_container.serialize()
-    arglist = [
-        (n, ccs, config.domain_size, pyargs.skel_method, pyargs.n_vertices)
-        for n in iterations
-    ]
+    arglist = tqdm(
+        [
+            (n, ccs, config.domain_size, pyargs.skel_method, pyargs.n_vertices)
+            for n in iterations
+        ],
+        total=len(iterations),
+    )
 
     if not pyargs.skip_graph or not pyargs.skip_distribution:
         crm.plotting.set_mpl_rc_params()

### 2025-07-14 plots/fitting_extract_positions.py
change styles of plots to fit to rest of paper

@@ -253,19 +253,19 @@ if __name__ == "__main__":
         x = np.arange(len(distances)) * config.t_max / (config.n_saves + 1)
         ax.plot(
             x,
-            [np.mean(li) for li in lengths_extracted],
-            # yerr=[np.std(li) for li in lengths1],
-            linestyle="--",
-            color=crm.plotting.COLOR3,
-            label="Extracted",
+            [np.mean(li) for li in lengths_exact],
+            # yerr=[np.std(li) for li in lengths2],
+            linestyle="-",
+            color=crm.plotting.COLOR5,
+            label="Exact",
         )
         ax.plot(
             x,
-            [np.mean(li) for li in lengths_exact],
-            # yerr=[np.std(li) for li in lengths2],
+            [np.mean(li) for li in lengths_extracted],
+            # yerr=[np.std(li) for li in lengths1],
             linestyle=":",
-            color=crm.plotting.COLOR5,
-            label="Exact",
+            color=crm.plotting.COLOR3,
+            label="Extracted",
         )
         ax.fill_between(
             x,
@@ -278,8 +278,7 @@ if __name__ == "__main__":
                 for i in range(len(lengths_extracted))
             ],
             alpha=0.3,
-            color=crm.plotting.COLOR5,
-            label="Vertex Diff",
+            color=crm.plotting.COLOR1,
         )
         ax.legend(
             loc="upper center",

### 2025-07-14 plots/fitting_penalty_over_time.py

@@ -76,35 +76,39 @@ if __name__ == "__main__":
         x[1:],
         penalties_area_diff,
         label="Area Difference",
-        linestyle=(0, (6, 4)),
-        color="gray",
+        color=crm.plotting.COLOR3,
     )
     ax1.plot(
         x[1:],
         penalties_parents,
-        label="Account for Parents",
-        linestyle="-",
-        color="k",
+        label="with Parents",
+        linestyle="--",
+        color=crm.plotting.COLOR2,
+    )
+    ax1.plot(
+        x[1:],
+        exponential(x[1:], *popt),
+        color=crm.plotting.COLOR1,
+        label="ER",
     )
     ax1.fill_between(
         x[1:],
         exponential(x[1:], *[popt[i] - pcov[i][i] ** 0.5 for i in range(len(popt))]),
         exponential(x[1:], *[popt[i] + pcov[i][i] ** 0.5 for i in range(len(popt))]),
-        label="Fit $A e^{{\\lambda t}}$",
-        color=(0.85, 0.85, 0.85),
+        color=crm.plotting.COLOR1,
+        alpha=0.6,
     )
     ax1.set_xlabel("Time [min]")
     ax1.set_ylabel("Penalty [1/min]")
     ax2 = ax1.twinx()
-    ax2.plot(x, n_cells, label="Number of Cells", linestyle=(0, (1, 1)), color="k")
+    ax2.plot(x, n_cells, label="Cells", linestyle=(0, (1, 1)), color="k")
     ax2.set_ylabel("Number of Cells")
     ax2.set_ylim(1, 100)
     ax1.set_yscale("log")
     ax2.set_yscale("log")
-    fig.tight_layout()
 
     handles1, labels1 = ax1.get_legend_handles_labels()
-    handles2, labels2 = ax1.get_legend_handles_labels()
+    handles2, labels2 = ax2.get_legend_handles_labels()
     handles = handles1 + handles2
     labels = labels1 + labels2
     ax1.legend(
@@ -112,7 +116,7 @@ if __name__ == "__main__":
         labels,
         loc="upper center",
         bbox_to_anchor=(0.5, 1.18),
-        ncol=4,
+        ncol=2,
         frameon=False,
     )
 

### 2025-07-14 plots/fitting_progression.py
remove resolution argument

@@ -13,7 +13,7 @@ if __name__ == "__main__":
     i1 = iterations[25]
     i2 = iterations[35]
 
-    rs = crm.RenderSettings(resolution=800)
+    rs = crm.RenderSettings()
     mask1 = crm.render_mask(all_cells[i1], colors, config.domain_size, rs)
     mask2 = crm.render_mask(all_cells[i2], colors, config.domain_size, rs)
     mask3 = crm.area_diff_mask(mask1, mask2)

### 2025-07-14 tests/test_counter_color_conversion.py
fix test
- use new functions
- start counting with minimum counter instead of 0

@@ -19,10 +19,13 @@ def test_color_counter_conversion():
                 assert color == color_new
 
 
-def est_assign_colors():
+def test_assign_colors():
     config = crm.Configuration()
     agent_settings = crm.AgentSettings()
-    sim_result = crm.run_simulation(config, agent_settings)
+    positions = crm.generate_positions(8, agent_settings, config)
+    rod_args = agent_settings.to_rod_agent_dict()
+    agents = [crm.RodAgent(pos=p, vel=p * 0.0, **rod_args) for p in positions]
+    sim_result = crm.run_simulation_with_agents(config, agents)
 
     cell_to_color = sim_result.cell_to_color
     iterations = sim_result.get_all_iterations()
@@ -39,9 +42,9 @@ def est_assign_colors():
                 color = (color[0], color[1], color[2])
                 if color not in all_colors:
                     all_colors.add(color)
-                    counter = crm.color_to_counter(list(color))
+                    counter = crm.color_to_counter(color)
                     all_counters.add(counter)
-    for i, cell in enumerate(cells):
-        color_expected = tuple(crm.counter_to_color(i + 1))
+    for i, cell in enumerate(cells, np.min(np.array([a for a in all_counters]))):
+        color_expected = tuple(crm.counter_to_color(i))
         assert color_expected in all_colors
         cell_to_color[cell]

### 2025-07-14 paper/main.tex
position labels to avoid interference with plots
include hint for discussion around loss functions of semantic segmentation
rename; relabel
add tableofcontents to create overivew for iterations
write section on fundamental assumptions

@@
 % \keywords{}
+
+\tableofcontents
 
@@
 %###################################################################################################
-\section{Modeling Prerequesites}
+\section{Fundamental Assumptions on the Spatial Representation of Rod-Shaped Bacteria}
+% TODO CITATIONS
+This work makes use of a few but powerfull assumptions concerning the mathematical representation of
+the cells in question.
+In order to describe the spatial dynamics of individual agents, we need to define their position,
+shape and how they interact with each other.
+In order to provide as much flexibility as possible when constructing new models, we restrict
+ourselves to assumptions concerning the position such that researchers can freely experiment with
+various shapes and interaction types.
+Furthermore, it is very important to strike a good balance between the restrictiveness of our
+assumptions and generality in order to provide meaningful results and still be able to apply our
+approach to a broader class of models.
+Taking into account the natural shape of rod-like bacteria, it is clear that they can not be
+represented with a singular point $\vec{v}$ in space.
+We could extend this positional definition with a vector $\vec{w}$, indicating the direction of
+extension but this would restrict ourselves to fully straight rod-shaped bacteria.
+However, multiple experiments have shown that MReB is responsible for the growth and maintenance of
+the elongated shape of bacteria~\cite{Erickson2001}.
+By means of this protein, rod-shaped bacteria can bend elastically on short time-scales and
+deform plastically over longer periods of time.
+Our goal is to provide a model which can encapsulate these effects.
+Thus, we extend our model from a single point to multiple vertices which are part of a 1-dimensional
+polygon with two endpoints $\vec{v}_i$.
+In essence, we discretize the (possibly non-straight) rod of the agent.
+Figure~\ref{fig:mechanics-bacterium} shows a visualization of this approach.
+The angles $\alpha_i$ between vertices $\vec{v}_{i-1},\vec{v}_i$ can be used when modeling the
+dynamics of these bacteria.
+
+With these fundamental thoughts layed out, we retain the flexibility to represent many models such
+as varying shapes since we do not have to make any assumptions on the type of interactions between
+two bacteria.
+Even concerning the shape of bacteria, this assumption still allows us to mdoel elliptical cells by
+modifying the interaction type as well as other effects such as polarity of
+interactions~\cite{Duvernoy2018}.
+Furthermore, we also did not have to assume any particular mechanical dynamics of the agents, only
+their spatial representation, thus leaving in particular room for plastical or elastic deformations
+of the rod and other more intricate mechanisms.
+
 \begin{enumerate}
@@
         {docs/source/_static/fitting-methods/displacement-calculations.png}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/displacement-distribution.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
     \end{tikzonimage}
@@
 % --------------------------------------------------------------------------------------------------
-\subsection{Constructing a Cost Function}
+\subsection{Constructing a Loss Function}
 
@@
 than the other result, we are not able to compare them with this approach.
+
+\paragraph{Comparison with Semantic Segmentation}
+\label{paragraph:comparison-semantic-segmentation}
+\cite{Jadon2020}
 
@@
     \caption{Parameter estimation of growth rates for individual cells}%TODO extend caption}
-    \label{fig:parameter-estimates-supplement}
+    \label{fig:parameter-estimates-supplement-growth-rates}
 \end{figure}
@@
     \caption{Parameter estimation of radii for individual cells}%TODO extend caption}
-    \label{fig:parameter-estimates-supplement}
+    \label{fig:parameter-estimates-supplement-radii}
 \end{figure}

### 2025-07-14 plots/position_extraction_algorithm.py
use correct file; plot positions correctly

@@ -10,7 +10,7 @@ def get_skeleton(submask, color):
 
 
 if __name__ == "__main__":
-    mask = np.loadtxt("data/crm_fit/0001/masks/image001042-markers.csv", delimiter=",")
+    mask = np.loadtxt("data/crm_fit/0001/masks/001042-markers.csv", delimiter=",")
 
     # Create new colors for mask
     new_colors = {int(i): crm.counter_to_color(int(i)) for i in np.unique(mask)}
@@ -98,7 +98,11 @@ if __name__ == "__main__":
     ax.set_axis_off()
     ax.imshow(submask)
     for p in pos:
-        ax.plot(p[:, 0] - xmin, p[:, 1] - ymin, color="white")
+        p1 = p[:, 0] - xmin - dx / 2
+        p2 = p[:, 1] - ymin - dy / 2
+        q1 = -(-p1 + p2) * (-1) + p1
+        q2 = -(-p1 + p2) * (+1) + p2
+        ax.plot(q1 + dx / 2, q2 + dy / 2, color="white", marker="+", markersize=15)
     fig.savefig(
         "docs/source/_static/fitting-methods/algorithm/interpolate-positions.png",
         bbox_inches="tight",

### 2025-07-14 paper/main.tex
restructure; try to unify paragraphs and subsections
add and rename table of contnents
more restructuring

@@
 
+\pagebreak
+\renewcommand{\contentsname}{Temporary! Table of Contents}
 \tableofcontents
+\pagebreak
 
@@
 %###################################################################################################
-\section{Fundamental Assumptions on the Spatial Representation of Rod-Shaped Bacteria}
+\section{Method Description}
+\subsection{Fundamental Assumptions on the Spatial Representation of Rod-Shaped Bacteria}
 % TODO CITATIONS
@@
 %###################################################################################################
-\section{Data Extraction}
+\subsection{Data Extraction}
 \label{section:fitting-and-comparing-masks}
@@
 \paragraph{Benchmarking the Extraction Algorithm}
-\label{subsection:benchmarking-extraction-algorithm}
+\label{paragraph-extraction-algorithm}
 
@@
 %###################################################################################################
-\section{Visualization}
+\subsection{Visualization}
 \label{section:generating-masks-and-images}
@@
 %###################################################################################################
-\section{Parameter Estimation}
+\subsection{Parameter Estimation}
 In order to effectively estimate the parameters of our model, we need to set up a cost function
@@
 % --------------------------------------------------------------------------------------------------
-\subsection{Constructing a Loss Function}
+\subsubsection{Challenges of Constructing a Loss Function}
 
@@
 
-\paragraph{Comparison with Semantic Segmentation}
+% --------------------------------------------------------------------------------------------------
+\subsubsection{Relationship to Semantic Segmentation}
 \label{paragraph:comparison-semantic-segmentation}
@@
 
-\paragraph{Comparison across Generations}
+% --------------------------------------------------------------------------------------------------
+\subsubsection{Comparison across Generations}
 \label{paragraph:comparison-across-generations}
@@
 
-\section{Applications}
 % --------------------------------------------------------------------------------------------------
-\subsection{A Mechanical Model of Rod-Shaped Bacteria}
+\section{Application to Mechanics of Rod-Shaped Bacteria}
 \label{subsubsection:mechanical-model-rod-shaped-bacteria}
@@
 % --------------------------------------------------------------------------------------------------
-\subsection{Interactions}
+\subsection{Physical Interactions}
 \label{subsection:mechanical-model-interactions}
@@
 % --------------------------------------------------------------------------------------------------
-\subsection{Cycle}
+\subsection{Cell-Cycle}
 \label{subsection:mechanical-model-cycle}

### 2025-07-14 paper/main.tex
use new plot in paper

@@
     \centering
-    \begin{tikzonimage}[width=0.32\textwidth]
+    \begin{tikzonimage}[width=0.36\textwidth]
         {docs/source/_static/fitting-methods/algorithm/image001042.png}
@@
     \end{tikzonimage}
-    \begin{tikzonimage}[width=0.32\textwidth]
+    \begin{tikzonimage}[width=0.36\textwidth]
         {docs/source/_static/fitting-methods/algorithm/mask-zoom.png}
@@
     \end{tikzonimage}
-    \begin{tikzonimage}[width=0.32\textwidth]
-        {docs/source/_static/fitting-methods/algorithm/mask-zoom.png}
+    \begin{tikzonimage}[width=0.255\textwidth]
+        {docs/source/_static/fitting-methods/algorithm/interpolate-positions.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
@@
         (A) An example of a microscopic image.
-        (B) The associated mask which assigns individual color values to each bacterium.
-        (C) The result of the  skeletonization algorithm when applying it to a single cell.
+        (B) The result of the  skeletonization algorithm when applying it to a single cell.
+        (C) Extrapolated vertices $\vec{v}_i$ which represent the position of our cell-agents.
     }

### 2025-07-15 paper/main.tex
add TODOs
format figure; use newly generated image

@@
             important.
+    \end{itemize}
+    \begin{itemize}
+        \item State of the Art
     \end{itemize}
@@
         Comparison of masks across generations. Compare estimated with exact rod lengths.
+        "Iceberg plot"
     }
@@
 
+\begin{itemize}
+    \item Semantic Segmentation is more general
+    \item Loss functions and some techniques can be reused
+    \item Discuss which types of loss functions we could consider
+    \item Possibly benchmark them later when doing the parameter estimation; compare loss functions
+\end{itemize}
+
 % --------------------------------------------------------------------------------------------------
@@
 
-\begin{figure}
-    \centering
-    \begin{tikzonimage}[width=0.3\textwidth]
+\begin{figure}[H]
+    \centering
+    \begin{minipage}{0.5\textwidth}
+    \begin{tikzonimage}[width=0.48\textwidth]
         {docs/source/_static/fitting-methods/progressions-1.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.3\textwidth]
+    \end{tikzonimage}%
+    \hspace{0.01\textwidth}%
+    \begin{tikzonimage}[width=0.48\textwidth]
         {docs/source/_static/fitting-methods/progressions-2.png}
@@
     \linebreak
-    \begin{tikzonimage}[width=0.3\textwidth]
+    \begin{tikzonimage}[width=0.48\textwidth]
         {docs/source/_static/fitting-methods/progressions-3.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.3\textwidth]
+    \end{tikzonimage}%
+    \hspace{0.01\textwidth}%
+    \begin{tikzonimage}[width=0.48\textwidth]
         {docs/source/_static/fitting-methods/progressions-4.png}
@@
     \end{tikzonimage}
+    \end{minipage}%
+    \begin{minipage}{0.49\textwidth}
+        \begin{tikzonimage}[width=\textwidth]
+            {docs/source/_static/fitting-methods/penalty-time-flow.png}%
+            \node at (0, 0.975)[anchor=north west, draw, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+        \end{tikzonimage}
+    \end{minipage}
     \caption{

### 2025-07-15 paper/main.tex
combine figures

@@
         This approach is displayed in (D) where pixels of cells that are related have been colored
-        in gray.
+        in gray.\\
+        Total cost between successive iteration steps $t_n,t_{n+1}$.
+        The overall shape is expected to have the slope of an exponential function.
+        We can clearly see that division events introduce jumps within this relationship.
+        In the case where we account for the parental relationship, these jumps are mitigated.
+        The penalty was set to $0$ for pixel-comparisons which belong to cells that are in a
+        parental relationship.
     }
@@
 exponential is again an exponential.
-Figure~\ref{fig:parameter-estimates-single-step} shows the total penalty for this approach.
+Figure~\ref{fig:mask-difference-metric} shows the total penalty for this approach.
 We can see that by accounting for the parental relationship, we get the overall desired effect.
@@
 should be preferred.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.5\textwidth]
-        {docs/source/_static/fitting-methods/penalty-time-flow.png}%
-    \caption{
-        Total cost between successive iteration steps $t_n,t_{n+1}$.
-        The overall shape is expected to have the slope of an exponential function.
-        We can clearly see that division events introduce jumps within this relationship.
-        In the case where we account for the parental relationship, these jumps are mitigated.
-        The penalty was set to $0$ for pixel-comparisons which belong to cells that are in a
-        parental relationship.
-    }
-    \label{fig:penalty-calculation-cell-division}
-\end{figure}
 

### 2025-07-18 paper/main.tex
insert table here
mark sections with TODO
extend introduction with more references and bullet points

@@
 %###################################################################################################
-\section{Introduction}
+\section{Introduction (TODO)}
 \label{section:introduction}
@@
 hoops~\cite{DePedro2003}.
-Despite their differences in wall-size, B.subtilis and E.coli follow growth rules which are
-comparable with respect to the extension of the rod~\cite{Chang2014}.
+Despite their differences in wall-size, \textit{B.subtilis} and \textit{E.coli} follow growth rules
+which are comparable with respect to the extension of the rod~\cite{Chang2014}.
+
+\paragraph{Experimental studies of mechanical properties}
+\textbf{Study of Individual Bacteria}
+\begin{itemize}
+    \item \cite{Robert2014} growth of individual rods, distribution of parameters
+    \item \cite{Koutsoumanis2013} inheritance of growth rates, growth parameter distributions
+    \item \cite{Ursell2014} MreB acts againts curvature during growth of cell
+    \item \cite{Takeuchi2005} cells in microchambers leads to plastic deformations
+    \item \cite{Amir2014} bending stiffness, plastic + elastic deformations
+\end{itemize}
+\textbf{Collective Effects \& Self-Organization}
+\begin{itemize}
+    \item \cite{Trejo2013} mechanical response of bacterial colony
+    \item \cite{Duvernoy2018} bacterial colonies form multilayers, interactions are polar
+    \item \cite{vanGestel2015} van Gogh bundles form from interplay of 2 cell-types
+\end{itemize}
+
+% --------------------------------------------------------------------------------------------------
+\subsection{Numerical Models of Rod-Shaped Bacteria}
+\textbf{Requirements}
+\begin{itemize}
+    \item capture bendiness; do not assume rigid bodies
+    \item model in 2D and 3D possible
+    \item individual-based model; model single cells
+    \item collective phenomena; describe many interacting cells, not just a single one
+\end{itemize}
+
+\textbf{Computational Modeling Frameworks}
+What has been done so far?
+\begin{itemize}
+    \item In general: not many frameworks which support some form of rod-shaped bacteria
+    \item \cite{Kang2014} Biocellion has support for cylindrically-shaped agents; no
+        bendiness/rigidity or polar interactions
+    \item \cite{Matyjaszkiewicz2017} BSim2.0 represents cells as rigid capsular cells made from a
+        cylindrical center part and two half-spheres, which are placed at the ends of the cylinder
+        to round out the shape. In order to calculate interactions between cells, possible overlaps
+        are determined and minimized, thus determining the position values of the next iteration
+        step.
+        BSim does not consider bending forces for individual cells or polar interactions.
+    \item \cite{Gutirrez2017} The gro programming language was designed to simulate the growth of
+        colonies and cell-cell communication. Its “physics computation has been optimized for rigid
+        rod-shaped bodies, like E. coli bacteria” (Gutiérrez et al. 2017). They recognize two types
+        of forces which are acting on the cellular agents: Local forces which are calculated between
+        adjacent bacteria and a global force which pushes bacteria outwards of the colony. The
+        latter of these is a phenomenological implementation of the observed colony expansion and
+        the associated central pressure with it. This assumption may yield incorrect results for
+        sparsely populated cases.
+        The engine is limited to 2D and does not consider polar interactions or bending of the rods.
+\end{itemize}
+
+% --------------------------------------------------------------------------------------------------
+\subsection{Purpose-built Models}
+\begin{itemize}
+    \item \cite{Abar2017} Review of various ABMs
+    \item \cite{Winkle2017} Modeling mechanical interations in growing populations of rod-shaped
+        bacteria
+    \item \cite{Doumic2020} "A purely mechanical model with asymmetric features for early
+        morphogenesis of rod-shaped bacteria micro-colony"
+    \begin{itemize}
+        \item Really good paper for referencing
+        \item comparison of distributions; not individual agents
+        \item no bending in model
+        \item uses steric force (see also \cite{Trejo2013})
+    \end{itemize}
+    \item \cite{Grant2014} purposely-built model written in C++
+    \begin{itemize}
+        \item describes bacteria as collection of overlapping spheres
+        \item spheres are coupled by non-linear springs (Euler-Bernoulli dynamic beam theory). This
+            assumption is werid but does not alter their results
+        \item Only model repulsive forces; no attraction, adhesion
+    \end{itemize}
+    \item \cite{Cho2007} only 2D model; no param estimation; based on work done in \cite{Jnsson2005}
+    \item \cite{Storck2014} no bending; 41 parameters for various cases; only 8 parameters taken
+        from literature values or quantified; generate growth rate from normal distribution at each
+        integration step; stochastically equivalient for collection of bacteria but unfounded for
+        individual bactera
+\end{itemize}
+
+%###################################################################################################
+\section{Overview (TODO)}
 
@@
 
-%###################################################################################################
-\subsection{Visualization}
+% --------------------------------------------------------------------------------------------------
+\subsection{Visualization (TODO)}
 \label{section:generating-masks-and-images}
@@
 % --------------------------------------------------------------------------------------------------
-\subsubsection{Relationship to Semantic Segmentation}
+\subsubsection{Relationship to Semantic Segmentation (TODO)}
 \label{paragraph:comparison-semantic-segmentation}
@@
 
-\begin{table}
+\begin{table}[H]
     \centering
@@
 %###################################################################################################
-\section{Discussion}
+\section{Discussion (TODO)}
 \label{section:discussion}
@@
 
-\section{Conclusion}
+\section{Conclusion (TODO)}
 \label{section:conclusion}

### 2025-07-18 paper/main.tex
move multilayer stuff to supplement

@@
             Rod-Shaped bacteria divide in the middle of the rod, thus forming two new
-            agents.
-            Bacteria can form multilayers during their growth phase \cite{Duvernoy2018}.\\
+            agents.\\
+            % Bacteria can form multilayers during their growth phase \cite{Duvernoy2018}.\\
         (4) & Variable Parameters &
@@
         %     Friction between cells \cite{Grant2014} can be asymmetrical \cite{Doumic2020}.\\
-        &\textbf{(DC) Domain-Cell Interactions}\\
-        \midrule
-        (6) & External Forces &
-            Bacteria which rub against surfaces tend to stick to them after some
-            time~\cite{vanLoosdrecht1989}.
-            The extracellular gel which surrounds the bacteria exerts a force onto the cells
-            \cite{Grant2014}.\\
+        % &\textbf{(DC) Domain-Cell Interactions}\\
+        % \midrule
+        % (6) & External Forces &
+        %     Bacteria which rub against surfaces tend to stick to them after some
+        %     time~\cite{vanLoosdrecht1989}.
+        %     The extracellular gel which surrounds the bacteria exerts a force onto the cells
+        %     \cite{Grant2014}.\\
         % (9) & Extracellular Reactions &
@@
 right after the division event.
-
-% --------------------------------------------------------------------------------------------------
-\subsection{External Forces}
-\label{subsection:mechanical-model-external-forces}
-TODO
-\begin{itemize}
-    \item Friction force against surface. perpendicular to motion; proportional to velocity.
-    \item gel pressure coming from experiment (simply downwards force $F_z$)
-\end{itemize}
 
@@
 % TODO
+
+\begin{itemize}
+    \item Friction force against surface. perpendicular to motion; proportional to velocity.
+    \item gel pressure coming from experiment (simply downwards force $F_z$)
+\end{itemize}
+
 \begin{itemize}

### 2025-07-18 paper/main.tex
formatting
extend descriptino of figure

@@
         text centered,
-        draw=black,
         fill=color4!70,
@@
 %###################################################################################################
-\section{Method Description}
+\section{Methodology}
+% --------------------------------------------------------------------------------------------------
 \subsection{Fundamental Assumptions on the Spatial Representation of Rod-Shaped Bacteria}
@@
 
-%###################################################################################################
+% --------------------------------------------------------------------------------------------------
 \subsection{Data Extraction}
@@
     \caption{
-        TODO change 2nd image!\\
-        (A) An example of a microscopic image.
-        (B) The result of the  skeletonization algorithm when applying it to a single cell.
-        (C) Extrapolated vertices $\vec{v}_i$ which represent the position of our cell-agents.
+        (A) Input image.
+        (B) Result of the skeletonization algorithm for a single cell.
+        (C) Vertices $\vec{v}_i$ which have been interpolated from the skeletonized cell.
+        They represent the position of our cell-agents.
     }
@@
 
-%###################################################################################################
+% --------------------------------------------------------------------------------------------------
 \subsection{Parameter Estimation}
@@
 
-% --------------------------------------------------------------------------------------------------
-\section{Application to Mechanics of Rod-Shaped Bacteria}
+%###################################################################################################
+\section{A Mechanical Model for Rod-Shaped Bacteria}
 \label{subsubsection:mechanical-model-rod-shaped-bacteria}

### 2025-07-23 paper/main.tex
add figure for final and initial states
remove section on data generation
move orcid.pdf to figures folder
create standalone file with overview

@@
 \usepackage{listings}
+\usepackage[mode=buildnew]{standalone}
 \usepackage{tikz}
@@
 
-\usetikzlibrary{trees}
-\usetikzlibrary{shapes.geometric, arrows}
-
-\definecolor{color1}{HTML}{6bd2db}
-\definecolor{color2}{HTML}{0ea7b5}
-\definecolor{color3}{HTML}{0c457d}
-\definecolor{color4}{HTML}{ffbe4f}
-\definecolor{color5}{HTML}{e8702a}
+\input{figures/tikz-preamble.tex}
 
@@
     \href{https://orcid.org/0009-0001-0613-7978}{
-        \includegraphics[scale=0.06]{orcid.pdf}
+        \includegraphics[scale=0.06]{figures/orcid.pdf}
         \hspace{1mm}Jonas Pleyer
@@
 	\href{https://orcid.org/0000-0002-6371-4495}{
-        \includegraphics[scale=0.06]{orcid.pdf}
+        \includegraphics[scale=0.06]{figures/orcid.pdf}
         \hspace{1mm}Christian Fleck
@@
     \centering
-    \tikzstyle{rect} = [
-        rectangle,
-        % rounded corners,
-        minimum width=4cm,
-        minimum height=1cm,
-        text centered,
-        fill=color4!70,
-    ]
-    \tikzstyle{brect} = [rect, fill=color5!50]
-    \tikzstyle{frect} = [rect, fill=color3!40]
-    \tikzstyle{arrow} = [thick,->,>=stealth]
-    \tikzstyle{darrow} = [arrow, dashed]
-    \begin{tikzpicture}
-        \node (model) [brect, align=center] at (0,0) {Mathematical \&\\ Computational Model};
-        \node (extraction) [rect] at (0, -2) {Data Extraction};
-        \node (params) [brect] at (6, -2) {Parameter Estimation};
-        \node (visualization) [rect] at (6, 0) {Visualization};
-
-        \draw[arrow] (model) -- (extraction);
-        \draw[darrow] (params) -- node[anchor=south, rotate=-18.43] {validate} (model);
-        \draw[arrow] (model) -- (visualization);
-        \draw[arrow] (visualization) -- (params);
-        \draw[arrow] (extraction) -- (params);
-
-        % Future stuff
-        \node (realimages) [frect] at (12, 0) {Data Generation};
-        \node (segmentation) [frect] at (12, -2) {Cell Segmentation};
-        \node (tracking) [frect] at (12, -3) {Cell Tracking};
-        \node (analysis) [frect] at (0, -4) {Analysis Methods};
-
-        \draw[arrow] (visualization) -- (realimages);
-        \draw[arrow] (realimages) -- (segmentation);
-        \draw[arrow] (extraction) -- (analysis);
-        \draw[arrow] (params) -- (realimages);
-    \end{tikzpicture}
+    % \includegraphics[width=0.8\textwidth]{overview.pdf}
+    \includestandalone[width=0.8\textwidth]{figures/concept/overview}
     \caption{
-        The project can be divided into 4 subparts (orange + yellow) that interact with and depend
-        on each other.
-        The orange blocks require input by the user by specifying the model and supplying input
-        data.
+        The project can be divided into 4 subparts (yellow) that interact with and depend on each
+        other.
         Additional blocks (blue-gray) indicate possible future extensions of this project.
@@
     \begin{tikzonimage}[width=0.33\textwidth]
+        {figures/crm_fit/snapshots/exact-004444.png}%
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.33\textwidth]
+        {figures/crm_fit/snapshots/exact-004444.png}%
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \end{tikzonimage}%
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \begin{tikzonimage}[width=0.33\textwidth]
         {figures/crm_fit/profiles/damping.pdf}%
@@
 \end{figure}
-
-% --------------------------------------------------------------------------------------------------
-\subsection{Data Generation}
 

### 2025-07-26 paper/main.tex
write visualization part

@@
 % --------------------------------------------------------------------------------------------------
-\subsection{Visualization (TODO)}
+\subsection{Visualization}
 \label{section:generating-masks-and-images}
-\begin{itemize}
-    \item use pyvista (VTK) for Visualization
-    \item always do 3D render. This way, we can have overalp between agents and produce more
-        realistic masks.
-    \item Assign unique color values for idividual agents.
-        We have a 1:1 mapping between agent identifier and color! (for each simulation different
-        though)
-    \item use parallel projection; this should be most close to what microscopes measure due to the
-        relatively large distance between object and lens.
-    \item use some noise to smear out the rods; this is already similar to microscopic images; use
-        more like this and deal with real-world defects to make it more realistic (future, not now)
-    \item Figure~\ref{fig:progression-image-generation} shows steps in generating image.
-    \item Reflectivity, placement of light etc. can already introduce some heterogeneity in the
-        visual of the rod.
-    \item Plan to implement noise on the mesh itself; make it have rough surface (like real rods as
-        seen under electron microscopy)
-\end{itemize}
-
-\paragraph{Algorithm}
-\begin{enumerate}
-    \item Get position of cell
-    \item Insert spheres at each vertex
-    \item Insert cylinder along edges
-    \item Combine individual objects into one big mesh
-    \item Use \ac{pbr}
-\end{enumerate}
+
+For visualization we use pyvista, which provides high-level wrappers around the popular VTK
+library~\cite{vtkBook,Sullivan2019}.
+The first stage of our visualization is to generate 3D meshes for each bacterium, which we consider
+an intermediate representation.
+We insert spheres of a given radius $r$ at each vertex $\vec{v}_i$ and connect adjacent vertices
+$\vec{v}_i,\vec{v}_{i+1}$ with a cylinder, also of radius $r$ which is positioned such that its core
+is identical to the connecting line segment between the two vertices.
+Finally, we combine the meshes of all of these objects to obtain a single mesh of one bacterium.
+A render of this result can be seen in figure~\ref{fig:progression-image-generation}.
+Since we use a 3D rendering engine as an intermediate representation, the resulting image can have
+overlaps of bacteria which are placed at higher or lower values along the z-direction.
+
+To generate masks, we assign unique colors to every mesh of the respective cellular agent.
+This mapping is unique, ensuring no two agents will use identical colors.
+We then use a parallel projection and remove any existing light sources in order to obtain a solid
+color for every cell-agent.
+The resulting image is shown in subfigure (B).
+Subfigure (C) shows an example of a simplified way to generate microscopic images.
+The properties of the meshes such as roughness and reflectivity have been adjusted and we use
+physically-based rendering of VTK.
+The result of this render is augmented by introducing white noise on the pixel values at each point.
+This approach is very rudimentary and we hope to extend these methods in the future.
+In order to properly generate data, we need to deal with microscopic imaging techniques and possible
+defects that can occurr.
+It is crucial to get this part of the image-generation process right in order to obtain
+near-realistic images that can be further reused.
+
+% \begin{itemize}
+%     \item use pyvista (VTK) for Visualization
+%     \item always do 3D render. This way, we can have overalp between agents and produce more
+%         realistic masks.
+%     \item Assign unique color values for idividual agents.
+%         We have a 1:1 mapping between agent identifier and color! (for each simulation different
+%         though)
+%     \item use parallel projection; this should be most close to what microscopes measure due to the
+%         relatively large distance between object and lens.
+%     \item use some noise to smear out the rods; this is already similar to microscopic images; use
+%         more like this and deal with real-world defects to make it more realistic (future, not now)
+%     \item Figure~\ref{fig:progression-image-generation} shows steps in generating image.
+%     \item Reflectivity, placement of light etc. can already introduce some heterogeneity in the
+%         visual of the rod.
+%     \item Plan to implement noise on the mesh itself; make it have rough surface (like real rods as
+%         seen under electron microscopy)
+% \end{itemize}
+% 
+% \paragraph{Algorithm}
+% \begin{enumerate}
+%     \item Get position of cell
+%     \item Insert spheres at each vertex
+%     \item Insert cylinder along edges
+%     \item Combine individual objects into one big mesh
+%     \item Use \ac{pbr}
+% \end{enumerate}
 
@@
     \centering
+    % \begin{tikzonimage}[width=0.9\textwidth]{visualization.pdf}
+    %     \node at (0.025, 0.975)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    % \end{tikzonimage}
     \begin{tikzonimage}[width=0.3\textwidth]
@@
     \begin{tikzonimage}[width=0.3\textwidth]
-        {docs/source/_static/09395645494836445480/images/000000400.png}
+        {docs/source/_static/09395645494836445480/masks/000000400.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
@@
     \begin{tikzonimage}[width=0.3\textwidth]
-        {docs/source/_static/09395645494836445480/masks/000000400.png}
+        {docs/source/_static/09395645494836445480/images/000000400.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
@@
     \caption{
-        (A) shows the raw result from combining sphere and cylinder meshes to obtain the shape of a
+        (A) shows the result from combining sphere and cylinder meshes to obtain the shape of a
         bacterial rod.
@@
         on the upper part of the meshes.
-        Subfigure (B) takes this image and applies noise to it in order to make it look more
-        like an actual microscopic image.
-        (C) The render is projected along the z-axis and flat numerical values are assigned to the
-        bacteria.
-        The resuling render can have overlaps in which case the color of the rod is displayed that
-        is higher along the z-direction.
+        (B) Unique color values are assigned to each agent and light sources are removed.
+        Afterwards, the image is rendered by projecting along the z-axis.
+        (C) We use physically-based rendering and assign properties such as roughness and
+        reflectivity to generate an initial render.
+        Afterwards, we apply white noise to the pixel values of image.
+        In the future, we hope to extend this approach in order to construct near-realistic
+        microscopic images from synthetic data.
     }

### 2025-07-26 paper/main.tex
rename toc
write initial version of overview

@@
 \pagebreak
-\renewcommand{\contentsname}{Temporary! Table of Contents}
+\renewcommand{\contentsname}{Table of Contents (remove before submission)}
 \tableofcontents
@@
 \section{Overview (TODO)}
-
-\begin{itemize}
-    \item This publication contains 5 things:
-    \begin{enumerate}
-        \item Methods of how to obtain data from microscopic images and perform parameter
-            estimations; $\rightarrow$ also possibly across multiple generations
-        \item Showcase the usefulness of the methods and the model by applying it to parameter
-            estimation via microscopic images and multilayer formation
-        \item Initial (incomplete) model of rod-shaped bacteria
-        \item Generate 3D Images (masks)
-        \item A python package which can do all of that and can be reused and extended
-    \end{enumerate}
-    \item Overall structure
-    \begin{enumerate}
-        \item users bring model + data
-        \item this paper provides methods to do the parameter estimation + visualization
-    \end{enumerate}
-    \item For complete model, more work has to be done
-    \begin{itemize}
-        \item More aspects (see Review; put in discussion)
-        \item Note that we only care about \textit{E.Coli} and \textit{B.Subtilits}! This is
-            important.
-    \end{itemize}
-    \begin{itemize}
-        \item State of the Art
-    \end{itemize}
-\end{itemize}
+This work consists of many subcomponents which can be combined in order to achieve various tasks.
+Figure~\ref{fig:flowchart-project-structure} shows a high-level overview of all components.
+This work assumes that bacterial rods can be represented by a colleciton of vertices $\vec{v}_i$.
+This assumption alone is enough to allow us to visualize the bacteria and extract positional data
+of bacterial agents from microscopic images.
+The dynamics of the vertices $\vec{v}_i$ are determined by a computational model which is defined
+by the respective researcher.
+It is important to note that the components visualization and data extraction do not require any
+knowledge about the computational model, thus allowing us to reuse their functionality across
+various models.
+
+Constructing a computational model requires knowledge about the target system and thus allows us to
+interpret the results gathered by the provided data extraction methods which can then further be
+used to restrict or determine the paramters of our model.
+For example, by extracing the positions of bacteria across a time-series of microscopic images, we
+can also calculate their lengths at these time-points and thus determine their rate of growth and
+type of growth (see section~\ref{subsec:parameter-estimation-individual-treatment}).
+
+The visualization methods are an integral part of any modeling endevaour.
+They allow researchers to obtain feedback quickly and improve iteration times when constructing new
+models.
+Furthermore, they are an integral part of our advanced parameter estimation techniques
+(see section~\ref{subsec:comparison-across-generations}) and enable us to effectively compare states
+across generations of cells.
+In the future, we are planning to extend this work and generate realistic microscopic images and
+time-series.
+Together with the parameter estimation methods, they can then be used to provide large amounts of
+mechanistic data for the training cell-segmentation and cell-tracking machine learning algorithms.
+
+% \begin{itemize}
+%     \item This publication contains 5 things:
+%     \begin{enumerate}
+%         \item Methods of how to obtain data from microscopic images and perform parameter
+%             estimations; $\rightarrow$ also possibly across multiple generations
+%         \item Showcase the usefulness of the methods and the model by applying it to parameter
+%             estimation via microscopic images and multilayer formation
+%         \item Initial (incomplete) model of rod-shaped bacteria
+%         \item Generate 3D Images (masks)
+%         \item A python package which can do all of that and can be reused and extended
+%     \end{enumerate}
+%     \item Overall structure
+%     \begin{enumerate}
+%         \item users bring model + data
+%         \item this paper provides methods to do the parameter estimation + visualization
+%     \end{enumerate}
+%     \item For complete model, more work has to be done
+%     \begin{itemize}
+%         \item More aspects (see Review; put in discussion)
+%         \item Note that we only care about \textit{E.Coli} and \textit{B.Subtilits}! This is
+%             important.
+%     \end{itemize}
+%     \begin{itemize}
+%         \item State of the Art
+%     \end{itemize}
+% \end{itemize}
 
@@
     \caption{
-        The project can be divided into 4 subparts (yellow) that interact with and depend on each
-        other.
-        Additional blocks (blue-gray) indicate possible future extensions of this project.
-        Arrows with solid lines indicate dependencies between the subparts while dashed arrows carry
-        special meaning, which is specifically annotated.
-        The mathematical models provides the theoretical framework and is thus the parent of all
-        following work.
-        "Data Extraction" allows us to read variable and parameter values directly from microscopic
-        images.
-        The obtained data can then be used to initialize a numerical simulations and thus estimate
-        parameters.
-        On the other hand, results produced by our model can be visualized as 3D renders and
-        projected to cell-masks.
-        These algorithms can be exploited to construct cost-functions which can be used in the
-        parameter estimation process.
-        In future work, this visualization scheme could be extended to generate realistic
-        microscopic images with cell-masks and thus train cell-segmentation and cell-tracking
-        algorithms.
+        The project can be divided into subcomponents (yellow, red) that interact with and depend on
+        each other.
+        Yellow blocks are provided by this work and red blocks require additional input such as
+        modeling assumptions or input data.
+        Blue blocks indicate possible future extensions of this project.
+        Every component builds on the spatial representation of rod-shaped bacteria as vertices
+        $\vec{v}_i$ for all proceeding work.
     }

### 2025-07-26 paper/main.tex
insert figures in place
remove bullet points
introduction to model

@@
 
-\begin{enumerate}
-    \item represent bacteria as polygon with vertices $\vec{v}_i$
-    \item Assumes that these vertices are not fundamental but rather a discretization of the actual
-        problem
-\end{enumerate}
+% \begin{enumerate}
+%     \item represent bacteria as polygon with vertices $\vec{v}_i$
+%     \item Assumes that these vertices are not fundamental but rather a discretization of the actual
+%         problem
+% \end{enumerate}
 
@@
 
-%###################################################################################################
-
-In order to construct a mechanical model of rod-shaped bacteria, we have to take care of a number of
-different aspects.
 The mechanical behaviour of rod-shaped bacteria is governed by a plethora of effects.
-Although we set out to construct a model which is able to describe as many of these effects as
-possible, we will fall short in this paper and restrict ourselves to the aspects presented
-in~\ref{table:simulation-aspects}.
-We will discuss, how this model can be extended in order to capture more effects and how we can
-validate our approach given the existing literaturesection:supplement-more-simulation-aspects
-(see section~\ref{section:discussion}, and~\ref{section:supplement-more-simulation-aspects}).
+In this work, we aim to establish a model which is abstract enough such that it can be reused for
+many types of rod-shaped bacteria but also explicit enough such that we can quantify its parameters.
+We are thus aiming to work with a set of assumptions that is as general as possible but still
+specific to rod-shaped bacteria.
+With this delicate balance in mind, it is clear that many cellular processes which might be specific
+to one particular celltype or simply not necessary to assume in order to construct our model, will
+be omitted alltogether.
+This concerns effects such as polar interactions~\cite{Duvernoy2018}, extracellular
+reactions~\cite{Wang2010} or differentiation processes~\cite{vanGestel2015,Lpez2010} and many other.
+We restrict ourselves to the aspects presented in table~\ref{table:simulation-aspects}.
+For a collection of other known effects which could be considered, we refer to the supplementary
+section~\ref{section:discussion}, and~\ref{section:supplement-more-simulation-aspects}.
 

### 2025-07-26 paper/main.tex
split applications part into multiple subsections

@@
 
-% --------------------------------------------------------------------------------------------------
-\subsection{Estimating Mechanical Parameters of Growing \textit{E.Coli.} on Agar}
+%###################################################################################################
+\section{Parameter Estimation of \textit{E.Coli.} growing on Agar}
+% --------------------------------------------------------------------------------------------------
+\subsection{Data Preparation}
+TODO: Decide if this should be Supplement
+\begin{enumerate}
+    \item Have masks of data; reference image segmantation tools; or do it by hand
+    \item ensure that colors are matching for bacteria between time-steps
+    \item clean borders of images if needed
+    \item know domain size; otherwise use pixels
+\end{enumerate}
+
+% --------------------------------------------------------------------------------------------------
+\subsection{Non-Individual Treatment}
+% --------------------------------------------------------------------------------------------------
+\subsection{Individual Treatment}
+\label{subsec:parameter-estimation-individual-treatment}
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.33\textwidth]
+        {figures/crm_estimate_params/IWF-Goettingen/rod-lengths-individual.pdf}%
+    \includegraphics[width=0.33\textwidth]
+        {figures/crm_estimate_params/IWF-Goettingen/rod-lengths-average.pdf}%
+    \includegraphics[width=0.33\textwidth]
+        {figures/crm_estimate_params/IWF-Goettingen/parameter-distribution.pdf}%
+    \caption{}
+    \label{fig:estimated-growth-rates}
+\end{figure}
+% --------------------------------------------------------------------------------------------------
+\subsection{With Division}
 We will now focus on using the presented methods for parameter estimation via microscopic images.

### 2025-07-26 paper/main.tex
fix appendix
add ugly intermediate introduction

@@
 
-Bacteria exist in many physical shapes such as spheroidal, rod-shaped or spiral
-\cite{Zapun2008,Young2006}.
 The last decades have fundamentally challenged and changed our understanding of how bacteria retain
 their shape.
+They exist in many physical shapes such as spheroidal, rod-shaped or spiral
+\cite{Zapun2008,Young2006}.
 Rod-shaped bacteria can grow by extending their circular part, or by inserting new material at the
 tip of the rod.
-Species such as E.coli and B.subtilis~\cite{Errington2020} fall under the first category while
+Species such as \textit{E.coli} and \textit{B.subtilis}~\cite{Errington2020} fall under the first
+category while
 S.pombe represents the latter.
@@
 hoops~\cite{DePedro2003}.
-Despite their differences in wall-size, \textit{B.subtilis} and \textit{E.coli} follow growth rules
-which are comparable with respect to the extension of the rod~\cite{Chang2014}.
+Despite their differences such as wall-size or length \textit{B.subtilis} and \textit{E.coli} follow
+growth rules which are comparable with respect to the extension of the rod~\cite{Chang2014}.
 
 \paragraph{Experimental studies of mechanical properties}
-\textbf{Study of Individual Bacteria}
-\begin{itemize}
-    \item \cite{Robert2014} growth of individual rods, distribution of parameters
-    \item \cite{Koutsoumanis2013} inheritance of growth rates, growth parameter distributions
-    \item \cite{Ursell2014} MreB acts againts curvature during growth of cell
-    \item \cite{Takeuchi2005} cells in microchambers leads to plastic deformations
-    \item \cite{Amir2014} bending stiffness, plastic + elastic deformations
-\end{itemize}
-\textbf{Collective Effects \& Self-Organization}
-\begin{itemize}
-    \item \cite{Trejo2013} mechanical response of bacterial colony
-    \item \cite{Duvernoy2018} bacterial colonies form multilayers, interactions are polar
-    \item \cite{vanGestel2015} van Gogh bundles form from interplay of 2 cell-types
-\end{itemize}
+It was long believed that bacteria lack cytoskeletal filaments but with the end of the last
+century, it was discovered that the MreB protein fills this role since it "[..] forms an actin-like
+cytoskeleton in bacteria [..]" \cite{Erickson2001}.
+It can polymerize and form filaments which behave similar to actin microfilaments \cite{Dersch2020}.
+Studies of its crystal structure further showed the similarity to actin \cite{vandenEnt2001} and the
+MreB, MreC and MreD proteins were identified as homologues of actin \cite{Lowe2017_lj}.\
+Jones et al. studied the species of different bacterial sub-kingdoms which have the MreB
+gene~\cite{Jones2001}.
+They concluded: "Many of the organisms are rod shaped, similar to \textit{B. subtilis} and
+\textit{E. coli}.
+Organisms with more complex shapes, including curved, filamentous, and helical bacteria, were all
+abundantly represented." \cite{Jones2001}.
+Furthermore, \cite{Wachi1987} found that mutants of \textit{E.Coli} which create defective MreB
+proteins will be spherical instead of their natural rod-shaped form.
+
+% \begin{itemize}
+%     \item \cite{Robert2014} growth of individual rods, distribution of parameters
+%     \item \cite{Koutsoumanis2013} inheritance of growth rates, growth parameter distributions
+%     \item \cite{Ursell2014} MreB acts againts curvature during growth of cell
+%     \item \cite{Takeuchi2005} cells in microchambers leads to plastic deformations
+%     \item \cite{Amir2014} bending stiffness, plastic + elastic deformations
+% \end{itemize}
+Other experiments found that individual rods of \textit{E.Coli} grow exponentially until the
+division event which is triggered by a size-sensing rather than a time-sensing
+mechanism~\cite{Robert2014}.
+These results also showed that the growth parameters were randomly distributed but did not quantify
+said distribution.
+Using $220$ \textit{Salmonella enterica}, Koutsoumanis et al. analysed the distribution of division
+times and was able to quantify their distributions~\cite{Koutsoumanis2013}.
+They found that while the initial division times were longer, there was no correlation between
+division times of mother and daughter cells.
+The first finding can be explained by the bacterias lag-phase~\cite{Bertrand2019} which is a
+temporary period during which bacteria do not grow.
+The second observation is an example of principle of thought which underlyies guides this work:
+Cells are individual agents with individual parameters and thus individual behaviour and need to be
+treated as such.
+This is especially important if we want to determine parameters of our model for individual agents.
+
+% \paragraph{Collective Effects \& Self-Organization}
+% are emergent phenomena that result from the combined interactions of multiple cells.
+% Although they do not allow us to fundamentally understand, the individual behaviour of cells
+% directly, they can provide a way to falsify our results and thus further guide our modeling
+% approach.
+% \cite{Trejo2013} investigated the mechanical response of a tightly packed colony of \textit{Bacillus
+% Subtilis}.
+% \cite{Trejo2013} studied the formation of macroscopic wrinkles when growing bacteria in a confined
+% space.
+% They were able to model the strain and stress of the collection of bacteria which requires that
+% cells are connected by a (to first order) spring-like force.
+% The developing wrinkles could then be described by a buckling instability.
+% 
+% Duvernoy et al. used laser techniques and force microscopy to investigate adhesion of individual
+% \textit{E.coli} and \textit{Pseudomonas aeruginosa} cells~\cite{Duvernoy2018}.
+% They found that the adhesion forces of the rods are polar, resulting in mechanical tension which in
+% turn determines daughter cell arrangement.
+% They measured the size of the grown bacterial colonies, fitted ellipses to their shapes and found
+% that adhesion and polar adhesion in particular resulted in a greater difference between the long and
+% short axis of the fitted ellipse and thus a more oval shape rather than circular (WT).
+% By extending the work of \cite{Grant2014}, they showed, that the critical value of transitioning
+% from a monolayer to multiple layers of bacteria, depends on the adhesive strength and polarity of
+% the interaction.
+% \cite{Grant2014} described the growing bacterial colony as a disc with radius $R$ which is pressed
+% into the agarose slab.
+% The authors created a purposely-designed simulation in `C++` in order to model the effects of the
+% gel on the collection of bacteria.
+% 
+% Van Gestel et al. explored how \textit{B.subtilis} migrates over a surface by forming multicellular
+% structures~\cite{vanGestel2015}.
+% The cells organize themselves into "van Gogh bundles" of cells which are tightly aligned in chains
+% and form filamentous loops.
+% This phenomenon occurs at the border of the bacterial colony and the migration is driven by two
+% phenotypically different cell types~\cite{Lpez2010}.
+% \begin{itemize}
+%     \item \cite{Trejo2013} mechanical response of bacterial colony
+%     \item \cite{Duvernoy2018} bacterial colonies form multilayers, interactions are polar
+%     \item \cite{vanGestel2015} van Gogh bundles form from interplay of 2 cell-types
+% \end{itemize}
 
@@
 
+\appendix
 %###################################################################################################
@@
 \renewcommand{\thesection}{S\arabic{section}}
-\setcounter{section}{0}
 

### 2025-07-27 paper/main.tex
relabel
first version of introduction

@@
 their shape.
-They exist in many physical shapes such as spheroidal, rod-shaped or spiral
-\cite{Zapun2008,Young2006}.
-Rod-shaped bacteria can grow by extending their circular part, or by inserting new material at the
-tip of the rod.
+They exist in many physical forms such as spheroidal, rod-shaped or
+spiral~\cite{Zapun2008,Young2006}.
+Rod-shaped bacteria can grow by extending their cylindrical part, or by inserting new material at
+the tip of the rod.
 Species such as \textit{E.coli} and \textit{B.subtilis}~\cite{Errington2020} fall under the first
-category while
-S.pombe represents the latter.
+category while \textit{S.pombe} represents the latter.
 New material is inserted in small bursts on the nanoscale in forms of patches, bands or
@@
 % Subtilis}.
-% \cite{Trejo2013} studied the formation of macroscopic wrinkles when growing bacteria in a confined
-% space.
-% They were able to model the strain and stress of the collection of bacteria which requires that
-% cells are connected by a (to first order) spring-like force.
-% The developing wrinkles could then be described by a buckling instability.
-% 
-% Duvernoy et al. used laser techniques and force microscopy to investigate adhesion of individual
-% \textit{E.coli} and \textit{Pseudomonas aeruginosa} cells~\cite{Duvernoy2018}.
-% They found that the adhesion forces of the rods are polar, resulting in mechanical tension which in
-% turn determines daughter cell arrangement.
+Trejo et al. studied the formation of macroscopic wrinkles when growing bacteria in a confined
+space~\cite{Trejo2013}.
+They were able to model the strain and stress of the collection of bacteria which requires that
+cells are connected by a (to first order) spring-like force.
+The developing wrinkles could then be described by a buckling instability.
+
+Duvernoy et al. used laser techniques and force microscopy to investigate adhesion of individual
+\textit{E.coli} and \textit{Pseudomonas aeruginosa} cells~\cite{Duvernoy2018}.
+They found that the adhesion forces of the rods are polar, resulting in mechanical tension which in
+turn determines daughter cell arrangement.
 % They measured the size of the grown bacterial colonies, fitted ellipses to their shapes and found
@@
 % short axis of the fitted ellipse and thus a more oval shape rather than circular (WT).
-% By extending the work of \cite{Grant2014}, they showed, that the critical value of transitioning
-% from a monolayer to multiple layers of bacteria, depends on the adhesive strength and polarity of
-% the interaction.
-% \cite{Grant2014} described the growing bacterial colony as a disc with radius $R$ which is pressed
-% into the agarose slab.
-% The authors created a purposely-designed simulation in `C++` in order to model the effects of the
-% gel on the collection of bacteria.
-% 
-% Van Gestel et al. explored how \textit{B.subtilis} migrates over a surface by forming multicellular
-% structures~\cite{vanGestel2015}.
-% The cells organize themselves into "van Gogh bundles" of cells which are tightly aligned in chains
-% and form filamentous loops.
-% This phenomenon occurs at the border of the bacterial colony and the migration is driven by two
-% phenotypically different cell types~\cite{Lpez2010}.
+By extending the work of \cite{Grant2014}, they further showed, that the critical value of
+transitioning from a monolayer to multiple layers of bacteria, depends on the adhesive strength and
+polarity of the interaction.
+Grant et al. described the growing bacterial colony as a disc with radius $R$ which is pressed
+into the agarose slab~\cite{Grant2014}.
+The authors created a purposely-designed simulation in `C++` in order to model the effects of the
+gel on the collection of bacteria.
+
+Van Gestel et al. explored how \textit{B.subtilis} migrates over a surface by forming multicellular
+structures~\cite{vanGestel2015}.
+The cells organize themselves into "van Gogh bundles" of cells which are tightly aligned in chains
+and form filamentous loops.
+This phenomenon occurs at the border of the bacterial colony and the migration is driven by two
+phenotypically different cell types~\cite{Lpez2010}.
 % \begin{itemize}
@@
 
-% --------------------------------------------------------------------------------------------------
-\subsection{Numerical Models of Rod-Shaped Bacteria}
-\textbf{Requirements}
-\begin{itemize}
-    \item capture bendiness; do not assume rigid bodies
-    \item model in 2D and 3D possible
-    \item individual-based model; model single cells
-    \item collective phenomena; describe many interacting cells, not just a single one
-\end{itemize}
-
-\textbf{Computational Modeling Frameworks}
-What has been done so far?
-\begin{itemize}
-    \item In general: not many frameworks which support some form of rod-shaped bacteria
-    \item \cite{Kang2014} Biocellion has support for cylindrically-shaped agents; no
-        bendiness/rigidity or polar interactions
-    \item \cite{Matyjaszkiewicz2017} BSim2.0 represents cells as rigid capsular cells made from a
-        cylindrical center part and two half-spheres, which are placed at the ends of the cylinder
-        to round out the shape. In order to calculate interactions between cells, possible overlaps
-        are determined and minimized, thus determining the position values of the next iteration
-        step.
-        BSim does not consider bending forces for individual cells or polar interactions.
-    \item \cite{Gutirrez2017} The gro programming language was designed to simulate the growth of
-        colonies and cell-cell communication. Its “physics computation has been optimized for rigid
-        rod-shaped bodies, like E. coli bacteria” (Gutiérrez et al. 2017). They recognize two types
-        of forces which are acting on the cellular agents: Local forces which are calculated between
-        adjacent bacteria and a global force which pushes bacteria outwards of the colony. The
-        latter of these is a phenomenological implementation of the observed colony expansion and
-        the associated central pressure with it. This assumption may yield incorrect results for
-        sparsely populated cases.
+\paragraph{Computational Modeling Frameworks which support Rod-Shaped Bacteria}
+% \textbf{Requirements}
+% \begin{itemize}
+%     \item capture bendiness; do not assume rigid bodies
+%     \item model in 2D and 3D possible
+%     \item individual-based model; model single cells
+%     \item collective phenomena; describe many interacting cells, not just a single one
+% \end{itemize}
+There have been many variations in modeling rod-shaped bacteria.
+In order to provide enough flexibility and generality, we are in particular interested in models
+that can capture non-rigid rods and work both in 2D and 3D scenarios.
+Furthermore, we require that our model can describe bacteria individually and that a collective
+model is simply the propduct of multiple individual agents interacting with each other.
+In order to be able to investigate more specialized models, we need to be able to adjust and modify
+any existing model.
+
+In general, there are two classes of existing models which could be considered to fit these
+criteria: purpose-built (onf-off) solutions and computational modeling frameworks which have been
+designed to provide predefined cellular representations for various use-cases~\cite{Pleyer2023}.
+However, explicit support for rod-shaped bacteria is rather sparse.
+Biocellion has support for cylindrically-shaped agents but does not provide any mechanisms of how to
+model bendiness or flexibility of the rod~\cite{Kang2014}.
+BSim2.0 represents cells as rigid capsular cells made from a cylindrical center part and two
+half-spheres, which are placed at the ends of the cylinder to round out the
+shape~\cite{Matyjaszkiewicz2017}.
+In order to calculate interactions between cells, possible overlaps are determined and minimized,
+thus determining the position values of the next iteration step.
+However BSim does not consider bending of the rods.
+The \texttt{gro} programming language was designed to simulate the growth of colonies and cell-cell
+communication~\cite{Gutirrez2017}.
+Its “physics computation has been optimized for rigid rod-shaped bodies, like E. coli
+bacteria”~\cite{Gutirrez2017}.
+They recognize two types of forces which are acting on the cellular agents:
+Local forces which are calculated between adjacent bacteria and a global force which pushes bacteria
+outwards of the colony.
+The latter of these is a phenomenological implementation of the observed colony expansion and the
+associated central pressure with it.
+This assumption may yield incorrect results for sparsely populated cases.
 The engine is limited to 2D and does not consider polar interactions or bending of the rods.
-\end{itemize}
-
-% --------------------------------------------------------------------------------------------------
-\subsection{Purpose-built Models}
-\begin{itemize}
-    \item \cite{Abar2017} Review of various ABMs
-    \item \cite{Winkle2017} Modeling mechanical interations in growing populations of rod-shaped
-        bacteria
-    \item \cite{Doumic2020} "A purely mechanical model with asymmetric features for early
-        morphogenesis of rod-shaped bacteria micro-colony"
-    \begin{itemize}
-        \item Really good paper for referencing
-        \item comparison of distributions; not individual agents
-        \item no bending in model
-        \item uses steric force (see also \cite{Trejo2013})
-    \end{itemize}
-    \item \cite{Grant2014} purposely-built model written in C++
-    \begin{itemize}
-        \item describes bacteria as collection of overlapping spheres
-        \item spheres are coupled by non-linear springs (Euler-Bernoulli dynamic beam theory). This
-            assumption is werid but does not alter their results
-        \item Only model repulsive forces; no attraction, adhesion
-    \end{itemize}
-    \item \cite{Cho2007} only 2D model; no param estimation; based on work done in \cite{Jnsson2005}
-    \item \cite{Storck2014} no bending; 41 parameters for various cases; only 8 parameters taken
-        from literature values or quantified; generate growth rate from normal distribution at each
-        integration step; stochastically equivalient for collection of bacteria but unfounded for
-        individual bactera
-\end{itemize}
+
+% \textbf{Computational Modeling Frameworks}
+% What has been done so far?
+% \begin{itemize}
+%     \item In general: not many frameworks which support some form of rod-shaped bacteria
+%     \item \cite{Kang2014} Biocellion has support for cylindrically-shaped agents; no
+%         bendiness/rigidity or polar interactions
+%     \item \cite{Matyjaszkiewicz2017} BSim2.0 represents cells as rigid capsular cells made from a
+%         cylindrical center part and two half-spheres, which are placed at the ends of the cylinder
+%         to round out the shape. In order to calculate interactions between cells, possible overlaps
+%         are determined and minimized, thus determining the position values of the next iteration
+%         step.
+%         BSim does not consider bending forces for individual cells or polar interactions.
+%     \item \cite{Gutirrez2017} The gro programming language was designed to simulate the growth of
+%         colonies and cell-cell communication. Its “physics computation has been optimized for rigid
+%         rod-shaped bodies, like E. coli bacteria” (Gutiérrez et al. 2017). They recognize two types
+%         of forces which are acting on the cellular agents: Local forces which are calculated between
+%         adjacent bacteria and a global force which pushes bacteria outwards of the colony. The
+%         latter of these is a phenomenological implementation of the observed colony expansion and
+%         the associated central pressure with it. This assumption may yield incorrect results for
+%         sparsely populated cases.
+%         The engine is limited to 2D and does not consider polar interactions or bending of the rods.
+% \end{itemize}
+
+\paragraph{Applications \& Parameter Estimation}
+% \begin{itemize}
+%     \item \cite{Winkle2017} Modeling mechanical interations in growing populations of rod-shaped
+%         bacteria
+%     \item \cite{Doumic2020} "A purely mechanical model with asymmetric features for early
+%         morphogenesis of rod-shaped bacteria micro-colony"
+%     \begin{itemize}
+%         \item Really good paper for referencing
+%         \item comparison of distributions; not individual agents
+%         \item no bending in model
+%         \item uses steric force (see also \cite{Trejo2013})
+%     \end{itemize}
+%     \item \cite{Grant2014} purposely-built model written in C++
+%     \begin{itemize}
+%         \item describes bacteria as collection of overlapping spheres
+%         \item spheres are coupled by non-linear springs (Euler-Bernoulli dynamic beam theory). This
+%             assumption is werid but does not alter their results
+%         \item Only model repulsive forces; no attraction, adhesion
+%     \end{itemize}
+%     \item \cite{Cho2007} only 2D model; no param estimation; based on work done in \cite{Jnsson2005}
+%     \item \cite{Storck2014} no bending; 41 parameters for various cases; only 8 parameters taken
+%         from literature values or quantified; generate growth rate from normal distribution at each
+%         integration step; stochastically equivalient for collection of bacteria but unfounded for
+%         individual bactera
+%     \item \cite{You2018} TODO
+% \end{itemize}
+
+Cho et al.~\cite{Cho2007} extended the work done by~\cite{Jnsson2005} and construct a 2D model to
+study self-organization of colonies within restricted geometries of microfluidic devices.
+They do not provide any estimates of the parameters that were used.
+
+You et al. assumed a hard-rod model to study the emergence of order within microdomains  of
+rod-shaped bacteria~\cite{You2018}.
+They were able to show that the hydrodynamic equations of active nematic liquid crystals can serve
+as a continuum limit of their model.
+
+Winkle et al. represented cells with "[..] two acially independent cell halves that attach through a
+compressible, stiff spring [..]"~\cite{Winkle2017} and leveraged parts of the \texttt{gro}
+infrastructure.
+However, they only discuss results qualitatively without quantifying any of the used parameters.
+
+Doumic et al. constructed a model which takes into account asymmetries of the individual bacteria
+such as length, weight and friction~\cite{Doumic2020}.
+They were able to estimate some of the used parameters by comparing distributions of simulated data
+to distributions of experimental data.
+However, in doing so, no optimization optimization methods were used, but parameters were picked by
+hand in an attempt to best fit the observed results.
+Furthermore, their model is not able to describe any deformations such as bending of the rods.
+
+Grant et al. designed a custom \texttt{C++} program to simulate individual
+bacteria~\cite{Grant2014} which are represented as elastic rods of overlapping spheres.
+Their model does not include any attractive forces between agents but is able to describe bending of
+the rods in 2D and 3D.
+When it comes to the task of estimating parameters of the model, values are picked by hand in order
+to mimick the observed experimental results.
+
+Storck et al. constructed a flexible model which is based on particles and springs and thus able to
+describe a variety of cells~\cite{Storck2014}.
+Since this model incorporates many effects, it consists of $41$ parameters of which only $8$ have
+values that are taken from litearture or otherwise quantified.
+Another choice of their modeling approach was to stochastically assign growth bursts at each
+integration step of the numerical solving process.
+When generating from a normal distribution and observing on a collective level, this is
+stochastically equivalent to picking and fixing growth rates for each cell from the same
+distribution but is not identical when observing individual agents.
+
+\paragraph{The missing Link}
+It is clear that researchers are interested in studying mechanical properties of rod-like bacteria
+in order to better understand phenomena such as biofilm formation.
+The major downside of existing computational models is that there are no methods to properly
+quantify their parameters which diminishes their explanatory power.
+By leveraging a simple assumption about the cellular representation, this work aims to solve this
+particular problem.
 
@@
 \subsubsection{Comparison across Generations}
-\label{paragraph:comparison-across-generations}
+\label{subsec:comparison-across-generations}
 

### 2025-07-27 paper/main.tex
first version of overview

@@
 Figure~\ref{fig:flowchart-project-structure} shows a high-level overview of all components.
-This work assumes that bacterial rods can be represented by a colleciton of vertices $\vec{v}_i$.
-This assumption alone is enough to allow us to visualize the bacteria and extract positional data
-of bacterial agents from microscopic images.
-The dynamics of the vertices $\vec{v}_i$ are determined by a computational model which is defined
-by the respective researcher.
-It is important to note that the components visualization and data extraction do not require any
-knowledge about the computational model, thus allowing us to reuse their functionality across
-various models.
+We assume that bacterial rods can be represented by a colleciton of vertices $\vec{v}_i$.
+This assumption alone is enough to enable a cascade of other components.
+Using the spatial representation, we provide visualization methods that create 3D models from the
+given vertices which can then either be rendered or projected to a 2D plane.
+The data extraction component determines the vertices $\vec{v}_i$ from given cell masks which
+correspond to microscopic images.
+This allows us in particular to turn any sequence of images (movie) into a sequence of bacterial
+positions.
+The dynamics of the vertices $\vec{v}_i$ are determined by a computational model which is specific
+to the respective use-case and should be defined by the researcher.
+It merely needs to utilize the aforementioned spatial representation of bacterial rods as a
+collection of vertices $\vec{v}_i$.
+This work will provide one example of how such a model can be defined.
+Notably, the visualization and data extraction components do not require any knowledge about the
+computational model, thus allowing us to reuse their functionality across multiple models.
 
 Constructing a computational model requires knowledge about the target system and thus allows us to
-interpret the results gathered by the provided data extraction methods which can then further be
-used to restrict or determine the paramters of our model.
+interpret the results gathered by the provided data extraction methods.
+These results can then further be used to restrict or determine the parameters of our model.
 For example, by extracing the positions of bacteria across a time-series of microscopic images, we
 can also calculate their lengths at these time-points and thus determine their rate of growth and
-type of growth (see section~\ref{subsec:parameter-estimation-individual-treatment}).
-
-The visualization methods are an integral part of any modeling endevaour.
+type of growth (see section~\ref{subsec:parameter-estimation-individual-treatment}), thus
+effectively guiding our modeling steps.
+
+Visualization methods are an integral part of any modeling endevaour.
 They allow researchers to obtain feedback quickly and improve iteration times when constructing new
 models.
-Furthermore, they are an integral part of our advanced parameter estimation techniques
-(see section~\ref{subsec:comparison-across-generations}) and enable us to effectively compare states
-across generations of cells.
-In the future, we are planning to extend this work and generate realistic microscopic images and
-time-series.
+Furthermore, we utilize them in our more advanced parameter estimation methods (see
+section~\ref{subsec:comparison-across-generations}) where they enable us to effectively compare
+two bacterial configurations across division events.
+In the future, we are planning to build on the current state and generate realistic microscopic
+images and time-series.
 Together with the parameter estimation methods, they can then be used to provide large amounts of
-mechanistic data for the training cell-segmentation and cell-tracking machine learning algorithms.
+mechanistic data for the training cell-segmentation and cell-tracking machine-learning algorithms.
 
@@
         Yellow blocks are provided by this work and red blocks require additional input such as
-        modeling assumptions or input data.
+        modeling assumptions or data.
         Blue blocks indicate possible future extensions of this project.
-        Every component builds on the spatial representation of rod-shaped bacteria as vertices
-        $\vec{v}_i$ for all proceeding work.
+        Every component implicitly builds on the spatial representation of rod-shaped bacteria as
+        vertices $\vec{v}_i$.
     }

### 2025-07-27 paper/main.tex
rename section
formatting of figure
update catpion

@@
 % --------------------------------------------------------------------------------------------------
-\subsection{Fundamental Assumptions on the Spatial Representation of Rod-Shaped Bacteria}
+\subsection{Spatial Representation of Rod-Shaped Bacteria}
 % TODO CITATIONS
@@
     \linebreak
+    \vspace{0.01\textwidth}
     \begin{tikzonimage}[width=0.48\textwidth]
@@
         The figure shows two masks (A,B) which will be compared against each other.
-        As we can see, image (B) contains more cells which has resulted from a division process that
-        is not yet occurred in image (A).
-        In (C) we can see the "naive" comparison, simply identifying non-matching colors.
-        By assigning penalties to all highlighted pixels, we would not take into consideration the
-        parental relationship.
-        This approach is displayed in (D) where pixels of cells that are related have been colored
-        in gray.\\
-        Total cost between successive iteration steps $t_n,t_{n+1}$.
-        The overall shape is expected to have the slope of an exponential function.
+        Image (B) contains more cells which are a result of cellular division events that have not
+        yet occurred in image (A).
+        (C) "Naive" comparison of cell masks, which identifies pixels of non-matching colors and
+        assigns a cost by counting these pixels.
+        This simple approach does not take into consideration the parental relationship.
+        This approach is displayed in (D) where pixels of cells that are directly related have been
+        colored in gray.\\
+        (E) Comparison of cost functions, exponential regression (ER) and number of cells.
+        % Total cost between successive iteration steps $t_n,t_{n+1}$.
+        The overall shape of the cost is expected to have the slope of an exponential function.
         We can clearly see that division events introduce jumps within this relationship.

### 2025-07-27 paper/main.tex
rename sections
remove label
add citation as reminder
insert citation

@@
 \subsubsection{Relationship to Semantic Segmentation (TODO)}
-\label{paragraph:comparison-semantic-segmentation}
 \cite{Jadon2020}
@@
 %###################################################################################################
-\section{A Mechanical Model for Rod-Shaped Bacteria}
+\section{A Computational Model for Rod-Shaped Bacteria}
 \label{subsubsection:mechanical-model-rod-shaped-bacteria}
@@
 The details of the division event (i.e. Z-Ring formation) are very intricate and are heavily
-studied as a subject on their own (TODO citation).
+studied as a subject on their own~\cite{Harry2001}.
 On a higher level view, the rod is split apart at a point close to the center of the cell, thus
@@
 %###################################################################################################
-\section{Parameter Estimation of \textit{E.Coli.} growing on Agar}
+\section{Application to \textit{E.Coli.} growing on Agar (TODO)}
 % --------------------------------------------------------------------------------------------------
@@
 
+\cite{You2019}
+
 \begin{figure}[H]

### 2025-07-27 paper/main.tex
switch subsections
remove some TODOs

@@
 %###################################################################################################
-\section{Introduction (TODO)}
+\section{Introduction}
 \label{section:introduction}
@@
 %###################################################################################################
-\section{Overview (TODO)}
+\section{Overview}
 This work consists of many subcomponents which can be combined in order to achieve various tasks.
@@
 
-% TODO write more text here
-% Use some parts of the proposal text
-
 %###################################################################################################
@@
 \subsection{Spatial Representation of Rod-Shaped Bacteria}
-% TODO CITATIONS
 This work makes use of a few but powerfull assumptions concerning the mathematical representation of
@@
     \label{fig:mechanics-bacterium}
+\end{figure}
+
+% --------------------------------------------------------------------------------------------------
+\subsection{Visualization}
+\label{section:generating-masks-and-images}
+
+For visualization we use pyvista, which provides high-level wrappers around the popular VTK
+library~\cite{vtkBook,Sullivan2019}.
+The first stage of our visualization is to generate 3D meshes for each bacterium, which we consider
+an intermediate representation.
+We insert spheres of a given radius $r$ at each vertex $\vec{v}_i$ and connect adjacent vertices
+$\vec{v}_i,\vec{v}_{i+1}$ with a cylinder, also of radius $r$ which is positioned such that its core
+is identical to the connecting line segment between the two vertices.
+Finally, we combine the meshes of all of these objects to obtain a single mesh of one bacterium.
+A render of this result can be seen in figure~\ref{fig:progression-image-generation}.
+Since we use a 3D rendering engine as an intermediate representation, the resulting image can have
+overlaps of bacteria which are placed at higher or lower values along the z-direction.
+
+To generate masks, we assign unique colors to every mesh of the respective cellular agent.
+This mapping is unique, ensuring no two agents will use identical colors.
+We then use a parallel projection and remove any existing light sources in order to obtain a solid
+color for every cell-agent.
+The resulting image is shown in subfigure (B).
+Subfigure (C) shows an example of a simplified way to generate microscopic images.
+The properties of the meshes such as roughness and reflectivity have been adjusted and we use
+physically-based rendering of VTK.
+The result of this render is augmented by introducing white noise on the pixel values at each point.
+This approach is very rudimentary and we hope to extend these methods in the future.
+In order to properly generate data, we need to deal with microscopic imaging techniques and possible
+defects that can occurr.
+It is crucial to get this part of the image-generation process right in order to obtain
+near-realistic images that can be further reused.
+
+% \begin{itemize}
+%     \item use pyvista (VTK) for Visualization
+%     \item always do 3D render. This way, we can have overalp between agents and produce more
+%         realistic masks.
+%     \item Assign unique color values for idividual agents.
+%         We have a 1:1 mapping between agent identifier and color! (for each simulation different
+%         though)
+%     \item use parallel projection; this should be most close to what microscopes measure due to the
+%         relatively large distance between object and lens.
+%     \item use some noise to smear out the rods; this is already similar to microscopic images; use
+%         more like this and deal with real-world defects to make it more realistic (future, not now)
+%     \item Figure~\ref{fig:progression-image-generation} shows steps in generating image.
+%     \item Reflectivity, placement of light etc. can already introduce some heterogeneity in the
+%         visual of the rod.
+%     \item Plan to implement noise on the mesh itself; make it have rough surface (like real rods as
+%         seen under electron microscopy)
+% \end{itemize}
+%
+% \paragraph{Algorithm}
+% \begin{enumerate}
+%     \item Get position of cell
+%     \item Insert spheres at each vertex
+%     \item Insert cylinder along edges
+%     \item Combine individual objects into one big mesh
+%     \item Use \ac{pbr}
+% \end{enumerate}
+
+\begin{figure}[H]
+    \centering
+    % \begin{tikzonimage}[width=0.9\textwidth]{visualization.pdf}
+    %     \node at (0.025, 0.975)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    % \end{tikzonimage}
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {docs/source/_static/09395645494836445480/raw_pv/000000400.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {docs/source/_static/09395645494836445480/masks/000000400.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+    \end{tikzonimage}
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {docs/source/_static/09395645494836445480/images/000000400.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+    \end{tikzonimage}
+    \caption{
+        (A) shows the result from combining sphere and cylinder meshes to obtain the shape of a
+        bacterial rod.
+        This render contains lighting coming from 2 light sources which thus produces a small glow
+        on the upper part of the meshes.
+        (B) Unique color values are assigned to each agent and light sources are removed.
+        Afterwards, the image is rendered by projecting along the z-axis.
+        (C) We use physically-based rendering and assign properties such as roughness and
+        reflectivity to generate an initial render.
+        Afterwards, we apply white noise to the pixel values of image.
+        In the future, we hope to extend this approach in order to construct near-realistic
+        microscopic images from synthetic data.
+    }
+    \label{fig:progression-image-generation}
 \end{figure}
@@
     \label{fig:benchmarking-extraction-algorithm}
-\end{figure}
-
-% --------------------------------------------------------------------------------------------------
-\subsection{Visualization}
-\label{section:generating-masks-and-images}
-
-For visualization we use pyvista, which provides high-level wrappers around the popular VTK
-library~\cite{vtkBook,Sullivan2019}.
-The first stage of our visualization is to generate 3D meshes for each bacterium, which we consider
-an intermediate representation.
-We insert spheres of a given radius $r$ at each vertex $\vec{v}_i$ and connect adjacent vertices
-$\vec{v}_i,\vec{v}_{i+1}$ with a cylinder, also of radius $r$ which is positioned such that its core
-is identical to the connecting line segment between the two vertices.
-Finally, we combine the meshes of all of these objects to obtain a single mesh of one bacterium.
-A render of this result can be seen in figure~\ref{fig:progression-image-generation}.
-Since we use a 3D rendering engine as an intermediate representation, the resulting image can have
-overlaps of bacteria which are placed at higher or lower values along the z-direction.
-
-To generate masks, we assign unique colors to every mesh of the respective cellular agent.
-This mapping is unique, ensuring no two agents will use identical colors.
-We then use a parallel projection and remove any existing light sources in order to obtain a solid
-color for every cell-agent.
-The resulting image is shown in subfigure (B).
-Subfigure (C) shows an example of a simplified way to generate microscopic images.
-The properties of the meshes such as roughness and reflectivity have been adjusted and we use
-physically-based rendering of VTK.
-The result of this render is augmented by introducing white noise on the pixel values at each point.
-This approach is very rudimentary and we hope to extend these methods in the future.
-In order to properly generate data, we need to deal with microscopic imaging techniques and possible
-defects that can occurr.
-It is crucial to get this part of the image-generation process right in order to obtain
-near-realistic images that can be further reused.
-
-% \begin{itemize}
-%     \item use pyvista (VTK) for Visualization
-%     \item always do 3D render. This way, we can have overalp between agents and produce more
-%         realistic masks.
-%     \item Assign unique color values for idividual agents.
-%         We have a 1:1 mapping between agent identifier and color! (for each simulation different
-%         though)
-%     \item use parallel projection; this should be most close to what microscopes measure due to the
-%         relatively large distance between object and lens.
-%     \item use some noise to smear out the rods; this is already similar to microscopic images; use
-%         more like this and deal with real-world defects to make it more realistic (future, not now)
-%     \item Figure~\ref{fig:progression-image-generation} shows steps in generating image.
-%     \item Reflectivity, placement of light etc. can already introduce some heterogeneity in the
-%         visual of the rod.
-%     \item Plan to implement noise on the mesh itself; make it have rough surface (like real rods as
-%         seen under electron microscopy)
-% \end{itemize}
-% 
-% \paragraph{Algorithm}
-% \begin{enumerate}
-%     \item Get position of cell
-%     \item Insert spheres at each vertex
-%     \item Insert cylinder along edges
-%     \item Combine individual objects into one big mesh
-%     \item Use \ac{pbr}
-% \end{enumerate}
-
-\begin{figure}[H]
-    \centering
-    % \begin{tikzonimage}[width=0.9\textwidth]{visualization.pdf}
-    %     \node at (0.025, 0.975)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{A}};
-    % \end{tikzonimage}
-    \begin{tikzonimage}[width=0.3\textwidth]
-        {docs/source/_static/09395645494836445480/raw_pv/000000400.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.3\textwidth]
-        {docs/source/_static/09395645494836445480/masks/000000400.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.3\textwidth]
-        {docs/source/_static/09395645494836445480/images/000000400.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
-    \end{tikzonimage}
-    \caption{
-        (A) shows the result from combining sphere and cylinder meshes to obtain the shape of a
-        bacterial rod.
-        This render contains lighting coming from 2 light sources which thus produces a small glow
-        on the upper part of the meshes.
-        (B) Unique color values are assigned to each agent and light sources are removed.
-        Afterwards, the image is rendered by projecting along the z-axis.
-        (C) We use physically-based rendering and assign properties such as roughness and
-        reflectivity to generate an initial render.
-        Afterwards, we apply white noise to the pixel values of image.
-        In the future, we hope to extend this approach in order to construct near-realistic
-        microscopic images from synthetic data.
-    }
-    \label{fig:progression-image-generation}
 \end{figure}

### 2025-07-29 paper/main.tex
add reminder todo for abstract
introduce linebreak in title

@@
 
-\title{Methods for Analyzing Models of Rod-Shaped Bacteria}
+\title{Methods for Analyzing Models of\\ Rod-Shaped Bacteria}
 
@@
 \begin{abstract}
+    \textbf{!TODO!}
     This paper establishes the basis for a novel individual-based mechanical model of rod-shaped

### 2025-07-29 paper/main.tex
remove abstract and insert todo

@@
 \begin{abstract}
-    \textbf{!TODO!}
-    This paper establishes the basis for a novel individual-based mechanical model of rod-shaped
-    bacteria.
-    It describes methods of how to obtain data from microscopic images such that they can be used
-    for parameter estimation purposes.
-    Furthermore, we show initial results for parameter estimates and an application of the model to
-    the formation of multilayers within bacterial colonies.
+    \textbf{TODO: ABSTRACT}
+    % This paper establishes the basis for a novel individual-based mechanical model of rod-shaped
+    % bacteria.
+    % It describes methods of how to obtain data from microscopic images such that they can be used
+    % for parameter estimation purposes.
+    % Furthermore, we show initial results for parameter estimates and an application of the model to
+    % the formation of multilayers within bacterial colonies.
 \end{abstract}

### 2025-07-29 paper/main.tex
rename vertices from v to x

@@
 Figure~\ref{fig:flowchart-project-structure} shows a high-level overview of all components.
-We assume that bacterial rods can be represented by a colleciton of vertices $\vec{v}_i$.
+We assume that bacterial rods can be represented by a colleciton of vertices $\vec{x}_i$.
 This assumption alone is enough to enable a cascade of other components.
@@
 given vertices which can then either be rendered or projected to a 2D plane.
-The data extraction component determines the vertices $\vec{v}_i$ from given cell masks which
+The data extraction component determines the vertices $\vec{x}_i$ from given cell masks which
 correspond to microscopic images.
@@
 positions.
-The dynamics of the vertices $\vec{v}_i$ are determined by a computational model which is specific
+The dynamics of the vertices $\vec{x}_i$ are determined by a computational model which is specific
 to the respective use-case and should be defined by the researcher.
 It merely needs to utilize the aforementioned spatial representation of bacterial rods as a
-collection of vertices $\vec{v}_i$.
+collection of vertices $\vec{x}_i$.
 This work will provide one example of how such a model can be defined.
@@
         Every component implicitly builds on the spatial representation of rod-shaped bacteria as
-        vertices $\vec{v}_i$.
+        vertices $\vec{x}_i$.
     }
@@
 Taking into account the natural shape of rod-like bacteria, it is clear that they can not be
-represented with a singular point $\vec{v}$ in space.
+represented with a singular point $\vec{x}$ in space.
 We could extend this positional definition with a vector $\vec{w}$, indicating the direction of
@@
 Thus, we extend our model from a single point to multiple vertices which are part of a 1-dimensional
-polygon with two endpoints $\vec{v}_i$.
+polygon with two endpoints $\vec{x}_i$.
 In essence, we discretize the (possibly non-straight) rod of the agent.
 Figure~\ref{fig:mechanics-bacterium} shows a visualization of this approach.
-The angles $\alpha_i$ between vertices $\vec{v}_{i-1},\vec{v}_i$ can be used when modeling the
+The angles $\alpha_i$ between vertices $\vec{x}_{i-1},\vec{x}_i$ can be used when modeling the
 dynamics of these bacteria.
@@
 % \begin{enumerate}
-%     \item represent bacteria as polygon with vertices $\vec{v}_i$
+%     \item represent bacteria as polygon with vertices $\vec{x}_i$
 %     \item Assumes that these vertices are not fundamental but rather a discretization of the actual
@@
 an intermediate representation.
-We insert spheres of a given radius $r$ at each vertex $\vec{v}_i$ and connect adjacent vertices
-$\vec{v}_i,\vec{v}_{i+1}$ with a cylinder, also of radius $r$ which is positioned such that its core
+We insert spheres of a given radius $r$ at each vertex $\vec{x}_i$ and connect adjacent vertices
+$\vec{x}_i,\vec{x}_{i+1}$ with a cylinder, also of radius $r$ which is positioned such that its core
 is identical to the connecting line segment between the two vertices.
@@
 
-This section will explain how we can determine the collection of vertices $\vec{v}_i$ introduced in
+This section will explain how we can determine the collection of vertices $\vec{x}_i$ introduced in
 subsection~\ref{subsection:mechanical-model-mechanics} from microscopic images
@@
     \item Sort points such that they form a polygon.
-    \item Interpolate vertices $\vec{v}_i$ from obtained polygon.
+    \item Interpolate vertices $\vec{x}_i$ from obtained polygon.
 \end{enumerate}
@@
 In the last step, we take the sorted, extracted points $\vec{p}_i$ and interpolate $n$ values for
-the desired number of vertices $\vec{v}_i$.
+the desired number of vertices $\vec{x}_i$.
 
@@
         (B) Result of the skeletonization algorithm for a single cell.
-        (C) Vertices $\vec{v}_i$ which have been interpolated from the skeletonized cell.
+        (C) Vertices $\vec{x}_i$ which have been interpolated from the skeletonized cell.
         They represent the position of our cell-agents.
@@
 \end{equation}
-Afterwards, we start at one endpoint $\vec{v}_0$ and travel along the polygon defined by the
+Afterwards, we start at one endpoint $\vec{x}_0$ and travel along the polygon defined by the
 bacterial rod.

### 2025-07-29 paper/main.tex
remove todo
rewordings

@@
 %         individual bactera
-%     \item \cite{You2018} TODO
+%     \item \cite{You2018}
 % \end{itemize}
@@
 in order to better understand phenomena such as biofilm formation.
-The major downside of existing computational models is that there are no methods to properly
+The major downside of existing computational models is that there are almost no methods to properly
 quantify their parameters which diminishes their explanatory power.
@@
 Furthermore, we utilize them in our more advanced parameter estimation methods (see
-section~\ref{subsec:comparison-across-generations}) where they enable us to effectively compare
+section~\ref{section:parameter-estimation}) where they enable us to effectively compare
 two bacterial configurations across division events.

### 2025-07-29 paper/main.tex
relabel

@@
 \subsection{Spatial Representation of Rod-Shaped Bacteria}
+\label{subsec:spatial-representation}
 This work makes use of a few but powerfull assumptions concerning the mathematical representation of
@@
 \subsection{Visualization}
-\label{section:generating-masks-and-images}
+\label{subsubsection:visualization}
 
@@
 advantage of a functionality which will be introduced later in
-section~\ref{section:generating-masks-and-images}.
+section~\ref{subsubsection:visualization}.
 Our methodology makes use of the fact that we have already set up our model and solving scheme and

### 2025-07-29 paper/main.tex
completely rewrite parameter estimation section

@@
 \subsection{Parameter Estimation}
+\label{section:parameter-estimation}
 In order to effectively estimate the parameters of our model, we need to set up a cost function
@@
 In doing so, we hope to identify all parameters, given relevant experimental data.
-We use microscopic images and construct a novel algorithm that can extract the position of the agent
-from a provided cell-mask.
-These cell-masks are easily obtainable by either using existing segmentation tools or manually
-annottating images.
-Following this, we can define a cost-function and an optimization scheme to minimize said function.
-
-% --------------------------------------------------------------------------------------------------
-\subsubsection{Challenges of Constructing a Loss Function}
-
-As within any optimization scheme, it can be beneficial to construct custom cost functions
-which are tailored to the problem at hand.
-With the methods that we have already estalished, we need to compare a collection of agents which
-were numerically calcualted through our \ac{abm} modeling approach whith microscopic images.
-Utilizing the extraction algorithm, we can simply compare the numerically obtained positions with
-the extracted positions of the data.
-We choose a typical least-squares approach for the comparison of the individual vertices.
-However, this procedure is only valid for the case where cell-division events do not play a role.
-In the case where we either the numerical model or the experimental data contain more or less cells
-than the other result, we are not able to compare them with this approach.
-
-% --------------------------------------------------------------------------------------------------
-\subsubsection{Relationship to Semantic Segmentation (TODO)}
-\cite{Jadon2020}
-
-\begin{itemize}
-    \item Semantic Segmentation is more general
-    \item Loss functions and some techniques can be reused
-    \item Discuss which types of loss functions we could consider
-    \item Possibly benchmark them later when doing the parameter estimation; compare loss functions
-\end{itemize}
-
-% --------------------------------------------------------------------------------------------------
-\subsubsection{Comparison across Generations}
-\label{subsec:comparison-across-generations}
-
+We utilize the preceding algorithms to extract data from cell-masks, simulate from an initial state
+and then compare both results.
+The cell-masks are easily obtainable by either using existing segmentation
+tools~\cite{Cutler2022,Stringer2020,Hardo2022} or manually annottating microscopic image series.
+
+\paragraph{No Cell Division}
+In the scenario where no cells divide during the observed interval, we can use our extraction
+algorithm to obtain the position of the cells and then compare them directly to the positions
+calculated from the computational results.
+For all cells $j=1..M$, we have a collection of $i=0..N$ vertices $\vec{x}_{i,j}$.
+Thus, the total amount of information is an array of shape $(M,N,2)$
+Assuming that we have $N$ cells, each of which is represented by $M$ vertcies, the total amount of
+information is an array of shape $(N,M,2)$, containing all positions of all agents.
+It is thus only a matter of comparing two of these arrays with each other and many methods and
+approaches which already exist~\cite{Wang2020} can be leveraged.
+For now, we simply calculate the distance of each vertex pair, which is equivalent to the
+well-known loss function
+
+\begin{equation}
+    L_{\text{LS}}(\vec{x}, \vec{y})
+        = \sqrt{\sum\limits_{i,j}||\vec{x}_{i,j}-\vec{y}_{i,j}||^2}
+        = \sqrt{\sum\limits_{i,j,k}|x_{i,j,k}-y_{i,j,k}|^2}.
+\end{equation}
+
+of least squares (LS).
+One important detail is that it must be guaranteed that the ordering of vertices is identical in the
+numerical and experimentally obtained results.
+
+\paragraph{Comparison across Generations}
 In order to compare a variable amount of cells, we can not utilize the positions of these cells
-directly.
-In this approach, we propose a method which is similar to already existing routines (TODO mention
-machine learning, some vision algorithms use this).
-From the numerical results of the \ac{abm}, we can again generate an image of cell-masks.
-We then compare this image together to the experimental cell-mask pixel by pixel with each other.
-Whenever two pixel colors are not matching, a penalty value is assigned.
+for comparison directly since one particular state of our simulation might contain cells which are
+not present in the experimental results (or vice-versa).
+We thus need additional methods which can be used for comparisons across division events.
+To achieve this, we use the numerical results of the \ac{abm} simulation and our visualization
+algorithms to generate cell-masks (see section~\ref{subsubsection:visualization}).
+We then compare the generated masks to the experimentally obtained ones pixel by pixel.
+Whenever two pixel colors are not matching, a loss value is assigned.
 The sum of all those penalties is then the total cost.
@@
 It is very important to take into account the parental relationship between the cells, checking if
-if either one of the cells is a daughter of the other cell.
+either one of the cells is a daughter of the other cell.
+Figure~\ref{fig:mask-difference-metric} (A,B) shows two successive states of a numerical simulation
+with varying amounts of cells.
+We can see that some division events have occurred but other cells have only grown without
+completing a division step.
+The resulting penalties without accounting for the parental relationships are displayed in
+Figure~\ref{fig:mask-difference-metric} (C).
+Black indicates that the pixels are matching and no loss is assigned while a white pixel
+indicates that the original pixels are not matching, thus assigning a loss value.
+Similarly, Figure~\ref{fig:mask-difference-metric} (D) takes the parental relationship into account
+and highlights these in gray.
+It is possible to assign a custom value for the parental loss value.
+If we set this loss value to $0$, this algorithm ensures the contiuity of the loss function in time.
+
+But it is also biologically motivated.
 If a cell has just divided, the resulting two daughter cells will cover precisely the space that the
-preceding agent had taken up.
-Let's assume as an example that one result has more cells than the other
-(Figure~\ref{fig:mask-difference-metric} (A,B)).
-When we compare the pixels of both images, we can relate the given color of the pixel to a certain
-cell.
-Whenever the colors of two distinct cells are in a parental relationship, we reduce the penalty
-which would be assigned or set it to zero.
+mother had taken up.
+Often times, the exact time at which the fission event is over can not be determined accurately,
+thus leading not only to uncertainties but also some fundamental problems.
+Another approach to construct a comparison algorithm across generations would be to use the
+previously explained position comparison algorithm and use it on intervals in which no division
+event occurrs.
+But this attempt would quickly lead to multiple problems.
+How can we deal with the uncertainties of the timing of the division event?
+And the more cells we want to investigate, the more division events will occurr even in a fixed
+amount of time thus greatly decreasing the effective interval at which this alternative algorithm
+could be used.
+Furthermore, what if there are only two images and there is a division event which is occurring
+between them?
+
+It further needs to be noted that the constructed algorithm can also be adapted by users in order to
+provide more complex loss functions (such as considering time-evolvement, etc.).
+Our method can make use of existing machine leanring algorithms since it is only based on the
+principle of comparing two images with each other.
+There already exists a variety of methods on how to assign loss values in these
+circumstances~\cite{Dice1945}.
 
@@
         In the case where we account for the parental relationship, these jumps are mitigated.
-        The penalty was set to $0$ for pixel-comparisons which belong to cells that are in a
+        The loss value was set to $0$ for pixel-comparisons which belong to cells that are in a
         parental relationship.
@@
 
-We can use this procedure to calculate the cost between successive iteration steps
+\paragraph{Testing the Comparison Algorithm with Division Events}
+When utilizing a computational model (see section~\ref{sec:computational-model}), we can generate
+data and use the generated masks to provide a benchmark of our algorithm.
+To do this, we set out to calculate the loss between two successive iteration steps
 $t_n$ with $t_{n+1}$.
-We expect to obtain an exponential growth in the penalty.
-This can be explained by to the fact that the growth of the total area is exponential in time.
-Furthremore, the temporal area difference can be viewed as a derivative but the derivative of an
-exponential is again an exponential.
-Figure~\ref{fig:mask-difference-metric} shows the total penalty for this approach.
+We expect to obtain an exponential growth in the loss.
+This can be explained by the fact that the growth of the total area is exponential in time.
+Furthermore, the temporal area difference can be viewed as a derivative of the total area in time,
+which is again an exponential function.
+Figure~\ref{fig:mask-difference-metric} (E) shows the total loss for this approach.
 We can see that by accounting for the parental relationship, we get the overall desired effect.
-This method allows us to compare results between states that have a different number of cells
-inside as long as their origin stems from the a shared initial state.
+This method allows us to compare results between states that have a different number of cells.
 However, since this algorithm requires us to calculate a cell-mask for each comparison and then also
 process this mask as well as the comparator, it is much more computationally demanding.
-In cases where the simpler approach of comparing positions of the agents is reasonable, this method
+After all, for each comparison step, we need to run a full 3D visualization pipeline.
+In cases where the simpler approach of comparing positions of the agents is possible, this method
 should be preferred.

### 2025-07-29 paper/main.tex
fixup! relabel
adjust tables containing simulation aspects

@@
 \section{A Computational Model for Rod-Shaped Bacteria}
-\label{subsubsection:mechanical-model-rod-shaped-bacteria}
+\label{sec:computational-model}
 
@@
         (2) & Growth &
-            Cells grow exponentially by inserting new material along the circular part of the rod
-            \cite{Robert2014,Takeuchi2005}.\\
-        % (3) & Differentiation &
-        %     \textit{B.subtilis} is known to differentiate into matrix-producing and
-        %     surfactin-producing cells \cite{vanGestel2015,Lpez2010}.\\
+            Cells grow exponentially by inserting new material either along the circular part of the
+            rod or at the tip~\cite{Robert2014,Takeuchi2005}.\\
         (3) & Division &
-            Rod-Shaped bacteria divide in the middle of the rod, thus forming two new
-            agents.\\
+            Rod-Shaped bacteria divide in the middle of the rod into two new agents.\\
             % Bacteria can form multilayers during their growth phase \cite{Duvernoy2018}.\\
@@
         (5) & Adhesion &
-            Bacteria adhere to each other at longer distances and attach when in close contact
-            \cite{Verwey1947,Trejo2013}.\\
-            % The interaction between rods can be polarized \cite{Duvernoy2018}.\\
-        % (7) & Friction &
-        %     Friction between cells \cite{Grant2014} can be asymmetrical \cite{Doumic2020}.\\
-        % &\textbf{(DC) Domain-Cell Interactions}\\
-        % \midrule
-        % (6) & External Forces &
-        %     Bacteria which rub against surfaces tend to stick to them after some
-        %     time~\cite{vanLoosdrecht1989}.
-        %     The extracellular gel which surrounds the bacteria exerts a force onto the cells
-        %     \cite{Grant2014}.\\
-        % (9) & Extracellular Reactions &
-        %     Bacteria can take up nutrients or possible secrete/take up signalling molecules
-        %     \cite{Li2025}.\\
+            Bacteria attract at longer distances and adhere to each other at shorter
+            distances~\cite{Verwey1947,Trejo2013}.\\
         \bottomrule
@@
     \label{table:simulation-aspects}
-    \caption{TODO Insert caption}
+    \caption{
+        This table lists aspects of cellular behaviour which we consider in our computational model.
+        It is split into cellular aspects (C) which concern only an individual bacterium and
+        interactions between cells (CC).
+        These aspects are generalized in order to fit to as many types of rod-shaped bacteria as
+        possible.
+        These are the assumptions for the basic model that we are presenting here.
+        In order for our methods to work, we are really only required to satisfy the assumption of
+        our spatial representation (see section~\ref{subsec:spatial-representation}) and thus, these
+        assumptions can also be loosened if required.
+    }
 \end{table}
@@
         (2) & Growth &
-            Cells grow exponentially by inserting new material along the circular part of the rod
-            \cite{Robert2014,Takeuchi2005}.\\
+            Cells grow exponentially by inserting new material either along the circular part of the
+            rod or at the tip~\cite{Robert2014,Takeuchi2005}.\\
         (3) & Differentiation &
@@
     \label{table:simulation-aspects-supplement}
-    \caption{TODO Insert caption}
+    \caption{
+        This table extends table~\ref{table:simulation-aspects} with additional observations.
+        It has to be noted that this table includes more specific behaviour that might not be
+        generalizable across all types of rod-shaped bacteria.
+    }
 \end{table}

### 2025-07-29 paper/main.tex
include previously unused citation

@@
 force.
-In order to construct a model which can be parametrized and by method of Occam's Razor, we chose to
-model the bacterium as a bending beam, thus utilizing the simplest solution which can still produce
-the desired effects.
+In principle, this stress will be dependent on the position inside the cell~\cite{Chatterjee1988}
+but in order to construct a model which can be parametrized and by method of Occam's Razor, we chose
+to model the bacterium as a bending beam, thus utilizing the simplest solution which can still
+produce the desired effects.
 Such a beam can be represented by a collection of vertices $\vec{x}_i$ which are connected by a
@@
 recurring collisions with small particles of the surrounding medium or other minor fluctuations.
-
-%TODO use these citations
-\cite{Chatterjee1988},\cite{Ursell2014}
 

### 2025-07-29 paper/main.tex
relabel (sub)sections

@@
 \subsection{Visualization}
-\label{subsubsection:visualization}
+\label{subsection:visualization}
 
@@
 \subsection{Data Extraction}
-\label{section:fitting-and-comparing-masks}
+\label{section:data-extraction}
 
@@
 advantage of a functionality which will be introduced later in
-section~\ref{subsubsection:visualization}.
+section~\ref{subsection:visualization}.
 Our methodology makes use of the fact that we have already set up our model and solving scheme and
@@
 To achieve this, we use the numerical results of the \ac{abm} simulation and our visualization
-algorithms to generate cell-masks (see section~\ref{subsubsection:visualization}).
+algorithms to generate cell-masks (see section~\ref{subsection:visualization}).
 We then compare the generated masks to the experimentally obtained ones pixel by pixel.
@@
 % --------------------------------------------------------------------------------------------------
-\subsection{With Division}
+\subsection{Without Division}
 We will now focus on using the presented methods for parameter estimation via microscopic images.
@@
 \end{figure}
+
+% --------------------------------------------------------------------------------------------------
+\subsection{With Division (TODO)}
 

### 2025-07-29 paper/main.tex
rename to individual parameters

@@
             % Bacteria can form multilayers during their growth phase \cite{Duvernoy2018}.\\
-        (4) & Variable Parameters &
+        (4) & Individual Parameters &
             Parameters for individual cells are not fixed values but rather taken from a

### 2025-07-29 paper/main.tex
move section to supplement
write about individual treatment of cells

@@
 \section{Application to \textit{E.Coli.} growing on Agar (TODO)}
-% --------------------------------------------------------------------------------------------------
-\subsection{Data Preparation}
-TODO: Decide if this should be Supplement
-\begin{enumerate}
-    \item Have masks of data; reference image segmantation tools; or do it by hand
-    \item ensure that colors are matching for bacteria between time-steps
-    \item clean borders of images if needed
-    \item know domain size; otherwise use pixels
-\end{enumerate}
-
-% --------------------------------------------------------------------------------------------------
-\subsection{Non-Individual Treatment}
+
+With the previous works of extracting the positions (see section~\ref{section:data-extraction}) for
+individual cell-agents and the methods to estiamte parameters of our agents (see section~\ref{section:parameter-estimation}), we are now ready to apply them to showcase their usefulness.
+
+TODO explain what data we are using
+
 % --------------------------------------------------------------------------------------------------
 \subsection{Individual Treatment}
+We have already suspected (see table~\ref{table:simulation-aspects}) that in order to properly
+estimate parameters of our system, we need to treat every cell-agent individually, thus assigning
+parameters individually to them.
+This means that we increase the total number of parameters that needs to be sampled within an
+optimization routine.
+Given $N_c$ cells and $N_p$ parameters per cell and $N_u$ unique parameters (where $N_u\leq N_p$),
+the total amount of parameters is
+
+\begin{equation}
+    N_{p,\text{total}} = N_u + N_c (N_p - N_u) = N_C N_p - (N_c - 1) N_u.
+\end{equation}
+
+It is clear, that we should avoid to needlessly assign every parameter individually since the space
+which is needed to be sampled scales exponentially with the dimension ie. the number of parameters.
+However, we will now present data that indicate that an individual treatment is necessary.
+Figure~\ref{fig:estimated-growth-rates} shows the rod-lengths of individual bacteria which have been
+calculated from the extracted positions.
+We fit a delayed exponential function to them which is defined by
+
+\begin{equation}
+    x(t) =
+    \left\{\begin{array}{ll}
+            x_0 \exp(\lambda (t-t_0)) & t\geq t_0\\
+            x_0 & \text{ else}
+    \end{array}\right.
+\end{equation}
+
+and where $\lambda$ is the growth rate $t_0$ the delay.
+The fits to invidiual growth-curves of figure~\ref{fig:estimated-growth-rates} (A) result in good
+matches and so does the fit of subfigure (B).
+However, the latter exhibits large amounts of uncertainty within the data which is also illustrated
+in subfigure (C).
+There, we can see much more clearly the difference between considering invidiually assigned
+parameters and global ones.
+It is clear from this observation that in order for us to accurately estimate the parameters of this
+system, we need to treat the bacteria individually.
+
+The individual treatment of the bacteria allows us in principle to estimate the distribution of the
+parameters directly without having to resort to any assumptions on its form.
+This is true for any parameter, not just the growth rates, since we aim to estimate all mechanical
+parameters of our cellular model.
+
+\textbf{TODO is this discussion?}\\
+The total parameter space $P$ is a product of all single spaces $P_i$ in which the parameters
+$p_i$ live $P=P_0\times P_1\times\dots$ but some of these parameters share the same space (let's
+assume $p_0,p_1\in \tilde{P}$).
+This leaves the open question as to weather the optimization algorithm could be modified to leverage
+this knowledge and obtain increased performance.
+For instance, it might be valuable to initially consider individual parameters as globally defined
+and start treating the individually after a certain amount of optimization steps.
+
+% \begin{enumerate}
+%     \item motivate individual treatment of cells
+%     \item pro: get at distributions of parameters; even more than growth rates
+%     \item pro: probably good idea to do this in this case, since more or less required (see figure)
+%     \item con: computationally more demanding
+%     \item can we improve the optimization algorithm to somehow "know" that some parameters are for
+%         different cells and thus do not "interact" as other parameters with each other?
+%         ie. they are part of the same subspace.
+% \end{enumerate}
+
 \label{subsec:parameter-estimation-individual-treatment}
@@
     \centering
-    \includegraphics[width=0.33\textwidth]
+    \begin{tikzonimage}[width=0.33\textwidth]
         {figures/crm_estimate_params/IWF-Goettingen/rod-lengths-individual.pdf}%
-    \includegraphics[width=0.33\textwidth]
+        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.33\textwidth]
         {figures/crm_estimate_params/IWF-Goettingen/rod-lengths-average.pdf}%
-    \includegraphics[width=0.33\textwidth]
+        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.33\textwidth]
         {figures/crm_estimate_params/IWF-Goettingen/parameter-distribution.pdf}%
-    \caption{}
+        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+    \end{tikzonimage}%
+    \caption{
+        % TODO use same y-axis across the plots
+        (A) The subfigure shows the time-evolution of the rod-lengths for every individual
+        bacterium.
+        These growth curves are fitted individually with delayed exponential function.
+        (B) The datapoints are created by averaging the rod-lengths shown in subfigure (A) and then
+        fitted with an exponential function.
+        The uncertainty band is the standard deviation of the averaged datapoints.
+        (C) Blue dots show the combination of estimated growth rates and delays for individual
+        bacteria while the orange dot corresponds to the average.
+        Ellipses around the datapoints are taken from the uncertainty of the fits.
+        Time is given in units of frames of the movie from which the data have been extracted.
+    }
     \label{fig:estimated-growth-rates}
 \end{figure}
+
 % --------------------------------------------------------------------------------------------------
@@
 \renewcommand{\thesection}{S\arabic{section}}
+
+%###################################################################################################
+\section{Data Preparation}
+\begin{enumerate}
+    \item Have masks of data; reference image segmantation tools; or do it by hand
+    \item ensure that colors are matching for bacteria between time-steps
+    \item clean borders of images if needed
+    \item know domain size; otherwise use pixels
+\end{enumerate}
 

### 2025-07-29 paper/main.tex
remove cost to loss

@@
 \label{section:parameter-estimation}
-In order to effectively estimate the parameters of our model, we need to set up a cost function
+In order to effectively estimate the parameters of our model, we need to set up a loss function
 which can be minimized.
@@
 Whenever two pixel colors are not matching, a loss value is assigned.
-The sum of all those penalties is then the total cost.
+The sum of all those penalties is then the total loss.
 This algorithm is an implementation of a signed area measure.
@@
         (C) "Naive" comparison of cell masks, which identifies pixels of non-matching colors and
-        assigns a cost by counting these pixels.
+        assigns a loss by counting these pixels.
         This simple approach does not take into consideration the parental relationship.
@@
         colored in gray.\\
-        (E) Comparison of cost functions, exponential regression (ER) and number of cells.
-        % Total cost between successive iteration steps $t_n,t_{n+1}$.
-        The overall shape of the cost is expected to have the slope of an exponential function.
+        (E) Comparison of loss functions, exponential regression (ER) and number of cells.
+        % Total loss between successive iteration steps $t_n,t_{n+1}$.
+        The overall shape of the loss is expected to have the slope of an exponential function.
         We can clearly see that division events introduce jumps within this relationship.
@@
 
-Figure~\ref{fig:parameter-estimates-single-step} shows the cost function of multiple simulations.
+Figure~\ref{fig:parameter-estimates-single-step} shows the loss function of multiple simulations.
 The curves show the value at the found optimium and how it changes when varying the value of the
@@
     \caption{
-        The curves show the value of the cost function at the optimium as determined by our
+        The curves show the value of the loss function at the optimium as determined by our
         otimization scheme (blue dot).
-        We have chosen a single parameter and calculated the resulting cost function across a range
+        We have chosen a single parameter and calculated the resulting loss function across a range
         of values and plotted it.

### 2025-07-29 paper/main.tex
finish chapter on param estimation without divisions

@@
 \subsection{Without Division}
-We will now focus on using the presented methods for parameter estimation via microscopic images.
-Our initial state is given by TODO~\cite{}.% TODO
-We use our extraction algorithm to obtain the initial positions of our cells.
-From there, we supply the model with parameters such that we can make a prediction for the final
-positions of the bacteria.
-We chose a time-interval which does not contain any division events.
-By doing so, we can utilize the simpler and more performant position comparison scheme for our cost
+For now, we will focus on a time-interval which does not contain any division events.
+By doing so, we can utilize the simpler and more performant position comparison scheme for our loss
 function.
-Another important detail is the individuality of the agents.
-In our case, this means that each agent might have different parameters associated to it.
 For this dataset, we assumed that all interaction parameters which specify the interaction
 potential are identical across all agents but growth rates and the physical radii of the cells are
-individual and we also optimize them.
+individual and we optimize them like any other global parameter.
 We chose to optimize for 3 interaction parameters of the Mie potential $n,m,V_0$ and two parameters
@@
 This leads to a total of $4+2\times 6=16$ parameters.
-To estimate these parameters we used a latin-hypercube sweep across the parameter space and
-performed a local minimization with the Nelder-Mead method (TODO citations).% CITATIONS
-
-Figure~\ref{fig:parameter-estimates-single-step} shows the loss function of multiple simulations.
-The curves show the value at the found optimium and how it changes when varying the value of the
-parameter in question.
-We have exemplified the values for radii $r_i$ and growth rates $\mu_i$ with values of the first
-cell.
-The remaining plots can be found in the supplementary material in
-section~\ref{sec:supplement-parameter-profiles}.
-
-This depiction is similar to the method of profile likelihood~\cite{Raue2009}.
+To estimate these parameters we used a latin-hypercube~\cite{McKay1979} sweep across the parameter
+space and performed a local minimization with the Nelder-Mead~\cite{Powell1973} method.
+
+In figure~\ref{fig:parameter-estimates-single-step} we performed a sensitivity-analysis for every
+parameter.
 We can see that almost all parameters can be estimated by our optimization approach.
-However, the damping constant exhibits the well-known phenomenon of
-unidentifiability~\cite{Raue2009}.
-This is of course problematic since the real value of $\lambda$ can not be idenified in this case.
-But it also shows that the approach which we used is similar in its ability to detect these defects
-in the first place.
+However, the damping constant (A) does not show any changes in the loss function when increasing its
+value.
+It hints at a non-identifiability~\cite{Raue2009} of the parameter $\lambda$.
+But it also shows that the approach which we used is able to detect these problems.
 As we continue to develop these methods, we expect to be able to reuse many of the existing
-theoretical approaches which have already been developed for less complex systems such as \acp{ode}.
+theoretical approaches such as likelihood estimation and uncertainty quantification which have
+already been developed extensively for less complex systems such as \acp{ode}.
+\textbf{TODO citations, also: is this discussion?}
 Fortunately, \texttt{cellular\_raza} was chosen to construct our model such that we are flexible in
@@
     \centering
-    \begin{tikzonimage}[width=0.33\textwidth]
-        {figures/crm_fit/snapshots/exact-004444.png}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \begin{tikzonimage}[width=0.54\textwidth]
+        {figures/crm_fit/snapshots/predicted-031111.png}%
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
     \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.33\textwidth]
-        {figures/crm_fit/snapshots/exact-004444.png}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \hspace{0.01\textwidth}%
+    \begin{tikzonimage}[width=0.41\textwidth]
+        {figures/crm_fit/celldifss/cell-031111.png}%
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
     \end{tikzonimage}%
+    \caption{
+        \textbf{TODO do we need this figure?}
+        (A) A microscopic image with the predicted positions of the cells overlayed on top.
+        (B) Differences between extracted and predicted positions.
+        % TODO make this figure to scale
+    }
+    \label{fig:celldiffs-individual}
 \end{figure}
@@
     \caption{
-        The curves show the value of the loss function at the optimium as determined by our
-        otimization scheme (blue dot).
-        We have chosen a single parameter and calculated the resulting loss function across a range
-        of values and plotted it.
+        Sensitivity analyses werer performed for every optimized parameter of our model.
+        (A-F) show the curves of loss functions when varying this respective parameter and keeping
+        all others constant at the value of the found optimum (blue dot).
         In the cases of (B-F), we can see that a minimum was achieved and thus the parameters could
-        be identified.
-        However, for the case of the damping constant $\lambda$ in (A), a non-identifiability
-        arises.
+        be estimated.
+        However, for the case of the damping constant $\lambda$ in (A), we see that the upper bound
+        is open.
+        This suggests that our model could be reduced with regard to this parameter.
+        We have exemplified the values for radii $r_i$ and growth rates $\mu_i$ with values of the
+        first cell.
+        The remaining plots can be found in the supplementary material in
+        section~\ref{sec:supplement-parameter-profiles}.
     }

### 2025-08-29 paper/main.tex
add idea to discussion
remove todo in heading
add more notes for introduction

@@
 \section{Introduction}
-\label{section:introduction}
+
+General comments on \acp{abm} and their problems
+\begin{itemize}
+    \item \aclp{abm} are used for many purposes
+    \item describe current way of estimating parameters; compare distributions to each other; no idea if
+          individual parameters are correct
+    \item challenge: estimate their parameters
+    \item challenge: construct minimal models (requires flexibility)
+\end{itemize}
+
+General remarks about the steps in this work
+\begin{itemize}
+    \item Use rod-shaped bacteria as model system
+    \item Clear separation between different components (link to overview)
+    \item This separation allows us to generalize the type of approach and apply to other models
+    \item think of as a "blueprint" for estimating parameters for \acp{abm}
+\end{itemize}
+
+Experimental observations: rod-shaped bacteria
+\begin{itemize}
+    \item growth + division "distinct mechanical properties", MreB
+    \item bendiness (not much studied)
+    \item individual parameters (i.e. growth rates)
+    \item adhesion to each other
+\end{itemize}
+
+Computational studies of rod-shaped bacteria
+\begin{itemize}
+    \item not so many models for rod-shaped bacteria
+    \item could include more effects such as bending
+    \item almost no efforts to estimate parameters with individual-based models
+    \item although many things known no "generalized mechanical model" constructed for rod-shaped bacteria
+\end{itemize}
+
+\section{Introduction (old)}
 
@@
 %###################################################################################################
-\section{Application to \textit{E.Coli.} growing on Agar (TODO)}
+\section{Application to \textit{E.Coli.} growing on Agar}
 
@@
     \item Constructed Model which already combines many effects; extend this even further
+    \item can now use classical methods: sensitivity-analysis, likelihood estimators and uncertainty quantification
     \item Proof of concept: Able to fit individual parameters of the agents; can be extend this?

### 2025-08-31 paper/main.tex
formulate first part of the introduction

@@
 \section{Introduction}
+
+\acp{abm} have proven to be powerfull tools in the study of complex cellular systems.
+They have been applied in various scenarios, such as cancer
+research~\cite{Ghaffarizadeh2018,Cooper2020}, Immuno-Oncology~\cite{Karolak2021} or the modeling
+of mechanical intercellular connections in plant cells.
+By treating cells as individual entities, these tools allow researchers to think in terms of
+cellular properties and processes thus allowing for a better understanding of the biological
+reality.
+
+Despite their popularity, \acp{abm} face multiple challenges in practise.
+Constructing a fully working \ac{abm} simulation from scratch is an intricate process that requires
+a large set of skills and time.
+It is therefore common practise to reuse existing simulation tools, which brings along a number of
+other problems.
+Even if the existing tool is designed to be generalizable, it often comes with a pre-defined set
+of parameters that need to be specified.
+Furthermore, as we have discussed in our previous work~\cite{Pleyer2023}, these tools were most
+often not designed to be flexible but rather for specific purposes thus limiting their
+generalizability.
+% TODO citations
+Typically researchers aim to estimate every individual parameter of their model.
+This is achieved using methods such as the profile-likelihood and model reduction techniques.
+% TODO citaitons
+When it comes to \acp{abm} however, it is often only shown that the results which have been
+generated by an \ac{abm} lead to the phenomenon which aligns with the observed biological reality.
+% TODO citations
+This reduces the confidence in the constructed models and raises concerns with respect to the
+validity of the obtained results.
 

### 2025-08-31 paper/introduction_old.tex
move old introduction to separate file

--- 2025-08-29 paper/main.tex
@@
-\documentclass{article}
-
-\usepackage{arxiv}
-
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage[hidelinks]{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsmath,amssymb,amsthm}
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{float}
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{mathrsfs}
-\usepackage{onimage}
-\usepackage{graphicx}
-\usepackage{doi}
-\usepackage{acronym}
-\usepackage{listings}
-\usepackage[mode=buildnew]{standalone}
-\usepackage{tikz}
-\usepackage{tabularx}
-\usepackage{siunitx}
-\usepackage{cite}
-\usepackage{acronym}
-\usepackage{xcolor}
-
-\input{figures/tikz-preamble.tex}
-
-\title{Methods for Analyzing Models of\\ Rod-Shaped Bacteria}
-
-%\date{September 9, 1985}	% Here you can change the date presented in the paper title
-%\date{} 					% Or removing it
-
-\author{
-    \href{https://orcid.org/0009-0001-0613-7978}{
-        \includegraphics[scale=0.06]{figures/orcid.pdf}
-        \hspace{1mm}Jonas Pleyer
-    }
-    \thanks{
-        \href{https://jonas.pleyer.org}{jonas.pleyer.org},
-        \href{https://cellular-raza.com}{cellular-raza.com}
-    }\\
-	Freiburg Center for Data-Analysis and Modeling\\
-	University of Freiburg\\
-	\texttt{jonas.pleyer@fdm.uni-freiburg.de} \\
-	%% examples of more authors
-	\And
-	\href{https://orcid.org/0000-0002-6371-4495}{
-        \includegraphics[scale=0.06]{figures/orcid.pdf}
-        \hspace{1mm}Christian Fleck
-    }\\
-	Freiburg Center for Data-Analysis and Modeling\\
-	University of Freiburg
-}
-
-\newacro{abm}[ABM]{Agent-Based Model}
-\newacro{ode}[ODE]{Ordinary Differential Equation}
-\newacro{pbr}[PBR]{Physically-Based Rendering}
-
-% Uncomment to remove the date
-%\date{}
-
-% Uncomment to override  the `A preprint' in the header
-\renewcommand{\headeright}{Preprint}
-%\renewcommand{\undertitle}{Technical Report}
-\renewcommand{\shorttitle}{Methods for Analyzing Models of Rod-Shaped Bacteria}
-
-\usepackage{enumitem}
-\setlist{nolistsep}
-
-%%% Add PDF metadata to help others organize their library
-%%% Once the PDF is generated, you can check the metadata with
-%%% $ pdfinfo template.pdf
-\hypersetup{
-pdftitle={% TODO
-},
-pdfsubject={q-bio.NC, q-bio.QM},
-pdfauthor={Jonas Pleyer, Christian Fleck},
-pdfkeywords={},
-}
-
-% Change numbering of equations
-% \numberwithin{equation}{section}
-
-% MAKE TITLES IN THEOREMS BOLD
-\makeatletter
-\def\th@plain{%
-  \thm@notefont{}% same as heading font
-  \itshape % body font
-}
-\def\th@definition{%
-  \thm@notefont{}% same as heading font
-  \normalfont % body font
-}
-\makeatother
-
-\begin{document}
-\maketitle
-
-%###################################################################################################
-\begin{abstract}
-    \textbf{TODO: ABSTRACT}
-    % This paper establishes the basis for a novel individual-based mechanical model of rod-shaped
-    % bacteria.
-    % It describes methods of how to obtain data from microscopic images such that they can be used
-    % for parameter estimation purposes.
-    % Furthermore, we show initial results for parameter estimates and an application of the model to
-    % the formation of multilayers within bacterial colonies.
-\end{abstract}
-
-% keywords can be removed
-% \keywords{}
-
-\pagebreak
-\renewcommand{\contentsname}{Table of Contents (remove before submission)}
-\tableofcontents
-\pagebreak
-
-%###################################################################################################
-\section{Introduction}
-
 General comments on \acp{abm} and their problems
@@
 particular problem.
-
-%###################################################################################################
-\section{Overview}
-This work consists of many subcomponents which can be combined in order to achieve various tasks.
-Figure~\ref{fig:flowchart-project-structure} shows a high-level overview of all components.
-We assume that bacterial rods can be represented by a colleciton of vertices $\vec{x}_i$.
-This assumption alone is enough to enable a cascade of other components.
-Using the spatial representation, we provide visualization methods that create 3D models from the
-given vertices which can then either be rendered or projected to a 2D plane.
-The data extraction component determines the vertices $\vec{x}_i$ from given cell masks which
-correspond to microscopic images.
-This allows us in particular to turn any sequence of images (movie) into a sequence of bacterial
-positions.
-The dynamics of the vertices $\vec{x}_i$ are determined by a computational model which is specific
-to the respective use-case and should be defined by the researcher.
-It merely needs to utilize the aforementioned spatial representation of bacterial rods as a
-collection of vertices $\vec{x}_i$.
-This work will provide one example of how such a model can be defined.
-Notably, the visualization and data extraction components do not require any knowledge about the
-computational model, thus allowing us to reuse their functionality across multiple models.
-
-Constructing a computational model requires knowledge about the target system and thus allows us to
-interpret the results gathered by the provided data extraction methods.
-These results can then further be used to restrict or determine the parameters of our model.
-For example, by extracing the positions of bacteria across a time-series of microscopic images, we
-can also calculate their lengths at these time-points and thus determine their rate of growth and
-type of growth (see section~\ref{subsec:parameter-estimation-individual-treatment}), thus
-effectively guiding our modeling steps.
-
-Visualization methods are an integral part of any modeling endevaour.
-They allow researchers to obtain feedback quickly and improve iteration times when constructing new
-models.
-Furthermore, we utilize them in our more advanced parameter estimation methods (see
-section~\ref{section:parameter-estimation}) where they enable us to effectively compare
-two bacterial configurations across division events.
-In the future, we are planning to build on the current state and generate realistic microscopic
-images and time-series.
-Together with the parameter estimation methods, they can then be used to provide large amounts of
-mechanistic data for the training cell-segmentation and cell-tracking machine-learning algorithms.
-
-% \begin{itemize}
-%     \item This publication contains 5 things:
-%     \begin{enumerate}
-%         \item Methods of how to obtain data from microscopic images and perform parameter
-%             estimations; $\rightarrow$ also possibly across multiple generations
-%         \item Showcase the usefulness of the methods and the model by applying it to parameter
-%             estimation via microscopic images and multilayer formation
-%         \item Initial (incomplete) model of rod-shaped bacteria
-%         \item Generate 3D Images (masks)
-%         \item A python package which can do all of that and can be reused and extended
-%     \end{enumerate}
-%     \item Overall structure
-%     \begin{enumerate}
-%         \item users bring model + data
-%         \item this paper provides methods to do the parameter estimation + visualization
-%     \end{enumerate}
-%     \item For complete model, more work has to be done
-%     \begin{itemize}
-%         \item More aspects (see Review; put in discussion)
-%         \item Note that we only care about \textit{E.Coli} and \textit{B.Subtilits}! This is
-%             important.
-%     \end{itemize}
-%     \begin{itemize}
-%         \item State of the Art
-%     \end{itemize}
-% \end{itemize}
-
-\begin{figure}[H]
-    \centering
-    % \includegraphics[width=0.8\textwidth]{overview.pdf}
-    \includestandalone[width=0.8\textwidth]{figures/concept/overview}
-    \caption{
-        The project can be divided into subcomponents (yellow, red) that interact with and depend on
-        each other.
-        Yellow blocks are provided by this work and red blocks require additional input such as
-        modeling assumptions or data.
-        Blue blocks indicate possible future extensions of this project.
-        Every component implicitly builds on the spatial representation of rod-shaped bacteria as
-        vertices $\vec{x}_i$.
-    }
-    \label{fig:flowchart-project-structure}
-\end{figure}
-
-%###################################################################################################
-\section{Methodology}
-% --------------------------------------------------------------------------------------------------
-\subsection{Spatial Representation of Rod-Shaped Bacteria}
-\label{subsec:spatial-representation}
-This work makes use of a few but powerfull assumptions concerning the mathematical representation of
-the cells in question.
-In order to describe the spatial dynamics of individual agents, we need to define their position,
-shape and how they interact with each other.
-In order to provide as much flexibility as possible when constructing new models, we restrict
-ourselves to assumptions concerning the position such that researchers can freely experiment with
-various shapes and interaction types.
-Furthermore, it is very important to strike a good balance between the restrictiveness of our
-assumptions and generality in order to provide meaningful results and still be able to apply our
-approach to a broader class of models.
-Taking into account the natural shape of rod-like bacteria, it is clear that they can not be
-represented with a singular point $\vec{x}$ in space.
-We could extend this positional definition with a vector $\vec{w}$, indicating the direction of
-extension but this would restrict ourselves to fully straight rod-shaped bacteria.
-However, multiple experiments have shown that MReB is responsible for the growth and maintenance of
-the elongated shape of bacteria~\cite{Erickson2001}.
-By means of this protein, rod-shaped bacteria can bend elastically on short time-scales and
-deform plastically over longer periods of time.
-Our goal is to provide a model which can encapsulate these effects.
-Thus, we extend our model from a single point to multiple vertices which are part of a 1-dimensional
-polygon with two endpoints $\vec{x}_i$.
-In essence, we discretize the (possibly non-straight) rod of the agent.
-Figure~\ref{fig:mechanics-bacterium} shows a visualization of this approach.
-The angles $\alpha_i$ between vertices $\vec{x}_{i-1},\vec{x}_i$ can be used when modeling the
-dynamics of these bacteria.
-
-With these fundamental thoughts layed out, we retain the flexibility to represent many models such
-as varying shapes since we do not have to make any assumptions on the type of interactions between
-two bacteria.
-Even concerning the shape of bacteria, this assumption still allows us to mdoel elliptical cells by
-modifying the interaction type as well as other effects such as polarity of
-interactions~\cite{Duvernoy2018}.
-Furthermore, we also did not have to assume any particular mechanical dynamics of the agents, only
-their spatial representation, thus leaving in particular room for plastical or elastic deformations
-of the rod and other more intricate mechanisms.
-
-% \begin{enumerate}
-%     \item represent bacteria as polygon with vertices $\vec{x}_i$
-%     \item Assumes that these vertices are not fundamental but rather a discretization of the actual
-%         problem
-% \end{enumerate}
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.5\textwidth]{docs/source/_static/mechanics.png}
-    \caption{
-        Example of an elonged bacterial rod represented by a collection of vertices which are
-        connected by springs.
-        The angles $\alpha_i$ have been exaggerated such that they can be identified more easily.
-        A bacterial agent which is not interacting with any external objects will tend to the
-        equilibrium state of a straight rod with $0$ curvature.
-    }
-    \label{fig:mechanics-bacterium}
-\end{figure}
-
-% --------------------------------------------------------------------------------------------------
-\subsection{Visualization}
-\label{subsection:visualization}
-
-For visualization we use pyvista, which provides high-level wrappers around the popular VTK
-library~\cite{vtkBook,Sullivan2019}.
-The first stage of our visualization is to generate 3D meshes for each bacterium, which we consider
-an intermediate representation.
-We insert spheres of a given radius $r$ at each vertex $\vec{x}_i$ and connect adjacent vertices
-$\vec{x}_i,\vec{x}_{i+1}$ with a cylinder, also of radius $r$ which is positioned such that its core
-is identical to the connecting line segment between the two vertices.
-Finally, we combine the meshes of all of these objects to obtain a single mesh of one bacterium.
-A render of this result can be seen in figure~\ref{fig:progression-image-generation}.
-Since we use a 3D rendering engine as an intermediate representation, the resulting image can have
-overlaps of bacteria which are placed at higher or lower values along the z-direction.
-
-To generate masks, we assign unique colors to every mesh of the respective cellular agent.
-This mapping is unique, ensuring no two agents will use identical colors.
-We then use a parallel projection and remove any existing light sources in order to obtain a solid
-color for every cell-agent.
-The resulting image is shown in subfigure (B).
-Subfigure (C) shows an example of a simplified way to generate microscopic images.
-The properties of the meshes such as roughness and reflectivity have been adjusted and we use
-physically-based rendering of VTK.
-The result of this render is augmented by introducing white noise on the pixel values at each point.
-This approach is very rudimentary and we hope to extend these methods in the future.
-In order to properly generate data, we need to deal with microscopic imaging techniques and possible
-defects that can occurr.
-It is crucial to get this part of the image-generation process right in order to obtain
-near-realistic images that can be further reused.
-
-% \begin{itemize}
-%     \item use pyvista (VTK) for Visualization
-%     \item always do 3D render. This way, we can have overalp between agents and produce more
-%         realistic masks.
-%     \item Assign unique color values for idividual agents.
-%         We have a 1:1 mapping between agent identifier and color! (for each simulation different
-%         though)
-%     \item use parallel projection; this should be most close to what microscopes measure due to the
-%         relatively large distance between object and lens.
-%     \item use some noise to smear out the rods; this is already similar to microscopic images; use
-%         more like this and deal with real-world defects to make it more realistic (future, not now)
-%     \item Figure~\ref{fig:progression-image-generation} shows steps in generating image.
-%     \item Reflectivity, placement of light etc. can already introduce some heterogeneity in the
-%         visual of the rod.
-%     \item Plan to implement noise on the mesh itself; make it have rough surface (like real rods as
-%         seen under electron microscopy)
-% \end{itemize}
-%
-% \paragraph{Algorithm}
-% \begin{enumerate}
-%     \item Get position of cell
-%     \item Insert spheres at each vertex
-%     \item Insert cylinder along edges
-%     \item Combine individual objects into one big mesh
-%     \item Use \ac{pbr}
-% \end{enumerate}
-
-\begin{figure}[H]
-    \centering
-    % \begin{tikzonimage}[width=0.9\textwidth]{visualization.pdf}
-    %     \node at (0.025, 0.975)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{A}};
-    % \end{tikzonimage}
-    \begin{tikzonimage}[width=0.3\textwidth]
-        {docs/source/_static/09395645494836445480/raw_pv/000000400.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.3\textwidth]
-        {docs/source/_static/09395645494836445480/masks/000000400.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.3\textwidth]
-        {docs/source/_static/09395645494836445480/images/000000400.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
-    \end{tikzonimage}
-    \caption{
-        (A) shows the result from combining sphere and cylinder meshes to obtain the shape of a
-        bacterial rod.
-        This render contains lighting coming from 2 light sources which thus produces a small glow
-        on the upper part of the meshes.
-        (B) Unique color values are assigned to each agent and light sources are removed.
-        Afterwards, the image is rendered by projecting along the z-axis.
-        (C) We use physically-based rendering and assign properties such as roughness and
-        reflectivity to generate an initial render.
-        Afterwards, we apply white noise to the pixel values of image.
-        In the future, we hope to extend this approach in order to construct near-realistic
-        microscopic images from synthetic data.
-    }
-    \label{fig:progression-image-generation}
-\end{figure}
-
-% --------------------------------------------------------------------------------------------------
-\subsection{Data Extraction}
-\label{section:data-extraction}
-
-This section will explain how we can determine the collection of vertices $\vec{x}_i$ introduced in
-subsection~\ref{subsection:mechanical-model-mechanics} from microscopic images
-(Figure~\ref{fig:position-extraction-algorithm} (A)).
-They will serve as initial, final and intermediate values with which we can compare our predictions.
-The algorithms which we devised requires the input of cell-masks which assign a unique color for
-every cell in the image.
-Such masks can be generated by cell-segmentation tools or obtained by annottating microscopic images
-by hand (Figure~\ref{fig:position-extraction-algorithm} (B)).
-Afterwards, we proceed with the following 5 steps.
-\begin{enumerate}
-    \item Obtain submasks for each cell individually.
-    \item Skeletonize submask with algorithm from Lee~\cite{Lee1994}
-    \item Determine endpoints $\vec{q}_0,\vec{q}_1$.
-        Assert that there are only 2 endpoints.
-    \item Sort points such that they form a polygon.
-    \item Interpolate vertices $\vec{x}_i$ from obtained polygon.
-\end{enumerate}
-In the first step, the image will be split into a collection of subimages that only contain
-information for one single cell.
-From there, we apply a skeletonization algorithm~\cite{Lee1994}
-(Figure~\ref{fig:position-extraction-algorithm} (C)).
-It has to be noted that there are multiple variants of skeletonization and in general, these classes
-of algorithms can produce results which have intersection points and multiple endpoints.
-This described behaviour is undesirable in our case.
-We have found that for most practical examples, the algorithm devised by~\cite{Lee1994} has worked
-well.\\
-The skeletonization algorithm supplies us with a collection of unordered lattice points $\vec{p}_i$
-(pixel coordinates) where the skeleton lives.
-We identify the endpoints $\vec{q}_0,\vec{q}_1$ by counting the number of neighbors in the von
-Neumann neighborhood for each lattice point.
-Any point with exactly one neighboring point must be and endpoint.
-If we encounter more or less than two points which fit this criterion, the routine can not commence
-and is aborted.
-Next, we start with one of the endpoints and determine the next neighbor which was not yet picked.
-We list the points in this order, thus forming a sorted polygon of pixel coordinates.\\
-In the last step, we take the sorted, extracted points $\vec{p}_i$ and interpolate $n$ values for
-the desired number of vertices $\vec{x}_i$.
-
-\begin{figure}[H]
-    \centering
-    \begin{tikzonimage}[width=0.36\textwidth]
-        {docs/source/_static/fitting-methods/algorithm/image001042.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.36\textwidth]
-        {docs/source/_static/fitting-methods/algorithm/mask-zoom.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.255\textwidth]
-        {docs/source/_static/fitting-methods/algorithm/interpolate-positions.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
-    \end{tikzonimage}
-    \caption{
-        (A) Input image.
-        (B) Result of the skeletonization algorithm for a single cell.
-        (C) Vertices $\vec{x}_i$ which have been interpolated from the skeletonized cell.
-        They represent the position of our cell-agents.
-    }
-    \label{fig:position-extraction-algorithm}
-\end{figure}
-
-% --------------------------------------------------------------------------------------------------
-\paragraph{Benchmarking the Extraction Algorithm}
-\label{paragraph-extraction-algorithm}
-
-In order to determine if our extraction algorithm provides reliable results, we have to take
-advantage of a functionality which will be introduced later in
-section~\ref{subsection:visualization}.
-Our methodology makes use of the fact that we have already set up our model and solving scheme and
-are able to produce cell masks from these numerical results.
-In doing so, we have exact knowledge about the position of our agents since we can simply obtain
-this information from our model.
-But on the other hand we can also try to extract this position from the generated cell-masks.
-By comparing the differences between them, we get a good estimate of how well our extraction
-algorithm can determine the true position of the bacteria.
-
-In figure~\ref{fig:benchmarking-extraction-algorithm} (A-C) we see snapshots of a time series of
-masks.
-This particular simulation includes division events between images (B) and (C).
-For all images (A-C), we have drawn the exact position of the agents with white lines and indicated
-the extracted positions with crosses.
-They align very well visually.
-It may be hard to spot with the bare eye, but the agreement between extracted and exact positions is
-visually worse at the endpoints of the rods.
-To quantify this disparity, we calculate the distance between these individual vertices.
-Figure~\ref{fig:benchmarking-extraction-algorithm} (D) shows the estimated length of the rods from
-the extracted and exact vertices which is surrounded by the average vertex distance.
-Division events lead to steps in our rod-length function since the bacterial length is halved.
-We can clearly see that the agreement is very good in the beginning but right after every division
-event, the uncertainty becomes large.
-To illustrate this, we can see that in figure~\ref{fig:benchmarking-extraction-algorithm} (C) some
-of the rods are partially overlapping which hides parts of the individual cell-masks and thus
-changes the extraction result.
-We also plotted the distribution of individual vertex distances in
-figure~\ref{fig:benchmarking-extraction-algorithm} (E).
-Darker areas in the plot correspond to distances at time-points earlier in the simulatin time.
-We can see that brighter areas tend to shift towards the right, thus increasing the uncertainty of
-the extraction.
-However, the overall distribution shows small values of distances for a large number of vertices and
-is not dominated by outliers.
-
-\begin{figure}[H]
-    \centering
-    \begin{tikzonimage}[width=0.32\textwidth]
-        {docs/source/_static/fitting-methods/extract_positions-004000.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.32\textwidth]
-        {docs/source/_static/fitting-methods/extract_positions-006000.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.32\textwidth]
-        {docs/source/_static/fitting-methods/extract_positions-008000.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.5\textwidth]
-        {docs/source/_static/fitting-methods/displacement-calculations.png}%
-        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
-    \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.5\textwidth]
-        {docs/source/_static/fitting-methods/displacement-distribution.png}
-        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
-    \end{tikzonimage}
-    \caption{
-        Progression of generated masks with estimated and exact positions.
-        Comparison of masks across generations. Compare estimated with exact rod lengths.
-"Iceberg plot"
-    }
-    \label{fig:benchmarking-extraction-algorithm}
-\end{figure}
-
-% --------------------------------------------------------------------------------------------------
-\subsection{Parameter Estimation}
-\label{section:parameter-estimation}
-In order to effectively estimate the parameters of our model, we need to set up a loss function
-which can be minimized.
-In doing so, we hope to identify all parameters, given relevant experimental data.
-We utilize the preceding algorithms to extract data from cell-masks, simulate from an initial state
-and then compare both results.
-The cell-masks are easily obtainable by either using existing segmentation
-tools~\cite{Cutler2022,Stringer2020,Hardo2022} or manually annottating microscopic image series.
-
-\paragraph{No Cell Division}
-In the scenario where no cells divide during the observed interval, we can use our extraction
-algorithm to obtain the position of the cells and then compare them directly to the positions
-calculated from the computational results.
-For all cells $j=1..M$, we have a collection of $i=0..N$ vertices $\vec{x}_{i,j}$.
-Thus, the total amount of information is an array of shape $(M,N,2)$
-Assuming that we have $N$ cells, each of which is represented by $M$ vertcies, the total amount of
-information is an array of shape $(N,M,2)$, containing all positions of all agents.
-It is thus only a matter of comparing two of these arrays with each other and many methods and
-approaches which already exist~\cite{Wang2020} can be leveraged.
-For now, we simply calculate the distance of each vertex pair, which is equivalent to the
-well-known loss function
-
-\begin{equation}
-    L_{\text{LS}}(\vec{x}, \vec{y})
-        = \sqrt{\sum\limits_{i,j}||\vec{x}_{i,j}-\vec{y}_{i,j}||^2}
-        = \sqrt{\sum\limits_{i,j,k}|x_{i,j,k}-y_{i,j,k}|^2}.
-\end{equation}
-
-of least squares (LS).
-One important detail is that it must be guaranteed that the ordering of vertices is identical in the
-numerical and experimentally obtained results.
-
-\paragraph{Comparison across Generations}
-In order to compare a variable amount of cells, we can not utilize the positions of these cells
-for comparison directly since one particular state of our simulation might contain cells which are
-not present in the experimental results (or vice-versa).
-We thus need additional methods which can be used for comparisons across division events.
-To achieve this, we use the numerical results of the \ac{abm} simulation and our visualization
-algorithms to generate cell-masks (see section~\ref{subsection:visualization}).
-We then compare the generated masks to the experimentally obtained ones pixel by pixel.
-Whenever two pixel colors are not matching, a loss value is assigned.
-The sum of all those penalties is then the total loss.
-This algorithm is an implementation of a signed area measure.
-It is very important to take into account the parental relationship between the cells, checking if
-either one of the cells is a daughter of the other cell.
-Figure~\ref{fig:mask-difference-metric} (A,B) shows two successive states of a numerical simulation
-with varying amounts of cells.
-We can see that some division events have occurred but other cells have only grown without
-completing a division step.
-The resulting penalties without accounting for the parental relationships are displayed in
-Figure~\ref{fig:mask-difference-metric} (C).
-Black indicates that the pixels are matching and no loss is assigned while a white pixel
-indicates that the original pixels are not matching, thus assigning a loss value.
-Similarly, Figure~\ref{fig:mask-difference-metric} (D) takes the parental relationship into account
-and highlights these in gray.
-It is possible to assign a custom value for the parental loss value.
-If we set this loss value to $0$, this algorithm ensures the contiuity of the loss function in time.
-
-But it is also biologically motivated.
-If a cell has just divided, the resulting two daughter cells will cover precisely the space that the
-mother had taken up.
-Often times, the exact time at which the fission event is over can not be determined accurately,
-thus leading not only to uncertainties but also some fundamental problems.
-Another approach to construct a comparison algorithm across generations would be to use the
-previously explained position comparison algorithm and use it on intervals in which no division
-event occurrs.
-But this attempt would quickly lead to multiple problems.
-How can we deal with the uncertainties of the timing of the division event?
-And the more cells we want to investigate, the more division events will occurr even in a fixed
-amount of time thus greatly decreasing the effective interval at which this alternative algorithm
-could be used.
-Furthermore, what if there are only two images and there is a division event which is occurring
-between them?
-
-It further needs to be noted that the constructed algorithm can also be adapted by users in order to
-provide more complex loss functions (such as considering time-evolvement, etc.).
-Our method can make use of existing machine leanring algorithms since it is only based on the
-principle of comparing two images with each other.
-There already exists a variety of methods on how to assign loss values in these
-circumstances~\cite{Dice1945}.
-
-\begin{figure}[H]
-    \centering
-    \begin{minipage}{0.5\textwidth}
-    \begin{tikzonimage}[width=0.48\textwidth]
-        {docs/source/_static/fitting-methods/progressions-1.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
-    \end{tikzonimage}%
-    \hspace{0.01\textwidth}%
-    \begin{tikzonimage}[width=0.48\textwidth]
-        {docs/source/_static/fitting-methods/progressions-2.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
-    \end{tikzonimage}
-    \linebreak
-    \vspace{0.01\textwidth}
-    \begin{tikzonimage}[width=0.48\textwidth]
-        {docs/source/_static/fitting-methods/progressions-3.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
-    \end{tikzonimage}%
-    \hspace{0.01\textwidth}%
-    \begin{tikzonimage}[width=0.48\textwidth]
-        {docs/source/_static/fitting-methods/progressions-4.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{D}};
-    \end{tikzonimage}
-    \end{minipage}%
-    \begin{minipage}{0.49\textwidth}
-        \begin{tikzonimage}[width=\textwidth]
-            {docs/source/_static/fitting-methods/penalty-time-flow.png}%
-            \node at (0, 0.975)[anchor=north west, draw, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{E}};
-        \end{tikzonimage}
-    \end{minipage}
-    \caption{
-        The figure shows two masks (A,B) which will be compared against each other.
-        Image (B) contains more cells which are a result of cellular division events that have not
-        yet occurred in image (A).
-        (C) "Naive" comparison of cell masks, which identifies pixels of non-matching colors and
-        assigns a loss by counting these pixels.
-        This simple approach does not take into consideration the parental relationship.
-        This approach is displayed in (D) where pixels of cells that are directly related have been
-        colored in gray.\\
-        (E) Comparison of loss functions, exponential regression (ER) and number of cells.
-        % Total loss between successive iteration steps $t_n,t_{n+1}$.
-        The overall shape of the loss is expected to have the slope of an exponential function.
-        We can clearly see that division events introduce jumps within this relationship.
-        In the case where we account for the parental relationship, these jumps are mitigated.
-        The loss value was set to $0$ for pixel-comparisons which belong to cells that are in a
-        parental relationship.
-    }
-    \label{fig:mask-difference-metric}
-\end{figure}
-
-\paragraph{Testing the Comparison Algorithm with Division Events}
-When utilizing a computational model (see section~\ref{sec:computational-model}), we can generate
-data and use the generated masks to provide a benchmark of our algorithm.
-To do this, we set out to calculate the loss between two successive iteration steps
-$t_n$ with $t_{n+1}$.
-We expect to obtain an exponential growth in the loss.
-This can be explained by the fact that the growth of the total area is exponential in time.
-Furthermore, the temporal area difference can be viewed as a derivative of the total area in time,
-which is again an exponential function.
-Figure~\ref{fig:mask-difference-metric} (E) shows the total loss for this approach.
-We can see that by accounting for the parental relationship, we get the overall desired effect.
-This method allows us to compare results between states that have a different number of cells.
-However, since this algorithm requires us to calculate a cell-mask for each comparison and then also
-process this mask as well as the comparator, it is much more computationally demanding.
-After all, for each comparison step, we need to run a full 3D visualization pipeline.
-In cases where the simpler approach of comparing positions of the agents is possible, this method
-should be preferred.
-
-%###################################################################################################
-\section{A Computational Model for Rod-Shaped Bacteria}
-\label{sec:computational-model}
-
-The mechanical behaviour of rod-shaped bacteria is governed by a plethora of effects.
-In this work, we aim to establish a model which is abstract enough such that it can be reused for
-many types of rod-shaped bacteria but also explicit enough such that we can quantify its parameters.
-We are thus aiming to work with a set of assumptions that is as general as possible but still
-specific to rod-shaped bacteria.
-With this delicate balance in mind, it is clear that many cellular processes which might be specific
-to one particular celltype or simply not necessary to assume in order to construct our model, will
-be omitted alltogether.
-This concerns effects such as polar interactions~\cite{Duvernoy2018}, extracellular
-reactions~\cite{Wang2010} or differentiation processes~\cite{vanGestel2015,Lpez2010} and many other.
-We restrict ourselves to the aspects presented in table~\ref{table:simulation-aspects}.
-For a collection of other known effects which could be considered, we refer to the supplementary
-section~\ref{section:discussion}, and~\ref{section:supplement-more-simulation-aspects}.
-
-\paragraph{(C) Cellular}
-The key component of this model consists of the cellular representation.
-
-It was long believed that bacteria lack cytoskeletal filaments but with the end of the last century,
-it was discovered that the MreB protein fills this role since it “[..] forms an actin-like
-cytoskeleton in bacteria [..]”~\cite{Erickson2001}.
-It can polymerize and form filaments which behave similar to actin microfilaments~\cite{Dersch2020}.
-The effect of MreB on the shape of cells during growth was shown in more detail
-by~\cite{Ursell2014}.
-They found that MreB localizes specifically at regions of negative curvature and thus introduces
-heterogeneous cell growth which results in a bending force during the growth of cells.
-It also means that the cells are able to sense the curvature of the cell wall through cytoskeleteal
-MreB filaments during growth.
-\cite{Bratton2018} showed that the trans-membrane protein RodZ can modulate MreB density and
-facilitates local MreB assembly.
-This feedback mechanism will result in straight rod-shaped bacteria if not interfered from the
-external environment.
-These results were also confirmed by the studies of~\cite{Wang2010}.
-In their experiments, a single bacterium was fixed at the bottom pole while an optically trapped
-bead was placed on the upper pole.
-They could now exert a force on the free part and thus study the response of the cell under
-different conditions.
-When treating cells with A22, an antibiotic which disassembles
-MreB~\cite{IWAI2002,Gitai2005,Karczmarek2007,Bean2009}, a higher displacement with identical force
-could be measured, thus indicating that MreB increases mechanical rigidity of the cell.
-
-While some rod-shaped bacteria grow by inserting new material at the ends, \textit{E.Coli} and
-\textit{B.Subtilis} utilize a method where new material is inserted along the cylindrical part of
-the rod.
-The details of the division event (i.e. Z-Ring formation) are very intricate and are heavily
-studied as a subject on their own~\cite{Harry2001}.
-On a higher level view, the rod is split apart at a point close to the center of the cell, thus
-creating two new daughter cells.
-We know that the best predictor for when the division event occurs, is the length of the rod as
-opposed to the age of bacterium~\cite{Robert2014}.
-Furthermore, only the growth of the whole colony but also the individual growth of the bacteria is
-exponential~\cite{Amir2014,Takeuchi2005}.
-
-\paragraph{(CC) Cell-Cell Interactions \& (DC) Domain-Cell Interactions}
-A colony of bacteria is known to form biofilms (Dunne 2002). This strategy is is beneficial for the
-survival of the collective as it allows it to harness growth factors and migrate when necessary and
-also plays an important role in bacterial infections in the biomedical sector~\cite{Ong1999}.
-The formation of biofilms is only possible due to the attractive nature of bacterial interactions
-towards each other and surfaces (Berne et al. 2018). To analyze bacterial adhesion to surfaces,
-\cite{vanLoosdrecht1989} investigated the interactions of various bacteria with negatively charged
-polystyrene.
-They found that the interaction is characterized by a low Gibbs energy of $(2 - 3)kT$ per cell
-which can be described by a secondary minimum of the DLVO-theory~\cite{Verwey1947,Derjaguin1993}
-which is relevant for distances $\geq\SI{1}{\nano\meter}$.
-At these distances, the DLVO potential acts as a generic adhesive potential.
-Particles which stay in the secondary minimum are held together but even weak interactions are
-enough to redisperse them which means that this state is reversible.
-When approaching closer distances, more intricate mechanisms come into play in which the
-heterogeneity of the materials in question needs to be considered.
-\cite{Hori2010} described bacterial adhesion as a “two-phase process including an initial,
-instantaneous and reversible physical phase (phase one) and a time-dependent and irreversible
-molecular and cellular phase (phase two)”.
-
-All these considerations become even more intricate when not considering interactions of bacteria
-with surfaces but bacteria with each other.
-Gram-negative bacteria (eg. E.Coli) express outer membrane proteins (adhesins) which allow them to
-attach to various hosts (Beachey 1981, Vaca et al. 2019).
-In addition, the flagella can play an important role in the direct adhesion between
-cells~\cite{Haiko2013}.
-Thus it is not correct to blindly apply the results for adhesion to surfaces to the interactions
-between bacteria.
-
-\begin{table}[H]
-    \centering
-    \def\arraystretch{1.3}
-    \begin{tabularx}{\textwidth}{c l X}
-        &\textbf{Aspect} & \textbf{Description}\\
-        \toprule
-        &\textbf{(C) Cellular}\\
-        \midrule
-        (1) & Rod-Shaped Mechanics &
-            Rod-shaped bacteria are flexible rods which are able to freely move around (off-lattice
-            approach) \cite{Takeuchi2005,Ursell2014,Amir2014_2}.\\
-        (2) & Growth &
-            Cells grow exponentially by inserting new material either along the circular part of the
-            rod or at the tip~\cite{Robert2014,Takeuchi2005}.\\
-        (3) & Division &
-            Rod-Shaped bacteria divide in the middle of the rod into two new agents.\\
-            % Bacteria can form multilayers during their growth phase \cite{Duvernoy2018}.\\
-        (4) & Individual Parameters &
-            Parameters for individual cells are not fixed values but rather taken from a
-            distribution \cite{Koutsoumanis2013}.\\
-        &\textbf{(CC) Cell-Cell Interactions}\\
-        \midrule
-        (5) & Adhesion &
-            Bacteria attract at longer distances and adhere to each other at shorter
-            distances~\cite{Verwey1947,Trejo2013}.\\
-        \bottomrule
-    \end{tabularx}
-    \label{table:simulation-aspects}
-    \caption{
-        This table lists aspects of cellular behaviour which we consider in our computational model.
-        It is split into cellular aspects (C) which concern only an individual bacterium and
-        interactions between cells (CC).
-        These aspects are generalized in order to fit to as many types of rod-shaped bacteria as
-        possible.
-        These are the assumptions for the basic model that we are presenting here.
-        In order for our methods to work, we are really only required to satisfy the assumption of
-        our spatial representation (see section~\ref{subsec:spatial-representation}) and thus, these
-        assumptions can also be loosened if required.
-    }
-\end{table}
-
-To model the spatial mechanics of elongated bacteria \cite{Billaudeau2017}, we represent them as a
-collection of auxiliary vertices $\{\vec{x}_i\}$ which are connected by springs in
-ascending order.
-Furthermore, we assume that the cells are flexible described by their rigidity.
-A force $\vec{F}$ interacting between cellular agents determines the radius (thickness) of the
-rods and an attractive component can model adhesion between cells.
-
-% --------------------------------------------------------------------------------------------------
-\subsection{Mechanics}
-\label{subsection:mechanical-model-mechanics}
-
-We use principles from discrete differential geometry~\cite{bobenko2008discrete} to describe the
-stiffness and flexibility of the bacterial rod.
-As explained previously, MReB will counteract the bending of the rod, resulting in a stiffening
-force.
-In principle, this stress will be dependent on the position inside the cell~\cite{Chatterjee1988}
-but in order to construct a model which can be parametrized and by method of Occam's Razor, we chose
-to model the bacterium as a bending beam, thus utilizing the simplest solution which can still
-produce the desired effects.
-Such a beam can be represented by a collection of vertices $\vec{x}_i$ which are connected by a
-springs of lengths $l_i$.
-We use the notation $\vec{c}_i:=\vec{x}_i-\vec{x}_{i-1}$ with which the resulting force acting
-between two vertices if given by
-
-\begin{align}
-    \vec{F}_{i,\text{springs}} =
-        &-\gamma\left(1 - \frac{l}{\left|\vec{c}_i\right|}\right)
-        \vec{c}_i\\
-        &+ \gamma\left(1 - \frac{l}{\left|\vec{c}_{i+1}\right|}\right)
-        \vec{c}_{i+1}.
-\end{align}
-
-In addition to springs between individual vertices $\vec{x}_i$, we assume that each angle at a
-vertex between edges is subject to a force indiced by curvature.
-% We assume that we can model the mechanical properties of the bacterium as an elastic rod.
-Given the angle $\alpha_i = \sphericalangle(\vec{c}_{i-1},\vec{c}_i)$ between adjacent edges,
-the bending force is proportional to the curvature $\kappa_i$ at each vertex
-$\kappa_i = 2\tan(\alpha_i/2)$.
-The resulting force acts along the angle bisector which can be calculated from the edge vectors.
-Figure~\ref{fig:mechanics-bacterium} shows an elongated rod represented by $6$ vertices.
-The forces acting on vertices $\vec{x}_i,\vec{x}_{i-1},\vec{x}_{i+1}$ are given by
-
-\begin{align}
-    \vec{F}_{i,\text{curvature}} &= \eta\kappa_i
-        \frac{\vec{c}_i - \vec{c}_{i+1}}{|\vec{c}_i-\vec{c}_{i+1}|}\\
-    \vec{F}_{i-1,\text{curvature}} &= -\frac{1}{2}\vec{F}_{i,\text{curvature}}\\
-    \vec{F}_{i+1,\text{curvature}} &= -\frac{1}{2}\vec{F}_{i,\text{curvature}}
-\end{align}
-
-where $\eta_i$ is the rigidity at vertex $\vec{x}_i$ (see Figure~\ref{fig:mechanics-bacterium}).
-We can see that the curvature force does not move the overall center of the rod in space.
-The total force is the sum of external and interal forces.
-
-\begin{equation}
-    \vec{F}_{i,\text{total}} = \vec{F}_{i,\text{springs}}+ \vec{F}_{i,\text{curvature}}
-        + \vec{F}_{i,\text{external}}
-\end{equation}
-
-and are integrated via
-
-\begin{align}
-    \partial_t^2 \vec{x}_i &= \partial_t\vec{x}_i + \sqrt{2D}\vec{\xi}_i\\
-    \partial_t\vec{x}_i &= \vec{F}_{i,\text{total}} - \lambda \partial_t\vec{x}_i
-\end{align}
-
-where $D$ is the diffusion constant and $\vec{\xi}$ is the wiener process.
-Together they determine the stochastic motion of the individual vertices which can arise from
-recurring collisions with small particles of the surrounding medium or other minor fluctuations.
-
-% --------------------------------------------------------------------------------------------------
-\subsection{Physical Interactions}
-\label{subsection:mechanical-model-interactions}
-
-When calculating forces acting between the cells, we can use a simplified model to circumvent the
-numerically expensive integration over the complete length of the rod.
-Given a vertex $\vec{x}_i$ on one cell, we calculate the closest point $\vec{p}$ on the polygonal
-line given by the vertices $\{\vec{w}_j\}$ of the interacting cell.
-Furthermore we determine the value $q\in[0,1]$ such that $\vec{p} = (1-q)\vec{w}_j + q\vec{w}_{j+1}$
-for some specific $j$.
-The force is then calculated between the points $\vec{x}_i$ and $\vec{p}_i$ and acts on
-the vertex $\vec{w}_i,\vec{w}_{i+1}$ with relative strength $(1-q)$ and $q$.
-
-\begin{equation}
-    \vec{F}_{i,\text{External}} = \vec{F}(\vec{x}_i,\vec{p})
-    \label{eq:cell-force-external}
-\end{equation}
-
-The specific choice of interaction potential plays of course a crucial role for the dynamics of the
-agents.
-But we cannot expect to simply know the shape and parameters of this potential.
-Thus, we would like to experiment with various interaction potentials in order to quantify their
-ability to produce accurate results.
-We utilize two distinct potentials that are both able to capture the essence of our interactions by
-providing a short-range repulsive part and a long-ranged attractive component.
-
-\paragraph{Morse Potential}
-The morse potential~\cite{Morse1929} is a common example of numerically inexpensive interaction
-potential.
-It was originally designed to describe the interaction of diatomic molecules but has since found use
-in other areas~\cite{Breitwieser2021}.
-It contains a repulsive and attractive part and is bound from above which makes it particularly
-suitable for numerical simulations since unbound potentials can lead to large numerical deviations
-thus compromising the prediction of the model.
-The potential is of the form
-\begin{equation}
-    V(r) = V_0\left(1-e^{-\lambda(r-(R_1+R_2))}\right)^2
-\end{equation}
-and has 3 parameters.
-The overall potential strength $V_0$, the radii of the particles $R_1,R_2$ and the potential
-stiffness $\lambda$ which controls the rate of change as one travels along the radial axis.
-We employ a spatial cutoff which sets the interaction strength to $0$ when exceeding a distance
-$r\geq\xi$.
-
-\paragraph{Mie Potential}
-The Mie~\cite{Mie1903} potential is the generalization of the well-known
-Lennard-Jones~\cite{Jones1924} potential.
-It was also designed to describe the interactions of molecules but due to the flexible nature of its
-parameters, can be easily adapted to other situations.
-It is able to describe a wider range of shapes compared to the Morse potential but is also stiffer
-and not bound from above.
-In our simulation, we enforce an upper bound to guarantee numerical regularity.
-It has 4 parameters and is given by
-\begin{align}
-    U(r) &= C\epsilon\left[ \left(\frac{\sigma}{r}\right)^n -
-        \left(\frac{\sigma}{r}\right)^m\right]\\
-    C &= \frac{n}{n-m}\left(\frac{n}{m}\right)^{\frac{n}{n-m}}\\
-    V(r) &= \min(U(r), \beta)\theta(r-\zeta)
-\end{align}
-where $C$ is a constant that depends on the two exponents $n,m$ that describe the overall shape of
-the potential.
-The strength is given by $\epsilon$ and the radius is encoded in $\sigma =
-(m/n)^{1/(n-m)}(R_1+R_2)$ where $R_1,R_2$ are the radii of the two interacting particles.
-
-% --------------------------------------------------------------------------------------------------
-\subsection{Cell-Cycle}
-\label{subsection:mechanical-model-cycle}
-
-To simulate proliferation, we introduce a growth term for the spring lengths $l_i$.
-In most of our simulations, we choose to use identical lengths for all segments (i.e. $l_i=l$).
-However, in order to generalize our approach we assume that individual lengths of segments could
-vary.
-We reduce the growth rate with increasing number of neighbors $N$.
-This behaviour can be toggled on or off thus removing the additional term in brackets of
-equation~\ref{eq:growth-rod-lengths}.
-
-\begin{equation}
-    \partial_t l_i = \mu\left(1-\frac{\text{min}(n,N)}{N}\right)l_i
-    \label{eq:growth-rod-lengths}
-\end{equation}
-
-Note that the growth rate is proportional to the rod length $l$ which leads to exponential growth.
-This process will increase the length of the cell indefenitely unless we impose a condition for the
-division event.
-We define a fixed length threshold for the total length of the cell at which it divides.
-To construct a new cell, we cannot simply copy the existing one twice, but we also need to adjust
-internal parameters in the process.
-The following actions need to be taken for the old and new agent.
-We first calculate the new spring lengths of the two resulting rods.
-\begin{equation}
-    \tilde{l} = l_i\left(\frac{1}{2} - \frac{r}{\sum\limits_i l_i}\right)
-    \label{eq:growth-rod-new-lengths}
-\end{equation}
-Afterwards, we start at one endpoint $\vec{x}_0$ and travel along the polygon defined by the
-bacterial rod.
-We stop when we have reached a length that matches the length of the new spring length $\tilde{l}$.
-From here, we continue and iterate this procedure until we have found all new vertices $\vec{w}_i$.
-Equation~\ref{eq:growth-rod-new-lengths} ensures that that the tips of the two new rods are touching
-right after the division event.
-
-% --------------------------------------------------------------------------------------------------
-\subsection{Numerical Implementation}
-
-We implemented the described model in \texttt{cellular\_raza}~\cite{Pleyer2025}.
-Doing so has many benefits.
-Not only, was a more simpler model already available which could be adapted to our needs.
-But furthermore, due to the nature of \texttt{cellular\_raza} in that it does not assume any
-particular cellular representation, we are free to change the underlying model at any time, even
-making fundamental changes.
-It is an \ac{abm} which is able to natively track cells through the course of a numerical simulation
-and also keeps track of parent/daughter relationships.
-This allows us to to experiment with different model implementations and still retain performance.
-By leveraging the strong safety guarantees provided by the Rust programming language, we eliminate a
-whole class of problems from the start.
-% TODO CITATION
-In addition, we are using the \texttt{pyo3} and \texttt{maturin} crates which allow us to easily
-generate python bindings and thus interoperate between the two ecosystems.
-While this particular model is more complex due to the nature of non-trivial (asymmetric)
-interactions, we were still able to properly utilize the speedup from executing simulations in
-parallalel using the built-in parallelization methods of \texttt{cellular\_raza} (see
-Supplementary Material, Figure~\ref{fig:performance}).
-
-%###################################################################################################
-\section{Application to \textit{E.Coli.} growing on Agar}
-
-With the previous works of extracting the positions (see section~\ref{section:data-extraction}) for
-individual cell-agents and the methods to estiamte parameters of our agents (see section~\ref{section:parameter-estimation}), we are now ready to apply them to showcase their usefulness.
-
-TODO explain what data we are using
-
-% --------------------------------------------------------------------------------------------------
-\subsection{Individual Treatment}
-We have already suspected (see table~\ref{table:simulation-aspects}) that in order to properly
-estimate parameters of our system, we need to treat every cell-agent individually, thus assigning
-parameters individually to them.
-This means that we increase the total number of parameters that needs to be sampled within an
-optimization routine.
-Given $N_c$ cells and $N_p$ parameters per cell and $N_u$ unique parameters (where $N_u\leq N_p$),
-the total amount of parameters is
-
-\begin{equation}
-    N_{p,\text{total}} = N_u + N_c (N_p - N_u) = N_C N_p - (N_c - 1) N_u.
-\end{equation}
-
-It is clear, that we should avoid to needlessly assign every parameter individually since the space
-which is needed to be sampled scales exponentially with the dimension ie. the number of parameters.
-However, we will now present data that indicate that an individual treatment is necessary.
-Figure~\ref{fig:estimated-growth-rates} shows the rod-lengths of individual bacteria which have been
-calculated from the extracted positions.
-We fit a delayed exponential function to them which is defined by
-
-\begin{equation}
-    x(t) =
-    \left\{\begin{array}{ll}
-            x_0 \exp(\lambda (t-t_0)) & t\geq t_0\\
-            x_0 & \text{ else}
-    \end{array}\right.
-\end{equation}
-
-and where $\lambda$ is the growth rate $t_0$ the delay.
-The fits to invidiual growth-curves of figure~\ref{fig:estimated-growth-rates} (A) result in good
-matches and so does the fit of subfigure (B).
-However, the latter exhibits large amounts of uncertainty within the data which is also illustrated
-in subfigure (C).
-There, we can see much more clearly the difference between considering invidiually assigned
-parameters and global ones.
-It is clear from this observation that in order for us to accurately estimate the parameters of this
-system, we need to treat the bacteria individually.
-
-The individual treatment of the bacteria allows us in principle to estimate the distribution of the
-parameters directly without having to resort to any assumptions on its form.
-This is true for any parameter, not just the growth rates, since we aim to estimate all mechanical
-parameters of our cellular model.
-
-\textbf{TODO is this discussion?}\\
-The total parameter space $P$ is a product of all single spaces $P_i$ in which the parameters
-$p_i$ live $P=P_0\times P_1\times\dots$ but some of these parameters share the same space (let's
-assume $p_0,p_1\in \tilde{P}$).
-This leaves the open question as to weather the optimization algorithm could be modified to leverage
-this knowledge and obtain increased performance.
-For instance, it might be valuable to initially consider individual parameters as globally defined
-and start treating the individually after a certain amount of optimization steps.
-
-% \begin{enumerate}
-%     \item motivate individual treatment of cells
-%     \item pro: get at distributions of parameters; even more than growth rates
-%     \item pro: probably good idea to do this in this case, since more or less required (see figure)
-%     \item con: computationally more demanding
-%     \item can we improve the optimization algorithm to somehow "know" that some parameters are for
-%         different cells and thus do not "interact" as other parameters with each other?
-%         ie. they are part of the same subspace.
-% \end{enumerate}
-
-\label{subsec:parameter-estimation-individual-treatment}
-\begin{figure}[H]
-    \centering
-    \begin{tikzonimage}[width=0.33\textwidth]
-        {figures/crm_estimate_params/IWF-Goettingen/rod-lengths-individual.pdf}%
-        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{A}};
-    \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.33\textwidth]
-        {figures/crm_estimate_params/IWF-Goettingen/rod-lengths-average.pdf}%
-        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
-    \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.33\textwidth]
-        {figures/crm_estimate_params/IWF-Goettingen/parameter-distribution.pdf}%
-        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
-    \end{tikzonimage}%
-    \caption{
-        % TODO use same y-axis across the plots
-        (A) The subfigure shows the time-evolution of the rod-lengths for every individual
-        bacterium.
-        These growth curves are fitted individually with delayed exponential function.
-        (B) The datapoints are created by averaging the rod-lengths shown in subfigure (A) and then
-        fitted with an exponential function.
-        The uncertainty band is the standard deviation of the averaged datapoints.
-        (C) Blue dots show the combination of estimated growth rates and delays for individual
-        bacteria while the orange dot corresponds to the average.
-        Ellipses around the datapoints are taken from the uncertainty of the fits.
-        Time is given in units of frames of the movie from which the data have been extracted.
-    }
-    \label{fig:estimated-growth-rates}
-\end{figure}
-
-% --------------------------------------------------------------------------------------------------
-\subsection{Without Division}
-For now, we will focus on a time-interval which does not contain any division events.
-By doing so, we can utilize the simpler and more performant position comparison scheme for our loss
-function.
-For this dataset, we assumed that all interaction parameters which specify the interaction
-potential are identical across all agents but growth rates and the physical radii of the cells are
-individual and we optimize them like any other global parameter.
-We chose to optimize for 3 interaction parameters of the Mie potential $n,m,V_0$ and two parameters
-which were individual for each cell, namely the radii $r_i$ and growth rates $\mu_i$.
-We also included the damping constant $\lambda$ and removed any stochastic motion contributions
-($D=0$).
-This leads to a total of $4+2\times 6=16$ parameters.
-To estimate these parameters we used a latin-hypercube~\cite{McKay1979} sweep across the parameter
-space and performed a local minimization with the Nelder-Mead~\cite{Powell1973} method.
-
-In figure~\ref{fig:parameter-estimates-single-step} we performed a sensitivity-analysis for every
-parameter.
-We can see that almost all parameters can be estimated by our optimization approach.
-However, the damping constant (A) does not show any changes in the loss function when increasing its
-value.
-It hints at a non-identifiability~\cite{Raue2009} of the parameter $\lambda$.
-But it also shows that the approach which we used is able to detect these problems.
-As we continue to develop these methods, we expect to be able to reuse many of the existing
-theoretical approaches such as likelihood estimation and uncertainty quantification which have
-already been developed extensively for less complex systems such as \acp{ode}.
-\textbf{TODO citations, also: is this discussion?}
-Fortunately, \texttt{cellular\_raza} was chosen to construct our model such that we are flexible in
-changing the underlying representation if should find that such unidentifiabilities are structural
-(model-based) rather than practical (due to lack of data).
-This allows us to iterate and find the model formulation which suits the data best but still retain
-the mechanistic methodology of our approach.
-
-\begin{figure}[H]
-    \centering
-    \begin{tikzonimage}[width=0.54\textwidth]
-        {figures/crm_fit/snapshots/predicted-031111.png}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
-    \end{tikzonimage}%
-    \hspace{0.01\textwidth}%
-    \begin{tikzonimage}[width=0.41\textwidth]
-        {figures/crm_fit/celldifss/cell-031111.png}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
-    \end{tikzonimage}%
-    \caption{
-        \textbf{TODO do we need this figure?}
-        (A) A microscopic image with the predicted positions of the cells overlayed on top.
-        (B) Differences between extracted and predicted positions.
-        % TODO make this figure to scale
-    }
-    \label{fig:celldiffs-individual}
-\end{figure}
-
-\begin{figure}[H]
-    \centering
-    \begin{tikzonimage}[width=0.33\textwidth]
-        {figures/crm_fit/profiles/damping.pdf}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{A}};
-    \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.33\textwidth]
-        {figures/crm_fit/profiles/exponent-m.pdf}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
-    \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.33\textwidth]
-        {figures/crm_fit/profiles/exponent-n.pdf}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
-    \end{tikzonimage}\\
-    \begin{tikzonimage}[width=0.33\textwidth]
-        {figures/crm_fit/profiles/strength.pdf}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
-    \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.33\textwidth]
-        {figures/crm_fit/profiles/growth-rate-0.pdf}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
-    \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.33\textwidth]
-        {figures/crm_fit/profiles/radius-0.pdf}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
-    \end{tikzonimage}
-    \caption{
-        Sensitivity analyses werer performed for every optimized parameter of our model.
-        (A-F) show the curves of loss functions when varying this respective parameter and keeping
-        all others constant at the value of the found optimum (blue dot).
-        In the cases of (B-F), we can see that a minimum was achieved and thus the parameters could
-        be estimated.
-        However, for the case of the damping constant $\lambda$ in (A), we see that the upper bound
-        is open.
-        This suggests that our model could be reduced with regard to this parameter.
-        We have exemplified the values for radii $r_i$ and growth rates $\mu_i$ with values of the
-        first cell.
-        The remaining plots can be found in the supplementary material in
-        section~\ref{sec:supplement-parameter-profiles}.
-    }
-    \label{fig:parameter-estimates-single-step}
-\end{figure}
-
-% --------------------------------------------------------------------------------------------------
-\subsection{With Division (TODO)}
-
-%###################################################################################################
-\section{Discussion (TODO)}
-\label{section:discussion}
-
-\begin{itemize}
-    \item Model does not capture every aspect of rod-shaped bacteria. Needs to extended (see our
-        review).
-    \begin{enumerate}
-        \item polar interactions
-        \item differentiation for van Gogh bundles; surfactin-producing, matrix-producing cells.
-        \item Proper friction; this might have a huge implication on the estimation of
-            parameters~\cite{Grant2014}.
-            This paper has already implemented this~\cite{Doumic2020}.
-            We can build on top of that.
-        \item Extracellular Reactions~\cite{Li2025}
-        \item see supplement for full list \ref{table:simulation-aspects-supplement}
-    \end{enumerate}
-    \item Data Extraction: Skeletonization Algorithm ultimately not good enough for some scenarios
-        (i.e. overlapping of cells, endpoints can sometimes be "too short")
-    \begin{itemize}
-        \item Possibly use AI for new algorithm
-    \end{itemize}
-    \item Image generation; Produce high-quality near-realistic microscopic images with cell-masks
-        from scratch with mechanistic model; train cell-segmentation and more importantly
-        cell-tracking algorithms with that.
-\end{itemize}
-
-\section{Conclusion (TODO)}
-\label{section:conclusion}
-
-\begin{itemize}
-    \item Highlight methodological advancements
-    \item Constructed Model which already combines many effects; extend this even further
-    \item can now use classical methods: sensitivity-analysis, likelihood estimators and uncertainty quantification
-    \item Proof of concept: Able to fit individual parameters of the agents; can be extend this?
-    \item Provide software package which combines all functionality and is readily available
-\end{itemize}
-
-%###################################################################################################
-% EXCLUDE THIS FOR NOW
-% \section{Plastic Deformations}
-% 
-% \begin{figure}[H]
-%     \centering
-%     \includegraphics[width=0.49\textwidth]
-%         {docs/source/_static/scripts/crm_amir/0000000023.png}\\
-%     \includegraphics[width=0.49\textwidth]
-%         {docs/source/_static/scripts/crm_amir/angles.pdf}
-%     \includegraphics[width=0.49\textwidth]
-%         {docs/source/_static/scripts/crm_amir/endpoints.pdf}
-%     \caption{TODO}
-%     \label{fig:amir}
-% \end{figure}
-
-\bibliographystyle{IEEEtran}
-\bibliography{references}
-
-\appendix
-%###################################################################################################
-\section*{Supplementary Material}
-
-\renewcommand{\thesection}{S\arabic{section}}
-
-%###################################################################################################
-\section{Data Preparation}
-\begin{enumerate}
-    \item Have masks of data; reference image segmantation tools; or do it by hand
-    \item ensure that colors are matching for bacteria between time-steps
-    \item clean borders of images if needed
-    \item know domain size; otherwise use pixels
-\end{enumerate}
-
-%###################################################################################################
-\section{Parameter Estimation}
-\label{sec:supplement-parameter-profiles}
-%TODO {explain profiles}
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/growth-rate-0.pdf}%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/growth-rate-1.pdf}%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/growth-rate-2.pdf}\\%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/growth-rate-3.pdf}%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/growth-rate-4.pdf}%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/growth-rate-5.pdf}
-    \caption{Parameter estimation of growth rates for individual cells}%TODO extend caption}
-    \label{fig:parameter-estimates-supplement-growth-rates}
-\end{figure}
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/radius-0.pdf}%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/radius-1.pdf}%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/radius-2.pdf}\\%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/radius-3.pdf}%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/radius-4.pdf}%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/radius-5.pdf}%
-    \caption{Parameter estimation of radii for individual cells}%TODO extend caption}
-    \label{fig:parameter-estimates-supplement-radii}
-\end{figure}
-
-%###################################################################################################
-\section{More Simulation Aspects}
-\label{section:supplement-more-simulation-aspects}
-%TODO insert this into the discussion later
-\begin{table}[H]
-    \centering
-    \def\arraystretch{1.3}
-    \begin{tabularx}{\textwidth}{c l X}
-        &\textbf{Aspect} & \textbf{Description}\\
-        \toprule
-        &\textbf{(C) Cellular}\\
-        \midrule
-        (1) & Rod-Shaped Mechanics &
-            Rod-shaped bacteria are flexible rods which are able to freely move around (off-lattice
-            approach) \cite{Takeuchi2005,Ursell2014,Amir2014_2}.\\
-        (2) & Growth &
-            Cells grow exponentially by inserting new material either along the circular part of the
-            rod or at the tip~\cite{Robert2014,Takeuchi2005}.\\
-        (3) & Differentiation &
-            \textit{B.subtilis} is known to differentiate into matrix-producing and
-            surfactin-producing cells \cite{vanGestel2015,Lpez2010}.\\
-        (4) & Division &
-            The formation and continuation of van Gogh bundles is driven by cell division
-            \cite{vanGestel2015}.
-            Bacteria can form multilayers during their growth phase \cite{Duvernoy2018}.\\
-        (5) & Variable Parameters &
-            Parameters for individual cells are not fixed values but rather taken from a
-            distribution \cite{Koutsoumanis2013}.\\
-        &\textbf{(CC) Cell-Cell Interactions}\\
-        \midrule
-        (6) & Adhesion &
-            Bacteria adhere to each other at longer distances and attach when in close contact
-            \cite{Verwey1947,Trejo2013}.
-            The interaction between rods can be polarized \cite{Duvernoy2018}.\\
-        (7) & Friction &
-            Friction between cells \cite{Grant2014} can be asymmetrical \cite{Doumic2020}.\\
-        &\textbf{(DC) Domain-Cell Interactions}\\
-        \midrule
-        (8) & External Forces &
-            Bacteria tend to stick to surfaces \cite{vanLoosdrecht1989}.
-            The extracellular gel exerts a force onto the cells \cite{Grant2014}.\\
-        (9) & Extracellular Reactions &
-            Bacteria can take up nutrients or possible secrete/take up signalling molecules
-            \cite{Li2025}.\\
-        \bottomrule
-    \end{tabularx}
-    \label{table:simulation-aspects-supplement}
-    \caption{
-        This table extends table~\ref{table:simulation-aspects} with additional observations.
-        It has to be noted that this table includes more specific behaviour that might not be
-        generalizable across all types of rod-shaped bacteria.
-    }
-\end{table}
-
-%###################################################################################################
-\section{Performance}
-
-\begin{itemize}
-    \item Quadratic Scaling comes from more dense simulation
-    \item Overall number of agents which can be simulated is satisfactory for current purposes
-    \item Improvement from $1 \Rightarrow 2$ and $2\Rightarrow4$ threads; first is major; second
-        only minor
-    \item Flamegraph still shows much cycles running against hurdles::Barrier::wait $\Rightarrow$
-        this leaves room for improvement
-    \item Flamegraph also shows that $\approx 75\%$ of all cycles is related to actual work (not
-        just waiting).
-        In particular to update\_mechanics functions.
-\end{itemize}
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.49\textwidth]
-        {docs/source/_static/performance/computation-time-with-initial-agents.pdf}
-    \includegraphics[width=0.49\textwidth]{docs/source/_static/performance/flamegraph.pdf}
-    \caption{TODO}
-    \label{fig:performance}
-\end{figure}
-
-%###################################################################################################
-\section{Multilayer Predictions}
-% TODO
-
-\begin{itemize}
-    \item Friction force against surface. perpendicular to motion; proportional to velocity.
-    \item gel pressure coming from experiment (simply downwards force $F_z$)
-\end{itemize}
-
-\begin{itemize}
-    \item Do we need this in this publication?
-    \item Maybe this is out of scope as this is a more methodological paper.
-    \item If we want to omit it, we might also opt to change the model description in the beginning
-        and remove some parts of the external forces discussion.
-\end{itemize}
-
-\cite{You2019}
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.49\textwidth]
-        {docs/source/_static/scripts/crm_multilayer/multilayer-time-evolution.pdf}\\
-    \includegraphics[width=0.49\textwidth]
-        {docs/source/_static/scripts/crm_multilayer/colony-height-vs-gel_pressure.pdf}
-    \includegraphics[width=0.49\textwidth]
-        {docs/source/_static/scripts/crm_multilayer/colony-height-vs-time.pdf}
-    \caption{TODO}
-    \label{fig:multilayer}
-\end{figure}
-
-\end{document}

### 2025-08-31 paper/main.tex
move old introduction to separate file

@@
 
-General comments on \acp{abm} and their problems
-\begin{itemize}
-    \item \aclp{abm} are used for many purposes
-    \item describe current way of estimating parameters; compare distributions to each other; no idea if
-          individual parameters are correct
-    \item challenge: estimate their parameters
-    \item challenge: construct minimal models (requires flexibility)
-\end{itemize}
-
-General remarks about the steps in this work
-\begin{itemize}
-    \item Use rod-shaped bacteria as model system
-    \item Clear separation between different components (link to overview)
-    \item This separation allows us to generalize the type of approach and apply to other models
-    \item think of as a "blueprint" for estimating parameters for \acp{abm}
-\end{itemize}
-
-Experimental observations: rod-shaped bacteria
-\begin{itemize}
-    \item growth + division "distinct mechanical properties", MreB
-    \item bendiness (not much studied)
-    \item individual parameters (i.e. growth rates)
-    \item adhesion to each other
-\end{itemize}
-
-Computational studies of rod-shaped bacteria
-\begin{itemize}
-    \item not so many models for rod-shaped bacteria
-    \item could include more effects such as bending
-    \item almost no efforts to estimate parameters with individual-based models
-    \item although many things known no "generalized mechanical model" constructed for rod-shaped bacteria
-\end{itemize}
-
-\section{Introduction (old)}
-
-The last decades have fundamentally challenged and changed our understanding of how bacteria retain
-their shape.
-They exist in many physical forms such as spheroidal, rod-shaped or
-spiral~\cite{Zapun2008,Young2006}.
-Rod-shaped bacteria can grow by extending their cylindrical part, or by inserting new material at
-the tip of the rod.
-Species such as \textit{E.coli} and \textit{B.subtilis}~\cite{Errington2020} fall under the first
-category while \textit{S.pombe} represents the latter.
-New material is inserted in small bursts on the nanoscale in forms of patches, bands or
-hoops~\cite{DePedro2003}.
-Despite their differences such as wall-size or length \textit{B.subtilis} and \textit{E.coli} follow
-growth rules which are comparable with respect to the extension of the rod~\cite{Chang2014}.
-
-\paragraph{Experimental studies of mechanical properties}
-It was long believed that bacteria lack cytoskeletal filaments but with the end of the last
-century, it was discovered that the MreB protein fills this role since it "[..] forms an actin-like
-cytoskeleton in bacteria [..]" \cite{Erickson2001}.
-It can polymerize and form filaments which behave similar to actin microfilaments \cite{Dersch2020}.
-Studies of its crystal structure further showed the similarity to actin \cite{vandenEnt2001} and the
-MreB, MreC and MreD proteins were identified as homologues of actin \cite{Lowe2017_lj}.\
-Jones et al. studied the species of different bacterial sub-kingdoms which have the MreB
-gene~\cite{Jones2001}.
-They concluded: "Many of the organisms are rod shaped, similar to \textit{B. subtilis} and
-\textit{E. coli}.
-Organisms with more complex shapes, including curved, filamentous, and helical bacteria, were all
-abundantly represented." \cite{Jones2001}.
-Furthermore, \cite{Wachi1987} found that mutants of \textit{E.Coli} which create defective MreB
-proteins will be spherical instead of their natural rod-shaped form.
-
-% \begin{itemize}
-%     \item \cite{Robert2014} growth of individual rods, distribution of parameters
-%     \item \cite{Koutsoumanis2013} inheritance of growth rates, growth parameter distributions
-%     \item \cite{Ursell2014} MreB acts againts curvature during growth of cell
-%     \item \cite{Takeuchi2005} cells in microchambers leads to plastic deformations
-%     \item \cite{Amir2014} bending stiffness, plastic + elastic deformations
-% \end{itemize}
-Other experiments found that individual rods of \textit{E.Coli} grow exponentially until the
-division event which is triggered by a size-sensing rather than a time-sensing
-mechanism~\cite{Robert2014}.
-These results also showed that the growth parameters were randomly distributed but did not quantify
-said distribution.
-Using $220$ \textit{Salmonella enterica}, Koutsoumanis et al. analysed the distribution of division
-times and was able to quantify their distributions~\cite{Koutsoumanis2013}.
-They found that while the initial division times were longer, there was no correlation between
-division times of mother and daughter cells.
-The first finding can be explained by the bacterias lag-phase~\cite{Bertrand2019} which is a
-temporary period during which bacteria do not grow.
-The second observation is an example of principle of thought which underlyies guides this work:
-Cells are individual agents with individual parameters and thus individual behaviour and need to be
-treated as such.
-This is especially important if we want to determine parameters of our model for individual agents.
-
-% \paragraph{Collective Effects \& Self-Organization}
-% are emergent phenomena that result from the combined interactions of multiple cells.
-% Although they do not allow us to fundamentally understand, the individual behaviour of cells
-% directly, they can provide a way to falsify our results and thus further guide our modeling
-% approach.
-% \cite{Trejo2013} investigated the mechanical response of a tightly packed colony of \textit{Bacillus
-% Subtilis}.
-Trejo et al. studied the formation of macroscopic wrinkles when growing bacteria in a confined
-space~\cite{Trejo2013}.
-They were able to model the strain and stress of the collection of bacteria which requires that
-cells are connected by a (to first order) spring-like force.
-The developing wrinkles could then be described by a buckling instability.
-
-Duvernoy et al. used laser techniques and force microscopy to investigate adhesion of individual
-\textit{E.coli} and \textit{Pseudomonas aeruginosa} cells~\cite{Duvernoy2018}.
-They found that the adhesion forces of the rods are polar, resulting in mechanical tension which in
-turn determines daughter cell arrangement.
-% They measured the size of the grown bacterial colonies, fitted ellipses to their shapes and found
-% that adhesion and polar adhesion in particular resulted in a greater difference between the long and
-% short axis of the fitted ellipse and thus a more oval shape rather than circular (WT).
-By extending the work of \cite{Grant2014}, they further showed, that the critical value of
-transitioning from a monolayer to multiple layers of bacteria, depends on the adhesive strength and
-polarity of the interaction.
-Grant et al. described the growing bacterial colony as a disc with radius $R$ which is pressed
-into the agarose slab~\cite{Grant2014}.
-The authors created a purposely-designed simulation in `C++` in order to model the effects of the
-gel on the collection of bacteria.
-
-Van Gestel et al. explored how \textit{B.subtilis} migrates over a surface by forming multicellular
-structures~\cite{vanGestel2015}.
-The cells organize themselves into "van Gogh bundles" of cells which are tightly aligned in chains
-and form filamentous loops.
-This phenomenon occurs at the border of the bacterial colony and the migration is driven by two
-phenotypically different cell types~\cite{Lpez2010}.
-% \begin{itemize}
-%     \item \cite{Trejo2013} mechanical response of bacterial colony
-%     \item \cite{Duvernoy2018} bacterial colonies form multilayers, interactions are polar
-%     \item \cite{vanGestel2015} van Gogh bundles form from interplay of 2 cell-types
-% \end{itemize}
-
-\paragraph{Computational Modeling Frameworks which support Rod-Shaped Bacteria}
-% \textbf{Requirements}
-% \begin{itemize}
-%     \item capture bendiness; do not assume rigid bodies
-%     \item model in 2D and 3D possible
-%     \item individual-based model; model single cells
-%     \item collective phenomena; describe many interacting cells, not just a single one
-% \end{itemize}
-There have been many variations in modeling rod-shaped bacteria.
-In order to provide enough flexibility and generality, we are in particular interested in models
-that can capture non-rigid rods and work both in 2D and 3D scenarios.
-Furthermore, we require that our model can describe bacteria individually and that a collective
-model is simply the propduct of multiple individual agents interacting with each other.
-In order to be able to investigate more specialized models, we need to be able to adjust and modify
-any existing model.
-
-In general, there are two classes of existing models which could be considered to fit these
-criteria: purpose-built (onf-off) solutions and computational modeling frameworks which have been
-designed to provide predefined cellular representations for various use-cases~\cite{Pleyer2023}.
-However, explicit support for rod-shaped bacteria is rather sparse.
-Biocellion has support for cylindrically-shaped agents but does not provide any mechanisms of how to
-model bendiness or flexibility of the rod~\cite{Kang2014}.
-BSim2.0 represents cells as rigid capsular cells made from a cylindrical center part and two
-half-spheres, which are placed at the ends of the cylinder to round out the
-shape~\cite{Matyjaszkiewicz2017}.
-In order to calculate interactions between cells, possible overlaps are determined and minimized,
-thus determining the position values of the next iteration step.
-However BSim does not consider bending of the rods.
-The \texttt{gro} programming language was designed to simulate the growth of colonies and cell-cell
-communication~\cite{Gutirrez2017}.
-Its “physics computation has been optimized for rigid rod-shaped bodies, like E. coli
-bacteria”~\cite{Gutirrez2017}.
-They recognize two types of forces which are acting on the cellular agents:
-Local forces which are calculated between adjacent bacteria and a global force which pushes bacteria
-outwards of the colony.
-The latter of these is a phenomenological implementation of the observed colony expansion and the
-associated central pressure with it.
-This assumption may yield incorrect results for sparsely populated cases.
-The engine is limited to 2D and does not consider polar interactions or bending of the rods.
-
-% \textbf{Computational Modeling Frameworks}
-% What has been done so far?
-% \begin{itemize}
-%     \item In general: not many frameworks which support some form of rod-shaped bacteria
-%     \item \cite{Kang2014} Biocellion has support for cylindrically-shaped agents; no
-%         bendiness/rigidity or polar interactions
-%     \item \cite{Matyjaszkiewicz2017} BSim2.0 represents cells as rigid capsular cells made from a
-%         cylindrical center part and two half-spheres, which are placed at the ends of the cylinder
-%         to round out the shape. In order to calculate interactions between cells, possible overlaps
-%         are determined and minimized, thus determining the position values of the next iteration
-%         step.
-%         BSim does not consider bending forces for individual cells or polar interactions.
-%     \item \cite{Gutirrez2017} The gro programming language was designed to simulate the growth of
-%         colonies and cell-cell communication. Its “physics computation has been optimized for rigid
-%         rod-shaped bodies, like E. coli bacteria” (Gutiérrez et al. 2017). They recognize two types
-%         of forces which are acting on the cellular agents: Local forces which are calculated between
-%         adjacent bacteria and a global force which pushes bacteria outwards of the colony. The
-%         latter of these is a phenomenological implementation of the observed colony expansion and
-%         the associated central pressure with it. This assumption may yield incorrect results for
-%         sparsely populated cases.
-%         The engine is limited to 2D and does not consider polar interactions or bending of the rods.
-% \end{itemize}
-
-\paragraph{Applications \& Parameter Estimation}
-% \begin{itemize}
-%     \item \cite{Winkle2017} Modeling mechanical interations in growing populations of rod-shaped
-%         bacteria
-%     \item \cite{Doumic2020} "A purely mechanical model with asymmetric features for early
-%         morphogenesis of rod-shaped bacteria micro-colony"
-%     \begin{itemize}
-%         \item Really good paper for referencing
-%         \item comparison of distributions; not individual agents
-%         \item no bending in model
-%         \item uses steric force (see also \cite{Trejo2013})
-%     \end{itemize}
-%     \item \cite{Grant2014} purposely-built model written in C++
-%     \begin{itemize}
-%         \item describes bacteria as collection of overlapping spheres
-%         \item spheres are coupled by non-linear springs (Euler-Bernoulli dynamic beam theory). This
-%             assumption is werid but does not alter their results
-%         \item Only model repulsive forces; no attraction, adhesion
-%     \end{itemize}
-%     \item \cite{Cho2007} only 2D model; no param estimation; based on work done in \cite{Jnsson2005}
-%     \item \cite{Storck2014} no bending; 41 parameters for various cases; only 8 parameters taken
-%         from literature values or quantified; generate growth rate from normal distribution at each
-%         integration step; stochastically equivalient for collection of bacteria but unfounded for
-%         individual bactera
-%     \item \cite{You2018}
-% \end{itemize}
-
-Cho et al.~\cite{Cho2007} extended the work done by~\cite{Jnsson2005} and construct a 2D model to
-study self-organization of colonies within restricted geometries of microfluidic devices.
-They do not provide any estimates of the parameters that were used.
-
-You et al. assumed a hard-rod model to study the emergence of order within microdomains  of
-rod-shaped bacteria~\cite{You2018}.
-They were able to show that the hydrodynamic equations of active nematic liquid crystals can serve
-as a continuum limit of their model.
-
-Winkle et al. represented cells with "[..] two acially independent cell halves that attach through a
-compressible, stiff spring [..]"~\cite{Winkle2017} and leveraged parts of the \texttt{gro}
-infrastructure.
-However, they only discuss results qualitatively without quantifying any of the used parameters.
-
-Doumic et al. constructed a model which takes into account asymmetries of the individual bacteria
-such as length, weight and friction~\cite{Doumic2020}.
-They were able to estimate some of the used parameters by comparing distributions of simulated data
-to distributions of experimental data.
-However, in doing so, no optimization optimization methods were used, but parameters were picked by
-hand in an attempt to best fit the observed results.
-Furthermore, their model is not able to describe any deformations such as bending of the rods.
-
-Grant et al. designed a custom \texttt{C++} program to simulate individual
-bacteria~\cite{Grant2014} which are represented as elastic rods of overlapping spheres.
-Their model does not include any attractive forces between agents but is able to describe bending of
-the rods in 2D and 3D.
-When it comes to the task of estimating parameters of the model, values are picked by hand in order
-to mimick the observed experimental results.
-
-Storck et al. constructed a flexible model which is based on particles and springs and thus able to
-describe a variety of cells~\cite{Storck2014}.
-Since this model incorporates many effects, it consists of $41$ parameters of which only $8$ have
-values that are taken from litearture or otherwise quantified.
-Another choice of their modeling approach was to stochastically assign growth bursts at each
-integration step of the numerical solving process.
-When generating from a normal distribution and observing on a collective level, this is
-stochastically equivalent to picking and fixing growth rates for each cell from the same
-distribution but is not identical when observing individual agents.
-
-\paragraph{The missing Link}
-It is clear that researchers are interested in studying mechanical properties of rod-like bacteria
-in order to better understand phenomena such as biofilm formation.
-The major downside of existing computational models is that there are almost no methods to properly
-quantify their parameters which diminishes their explanatory power.
-By leveraging a simple assumption about the cellular representation, this work aims to solve this
-particular problem.
+% \input{introduction_old}
 

### 2025-09-02 paper/main.tex
formulate new introduction

@@
 Furthermore, as we have discussed in our previous work~\cite{Pleyer2023}, these tools were most
-often not designed to be flexible but rather for specific purposes thus limiting their
+often not designed to be flexible but rather built for specific purposes thus limiting their
 generalizability.
@@
 When it comes to \acp{abm} however, it is often only shown that the results which have been
-generated by an \ac{abm} lead to the phenomenon which aligns with the observed biological reality.
+generated by an \ac{abm} lead to the desired phenomenon which aligns with the observed biological
+reality thus omitting to actually estimate the necessary parameters directly.
 % TODO citations
@@
 validity of the obtained results.
+
+In this work, we address these challenges using a general-purpose mechanical model for rod-shaped
+bacteria.
+Our approach provides various methods which can be combined to systematically estimate the
+parameters of our model.
+We deliberately distinguish between conceptually separate components within this framework.
+This modularity highlights a "blueprint" which can be adapted and applied to other \acp{abm}.
+
+Mechanical properties of rod-shaped bacteria have been studied
+extensively~\cite{Chatterjee1988,Takeuchi2005,IWAI2002}.
+It was found that the MreB protein is the driving force in maintaining their distinct elongated
+shape~\cite{Ursell2014}, which plays an important role in the dynamics of
+growth~\cite{Billaudeau2017}, division~\cite{Harry2001} and their collective
+behaviour~\cite{vanGestel2015}.
+Furthermore, it has been discovered that the rod can be bent thus introducing temporary elastic or
+permanent plastic deformations~\cite{Amir2014_2}.
+They are an interesting target system since their physical form provides more surface information
+compared to spherical cells.
+In addition, parameters such as growth rates have been found to vary across multiple
+cells~\cite{Koutsoumanis2013} which highlights the need to incorporate this heterogeneity on a
+single-cell level.
+
+Computational models for rod-shaped bacteria allow researchers to systematically analyze their
+properties and infer emerging phenomena~\cite{Cho2007,Jnsson2005}.
+Existing tools can describe growth, division and many effects such as colony formation and
+expansion~\cite{Matyjaszkiewicz2017,Doumic2020} but they are limited in describing the
+aforementioned plastic or elastic deformations of the rod.
+Furthermore, the literature on how to estimate the parameters of such models is very sparse.
+
+This work adresses some of these limitations.
+We use microscopic image time series as inputs to obtain information about the spatial
+configuration of the cells.
+This information can be used as initial values for a numerical simulation.
+By varying the parameters of the model and comparing the results to the microscopic data, we can
+estimate and quantify the parameters of our model.
+Our approach takes a general stance which incorporates effects that are observed within
+\textit{B.Subtilis} and \textit{E.Coli}.
 

### 2025-09-02 paper/main.tex
write abstract

@@
 \begin{abstract}
-    \textbf{TODO: ABSTRACT}
-    % This paper establishes the basis for a novel individual-based mechanical model of rod-shaped
-    % bacteria.
-    % It describes methods of how to obtain data from microscopic images such that they can be used
-    % for parameter estimation purposes.
-    % Furthermore, we show initial results for parameter estimates and an application of the model to
-    % the formation of multilayers within bacterial colonies.
+    \aclp{abm} have become popular tools in the study of complex cellular systems.
+    However, despite their popularity, it is often difficult to systematically estimate their
+    parameters and perform advanced analyses e.g. using the profile-likelihood.
+    In this work, we study the mechanical properties of rod-shaped bacteria and provide techniques
+    for estimating their parameters and performing identifiability analysis on them.
+    Our approach follows a blueprint of a modular framework which can be adapted to other research
+    questions.
+    We use time series of microscopic images as input data and incorporate advanced effects as the
+    elasticity of the rod.
+    The developed model serves as a generalized foundation to describe multiple species of
+    rod-shaped bacteria such as \textit{E.Coli} or \textit{B.Subtilis}.
 \end{abstract}

### 2025-09-05 paper/main.tex
place figures freely
add todo

@@
 
-\begin{figure}[H]
+\begin{figure}
     \centering
@@
 \label{subsec:parameter-estimation-individual-treatment}
-\begin{figure}[H]
+\begin{figure}
     \centering
@@
 function.
-For this dataset, we assumed that all interaction parameters which specify the interaction
-potential are identical across all agents but growth rates and the physical radii of the cells are
-individual and we optimize them like any other global parameter.
+For this dataset, we assumed that all parameters which specify the interaction potential are
+identical across all agents but growth rates \textbf{TODO check this} and the physical radii of the
+cells are individual and we optimize them like any other global parameter.
 We chose to optimize for 3 interaction parameters of the Mie potential $n,m,V_0$ and two parameters
@@
 % 
-% \begin{figure}[H]
+% \begin{figure}
 %     \centering
@@
 %TODO {explain profiles}
-\begin{figure}[H]
+\begin{figure}
     \centering
@@
 \end{figure}
-\begin{figure}[H]
+\begin{figure}
     \centering

### 2025-09-05 paper/main.tex
add data and code availability statement

@@
 % \end{figure}
+
+%###################################################################################################
+\section{Data and Code Availability}
+All data and code which has been used in this study is available at the github repository
+\url{https://github.com/jonaspleyer/cr\_mech\_coli}.
+It includes detailed descriptions of how to obtain the used data and bring it into a usable format.
+The package which bundles the developed funtionality is published at
+\url{https://pypi.org/project/cr-mech-coli/} from which it is readily available as a python
+dependency.
 

### 2025-09-05 paper/main.tex
switch order of setions

@@
 %###################################################################################################
-\section{Methodology}
+\section{A Computational Model for Rod-Shaped Bacteria}
+\label{sec:computational-model}
+
+The mechanical behaviour of rod-shaped bacteria is governed by a plethora of effects.
+In this work, we aim to establish a model which is abstract enough such that it can be reused for
+many types of rod-shaped bacteria but also explicit enough such that we can quantify its parameters.
+We are thus aiming to work with a set of assumptions that is as general as possible but still
+specific to rod-shaped bacteria.
+With this delicate balance in mind, it is clear that many cellular processes which might be specific
+to one particular celltype or simply not necessary to assume in order to construct our model, will
+be omitted alltogether.
+This concerns effects such as polar interactions~\cite{Duvernoy2018}, extracellular
+reactions~\cite{Wang2010} or differentiation processes~\cite{vanGestel2015,Lpez2010} and many other.
+We restrict ourselves to the aspects presented in table~\ref{table:simulation-aspects}.
+For a collection of other known effects which could be considered, we refer to the supplementary
+section~\ref{section:discussion}, and~\ref{section:supplement-more-simulation-aspects}.
+
+\paragraph{(C) Cellular}
+It was long believed that bacteria lack cytoskeletal filaments but with the end of the last century,
+it was discovered that the MreB protein fills this role since it “[..] forms an actin-like
+cytoskeleton in bacteria [..]”~\cite{Erickson2001}.
+It can polymerize and form filaments which behave similar to actin microfilaments~\cite{Dersch2020}.
+The effect of MreB on the shape of cells during growth was shown in more detail
+by~\cite{Ursell2014}.
+They found that MreB localizes specifically at regions of negative curvature and thus introduces
+heterogeneous cell growth which results in a bending force during the growth of cells.
+It also means that the cells are able to sense the curvature of the cell wall through cytoskeleteal
+MreB filaments during growth.
+\cite{Bratton2018} showed that the trans-membrane protein RodZ can modulate MreB density and
+facilitates local MreB assembly.
+This feedback mechanism will result in straight rod-shaped bacteria if not interfered from the
+external environment.
+These results were also confirmed by the studies of~\cite{Wang2010}.
+In their experiments, a single bacterium was fixed at the bottom pole while an optically trapped
+bead was placed on the upper pole.
+They could now exert a force on the free part and thus study the response of the cell under
+different conditions.
+When treating cells with A22, an antibiotic which disassembles
+MreB~\cite{IWAI2002,Gitai2005,Karczmarek2007,Bean2009}, a higher displacement with identical force
+could be measured, thus indicating that MreB increases mechanical rigidity of the cell.
+
+While some rod-shaped bacteria grow by inserting new material at the ends, \textit{E.Coli} and
+\textit{B.Subtilis} utilize a method where new material is inserted along the cylindrical part of
+the rod.
+The details of the division event (i.e. Z-Ring formation) are very intricate and are heavily
+studied as a subject on their own~\cite{Harry2001}.
+On a higher level view, the rod is split apart at a point close to the center of the cell, thus
+creating two new daughter cells.
+We know that the best predictor for when the division event occurs, is the length of the rod as
+opposed to the age of bacterium~\cite{Robert2014}.
+Furthermore, only the growth of the whole colony but also the individual growth of the bacteria is
+exponential~\cite{Amir2014,Takeuchi2005}.
+
+\paragraph{(CC) Cell-Cell Interactions \& (DC) Domain-Cell Interactions}
+A colony of bacteria is known to form biofilms (Dunne 2002). This strategy is is beneficial for the
+survival of the collective as it allows it to harness growth factors and migrate when necessary and
+also plays an important role in bacterial infections in the biomedical sector~\cite{Ong1999}.
+The formation of biofilms is only possible due to the attractive nature of bacterial interactions
+towards each other and surfaces (Berne et al. 2018). To analyze bacterial adhesion to surfaces,
+\cite{vanLoosdrecht1989} investigated the interactions of various bacteria with negatively charged
+polystyrene.
+They found that the interaction is characterized by a low Gibbs energy of $(2 - 3)kT$ per cell
+which can be described by a secondary minimum of the DLVO-theory~\cite{Verwey1947,Derjaguin1993}
+which is relevant for distances $\geq\SI{1}{\nano\meter}$.
+At these distances, the DLVO potential acts as a generic adhesive potential.
+Particles which stay in the secondary minimum are held together but even weak interactions are
+enough to redisperse them which means that this state is reversible.
+When approaching closer distances, more intricate mechanisms come into play in which the
+heterogeneity of the materials in question needs to be considered.
+\cite{Hori2010} described bacterial adhesion as a “two-phase process including an initial,
+instantaneous and reversible physical phase (phase one) and a time-dependent and irreversible
+molecular and cellular phase (phase two)”.
+
+All these considerations become even more intricate when not considering interactions of bacteria
+with surfaces but bacteria with each other.
+Gram-negative bacteria (eg. E.Coli) express outer membrane proteins (adhesins) which allow them to
+attach to various hosts (Beachey 1981, Vaca et al. 2019).
+In addition, the flagella can play an important role in the direct adhesion between
+cells~\cite{Haiko2013}.
+Thus it is not correct to blindly apply the results for adhesion to surfaces to the interactions
+between bacteria.
+
+\begin{table}[H]
+    \centering
+    \def\arraystretch{1.3}
+    \begin{tabularx}{\textwidth}{c l X}
+        &\textbf{Aspect} & \textbf{Description}\\
+        \toprule
+        &\textbf{(C) Cellular}\\
+        \midrule
+        (1) & Rod-Shaped Mechanics &
+            Rod-shaped bacteria are flexible rods which are able to freely move around (off-lattice
+            approach) \cite{Takeuchi2005,Ursell2014,Amir2014_2}.\\
+        (2) & Growth &
+            Cells grow exponentially by inserting new material either along the circular part of the
+            rod or at the tip~\cite{Robert2014,Takeuchi2005}.\\
+        (3) & Division &
+            Rod-Shaped bacteria divide in the middle of the rod into two new agents.\\
+            % Bacteria can form multilayers during their growth phase \cite{Duvernoy2018}.\\
+        (4) & Individual Parameters &
+            Parameters for individual cells are not fixed values but rather taken from a
+            distribution \cite{Koutsoumanis2013}.\\
+        &\textbf{(CC) Cell-Cell Interactions}\\
+        \midrule
+        (5) & Adhesion &
+            Bacteria attract at longer distances and adhere to each other at shorter
+            distances~\cite{Verwey1947,Trejo2013}.\\
+        \bottomrule
+    \end{tabularx}
+    \label{table:simulation-aspects}
+    \caption{
+        This table lists aspects of cellular behaviour which we consider in our computational model.
+        It is split into cellular aspects (C) which concern only an individual bacterium and
+        interactions between cells (CC).
+        These aspects are generalized in order to fit to as many types of rod-shaped bacteria as
+        possible.
+        These are the assumptions for the basic model that we are presenting here.
+        In order for our methods to work, we are really only required to satisfy the assumption of
+        our spatial representation (see section~\ref{subsec:spatial-representation}) and thus, these
+        assumptions can also be loosened if required.
+    }
+\end{table}
+
+To model the spatial mechanics of elongated bacteria \cite{Billaudeau2017}, we represent them as a
+collection of auxiliary vertices $\{\vec{x}_i\}$ which are connected by springs in
+ascending order.
+Furthermore, we assume that the cells are flexible described by their rigidity.
+A force $\vec{F}$ interacting between cellular agents determines the radius (thickness) of the
+rods and an attractive component can model adhesion between cells.
+
 % --------------------------------------------------------------------------------------------------
-\subsection{Spatial Representation of Rod-Shaped Bacteria}
+\subsection{Spatial Representation}
 \label{subsec:spatial-representation}
@@
 \end{figure}
+
+% --------------------------------------------------------------------------------------------------
+\subsection{Mechanics}
+\label{subsection:mechanical-model-mechanics}
+
+We use principles from discrete differential geometry~\cite{bobenko2008discrete} to describe the
+stiffness and flexibility of the bacterial rod.
+As explained previously, MReB will counteract the bending of the rod, resulting in a stiffening
+force.
+In principle, this stress will be dependent on the position inside the cell~\cite{Chatterjee1988}
+but in order to construct a model which can be parametrized and by method of Occam's Razor, we chose
+to model the bacterium as a bending beam, thus utilizing the simplest solution which can still
+produce the desired effects.
+Such a beam can be represented by a collection of vertices $\vec{x}_i$ which are connected by a
+springs of lengths $l_i$.
+We use the notation $\vec{c}_i:=\vec{x}_i-\vec{x}_{i-1}$ with which the resulting force acting
+between two vertices if given by
+
+\begin{align}
+    \vec{F}_{i,\text{springs}} =
+        &-\gamma\left(1 - \frac{l}{\left|\vec{c}_i\right|}\right)
+        \vec{c}_i\\
+        &+ \gamma\left(1 - \frac{l}{\left|\vec{c}_{i+1}\right|}\right)
+        \vec{c}_{i+1}.
+\end{align}
+
+In addition to springs between individual vertices $\vec{x}_i$, we assume that each angle at a
+vertex between edges is subject to a force indiced by curvature.
+% We assume that we can model the mechanical properties of the bacterium as an elastic rod.
+Given the angle $\alpha_i = \sphericalangle(\vec{c}_{i-1},\vec{c}_i)$ between adjacent edges,
+the bending force is proportional to the curvature $\kappa_i$ at each vertex
+$\kappa_i = 2\tan(\alpha_i/2)$.
+The resulting force acts along the angle bisector which can be calculated from the edge vectors.
+Figure~\ref{fig:mechanics-bacterium} shows an elongated rod represented by $6$ vertices.
+The forces acting on vertices $\vec{x}_i,\vec{x}_{i-1},\vec{x}_{i+1}$ are given by
+
+\begin{align}
+    \vec{F}_{i,\text{curvature}} &= \eta\kappa_i
+        \frac{\vec{c}_i - \vec{c}_{i+1}}{|\vec{c}_i-\vec{c}_{i+1}|}\\
+    \vec{F}_{i-1,\text{curvature}} &= -\frac{1}{2}\vec{F}_{i,\text{curvature}}\\
+    \vec{F}_{i+1,\text{curvature}} &= -\frac{1}{2}\vec{F}_{i,\text{curvature}}
+\end{align}
+
+where $\eta_i$ is the rigidity at vertex $\vec{x}_i$ (see Figure~\ref{fig:mechanics-bacterium}).
+We can see that the curvature force does not move the overall center of the rod in space.
+The total force is the sum of external and interal forces.
+
+\begin{equation}
+    \vec{F}_{i,\text{total}} = \vec{F}_{i,\text{springs}}+ \vec{F}_{i,\text{curvature}}
+        + \vec{F}_{i,\text{external}}
+\end{equation}
+
+and are integrated via
+
+\begin{align}
+    \partial_t^2 \vec{x}_i &= \partial_t\vec{x}_i + \sqrt{2D}\vec{\xi}_i\\
+    \partial_t\vec{x}_i &= \vec{F}_{i,\text{total}} - \lambda \partial_t\vec{x}_i
+\end{align}
+
+where $D$ is the diffusion constant and $\vec{\xi}$ is the wiener process.
+Together they determine the stochastic motion of the individual vertices which can arise from
+recurring collisions with small particles of the surrounding medium or other minor fluctuations.
+
+% --------------------------------------------------------------------------------------------------
+\subsection{Physical Interactions}
+\label{subsection:mechanical-model-interactions}
+
+When calculating forces acting between the cells, we can use a simplified model to circumvent the
+numerically expensive integration over the complete length of the rod.
+Given a vertex $\vec{x}_i$ on one cell, we calculate the closest point $\vec{p}$ on the polygonal
+line given by the vertices $\{\vec{w}_j\}$ of the interacting cell.
+Furthermore we determine the value $q\in[0,1]$ such that $\vec{p} = (1-q)\vec{w}_j + q\vec{w}_{j+1}$
+for some specific $j$.
+The force is then calculated between the points $\vec{x}_i$ and $\vec{p}_i$ and acts on
+the vertex $\vec{w}_i,\vec{w}_{i+1}$ with relative strength $(1-q)$ and $q$.
+
+\begin{equation}
+    \vec{F}_{i,\text{External}} = \vec{F}(\vec{x}_i,\vec{p})
+    \label{eq:cell-force-external}
+\end{equation}
+
+The specific choice of interaction potential plays of course a crucial role for the dynamics of the
+agents.
+But we cannot expect to simply know the shape and parameters of this potential.
+Thus, we would like to experiment with various interaction potentials in order to quantify their
+ability to produce accurate results.
+We utilize two distinct potentials that are both able to capture the essence of our interactions by
+providing a short-range repulsive part and a long-ranged attractive component.
+
+\paragraph{Morse Potential}
+The morse potential~\cite{Morse1929} is a common example of numerically inexpensive interaction
+potential.
+It was originally designed to describe the interaction of diatomic molecules but has since found use
+in other areas~\cite{Breitwieser2021}.
+It contains a repulsive and attractive part and is bound from above which makes it particularly
+suitable for numerical simulations since unbound potentials can lead to large numerical deviations
+thus compromising the prediction of the model.
+The potential is of the form
+\begin{equation}
+    V(r) = V_0\left(1-e^{-\lambda(r-(R_1+R_2))}\right)^2
+\end{equation}
+and has 3 parameters.
+The overall potential strength $V_0$, the radii of the particles $R_1,R_2$ and the potential
+stiffness $\lambda$ which controls the rate of change as one travels along the radial axis.
+We employ a spatial cutoff which sets the interaction strength to $0$ when exceeding a distance
+$r\geq\xi$.
+
+\paragraph{Mie Potential}
+The Mie~\cite{Mie1903} potential is the generalization of the well-known
+Lennard-Jones~\cite{Jones1924} potential.
+It was also designed to describe the interactions of molecules but due to the flexible nature of its
+parameters, can be easily adapted to other situations.
+It is able to describe a wider range of shapes compared to the Morse potential but is also stiffer
+and not bound from above.
+In our simulation, we enforce an upper bound to guarantee numerical regularity.
+It has 4 parameters and is given by
+\begin{align}
+    U(r) &= C\epsilon\left[ \left(\frac{\sigma}{r}\right)^n -
+        \left(\frac{\sigma}{r}\right)^m\right]\\
+    C &= \frac{n}{n-m}\left(\frac{n}{m}\right)^{\frac{n}{n-m}}\\
+    V(r) &= \min(U(r), \beta)\theta(r-\zeta)
+\end{align}
+where $C$ is a constant that depends on the two exponents $n,m$ that describe the overall shape of
+the potential.
+The strength is given by $\epsilon$ and the radius is encoded in $\sigma =
+(m/n)^{1/(n-m)}(R_1+R_2)$ where $R_1,R_2$ are the radii of the two interacting particles.
+
+% --------------------------------------------------------------------------------------------------
+\subsection{Cell-Cycle}
+\label{subsection:mechanical-model-cycle}
+
+To simulate proliferation, we introduce a growth term for the spring lengths $l_i$.
+In most of our simulations, we choose to use identical lengths for all segments (i.e. $l_i=l$).
+However, in order to generalize our approach we assume that individual lengths of segments could
+vary.
+We reduce the growth rate with increasing number of neighbors $N$.
+This behaviour can be toggled on or off thus removing the additional term in brackets of
+equation~\ref{eq:growth-rod-lengths}.
+
+\begin{equation}
+    \partial_t l_i = \mu\left(1-\frac{\text{min}(n,N)}{N}\right)l_i
+    \label{eq:growth-rod-lengths}
+\end{equation}
+
+Note that the growth rate is proportional to the rod length $l$ which leads to exponential growth.
+This process will increase the length of the cell indefenitely unless we impose a condition for the
+division event.
+We define a fixed length threshold for the total length of the cell at which it divides.
+To construct a new cell, we cannot simply copy the existing one twice, but we also need to adjust
+internal parameters in the process.
+The following actions need to be taken for the old and new agent.
+We first calculate the new spring lengths of the two resulting rods.
+\begin{equation}
+    \tilde{l} = l_i\left(\frac{1}{2} - \frac{r}{\sum\limits_i l_i}\right)
+    \label{eq:growth-rod-new-lengths}
+\end{equation}
+Afterwards, we start at one endpoint $\vec{x}_0$ and travel along the polygon defined by the
+bacterial rod.
+We stop when we have reached a length that matches the length of the new spring length $\tilde{l}$.
+From here, we continue and iterate this procedure until we have found all new vertices $\vec{w}_i$.
+Equation~\ref{eq:growth-rod-new-lengths} ensures that that the tips of the two new rods are touching
+right after the division event.
+
+% --------------------------------------------------------------------------------------------------
+\subsection{Numerical Implementation}
+
+We implemented the described model in \texttt{cellular\_raza}~\cite{Pleyer2025}.
+Doing so has many benefits.
+Not only, was a more simpler model already available which could be adapted to our needs.
+But furthermore, due to the nature of \texttt{cellular\_raza} in that it does not assume any
+particular cellular representation, we are free to change the underlying model at any time, even
+making fundamental changes.
+It is an \ac{abm} which is able to natively track cells through the course of a numerical simulation
+and also keeps track of parent/daughter relationships.
+This allows us to to experiment with different model implementations and still retain performance.
+By leveraging the strong safety guarantees provided by the Rust programming language, we eliminate a
+whole class of problems from the start.
+% TODO CITATION
+In addition, we are using the \texttt{pyo3} and \texttt{maturin} crates which allow us to easily
+generate python bindings and thus interoperate between the two ecosystems.
+While this particular model is more complex due to the nature of non-trivial (asymmetric)
+interactions, we were still able to properly utilize the speedup from executing simulations in
+parallalel using the built-in parallelization methods of \texttt{cellular\_raza} (see
+Supplementary Material, Figure~\ref{fig:performance}).
+
+%###################################################################################################
+\section{Methodology}
+These sections introduce methods which can be combined in order to estimate the parameters of our
+computational model.
+It has to be noted that they only depend on the spatial representation described in
+section~\ref{subsec:spatial-representation}.
+Therefore, they can be applied to a whole class of mechanical models which could make different
+choices with respect to the cellular details.
 
@@
 should be preferred.
-
-%###################################################################################################
-\section{A Computational Model for Rod-Shaped Bacteria}
-\label{sec:computational-model}
-
-The mechanical behaviour of rod-shaped bacteria is governed by a plethora of effects.
-In this work, we aim to establish a model which is abstract enough such that it can be reused for
-many types of rod-shaped bacteria but also explicit enough such that we can quantify its parameters.
-We are thus aiming to work with a set of assumptions that is as general as possible but still
-specific to rod-shaped bacteria.
-With this delicate balance in mind, it is clear that many cellular processes which might be specific
-to one particular celltype or simply not necessary to assume in order to construct our model, will
-be omitted alltogether.
-This concerns effects such as polar interactions~\cite{Duvernoy2018}, extracellular
-reactions~\cite{Wang2010} or differentiation processes~\cite{vanGestel2015,Lpez2010} and many other.
-We restrict ourselves to the aspects presented in table~\ref{table:simulation-aspects}.
-For a collection of other known effects which could be considered, we refer to the supplementary
-section~\ref{section:discussion}, and~\ref{section:supplement-more-simulation-aspects}.
-
-\paragraph{(C) Cellular}
-The key component of this model consists of the cellular representation.
-
-It was long believed that bacteria lack cytoskeletal filaments but with the end of the last century,
-it was discovered that the MreB protein fills this role since it “[..] forms an actin-like
-cytoskeleton in bacteria [..]”~\cite{Erickson2001}.
-It can polymerize and form filaments which behave similar to actin microfilaments~\cite{Dersch2020}.
-The effect of MreB on the shape of cells during growth was shown in more detail
-by~\cite{Ursell2014}.
-They found that MreB localizes specifically at regions of negative curvature and thus introduces
-heterogeneous cell growth which results in a bending force during the growth of cells.
-It also means that the cells are able to sense the curvature of the cell wall through cytoskeleteal
-MreB filaments during growth.
-\cite{Bratton2018} showed that the trans-membrane protein RodZ can modulate MreB density and
-facilitates local MreB assembly.
-This feedback mechanism will result in straight rod-shaped bacteria if not interfered from the
-external environment.
-These results were also confirmed by the studies of~\cite{Wang2010}.
-In their experiments, a single bacterium was fixed at the bottom pole while an optically trapped
-bead was placed on the upper pole.
-They could now exert a force on the free part and thus study the response of the cell under
-different conditions.
-When treating cells with A22, an antibiotic which disassembles
-MreB~\cite{IWAI2002,Gitai2005,Karczmarek2007,Bean2009}, a higher displacement with identical force
-could be measured, thus indicating that MreB increases mechanical rigidity of the cell.
-
-While some rod-shaped bacteria grow by inserting new material at the ends, \textit{E.Coli} and
-\textit{B.Subtilis} utilize a method where new material is inserted along the cylindrical part of
-the rod.
-The details of the division event (i.e. Z-Ring formation) are very intricate and are heavily
-studied as a subject on their own~\cite{Harry2001}.
-On a higher level view, the rod is split apart at a point close to the center of the cell, thus
-creating two new daughter cells.
-We know that the best predictor for when the division event occurs, is the length of the rod as
-opposed to the age of bacterium~\cite{Robert2014}.
-Furthermore, only the growth of the whole colony but also the individual growth of the bacteria is
-exponential~\cite{Amir2014,Takeuchi2005}.
-
-\paragraph{(CC) Cell-Cell Interactions \& (DC) Domain-Cell Interactions}
-A colony of bacteria is known to form biofilms (Dunne 2002). This strategy is is beneficial for the
-survival of the collective as it allows it to harness growth factors and migrate when necessary and
-also plays an important role in bacterial infections in the biomedical sector~\cite{Ong1999}.
-The formation of biofilms is only possible due to the attractive nature of bacterial interactions
-towards each other and surfaces (Berne et al. 2018). To analyze bacterial adhesion to surfaces,
-\cite{vanLoosdrecht1989} investigated the interactions of various bacteria with negatively charged
-polystyrene.
-They found that the interaction is characterized by a low Gibbs energy of $(2 - 3)kT$ per cell
-which can be described by a secondary minimum of the DLVO-theory~\cite{Verwey1947,Derjaguin1993}
-which is relevant for distances $\geq\SI{1}{\nano\meter}$.
-At these distances, the DLVO potential acts as a generic adhesive potential.
-Particles which stay in the secondary minimum are held together but even weak interactions are
-enough to redisperse them which means that this state is reversible.
-When approaching closer distances, more intricate mechanisms come into play in which the
-heterogeneity of the materials in question needs to be considered.
-\cite{Hori2010} described bacterial adhesion as a “two-phase process including an initial,
-instantaneous and reversible physical phase (phase one) and a time-dependent and irreversible
-molecular and cellular phase (phase two)”.
-
-All these considerations become even more intricate when not considering interactions of bacteria
-with surfaces but bacteria with each other.
-Gram-negative bacteria (eg. E.Coli) express outer membrane proteins (adhesins) which allow them to
-attach to various hosts (Beachey 1981, Vaca et al. 2019).
-In addition, the flagella can play an important role in the direct adhesion between
-cells~\cite{Haiko2013}.
-Thus it is not correct to blindly apply the results for adhesion to surfaces to the interactions
-between bacteria.
-
-\begin{table}[H]
-    \centering
-    \def\arraystretch{1.3}
-    \begin{tabularx}{\textwidth}{c l X}
-        &\textbf{Aspect} & \textbf{Description}\\
-        \toprule
-        &\textbf{(C) Cellular}\\
-        \midrule
-        (1) & Rod-Shaped Mechanics &
-            Rod-shaped bacteria are flexible rods which are able to freely move around (off-lattice
-            approach) \cite{Takeuchi2005,Ursell2014,Amir2014_2}.\\
-        (2) & Growth &
-            Cells grow exponentially by inserting new material either along the circular part of the
-            rod or at the tip~\cite{Robert2014,Takeuchi2005}.\\
-        (3) & Division &
-            Rod-Shaped bacteria divide in the middle of the rod into two new agents.\\
-            % Bacteria can form multilayers during their growth phase \cite{Duvernoy2018}.\\
-        (4) & Individual Parameters &
-            Parameters for individual cells are not fixed values but rather taken from a
-            distribution \cite{Koutsoumanis2013}.\\
-        &\textbf{(CC) Cell-Cell Interactions}\\
-        \midrule
-        (5) & Adhesion &
-            Bacteria attract at longer distances and adhere to each other at shorter
-            distances~\cite{Verwey1947,Trejo2013}.\\
-        \bottomrule
-    \end{tabularx}
-    \label{table:simulation-aspects}
-    \caption{
-        This table lists aspects of cellular behaviour which we consider in our computational model.
-        It is split into cellular aspects (C) which concern only an individual bacterium and
-        interactions between cells (CC).
-        These aspects are generalized in order to fit to as many types of rod-shaped bacteria as
-        possible.
-        These are the assumptions for the basic model that we are presenting here.
-        In order for our methods to work, we are really only required to satisfy the assumption of
-        our spatial representation (see section~\ref{subsec:spatial-representation}) and thus, these
-        assumptions can also be loosened if required.
-    }
-\end{table}
-
-To model the spatial mechanics of elongated bacteria \cite{Billaudeau2017}, we represent them as a
-collection of auxiliary vertices $\{\vec{x}_i\}$ which are connected by springs in
-ascending order.
-Furthermore, we assume that the cells are flexible described by their rigidity.
-A force $\vec{F}$ interacting between cellular agents determines the radius (thickness) of the
-rods and an attractive component can model adhesion between cells.
-
-% --------------------------------------------------------------------------------------------------
-\subsection{Mechanics}
-\label{subsection:mechanical-model-mechanics}
-
-We use principles from discrete differential geometry~\cite{bobenko2008discrete} to describe the
-stiffness and flexibility of the bacterial rod.
-As explained previously, MReB will counteract the bending of the rod, resulting in a stiffening
-force.
-In principle, this stress will be dependent on the position inside the cell~\cite{Chatterjee1988}
-but in order to construct a model which can be parametrized and by method of Occam's Razor, we chose
-to model the bacterium as a bending beam, thus utilizing the simplest solution which can still
-produce the desired effects.
-Such a beam can be represented by a collection of vertices $\vec{x}_i$ which are connected by a
-springs of lengths $l_i$.
-We use the notation $\vec{c}_i:=\vec{x}_i-\vec{x}_{i-1}$ with which the resulting force acting
-between two vertices if given by
-
-\begin{align}
-    \vec{F}_{i,\text{springs}} =
-        &-\gamma\left(1 - \frac{l}{\left|\vec{c}_i\right|}\right)
-        \vec{c}_i\\
-        &+ \gamma\left(1 - \frac{l}{\left|\vec{c}_{i+1}\right|}\right)
-        \vec{c}_{i+1}.
-\end{align}
-
-In addition to springs between individual vertices $\vec{x}_i$, we assume that each angle at a
-vertex between edges is subject to a force indiced by curvature.
-% We assume that we can model the mechanical properties of the bacterium as an elastic rod.
-Given the angle $\alpha_i = \sphericalangle(\vec{c}_{i-1},\vec{c}_i)$ between adjacent edges,
-the bending force is proportional to the curvature $\kappa_i$ at each vertex
-$\kappa_i = 2\tan(\alpha_i/2)$.
-The resulting force acts along the angle bisector which can be calculated from the edge vectors.
-Figure~\ref{fig:mechanics-bacterium} shows an elongated rod represented by $6$ vertices.
-The forces acting on vertices $\vec{x}_i,\vec{x}_{i-1},\vec{x}_{i+1}$ are given by
-
-\begin{align}
-    \vec{F}_{i,\text{curvature}} &= \eta\kappa_i
-        \frac{\vec{c}_i - \vec{c}_{i+1}}{|\vec{c}_i-\vec{c}_{i+1}|}\\
-    \vec{F}_{i-1,\text{curvature}} &= -\frac{1}{2}\vec{F}_{i,\text{curvature}}\\
-    \vec{F}_{i+1,\text{curvature}} &= -\frac{1}{2}\vec{F}_{i,\text{curvature}}
-\end{align}
-
-where $\eta_i$ is the rigidity at vertex $\vec{x}_i$ (see Figure~\ref{fig:mechanics-bacterium}).
-We can see that the curvature force does not move the overall center of the rod in space.
-The total force is the sum of external and interal forces.
-
-\begin{equation}
-    \vec{F}_{i,\text{total}} = \vec{F}_{i,\text{springs}}+ \vec{F}_{i,\text{curvature}}
-        + \vec{F}_{i,\text{external}}
-\end{equation}
-
-and are integrated via
-
-\begin{align}
-    \partial_t^2 \vec{x}_i &= \partial_t\vec{x}_i + \sqrt{2D}\vec{\xi}_i\\
-    \partial_t\vec{x}_i &= \vec{F}_{i,\text{total}} - \lambda \partial_t\vec{x}_i
-\end{align}
-
-where $D$ is the diffusion constant and $\vec{\xi}$ is the wiener process.
-Together they determine the stochastic motion of the individual vertices which can arise from
-recurring collisions with small particles of the surrounding medium or other minor fluctuations.
-
-% --------------------------------------------------------------------------------------------------
-\subsection{Physical Interactions}
-\label{subsection:mechanical-model-interactions}
-
-When calculating forces acting between the cells, we can use a simplified model to circumvent the
-numerically expensive integration over the complete length of the rod.
-Given a vertex $\vec{x}_i$ on one cell, we calculate the closest point $\vec{p}$ on the polygonal
-line given by the vertices $\{\vec{w}_j\}$ of the interacting cell.
-Furthermore we determine the value $q\in[0,1]$ such that $\vec{p} = (1-q)\vec{w}_j + q\vec{w}_{j+1}$
-for some specific $j$.
-The force is then calculated between the points $\vec{x}_i$ and $\vec{p}_i$ and acts on
-the vertex $\vec{w}_i,\vec{w}_{i+1}$ with relative strength $(1-q)$ and $q$.
-
-\begin{equation}
-    \vec{F}_{i,\text{External}} = \vec{F}(\vec{x}_i,\vec{p})
-    \label{eq:cell-force-external}
-\end{equation}
-
-The specific choice of interaction potential plays of course a crucial role for the dynamics of the
-agents.
-But we cannot expect to simply know the shape and parameters of this potential.
-Thus, we would like to experiment with various interaction potentials in order to quantify their
-ability to produce accurate results.
-We utilize two distinct potentials that are both able to capture the essence of our interactions by
-providing a short-range repulsive part and a long-ranged attractive component.
-
-\paragraph{Morse Potential}
-The morse potential~\cite{Morse1929} is a common example of numerically inexpensive interaction
-potential.
-It was originally designed to describe the interaction of diatomic molecules but has since found use
-in other areas~\cite{Breitwieser2021}.
-It contains a repulsive and attractive part and is bound from above which makes it particularly
-suitable for numerical simulations since unbound potentials can lead to large numerical deviations
-thus compromising the prediction of the model.
-The potential is of the form
-\begin{equation}
-    V(r) = V_0\left(1-e^{-\lambda(r-(R_1+R_2))}\right)^2
-\end{equation}
-and has 3 parameters.
-The overall potential strength $V_0$, the radii of the particles $R_1,R_2$ and the potential
-stiffness $\lambda$ which controls the rate of change as one travels along the radial axis.
-We employ a spatial cutoff which sets the interaction strength to $0$ when exceeding a distance
-$r\geq\xi$.
-
-\paragraph{Mie Potential}
-The Mie~\cite{Mie1903} potential is the generalization of the well-known
-Lennard-Jones~\cite{Jones1924} potential.
-It was also designed to describe the interactions of molecules but due to the flexible nature of its
-parameters, can be easily adapted to other situations.
-It is able to describe a wider range of shapes compared to the Morse potential but is also stiffer
-and not bound from above.
-In our simulation, we enforce an upper bound to guarantee numerical regularity.
-It has 4 parameters and is given by
-\begin{align}
-    U(r) &= C\epsilon\left[ \left(\frac{\sigma}{r}\right)^n -
-        \left(\frac{\sigma}{r}\right)^m\right]\\
-    C &= \frac{n}{n-m}\left(\frac{n}{m}\right)^{\frac{n}{n-m}}\\
-    V(r) &= \min(U(r), \beta)\theta(r-\zeta)
-\end{align}
-where $C$ is a constant that depends on the two exponents $n,m$ that describe the overall shape of
-the potential.
-The strength is given by $\epsilon$ and the radius is encoded in $\sigma =
-(m/n)^{1/(n-m)}(R_1+R_2)$ where $R_1,R_2$ are the radii of the two interacting particles.
-
-% --------------------------------------------------------------------------------------------------
-\subsection{Cell-Cycle}
-\label{subsection:mechanical-model-cycle}
-
-To simulate proliferation, we introduce a growth term for the spring lengths $l_i$.
-In most of our simulations, we choose to use identical lengths for all segments (i.e. $l_i=l$).
-However, in order to generalize our approach we assume that individual lengths of segments could
-vary.
-We reduce the growth rate with increasing number of neighbors $N$.
-This behaviour can be toggled on or off thus removing the additional term in brackets of
-equation~\ref{eq:growth-rod-lengths}.
-
-\begin{equation}
-    \partial_t l_i = \mu\left(1-\frac{\text{min}(n,N)}{N}\right)l_i
-    \label{eq:growth-rod-lengths}
-\end{equation}
-
-Note that the growth rate is proportional to the rod length $l$ which leads to exponential growth.
-This process will increase the length of the cell indefenitely unless we impose a condition for the
-division event.
-We define a fixed length threshold for the total length of the cell at which it divides.
-To construct a new cell, we cannot simply copy the existing one twice, but we also need to adjust
-internal parameters in the process.
-The following actions need to be taken for the old and new agent.
-We first calculate the new spring lengths of the two resulting rods.
-\begin{equation}
-    \tilde{l} = l_i\left(\frac{1}{2} - \frac{r}{\sum\limits_i l_i}\right)
-    \label{eq:growth-rod-new-lengths}
-\end{equation}
-Afterwards, we start at one endpoint $\vec{x}_0$ and travel along the polygon defined by the
-bacterial rod.
-We stop when we have reached a length that matches the length of the new spring length $\tilde{l}$.
-From here, we continue and iterate this procedure until we have found all new vertices $\vec{w}_i$.
-Equation~\ref{eq:growth-rod-new-lengths} ensures that that the tips of the two new rods are touching
-right after the division event.
-
-% --------------------------------------------------------------------------------------------------
-\subsection{Numerical Implementation}
-
-We implemented the described model in \texttt{cellular\_raza}~\cite{Pleyer2025}.
-Doing so has many benefits.
-Not only, was a more simpler model already available which could be adapted to our needs.
-But furthermore, due to the nature of \texttt{cellular\_raza} in that it does not assume any
-particular cellular representation, we are free to change the underlying model at any time, even
-making fundamental changes.
-It is an \ac{abm} which is able to natively track cells through the course of a numerical simulation
-and also keeps track of parent/daughter relationships.
-This allows us to to experiment with different model implementations and still retain performance.
-By leveraging the strong safety guarantees provided by the Rust programming language, we eliminate a
-whole class of problems from the start.
-% TODO CITATION
-In addition, we are using the \texttt{pyo3} and \texttt{maturin} crates which allow us to easily
-generate python bindings and thus interoperate between the two ecosystems.
-While this particular model is more complex due to the nature of non-trivial (asymmetric)
-interactions, we were still able to properly utilize the speedup from executing simulations in
-parallalel using the built-in parallelization methods of \texttt{cellular\_raza} (see
-Supplementary Material, Figure~\ref{fig:performance}).
 

### 2025-09-05 paper/main.tex
format supplement subsection headers
adjust some formulations

@@
 
-Mechanical properties of rod-shaped bacteria have been studied
-extensively~\cite{Chatterjee1988,Takeuchi2005,IWAI2002}.
+Mechanical properties of rod-shaped bacteria have been studied in great detail
+experimentally~\cite{Chatterjee1988,Takeuchi2005,IWAI2002}.
 It was found that the MreB protein is the driving force in maintaining their distinct elongated
@@
 aforementioned plastic or elastic deformations of the rod.
-Furthermore, the literature on how to estimate the parameters of such models is very sparse.
+Furthermore, the literature on how to estimate the parameters of such models directly from data is
+very sparse.
 
@@
 Figure~\ref{fig:flowchart-project-structure} shows a high-level overview of all components.
-We assume that bacterial rods can be represented by a colleciton of vertices $\vec{x}_i$.
+We assume that bacterial rods can be represented by a colleciton of vertices $\vec{x}_i$ (see
+section~\ref{subsec:spatial-representation}).
 This assumption alone is enough to enable a cascade of other components.
@@
 reactions~\cite{Wang2010} or differentiation processes~\cite{vanGestel2015,Lpez2010} and many other.
+When describing our system, we distinguish between cellular aspects which ocurr within a single
+cell and cell-cell interactions between numerous bacteria.
 We restrict ourselves to the aspects presented in table~\ref{table:simulation-aspects}.
@@
 Doing so has many benefits.
-Not only, was a more simpler model already available which could be adapted to our needs.
-But furthermore, due to the nature of \texttt{cellular\_raza} in that it does not assume any
-particular cellular representation, we are free to change the underlying model at any time, even
-making fundamental changes.
-It is an \ac{abm} which is able to natively track cells through the course of a numerical simulation
-and also keeps track of parent/daughter relationships.
+Not only, was a more simpler model already available which we were able to adapt to our needs.
+But by design \texttt{cellular\_raza} aims to provide researchers with a great amount of flexibility
+and therefore does not assume any particular cellular representation, thus enabling us to make
+fundamental changes.
 This allows us to to experiment with different model implementations and still retain performance.
-By leveraging the strong safety guarantees provided by the Rust programming language, we eliminate a
-whole class of problems from the start.
-% TODO CITATION
-In addition, we are using the \texttt{pyo3} and \texttt{maturin} crates which allow us to easily
-generate python bindings and thus interoperate between the two ecosystems.
-While this particular model is more complex due to the nature of non-trivial (asymmetric)
-interactions, we were still able to properly utilize the speedup from executing simulations in
-parallalel using the built-in parallelization methods of \texttt{cellular\_raza} (see
-Supplementary Material, Figure~\ref{fig:performance}).
+Furthermore, it is an \ac{abm} which is able to natively track cells through the course of a
+numerical simulation and also keeps track of mother/daughter relationships.
+On the programming side, we are using the \texttt{pyo3} and \texttt{maturin} crates which allow us
+to easily generate Python bindings and enable us to interoperate between the two ecosystems.
+By using \texttt{cellular\_raza}, we can make use of multiple threads for running a single numerical
+simulation, using the built-in parallelization methods (see Supplementary Material,
+Figure~\ref{fig:performance}).
 
@@
 
-For visualization we use pyvista, which provides high-level wrappers around the popular VTK
-library~\cite{vtkBook,Sullivan2019}.
+To visualize the spatial configuration of our results we use pyvista, which provides high-level
+wrappers around the popular VTK library~\cite{vtkBook,Sullivan2019}.
 The first stage of our visualization is to generate 3D meshes for each bacterium, which we consider
@@
 
-To generate masks, we assign unique colors to every mesh of the respective cellular agent.
-This mapping is unique, ensuring no two agents will use identical colors.
+To generate masks, we assign unique colors to every mesh of the respective cellular agent, ensuring
+no two agents will use identical colors.
 We then use a parallel projection and remove any existing light sources in order to obtain a solid
@@
 In doing so, we have exact knowledge about the position of our agents since we can simply obtain
-this information from our model.
-But on the other hand we can also try to extract this position from the generated cell-masks.
+this information from our computational model.
+But on the other hand we can also try to extract these positions from the generated cell-masks.
 By comparing the differences between them, we get a good estimate of how well our extraction
@@
 %###################################################################################################
-\section*{Supplementary Material}
-
-\renewcommand{\thesection}{S\arabic{section}}
+\renewcommand{\thesection}{}
+\renewcommand{\thesubsection}{S\arabic{subsection}}
+\section{Supplementary Material}
 
 %###################################################################################################
-\section{Data Preparation}
+\subsection{Data Preparation}
 \begin{enumerate}
@@
 %###################################################################################################
-\section{Parameter Estimation}
+\subsection{Parameter Estimation}
 \label{sec:supplement-parameter-profiles}
@@
 %###################################################################################################
-\section{More Simulation Aspects}
+\subsection{More Simulation Aspects}
 \label{section:supplement-more-simulation-aspects}
@@
 %###################################################################################################
-\section{Performance}
+\subsection{Performance}
 
@@
 %###################################################################################################
-\section{Multilayer Predictions}
+\subsection{Multilayer Predictions}
 % TODO

### 2025-09-05 paper/main.tex
mark as todo
add bullet points
remove multilayer predictions

@@
 %###################################################################################################
-\subsection{Data Preparation}
+\subsection{Data Preparation (TODO)}
 \begin{enumerate}
@@
 %###################################################################################################
-\subsection{Parameter Estimation}
+\subsection{Parameter Estimation Results (TODO)}
 \label{sec:supplement-parameter-profiles}
@@
 %###################################################################################################
-\subsection{More Simulation Aspects}
+\subsection{More Simulation Aspects (TODO)}
 \label{section:supplement-more-simulation-aspects}
@@
 
+\begin{itemize}
+    \item new class of aspects: (DC) Domain-Cell Interactions
+    \item List which aspects are here additionally
+    \item discuss which of them might need specific models and can not be paired with others
+    \item which can be bundled together without problems?
+\end{itemize}
+
 %###################################################################################################
-\subsection{Performance}
+\subsection{Performance (TODO)}
 
@@
 %###################################################################################################
-\subsection{Multilayer Predictions}
-% TODO
-
-\begin{itemize}
-    \item Friction force against surface. perpendicular to motion; proportional to velocity.
-    \item gel pressure coming from experiment (simply downwards force $F_z$)
-\end{itemize}
-
-\begin{itemize}
-    \item Do we need this in this publication?
-    \item Maybe this is out of scope as this is a more methodological paper.
-    \item If we want to omit it, we might also opt to change the model description in the beginning
-        and remove some parts of the external forces discussion.
-\end{itemize}
-
-\cite{You2019}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.49\textwidth]
-        {docs/source/_static/scripts/crm_multilayer/multilayer-time-evolution.pdf}\\
-    \includegraphics[width=0.49\textwidth]
-        {docs/source/_static/scripts/crm_multilayer/colony-height-vs-gel_pressure.pdf}
-    \includegraphics[width=0.49\textwidth]
-        {docs/source/_static/scripts/crm_multilayer/colony-height-vs-time.pdf}
-    \caption{TODO}
-    \label{fig:multilayer}
-\end{figure}
+% \subsection{Multilayer Predictions}
+% % TODO
+%
+% \begin{itemize}
+%     \item Friction force against surface. perpendicular to motion; proportional to velocity.
+%     \item gel pressure coming from experiment (simply downwards force $F_z$)
+% \end{itemize}
+%
+% \begin{itemize}
+%     \item Do we need this in this publication?
+%     \item Maybe this is out of scope as this is a more methodological paper.
+%     \item If we want to omit it, we might also opt to change the model description in the beginning
+%         and remove some parts of the external forces discussion.
+% \end{itemize}
+%
+% \cite{You2019}
+%
+% \begin{figure}
+%     \centering
+%     \includegraphics[width=0.49\textwidth]
+%         {docs/source/_static/scripts/crm_multilayer/multilayer-time-evolution.pdf}\\
+%     \includegraphics[width=0.49\textwidth]
+%         {docs/source/_static/scripts/crm_multilayer/colony-height-vs-gel_pressure.pdf}
+%     \includegraphics[width=0.49\textwidth]
+%         {docs/source/_static/scripts/crm_multilayer/colony-height-vs-time.pdf}
+%     \caption{TODO}
+%     \label{fig:multilayer}
+% \end{figure}
 

### 2025-09-07 src/crm_fit/optimize.rs
return values of cost function evaluations from optimization as well

@@ -17,6 +17,7 @@ pub struct OptimizationResult {
     pub success: Option<bool>,
     pub neval: Option<usize>,
     pub niter: Option<usize>,
+    pub evals: Vec<f32>,
 }
 
 #[pymethods]
@@ -50,7 +51,7 @@ fn lhs_optimization(
     iterations_images: &[usize],
     positions_all: numpy::ndarray::ArrayView4<f32>,
     settings: &Settings,
-) -> PyResult<Option<(Vec<f32>, f32)>> {
+) -> PyResult<Option<(Vec<f32>, f32, Vec<f32>)>> {
     use kdam::{term::Colorizer, *};
     use rayon::prelude::*;
 
@@ -81,12 +82,12 @@ fn lhs_optimization(
             iterations_images,
         )
         .ok()
-        .map(|x| (parameters.to_vec(), x))
+        .map(|x| (parameters.to_vec(), x, vec![x]))
     })
     .filter(|x| x.1.is_finite())
     .reduce_with(|x, y| if x.1 < y.1 { x } else { y });
 
-    if let Some((_, cost)) = result {
+    if let Some((_, cost, _)) = result {
         println!("Final cost: {}", format!("{cost}").colorize("blue"));
     }
 
@@ -102,7 +103,7 @@ fn lhs_optimization_iter(
     positions_all: numpy::ndarray::ArrayView4<f32>,
     settings: &Settings,
     relative_reduction: f32,
-) -> PyResult<Option<(Vec<f32>, f32)>> {
+) -> PyResult<Option<(Vec<f32>, f32, Vec<f32>)>> {
     // Do lowering via repeated applications of LatinHypercube algorithm
     let mut bounds = bounds.clone();
     let mut result = None;
@@ -121,13 +122,16 @@ fn lhs_optimization_iter(
         // This will still decrease the overall size
         if result.is_none() && new_result.is_some() {
             result = new_result;
-        } else if let (Some((_, c0)), Some((p1, c1))) = (&result, new_result) {
+        } else if let (Some((_, c0, evals_old)), Some((p1, c1, evals_new))) = (&result, new_result)
+        {
             if c1 < *c0 {
-                result = Some((p1, c1));
+                let mut evals_combined = evals_old.clone();
+                evals_combined.extend(evals_new);
+                result = Some((p1, c1, evals_combined));
             }
         }
 
-        if let Some((params, _)) = &result {
+        if let Some((params, _, _)) = &result {
[…]

### 2025-09-07 cr_mech_coli/crm_fit/main.py
new function to plot optimization progression

@@ -259,6 +259,23 @@ def crm_fit_main():
 
     interval = time.time()
 
+    def plot_optimization_progression(evals):
+        fig, ax = plt.subplots(figsize=(8, 8))
+        crm.plotting.configure_ax(ax)
+        evals = np.sort(evals)[::-1]
+        ax.plot(np.arange(len(evals)), evals, label="Cost Function")
+        ax.legend(
+            loc="upper center",
+            bbox_to_anchor=(0.5, 1.1),
+            ncol=1,
+            frameon=False,
+        )
+        fig.savefig(out / "optimization-progression.png")
+        fig.savefig(out / "optimization-progression.pdf")
+        plt.close(fig)
+
+    plot_optimization_progression(optimization_result.evals)
+
     # Plot Cost function against varying parameters
     if not pyargs.skip_profiles:
         warnings.filterwarnings(

### 2025-09-07 paper/main.tex
completely remove multilayer stuff from paper
use new paths for figures

@@
     \centering
-    \begin{tikzonimage}[width=0.54\textwidth]
-        {figures/crm_fit/snapshots/predicted-031111.png}%
+    \begin{tikzonimage}[width=0.49\textwidth]
+        {figures/crm_fit/snapshots/predicted-001667.png}%
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
@@
     \hspace{0.01\textwidth}%
-    \begin{tikzonimage}[width=0.41\textwidth]
-        {figures/crm_fit/celldifss/cell-031111.png}%
+    \begin{tikzonimage}[width=0.49\textwidth]
+        {figures/crm_fit/celldiffs/diff-001667.png}%
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
@@
     \begin{tikzonimage}[width=0.33\textwidth]
-        {figures/crm_fit/profiles/radius-0.pdf}%
+        {figures/crm_fit/profiles/radius.pdf}%
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
@@
 
-%###################################################################################################
-% \subsection{Multilayer Predictions}
-% % TODO
-%
-% \begin{itemize}
-%     \item Friction force against surface. perpendicular to motion; proportional to velocity.
-%     \item gel pressure coming from experiment (simply downwards force $F_z$)
-% \end{itemize}
-%
-% \begin{itemize}
-%     \item Do we need this in this publication?
-%     \item Maybe this is out of scope as this is a more methodological paper.
-%     \item If we want to omit it, we might also opt to change the model description in the beginning
-%         and remove some parts of the external forces discussion.
-% \end{itemize}
-%
-% \cite{You2019}
-%
-% \begin{figure}
-%     \centering
-%     \includegraphics[width=0.49\textwidth]
-%         {docs/source/_static/scripts/crm_multilayer/multilayer-time-evolution.pdf}\\
-%     \includegraphics[width=0.49\textwidth]
-%         {docs/source/_static/scripts/crm_multilayer/colony-height-vs-gel_pressure.pdf}
-%     \includegraphics[width=0.49\textwidth]
-%         {docs/source/_static/scripts/crm_multilayer/colony-height-vs-time.pdf}
-%     \caption{TODO}
-%     \label{fig:multilayer}
-% \end{figure}
-
 \end{document}

### 2025-09-07 paper/main.tex
remove old individual radius plots

@@
     \label{fig:parameter-estimates-supplement-growth-rates}
-\end{figure}
-\begin{figure}
-    \centering
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/radius-0.pdf}%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/radius-1.pdf}%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/radius-2.pdf}\\%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/radius-3.pdf}%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/radius-4.pdf}%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/radius-5.pdf}%
-    \caption{Parameter estimation of radii for individual cells}%TODO extend caption}
-    \label{fig:parameter-estimates-supplement-radii}
 \end{figure}

### 2025-09-07 src/crm_fit/optimize.rs
remove print statement and usage of globals

@@ -189,7 +189,6 @@ pub fn run_optimizer(
     match settings.optimization.borrow(py).deref() {
         OptimizationMethod::DifferentialEvolution(de) => {
             let locals = pyo3::types::PyDict::new(py);
-            let globals = pyo3::types::PyDict::new(py);
 
             // Required
             locals.set_item("bounds", bounds.to_pyarray(py))?;
@@ -216,7 +215,6 @@ def callback(intermediate_result):
     fun = intermediate_result.fun
     global evals
     evals.append(float(fun))
-    print(evals)
 
 res = sp.optimize.differential_evolution(
     predict_calculate_cost,
@@ -234,11 +232,9 @@ res = sp.optimize.differential_evolution(
     rng=optimization.seed,
     callback=callback,
 )
-
-print(evals)
 "#
                 ),
-                Some(&globals),
+                None,
                 Some(&locals),
             )?;
             let res = locals.get_item("res")?.unwrap();
@@ -250,14 +246,7 @@ print(evals)
             let evals: Vec<f32> = locals
                 .get_item("evals")?
                 .and_then(|x| x.extract().ok())
-                .unwrap_or(
-                    globals
-                        .get_item("evals")?
-                        .and_then(|x| x.extract().ok())
-                        .unwrap(),
-                );
-
-            println!("{:?}", evals);
+                .unwrap_or_default();
 
             Ok(OptimizationResult {
                 params,

### 2025-09-07 cr_mech_coli/crm_fit/main.py
add color and xlabel to plot

@@ -263,13 +263,19 @@ def crm_fit_main():
         fig, ax = plt.subplots(figsize=(8, 8))
         crm.plotting.configure_ax(ax)
         evals = np.sort(evals)[::-1]
-        ax.plot(np.arange(len(evals)), evals, label="Cost Function")
+        ax.plot(
+            np.arange(len(evals)),
+            evals,
+            label="Cost Function",
+            color=crm.plotting.COLOR3,
+        )
         ax.legend(
             loc="upper center",
             bbox_to_anchor=(0.5, 1.1),
             ncol=1,
             frameon=False,
         )
+        ax.set_xlabel("Iterations")
         fig.savefig(out / "optimization-progression.png")
         fig.savefig(out / "optimization-progression.pdf")
         plt.close(fig)

### 2025-09-09 paper_segmentation_tracking/main.tex
template for secondary paper

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[hidelinks]{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsmath,amssymb,amsthm}
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{float}
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{mathrsfs}
\usepackage{onimage}
\usepackage{graphicx}
\usepackage{doi}
\usepackage{acronym}
\usepackage{listings}
\usepackage[mode=buildnew]{standalone}
\usepackage{tikz}
\usepackage{tabularx}
\usepackage{siunitx}
\usepackage{cite}
\usepackage{acronym}
\usepackage{xcolor}

\title{Improving Cell Segmentation and Tracking via Mechanistic Models of
Rod-Shaped Bacteria}

%\date{September 9, 1985}	% Here you can change the date presented in the paper title
%\date{} 					% Or removing it

\author{
    \href{https://orcid.org/0009-0001-0613-7978}{
        \includegraphics[scale=0.06]{figures/orcid.pdf}
        \hspace{1mm}Jonas Pleyer
    }$^1$
	% \texttt{jonas.pleyer@fdm.uni-freiburg.de} \\
	%% examples of more authors
	\href{https://orcid.org/0000-0002-6371-4495}{
        \includegraphics[scale=0.06]{figures/orcid.pdf}
        \hspace{1mm}Christian Fleck
    }$^1$\\\\
    $^1$Freiburg Center for Data Analysis, Modeling and AI, University of Freiburg\\
}

\newacro{abm}[ABM]{Agent-Based Model}
\newacro{ode}[ODE]{Ordinary Differential Equation}
\newacro{pbr}[PBR]{Physically-Based Rendering}

% Uncomment to remove the date
%\date{}

% Uncomment to override  the `A preprint' in the header
\renewcommand{\headeright}{Preprint}
%\renewcommand{\undertitle}{Technical Report}
\renewcommand{\shorttitle}{Improving Cell Segmentation and Tracking via Mechanistic Models of
Rod-Shaped Bacteria}

\usepackage{enumitem}
\setlist{nolistsep}

%%% Add PDF metadata to help others organize their library
%%% Once the PDF is generated, you can check the metadata with
%%% $ pdfinfo template.pdf
\hypersetup{
pdftitle={% TODO
},
pdfsubject={q-bio.NC, q-bio.QM},
pdfauthor={Jonas Pleyer, Christian Fleck},
pdfkeywords={},
}

% Change numbering of equations
% \numberwithin{equation}{section}

% MAKE TITLES IN THEOREMS BOLD
\makeatletter
\def\th@plain{%
  \thm@notefont{}% same as heading font
  \itshape % body font
}
\def\th@definition{%
  \thm@notefont{}% same as heading font
  \normalfont % body font
}
\makeatother

\begin{document}
\maketitle

%###################################################################################################
\begin{abstract}
\end{abstract}

% keywords can be removed
\keywords{Cell Segmentation, Cell Tracking, Mechanistic Modeling, Data Augmentation, Single-Cell}

\pagebreak
\renewcommand{\contentsname}{Table of Contents (remove before submission)}
\tableofcontents
\vfill
\pagebreak

%###################################################################################################
\section{(40\%) Introduction}

The field of biology has become increasingly intertwined with other disciplines such as mathematics,
physics and computation.
With their introduction, many processes have become increasingly automated, thus advancing our
capacity to analyse data at scale.
Many analysis schemes are based on microscopic images of cells which are being observed and possibly
tracked through time and space.
Often, these cells are genetically altered so that effects such as internal production parameters
can be observed from the outside via visual clues.
In order to be able to quantify single-cell effects, it is necessary that we can segment microscopic
images to obtain the location of individual cells.
This challenge of segmenting individual cells is a static problem by nature.
However by considering multiple microscopic images in succession, it is turned into a dynamic
problem of tracking single cells through time and space.
Going even further, we can include division events, thus again increasing the overall complexity of
the problem at hand.

Many algorithms have been conceived that are capable of dealing with these topics and they are being
applied actively within realms such as.
While many earlier iterations targeted 2D situations exclusively, more recent developments are
capable of processing 3D data as well.
Especially this latter subproblem is difficult to tackle since it allows for much larger
complexities within the analyzed dataset.
Furthermore, data in this case is often sparse, thus further complicating efficient training.
But while these tools are essential to get the task done, they can only be as good as the data with
which they are trained.
So far, these datasets are almost exclusively of experimental nature which requires lots of manual
labor in their construction.
The importance of data was also shown in various cases such as \textbf{TODO}.

\begin{itemize}
    \item Compare existing Cell Segmentation+Tracking tools (Cellpose, Omnipose, ..)
\end{itemize}

The goal of this study is to inspect if the process of data generation can be augmented by a
mechanistic single-cell model.
We will study this question along the example of rod-shaped bacteria.
With our recent developments, we have shown how \acp{abm} of rod-shaped bacteria can be fit to
microscopic image time-series, thus allowing us to construct mechanistic models.
From these single-cell models, we will first generate "sterile" snapshots which can then be used to
construct near-realistic brightfield microscopic images and perfect cell masks.
We will mimick known microscopic defects in order to provide realistic images variability that would
also ocurr naturally.
This data will be used to augment existing datasets and asses the effectivenss of this approach when
training existing cell-segmentation and cell-tracking tools on the expanded dataset.
With this procedure we also able to systematically study which of the introduced image defects poses
the biggest challenge for the investigated cell-tracking algorithms.
Furthermore, it will provide experimentalists with information, on which variables are most
important to control within their experimental setup in order to provide the best cell-segmentation
and cell-tracking results.

%###################################################################################################
\section{(70\%) Data Augmentation Strategy}
% --------------------------------------------------------------------------------------------------
\subsection{(90\%) Modeling Approach \& Parameter Estimation}

\begin{itemize}
    \item summarize contents from the other paper which I am preparing
    \item explain on which datasets we fitted the model
\end{itemize}

% --------------------------------------------------------------------------------------------------
\subsection{(80\%) Image Generation}
Using the model of the previous section, we can generate visualizations from its numerical results.
Figure~\ref{fig:progression-image-generation} shows intermediate steps and final results of a
particular generated image.
In the first step (subfigure A), we generate a 3D render which is a collection of meshes and
properties that can be rendered given a set of light sources and a particular view angle.
Using a parallel projection and by assigning uniue colors to individual agents, we can generate cell
masks similar to which would have been generated by a segmentation tool (subfigure B).
Due to the fact that we are utilizing a mechanistic model, we ensure that cells are colored
correctly and that our generated masks are in this sense "perfect".
This is one of the key foundations which allows this investigation in the first place.
Furthermore, since individual agents do not change their labels, the color values of cells are also
carried over correctly between individual time snapshots.
The last image (subfigure C) shows an example of a render where noise effects have been introduced.
These effects are the result of properties of the microscopic appartus, the cellular nature or other
external effects and ocurr naturally within most experimental setups.
We utilize various effects and implement their effect on the resulting image as summarized in
Table~\ref{tabular:microscopic-image-defects}.

Use various effects to make images realistic; mimick real microscopic effects for that
\begin{itemize}
    \item General (white) Noise
    \item Distort Images (lens distortion)
    \item Lighting for rendering in 3D
    \item Shaking for time-series
    \item ...
\end{itemize}

\begin{figure}
    \centering
    % \begin{tikzonimage}[width=0.9\textwidth]{visualization.pdf}
    %     \node at (0.025, 0.975)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    % \end{tikzonimage}
    \begin{tikzonimage}[width=0.3\textwidth]
        {docs/source/_static/09395645494836445480/raw_pv/000000400.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.3\textwidth]
        {docs/source/_static/09395645494836445480/masks/000000400.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.3\textwidth]
        {docs/source/_static/09395645494836445480/images/000000400.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}
    \caption{
        (A) shows the result from combining sphere and cylinder meshes to obtain the shape of a
        bacterial rod.
        This render contains lighting coming from 2 light sources which thus produces a small glow
        on the upper part of the meshes.
        (B) Unique color values are assigned to each agent and light sources are removed.
        Afterwards, the image is rendered by projecting along the z-axis.
        (C) We use physically-based rendering and assign properties such as roughness and
        reflectivity to generate an initial render.
        Afterwards, we apply white noise to the pixel values of image.
        In the future, we hope to extend this approach in order to construct near-realistic
        microscopic images from synthetic data.
    }
    \label{fig:progression-image-generation}
\end{figure}

\begin{table}[H]
    \begin{tabularx}{\columnwidth}{l l}
        \toprule
        Microscopic effect & Image Transformation\\
        \midrule
        Sensor Noise & White noise on pixel level\\
        Shaking Apparatus & Random spatial translations between time-steps\\
        Lighting & 3D Rendering: Introcue light sources; modify surface of cells; use \ac{pbr}\\
        Cellular Translucency & Make cells partially transparent\\
        Depth of Field & Included in \ac{pbr}\\
        \bottomrule
    \end{tabularx}
    \label{tabular:microscopic-image-defects}
\end{table}

%###################################################################################################
\section{(10\%) Methods}
% --------------------------------------------------------------------------------------------------
\subsection{Data Acquisition}
\begin{itemize}
    \item Use datasets from celltracking challenge
    \item possibly restrict to rod-shaped bacteria for now
\end{itemize}
% --------------------------------------------------------------------------------------------------
\subsection{Testing Methodology}
\begin{enumerate}
    \item Test model on Dataset
    \item Train model on additional data (possibly various configurations)
    \item Test again for same Dataset
\end{enumerate}

%###################################################################################################
\section{(0\%) Results}
% --------------------------------------------------------------------------------------------------
\subsection{Cell Segmentation}
\begin{itemize}
    \item Compare improvement from Cellpose -> Omnipose (generational improvement) with improvement
        in more (augmented) data
    \item Compare how various tools benefit from this approach
    \item How can the variation of generated data improve results?
    \item What is the impact of "noise introduction"? Which effects are more important than others?
\end{itemize}

% --------------------------------------------------------------------------------------------------
\subsection{Cell Tracking}
\begin{itemize}
    \item Use various "graining" of time-series data (larger time-steps between images) to asses the
        importance of this detail
    \item Can the improved ability to track cells through space be generalized to non-rod-shaped
        bacterial tracking?
\end{itemize}

%###################################################################################################
\section{(0\%) Discussion}
\begin{itemize}
    \item How can this be abstracted to other models (such as human stem cells, cocci, etc.)? Refer
        to "blueprint" which was talked about in previous paper
    \item 
\end{itemize}
\section{(0\%) Conclusion}

%###################################################################################################
\section{Data and Code Availability}
All data and code which has been used in this study is available at the github repository
\url{https://github.com/jonaspleyer/cr\_mech\_coli}.
It includes detailed descriptions of how to obtain the used data and bring it into a usable format.
The package which bundles the developed funtionality is published at
\url{https://pypi.org/project/cr-mech-coli/} from which it is readily available as a python
dependency.

\bibliographystyle{IEEEtran}
\bibliography{references}

\appendix
%###################################################################################################
\renewcommand{\thesection}{}
\renewcommand{\thesubsection}{S\arabic{subsection}}
\section{Supplementary Material}

%###################################################################################################
\subsection{Data Preparation}

\end{document}

### 2025-09-15 paper_segmentation_tracking/main.tex
add more authors
write short abstract

@@
     \href{https://orcid.org/0009-0001-0613-7978}{
+        % \includegraphics[scale=0.06]{figures/orcid.pdf}
+        \hspace{1mm}Student XYZ
+    }$^1$
+    \href{https://orcid.org/0009-0001-0613-7978}{
         \includegraphics[scale=0.06]{figures/orcid.pdf}
         \hspace{1mm}Jonas Pleyer
+    }$^2$
+    \href{https://orcid.org/0009-0001-0613-7978}{
+        % \includegraphics[scale=0.06]{figures/orcid.pdf}
+        \hspace{1mm}Jelena Bratulić
     }$^1$
@@
         \hspace{1mm}Christian Fleck
+    }$^2$
+	\href{https://orcid.org/0000-0002-6371-4495}{
+        % \includegraphics[scale=0.06]{figures/orcid.pdf}
+        \hspace{1mm}Thomas Brox
     }$^1$\\\\
-    $^1$Freiburg Center for Data Analysis, Modeling and AI, University of Freiburg\\
+    $^1$University of Freiburg, Germany\\
+    $^2$Freiburg Center for Data Analysis, Modeling and AI\\
 }
@@
 \begin{abstract}
+    The detection of individual cells from microscopic images is common practice with many
+    biological studies.
+    A variety of deep-learning algorithms are under active development to tackle these tasks.
+    This study discusses how the training of such algorithms can be improved by generated synthetic
+    data from a parametrized, mechanistic single-cell model.
+    We use our recent advancements where we have shown how to estimate the parameters of a
+    generalized mechanical model for rod-shaped bacteria to generate cell masks and extend the work
+    by generating synthetic, near-realistic microscopic images utilizing known effects such as noise
+    and other defects that ocurr naturally within these images.
+    Using these methods, we investigate how differing approaches benefit from the additional data
+    and how various defects within our microscopic images impact their performances.
 \end{abstract}

### 2025-09-15 paper_segmentation_tracking/fake-figures/main.py
add fake figures for paper

@@ -1,0 +1,87 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import cr_mech_coli as crm
+
+
+def configure_ax(ax):
+    crm.configure_ax(ax)
+    ax.text(
+        0.5,
+        0.5,
+        "FAKE DATA",
+        transform=ax.transAxes,
+        fontsize=60,
+        color="red",
+        alpha=0.5,
+        ha="center",
+        va="center",
+        rotation=30,
+    )
+
+
+if __name__ == "__main__":
+    fig, ax = plt.subplots(figsize=(8, 8))
+    configure_ax(ax)
+
+    fake_data = {
+        # Name, (performance before, performance after)
+        "Cellpose": (0.8, 0.85),
+        "Omnipose": (0.9, 0.92),
+        "Other ..": (0.5, 0.6),
+    }
+
+    width = 0.33
+    padding = 0.02
+    x = np.arange(2)
+    colors = ["black", "gray"]
+    labels = ["Before", "After"]
+
+    for n, (name, measurements) in enumerate(fake_data.items()):
+        rects = ax.bar(
+            [n - width / 2 - padding / 2, n + width / 2 + padding / 2],
+            measurements,
+            width,
+            label=labels,
+            color=colors,
+        )
+        ax.bar_label(rects, padding=3)
+
+    ax.set_xticks(np.arange(len(fake_data)), fake_data.keys())
+    ax.set_ylim(0, 1.1)
+    handles, labels = ax.get_legend_handles_labels()
+    ax.legend(handles[: len(colors)], labels[: len(colors)])
+
+    fig.savefig("paper_segmentation_tracking/fake-figures/segmentation-benchmark.pdf")
+    plt.close(fig)
+
+    fig, ax = plt.subplots(figsize=(8, 8))
+    configure_ax(ax)
+
+    width = 0.25
+    fake_data = {
+        "Cellpose": (5.1, 15.34, 60.3),
+        "Omnipose": (3.1, 16.34, 34.3),
+        "Other ..": (7.8, 26.1, 87.7),
+    }
+
+    colors = ["green", "orange", "blue"]
+    labels = ["Sensor Noise", "Smudges", "Shaking"]
+    x = np.arange(len(labels))
+
+    for n, (name, measurements) in enumerate(fake_data.items()):
+        rects = ax.bar(
+            n + x * width - width / 2,
+            measurements,
+            width,
+            label=labels,
[…]

### 2025-09-15 paper_segmentation_tracking/main.tex
formatting
move tabular further up; add smudges

@@
 
+% TODO comment this before submission
+\vfill
 \pagebreak
@@
 
+\begin{table}[H]
+    \begin{tabularx}{\textwidth}{l l}
+        \toprule
+        Microscopic effect & Image Transformation\\
+        \midrule
+        Sensor Noise & White noise on pixel level\\
+        Shaking Apparatus & Random spatial translations between time-steps\\
+        Smudges & Custom overlays (Rotations, Mirrors, Combinations, etc.)\\
+        Lighting & 3D Rendering: Introcue light sources; modify surface of cells; use \ac{pbr}\\
+        Cellular Translucency & Make cells partially transparent\\
+        Depth of Field & Included in \ac{pbr}\\
+        \bottomrule
+    \end{tabularx}
+    \label{tabular:microscopic-image-defects}
+\end{table}
+
 Use various effects to make images realistic; mimick real microscopic effects for that
@@
 \end{figure}
-
-\begin{table}[H]
-    \begin{tabularx}{\columnwidth}{l l}
-        \toprule
-        Microscopic effect & Image Transformation\\
-        \midrule
-        Sensor Noise & White noise on pixel level\\
-        Shaking Apparatus & Random spatial translations between time-steps\\
-        Lighting & 3D Rendering: Introcue light sources; modify surface of cells; use \ac{pbr}\\
-        Cellular Translucency & Make cells partially transparent\\
-        Depth of Field & Included in \ac{pbr}\\
-        \bottomrule
-    \end{tabularx}
-    \label{tabular:microscopic-image-defects}
-\end{table}
 

### 2025-09-15 paper_segmentation_tracking/main.tex
add more todos
add preliminary figure of "fake-data reults"

@@
     \item possibly restrict to rod-shaped bacteria for now
+    \item maybe use remaining data to check that there is no overall regression within other samples
 \end{itemize}
@@
     \item Test model on Dataset
-    \item Train model on additional data (possibly various configurations)
+    \item Train model on generated data (possibly various configurations)
     \item Test again for same Dataset
+    \item Compare results
+    \item Put in nice graphs
 \end{enumerate}
@@
 \end{itemize}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{fake-figures/segmentation-benchmark.pdf}%
+    \includegraphics[width=0.5\textwidth]{fake-figures/tracking-defects.pdf}
+    \caption{
+        Performance of various segmentation and tracking techniques.
+        We will need various of such plots.
+    }
+\end{figure}
 

### 2025-09-15 paper_segmentation_tracking/fake-figures/main.py
change style of fake figures

@@ -4,6 +4,7 @@ import cr_mech_coli as crm
 
 
 def configure_ax(ax):
+    crm.set_mpl_rc_params()
     crm.configure_ax(ax)
     ax.text(
         0.5,
@@ -16,6 +17,8 @@ def configure_ax(ax):
         ha="center",
         va="center",
         rotation=30,
+        fontfamily="sans-serif",
+        fontweight="bold",
     )
 
 
@@ -49,7 +52,14 @@ if __name__ == "__main__":
     ax.set_xticks(np.arange(len(fake_data)), fake_data.keys())
     ax.set_ylim(0, 1.1)
     handles, labels = ax.get_legend_handles_labels()
-    ax.legend(handles[: len(colors)], labels[: len(colors)])
+    ax.legend(
+        handles[: len(colors)],
+        labels[: len(colors)],
+        loc="upper center",
+        bbox_to_anchor=(0.5, 1.1),
+        ncol=2,
+        frameon=False,
+    )
 
     fig.savefig("paper_segmentation_tracking/fake-figures/segmentation-benchmark.pdf")
     plt.close(fig)
@@ -81,7 +91,14 @@ if __name__ == "__main__":
     ax.set_xticks(np.arange(len(fake_data)) + width / 2, fake_data.keys())
     ax.set_ylabel("Relative Impact [%]")
     handles, labels = ax.get_legend_handles_labels()
-    ax.legend(handles[: len(colors)], labels[: len(colors)])
+    ax.legend(
+        handles[: len(colors)],
+        labels[: len(colors)],
+        loc="upper center",
+        bbox_to_anchor=(0.5, 1.1),
+        ncol=3,
+        frameon=False,
+    )
 
     fig.savefig("paper_segmentation_tracking/fake-figures/tracking-defects.pdf")
     plt.close(fig)

### 2025-09-20 paper/main.tex
add comparison between estimated and fitted growth_rates
rephrasing

@@
 generated by an \ac{abm} lead to the desired phenomenon which aligns with the observed biological
-reality thus omitting to actually estimate the necessary parameters directly.
+reality.
+But this process does not actually estimate the necessary parameters directly
 % TODO citations
-This reduces the confidence in the constructed models and raises concerns with respect to the
+and reduces the confidence in the constructed models and raises concerns with respect to the
 validity of the obtained results.
@@
 permanent plastic deformations~\cite{Amir2014_2}.
-They are an interesting target system since their physical form provides more surface information
-compared to spherical cells.
+Rod-shaped Bacteria are an interesting target system since their physical form provides more surface
+information compared to other forms such as spherical cells.
 In addition, parameters such as growth rates have been found to vary across multiple
@@
 Computational models for rod-shaped bacteria allow researchers to systematically analyze their
-properties and infer emerging phenomena~\cite{Cho2007,Jnsson2005}.
+properties and infer emergent phenomena~\cite{Cho2007,Jnsson2005}.
 Existing tools can describe growth, division and many effects such as colony formation and
@@
 Together with the parameter estimation methods, they can then be used to provide large amounts of
-mechanistic data for the training cell-segmentation and cell-tracking machine-learning algorithms.
+mechanistic data for the training of cell-segmentation and cell-tracking machine-learning
+algorithms.
 
@@
 % --------------------------------------------------------------------------------------------------
-\subsection{Physical Interactions}
-\label{subsection:mechanical-model-interactions}
-
-When calculating forces acting between the cells, we can use a simplified model to circumvent the
+\subsection{Physical Interactions - Mie Potential}
+
+When calculating forces which act between the cells, we can use a simplified model to circumvent the
 numerically expensive integration over the complete length of the rod.
@@
 \end{figure}
+
+\begin{table}
+    \centering
+    \begin{tabularx}{\textwidth}{l c c c c c c}
+        \toprule
+        % \multicolumn{6}{Growth Rates}\\
+        Method & Cell $0$ & Cell $1$ & Cell $2$ & Cell $3$ & Cell $4$ & Cell $5$\\
+        \midrule
+        \ac{abm} & 0.0011527990 & 0.0014106040 & 0.0018761827 & 0.0016834959 & 0.0036106023 & 0.0015209642\\
+        Direct   & 0.0014057073 & 0.0010807540 & 0.0023177727 & 0.0015120675 & 0.0020488179 & 0.0016726263\\
+        % 0.0014057073 & 0.0010807540 & 0.0023177727 & 0.0015120675 & 0.0020488179 & 0.0016726263\\
+        % direct & 0.0009785162 & 0.0012641432 & 0.0021701728 & 0.0015695889 & 0.0036255661 & 0.0018419186\\
+        \bottomrule
+    \end{tabularx}
+    \caption{TODO}
+\end{table}
 

### 2025-09-20 paper/main.tex
move morse potential to supplement

@@
     \vec{F}_{i,\text{External}} = \vec{F}(\vec{x}_i,\vec{p})
-    \label{eq:cell-force-external}
 \end{equation}
 
-The specific choice of interaction potential plays of course a crucial role for the dynamics of the
-agents.
-But we cannot expect to simply know the shape and parameters of this potential.
-Thus, we would like to experiment with various interaction potentials in order to quantify their
-ability to produce accurate results.
-We utilize two distinct potentials that are both able to capture the essence of our interactions by
-providing a short-range repulsive part and a long-ranged attractive component.
-
-\paragraph{Morse Potential}
-The morse potential~\cite{Morse1929} is a common example of numerically inexpensive interaction
-potential.
-It was originally designed to describe the interaction of diatomic molecules but has since found use
-in other areas~\cite{Breitwieser2021}.
-It contains a repulsive and attractive part and is bound from above which makes it particularly
-suitable for numerical simulations since unbound potentials can lead to large numerical deviations
-thus compromising the prediction of the model.
-The potential is of the form
-\begin{equation}
-    V(r) = V_0\left(1-e^{-\lambda(r-(R_1+R_2))}\right)^2
-\end{equation}
-and has 3 parameters.
-The overall potential strength $V_0$, the radii of the particles $R_1,R_2$ and the potential
-stiffness $\lambda$ which controls the rate of change as one travels along the radial axis.
-We employ a spatial cutoff which sets the interaction strength to $0$ when exceeding a distance
-$r\geq\xi$.
-
-\paragraph{Mie Potential}
 The Mie~\cite{Mie1903} potential is the generalization of the well-known
@@
 (m/n)^{1/(n-m)}(R_1+R_2)$ where $R_1,R_2$ are the radii of the two interacting particles.
+
+The specific choice of interaction potential plays a crucial role for the dynamics of the agents.
+Since we cannot expect to simply know the correct shape and parameters of this potential, it is
+desirable to experiment with various interaction potentials in order to quantify their
+ability to produce accurate results.
+While this study, only relies on the Mie Potential, our python package \texttt{cr\_mech\_coli}
+also implements the popular Morse potential~\cite{Morse1929} to model physical interactions.
 
@@
 \end{table}
+
+%###################################################################################################
+\subsection{Alternative Interaction Potential - Morse Potential}
+The morse potential~\cite{Morse1929} is a common example of numerically inexpensive interaction
+potential.
+It was originally designed to describe the interaction of diatomic molecules but has since found use
+in other areas~\cite{Breitwieser2021}.
+It contains a repulsive and attractive part and is bound from above which makes it particularly
+suitable for numerical simulations since unbound potentials can lead to large numerical deviations
+thus compromising the prediction of the model.
+The potential is of the form
+\begin{equation}
+    V(r) = V_0\left(1-e^{-\lambda(r-(R_1+R_2))}\right)^2
+\end{equation}
+and has 3 parameters.
+The overall potential strength $V_0$, the radii of the particles $R_1,R_2$ and the potential
+stiffness $\lambda$ which controls the rate of change as one travels along the radial axis.
+We employ a spatial cutoff which sets the interaction strength to $0$ when exceeding a distance
+$r\geq\xi$.
 

### 2025-09-20 paper/main.tex
remove unused label
write short introductory notes for application chapter

@@
 \subsection{Cell-Cycle}
-\label{subsection:mechanical-model-cycle}
 
@@
 \section{Application to \textit{E.Coli.} growing on Agar}
-
-With the previous works of extracting the positions (see section~\ref{section:data-extraction}) for
-individual cell-agents and the methods to estiamte parameters of our agents (see section~\ref{section:parameter-estimation}), we are now ready to apply them to showcase their usefulness.
-
-TODO explain what data we are using
+\label{section:application}
+
+To show the usefulness of our previously developed methods (see
+section~\ref{section:parameter-estimation}), we will use them to estimate the parameters of a
+sequence of microscopic images of growing \texttt{E.Coli}.
+The data we use~\cite{https://doi.org/10.3203/iwf/k-129} (see
+section~\ref{section:data-code-availability}) has been automatically segmented with the
+omnipose~\cite{Cutler2022} algorithm which provides us with the necessary cell masks.
 

### 2025-09-20 paper/main.tex
rewrite individual treatment section

@@
 estimate parameters of our system, we need to treat every cell-agent individually, thus assigning
-parameters individually to them.
-This means that we increase the total number of parameters that needs to be sampled within an
+parameters on a single-cell level.
+This would mean to increase the total number of parameters that need to be sampled within an
 optimization routine.
-Given $N_c$ cells and $N_p$ parameters per cell and $N_u$ unique parameters (where $N_u\leq N_p$),
-the total amount of parameters is
+Given $N_c$ cells and $N_p$ parameters per cell and $N_u$ individual parameters (where
+$N_u\leq N_p$), the total amount of parameters is
 
@@
 
-It is clear, that we should avoid to needlessly assign every parameter individually since the space
-which is needed to be sampled scales exponentially with the dimension ie. the number of parameters.
-However, we will now present data that indicate that an individual treatment is necessary.
+It is clear, that we should avoid to needlessly assign every parameter individually since the volume
+of the space that needs to be sampled, scales exponentially with the dimension ie. the number of
+parameters.
+However, we will now present data that indicate that an individual treatment is in fact necessary.
 Figure~\ref{fig:estimated-growth-rates} shows the rod-lengths of individual bacteria which have been
@@
 
-and where $\lambda$ is the growth rate $t_0$ the delay.
-The fits to invidiual growth-curves of figure~\ref{fig:estimated-growth-rates} (A) result in good
-matches and so does the fit of subfigure (B).
+where $\lambda$ is the growth rate and $t_0$ the delay.
+Fitting this function to the individual growth curves of the bacteria and their mean growth curve,
+both results in good agreements (Figure~\ref{fig:estimated-growth-rates} A,B).
 However, the latter exhibits large amounts of uncertainty within the data which is also illustrated
@@
 system, we need to treat the bacteria individually.
-
-The individual treatment of the bacteria allows us in principle to estimate the distribution of the
-parameters directly without having to resort to any assumptions on its form.
-This is true for any parameter, not just the growth rates, since we aim to estimate all mechanical
-parameters of our cellular model.
-
-\textbf{TODO is this discussion?}\\
-The total parameter space $P$ is a product of all single spaces $P_i$ in which the parameters
-$p_i$ live $P=P_0\times P_1\times\dots$ but some of these parameters share the same space (let's
-assume $p_0,p_1\in \tilde{P}$).
-This leaves the open question as to weather the optimization algorithm could be modified to leverage
-this knowledge and obtain increased performance.
-For instance, it might be valuable to initially consider individual parameters as globally defined
-and start treating the individually after a certain amount of optimization steps.
 

### 2025-09-20 paper/main.tex
rewrite section without division

@@
 \subsection{Without Division}
-For now, we will focus on a time-interval which does not contain any division events.
+For now, we will focus on a time-interval which does not contain any division events and at which
+the initial lag-phase of our bacteria os already passed.
 By doing so, we can utilize the simpler and more performant position comparison scheme for our loss
 function.
+Furthermore, we need not to estimate the lag-time parameters which improves performance of our
+optimization algorithm.
 For this dataset, we assumed that all parameters which specify the interaction potential are
-identical across all agents but growth rates \textbf{TODO check this} and the physical radii of the
-cells are individual and we optimize them like any other global parameter.
-We chose to optimize for 3 interaction parameters of the Mie potential $n,m,V_0$ and two parameters
-which were individual for each cell, namely the radii $r_i$ and growth rates $\mu_i$.
-We also included the damping constant $\lambda$ and removed any stochastic motion contributions
-($D=0$).
-This leads to a total of $4+2\times 6=16$ parameters.
-To estimate these parameters we used a latin-hypercube~\cite{McKay1979} sweep across the parameter
-space and performed a local minimization with the Nelder-Mead~\cite{Powell1973} method.
-
-In figure~\ref{fig:parameter-estimates-single-step} we performed a sensitivity-analysis for every
-parameter.
-We can see that almost all parameters can be estimated by our optimization approach.
-However, the damping constant (A) does not show any changes in the loss function when increasing its
-value.
-It hints at a non-identifiability~\cite{Raue2009} of the parameter $\lambda$.
-But it also shows that the approach which we used is able to detect these problems.
-As we continue to develop these methods, we expect to be able to reuse many of the existing
-theoretical approaches such as likelihood estimation and uncertainty quantification which have
-already been developed extensively for less complex systems such as \acp{ode}.
-\textbf{TODO citations, also: is this discussion?}
-Fortunately, \texttt{cellular\_raza} was chosen to construct our model such that we are flexible in
-changing the underlying representation if should find that such unidentifiabilities are structural
-(model-based) rather than practical (due to lack of data).
-This allows us to iterate and find the model formulation which suits the data best but still retain
-the mechanistic methodology of our approach.
+identical across all agents but growth rates of the cells are individual.
+
+We chose to optimize for the damping of agents' movement within the medium, 3 interaction parameters
+of the Mie potential $n,m,V_0$ a single radius for all agents and growth rates which were chosen
+individually for each cell.
+We removed any stochastic motion contributions by setting $D=0$.
+This leads to a total of $5+1\times 6=11$ parameters.
+To estimate these parameters we used the differential evolution~\cite{Storn1997} global optimization
+algorithm which was initially seeded with a sample generated by the latin-hypercube~\cite{McKay1979}
+method.
+Finally, we performed a local minimization using the approach by Nelder-Mead~\cite{Powell1973}.
+
+Figure~\ref{fig:parameter-estimates-single-step} shows profile-likelihoods for each optimized
+parameter with respect to the optimization results described before.
+We can see that all parameters can be estimated by our optimization approach and that no
+non-identifiability~\cite{Raue2009} is present.
 
@@
     \begin{tikzonimage}[width=0.49\textwidth]
-        {figures/crm_fit/snapshots/predicted-001667.png}%
+        {figures/crm_fit/snapshots/predicted-003889.png}%
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
@@
     \begin{tikzonimage}[width=0.49\textwidth]
-        {figures/crm_fit/celldiffs/diff-001667.png}%
+        {figures/crm_fit/celldiffs/diff-003889.png}%
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
@@
     \caption{
-        \textbf{TODO do we need this figure?}
-        (A) A microscopic image with the predicted positions of the cells overlayed on top.
+        (A) Final microscopic image of the time series with the predicted positions of the cells
+        overlayed on top.
         (B) Differences between extracted and predicted positions.
-        % TODO make this figure to scale
-    }
-    \label{fig:celldiffs-individual}
+        Note that the imaging-based comparison algorithm was not used to optimize this system but is
+        chosen here as a tool to aid visualization of the prediction difference.
+    }
 \end{figure}
@@
         {figures/crm_fit/profiles/damping.pdf}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{A}};
     \end{tikzonimage}%
@@
         {figures/crm_fit/profiles/exponent-m.pdf}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
     \end{tikzonimage}%
@@
         {figures/crm_fit/profiles/exponent-n.pdf}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
     \end{tikzonimage}\\
@@
         {figures/crm_fit/profiles/strength.pdf}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
     \end{tikzonimage}%
@@
         {figures/crm_fit/profiles/growth-rate-0.pdf}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
     \end{tikzonimage}%
@@
         {figures/crm_fit/profiles/radius.pdf}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
+        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
     \end{tikzonimage}
     \caption{
-        Sensitivity analyses werer performed for every optimized parameter of our model.
-        (A-F) show the curves of loss functions when varying this respective parameter and keeping
-        all others constant at the value of the found optimum (blue dot).
-        In the cases of (B-F), we can see that a minimum was achieved and thus the parameters could
-        be estimated.
-        However, for the case of the damping constant $\lambda$ in (A), we see that the upper bound
-        is open.
-        This suggests that our model could be reduced with regard to this parameter.
-        We have exemplified the values for radii $r_i$ and growth rates $\mu_i$ with values of the
-        first cell.
-        The remaining plots can be found in the supplementary material in
-        section~\ref{sec:supplement-parameter-profiles}.
+        We calculated the likelihood profiles for every optimized parameter of our model.
+        All plots show how the value of the loss function varies when fixing a single parameter but
+        still optimizing for the remaining ones.
+        Every plot indicates that a minimum was achieved, thus every parameter could be identified.
+        We show only one of the 6 individually-chosen growth rate as an example.
+        The remaining plots show similar behaviour although with varying values and can be found in
+        the supplementary material in section~\ref{sec:supplement-parameter-profiles}.
     }

### 2025-09-20 paper/main.tex
fill section with division

@@
 \subsection{With Division (TODO)}
+
+In the next step, we extend our previously used time-series with additional images, which include
+cell division events which means that we now have to use the imaing-based comparison algorithm (see
+section~\ref{section:parameter-estimation}).
+A practical problem in extending the time series is that the segmentation algorithm was not able to
+properly segment the individual bacteria around the division event which meant that we had to omit
+parts of the snapshots, leaving us with only incomplete data.
+
+Since we have already optimized said system, we can reduce the computational demand by choosing the
+parameters which were obtained by our previous attempts and extend them with the parameters that the
+daughter cells will carry.
+In addition to assigning growth rates individually, but we also need to consider division lengths at
+which the bacteria split.
+Since our simulation involves the division of 4 cells into 8 new ones, in principle, this would lead
+us to a total of $8$ new growth rates and $4$ division lengths.
+But to decrease the size of the parameter space, we assume that the growth rates of the daughter
+cells are identical between every sibling, leaving us at only $4$ new growth rates and a total of
+$8$ parameters.
+Future approaches should aim to loosen this assumption, allowing to optimize every new growth rate.
+To optimize this system, we again used the same optimization approach by seed with samples generated
+by latin-hypercube~\cite{McKay1979}, then applying the differential evolution
+algorithm~\cite{Storn1997} and finally a local minimization with Nelder-Mead~\cite{Powell1973}.
+We used the value $0.5$ for the parent penalty, which aims to provide a smooth transition between
+not accounting for them and assigning full weighting to this difference.
+Figure~\ref{fig:celldiffs-progression-with-division} shows the difference between the cell mask as
+given by the data and the cell mask predicted by the best fit of our model.
+
+We can again calculate likelihood profiles for every optimized parameter.
+The results displayed in Figure~\ref{fig:likelihood-profiles-comparison-with-division} show that the
+values for the division lengths seem exhibit a decrease in steepness of the curve towards the left
+side of the estimated parameter.
+This can be attributed to the missing images which were not able to be segmented in our time series.
+These results highlight a reassuring finding, namely that the method of profile-likelihood is able
+to properly display this obvious lack of data.
+It could be argued that this leads to a non-identifiability for the first division length while the
+reamining parameters are still identifiable.
+Meanwhile, the growth rates of the daughters ...
+
+\begin{figure}
+    \centering
+    \begin{tikzonimage}[width=0.32\textwidth]
+        {figures/crm_divide/masks_diff/000000.png}%
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \end{tikzonimage}%
+    \hspace{0.01\textwidth}%
+    \begin{tikzonimage}[width=0.32\textwidth]
+        {figures/crm_divide/masks_diff/000010.png}%
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+    \end{tikzonimage}%
+    \hspace{0.01\textwidth}%
+    \begin{tikzonimage}[width=0.32\textwidth]
+        {figures/crm_divide/masks_diff/000020.png}%
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+    \end{tikzonimage}%
+    \caption{
+        Snapshots of differences between cell masks obtained from the data and predicted by the best
+        fit of our model at (A) the beginning (B) right before the division event (C) right after
+        the division event.
+        The optimization routine was able to fully mitigate the cost assigned due to parental
+        relationships which means that the fully-optimized result contains division lengths that
+        match the data.
+    }
+    \label{fig:celldiffs-progression-with-division}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \begin{tikzonimage}[width=0.25\textwidth]
+        {figures/crm_divide/profiles/profile-000000.png}%
+        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.25\textwidth]
+        {figures/crm_divide/profiles/profile-000001.png}%
+        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.25\textwidth]
+        {figures/crm_divide/profiles/profile-000002.png}%
+        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.25\textwidth]
+        {figures/crm_divide/profiles/profile-000003.png}%
+        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+    \end{tikzonimage}\\
+    \begin{tikzonimage}[width=0.25\textwidth]
+        {figures/crm_divide/profiles/profile-000004.png}%
+        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.25\textwidth]
+        {figures/crm_divide/profiles/profile-000005.png}%
+        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.25\textwidth]
+        {figures/crm_divide/profiles/profile-000005.png}%
+        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.25\textwidth]
+        {figures/crm_divide/profiles/profile-000005.png}%
+        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{H}};
+    \end{tikzonimage}%
+    \caption{TODO}
+    \label{fig:likelihood-profiles-comparison-with-division}
+\end{figure}
 

### 2025-09-20 paper/main.tex
include discussion which was in previous sections in final discussion

@@
 
+\textbf{Discussion on the Individual Treatment}\\
+The individual treatment of the bacteria allows us in principle to estimate the distribution of the
+parameters directly without having to resort to any assumptions on the form of said distribution.
+This is true for any parameter, not just the growth rates, since we aim to estimate all mechanical
+parameters of our cellular model.
+
+The total parameter space $P$ is a product of all single spaces $P_i$ in which the parameters
+$p_i$ live $P=P_0\times P_1\times\dots$ but some of these parameters share the same space (let's
+assume $p_0,p_1\in \tilde{P}$).
+This leaves the open question as to weather the optimization algorithm could be modified to leverage
+this knowledge and obtain increased performance.
+For instance, it might be valuable to initially consider individual parameters as globally defined
+and start treating the individually after a certain amount of optimization steps.
+
+\textbf{Discussion Identifiability and Model Reduction/Flexibility}\\
+But it also shows that the approach which we used is able to detect these problems.
+As we continue to develop these methods, we expect to be able to reuse many of the existing
+theoretical approaches such as likelihood estimation and uncertainty quantification which have
+already been developed extensively for less complex systems such as \acp{ode}.
+\textbf{TODO citations, also: is this discussion?}
+Fortunately, \texttt{cellular\_raza} was chosen to construct our model such that we are flexible in
+changing the underlying representation if should find that such unidentifiabilities are structural
+(model-based) rather than practical (due to lack of data).
+This allows us to iterate and find the model formulation which suits the data best but still retain
+the mechanistic methodology of our approach.
+
 \section{Conclusion (TODO)}
-\label{section:conclusion}
 

### 2025-09-20 paper/main.tex
add future work idea

@@
     \item Provide software package which combines all functionality and is readily available
+    \item Future work:
+    \begin{itemize}
+        \item Optimize multiple data sources (images) in parallel: requires some kind of
+        "normalization" to ensure that the dimensions, pixels etc. are the same such that parameters
+        are comparable; also requires identical experimental setup
+        \item Compare different species with various modifications
+        \item Modify model futher; asymmetric forces etc.
+    \end{itemize}
 \end{itemize}

### 2025-09-20 paper/main.tex
enhance data & code availability statement

@@
 \section{Data and Code Availability}
+\label{section:data-code-availability}
 All data and code which has been used in this study is available at the github repository
@@
 \url{https://pypi.org/project/cr-mech-coli/} from which it is readily available as a python
-dependency.
+dependency.\\
+The time series of microscopic images which has been studied in section~\ref{section:application}
+was extracted from individual frames of a video~\cite{https://doi.org/10.3203/iwf/k-129} that
+showcases the exponential growth of bacteria.
 

### 2025-09-20 cr_mech_coli/crm_divide.py
cleanup

@@ -313,16 +313,6 @@ def predict(
     return container
 
 
-def mask_iterator(iterations_data, new_masks, masks_predicted, return_all):
-    if return_all:
-        for iter, new_mask in zip(iterations_data, new_masks):
-            yield new_mask, masks_predicted[iter]
-
-    else:
-        for m1, m2 in zip(new_masks, masks_predicted):
-            yield m1, m2
-
-
 ERROR_COST = 1e6
 
 
@@ -432,7 +422,6 @@ def objective_function(
             color_to_cell,
             container,
             masks_predicted,
-            penalties,
         )
 
     if return_times:
@@ -524,9 +513,7 @@ def plot_mask_adjustment(
         color_to_cell,
         container,
         masks_predicted,
-        penalties,
     ) = objective_function(x0, *args, return_all=True, show_progressbar=True)
-    print(penalties)
 
     (output_dir / "mask_adjustments").mkdir(parents=True, exist_ok=True)
     for mask_predicted, mask_adjusted, mask_data, mask_iter in tqdm(
@@ -819,21 +806,6 @@ def run_optimizer(
         final_parameters = result[:-1]
         final_cost = result[-1]
     else:
-        # lhs = sp.stats.qmc.LatinHypercube(
-        #     d=len(spring_length_thresholds_and_new_growth_rates)
-        # )
-        # b = np.array(bounds)
-        # samples = b[:, 0] + lhs.random(n=50) * (b[:, 1] - b[:, 0])
-
-        # pool_args = [(s, *args) for s in samples]
-        # costs = process_map(
-        #     calculate_single,
-        #     pool_args,
-        #     max_workers=n_workers,
-        #     desc="Finding Global Optimum",
-        # )
-        # final_parameters, final_cost = min(costs, key=lambda x: x[1])
-
         res = sp.optimize.differential_evolution(
             objective_function,
             x0=spring_length_thresholds_and_new_growth_rates,
@@ -1002,7 +974,6 @@ def crm_divide_main():
         color_to_cell,
         container,
         masks_predicted,
-        _,
     ) = objective_function(
         final_parameters, *args, return_all=True, show_progressbar=True
     )

### 2025-09-20 paper/main.tex
replace nelder-mead with l-bfgs-b

@@
 method.
-Finally, we performed a local minimization using the approach by Nelder-Mead~\cite{Powell1973}.
+Finally, we performed a local minimization using the L-BFGS-B algorithm~\cite{Liu1989}.
 
@@
 by latin-hypercube~\cite{McKay1979}, then applying the differential evolution
-algorithm~\cite{Storn1997} and finally a local minimization with Nelder-Mead~\cite{Powell1973}.
+algorithm~\cite{Storn1997} and finally a local minimization via L-BFGS-B~\cite{Liu1989}.
 We used the value $0.5$ for the parent penalty, which aims to provide a smooth transition between

### 2025-09-20 paper/main.tex
add more profiles for growth rates

@@
     \centering
-    \begin{tikzonimage}[width=0.33\textwidth]
+    \begin{tikzonimage}[width=0.25\textwidth]
         {figures/crm_fit/profiles/damping.pdf}%
@@
     \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.33\textwidth]
+    \begin{tikzonimage}[width=0.25\textwidth]
         {figures/crm_fit/profiles/exponent-m.pdf}%
@@
     \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.33\textwidth]
+    \begin{tikzonimage}[width=0.25\textwidth]
         {figures/crm_fit/profiles/exponent-n.pdf}%
         \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
-    \end{tikzonimage}\\
-    \begin{tikzonimage}[width=0.33\textwidth]
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.25\textwidth]
         {figures/crm_fit/profiles/strength.pdf}%
         \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
-    \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.33\textwidth]
+    \end{tikzonimage}\\
+    \begin{tikzonimage}[width=0.25\textwidth]
+        {figures/crm_fit/profiles/radius.pdf}%
+        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.25\textwidth]
         {figures/crm_fit/profiles/growth-rate-0.pdf}%
-        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
-    \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.33\textwidth]
-        {figures/crm_fit/profiles/radius.pdf}%
         \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
-    \end{tikzonimage}
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.25\textwidth]
+        {figures/crm_fit/profiles/growth-rate-1.pdf}%
+        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.25\textwidth]
+        {figures/crm_fit/profiles/growth-rate-2.pdf}%
+        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{H}};
+    \end{tikzonimage}%
     \caption{

### 2025-09-20 paper/main.tex
finish With Division subsection
- add some information into appendix
  and link to it

@@
 % --------------------------------------------------------------------------------------------------
-\subsection{With Division (TODO)}
+\subsection{With Division}
 
@@
 The results displayed in Figure~\ref{fig:likelihood-profiles-comparison-with-division} show that the
-values for the division lengths seem exhibit a decrease in steepness of the curve towards the left
-side of the estimated parameter.
+values for the division lengths exhibit a decrease in steepness of the curve towards the left side
+of the estimated parameter.
 This can be attributed to the missing images which were not able to be segmented in our time series.
+It can be argued that this leads to a non-identifiability for the first division length while the
+remaining parameters are still identifiable although the effect is also notable.
 These results highlight a reassuring finding, namely that the method of profile-likelihood is able
-to properly display this obvious lack of data.
-It could be argued that this leads to a non-identifiability for the first division length while the
-reamining parameters are still identifiable.
-Meanwhile, the growth rates of the daughters ...
+to properly display this lack of data.
+It therefore further underlines the accuracy of our methods.
+In comparison, the growth rates of the daughter cells display larger variations within their
+likelihood profiles.
+\textbf{TODO finish this; need more results for otimization of all daughter growth rates as well}
+
+It has to be noted that this approach is computationally much more expensive compared to the simpler
+position-comparison algorithm.
+There are 4 main steps which are executed within any single optimization iteration that contribute
+to the total wall time.
+These steps are prediction via the computational model ($15.63(21)\%$), mask adjustment of the data
+masks ($14.49(19)\%$), mask rendering for the numerical results ($69.2(1.4)\%$) and finally
+comparison of the masks ($0.68(5)\%$) (see section \ref{section:supplement-performance}).
+The masks which are provided by the data need to be adjusted such that the identifiers of the cells
+which are directly related to the chosen colors match with the ones used within the masks provided
+by the data.
+It is clear that the generation of masks from the numerical results takes up the majority of the
+work.
+In addition, the step of mask adjustment would not have been required within the position-comparison
+algorithm.
+Combining these two values, the position-comparison algorithm is $\approx 6.13(8)$ times faster.
 
@@
 \subsection{Performance (TODO)}
+\label{section:supplement-performance}
 
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{figures/crm_divide/timings.pdf}%
+    \includegraphics[width=0.5\textwidth]{figures/crm_divide/time-evolution.pdf}%
+    \caption{TODO}
+    \label{fig:timings-crm_divide}
+\end{figure}
+
 \end{document}

### 2025-09-20 paper/main.tex
cleanup
use correct images
insert description for figure

@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_divide/profiles/profile-000005.png}%
+        {figures/crm_divide/profiles/profile-000006.png}%
         \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_divide/profiles/profile-000005.png}%
+        {figures/crm_divide/profiles/profile-000007.png}%
         \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{H}};
     \end{tikzonimage}%
-    \caption{TODO}
+    \caption{
+        Likelihood profiles of every estimated parameter.
+        (A-D) Division Lengths of initially placed cells.
+        (E-H) Growth rates of daughter cells.
+        The growth rates diplay larger values of noise within their profiles compared to the
+        division lengths.
+        \textbf{TODO possibly extend this description?}
+    }
     \label{fig:likelihood-profiles-comparison-with-division}
@@
 \end{itemize}
-
-%###################################################################################################
-% EXCLUDE THIS FOR NOW
-% \section{Plastic Deformations}
-% 
-% \begin{figure}
-%     \centering
-%     \includegraphics[width=0.49\textwidth]
-%         {docs/source/_static/scripts/crm_amir/0000000023.png}\\
-%     \includegraphics[width=0.49\textwidth]
-%         {docs/source/_static/scripts/crm_amir/angles.pdf}
-%     \includegraphics[width=0.49\textwidth]
-%         {docs/source/_static/scripts/crm_amir/endpoints.pdf}
-%     \caption{TODO}
-%     \label{fig:amir}
-% \end{figure}
 

### 2025-09-20 cr_mech_coli/crm_divide.py
new parameters, bounds and optimization settings

@@ -812,14 +812,16 @@ def run_optimizer(
             bounds=bounds,
             args=args,
             disp=True,
-            maxiter=50,
+            maxiter=100,
             popsize=15,
-            mutation=(0.3, 1.2),
+            mutation=(0.0, 1.5),
             recombination=0.6,
             tol=0.0001,
             workers=n_workers,
             updating="deferred",
             polish=True,
+            init="latinhypercube",
+            strategy="best1bin",
         )
         final_parameters = res.x
         final_cost = res.fun
@@ -933,22 +935,22 @@ def crm_divide_main():
             exit()
 
     spring_length_thresholds = [
-        8.1,
-        9.0,
-        7.2,
-        6.1,
+        9.405841188088112759e00,
+        1.103879179742345329e01,
+        9.277040994277374608e00,
+        7.600778468661159692e00,
     ]
     new_growth_rates = [
-        0.0023,
-        0.0012,
-        0.0020,
-        0.0025,
+        2.186928453188154847e-03,
+        4.834987621055546192e-04,
+        1.669060863394238470e-03,
+        2.300937616285135622e-03,
     ]
     x0 = [
         *spring_length_thresholds,
         *new_growth_rates,
     ]
-    bounds = [(5, 12)] * 4 + [(0.0000, 0.006)] * 4
+    bounds = [(5, 12)] * 4 + [(0.0000, 0.004)] * 4
     parent_penalty = 0.5
     args = (
         positions_all,

### 2025-09-20 paper/main.tex
add more details to model description

@@
 We use principles from discrete differential geometry~\cite{bobenko2008discrete} to describe the
-stiffness and flexibility of the bacterial rod.
+stiffness and flexibility of the bacterial rod~\cite{Amir2014_2}.
 As explained previously, MReB will counteract the bending of the rod, resulting in a stiffening
@@
 vary.
-We reduce the growth rate with increasing number of neighbors $N$.
-This behaviour can be toggled on or off thus removing the additional term in brackets of
+Our implemented model allows us to reduce the growth rate with increasing number of neighbors
+$n\leq N$.
+This option provides a phenomenological approach for modeling slower growth in larger colonies but
+is be toggled off by default thus removing the additional term in brackets of
 equation~\ref{eq:growth-rod-lengths}.
@@
 
-Note that the growth rate is proportional to the rod length $l$ which leads to exponential growth.
+Note that the growth rate is proportional to the rod length $l$ which leads to exponential
+growth~\cite{Robert2014,Takeuchi2005}.
 This process will increase the length of the cell indefenitely unless we impose a condition for the
@@
 We removed any stochastic motion contributions by setting $D=0$.
-This leads to a total of $5+1\times 6=11$ parameters.
+This leads to a total of $5+1\times 6=11$ variable parameters.
+In order to model non-flexible rods, we set the rigidity and spring tension of the rod to a high,
+fixed value.
 To estimate these parameters we used the differential evolution~\cite{Storn1997} global optimization

### 2025-09-23 paper/main.tex
add statement about axiliary nature of rigidity and spring tension

@@
 
+The value of the spring constants determines the elasticity with respect to elongation of the cell.
+In situations where this aspect is not relevant, they merely serve as a auxiliary parameters to
+maintain the length of the rod.
 In addition to springs between individual vertices $\vec{x}_i$, we assume that each angle at a
@@
 This leads to a total of $5+1\times 6=11$ variable parameters.
-In order to model non-flexible rods, we set the rigidity and spring tension of the rod to a high,
-fixed value.
+As highlighted before, in order to model non-flexible rods, we set the rigidity and spring tension
+of the rod to a high, fixed value.
 To estimate these parameters we used the differential evolution~\cite{Storn1997} global optimization

### 2025-09-23 paper/main.tex
add citation

@@
 growth~\cite{Billaudeau2017}, division~\cite{Harry2001} and their collective
-behaviour~\cite{vanGestel2015}.
+behaviour~\cite{vanGestel2015,Balagam2015}.
 Furthermore, it has been discovered that the rod can be bent thus introducing temporary elastic or

### 2025-09-23 cr_mech_coli/crm_amir/main.py
implement simulation of relaxation part

@@ -29,7 +29,7 @@ def calculate_angle(p: np.ndarray, parameters: crm_amir.Parameters) -> float:
     return angle
 
 
-def generate_parameters() -> crm_amir.Parameters:
+def generate_parameters(**kwargs) -> crm_amir.Parameters:
     parameters = crm_amir.Parameters()
     parameters.block_size = 25.0
     parameters.dt = 0.01
@@ -39,10 +39,13 @@ def generate_parameters() -> crm_amir.Parameters:
     parameters.n_vertices = n_vertices
     parameters.growth_rate = 0.03 * 7 / (n_vertices - 1)
     parameters.rod_rigidity = 20.0 * n_vertices / 20
-    parameters.save_interval = 1.0
+    parameters.save_interval = 20
     parameters.damping = 0.02
     parameters.spring_tension = 10.0
     parameters.drag_force = 0.03
+
+    for k, v in kwargs:
+        parameters.__setattr__(k, v)
     return parameters
 
 
@@ -213,15 +216,19 @@ def objective_function(
         return ERROR_COST
 
     # Get initial and final position of rod
-    p0 = rods[0][1].agent.pos[:, np.array([0, 2])]
-    p1 = rods[-1][1].agent.pos[:, np.array([0, 2])]
+    p_rods = np.array(
+        [
+            rods[0][1].agent.pos[:, np.array([0, 2])],
+            rods[1][1].agent.pos[:, np.array([0, 2])],
+            rods[2][1].agent.pos[:, np.array([0, 2])],
+        ]
+    )
 
-    p0[:, 1] = parameters.domain_size - p0[:, 1]
-    p1[:, 1] = parameters.domain_size - p1[:, 1]
+    p_rods[:, :, 1] = parameters.domain_size - p_rods[:, :, 1]
 
     # Shift such that start points align
 
-    positions_data = np.array(positions_data)
+    positions_data = np.array([*positions_data, positions_data[0]])
     positions_data = (
         np.array([parameters.domain_size, 0]) - np.array([1, -1]) * positions_data
     )
@@ -229,14 +236,14 @@ def objective_function(
     for i in range(positions_data.shape[0]):
         positions_data[i, 0, 0] -= positions_data[i, 0, 0]
     x_shift_positions0 = calculate_x_shift(positions_data[0], parameters.block_size)
-    x_shift_p0 = calculate_x_shift(p0, parameters.block_size)
[…]

### 2025-09-23 src/crm_amir.rs

@@ -344,10 +344,38 @@ fn run_sim(
     Ok(cells)
 }
 
+#[pyfunction]
+#[pyo3(signature = (parameters, initial_pos=None))]
+fn run_sim_with_relaxation(
+    py: Python,
+    parameters: &Parameters,
+    initial_pos: Option<Bound<numpy::PyArray2<f32>>>,
+) -> PyResult<Vec<(u64, FixedRod)>> {
+    // Run first iteration of simulation
+    let mut rods = run_sim(parameters.clone(), initial_pos)?;
+    // Obtain final position of rod
+    let p = rods[1].1.agent.mechanics.pos.clone();
+    let initial_pos =
+        ndarray::Array2::<f32>::from_shape_vec(p.shape(), p.into_iter().copied().collect())
+            .unwrap();
+    // Set final position of rod as initial position for second run
+    let initial_pos = numpy::PyArray2::from_owned_array(py, initial_pos);
+    let mut new_parameters = parameters.clone();
+    // Remove the drag force and run simulation again
+    new_parameters.drag_force = 0.0;
+    let (iter_final, agent) = run_sim(new_parameters, Some(initial_pos))?[1].clone();
+
+    // Append last rods to results
+    rods.push((rods[1].0 + iter_final, agent));
+    assert_eq!(rods.len(), 3);
+    Ok(rods)
+}
+
 /// A Python module implemented in Rust.
 pub fn crm_amir(py: Python) -> PyResult<Bound<PyModule>> {
     let m = PyModule::new(py, "cr_mech_coli.crm_amir.crm_amir_rs")?;
     m.add_function(wrap_pyfunction!(run_sim, &m)?)?;
+    m.add_function(wrap_pyfunction!(run_sim_with_relaxation, &m)?)?;
     m.add_class::<FixedRod>()?;
     m.add_class::<Parameters>()?;
     Ok(m)

### 2025-09-23 src/agent.rs
implement radius setter for agent

@@ -420,6 +420,15 @@ impl RodAgent {
             PhysInt::MiePotentialF32(pot) => pot.radius,
         }
     }
+
+    /// Sets the radius of the interaction potential
+    #[setter]
+    pub fn set_radius(&mut self, radius: f32) {
+        match &mut self.interaction.0 .0 {
+            PhysInt::MorsePotentialF32(pot) => pot.radius = radius,
+            PhysInt::MiePotentialF32(pot) => pot.radius = radius,
+        }
+    }
 }
 
 impl Cycle<RodAgent, f32> for RodAgent {

### 2025-09-23 cr_mech_coli/crm_amir/main.py
rework saving of snapshots

@@ -500,27 +500,22 @@ def __render_single_snapshot(iter, agent, parameters, render_settings, output_di
     img[:, :block_size][bg_filt[:, :block_size]] = int(
         render_settings.bg_brightness / 2
     )
-    cv.imwrite(f"out/crm_amir/{iter:010}.png", np.swapaxes(img, 0, 1)[::-1])
+    cv.imwrite(output_dir / f"{iter:010}.png", np.swapaxes(img, 0, 1)[::-1])
 
 
-def render_snapshots():
-    parameters = generate_parameters()
-    agents = crm_amir.run_sim(parameters)
-
-    n_saves = 10
-
-    save_points = np.clip(
-        np.round(np.linspace(0, len(agents), n_saves)).astype(int), 0, len(agents) - 1
-    )
-
+def render_snapshots(agents, parameters, output_dir):
     render_settings = crm.RenderSettings()
     render_settings.bg_brightness = 200
 
-    for save_point in tqdm(
-        save_points, total=len(save_points), desc="Rendering Snapshots"
+    for n, agent in tqdm(
+        enumerate(agents), total=len(agents), desc="Rendering Snapshots"
     ):
         __render_single_snapshot(
-            save_point, agents[save_point][1].agent, parameters, render_settings
+            n,
+            agent,
+            parameters,
+            render_settings,
+            output_dir,
         )
 
 
@@ -534,7 +529,10 @@ def crm_amir_main():
         pyargs.workers = mp.cpu_count()
 
     crm.plotting.set_mpl_rc_params()
-    # render_snapshots()
+
+    parameters = generate_parameters()
+    agents = [x[1].agent for x in crm_amir.run_sim(parameters)]
+    render_snapshots(agents, parameters, Path("out/crm_amir/"))
 
     # Define which parameters should be optimized
     x0_bounds = {

### 2025-09-23 cr_mech_coli/crm_amir/main.py
pick new parameters for optimization

@@ -537,18 +537,35 @@ def crm_amir_main():
     # Define which parameters should be optimized
     x0_bounds = {
         "rod_rigidity": (0.0001, 20.0, 250),  # rod_rigidity,
-        "drag_force": (0.0000, 0.1, 10.0),  # drag_force,
+        "drag_force": (0.0000, 0.1, 3.0),  # drag_force,
         "damping": (0.000, 1.0, 2.0),  # damping,
-        "growth_rate": (0.0, 0.01, 2.0),  # growth_rate,
+        "growth_rate": (0.0, 0.01, 0.15),  # growth_rate,
         "spring_tension": (0.0000, 0.01, 30.0),  # spring_tension
     }
     compare_with_data(x0_bounds, workers=pyargs.workers)
 
     x0_bounds_reduced = {
         "rod_rigidity": (0.0001, 20.0, 250),  # rod_rigidity,
-        "drag_force": (0.0000, 0.1, 10.0),  # drag_force,
+        "drag_force": (0.0000, 0.1, 3.0),  # drag_force,
+        "damping": (0.000, 1.0, 2.0),  # damping,
+        "growth_rate": (0.0, 0.01, 0.15),  # growth_rate,
+        # "spring_tension": (0.0000, 0.01, 30.0),  # spring_tension
+    }
+    set_params = {
+        "spring_tension": 20,
+    }
+    compare_with_data(
+        x0_bounds_reduced,
+        workers=pyargs.workers,
+        set_params=set_params,
+        output_dir="out/crm_amir/profiles-without-spring-tension/",
+    )
+
+    x0_bounds_reduced = {
+        "rod_rigidity": (0.0001, 20.0, 250),  # rod_rigidity,
+        "drag_force": (0.0000, 0.1, 3.0),  # drag_force,
         # "damping": (0.000, 1.0, 2.0),  # damping,
-        "growth_rate": (0.0, 0.01, 2.0),  # growth_rate,
+        "growth_rate": (0.0, 0.01, 0.15),  # growth_rate,
         # "spring_tension": (0.0000, 0.01, 30.0),  # spring_tension
     }
     set_params = {

### 2025-09-23 paper/main.tex
add more details to equations of motion
fill description for crm_amir part

@@
 
-and are integrated via
+and are integrated via damped brownian dynamics
 
@@
 
-where $D$ is the diffusion constant and $\vec{\xi}$ is the wiener process.
+where $D$ is the diffusion constant, $\vec{\xi}$ is the wiener process and $\lambda$ a damping
+constant.
 Together they determine the stochastic motion of the individual vertices which can arise from
@@
 %###################################################################################################
-\section{Application to \textit{E.Coli.} growing on Agar}
-\label{section:application}
+\section{Application 1: Elastic deformations of \textit{E.Coli} (TODO)}
+\label{section:application-1}
+
+We will use our previously developed computational model and apply it directly to the data provided
+by the experiments of Amir et. al~\cite{Amir2014,Amir2014_2}.
+These images show a singular \textit{E.Coli} which is fixed at the bottom and is subject to a
+horizontal fluid flow that exerts a force onto the rod.
+This setup is displayed in figure~\ref{fig:amir-bending-simulation} (A).
+In order to model this setup, we also fixate the lower-most vertex of our rod at the bottom of the
+simulation domain.
+Furthermore, we introduce a force $\vec{F}_i = C \sin(\alpha_i)\vec{e}_x$ that acts similarlyl to
+Stoke's Law \textbf{TODO Citation} onto the normal component of each rod-segment where $\alpha_i$ is
+the angle of the rod edge between $\vec{x}_i$ and $\vec{x}_{i+1}$ and the x-axis.
+We extract the position of the rod from the provided image and use it as intial value.
+Our simulation is split into two phases, one with an external force that initially bends the rod,
+and a second one during which the force is disabled and the cell returns to its original shape.
+\textbf{TODO maybe exclude second phase again}
+We compare the simulated positions with the extracted ones and calculate the euclidean distance
+between them.
+This value serves as the cost function that allows us to estimate the parameters of this model.
+Since this model is void of any interactions or cell division, we are left with the 5 parameters
+damping 
+Figure~\ref{fig:amir-bending-simulation} (B) shows a snapshot of the initial (orange) and final
+(blue) position of the rod using the optimized parameters.
+Unfortunately the images only contain snapshots of a fully-bent or fully-relaxed rod without any
+intermediate steps.
+This effectively limits us to 2 datapoints and prohibits us from accurately estimating the
+parameters involved in the dynamics of the rod-bending.
+This effect can be seen in the likelihood profiles of the individual parameters.
+While the dragging force, growth rate and damping can be identified properly, the same can not be
+said for the rigidity and spring tension.
+The rigidity of the rod and spring tension exhibit non-identifiabilities with respect to larger
+values.
+Both of these behaviours are expected since firstly, the shape of the rod is almost exclusively
+straight, thus adding no further information except that a lower threshold of rigidity and spring
+tension should be present.
+
+\begin{itemize}
+    \item discuss why damping and spring tension can not be estimated: damping - missing dynamics,
+        spring tension - no elasticity with respect to elongation, more tension = better shape
+        preservation
+    \item TODO: also model relaxation part
+\end{itemize}
+
+\begin{figure}
+    \centering
+    \begin{minipage}{0.33\textwidth}
+        \begin{tikzonimage}[width=\textwidth]
+            {figures/amir-elastic-000032-wider.png}%
+            \node at (0.01, 0.01+17pt)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+        \end{tikzonimage}%
+    \end{minipage}%
+    \begin{minipage}{0.66\textwidth}
+        \begin{tikzonimage}[width=0.33\textwidth]
+            {figures/crm_amir/profiles-full/fit-comparison.pdf}%
+            \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \end{tikzonimage}%
+        \begin{tikzonimage}[width=0.33\textwidth]
+            {figures/crm_amir/profiles-full/drag force.pdf}%
+            \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+        \end{tikzonimage}%
+        \begin{tikzonimage}[width=0.33\textwidth]
+            {figures/crm_amir/profiles-full/growth rate.pdf}%
+            \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+        \end{tikzonimage}\\
+        \begin{tikzonimage}[width=0.33\textwidth]
+            {figures/crm_amir/profiles-full/rod rigidity.pdf}%
+            \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+        \end{tikzonimage}%
+        \begin{tikzonimage}[width=0.33\textwidth]
+            {figures/crm_amir/profiles-full/spring tension.pdf}%
+            \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
+        \end{tikzonimage}%
+        \begin{tikzonimage}[width=0.33\textwidth]
+            {figures/crm_amir/profiles-full/damping.pdf}%
+            \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
+        \end{tikzonimage}%
+    \end{minipage}
+    \caption{
+        TODO \cite{Amir2014_2}
+    }
+    \label{fig:amir-bending-simulation}
+\end{figure}
+
+%###################################################################################################
+\section{Application 2: \textit{E.Coli.} growing on Agar}
+\label{section:application-2}
 

### 2025-09-25 paper/main.tex
fill description of figure

@@
     \caption{
-        Progression of generated masks with estimated and exact positions.
-        Comparison of masks across generations. Compare estimated with exact rod lengths.
-"Iceberg plot"
+        Subfigures (A-C) show rendered masks of a simulation where the extracted positions have been
+        drawn within the individual agents.
+        Subfigure (D) displayes the average rod length of the exact and extracted data.
+        The light blue region is the mean distance between vertices of extracted and exact
+        positions which captures the inherent uncertainty of our method.
+        Subfigure (E) displays all of these distances on a logarithmic scale.
+        The colorscale represents earlier (darker) and later (brighter) temporal states of our
+        simulation.
+        Overall, this plot shows a widening of the distance distributions for increasing simulation
+        times.
     }

### 2025-09-25 paper/main.tex
declare new unit pixels
finish writing about amir experiment

@@
 \newacro{pbr}[PBR]{Physically-Based Rendering}
+
+% Define new units
+\DeclareSIUnit\pixel{pix}
 
@@
 simulation domain.
-Furthermore, we introduce a force $\vec{F}_i = C \sin(\alpha_i)\vec{e}_x$ that acts similarlyl to
-Stoke's Law \textbf{TODO Citation} onto the normal component of each rod-segment where $\alpha_i$ is
-the angle of the rod edge between $\vec{x}_i$ and $\vec{x}_{i+1}$ and the x-axis.
+Furthermore, we introduce a force $\vec{F}_i = C |\vec{x}_i - \vec{x}_{i+1}|\sin(\alpha_i)\vec{e}_x$
+that acts similarlyl to Stoke's Law \textbf{TODO Citation} onto the normal component of each
+rod-segment where $\alpha_i$ is the angle of the rod edge between $\vec{x}_i$ and $\vec{x}_{i+1}$
+and the x-axis.
 We extract the position of the rod from the provided image and use it as intial value.
@@
 and a second one during which the force is disabled and the cell returns to its original shape.
-\textbf{TODO maybe exclude second phase again}
 We compare the simulated positions with the extracted ones and calculate the euclidean distance
@@
 This value serves as the cost function that allows us to estimate the parameters of this model.
-Since this model is void of any interactions or cell division, we are left with the 5 parameters
-damping 
+Since this model is void of any interactions or cell division, we are left with 5 parameters:
+Magnitude of the drag force $C$, growth rate $\alpha$, rigidity of the rod $\eta$, spring tension
+$\gamma$ and damping constant $\lambda$.
 Figure~\ref{fig:amir-bending-simulation} (B) shows a snapshot of the initial (orange) and final
 (blue) position of the rod using the optimized parameters.
-Unfortunately the images only contain snapshots of a fully-bent or fully-relaxed rod without any
-intermediate steps.
-This effectively limits us to 2 datapoints and prohibits us from accurately estimating the
-parameters involved in the dynamics of the rod-bending.
-This effect can be seen in the likelihood profiles of the individual parameters.
-While the dragging force, growth rate and damping can be identified properly, the same can not be
-said for the rigidity and spring tension.
-The rigidity of the rod and spring tension exhibit non-identifiabilities with respect to larger
-values.
-Both of these behaviours are expected since firstly, the shape of the rod is almost exclusively
-straight, thus adding no further information except that a lower threshold of rigidity and spring
-tension should be present.
-
-\begin{itemize}
-    \item discuss why damping and spring tension can not be estimated: damping - missing dynamics,
-        spring tension - no elasticity with respect to elongation, more tension = better shape
-        preservation
-    \item TODO: also model relaxation part
-\end{itemize}
+While the dragging force, growth rate and damping can be identified properly, the same
+can not be said for the rod rigidity and spring tension which both exhibit non-identifiabilities.
+However this behaviour is expected since, the shape of the rod is almost exclusively straight,
+thus adding no further information except that a lower threshold of rigidity and spring tension
+should be present.
+Note that each point represented in the subfigures~\ref{fig:amir-bending-simulation} (C-G) was
+optimized again, meaning that higher values of rigidity and spring tension can also lead to higher
+values of drag forces.
+Furthermore the used images only contain snapshots of a fully-bent or fully-relaxed rod without any
+intermediate steps that cannot fully capture its dynamic.
+This effectively limits us to 2 datapoints and is likely the key factor that leads to the two local
+minima observed in the damping profile.
+This is further supported by the observation that the contributions to the cost function is evenly
+distributed between the bent and relaxed position of our system.
+In order to better understand this system, we reduce it by inserting a new parameter that determines
+the values of rod rigidity and spring tension via $\gamma=\gamma_0/s$ and $\eta=\eta_0 s$.
+We then optimize for this parameter and pick $\gamma_0=\SI{15}{\pixel}$ with
+$\eta_0=\SI{150}{\pixel\per\second\squared}$ which corresponds to $60\%$ of the previously considered
+interval.
+With this change, all parameters are identifiable.
+The corresponding plots can be seen in \ref{sec:supplement-parameter-profiles}.
 

### 2025-09-25 paper/main.tex
link to correct subsection of supplement

@@
 With this change, all parameters are identifiable.
-The corresponding plots can be seen in \ref{sec:supplement-parameter-profiles}.
+The corresponding plots can be seen in \ref{subsec:supplement-parameter-profiles-amir}.
 
@@
     \caption{
-        TODO \cite{Amir2014_2}
+        Subfigure (A) shows a snapshot of the experiments performed by \cite{Amir2014_2}.
+        Division of the bacteria is inhibited and they are trapped within a mother machine such that
+        a controlled flow can be exerted on them.
+        (B) A comparison of the initial and bent state show good visual agreement of our prediction
+        with the experimental data.
+        Subfigures (C-G) show likelihood profiles for the individual parameters.
     }
@@
 \label{sec:supplement-parameter-profiles}
+
+\subsubsection{Amir}
+\label{subsec:supplement-parameter-profiles-amir}
+
+The parameter ranges were chosen such that a suitable 
+Plots are chosen for visual ease; parameter ranges have been widened to ensure results are
+accurate (maybe just leave this in the appendix).
+
 %TODO {explain profiles}

### 2025-09-25 cr_mech_coli/crm_amir/main.py
implement combined parameter

@@ -205,11 +205,19 @@ def objective_function(
     parameters, t_relax = create_default_parameters(positions_data, iterations_data)
 
     for k, v in set_params.items():
-        parameters.__setattr__(k, v)
+        if k == "rigidity_to_spring_tension_ratio":
+            parameters.rod_rigidity = set_params["rod_rigidity"] * v
+            parameters.spring_tension = set_params["spring_tension"] / v
+        else:
+            parameters.__setattr__(k, v)
 
     # Variable Parameters
     for name, value in zip(x0_bounds.keys(), params):
-        parameters.__setattr__(name, value)
+        if name == "rigidity_to_spring_tension_ratio":
+            parameters.rod_rigidity = set_params["rod_rigidity"] * value
+            parameters.spring_tension = set_params["spring_tension"] / value
+        else:
+            parameters.__setattr__(name, value)
 
     try:
         rods = crm_amir.run_sim_with_relaxation(parameters, t_relax)
@@ -684,10 +692,11 @@ def crm_amir_main():
         "rod_rigidity": rod_rigidity,
         "drag_force": drag_force,
         "growth_rate": growth_rate,
+        "rigidity_to_spring_tension_ratio": (0.0, 1.0, 10.0, "[pix]"),
     }
     set_params = {
-        "damping": popt["damping"],
-        "spring_tension": set_params["spring_tension"],
+        "rod_rigidity": 150.0,
+        "spring_tension": 18.0,
     }
     compare_with_data(
         x0_bounds_reduced,

### 2025-09-25 cr_mech_coli/crm_amir/main.py
update optimization arguments and fix function args

@@ -648,11 +648,11 @@ def crm_amir_main():
     positions_data, iterations_data = obtain_data("out/crm_amir/", n_vertices=20)
 
     # Define globals for all optimizations
-    rod_rigidity = (0.0, 20.0, 350, "[pix/s²]")
-    drag_force = (0.0000, 0.1, 0.5, "[1/s²]")
-    damping = (0.000, 1.0, 0.15, "[1/s]")
+    rod_rigidity = (0.0, 20.0, 250, "[pix/s²]")
+    drag_force = (0.0000, 0.1, 1.5, "[1/s²]")
+    damping = (0.000, 1.0, 1.0, "[1/s]")
     growth_rate = (0.0, 0.01, 0.05, "[1/s]")
-    spring_tension = (5, 30.0, 150.0, "[1/pix]")
+    spring_tension = (5, 30.0, 30.0, "[1/s²]")
 
     # Define which parameters should be optimized
     x0_bounds = {
@@ -662,35 +662,34 @@ def crm_amir_main():
         "growth_rate": growth_rate,
         "spring_tension": spring_tension,
     }
-    popt = compare_with_data(
+    compare_with_data(
         x0_bounds,
         positions_data,
         iterations_data,
         pyargs,
     )
-    exit()
 
     x0_bounds_reduced = {
-        "rod_rigidity": rod_rigidity,
         "drag_force": drag_force,
         "damping": damping,
         "growth_rate": growth_rate,
+        "spring_tension": spring_tension,
     }
     set_params = {
-        "spring_tension": popt["spring_tension"],
+        "rod_rigidity": 150,
     }
-    popt = compare_with_data(
+    compare_with_data(
         x0_bounds_reduced,
         positions_data,
         iterations_data,
-        workers=pyargs.workers,
+        pyargs,
         set_params=set_params,
-        output_dir="out/crm_amir/profiles-without-spring-tension/",
+        output_dir="out/crm_amir/profiles-fixed-rod-rigidity/",
     )
 
     x0_bounds_reduced = {
-        "rod_rigidity": rod_rigidity,
         "drag_force": drag_force,
+        "damping": damping,
         "growth_rate": growth_rate,
         "rigidity_to_spring_tension_ratio": (0.0, 1.0, 10.0, "[pix]"),
     }
@@ -702,8 +701,9 @@ def crm_amir_main():
         x0_bounds_reduced,
         positions_data,
         iterations_data,
-        workers=pyargs.workers,
[…]

### 2025-09-25 cr_mech_coli/crm_amir/main.py
fix import

@@ -10,6 +10,7 @@ from tqdm import tqdm
 from pathlib import Path
 import multiprocessing as mp
 import argparse
+import itertools
 
 GREEN_COLOR = np.array([95, 231, 76])
 

### 2025-09-25 cr_mech_coli/crm_amir/main.py
write comment

@@ -279,7 +279,7 @@ def plot_results(
     p0 = p_rods[0]
     p1 = p_rods[1]
 
-    # Plot distribution of cost between bending and relaxation
+    #
     fig, ax = plt.subplots(figsize=(8, 8))
     diffs = np.linalg.norm(positions_data[1:] - p_rods[1:], axis=(1, 2))
     labels = ["Bending", "Relaxation"]

### 2025-09-25 paper/main.tex
fix color
include profiles for daughter cells

@@
         {figures/crm_fit/celldiffs/diff-003889.png}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
     \end{tikzonimage}%
@@
         \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{H}};
+    \end{tikzonimage}\\
+    \begin{tikzonimage}[width=0.25\textwidth]
+        {figures/crm_divide/profiles/profile-000008.png}%
+        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{I}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.25\textwidth]
+        {figures/crm_divide/profiles/profile-000009.png}%
+        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{J}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.25\textwidth]
+        {figures/crm_divide/profiles/profile-000010.png}%
+        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{K}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.25\textwidth]
+        {figures/crm_divide/profiles/profile-000011.png}%
+        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{L}};
     \end{tikzonimage}%

### 2025-09-25 paper/main.tex
fix reference

@@
 dependency.\\
-The time series of microscopic images which has been studied in section~\ref{section:application}
+The time series of microscopic images which has been studied in section~\ref{section:application-2}
 was extracted from individual frames of a video~\cite{https://doi.org/10.3203/iwf/k-129} that

### 2025-09-29 paper/main.tex
add points for future ideas
remove from description
adjust formatting
adjust text to new optimization setup; include figure for amir

@@
         simulation.
-        Overall, this plot shows a widening of the distance distributions for increasing simulation
-        times.
     }
@@
     \centering
-    \begin{minipage}{0.33\textwidth}
+    \begin{minipage}{0.38\textwidth}
         \begin{tikzonimage}[width=\textwidth]
@@
     \end{minipage}%
-    \begin{minipage}{0.66\textwidth}
+    \begin{minipage}{0.62\textwidth}
         \begin{tikzonimage}[width=0.33\textwidth]
@@
 To show the usefulness of our previously developed methods (see
-section~\ref{section:parameter-estimation}), we will use them to estimate the parameters of a
-sequence of microscopic images of growing \texttt{E.Coli}.
+section~\ref{section:parameter-estimation}), we will use them to estimate the parameters individual
+growing \textit{E.Coli} using a sequence of microscopic images.
 The data we use~\cite{https://doi.org/10.3203/iwf/k-129} (see
@@
 We have already suspected (see table~\ref{table:simulation-aspects}) that in order to properly
-estimate parameters of our system, we need to treat every cell-agent individually, thus assigning
+estimate parameters of our system, we need to treat every cell individually, thus assigning
 parameters on a single-cell level.
-This would mean to increase the total number of parameters that need to be sampled within an
+This means that we increase the total number of parameters that need to be sampled within an
 optimization routine.
 Given $N_c$ cells and $N_p$ parameters per cell and $N_u$ individual parameters (where
-$N_u\leq N_p$), the total amount of parameters is
+$N_u\leq N_p$), the total amount of parameters to be estimated is
 
@@
 It is clear, that we should avoid to needlessly assign every parameter individually since the volume
-of the space that needs to be sampled, scales exponentially with the dimension ie. the number of
-parameters.
+of the parameter space that needs to be sampled, scales exponentially with the dimension ie. the
+number of parameters.
 However, we will now present data that indicate that an individual treatment is in fact necessary.
@@
 For now, we will focus on a time-interval which does not contain any division events and at which
-the initial lag-phase of our bacteria os already passed.
+the initial lag-phase of our bacteria has already passed.
 By doing so, we can utilize the simpler and more performant position comparison scheme for our loss
@@
 optimization algorithm.
-For this dataset, we assumed that all parameters which specify the interaction potential are
-identical across all agents but growth rates of the cells are individual.
-
+% For this dataset, we assumed that all parameters which specify the interaction potential are
+% identical across all agents but growth rates of the cells are individual.
 We chose to optimize for the damping of agents' movement within the medium, 3 interaction parameters
-of the Mie potential $n,m,V_0$ a single radius for all agents and growth rates which were chosen
+of the Mie potential $n,m,V_0$ a single radius $r$ for all agents and growth rates which were chosen
 individually for each cell.
@@
 Figure~\ref{fig:parameter-estimates-single-step} shows profile-likelihoods for each optimized
-parameter with respect to the optimization results described before.
+parameter with respect to the optimization results.
 We can see that all parameters can be estimated by our optimization approach and that no
 non-identifiability~\cite{Raue2009} is present.
+These plots also show that the individual growth rates of the bacteria can be estimated effectively.
 
@@
 
-In the next step, we extend our previously used time-series with additional images, which include
-cell division events which means that we now have to use the imaing-based comparison algorithm (see
+In the next step, we extend our previously used time-series with successive images including
+cell division events which means that we now have to use the imaging-based comparison algorithm (see
 section~\ref{section:parameter-estimation}).
@@
 
-Since we have already optimized said system, we can reduce the computational demand by choosing the
-parameters which were obtained by our previous attempts and extend them with the parameters that the
-daughter cells will carry.
-In addition to assigning growth rates individually, but we also need to consider division lengths at
-which the bacteria split.
-Since our simulation involves the division of 4 cells into 8 new ones, in principle, this would lead
-us to a total of $8$ new growth rates and $4$ division lengths.
-But to decrease the size of the parameter space, we assume that the growth rates of the daughter
-cells are identical between every sibling, leaving us at only $4$ new growth rates and a total of
-$8$ parameters.
-Future approaches should aim to loosen this assumption, allowing to optimize every new growth rate.
-To optimize this system, we again used the same optimization approach by seed with samples generated
-by latin-hypercube~\cite{McKay1979}, then applying the differential evolution
-algorithm~\cite{Storn1997} and finally a local minimization via L-BFGS-B~\cite{Liu1989}.
+% Since we have already optimized said system, we can reduce the computational demand by choosing the
+% parameters which were obtained by our previous attempts and extend them with the parameters that the
+% daughter cells will carry.
+We optimize for the same parameters as done before.
+In addition to assigning growth rates individually, we also need to consider division lengths at
+which the bacteria split and the new individual growth rates of the daughter cells as additional
+parameters.
+Since our simulation involves the division of 4 cells into 8 new ones, in principle, this leads
+to a total of $8$ new growth rates and $4$ division lengths.
+% But to decrease the size of the parameter space, we assume that the growth rates of the daughter
+% cells are identical between every sibling, leaving us at only $4$ new growth rates and a total of
+% $8$ parameters.
+% Future approaches should aim to loosen this assumption, allowing to optimize every new growth rate.
+To optimize this system, we used a similar optimization approach by again seeding with samples
+generated via latin-hypercube~\cite{McKay1979}, then applying the differential evolution
+algorithm~\cite{Storn1997} and finally a local minimization with Nelder-Mead~\cite{Gao2010}.
 We used the value $0.5$ for the parent penalty, which aims to provide a smooth transition between
@@
 given by the data and the cell mask predicted by the best fit of our model.
+As expected, the fully-optimized system does not produce differences where the parental penalty is
+required.
+However, its value is frequently used within the the optimization scheme itself.
 
 We can again calculate likelihood profiles for every optimized parameter.
-The results displayed in Figure~\ref{fig:likelihood-profiles-comparison-with-division} show that the
-values for the division lengths exhibit a decrease in steepness of the curve towards the left side
-of the estimated parameter.
-This can be attributed to the missing images which were not able to be segmented in our time series.
+The results displayed in Figure~\ref{fig:likelihood-profiles-comparison-with-division} show that
+some curves exhibit a plateau close to the calculated optimum of the estimated parameter.
+This concerns the division lengths and some growth rates of the mother bacteria which both determine
+if and when a cell divides.
+Such behaviour can be attributed to the missing images around the division events of the cells which
+were not able to be segmented.
+\textbf{TODO check this with data}
 It can be argued that this leads to a non-identifiability for the first division length while the
@@
 to properly display this lack of data.
-It therefore further underlines the accuracy of our methods.
-In comparison, the growth rates of the daughter cells display larger variations within their
-likelihood profiles.
-\textbf{TODO finish this; need more results for otimization of all daughter growth rates as well}
-
-It has to be noted that this approach is computationally much more expensive compared to the simpler
-position-comparison algorithm.
+It therefore further underlines the accuracy and predictive power of our methods.
+
+The imaging-based comparison algorithm is computationally much more expensive compared to the
+simpler position-comparison algorithm.
 There are 4 main steps which are executed within any single optimization iteration that contribute
@@
 comparison of the masks ($0.68(5)\%$) (see section \ref{section:supplement-performance}).
-The masks which are provided by the data need to be adjusted such that the identifiers of the cells
-which are directly related to the chosen colors match with the ones used within the masks provided
-by the data.
-It is clear that the generation of masks from the numerical results takes up the majority of the
-work.
+The second step alters the masks provided by the data for every individual numerical simulation.
+This is required since we need to ensure that the colors of the data-masks match the colors of the
+masks produced by our computational model.
+To this end, we need to construct a family tree of the data-bacteria and compare with the
+\textit{in-silico} family tree.
+We can see that the generation of masks from the numerical results takes up the majority of the work.
 In addition, the step of mask adjustment would not have been required within the position-comparison
@@
         cell-tracking algorithms with that.
+    \item Lack of data -> this study establishes only methods -> apply it much more broadly in
+        future
 \end{itemize}
@@
 the mechanistic methodology of our approach.
+
+\textbf{Discussion Advanced Optimizations}\\
+\begin{itemize}
+    \item 
+\end{itemize}
 
@@
 accurate (maybe just leave this in the appendix).
+
+\begin{figure}
+    \centering
+    \begin{tikzonimage}[width=0.33\textwidth]
+        {figures/crm_amir/profiles-reduced/fit-comparison.pdf}%
+        \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.33\textwidth]
+        {figures/crm_amir/profiles-reduced/drag force.pdf}%
+        \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.33\textwidth]
+        {figures/crm_amir/profiles-reduced/growth rate.pdf}%
+        \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+    \end{tikzonimage}\\
+    \begin{tikzonimage}[width=0.33\textwidth]
+        {figures/crm_amir/profiles-reduced/rod rigidity.pdf}%
+        \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.33\textwidth]
+        {figures/crm_amir/profiles-reduced/rigidity to spring tension ratio.pdf}%
+        \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.33\textwidth]
+        {figures/crm_amir/profiles-reduced/damping.pdf}%
+        \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
+    \end{tikzonimage}%
+    \caption{
+        \textbf{TODO}
+    }
+\end{figure}
+
+\subsubsection{With Division}
 

### 2025-09-29 paper/main.tex
write summary of discussion

@@
 \newacro{pbr}[PBR]{Physically-Based Rendering}
+\newacro{ib}[IB]{Individual-based}
 
@@
 \label{section:discussion}
+
+With this work we have layed out a "blueprint" to estimate the parameters of \acp{abm} using
+microscopic image time series.
+Our approach is modularized such that its structure can be applied to a variety of problems.
+We introduced a \ac{ib} computational model for rod-shaped bacteria that incorporates bendiness of
+the rod and is implemented with \texttt{cellular\_raza} allowing it to be flexible in design and
+readily applicable for computational studies.
+Furthermore, we devised novel methods to extract the spatial positions and other information of
+cell-agents from masks which have been segmented from microscopic images.
+The extracting method was validated using results from our computational model.
+We further implemented a pipeline that can produce 3D visualizations of simulation snapshots and can
+generate cell masks from data as produced by our computational model.
+These imaging techniques were also utilized to design a cost function that can compare snapshots
+of cellular states across cellular division events, allowing us to estimate parameters across a
+large time-span.
+Finally, the combination of these preceding advancements were applied to 2 case studies, which showed
+that the profile-likelihood method can be used to effectively evaluate the ability of our system to
+properly identify the involved parameters.
+In the first example, we fit a simplified model of a singular bent rod to microscopic images.
+These results highlighted that a reduction of the systems parameters is necessary in order to
+identify them.
+In our next step, the same system was fitted to a series of microscopic images in two secarios,
+excluding and incorporating division events.
+Both results showed the ability of our model to effectively estimate individual parameters of the
+cells.
+The case including division events produced parameter profiles with minima that were not fully able
+to estimate the chosen parameters.
+These results ca be traced back to a gap within the provided segmentation masks.
 

### 2025-09-29 paper/main.tex
first iteration of discussion

@@
 %###################################################################################################
-\section{Discussion (TODO)}
+\section{Discussion}
 \label{section:discussion}
@@
 
-\begin{itemize}
-    \item Model does not capture every aspect of rod-shaped bacteria. Needs to extended (see our
-        review).
-    \begin{enumerate}
-        \item polar interactions
-        \item differentiation for van Gogh bundles; surfactin-producing, matrix-producing cells.
-        \item Proper friction; this might have a huge implication on the estimation of
-            parameters~\cite{Grant2014}.
-            This paper has already implemented this~\cite{Doumic2020}.
-            We can build on top of that.
-        \item Extracellular Reactions~\cite{Li2025}
-        \item see supplement for full list \ref{table:simulation-aspects-supplement}
-    \end{enumerate}
-    \item Data Extraction: Skeletonization Algorithm ultimately not good enough for some scenarios
-        (i.e. overlapping of cells, endpoints can sometimes be "too short")
-    \begin{itemize}
-        \item Possibly use AI for new algorithm
-    \end{itemize}
-    \item Image generation; Produce high-quality near-realistic microscopic images with cell-masks
-        from scratch with mechanistic model; train cell-segmentation and more importantly
-        cell-tracking algorithms with that.
-    \item Lack of data -> this study establishes only methods -> apply it much more broadly in
-        future
-\end{itemize}
-
-\textbf{Discussion on the Individual Treatment}\\
-The individual treatment of the bacteria allows us in principle to estimate the distribution of the
+Our work shows that it is possible to estimate the parameters of \acp{abm} on a single-cell level.
+Together with the flexible nature of \texttt{cellular\_raza}, this opens up the possibility to
+design minimal \acp{abm} which can be fit to data.
+As such it formulates a blueprint for the parameter estimation of \acp{abm}.
+The developed methods are publicly available as a well-documented python package
+(\url{https://github.com/jonaspleyer/cr_mech_coli}).
+The individual treatment of the cells allows us in principle to estimate the distribution of the
 parameters directly without having to resort to any assumptions on the form of said distribution.
-This is true for any parameter, not just the growth rates, since we aim to estimate all mechanical
-parameters of our cellular model.
-
-The total parameter space $P$ is a product of all single spaces $P_i$ in which the parameters
-$p_i$ live $P=P_0\times P_1\times\dots$ but some of these parameters share the same space (let's
-assume $p_0,p_1\in \tilde{P}$).
-This leaves the open question as to weather the optimization algorithm could be modified to leverage
-this knowledge and obtain increased performance.
-For instance, it might be valuable to initially consider individual parameters as globally defined
-and start treating the individually after a certain amount of optimization steps.
-
-\textbf{Discussion Identifiability and Model Reduction/Flexibility}\\
-But it also shows that the approach which we used is able to detect these problems.
-As we continue to develop these methods, we expect to be able to reuse many of the existing
-theoretical approaches such as likelihood estimation and uncertainty quantification which have
-already been developed extensively for less complex systems such as \acp{ode}.
-\textbf{TODO citations, also: is this discussion?}
-Fortunately, \texttt{cellular\_raza} was chosen to construct our model such that we are flexible in
-changing the underlying representation if should find that such unidentifiabilities are structural
-(model-based) rather than practical (due to lack of data).
+This is true for any parameter, not just the growth rates, since our approach can be generalized to
+other individual cellular behaviours.
+
+Future work should focus on multiple aspects that target challenges with respect to computational
+efficiency and accuracy and extending the described model.
+Additional biological effects include polar interactions~\cite{Duvernoy2018},
+differentiation~\cite{vanGestel2015}, extracellular reactions~\cite{Li2025} (see also
+table~\ref{table:simulation-aspects-supplement}).
+When it comes to the designed algorithms, we have already discussed that the performance of
+generating masks from computational results is the limiting factor for the performance of the
+imaging-based comparison algorithm.
+This problem could be targeted by rewriting existing code in C++ using native VTK~\cite{vtkBook}
+instead of relying on the wrapper library pyvista~\cite{Sullivan2019}.
+Another area of improvement is the position extraction algorithm which works well for cells living
+in 2D or thin sheets but will not hold up when multiple bacteria are overlapping.
+This task is well suited to be tackled by a purpose-built machine-learning tool that can either
+reconstruct masks or accurately extrapolate positions from incomplete (possibly overlapping) masks.
+Finally, the results of our computational model can be used to generate synthetic near-realistic
+microscopic brightfield images together with their corresponding cell masks.
+This combination of data allows us to train segmentation and more importantly tracking algorithms
+which require large amounts of data.
+
+In order to speed up the optimization process, it could also be valuable to initially consider all
+parameters as global ones and once a certain threshold of tolerance is matched, losening this
+restriction and treating cells individually.
+This will allow the algorithm to initially search within a smaller space and possibly reach the
+optimium more quickly in total.
+
+Our approach furthermore highlighted that we are able to detect non-identifiabilities.
+By choosing \texttt{cellular\_raza} as a modeling framework, we retain flexibility with respect to
+model design thus effectively allowing us to reduce our model as we continue working on it.
+\textt{cellular\_raza} allows us to change the underlying representation if such unidentifiabilities
+are determined to be structural (model-based) rather than practical (due to lack of
+data).\textbf{TODO CITATION}
 This allows us to iterate and find the model formulation which suits the data best but still retain
@@
 
-\textbf{Discussion Advanced Optimizations}\\
-\begin{itemize}
-    \item 
-\end{itemize}
+As we continue to develop these methods, we expect to be able to reuse many of the existing
+workflows\textbf{TODO better wording} such as likelihood estimation and uncertainty quantification
+which have already been developed extensively for less complex systems such as \acp{ode}
+\textbf{TODO CITATION}.
+
+% \begin{itemize}
+%     \item Summary:
+%         "blueprint" to estimate parameters of \ac{abm},
+%         bending of bacteria,
+%         individual nature of cells,
+%         fit model to data,
+%         without+with division
+%
+%     \item Model does not capture every aspect of rod-shaped bacteria. Needs to extended (see our
+%         review).
+%     \begin{enumerate}
+%         \item polar interactions
+%         \item differentiation for van Gogh bundles; surfactin-producing, matrix-producing cells.
+%         \item Proper friction; this might have a huge implication on the estimation of
+%             parameters~\cite{Grant2014}.
+%             This paper has already implemented this~\cite{Doumic2020}.
+%             We can build on top of that.
+%         \item Extracellular Reactions~\cite{Li2025}
+%         \item see supplement for full list \ref{table:simulation-aspects-supplement}
+%     \end{enumerate}
+%     \item Data Extraction: Skeletonization Algorithm ultimately not good enough for some scenarios
+%         (i.e. overlapping of cells, endpoints can sometimes be "too short")
+%     \begin{itemize}
+%         \item Possibly use AI for new algorithm
+%     \end{itemize}
+%     \item Image generation; Produce high-quality near-realistic microscopic images with cell-masks
+%         from scratch with mechanistic model; train cell-segmentation and more importantly
+%         cell-tracking algorithms with that.
+%     \item Lack of data -> this study establishes only methods -> apply it much more broadly in
+%         future
+% \end{itemize}
+%
+% \textbf{Discussion Identifiability and Model Reduction/Flexibility}\\
+%
+% \textbf{Discussion Advanced Optimizations}\\
+% \begin{itemize}
+%     \item 
+% \end{itemize}
 

### 2025-09-29 paper/main.tex
add more details; fix command

@@
 optimium more quickly in total.
+Another factor that could be considered to improve performance is that individual parameters of
+cells which are spatially separated will not have an immediate effect on each other, thus presenting
+a loosely coupled or fully decoupled relationship.
+Depening on the solver, this information could be exploited.
 
@@
 model design thus effectively allowing us to reduce our model as we continue working on it.
-\textt{cellular\_raza} allows us to change the underlying representation if such unidentifiabilities
+\texttt{cellular\_raza} allows us to change the underlying representation if such unidentifiabilities
 are determined to be structural (model-based) rather than practical (due to lack of

### 2025-09-29 paper/main.tex
update todos

@@
 often not designed to be flexible but rather built for specific purposes thus limiting their
-generalizability.
-% TODO citations
+generalizability \textbf{TODO CITATION}.
 Typically researchers aim to estimate every individual parameter of their model.
-This is achieved using methods such as the profile-likelihood and model reduction techniques.
-% TODO citaitons
+This is achieved using methods such as the profile-likelihood and model reduction
+techniques \textbf{TODO CITATION}.
 When it comes to \acp{abm} however, it is often only shown that the results which have been
@@
 reality.
-But this process does not actually estimate the necessary parameters directly
-% TODO citations
+But this process does not actually estimate the necessary parameters directly \textbf{TODO CITATION}
 and reduces the confidence in the constructed models and raises concerns with respect to the
@@
 %###################################################################################################
-\section{Application 1: Elastic deformations of \textit{E.Coli} (TODO)}
+\section{Application 1: Elastic deformations of \textit{E.Coli}}
 \label{section:application-1}
@@
 were not able to be segmented.
-\textbf{TODO check this with data}
+\textbf{TODO check this with results}
 It can be argued that this leads to a non-identifiability for the first division length while the

### 2025-09-29 paper/main.tex
fix link
finish conclusion

@@
 The developed methods are publicly available as a well-documented python package
-(\url{https://github.com/jonaspleyer/cr_mech_coli}).
+(\href{https://github.com/jonaspleyer/cr_mech_coli}{cr\_mech\_coli}).
 The individual treatment of the cells allows us in principle to estimate the distribution of the
@@
 Depening on the solver, this information could be exploited.
+Ultimately, we desire to apply these methods to large amounts of data, which is done most
+effectively by fitting our model to various image time-series simultanously which is yet to be
+implemented.
 
@@
 
-\section{Conclusion (TODO)}
-
-\begin{itemize}
-    \item Highlight methodological advancements
-    \item Constructed Model which already combines many effects; extend this even further
-    \item can now use classical methods: sensitivity-analysis, likelihood estimators and uncertainty quantification
-    \item Proof of concept: Able to fit individual parameters of the agents; can be extend this?
-    \item Provide software package which combines all functionality and is readily available
-    \item Future work:
-    \begin{itemize}
-        \item Optimize multiple data sources (images) in parallel: requires some kind of
-        "normalization" to ensure that the dimensions, pixels etc. are the same such that parameters
-        are comparable; also requires identical experimental setup
-        \item Compare different species with various modifications
-        \item Modify model futher; asymmetric forces etc.
-    \end{itemize}
-\end{itemize}
+\section{Conclusion}
+
+This work has introduced a novel methodological framework which allows us to estimate the parameters
+of models describing rod-shaped bacteria on the single-cell level by using data obtained from
+microscopic images.
+The approach is agnostic with respect to details of the underlying model and instead only makes use
+of the fundamental spatial representation of our bacteria as a collection of vertices.
+In this way it highlights a modular blueprint that can be followed by researchers applied to models
+of rod-shaped bacteria which incorporate other and more complex behaviours.
+Combining these methods, our workflow links experimental data to the parameter estimation of a
+mechanistic, computational model and allows us to estimate the individual parameters of the
+cell-agents.
+
+Our framework allow researchers to employ well-known methods such as profile-likelihood
+\textbf{TODO CITATION} and sensitivity-analysis.
+The code is publicly available as the python package
+\href{https://github.com/jonaspleyer/cr_mech_coli}{cr\_mech\_coli}.
+The utilized cellular properties are contained within the \texttt{cellular\_raza}~\cite{Pleyer2025}
+package which provides flexibility in model design, thus enabling researchers to easily reuse,
+extend and combine existing building blocks.
+This allows researchers to extend the model and incorporate additional biological properties.
+
+Future work will focus on the data-generation capabilities of the software.
+We will aim to generate near-realistic microscopic brightfield images together with corresponding
+segmentation masks in order to train segmentation and tracking algorithms.
+On the biological side, a plethora of cellular properties waits to be investigated using this novel
+tecnhique.
+Another open question is how to efficiently scale up this operation in order to treat large amounts
+of cells such that distributions of parameters themselves can be estimated.
+Put together, this work establishes a promising platform which is able to directly combine
+microscopic data with single-cell agent-based modeling techniques and offers a wide range of
+opportunities for future quantitative studies in cellular biolgoy.
+
+% \begin{itemize}
+%     \item Highlight methodological advancements
+%     \item Constructed Model which already combines many effects; possibility to extend this even further
+%     \item can now use classical methods: sensitivity-analysis, likelihood estimators and uncertainty quantification
+%     \item Proof of concept: Able to fit individual parameters of the agents; can be extend this?
+%     \item Provide software package which combines all functionality and is readily available
+%     \item Future work:
+%     \begin{itemize}
+%         \item Optimize multiple data sources (images) in parallel: requires some kind of
+%         "normalization" to ensure that the dimensions, pixels etc. are the same such that parameters
+%         are comparable; also requires identical experimental setup
+%         \item Compare different species with various modifications
+%         \item Modify model futher; asymmetric forces etc.
+%     \end{itemize}
+% \end{itemize}
 

### 2025-10-17 paper/main.tex
format toc
add todo
add comments from christian
Co-authored-by: Christian Fleck <christian.fleck@fdm.uni-freiburg.de>

@@
     However, despite their popularity, it is often difficult to systematically estimate their
-    parameters and perform advanced analyses e.g. using the profile-likelihood.
+    parameters and perform advanced analyses, e.g., using the profile-likelihood.
     In this work, we study the mechanical properties of rod-shaped bacteria and provide techniques
     for estimating their parameters and performing identifiability analysis on them.
-    Our approach follows a blueprint of a modular framework which can be adapted to other research
-    questions.
+    Our approach may serve as a blueprint for a modular framework which can be adapted to other
+    research questions.
     We use time series of microscopic images as input data and incorporate advanced effects as the
@@
 \tableofcontents
+\vfill
 \pagebreak
@@
 \section{Introduction}
-
-\acp{abm} have proven to be powerfull tools in the study of complex cellular systems.
+\textbf{
+    Mein Vorgehen: ich passe Formulierungen an, wenn ich im Prinzip mit der Aussage einverstanden
+    bin, aber den Wortlaut verbesserungswürdig finde.
+    Ich schreibe Kommentare, dort wo ich eher grundsätzliche Schwierigkeiten sehe.
+    Dort wo ich nichts geändert habe und keinen Kommentar gegeben habe, kann ich zu einem späteren
+    Zeitpunkt doch Änderungen vornehmen wollen.
+}
+
+\acp{abm} have proven to be powerfull tools in the study of complex cellular
+systems~\cite{Pleyer2023}.
 They have been applied in various scenarios, such as cancer
 research~\cite{Ghaffarizadeh2018,Cooper2020}, Immuno-Oncology~\cite{Karolak2021} or the modeling
-of mechanical intercellular connections in plant cells.
+of mechanical intercellular connections in plant cells \textbf{TODO CITATION}.
 By treating cells as individual entities, these tools allow researchers to think in terms of
@@
 Despite their popularity, \acp{abm} face multiple challenges in practise.
-Constructing a fully working \ac{abm} simulation from scratch is an intricate process that requires
-a large set of skills and time.
-It is therefore common practise to reuse existing simulation tools, which brings along a number of
+The construction of a fully functional \ac{abm} simulation from the ground up constitutes an
+intricate and laborious process, necessitating a comprehensive array of skills. It is therefore
+common practise to reuse existing simulation tools, which, in turn, brings along a number of
 other problems.
@@
 often not designed to be flexible but rather built for specific purposes thus limiting their
-generalizability \textbf{TODO CITATION}.
-Typically researchers aim to estimate every individual parameter of their model.
-This is achieved using methods such as the profile-likelihood and model reduction
-techniques \textbf{TODO CITATION}.
-When it comes to \acp{abm} however, it is often only shown that the results which have been
-generated by an \ac{abm} lead to the desired phenomenon which aligns with the observed biological
-reality.
-But this process does not actually estimate the necessary parameters directly \textbf{TODO CITATION}
-and reduces the confidence in the constructed models and raises concerns with respect to the
-validity of the obtained results.
-
+generalizability \textbf{TODO CITATION}. Typically researchers estimate the parameter of their model
+using methods such as Maximum Likelihood or Maximum a Posteriori.
+
+[Gabor, Attila, and Julio R.
+Banga.
+‘Robust and Efficient Parameter Estimation in Dynamic Models of Biological Systems’.
+BMC Systems Biology 9, no.
+1 (2015): 74.
+https://doi.org/10.1186/s12918-015-0219-2.,
+Banga, Julio R, and Eva Balsa-Canto. ‘Parameter Estimation and Optimal Experimental Design’. Essays in Biochemistry 45 (2008): 195–209. https://doi.org/10.1042/BSE0450195., 
+Ashyraliyev, Maksat, Yves Fomekong-Nanfack, Jaap A Kaandorp, and Joke G Blom.
+‘Systems Biology: Parameter Estimation for Biochemical Models’.
+FEBS Journal 276, no.
+4 (2009): 886–902.
+https://doi.org/10.1111/j.1742-4658.2008.06844.x.] \\
+
+check for parameter identifiability using methods like profile-likelihood or apply model reduction techniques \\
+
+[Kreutz, Clemens, Andreas Raue, Daniel Kaschek, and Jens Timmer. ‘Profile Likelihood in Systems Biology’. The FEBS Journal, May 2013, n/a-n/a. https://doi.org/10.1111/febs.12276.
+Raue, Andreas, Johan Karlsson, Maria Pia Saccomani, Mats Jirstrand, and Jens Timmer.
+‘Comparison of Approaches for Parameter Identifiability Analysis of Biological Systems.’ Bioinformatics (Oxford, England), ahead of print, January 2014.
+https://doi.org/10.1093/bioinformatics/btu006.
+Simpson, Matthew J. ‘Data-Informed Model Reduction for Inference and Prediction from Non-Identifiable Models’. Preprint, Synthetic Biology, 21 March 2025. https://doi.org/10.1101/2025.03.20.644462.]\\
+
+In the context of \acp{abm}, however, it is often only shown that the generated results lead to the
+desired phenomenon aligning with the observed biological reality.
+But this process does not actually estimate the necessary parameters directly \textbf{TODO CITATION}.
+Consequently, it reduces the confidence in the constructed models and raises concerns with respect
+to the validity of the obtained results.
 In this work, we address these challenges using a general-purpose mechanical model for rod-shaped
-bacteria.
+bacteria as an example.
 Our approach provides various methods which can be combined to systematically estimate the
-parameters of our model.
-We deliberately distinguish between conceptually separate components within this framework.
+parameters of our model. We deliberately distinguish between conceptually separate components within
+this framework.
 This modularity highlights a "blueprint" which can be adapted and applied to other \acp{abm}.
@@
 expansion~\cite{Matyjaszkiewicz2017,Doumic2020} but they are limited in describing the
-aforementioned plastic or elastic deformations of the rod.
-Furthermore, the literature on how to estimate the parameters of such models directly from data is
-very sparse.
-
-This work adresses some of these limitations.
-We use microscopic image time series as inputs to obtain information about the spatial
-configuration of the cells.
+aforementioned plastic or elastic deformations of the rod. Furthermore, the literature on how to
+estimate the parameters of such models directly from data is very sparse.
+
+\textbf{[Was mir hier nicht gefällt, ist, dass du von Parameter schätzen bei ABMs kommend ziemlich detailliert über Mechanik redest.
+Das passt nicht.
+Man erwartet mehr Infos über den Blueprint und ist verwirrt.
+Worum geht es hier?]}
+
+This work addresses some of these limitations.
+We use microscopic image time series as inputs to obtain information about the spatial configuration
+of the cells.
 This information can be used as initial values for a numerical simulation.
@@
 They allow researchers to obtain feedback quickly and improve iteration times when constructing new
-models.
-Furthermore, we utilize them in our more advanced parameter estimation methods (see
+models. Furthermore, we utilize them in our more advanced parameter estimation methods (see
 section~\ref{section:parameter-estimation}) where they enable us to effectively compare
@@
 algorithms.
+
+\textbf{[Ich denke nicht, dass $\vec{x}_i$ hier nötig ist.
+Es bietet keinerlei Information.
+Du kannst alles so schreiben ohne das mathematische Symbol.
+Du musst den Bogen hinbekommen von dem allgemeinen Problem des Parameterschätzen bei ABMs zu dem spezifischen Problem, welches du hier verhandelst.]}
 
@@
 
+We will go into the details of the model for rod-shaped bacteria.
 The mechanical behaviour of rod-shaped bacteria is governed by a plethora of effects.
@@
 many types of rod-shaped bacteria but also explicit enough such that we can quantify its parameters.
-We are thus aiming to work with a set of assumptions that is as general as possible but still
-specific to rod-shaped bacteria.
-With this delicate balance in mind, it is clear that many cellular processes which might be specific
-to one particular celltype or simply not necessary to assume in order to construct our model, will
-be omitted alltogether.
-This concerns effects such as polar interactions~\cite{Duvernoy2018}, extracellular
-reactions~\cite{Wang2010} or differentiation processes~\cite{vanGestel2015,Lpez2010} and many other.
-When describing our system, we distinguish between cellular aspects which ocurr within a single
-cell and cell-cell interactions between numerous bacteria.
-We restrict ourselves to the aspects presented in table~\ref{table:simulation-aspects}.
-For a collection of other known effects which could be considered, we refer to the supplementary
+We are thus aiming to work with a minimal set of assumptions that is as general as possible but
+still specific to rod-shaped bacteria.
+
+Given this delicate balance, it is evident that numerous cellular processes will be excluded due to
+their specificity to particular cell types or their irrelevance to the construction of our model.
+This encompasses a range of effects, including polar interactions~\cite{Duvernoy2018}, extracellular
+reactions~\cite{Wang2010} or differentiation processes~\cite{vanGestel2015,Lpez2010}, among numerous
+others.
+When describing the bacterial system, we distinguish between intra-cellular processes and cell-cell
+interactions.
+We restrict ourselves to the features discussed below and summarized in
+Table~\ref{table:simulation-aspects}.
+For a collection of other known aspects which could be considered, we refer to the supplementary
 section~\ref{section:discussion}, and~\ref{section:supplement-more-simulation-aspects}.
@@
 \paragraph{(C) Cellular}
-It was long believed that bacteria lack cytoskeletal filaments but with the end of the last century,
-it was discovered that the MreB protein fills this role since it “[..] forms an actin-like
-cytoskeleton in bacteria [..]”~\cite{Erickson2001}.
+In bacteria the MreB protein forms an actin-like cytoskeleton~\cite{Erickson2001}.
 It can polymerize and form filaments which behave similar to actin microfilaments~\cite{Dersch2020}.
 The effect of MreB on the shape of cells during growth was shown in more detail
-by~\cite{Ursell2014}.
-They found that MreB localizes specifically at regions of negative curvature and thus introduces
+by~\cite{Ursell2014}. They found that MreB localizes specifically at regions of negative curvature
+and thus introduces
 heterogeneous cell growth which results in a bending force during the growth of cells.
@@
 This feedback mechanism will result in straight rod-shaped bacteria if not interfered from the
-external environment.
-These results were also confirmed by the studies of~\cite{Wang2010}.
+external environment. These results were also confirmed by the studies of~\cite{Wang2010}.
 In their experiments, a single bacterium was fixed at the bottom pole while an optically trapped
-bead was placed on the upper pole.
-They could now exert a force on the free part and thus study the response of the cell under
-different conditions.
-When treating cells with A22, an antibiotic which disassembles
-MreB~\cite{IWAI2002,Gitai2005,Karczmarek2007,Bean2009}, a higher displacement with identical force
-could be measured, thus indicating that MreB increases mechanical rigidity of the cell.
+bead was placed on the upper pole. They could now exert a force on the free part and thus study the
+response of the cell under different conditions. When treating cells with A22, an antibiotic which
+disassembles MreB~\cite{IWAI2002,Gitai2005,Karczmarek2007,Bean2009}, a higher displacement with
+identical force could be measured, thus indicating that MreB increases mechanical rigidity of the
+cell.
 
@@
 \textit{B.Subtilis} utilize a method where new material is inserted along the cylindrical part of
-the rod.
-The details of the division event (i.e. Z-Ring formation) are very intricate and are heavily
+the rod. The details of the division event (i.e. Z-Ring formation) are very intricate and are heavily
 studied as a subject on their own~\cite{Harry2001}.
@@
 opposed to the age of bacterium~\cite{Robert2014}.
-Furthermore, only the growth of the whole colony but also the individual growth of the bacteria is
+Furthermore, not only the growth of the whole colony but also the growth of individual bacteria is
 exponential~\cite{Amir2014,Takeuchi2005}.
+\textbf{[Bitte etwas kürzen, da du hier keinen Review schreibst und dich deswegen um Prägnanz bemühen musst.]}
 
@@
 
-All these considerations become even more intricate when not considering interactions of bacteria
-with surfaces but bacteria with each other.
+All these considerations become even more intricate when interactions between bacteria are considered.
 Gram-negative bacteria (eg. E.Coli) express outer membrane proteins (adhesins) which allow them to
@@
 between bacteria.
+
+\textbf{[Man versteht nicht so recht, warum du das alles hier erzählst und was daraus folgt. Die Verbindung zu deinem Modell fehlt.]}
 
@@
             % Bacteria can form multilayers during their growth phase \cite{Duvernoy2018}.\\
-        (4) & Individual Parameters &
-            Parameters for individual cells are not fixed values but rather taken from a
-            distribution \cite{Koutsoumanis2013}.\\
+        (4) & Process Parameters &
+            We take a potential heterogeneity of the bacteria into account by \textbf{[wir sollten über diesen Punkt nochmals reden. 
+            Das ist hier leicht missverständlich. Parameter sind auch kein intrazellularer Aspekt.]} \cite{Koutsoumanis2013}.\\
         &\textbf{(CC) Cell-Cell Interactions}\\
@@
 To model the spatial mechanics of elongated bacteria \cite{Billaudeau2017}, we represent them as a
-collection of auxiliary vertices $\{\vec{x}_i\}$ which are connected by springs in
-ascending order.
-Furthermore, we assume that the cells are flexible described by their rigidity.
-A force $\vec{F}$ interacting between cellular agents determines the radius (thickness) of the
-rods and an attractive component can model adhesion between cells.
+collection of auxiliary vertices $\{\vec{x}_i\}$ which are connected by springs in ascending order.
+Furthermore, we assume that the cells are flexible described by their rigidity. A force $\vec{F}$
+interacting between cellular agents determines the radius (thickness) of the rods and an attractive
+component can model adhesion between cells.
 
@@
 \label{subsec:spatial-representation}
-This work makes use of a few but powerfull assumptions concerning the mathematical representation of
-the cells in question.
-In order to describe the spatial dynamics of individual agents, we need to define their position,
-shape and how they interact with each other.
+This work makes use of a few but powerful assumptions concerning the mathematical representation of
+the cells. In order to describe the spatial dynamics of individual cells or agents, we need to
+define their position, shape and how they interact with each other.
+
 In order to provide as much flexibility as possible when constructing new models, we restrict
@@
 various shapes and interaction types.
+\textbf{[Dieser Satz ist leider unverständlich. Je öfter ich den Satz lese, umso weniger ist mir klar, was du sagen möchtest.]}\\
+
 Furthermore, it is very important to strike a good balance between the restrictiveness of our
@@
 approach to a broader class of models.
+\textbf{[Ebenso unklar warum du das hier an dieser Stelle sagst. Vielleicht versuchst du die Aussagen zu bündeln und geeigneter Stelle zu bringen. So ist man leider eher verwirrt.]}\\
+
 Taking into account the natural shape of rod-like bacteria, it is clear that they can not be
 represented with a singular point $\vec{x}$ in space.
+\textbf{[Hast du nicht bereits oben die Hilfsvariablen $x_i$ eingeführt?]}\\
+
 We could extend this positional definition with a vector $\vec{w}$, indicating the direction of
 extension but this would restrict ourselves to fully straight rod-shaped bacteria.
+\textbf{[Nicht spekulieren was du hättest machen können, sondern einfach sagen, was du gemacht hast.]}
+
 However, multiple experiments have shown that MReB is responsible for the growth and maintenance of
 the elongated shape of bacteria~\cite{Erickson2001}.
-By means of this protein, rod-shaped bacteria can bend elastically on short time-scales and
-deform plastically over longer periods of time.
-Our goal is to provide a model which can encapsulate these effects.
-Thus, we extend our model from a single point to multiple vertices which are part of a 1-dimensional
-polygon with two endpoints $\vec{x}_i$.
-In essence, we discretize the (possibly non-straight) rod of the agent.
+\textbf{[Wiederholung?]}
+
+By means of this protein, rod-shaped bacteria can bend elastically on short time-scales and deform
+plastically over longer periods of time. Our goal is to provide a model that can encapsulate these effects.
+\textbf{[Das beschriebene Verhalten ist visko-elastisch.
+Soweit ich sehe, hast du das nicht implementiert.
+Könnte es sein, dass das elastisches Modell mit Wachstum äquivalent zum Maxwell-Modell ist?
+Eigene Antwort: im allgemeinen nein.]}\\
+
+We discretize the (possibly non-straight) rod of a bacteria and model it by using multiple vertices
+which are part of a 1-dimensional
+polygon with two endpoints $\vec{x}_i$. \textbf{[Was ist denn die Diskretisierungslänge? Two endpoints $x_i$ macht keinen Sinn.]}\\
+
 Figure~\ref{fig:mechanics-bacterium} shows a visualization of this approach.
@@
 dynamics of these bacteria.
+\textbf{[Klingt kryptisch. "Can be used ..". Für was, wie?]}\\
 
@@
 two bacteria.
-Even concerning the shape of bacteria, this assumption still allows us to mdoel elliptical cells by
+Even concerning the shape of bacteria, this assumption still allows us to model elliptical cells by
 modifying the interaction type as well as other effects such as polarity of
@@
 of the rod and other more intricate mechanisms.
+\textbf{[Ich glaube, ich verstehe warum du das hier sagst.
+Aber das kommt nicht richtig rüber.
+Der Absatz ist vielleicht nicht nötig.
+Der Kommentar über elliptische Zellen, sollte eher unter Potential stehen?]}\\
 
@@
         A bacterial agent which is not interacting with any external objects will tend to the
-        equilibrium state of a straight rod with $0$ curvature.
+        equilibrium state of a straight rod with no curvature.
     }
@@
 \label{subsection:mechanical-model-mechanics}
-
 We use principles from discrete differential geometry~\cite{bobenko2008discrete} to describe the
-stiffness and flexibility of the bacterial rod~\cite{Amir2014_2}.
-As explained previously, MReB will counteract the bending of the rod, resulting in a stiffening
-force.
-In principle, this stress will be dependent on the position inside the cell~\cite{Chatterjee1988}
-but in order to construct a model which can be parametrized and by method of Occam's Razor, we chose
-to model the bacterium as a bending beam, thus utilizing the simplest solution which can still
-produce the desired effects.
+stiffness of the bacterial rod~\cite{Amir2014_2}.
+As explained above, MReB will counteract the bending of the rod, resulting in a stiffening force.
+In principle, this stress will be dependent on the position inside the cell~\cite{Chatterjee1988},
+but for sake of simplicity we chose to model the bacterium as a bending beam.
+\textbf{[", thus utilizing the simplest solution which can still produce the desired effects." Es ist unklar was die "desired effects" sind. ]}\\
+
 Such a beam can be represented by a collection of vertices $\vec{x}_i$ which are connected by a
 springs of lengths $l_i$.
+\textbf{[Das klingt nach einer Wiederholung von Spatial Representation. Du erklärst ziemlich umständlich. Konzentriere dich auf das wesentliche und überlege, was du eigentlich sagen möchtest.]}\\
+
 We use the notation $\vec{c}_i:=\vec{x}_i-\vec{x}_{i-1}$ with which the resulting force acting
@@
 \end{align}
-
-where $D$ is the diffusion constant, $\vec{\xi}$ is the wiener process and $\lambda$ a damping
-constant.
+\textbf{[Ich verstehe die Gleichungen nicht. Eine überdämpfte Langevin Gleichung bezeichnet man doch als Brownsche Bewegung, oder? Und sollte die Brownsche Gleichung nicht so lauten?
+\begin{align}
+   \partial_t \vec{x}_i &= \frac{D}{k_B T} \vec{F}_{i,\text{total}} + \sqrt{2D}\vec{\xi}_i
+\end{align}
+Die Einheiten stimmen auch überhaupt nicht. $[x_i]=$m, $[\lambda]=$kg/s, $[F]=$(kg m)/s$^2$
+]}\\
+
+where $D$ is the diffusion constant, $\vec{\xi}$ is a Wiener process and $\lambda$ a damping constant.
 Together they determine the stochastic motion of the individual vertices which can arise from
@@
 
-When calculating forces which act between the cells, we can use a simplified model to circumvent the
-numerically expensive integration over the complete length of the rod.
-Given a vertex $\vec{x}_i$ on one cell, we calculate the closest point $\vec{p}$ on the polygonal
-line given by the vertices $\{\vec{w}_j\}$ of the interacting cell.
+Calculating the force between cells resulting  from long ranged interactions requires performing a
+double integral over the length of the cells. 
+To reduce the computational costs we use a approximative, simplified model Given a vertex
+$\vec{x}_i$ on one cell, we determine the closest point $\vec{p}$ on the polygonal line given by the
+vertices $\{\vec{w}_j\}$ of the interacting cell. 
+\textbf{[Ich bin etwas verwirrt. Warum änderst du die Bezeichnung der Vertices?]}
+
 Furthermore we determine the value $q\in[0,1]$ such that $\vec{p} = (1-q)\vec{w}_j + q\vec{w}_{j+1}$
@@
 
-The Mie~\cite{Mie1903} potential is the generalization of the well-known
-Lennard-Jones~\cite{Jones1924} potential.
-It was also designed to describe the interactions of molecules but due to the flexible nature of its
-parameters, can be easily adapted to other situations.
-It is able to describe a wider range of shapes compared to the Morse potential but is also stiffer
-and not bound from above.
-In our simulation, we enforce an upper bound to guarantee numerical regularity.
-It has 4 parameters and is given by
+The interaction between cells can be described by the Mie potential~\cite{Mie1903}.
+It is the generalization of the well-known Lennard-Jones~\cite{Jones1924} potential and was designed
+to describe the interactions between molecules.
+The Mie potential is given by:
+
 \begin{align}
@@
 \end{align}
-where $C$ is a constant that depends on the two exponents $n,m$ that describe the overall shape of
-the potential.
-The strength is given by $\epsilon$ and the radius is encoded in $\sigma =
-(m/n)^{1/(n-m)}(R_1+R_2)$ where $R_1,R_2$ are the radii of the two interacting particles.
+
+The strength of the potential is given by $\epsilon$ and the radius \textbf{[Welcher Radius?]} is
+encoded in $\sigma = (m/n)^{1/(n-m)}(R_1+R_2)$ where $R_1,R_2$ are the radii of the two interacting
+particles.
+In our simulation, we use $V(r)$ to guarantee numerical regularity.
+\textbf{[Beschreibe, welche Rolle $\beta$ und $\zeta$ spielen.]}
 
@@
 
+\textbf{[Du hast hier von shape geredet.
+Das Potential ist doch immer radial-symmetrisch, oder?
+Vielleicht könnte es helfen, Bilder zu zeigen?
+Darin könnte man die Bedeutung der Parameter verdeutlichen.
+Wenn du schreibst, das Potential ist sehr wichtig, dann musst du viel besser begründen, warum du das Potential gewählt hast.
+Ich denke, du solltest das etwas abschwächen.]}
+
 % --------------------------------------------------------------------------------------------------
@@
 However, in order to generalize our approach we assume that individual lengths of segments could
-vary.
-Our implemented model allows us to reduce the growth rate with increasing number of neighbors
-$n\leq N$.
-This option provides a phenomenological approach for modeling slower growth in larger colonies but
-is be toggled off by default thus removing the additional term in brackets of
-equation~\ref{eq:growth-rod-lengths}.
+vary. Our implemented model allows us to reduce the growth rate with increasing number of neighbours
+$n\leq N$. This option provides a phenomenological approach for modelling slower growth in larger colonies but
+is be toggled off by default thus removing the additional term in brackets of equation~\ref{eq:growth-rod-lengths}.
+
+\textbf{[Ich empfinde deinen Text als uneinheitlich.
+Du solltest hier nicht deine Softwareimplementation beschreiben, sondern das was du tatsächlich in den Simulationen benutzt hast.
+Also Möglichkeiten, die du in deinem Code angelegt hast, aber für das Paper nicht verwendet hast, brauchst du nicht zu erwähnen.
+Das führt zu Verwirrung.]}
 
@@
 growth~\cite{Robert2014,Takeuchi2005}.
-This process will increase the length of the cell indefenitely unless we impose a condition for the
-division event.
-We define a fixed length threshold for the total length of the cell at which it divides.
+\textbf{[Das zu erwähnen ist eine Trivialität für ein Forschungspapier. Du kannst begründen warum du exp. Wachstum verwendest und wie du den Teilungszeitpunkt bestimmst.]}
+
+This process will increase the length of the cell indefinitely unless we impose a condition for the
+division event. We define a fixed length threshold for the total length of the cell at which it divides.
 To construct a new cell, we cannot simply copy the existing one twice, but we also need to adjust
@@
 The following actions need to be taken for the old and new agent.
-We first calculate the new spring lengths of the two resulting rods.
+We first calculate the new spring lengths of the two resulting rods, according to the equation:
+
 \begin{equation}
@@
 \end{equation}
-Afterwards, we start at one endpoint $\vec{x}_0$ and travel along the polygon defined by the
-bacterial rod.
-We stop when we have reached a length that matches the length of the new spring length $\tilde{l}$.
+
+Afterwards, we start at one endpoint $\vec{x}_0$ and move along the polygon defined by the
+bacterial rod. We stop when we have reached a length that matches the length of the new spring length $\tilde{l}$.
 From here, we continue and iterate this procedure until we have found all new vertices $\vec{w}_i$.
@@
 right after the division event.
+\textbf{[Was ist $r$? Was ist mit den Parametern und mit anderen intro-zellularen Zuständen?]}
 
@@
 
-We implemented the described model in \texttt{cellular\_raza}~\cite{Pleyer2025}.
-Doing so has many benefits.
-Not only, was a more simpler model already available which we were able to adapt to our needs.
-But by design \texttt{cellular\_raza} aims to provide researchers with a great amount of flexibility
-and therefore does not assume any particular cellular representation, thus enabling us to make
-fundamental changes.
-This allows us to to experiment with different model implementations and still retain performance.
+We implemented the described model in our recently developed \ac{abm} framework
+\texttt{cellular\_raza}~\cite{Pleyer2025}.
+By design \texttt{cellular\_raza} does not assume any particular cellular representation, thus makes
+it flexible and easy to adapt the simulation to particular needs. This allows us to experiment with
+different model implementations and still retain performance.
+\textbf{[Der Zusammenhang zwischen der Flexibilität und Performance ist unklar.]}
+
 Furthermore, it is an \ac{abm} which is able to natively track cells through the course of a
 numerical simulation and also keeps track of mother/daughter relationships.
+\textbf{[Das scheint hier trivial zu sein.]}
+
 On the programming side, we are using the \texttt{pyo3} and \texttt{maturin} crates which allow us
 to easily generate Python bindings and enable us to interoperate between the two ecosystems.
+\textbf{[Das kennt hier keiner und gehört eher in den Anhang.]}
+
 By using \texttt{cellular\_raza}, we can make use of multiple threads for running a single numerical
@@
 Figure~\ref{fig:performance}).
+\textbf{[Ist alles eher für den Anhang. Vielleicht solltest du mehr darüber reden, warum ABMs für die Fragestellungen gut geeignet sind.]}
 
@@
 section~\ref{subsec:spatial-representation}.
-Therefore, they can be applied to a whole class of mechanical models which could make different
-choices with respect to the cellular details.
+Consequently, these models can be applied to a broader class of mechanical models that may exhibit
+varied cellular behaviours.
 
@@
 \label{subsection:visualization}
-
-To visualize the spatial configuration of our results we use pyvista, which provides high-level
+To visualize the spatial configuration of the results we employ pyvista, which provides high-level
 wrappers around the popular VTK library~\cite{vtkBook,Sullivan2019}.
-The first stage of our visualization is to generate 3D meshes for each bacterium, which we consider
-an intermediate representation.
+The initial phase of the visualization process entails the generation of three-dimensional meshes
+for each bacterium as an intermediate representation.
 We insert spheres of a given radius $r$ at each vertex $\vec{x}_i$ and connect adjacent vertices
@@
 Finally, we combine the meshes of all of these objects to obtain a single mesh of one bacterium.
+
 A render of this result can be seen in figure~\ref{fig:progression-image-generation}.
@@
 overlaps of bacteria which are placed at higher or lower values along the z-direction.
-
-To generate masks, we assign unique colors to every mesh of the respective cellular agent, ensuring
-no two agents will use identical colors.
+To generate masks, we assign unique colours to every mesh of the respective cellular agent, ensuring
+no two agents will use identical colours.
 We then use a parallel projection and remove any existing light sources in order to obtain a solid
@@
 The properties of the meshes such as roughness and reflectivity have been adjusted and we use
-physically-based rendering of VTK.
-The result of this render is augmented by introducing white noise on the pixel values at each point.
+physically-based rendering of VTK. The result of this render is augmented by introducing white noise
+on the pixel values at each point.
+\textbf{[Warum?]}
+
 This approach is very rudimentary and we hope to extend these methods in the future.
+\textbf{[So sollte man das nicht formulieren. Hört sich etwa so an wie: jetzt hatte ich keine Lust es richtig zu machen, aber vielleicht mache ich das in der Zukunft.]}
 In order to properly generate data, we need to deal with microscopic imaging techniques and possible
-defects that can occurr.
-It is crucial to get this part of the image-generation process right in order to obtain
-near-realistic images that can be further reused.
+defects that can occur. It is crucial to get this part of the image-generation process right in
+order to obtain near-realistic images that can be further reused.
+\textbf{[Kann wenig mit diesen Kommentaren anfangen. Verstehe nicht warum du das hier sagst.]}
 
@@
 They will serve as initial, final and intermediate values with which we can compare our predictions.
+\textbf{[Beziehen sich initial und final, usw. auf räumliches? Ist missverständlich, bitte umformulieren.]}
 The algorithms which we devised requires the input of cell-masks which assign a unique color for
 every cell in the image.
+
 Such masks can be generated by cell-segmentation tools or obtained by annottating microscopic images
@@
         Assert that there are only 2 endpoints.
+        \textbf{[Heissen die Endpoints immer $q$? Kontrolliere bitte, ob deine Notation einheitlich ist.]}
     \item Sort points such that they form a polygon.
@@
 \end{enumerate}
+
 In the first step, the image will be split into a collection of subimages that only contain
@@
 It has to be noted that there are multiple variants of skeletonization and in general, these classes
-of algorithms can produce results which have intersection points and multiple endpoints.
-This described behaviour is undesirable in our case.
-We have found that for most practical examples, the algorithm devised by~\cite{Lee1994} has worked
-well.\\
+of algorithms can produce results which have intersection points and multiple endpoints, which is
+undesirable in our case.
+We have found that for most practical examples, the algorithm devised by~\cite{Lee1994} works well.\\
+
 The skeletonization algorithm supplies us with a collection of unordered lattice points $\vec{p}_i$
@@
 advantage of a functionality which will be introduced later in
-section~\ref{subsection:visualization}.
+section~\ref{subsection:visualization} \textbf{[Hier stimmt die Referenz nicht.]}.
 Our methodology makes use of the fact that we have already set up our model and solving scheme and
@@
 
+\textbf{
+    Die Verständlichkeit und die Formulierungen hier bedürfen noch der Verbesserung.
+    Damit ich aber schneller durchkomme, verschieben wir das auf später.
+}
+
 \begin{figure}
@@
 In order to effectively estimate the parameters of our model, we need to set up a loss function
-which can be minimized.
-In doing so, we hope to identify all parameters, given relevant experimental data.
+which can be minimized. In doing so, we hope to identify all parameters, given relevant experimental data.
 We utilize the preceding algorithms to extract data from cell-masks, simulate from an initial state
@@
 The cell-masks are easily obtainable by either using existing segmentation
-tools~\cite{Cutler2022,Stringer2020,Hardo2022} or manually annottating microscopic image series.
+tools~\cite{Cutler2022,Stringer2020,Hardo2022} or manually annotating microscopic image series.
+
+\textbf{Du solltest hier erwähnen, dass du die Parameter per Bakterium fittest und nicht globale Parameter.}
 
@@
 For all cells $j=1..M$, we have a collection of $i=0..N$ vertices $\vec{x}_{i,j}$.
-Thus, the total amount of information is an array of shape $(M,N,2)$
-Assuming that we have $N$ cells, each of which is represented by $M$ vertcies, the total amount of
-information is an array of shape $(N,M,2)$, containing all positions of all agents.
+\textbf{[Warum jetzt zweifach indiziert?]} Thus, the total amount of information is an array of shape $(M,N,2)$ Assuming that we have $N$ cells, each of which is represented by $M$ vertices, the total amount of information is an array of shape $(N,M,2)$, containing all positions of all agents.
 It is thus only a matter of comparing two of these arrays with each other and many methods and
-approaches which already exist~\cite{Wang2020} can be leveraged.
-For now, we simply calculate the distance of each vertex pair, which is equivalent to the
-well-known loss function
+approaches which already exist~\cite{Wang2020} can be leveraged. For now, we simply calculate the distance of each vertex pair, which is equivalent to the
+well-known loss function:
 
@@
         = \sqrt{\sum\limits_{i,j}||\vec{x}_{i,j}-\vec{y}_{i,j}||^2}
-        = \sqrt{\sum\limits_{i,j,k}|x_{i,j,k}-y_{i,j,k}|^2}.
+        = \sqrt{\sum\limits_{i,j,k}|x_{i,j,k}-y_{i,j,k}|^2}
 \end{equation}
@@
 not present in the experimental results (or vice-versa).
+\textbf{[Es bleibt unklar, was du hier verhandeln möchtest.]}
+
 We thus need additional methods which can be used for comparisons across division events.
@@
 circumstances~\cite{Dice1945}.
+
+\textbf{[Wir müssen darüber reden, mir ist hier vieles unklar.]}
 
@@
 should be preferred.
+\textbf{[Wir müssen darüber reden, mir ist hier vieles unklar. Es könnte nützlich sein, vor der Application Sektion, eine Zusammenfassung zu setzen.]}
 
@@
 by the experiments of Amir et. al~\cite{Amir2014,Amir2014_2}.
-These images show a singular \textit{E.Coli} which is fixed at the bottom and is subject to a
-horizontal fluid flow that exerts a force onto the rod.
+\textbf{[Es wäre gut zu sagen, was für ein Ziel du verfolgst. Warum machst du das?]}
+They studied how a single \textit{E.Coli}, which is fixed at the bottom, reacts to a horizontal
+fluid flow that exerts a force onto the bacteria.
 This setup is displayed in figure~\ref{fig:amir-bending-simulation} (A).
-In order to model this setup, we also fixate the lower-most vertex of our rod at the bottom of the
-simulation domain.
+To model this configuration, the lowermost vertex of the rod is fixed at the base of the simulation
+domain.
+
 Furthermore, we introduce a force $\vec{F}_i = C |\vec{x}_i - \vec{x}_{i+1}|\sin(\alpha_i)\vec{e}_x$
-that acts similarlyl to Stoke's Law \textbf{TODO Citation} onto the normal component of each
+that acts similarly to Stoke's Law \textbf{TODO Citation} onto the normal component of each
 rod-segment where $\alpha_i$ is the angle of the rod edge between $\vec{x}_i$ and $\vec{x}_{i+1}$
 and the x-axis.
-We extract the position of the rod from the provided image and use it as intial value.
+We extract the position of the rod from the provided image and use it as initial value.
 Our simulation is split into two phases, one with an external force that initially bends the rod,
@@
 We compare the simulated positions with the extracted ones and calculate the euclidean distance
-between them.
-This value serves as the cost function that allows us to estimate the parameters of this model.
+between them, which we use as the cost function to estimate the parameters of this model.
+
 Since this model is void of any interactions or cell division, we are left with 5 parameters:
@@
 (blue) position of the rod using the optimized parameters.
-While the dragging force, growth rate and damping can be identified properly, the same
-can not be said for the rod rigidity and spring tension which both exhibit non-identifiabilities.
-However this behaviour is expected since, the shape of the rod is almost exclusively straight,
-thus adding no further information except that a lower threshold of rigidity and spring tension
-should be present.
+While the parameters for the dragging force, growth rate and damping can be identified, the same
+can not be said for the rod rigidity and spring tension, which both are not identifiable.
+However this behaviour is expected since, the shape of the rod is almost exclusively straight, thus
+adding no further information except that a lower threshold of rigidity and spring tension should be
+present.
+\textbf{[Ist mir nicht ganz klar]}
+
 Note that each point represented in the subfigures~\ref{fig:amir-bending-simulation} (C-G) was
@@
 values of drag forces.
+\textbf{[Ist mir nicht ganz klar.]}
+
 Furthermore the used images only contain snapshots of a fully-bent or fully-relaxed rod without any
-intermediate steps that cannot fully capture its dynamic.
+intermediate steps.
 This effectively limits us to 2 datapoints and is likely the key factor that leads to the two local
@@
 distributed between the bent and relaxed position of our system.
-In order to better understand this system, we reduce it by inserting a new parameter that determines
-the values of rod rigidity and spring tension via $\gamma=\gamma_0/s$ and $\eta=\eta_0 s$.
+
+To remove this non-identifiability we reduce the number of parameters by setting $\gamma=\gamma_0/s$
+and $\eta=\eta_0 s$ for the rod rigidity and spring tension, resp.
 We then optimize for this parameter and pick $\gamma_0=\SI{15}{\pixel}$ with
 $\eta_0=\SI{150}{\pixel\per\second\squared}$ which corresponds to $60\%$ of the previously considered
-interval.
-With this change, all parameters are identifiable.
+interval. \textbf{[Unklar, wie die Pixel hier auftauchen.]}
 The corresponding plots can be seen in \ref{subsec:supplement-parameter-profiles-amir}.
@@
 \label{section:application-2}
-
-To show the usefulness of our previously developed methods (see
-section~\ref{section:parameter-estimation}), we will use them to estimate the parameters individual
-growing \textit{E.Coli} using a sequence of microscopic images.
+To show the usefulness of our previously developed methods
+(see section~\ref{section:parameter-estimation}), we will use them to estimate the parameters
+individual growing \textit{E.Coli} using a sequence of microscopic images.
 The data we use~\cite{https://doi.org/10.3203/iwf/k-129} (see
 section~\ref{section:data-code-availability}) has been automatically segmented with the
-omnipose~\cite{Cutler2022} algorithm which provides us with the necessary cell masks.
+omnipose~\cite{Cutler2022} algorithm by mean of which the necessary cell masks are obtained.
 
@@
 \subsection{Individual Treatment}
+
 We have already suspected (see table~\ref{table:simulation-aspects}) that in order to properly
@@
 optimization routine.
-Given $N_c$ cells and $N_p$ parameters per cell and $N_u$ individual parameters (where
-$N_u\leq N_p$), the total amount of parameters to be estimated is
-
+Given $N_c$ cells and $N_p$ parameters per cell and $N_u$ individual parameters (where $N_u\leq N_p$), the total amount of parameters to be estimated is:
 \begin{equation}
@@
 \end{equation}
-
 It is clear, that we should avoid to needlessly assign every parameter individually since the volume
@@
 number of parameters.
+\textbf{[Das ist leider komplett unverständlich.
+Ich glaube, ich weiss schon, was du sagen möchtest, aber das muss umformuliert werden.
+Stichworte sind hier: Cell Variability, Heterogeneity, latent variables.
+Parameter per cell and individual parameters versteht auch niemand den Unterschied.]}
+
 However, we will now present data that indicate that an individual treatment is in fact necessary.
@@
 in subfigure (C).
-There, we can see much more clearly the difference between considering invidiually assigned
+There, we can see much more clearly the difference between considering individually assigned
 parameters and global ones.
@@
 system, we need to treat the bacteria individually.
+\textbf{[Man ahnt was du meinst, doch leider ahnt man es nur. Vermeide Aussagen wie "it is clear".]}
 
@@
 \subsection{Without Division}
-For now, we will focus on a time-interval which does not contain any division events and at which
+In section we will focus on a time-interval which does not contain any division events and at which
 the initial lag-phase of our bacteria has already passed.
-By doing so, we can utilize the simpler and more performant position comparison scheme for our loss
-function.
+\textbf{[Sprachlich wieder fragwürdig. "A time interval at which the initial lag-phase has passed" ist kein irgendwie korrekter englischer Satz.]}
+
+By doing so, we can utilize the simpler and more performant position comparison scheme for our loss function.
+\textbf{[Hier weiss man nicht worüber du sprichst.]}
+
 Furthermore, we need not to estimate the lag-time parameters which improves performance of our
 optimization algorithm.
+\textbf{[Verweise auf die entsprechende Gleichung.]}
+\textbf{[Ich werde nun nicht mehr das Sprachliche kommentieren oder verbessern. Zuerst muss die gesamte Konzeption verbessert werden.]}
+
 % For this dataset, we assumed that all parameters which specify the interaction potential are
@@
 microscopic image time series.
+\textbf{
+    Es wird leider überhaupt nicht deutlich, was die Blaupause sein könnte.
+    Wenn man das MS liesst, gewinnt man den Eindruck, du machst dies und das, aber ein einheitliches Konzept fehlt.
+    Wir müssen zuerst eine Konzeption für das Paper entwickeln, bevor wir über weitere Details reden können.
+}
+
 Our approach is modularized such that its structure can be applied to a variety of problems.

### 2025-10-20 paper/main.tex
include citations from christian
Co-authored-by: Christian Fleck <christian.fleck@fdm.uni-freiburg.de>

@@
 often not designed to be flexible but rather built for specific purposes thus limiting their
-generalizability \textbf{TODO CITATION}. Typically researchers estimate the parameter of their model
-using methods such as Maximum Likelihood or Maximum a Posteriori.
-
-[Gabor, Attila, and Julio R.
-Banga.
-‘Robust and Efficient Parameter Estimation in Dynamic Models of Biological Systems’.
-BMC Systems Biology 9, no.
-1 (2015): 74.
-https://doi.org/10.1186/s12918-015-0219-2.,
-Banga, Julio R, and Eva Balsa-Canto. ‘Parameter Estimation and Optimal Experimental Design’. Essays in Biochemistry 45 (2008): 195–209. https://doi.org/10.1042/BSE0450195., 
-Ashyraliyev, Maksat, Yves Fomekong-Nanfack, Jaap A Kaandorp, and Joke G Blom.
-‘Systems Biology: Parameter Estimation for Biochemical Models’.
-FEBS Journal 276, no.
-4 (2009): 886–902.
-https://doi.org/10.1111/j.1742-4658.2008.06844.x.] \\
-
-check for parameter identifiability using methods like profile-likelihood or apply model reduction techniques \\
-
-[Kreutz, Clemens, Andreas Raue, Daniel Kaschek, and Jens Timmer. ‘Profile Likelihood in Systems Biology’. The FEBS Journal, May 2013, n/a-n/a. https://doi.org/10.1111/febs.12276.
-Raue, Andreas, Johan Karlsson, Maria Pia Saccomani, Mats Jirstrand, and Jens Timmer.
-‘Comparison of Approaches for Parameter Identifiability Analysis of Biological Systems.’ Bioinformatics (Oxford, England), ahead of print, January 2014.
-https://doi.org/10.1093/bioinformatics/btu006.
-Simpson, Matthew J. ‘Data-Informed Model Reduction for Inference and Prediction from Non-Identifiable Models’. Preprint, Synthetic Biology, 21 March 2025. https://doi.org/10.1101/2025.03.20.644462.]\\
+generalizability \textbf{TODO CITATION}.
+Typically researchers estimate the parameter of their model using methods such as Maximum Likelihood
+or Maximum a Posteriori~\cite{Gbor2015,Banga2008,Ashyraliyev2009}, check for parameter
+identifiability using methods like profile-likelihood or apply model reduction
+techniques~\cite{Kreutz2013,Raue2014,Simpson2025}.
 

### 2025-10-20 paper/main.tex
formatting of comments and cleanup

@@
 
-\textbf{[Was mir hier nicht gefällt, ist, dass du von Parameter schätzen bei ABMs kommend ziemlich detailliert über Mechanik redest.
+\textbf{
+    Was mir hier nicht gefällt, ist, dass du von Parameter schätzen bei ABMs kommend ziemlich
+    detailliert über Mechanik redest.
     Das passt nicht.
     Man erwartet mehr Infos über den Blueprint und ist verwirrt.
-Worum geht es hier?]}
+    Worum geht es hier?
+}
 
@@
 \textit{B.Subtilis} and \textit{E.Coli}.
-
-% \input{introduction_old}
 
@@
 
-\textbf{[Ich denke nicht, dass $\vec{x}_i$ hier nötig ist.
-Es bietet keinerlei Information.
-Du kannst alles so schreiben ohne das mathematische Symbol.
-Du musst den Bogen hinbekommen von dem allgemeinen Problem des Parameterschätzen bei ABMs zu dem spezifischen Problem, welches du hier verhandelst.]}
-
-% \begin{itemize}
-%     \item This publication contains 5 things:
-%     \begin{enumerate}
-%         \item Methods of how to obtain data from microscopic images and perform parameter
-%             estimations; $\rightarrow$ also possibly across multiple generations
-%         \item Showcase the usefulness of the methods and the model by applying it to parameter
-%             estimation via microscopic images and multilayer formation
-%         \item Initial (incomplete) model of rod-shaped bacteria
-%         \item Generate 3D Images (masks)
-%         \item A python package which can do all of that and can be reused and extended
-%     \end{enumerate}
-%     \item Overall structure
-%     \begin{enumerate}
-%         \item users bring model + data
-%         \item this paper provides methods to do the parameter estimation + visualization
-%     \end{enumerate}
-%     \item For complete model, more work has to be done
-%     \begin{itemize}
-%         \item More aspects (see Review; put in discussion)
-%         \item Note that we only care about \textit{E.Coli} and \textit{B.Subtilits}! This is
-%             important.
-%     \end{itemize}
-%     \begin{itemize}
-%         \item State of the Art
-%     \end{itemize}
-% \end{itemize}
+\textbf{
+    Du musst den Bogen hinbekommen von dem allgemeinen Problem des Parameterschätzen bei ABMs zu dem
+    spezifischen Problem, welches du hier verhandelst.
+}
 
@@
 exponential~\cite{Amir2014,Takeuchi2005}.
-\textbf{[Bitte etwas kürzen, da du hier keinen Review schreibst und dich deswegen um Prägnanz bemühen musst.]}
+\textbf{
+    Bitte etwas kürzen, da du hier keinen Review schreibst und dich deswegen um Prägnanz bemühen
+    musst.
+}
 

### 2025-10-20 paper/main.tex
remove mathematical variable from this text

@@
 Figure~\ref{fig:flowchart-project-structure} shows a high-level overview of all components.
-We assume that bacterial rods can be represented by a colleciton of vertices $\vec{x}_i$ (see
+We assume that bacterial rods can be represented by a colleciton of vertices (see
 section~\ref{subsec:spatial-representation}).
@@
 given vertices which can then either be rendered or projected to a 2D plane.
-The data extraction component determines the vertices $\vec{x}_i$ from given cell masks which
+The data extraction component determines the vertices from given cell masks which
 correspond to microscopic images.
@@
 positions.
-The dynamics of the vertices $\vec{x}_i$ are determined by a computational model which is specific
+The dynamics of the vertices are determined by a computational model which is specific
 to the respective use-case and should be defined by the researcher.
 It merely needs to utilize the aforementioned spatial representation of bacterial rods as a
-collection of vertices $\vec{x}_i$.
+collection of vertices.
 This work will provide one example of how such a model can be defined.
@@
         Every component implicitly builds on the spatial representation of rod-shaped bacteria as
-        vertices $\vec{x}_i$.
+        vertices.
     }

### 2025-10-20 paper/main.tex
formatting

@@
 
-We will go into the details of the model for rod-shaped bacteria.
 The mechanical behaviour of rod-shaped bacteria is governed by a plethora of effects.
-In this work, we aim to establish a model which is abstract enough such that it can be reused for
-many types of rod-shaped bacteria but also explicit enough such that we can quantify its parameters.
-We are thus aiming to work with a minimal set of assumptions that is as general as possible but
-still specific to rod-shaped bacteria.
-
-Given this delicate balance, it is evident that numerous cellular processes will be excluded due to
-their specificity to particular cell types or their irrelevance to the construction of our model.
-This encompasses a range of effects, including polar interactions~\cite{Duvernoy2018}, extracellular
-reactions~\cite{Wang2010} or differentiation processes~\cite{vanGestel2015,Lpez2010}, among numerous
-others.
+In this work, we aim to establish a model which is general enough such that it can be reused as the
+basis for many variants of rod-shaped bacteria but also explicit enough such that we can quantify
+its parameters.
+We thus utilize only a minimal set of assumptions.
+
 When describing the bacterial system, we distinguish between intra-cellular processes and cell-cell
 interactions.
-We restrict ourselves to the features discussed below and summarized in
+We restrict ourselves to the features discussed below and summarize them in
 Table~\ref{table:simulation-aspects}.
-For a collection of other known aspects which could be considered, we refer to the supplementary
-section~\ref{section:discussion}, and~\ref{section:supplement-more-simulation-aspects}.
+% For a collection of other known aspects which could be considered, we refer to the supplementary
+% section~\ref{section:discussion}, and~\ref{section:supplement-more-simulation-aspects}.
 
@@
 The effect of MreB on the shape of cells during growth was shown in more detail
-by~\cite{Ursell2014}. They found that MreB localizes specifically at regions of negative curvature
-and thus introduces
+by~\cite{Ursell2014}.
+They found that MreB localizes specifically at regions of negative curvature and thus introduces
 heterogeneous cell growth which results in a bending force during the growth of cells.
@@
 This feedback mechanism will result in straight rod-shaped bacteria if not interfered from the
-external environment. These results were also confirmed by the studies of~\cite{Wang2010}.
+external environment.
+These results were also confirmed by the studies of~\cite{Wang2010}.
 In their experiments, a single bacterium was fixed at the bottom pole while an optically trapped
-bead was placed on the upper pole. They could now exert a force on the free part and thus study the
-response of the cell under different conditions. When treating cells with A22, an antibiotic which
-disassembles MreB~\cite{IWAI2002,Gitai2005,Karczmarek2007,Bean2009}, a higher displacement with
-identical force could be measured, thus indicating that MreB increases mechanical rigidity of the
-cell.
+bead was placed on the upper pole.
+They could now exert a force on the free part and thus study the response of the cell under
+different conditions.
+When treating cells with A22, an antibiotic which disassembles
+MreB~\cite{IWAI2002,Gitai2005,Karczmarek2007,Bean2009}, a higher displacement with identical force
+could be measured, thus indicating that MreB increases mechanical rigidity of the cell.
 

### 2025-10-22 paper/main.tex
restructure part according to christians notes

@@
 
-\paragraph{(C) Cellular}
-In bacteria the MreB protein forms an actin-like cytoskeleton~\cite{Erickson2001}.
-It can polymerize and form filaments which behave similar to actin microfilaments~\cite{Dersch2020}.
-The effect of MreB on the shape of cells during growth was shown in more detail
-by~\cite{Ursell2014}.
-They found that MreB localizes specifically at regions of negative curvature and thus introduces
-heterogeneous cell growth which results in a bending force during the growth of cells.
-It also means that the cells are able to sense the curvature of the cell wall through cytoskeleteal
-MreB filaments during growth.
-\cite{Bratton2018} showed that the trans-membrane protein RodZ can modulate MreB density and
-facilitates local MreB assembly.
-This feedback mechanism will result in straight rod-shaped bacteria if not interfered from the
-external environment.
-These results were also confirmed by the studies of~\cite{Wang2010}.
-In their experiments, a single bacterium was fixed at the bottom pole while an optically trapped
-bead was placed on the upper pole.
-They could now exert a force on the free part and thus study the response of the cell under
-different conditions.
-When treating cells with A22, an antibiotic which disassembles
-MreB~\cite{IWAI2002,Gitai2005,Karczmarek2007,Bean2009}, a higher displacement with identical force
-could be measured, thus indicating that MreB increases mechanical rigidity of the cell.
-
 While some rod-shaped bacteria grow by inserting new material at the ends, \textit{E.Coli} and
@@
 creating two new daughter cells.
-We know that the best predictor for when the division event occurs, is the length of the rod as
+It was shown that the best predictor for when the division event occurs, is the length of the rod as
 opposed to the age of bacterium~\cite{Robert2014}.
@@
 exponential~\cite{Amir2014,Takeuchi2005}.
-\textbf{
-    Bitte etwas kürzen, da du hier keinen Review schreibst und dich deswegen um Prägnanz bemühen
-    musst.
-}
-
-\paragraph{(CC) Cell-Cell Interactions \& (DC) Domain-Cell Interactions}
-A colony of bacteria is known to form biofilms (Dunne 2002). This strategy is is beneficial for the
-survival of the collective as it allows it to harness growth factors and migrate when necessary and
-also plays an important role in bacterial infections in the biomedical sector~\cite{Ong1999}.
-The formation of biofilms is only possible due to the attractive nature of bacterial interactions
-towards each other and surfaces (Berne et al. 2018). To analyze bacterial adhesion to surfaces,
-\cite{vanLoosdrecht1989} investigated the interactions of various bacteria with negatively charged
-polystyrene.
-They found that the interaction is characterized by a low Gibbs energy of $(2 - 3)kT$ per cell
-which can be described by a secondary minimum of the DLVO-theory~\cite{Verwey1947,Derjaguin1993}
-which is relevant for distances $\geq\SI{1}{\nano\meter}$.
-At these distances, the DLVO potential acts as a generic adhesive potential.
-Particles which stay in the secondary minimum are held together but even weak interactions are
-enough to redisperse them which means that this state is reversible.
-When approaching closer distances, more intricate mechanisms come into play in which the
-heterogeneity of the materials in question needs to be considered.
-\cite{Hori2010} described bacterial adhesion as a “two-phase process including an initial,
-instantaneous and reversible physical phase (phase one) and a time-dependent and irreversible
-molecular and cellular phase (phase two)”.
-
-All these considerations become even more intricate when interactions between bacteria are considered.
-Gram-negative bacteria (eg. E.Coli) express outer membrane proteins (adhesins) which allow them to
-attach to various hosts (Beachey 1981, Vaca et al. 2019).
-In addition, the flagella can play an important role in the direct adhesion between
-cells~\cite{Haiko2013}.
-Thus it is not correct to blindly apply the results for adhesion to surfaces to the interactions
-between bacteria.
-
-\textbf{[Man versteht nicht so recht, warum du das alles hier erzählst und was daraus folgt. Die Verbindung zu deinem Modell fehlt.]}
+The growth rates of individual bacteria and other process parameters can be heterogeneous which
+should be accounted for as well \cite{Koutsoumanis2013}.
+
+Many bacteria organize into biofilms~\cite{Dunne2002,Ong1999} which play a key role in the survival
+of the collective.
+Their formation is only possible due to the attractive nature of bacterial interactions between
+each other~\cite{Berne2018}.
+This interaction is characterized by a long-range attraction and short-ranged
+adherance~\cite{Verwey1947,Trejo2013}.
+We therefore need account for this phenomenon as well.
 
@@
             % Bacteria can form multilayers during their growth phase \cite{Duvernoy2018}.\\
-        (4) & Process Parameters &
-            We take a potential heterogeneity of the bacteria into account by \textbf{[wir sollten über diesen Punkt nochmals reden. 
-            Das ist hier leicht missverständlich. Parameter sind auch kein intrazellularer Aspekt.]} \cite{Koutsoumanis2013}.\\
         &\textbf{(CC) Cell-Cell Interactions}\\
         \midrule
-        (5) & Adhesion &
+        (4) & Adhesion &
             Bacteria attract at longer distances and adhere to each other at shorter
@@
     \end{tabularx}
-    \label{table:simulation-aspects}
     \caption{
@@
     }
+    \label{table:simulation-aspects}
 \end{table}
-
-To model the spatial mechanics of elongated bacteria \cite{Billaudeau2017}, we represent them as a
-collection of auxiliary vertices $\{\vec{x}_i\}$ which are connected by springs in ascending order.
-Furthermore, we assume that the cells are flexible described by their rigidity. A force $\vec{F}$
-interacting between cellular agents determines the radius (thickness) of the rods and an attractive
-component can model adhesion between cells.
 

### 2025-10-22 paper/main.tex
add notes for christian to read
restructure mechanics and interactions section

@@
 
+@Christian
+\begin{itemize}
+    \item shortened everything
+    \item leave out parts which are not directly used by the model
+    \item move biological context to respective subsections to link more directly to model
+    \item remove duplicates and useless text
+    \item move heterogeneity considerations from table in main text
+\end{itemize}
+
 The mechanical behaviour of rod-shaped bacteria is governed by a plethora of effects.
@@
 We thus utilize only a minimal set of assumptions.
-
 When describing the bacterial system, we distinguish between intra-cellular processes and cell-cell
@@
 \label{subsec:spatial-representation}
-This work makes use of a few but powerful assumptions concerning the mathematical representation of
-the cells. In order to describe the spatial dynamics of individual cells or agents, we need to
-define their position, shape and how they interact with each other.
-
-In order to provide as much flexibility as possible when constructing new models, we restrict
-ourselves to assumptions concerning the position such that researchers can freely experiment with
-various shapes and interaction types.
-\textbf{[Dieser Satz ist leider unverständlich. Je öfter ich den Satz lese, umso weniger ist mir klar, was du sagen möchtest.]}\\
-
-Furthermore, it is very important to strike a good balance between the restrictiveness of our
-assumptions and generality in order to provide meaningful results and still be able to apply our
-approach to a broader class of models.
-\textbf{[Ebenso unklar warum du das hier an dieser Stelle sagst. Vielleicht versuchst du die Aussagen zu bündeln und geeigneter Stelle zu bringen. So ist man leider eher verwirrt.]}\\
-
-Taking into account the natural shape of rod-like bacteria, it is clear that they can not be
-represented with a singular point $\vec{x}$ in space.
-\textbf{[Hast du nicht bereits oben die Hilfsvariablen $x_i$ eingeführt?]}\\
-
-We could extend this positional definition with a vector $\vec{w}$, indicating the direction of
-extension but this would restrict ourselves to fully straight rod-shaped bacteria.
-\textbf{[Nicht spekulieren was du hättest machen können, sondern einfach sagen, was du gemacht hast.]}
-
-However, multiple experiments have shown that MReB is responsible for the growth and maintenance of
-the elongated shape of bacteria~\cite{Erickson2001}.
-\textbf{[Wiederholung?]}
-
-By means of this protein, rod-shaped bacteria can bend elastically on short time-scales and deform
-plastically over longer periods of time. Our goal is to provide a model that can encapsulate these effects.
-\textbf{[Das beschriebene Verhalten ist visko-elastisch.
-Soweit ich sehe, hast du das nicht implementiert.
-Könnte es sein, dass das elastisches Modell mit Wachstum äquivalent zum Maxwell-Modell ist?
-Eigene Antwort: im allgemeinen nein.]}\\
-
-We discretize the (possibly non-straight) rod of a bacteria and model it by using multiple vertices
-which are part of a 1-dimensional
-polygon with two endpoints $\vec{x}_i$. \textbf{[Was ist denn die Diskretisierungslänge? Two endpoints $x_i$ macht keinen Sinn.]}\\
-
-Figure~\ref{fig:mechanics-bacterium} shows a visualization of this approach.
-The angles $\alpha_i$ between vertices $\vec{x}_{i-1},\vec{x}_i$ can be used when modeling the
-dynamics of these bacteria.
-\textbf{[Klingt kryptisch. "Can be used ..". Für was, wie?]}\\
-
-With these fundamental thoughts layed out, we retain the flexibility to represent many models such
-as varying shapes since we do not have to make any assumptions on the type of interactions between
-two bacteria.
-Even concerning the shape of bacteria, this assumption still allows us to model elliptical cells by
-modifying the interaction type as well as other effects such as polarity of
-interactions~\cite{Duvernoy2018}.
-Furthermore, we also did not have to assume any particular mechanical dynamics of the agents, only
-their spatial representation, thus leaving in particular room for plastical or elastic deformations
-of the rod and other more intricate mechanisms.
-\textbf{[Ich glaube, ich verstehe warum du das hier sagst.
-Aber das kommt nicht richtig rüber.
-Der Absatz ist vielleicht nicht nötig.
-Der Kommentar über elliptische Zellen, sollte eher unter Potential stehen?]}\\
+This work makes use of a one powerful assumption concerning the mathematical representation of the
+cells.
+To model the spatial mechanics of elongated bacteria \cite{Billaudeau2017}, we represent them as a
+collection of auxiliary vertices $\{\vec{x}_i\}$.
+We treat this collection of vertices as a discretization of the rod where the individual segments
+between the vertices carry no explicit physical or biological meaning.
+This assumption alone is enough to be able to build visualization and data-extraction pipelines (see
+\ref{fig:flowchart-project-structure}).
+Our computational model builds on top of this assumption and utilizes this representation to
+define the dynamics of the vertices $\{\vec{x}_i\}$.
+For this study, we will choose a particular type of model incorporating a chosen set effects.
+But researchers are in general free to adapt the model for the specific problem at hand.
+By building on our discretization scheme, future modeling approaches will be able to directly reuse
+our established theoretical methods and computational functionalities.
+
+\textbf{TODO mention this in sections below}
+\begin{itemize}
+    \item plastic/elastic deformations
+    \item non-uniform thickness of the rod
+    \item polar interactions \cite{Duvernoy2018}
+    \item other dynamics
+\end{itemize}
 
@@
         connected by springs.
-        The angles $\alpha_i$ have been exaggerated such that they can be identified more easily.
+        The angles $\alpha_i$ will be used to model the rigidity of the rod.
         A bacterial agent which is not interacting with any external objects will tend to the
@@
 \label{subsection:mechanical-model-mechanics}
+In rod-shaped bacteria, the MreB protein plays a key role for the maintenance of the cellular shape.
+It forms an actin-like cytoskeleton~\cite{Erickson2001,Dersch2020} and localizes specifically at
+regions of negative curvature thus introducing heterogeneous cell growth~\cite{Ursell2014} which
+results in a bending force during growth.
+This feedback mechanism will result in straight rod-shaped bacteria if not interfered from the
+external environment~\cite{Wang2010}.
+
 We use principles from discrete differential geometry~\cite{bobenko2008discrete} to describe the
@@
 As explained above, MReB will counteract the bending of the rod, resulting in a stiffening force.
-In principle, this stress will be dependent on the position inside the cell~\cite{Chatterjee1988},
-but for sake of simplicity we chose to model the bacterium as a bending beam.
-\textbf{[", thus utilizing the simplest solution which can still produce the desired effects." Es ist unklar was die "desired effects" sind. ]}\\
-
-Such a beam can be represented by a collection of vertices $\vec{x}_i$ which are connected by a
-springs of lengths $l_i$.
-\textbf{[Das klingt nach einer Wiederholung von Spatial Representation. Du erklärst ziemlich umständlich. Konzentriere dich auf das wesentliche und überlege, was du eigentlich sagen möchtest.]}\\
-
-We use the notation $\vec{c}_i:=\vec{x}_i-\vec{x}_{i-1}$ with which the resulting force acting
+We model the bacterium as a bending beam.
+% In principle, this stress will be dependent on the position inside the cell~\cite{Chatterjee1988},
+% but for sake of simplicity we chose to model the bacterium as a bending beam.
+% \textbf{[", thus utilizing the simplest solution which can still produce the desired effects." Es ist unklar was die "desired effects" sind. ]}\\
+Using our discretization $\vec{x}_i$, we connect the vertices by springs of length $l_i$.
+% which can be represented by a collection of vertices $\vec{x}_i$ which are connected by a
+% springs of lengths $l_i$.
+% \textbf{[Das klingt nach einer Wiederholung von Spatial Representation. Du erklärst ziemlich umständlich. Konzentriere dich auf das wesentliche und überlege, was du eigentlich sagen möchtest.]}\\
+We use the notation $\vec{w}i:=\vec{x}_i-\vec{x}_{i-1}$ with which the resulting force acting
 between two vertices if given by
 
-\begin{align}
+\begin{equation}
     \vec{F}_{i,\text{springs}} =
-        &-\gamma\left(1 - \frac{l}{\left|\vec{c}_i\right|}\right)
-        \vec{c}_i\\
-        &+ \gamma\left(1 - \frac{l}{\left|\vec{c}_{i+1}\right|}\right)
-        \vec{c}_{i+1}.
-\end{align}
+        -\gamma\left(1 - \frac{l}{\left|\vec{w}i\right|}\right) \vec{w}i
+        + \gamma\left(1 - \frac{l}{\left|\vec{w}{i+1}\right|}\right) \vec{w}{i+1}.
+\end{equation}
 
@@
 % We assume that we can model the mechanical properties of the bacterium as an elastic rod.
-Given the angle $\alpha_i = \sphericalangle(\vec{c}_{i-1},\vec{c}_i)$ between adjacent edges,
+Given the angle $\alpha_i = \sphericalangle(\vec{w}{i-1},\vec{w}i)$ between adjacent edges,
 the bending force is proportional to the curvature $\kappa_i$ at each vertex
@@
 Figure~\ref{fig:mechanics-bacterium} shows an elongated rod represented by $6$ vertices.
-The forces acting on vertices $\vec{x}_i,\vec{x}_{i-1},\vec{x}_{i+1}$ are given by
-
-\begin{align}
-    \vec{F}_{i,\text{curvature}} &= \eta\kappa_i
-        \frac{\vec{c}_i - \vec{c}_{i+1}}{|\vec{c}_i-\vec{c}_{i+1}|}\\
-    \vec{F}_{i-1,\text{curvature}} &= -\frac{1}{2}\vec{F}_{i,\text{curvature}}\\
-    \vec{F}_{i+1,\text{curvature}} &= -\frac{1}{2}\vec{F}_{i,\text{curvature}}
-\end{align}
+The force acting on vertex $\vec{x}_i$ is given by
+
+\begin{equation}
+    \vec{F}_{i,\text{curvature}} = \eta\kappa_i
+        \frac{\vec{w}i - \vec{w}{i+1}}{|\vec{w}i-\vec{w}{i+1}|}
+\end{equation}
 
 where $\eta_i$ is the rigidity at vertex $\vec{x}_i$ (see Figure~\ref{fig:mechanics-bacterium}).
+The adjoining vertices $\vec{x}_{i-1},\vec{x}_{i+1}$ are subject to half of the same force in the
+opposite direction $\vec{F}_{i\pm 1,\text{curvature}} = -0.5\vec{F}_{i,\text{curvature}}$.
 We can see that the curvature force does not move the overall center of the rod in space.
-The total force is the sum of external and interal forces.
+The total force is the sum of interal and external forces.
 
@@
 
-and are integrated via damped brownian dynamics
-
-\begin{align}
-    \partial_t^2 \vec{x}_i &= \partial_t\vec{x}_i + \sqrt{2D}\vec{\xi}_i\\
-    \partial_t\vec{x}_i &= \vec{F}_{i,\text{total}} - \lambda \partial_t\vec{x}_i
-\end{align}
-\textbf{[Ich verstehe die Gleichungen nicht. Eine überdämpfte Langevin Gleichung bezeichnet man doch als Brownsche Bewegung, oder? Und sollte die Brownsche Gleichung nicht so lauten?
-\begin{align}
-   \partial_t \vec{x}_i &= \frac{D}{k_B T} \vec{F}_{i,\text{total}} + \sqrt{2D}\vec{\xi}_i
-\end{align}
-Die Einheiten stimmen auch überhaupt nicht. $[x_i]=$m, $[\lambda]=$kg/s, $[F]=$(kg m)/s$^2$
-]}\\
-
-where $D$ is the diffusion constant, $\vec{\xi}$ is a Wiener process and $\lambda$ a damping constant.
-Together they determine the stochastic motion of the individual vertices which can arise from
-recurring collisions with small particles of the surrounding medium or other minor fluctuations.
+and are integrated via damped newtonian dynamics
+
+\begin{equation}
+    \partial_t^2 \vec{x}_i = \vec{F}_{i,\text{total}} - \lambda \partial_t\vec{x}_i
+\end{equation}
+
+where $\lambda$ a damping constant and the mass was set to $1$ to simplify the model.
 
@@
 
-Calculating the force between cells resulting  from long ranged interactions requires performing a
-double integral over the length of the cells. 
-To reduce the computational costs we use a approximative, simplified model Given a vertex
-$\vec{x}_i$ on one cell, we determine the closest point $\vec{p}$ on the polygonal line given by the
-vertices $\{\vec{w}_j\}$ of the interacting cell. 
-\textbf{[Ich bin etwas verwirrt. Warum änderst du die Bezeichnung der Vertices?]}
-
-Furthermore we determine the value $q\in[0,1]$ such that $\vec{p} = (1-q)\vec{w}_j + q\vec{w}_{j+1}$
+Calculating the force between cells requires performing a double integral over the length of the
+cells.
+To reduce the computational costs we use an approximative, simplified model.
+Given a vertex $\vec{x}_i$ on one cell, we determine the closest point $\vec{p}$ on the polygonal
+line given by the vertices $\{\vec{y}_j\}$ of the interacting cell.
+Furthermore we determine the value $q\in[0,1]$ such that $\vec{p} = (1-q)\vec{y}_j + q\vec{y}_{j+1}$
 for some specific $j$.
-The force is then calculated between the points $\vec{x}_i$ and $\vec{p}_i$ and acts on
-the vertex $\vec{w}_i,\vec{w}_{i+1}$ with relative strength $(1-q)$ and $q$.
+The force is then calculated between the points $\vec{x}_i$ and $\vec{p}$ and acts on
+the vertex $\vec{y}_j,\vec{y}_{j+1}$ with relative strength $(1-q)$ and $q$.
 
@@
 
-The interaction between cells can be described by the Mie potential~\cite{Mie1903}.
+We utilize the Mie potential~\cite{Mie1903} to describe the attraction between two cells.
 It is the generalization of the well-known Lennard-Jones~\cite{Jones1924} potential and was designed
@@
         \left(\frac{\sigma}{r}\right)^m\right]\\
-    C &= \frac{n}{n-m}\left(\frac{n}{m}\right)^{\frac{n}{n-m}}\\
-    V(r) &= \min(U(r), \beta)\theta(r-\zeta)
+    C &= \frac{n}{n-m}\left(\frac{n}{m}\right)^{\frac{n}{n-m}}
 \end{align}
 
-The strength of the potential is given by $\epsilon$ and the radius \textbf{[Welcher Radius?]} is
-encoded in $\sigma = (m/n)^{1/(n-m)}(R_1+R_2)$ where $R_1,R_2$ are the radii of the two interacting
-particles.
-In our simulation, we use $V(r)$ to guarantee numerical regularity.
-\textbf{[Beschreibe, welche Rolle $\beta$ und $\zeta$ spielen.]}
-
-The specific choice of interaction potential plays a crucial role for the dynamics of the agents.
-Since we cannot expect to simply know the correct shape and parameters of this potential, it is
-desirable to experiment with various interaction potentials in order to quantify their
-ability to produce accurate results.
-While this study, only relies on the Mie Potential, our python package \texttt{cr\_mech\_coli}
-also implements the popular Morse potential~\cite{Morse1929} to model physical interactions.
-
-\textbf{[Du hast hier von shape geredet.
-Das Potential ist doch immer radial-symmetrisch, oder?
-Vielleicht könnte es helfen, Bilder zu zeigen?
-Darin könnte man die Bedeutung der Parameter verdeutlichen.
-Wenn du schreibst, das Potential ist sehr wichtig, dann musst du viel besser begründen, warum du das Potential gewählt hast.
-Ich denke, du solltest das etwas abschwächen.]}
+The strength of the potential is given by $\epsilon$ and the thickness of our rods
+is encoded in $\sigma = (m/n)^{1/(n-m)}(R_1+R_2)$ where $R_1,R_2$ are the thicknesses of the two
+interacting bacteria.
+We utilize the shorthand notation $r=|\vec{x}_i - \vec{p}|$ for the distance between the two points
+in question.
+We numerically regularize the potential by imposing an upper bound $\beta$ on the norm of the
+resulting force $\vec{F} = \vec{\nabla} V$.
+Furthermore, we impose a cutoff parameter $\zeta$ which sets sets the force to zero when the
+distance $r$ between the interacting points exceeds this threshold.
+The resulting force is given by
+
+\begin{equation}
+    \vec{F}_{i,\text{external}} = \vec{F} \theta(r-\zeta) \min\left(1, \beta / |\vec{F}|\right)
+\end{equation}
+
+where $\theta$ is the heaviside step-function and the $\min$ term ensures the upper bound $\beta$.
+
+The physical interaction between the bacteria plays a crucial role for the overall dynamics of our
+simulation.
+The Mie potential is capable of describing a wide range of interaction types from soft to rigid, by
+adjusting the parameters of the potential accordingly.
+In contrast, a popular alternative is the Morse potential~\cite{Morse1929}, which is less
+customizable due to the lower amount of parameters and more restrictive mathematical formulation.
+Therefore, the Mie potential is a good choice for a simple, interpretable interaction potential with
+low amount of parameters that can be estimated by a data-driven approach.
 
@@
 bacterial rod. We stop when we have reached a length that matches the length of the new spring length $\tilde{l}$.
-From here, we continue and iterate this procedure until we have found all new vertices $\vec{w}_i$.
+From here, we continue and iterate this procedure until we have found all new vertices $\vec{y}_i$.
 Equation~\ref{eq:growth-rod-new-lengths} ensures that that the tips of the two new rods are touching

### 2025-10-23 paper/main.tex
fix indices

@@
 % \textbf{[Das klingt nach einer Wiederholung von Spatial Representation. Du erklärst ziemlich umständlich. Konzentriere dich auf das wesentliche und überlege, was du eigentlich sagen möchtest.]}\\
-We use the notation $\vec{w}i:=\vec{x}_i-\vec{x}_{i-1}$ with which the resulting force acting
+We use the notation $\vec{w}_i:=\vec{x}_i-\vec{x}_{i-1}$ with which the resulting force acting
 between two vertices if given by
@@
     \vec{F}_{i,\text{springs}} =
-        -\gamma\left(1 - \frac{l}{\left|\vec{w}i\right|}\right) \vec{w}i
-        + \gamma\left(1 - \frac{l}{\left|\vec{w}{i+1}\right|}\right) \vec{w}{i+1}.
+        -\gamma\left(1 - \frac{l}{\left|\vec{w}_i\right|}\right) \vec{w}_i
+        + \gamma\left(1 - \frac{l}{\left|\vec{w}_{i+1}\right|}\right) \vec{w}_{i+1}.
 \end{equation}
@@
 % We assume that we can model the mechanical properties of the bacterium as an elastic rod.
-Given the angle $\alpha_i = \sphericalangle(\vec{w}{i-1},\vec{w}i)$ between adjacent edges,
+Given the angle $\alpha_i = \sphericalangle(\vec{w}_{i-1},\vec{w}_i)$ between adjacent edges,
 the bending force is proportional to the curvature $\kappa_i$ at each vertex
@@
     \vec{F}_{i,\text{curvature}} = \eta\kappa_i
-        \frac{\vec{w}i - \vec{w}{i+1}}{|\vec{w}i-\vec{w}{i+1}|}
+        \frac{\vec{w}_i - \vec{w}_{i+1}}{|\vec{w}_i-\vec{w}_{i+1}|}
 \end{equation}

### 2025-10-23 paper/main.tex
adapt model text; clear some of christian thoughts; add table with parameters and variables

@@
 % \textbf{[", thus utilizing the simplest solution which can still produce the desired effects." Es ist unklar was die "desired effects" sind. ]}\\
-Using our discretization $\vec{x}_i$, we connect the vertices by springs of length $l_i$.
+Using our discretization $\vec{x}_i$, we connect the vertices by springs of length $l$.
 % which can be represented by a collection of vertices $\vec{x}_i$ which are connected by a
-% springs of lengths $l_i$.
+% springs of lengths $l$.
 % \textbf{[Das klingt nach einer Wiederholung von Spatial Representation. Du erklärst ziemlich umständlich. Konzentriere dich auf das wesentliche und überlege, was du eigentlich sagen möchtest.]}\\
@@
 The strength of the potential is given by $\epsilon$ and the thickness of our rods
-is encoded in $\sigma = (m/n)^{1/(n-m)}(R_1+R_2)$ where $R_1,R_2$ are the thicknesses of the two
-interacting bacteria.
+is encoded in $\sigma = (m/n)^{1/(n-m)}(R_1+R_2)$ where $R_1,R_2$ are the thicknesses (radii) of the
+two interacting rod-like bacteria.
 We utilize the shorthand notation $r=|\vec{x}_i - \vec{p}|$ for the distance between the two points
@@
 
-The physical interaction between the bacteria plays a crucial role for the overall dynamics of our
-simulation.
-The Mie potential is capable of describing a wide range of interaction types from soft to rigid, by
-adjusting the parameters of the potential accordingly.
+The physical interaction between the bacteria plays a crucial role for the overall dynamics of the
+system.
+By varying its parameters, the Mie potential is capable of describing a wide range of interaction
+types from softer ones that permit some level of overlap to rigid ones with steep ascends for
+distance valuels that correspond to overlaps.
 In contrast, a popular alternative is the Morse potential~\cite{Morse1929}, which is less
@@
 Therefore, the Mie potential is a good choice for a simple, interpretable interaction potential with
-low amount of parameters that can be estimated by a data-driven approach.
+low amount of parameters that can be estimated by our data-driven approach.
 
@@
 
-To simulate proliferation, we introduce a growth term for the spring lengths $l_i$.
-In most of our simulations, we choose to use identical lengths for all segments (i.e. $l_i=l$).
-However, in order to generalize our approach we assume that individual lengths of segments could
-vary. Our implemented model allows us to reduce the growth rate with increasing number of neighbours
-$n\leq N$. This option provides a phenomenological approach for modelling slower growth in larger colonies but
-is be toggled off by default thus removing the additional term in brackets of equation~\ref{eq:growth-rod-lengths}.
-
-\textbf{[Ich empfinde deinen Text als uneinheitlich.
-Du solltest hier nicht deine Softwareimplementation beschreiben, sondern das was du tatsächlich in den Simulationen benutzt hast.
-Also Möglichkeiten, die du in deinem Code angelegt hast, aber für das Paper nicht verwendet hast, brauchst du nicht zu erwähnen.
-Das führt zu Verwirrung.]}
+To simulate cellular growth, we introduce an exponential growth term for the spring lengths $l$
+\cite{Takeuchi2005}.
 
 \begin{equation}
-    \partial_t l_i = \mu\left(1-\frac{\text{min}(n,N)}{N}\right)l_i
-    \label{eq:growth-rod-lengths}
+    \partial_t l = \mu l
 \end{equation}
 
-Note that the growth rate is proportional to the rod length $l$ which leads to exponential
-growth~\cite{Robert2014,Takeuchi2005}.
-\textbf{[Das zu erwähnen ist eine Trivialität für ein Forschungspapier. Du kannst begründen warum du exp. Wachstum verwendest und wie du den Teilungszeitpunkt bestimmst.]}
-
 This process will increase the length of the cell indefinitely unless we impose a condition for the
-division event. We define a fixed length threshold for the total length of the cell at which it divides.
+division event.
+The cell divides when the spring length $l$ has reached a fixed length threshold~\cite{Robert2014}
+$l_\text{thresh}$.
 To construct a new cell, we cannot simply copy the existing one twice, but we also need to adjust
@@
 The following actions need to be taken for the old and new agent.
-We first calculate the new spring lengths of the two resulting rods, according to the equation:
+We first calculate the new spring lengths of the two resulting rods via
 
 \begin{equation}
-    \tilde{l} = l_i\left(\frac{1}{2} - \frac{r}{\sum\limits_i l_i}\right)
-    \label{eq:growth-rod-new-lengths}
+    \tilde{l} = l\left(\frac{1}{2} - \frac{R}{n_v l}\right)
 \end{equation}
 
-Afterwards, we start at one endpoint $\vec{x}_0$ and move along the polygon defined by the
-bacterial rod. We stop when we have reached a length that matches the length of the new spring length $\tilde{l}$.
-From here, we continue and iterate this procedure until we have found all new vertices $\vec{y}_i$.
-Equation~\ref{eq:growth-rod-new-lengths} ensures that that the tips of the two new rods are touching
-right after the division event.
-\textbf{[Was ist $r$? Was ist mit den Parametern und mit anderen intro-zellularen Zuständen?]}
+where $n_v$ is the number of vertices and $R$ the previously introduced thickness of the rod.
+This relation ensures that that the tips of the two new rods are touching right after the division
+event.
+We subdivide the polygon given by the vertices of the mother cell in $n$ segments of length
+$\tilde{l}$, one segment of length $2R$ and again $n$ segments of length $\tilde{l}$.
+The first and latter groups determine the new vertices of the two daughter cells.
+This approach assumes cells where the total length of the polygon is greater than its diameter
+$n_vl>2R$.
+We set the growth rates of the daughter cells explicitly during the division event.
+This approach allows us to treat the growth rates of the daughter cells as full individual
+parameters within our model.
+In order to reuse this model, we also provide the option to assign growth rates by sampling from a
+normal distribution with user-defined parameters.
+Every other intracellular parameter is copied identically to the daughter cells.
+Table~\ref{table:list-of-parameters} shows a list of all parameters used within our model.
+
+\begin{table}
+    \centering
+    \def\arraystretch{1.3}
+    \begin{tabular}{l c l c}
+        \textbf{Parameters} && \textbf{Variables}\\
+        \toprule
+        \multicolumn{4}{l}{\textbf{Mechanics}}\\
+        \midrule
+        Spring tension & $\gamma$ & Position (vertices) & $\vec{x}_i$\\
+        Rigidity & $\eta$ & Velocity & $\partial_t\vec{x}_i$\\
+        Damping & $\lambda$ & Spring length & $l$ \\
+        \multicolumn{4}{l}{\textbf{Interaction}}\\
+        \midrule
+        Thickness & $R$\\
+        Strength & $\epsilon$\\
+        Exponent 1 & $n$\\
+        Exponent 2 & $m$\\
+        Bound & $\beta$\\
+        Cutoff & $\zeta$\\
+        \multicolumn{4}{l}{\textbf{Growth \& Division}}\\
+        \midrule
+        Growth rate & $\mu$\\
+        Spring length threshold & $l_\text{thresh}$\\
+        \bottomrule
+    \end{tabular}
+    \caption{
+        List of all parameters and variables.
+    }
+    \label{table:list-of-parameters}
+\end{table}
 
@@
 By design \texttt{cellular\_raza} does not assume any particular cellular representation, thus makes
-it flexible and easy to adapt the simulation to particular needs. This allows us to experiment with
-different model implementations and still retain performance.
+it flexible and easy to adapt the simulation to particular needs.
+Its design allows us to experiment with different model implementations while retaining fast
+execution times.
 \textbf{[Der Zusammenhang zwischen der Flexibilität und Performance ist unklar.]}
+\textit{Oft muss man das eine für das andere opfern. Hier nicht/kaum.}
 
@@
 \textbf{[Das scheint hier trivial zu sein.]}
-
-On the programming side, we are using the \texttt{pyo3} and \texttt{maturin} crates which allow us
-to easily generate Python bindings and enable us to interoperate between the two ecosystems.
-\textbf{[Das kennt hier keiner und gehört eher in den Anhang.]}
-
-By using \texttt{cellular\_raza}, we can make use of multiple threads for running a single numerical
-simulation, using the built-in parallelization methods (see Supplementary Material,
-Figure~\ref{fig:performance}).
-\textbf{[Ist alles eher für den Anhang. Vielleicht solltest du mehr darüber reden, warum ABMs für die Fragestellungen gut geeignet sind.]}
+\textit{Machen aber viele (zB. PhysiCell) einfach nicht. (Mother-Daughter relation)}
 
@@
 and $\eta=\eta_0 s$ for the rod rigidity and spring tension, resp.
-We then optimize for this parameter and pick $\gamma_0=\SI{15}{\pixel}$ with
-$\eta_0=\SI{150}{\pixel\per\second\squared}$ which corresponds to $60\%$ of the previously considered
-interval. \textbf{[Unklar, wie die Pixel hier auftauchen.]}
+We then optimize for this parameter and pick $\gamma_0=\SI{15}{\micro\metre}$ with
+$\eta_0=\SI{150}{\micro\metre\per\second\squared}$ which corresponds to $60\%$ of the previously considered
+interval.
 The corresponding plots can be seen in \ref{subsec:supplement-parameter-profiles-amir}.
@@
 %###################################################################################################
-\subsection{Performance (TODO)}
+\subsection{Implementation \& Performance (TODO)}
 \label{section:supplement-performance}
+
+On the programming side, we are using the \texttt{pyo3} and \texttt{maturin} crates which allow us
+to easily generate Python bindings and enable us to interoperate between the two ecosystems.
+
+By using \texttt{cellular\_raza}, we can make use of multiple threads for running a single numerical
+simulation, using the built-in parallelization methods (see Supplementary Material,
+Figure~\ref{fig:performance}).
 

### 2025-10-23 paper/main.tex
fix typo
remove some useless text; fix some formuations and comments from christian

@@
 section~\ref{subsec:spatial-representation}.
-Consequently, these models can be applied to a broader class of mechanical models that may exhibit
+Consequently, these methods can be applied to a broader class of mechanical models that may exhibit
 varied cellular behaviours.
@@
 for each bacterium as an intermediate representation.
-We insert spheres of a given radius $r$ at each vertex $\vec{x}_i$ and connect adjacent vertices
-$\vec{x}_i,\vec{x}_{i+1}$ with a cylinder, also of radius $r$ which is positioned such that its core
-is identical to the connecting line segment between the two vertices.
+We insert spheres matching the thickness of the rod $R$ at each vertex $\vec{x}_i$ and connect
+adjacent vertices $\vec{x}_i,\vec{x}_{i+1}$ with a cylinder, also of radius $r$ which is positioned
+such that its core is identical to the connecting line segment between the two vertices.
 Finally, we combine the meshes of all of these objects to obtain a single mesh of one bacterium.
@@
 A render of this result can be seen in figure~\ref{fig:progression-image-generation}.
-Since we use a 3D rendering engine as an intermediate representation, the resulting image can have
+Since we use a 3D rendering engine for our visualization, the resulting image can have
 overlaps of bacteria which are placed at higher or lower values along the z-direction.
 To generate masks, we assign unique colours to every mesh of the respective cellular agent, ensuring
-no two agents will use identical colours.
+no two agents will use identical colours. \textbf{TODO semantic segmentation Citation}
 We then use a parallel projection and remove any existing light sources in order to obtain a solid
@@
 The resulting image is shown in subfigure (B).
-Subfigure (C) shows an example of a simplified way to generate microscopic images.
-The properties of the meshes such as roughness and reflectivity have been adjusted and we use
-physically-based rendering of VTK. The result of this render is augmented by introducing white noise
-on the pixel values at each point.
-\textbf{[Warum?]}
-
-This approach is very rudimentary and we hope to extend these methods in the future.
-\textbf{[So sollte man das nicht formulieren. Hört sich etwa so an wie: jetzt hatte ich keine Lust es richtig zu machen, aber vielleicht mache ich das in der Zukunft.]}
-In order to properly generate data, we need to deal with microscopic imaging techniques and possible
-defects that can occur. It is crucial to get this part of the image-generation process right in
-order to obtain near-realistic images that can be further reused.
-\textbf{[Kann wenig mit diesen Kommentaren anfangen. Verstehe nicht warum du das hier sagst.]}
-
-% \begin{itemize}
-%     \item use pyvista (VTK) for Visualization
-%     \item always do 3D render. This way, we can have overalp between agents and produce more
-%         realistic masks.
-%     \item Assign unique color values for idividual agents.
-%         We have a 1:1 mapping between agent identifier and color! (for each simulation different
-%         though)
-%     \item use parallel projection; this should be most close to what microscopes measure due to the
-%         relatively large distance between object and lens.
-%     \item use some noise to smear out the rods; this is already similar to microscopic images; use
-%         more like this and deal with real-world defects to make it more realistic (future, not now)
-%     \item Figure~\ref{fig:progression-image-generation} shows steps in generating image.
-%     \item Reflectivity, placement of light etc. can already introduce some heterogeneity in the
-%         visual of the rod.
-%     \item Plan to implement noise on the mesh itself; make it have rough surface (like real rods as
-%         seen under electron microscopy)
-% \end{itemize}
-%
-% \paragraph{Algorithm}
-% \begin{enumerate}
-%     \item Get position of cell
-%     \item Insert spheres at each vertex
-%     \item Insert cylinder along edges
-%     \item Combine individual objects into one big mesh
-%     \item Use \ac{pbr}
-% \end{enumerate}
 
@@
     \centering
-    % \begin{tikzonimage}[width=0.9\textwidth]{visualization.pdf}
-    %     \node at (0.025, 0.975)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{A}};
-    % \end{tikzonimage}
-    \begin{tikzonimage}[width=0.3\textwidth]
+    \begin{tikzonimage}[width=0.4\textwidth]
         {docs/source/_static/09395645494836445480/raw_pv/000000400.png}
@@
     \end{tikzonimage}
-    \begin{tikzonimage}[width=0.3\textwidth]
+    \hspace{0.01\textwidth}
+    \begin{tikzonimage}[width=0.4\textwidth]
         {docs/source/_static/09395645494836445480/masks/000000400.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.3\textwidth]
-        {docs/source/_static/09395645494836445480/images/000000400.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
     \end{tikzonimage}
@@
         Afterwards, the image is rendered by projecting along the z-axis.
-        (C) We use physically-based rendering and assign properties such as roughness and
-        reflectivity to generate an initial render.
-        Afterwards, we apply white noise to the pixel values of image.
-        In the future, we hope to extend this approach in order to construct near-realistic
-        microscopic images from synthetic data.
     }
@@
 \subsection{Data Extraction}
-\label{section:data-extraction}
 
@@
 (Figure~\ref{fig:position-extraction-algorithm} (A)).
-They will serve as initial, final and intermediate values with which we can compare our predictions.
-\textbf{[Beziehen sich initial und final, usw. auf räumliches? Ist missverständlich, bitte umformulieren.]}
-The algorithms which we devised requires the input of cell-masks which assign a unique color for
+The extracted vertices for multiple time points will be used by our parameter estimation methods as
+input data.
+The algorithms which we devised requires the input of cell-masks which utilize a unique color for
 every cell in the image.
@@
         Assert that there are only 2 endpoints.
-        \textbf{[Heissen die Endpoints immer $q$? Kontrolliere bitte, ob deine Notation einheitlich ist.]}
     \item Sort points such that they form a polygon.
@@
 \paragraph{Benchmarking the Extraction Algorithm}
-\label{paragraph-extraction-algorithm}
-
-In order to determine if our extraction algorithm provides reliable results, we have to take
-advantage of a functionality which will be introduced later in
-section~\ref{subsection:visualization} \textbf{[Hier stimmt die Referenz nicht.]}.
+
+In order to determine if our extraction algorithm provides reliable results, we set out to compare
+extracted data with known real data.
 Our methodology makes use of the fact that we have already set up our model and solving scheme and
@@
 \section{Discussion}
-\label{section:discussion}
 

### 2025-10-24 paper/main.tex
explain more clearly what we are doing in the introductory part

@@
 \label{section:parameter-estimation}
-In order to effectively estimate the parameters of our model, we need to set up a loss function
-which can be minimized. In doing so, we hope to identify all parameters, given relevant experimental data.
+In order to effectively estimate the parameters of our model, we need to set up a cost function
+which can be minimized.
+In doing so, we hope to identify all parameters, given relevant experimental data.
+Parameters may be defined as global ones, which are identical across all cells or individual per
+cell.
 We utilize the preceding algorithms to extract data from cell-masks, simulate from an initial state
@@
 \textbf{Du solltest hier erwähnen, dass du die Parameter per Bakterium fittest und nicht globale Parameter.}
+\textit{Tue ich aber nicht grundsätzlich. Es gibt globale und individuelle parameter; also ein "mixed approach". Ich habe aber einen Hinweis eingefügt.}
 

### 2025-10-24 paper/main.tex
rename section to make purpose more clear
reformulate cost function sections

@@
 % --------------------------------------------------------------------------------------------------
-\subsection{Data Extraction}
+\subsection{Position Extraction}
 
@@
 
-\paragraph{No Cell Division}
+\paragraph{Direct Comparison}
 In the scenario where no cells divide during the observed interval, we can use our extraction
@@
 calculated from the computational results.
-For all cells $j=1..M$, we have a collection of $i=0..N$ vertices $\vec{x}_{i,j}$.
-\textbf{[Warum jetzt zweifach indiziert?]} Thus, the total amount of information is an array of shape $(M,N,2)$ Assuming that we have $N$ cells, each of which is represented by $M$ vertices, the total amount of information is an array of shape $(N,M,2)$, containing all positions of all agents.
-It is thus only a matter of comparing two of these arrays with each other and many methods and
-approaches which already exist~\cite{Wang2020} can be leveraged. For now, we simply calculate the distance of each vertex pair, which is equivalent to the
-well-known loss function:
+Each cell contains a collection of vertices.
+Given $M$ cells with $N$ vertices each and dimension $d$, the total information can be stored in an
+array $\nu_{i,j,k}$ of shape $(M,N,2)$.
+Comparing the results of a simulation $\nu$ with experimental values $\tilde{\nu}$ is thus only a
+matter of comparing two of these arrays within a suitable cost function~\cite{Wang2020}.
+In our approach we utilize a least squares approach given by
 
@@
     L_{\text{LS}}(\vec{x}, \vec{y})
-        = \sqrt{\sum\limits_{i,j}||\vec{x}_{i,j}-\vec{y}_{i,j}||^2}
-        = \sqrt{\sum\limits_{i,j,k}|x_{i,j,k}-y_{i,j,k}|^2}
+        = \sqrt{\sum\limits_{i,j,k}|\nu_{i,j,k} - \tilde{\nu}_{i,j,k}|^2}
 \end{equation}
 
-of least squares (LS).
-One important detail is that it must be guaranteed that the ordering of vertices is identical in the
-numerical and experimentally obtained results.
+which minimizes the distance between the corresponding vertices.
+One important detail is that the ordering of vertices must be identical in the numerical and
+experimental values.
 
 \paragraph{Comparison across Generations}
-In order to compare a variable amount of cells, we can not utilize the positions of these cells
-for comparison directly since one particular state of our simulation might contain cells which are
-not present in the experimental results (or vice-versa).
+The preceding algorithm is performant and effective but can not be applied in the event of a
+division process.
+Due to possibly differing division times, a cell which is present in the simulation might not yet
+have a corresponding cell in the experimental data at this time point yet (or vice-versa).
+In order to still be able to compare these datapoints, we require additional methods which can be
+used for comparisons across division events.
 \textbf{[Es bleibt unklar, was du hier verhandeln möchtest.]}
-
-We thus need additional methods which can be used for comparisons across division events.
-To achieve this, we use the numerical results of the \ac{abm} simulation and our visualization
+\textit{Jetzt klarer?}
+
+To this end we use the numerical results of the \ac{abm} simulation and our visualization
 algorithms to generate cell-masks (see section~\ref{subsection:visualization}).
-We then compare the generated masks to the experimentally obtained ones pixel by pixel.
-Whenever two pixel colors are not matching, a loss value is assigned.
-The sum of all those penalties is then the total loss.
-This algorithm is an implementation of a signed area measure.
-It is very important to take into account the parental relationship between the cells, checking if
-either one of the cells is a daughter of the other cell.
+Each pixel has a unique color value which is directly related to one particular cell or background
+(black).
+We compare the generated masks to the experimentally obtained ones pixel by pixel.
+Whenever two colors are not matching, a penalty of $1$ is assigned.
+The total cost is the sum of all penalties.
+% This algorithm is an implementation of a signed area measure.
+It is important to take into account the parental relationship between the cells, by checking if
+either one of the colors corresponding to the compared colors is a daughter of the other cell.
 Figure~\ref{fig:mask-difference-metric} (A,B) shows two successive states of a numerical simulation
 with varying amounts of cells.
-We can see that some division events have occurred but other cells have only grown without
-completing a division step.
+We can see that some cells have divided while others have only grown without completing a division
+step.
 The resulting penalties without accounting for the parental relationships are displayed in
 Figure~\ref{fig:mask-difference-metric} (C).
-Black indicates that the pixels are matching and no loss is assigned while a white pixel
-indicates that the original pixels are not matching, thus assigning a loss value.
-Similarly, Figure~\ref{fig:mask-difference-metric} (D) takes the parental relationship into account
-and highlights these in gray.
-It is possible to assign a custom value for the parental loss value.
-If we set this loss value to $0$, this algorithm ensures the contiuity of the loss function in time.
-
-But it is also biologically motivated.
-If a cell has just divided, the resulting two daughter cells will cover precisely the space that the
-mother had taken up.
-Often times, the exact time at which the fission event is over can not be determined accurately,
-thus leading not only to uncertainties but also some fundamental problems.
-Another approach to construct a comparison algorithm across generations would be to use the
-previously explained position comparison algorithm and use it on intervals in which no division
-event occurrs.
+Black indicates that the pixels are matching and no penalty is assigned while a white pixel
+indicates that the original pixels are not matching, thus adding to the overall cost.
+In contrast, Figure~\ref{fig:mask-difference-metric} (D) takes the parental relationship into
+account and highlights regions where compared pixels belong to mother-daughter cells in gray.
+We assign a penalty $0\leq p \leq 1$ which is a hyperparameter for this optimization algorithm.
+% It is possible to assign a custom value for the parental cost value.
+Settings its value to $0$ results in a contiuous cost function in time.
+
+Future extensions should introduce a time-dependent parent penalty $p=p(t)$ which grows with the age
+of the daughter cells.
+Furthermore, the developed algorithm relies on the comparison of two images and can as such make use
+of existing machine-learning workflows that construct cost functions for such
+scenarios~\cite{Dice1945}.\\
+\textbf{@Christian are these two sentences Discussion?}
+
+Although this optimization scheme was driven by technical challenges, it is also biologically
+motivated.
+In the infenitisemal time point after the division event the resulting two daughter cells will cover
+precisely the space that the mother had taken up.
+Often times, the exact time at which the fission event has commenced can not be determined
+accurately, thus leading to uncertainties with respect to the question when the life of the daughter
+cells begins and the one of the mother ends.
+
+In a different approach, one might be tempted to utilize the previously constructed direct
+comparison algorithm for time intervals that do not contain any division events.
 But this attempt would quickly lead to multiple problems.
-How can we deal with the uncertainties of the timing of the division event?
-And the more cells we want to investigate, the more division events will occurr even in a fixed
-amount of time thus greatly decreasing the effective interval at which this alternative algorithm
-could be used.
-Furthermore, what if there are only two images and there is a division event which is occurring
-between them?
-
-It further needs to be noted that the constructed algorithm can also be adapted by users in order to
-provide more complex loss functions (such as considering time-evolvement, etc.).
-Our method can make use of existing machine leanring algorithms since it is only based on the
-principle of comparing two images with each other.
-There already exists a variety of methods on how to assign loss values in these
-circumstances~\cite{Dice1945}.
+It will not be able to deal with the uncertainties of the timing of the division event.
+As the number of investigated cells grows, more division events will occurr even in a fixed
+amount of time thus greatly decreasing or completely vanishing the effective interval at which no
+division event is observed.
+The latter situation is equivalent to two microscopic images as input data between which a division
+event has occured.
+
+\textbf{@Christian is this Supplement?}
 
@@
         (C) "Naive" comparison of cell masks, which identifies pixels of non-matching colors and
-        assigns a loss by counting these pixels.
+        assigns a cost by counting these pixels.
         This simple approach does not take into consideration the parental relationship.
@@
         colored in gray.\\
-        (E) Comparison of loss functions, exponential regression (ER) and number of cells.
-        % Total loss between successive iteration steps $t_n,t_{n+1}$.
-        The overall shape of the loss is expected to have the slope of an exponential function.
+        (E) Comparison of cost functions, exponential regression (ER) and number of cells.
+        % Total cost between successive iteration steps $t_n,t_{n+1}$.
+        The overall shape of the cost is expected to have the slope of an exponential function.
         We can clearly see that division events introduce jumps within this relationship.
         In the case where we account for the parental relationship, these jumps are mitigated.
-        The loss value was set to $0$ for pixel-comparisons which belong to cells that are in a
+        The cost value was set to $0$ for pixel-comparisons which belong to cells that are in a
         parental relationship.
@@
 data and use the generated masks to provide a benchmark of our algorithm.
-To do this, we set out to calculate the loss between two successive iteration steps
+To do this, we set out to calculate the cost between two successive iteration steps
 $t_n$ with $t_{n+1}$.
-We expect to obtain an exponential growth in the loss.
+We expect to obtain an exponential growth in the cost.
 This can be explained by the fact that the growth of the total area is exponential in time.
@@
 which is again an exponential function.
-Figure~\ref{fig:mask-difference-metric} (E) shows the total loss for this approach.
+Figure~\ref{fig:mask-difference-metric} (E) shows the total cost for this approach.
 We can see that by accounting for the parental relationship, we get the overall desired effect.
@@
 
-By doing so, we can utilize the simpler and more performant position comparison scheme for our loss function.
+By doing so, we can utilize the simpler and more performant direct comparison scheme for our cost function.
 \textbf{[Hier weiss man nicht worüber du sprichst.]}
@@
         We calculated the likelihood profiles for every optimized parameter of our model.
-        All plots show how the value of the loss function varies when fixing a single parameter but
+        All plots show how the value of the cost function varies when fixing a single parameter but
         still optimizing for the remaining ones.

### 2025-10-24 paper/main.tex
apply plos template

@@
-\documentclass{article}
-
-\usepackage{arxiv}
+\documentclass[10pt,letterpaper]{article}
+
+\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
 
@@
 \usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
+% \usepackage{microtype}      % microtypography
 \usepackage{mathrsfs}
@@
 
+% Plos usepackages
+% amsmath and amssymb packages, useful for mathematical formulas and symbols
+\usepackage{amsmath,amssymb}
+
+% Use adjustwidth environment to exceed column width (see example table in text)
+\usepackage{changepage}
+
+% textcomp package and marvosym package for additional characters
+\usepackage{textcomp,marvosym}
+
+% cite package, to clean up citations in the main text. Do not remove.
+\usepackage{cite}
+
+% Use nameref to cite supporting information files (see Supporting Information section for more info)
+\usepackage{nameref,hyperref}
+
+% line numbers
+\usepackage[right]{lineno}
+
+% ligatures disabled
+\usepackage[nopatch=eqnum]{microtype}
+\DisableLigatures[f]{encoding = *, family = * }
+
+% color can be used to apply background shading to table cells only
+\usepackage[table]{xcolor}
+
+% array package and thick rules for tables
+\usepackage{array}
+
+% create "+" rule type for thick vertical lines
+\newcolumntype{+}{!{\vrule width 2pt}}
 \title{Methods for Analyzing Models of\\ Rod-Shaped Bacteria}
@@
 
-\author{
-    \href{https://orcid.org/0009-0001-0613-7978}{
-        \includegraphics[scale=0.06]{figures/orcid.pdf}
-        \hspace{1mm}Jonas Pleyer
-    }
-    \thanks{
-        \href{https://jonas.pleyer.org}{jonas.pleyer.org},
-        \href{https://cellular-raza.com}{cellular-raza.com}
-    }\\
-	Freiburg Center for Data-Analysis and Modeling\\
-	University of Freiburg\\
-	\texttt{jonas.pleyer@fdm.uni-freiburg.de} \\
-	%% examples of more authors
-	\And
-	\href{https://orcid.org/0000-0002-6371-4495}{
-        \includegraphics[scale=0.06]{figures/orcid.pdf}
-        \hspace{1mm}Christian Fleck
-    }\\
-	Freiburg Center for Data-Analysis and Modeling\\
-	University of Freiburg
-}
+\author{Jonas Pleyer and Christian Fleck}
 
@@
 % Uncomment to override  the `A preprint' in the header
-\renewcommand{\headeright}{Preprint}
+% \renewcommand{\headeright}{Preprint}
 %\renewcommand{\undertitle}{Technical Report}
-\renewcommand{\shorttitle}{Methods for Analyzing Models of Rod-Shaped Bacteria}
+% \renewcommand{\shorttitle}{Methods for Analyzing Models of Rod-Shaped Bacteria}
 
@@
 
-% MAKE TITLES IN THEOREMS BOLD
+% create "+" rule type for thick vertical lines
+\newcolumntype{+}{!{\vrule width 2pt}}
+
+% create \thickcline for thick horizontal lines of variable length
+\newlength\savedwidth
+\newcommand\thickcline[1]{%
+  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+  \cline{#1}%
+  \noalign{\vskip\arrayrulewidth}%
+  \noalign{\global\arrayrulewidth\savedwidth}%
+}
+
+% \thickhline command for thick horizontal lines that span the table
+\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+\hline
+\noalign{\global\arrayrulewidth\savedwidth}}
+
+% Remove comment for double spacing
+%\usepackage{setspace} 
+%\doublespacing
+
+% Text layout
+\raggedright
+\setlength{\parindent}{0.5cm}
+\textwidth 5.25in
+\textheight 8.75in
+
+% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
+% Captions will be left justified
+\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
+\renewcommand{\figurename}{Fig}
+
+% Remove brackets from numbering in List of References
 \makeatletter
-\def\th@plain{%
-  \thm@notefont{}% same as heading font
-  \itshape % body font
-}
-\def\th@definition{%
-  \thm@notefont{}% same as heading font
-  \normalfont % body font
-}
+\renewcommand{\@biblabel}[1]{\quad#1.}
 \makeatother
 
+% Header and Footer with logo
+\usepackage{lastpage,fancyhdr,graphicx}
+\usepackage{epstopdf}
+%\pagestyle{myheadings}
+\pagestyle{fancy}
+\fancyhf{}
+%\setlength{\headheight}{27.023pt}
+%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+\rfoot{\thepage/\pageref{LastPage}}
+\renewcommand{\headrulewidth}{0pt}
+\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
+\fancyheadoffset[L]{2.25in}
+\fancyfootoffset[L]{2.25in}
+\lfoot{\today}
+
+%% Include all user-defined macros below
+
+\newcommand{\lorem}{\textbf{LOREM}}
+\newcommand{\ipsum}{\textbf{IPSUM}}
+
+%% END MACROS SECTION
+
 \begin{document}
-\maketitle
+\vspace*{0.2in}
+
+% Title must be 250 characters or less.
+\begin{flushleft}
+{\Large
+\textbf{Methods for analyzing models of rod-shaped bacteria}
+% Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+}
+\newline
+% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
+\\
+\href{https://orcid.org/0009-0001-0613-7978}{
+    \includegraphics[scale=0.06]{figures/orcid.pdf}
+    \hspace{1mm}Jonas Pleyer\textsuperscript{1},
+}
+\href{https://orcid.org/0000-0002-6371-4495}{
+    \includegraphics[scale=0.06]{figures/orcid.pdf}
+    \hspace{1mm}Christian Fleck\textsuperscript{1}
+}
+\\
+\bigskip
+\textbf{1} Freiburg Center for Data-Analysis and Modeling, University of Freiburg, Freiburg Germany
+\\
+\bigskip
+
+\end{flushleft}
 
 %###################################################################################################
-\begin{abstract}
+\section*{Abstract}
 \aclp{abm} have become popular tools in the study of complex cellular systems.
@@
 rod-shaped bacteria such as \textit{E.Coli} or \textit{B.Subtilis}.
-\end{abstract}
 
@@
 
-\bibliographystyle{IEEEtran}
+\bibliographystyle{plos2025}
 \bibliography{references}
@@
 
+\label{LastPage}
 \end{document}

### 2025-10-24 paper/main.tex
remove everything related to blueprint

@@
 for estimating their parameters and performing identifiability analysis on them.
-Our approach may serve as a blueprint for a modular framework which can be adapted to other
-research questions.
+Our approach provides a modular framework which can be adapted to a broad range of research
+questions regarding models of rod-shaped bacteria.
 We use time series of microscopic images as input data and incorporate advanced effects as the
@@
 this framework.
-This modularity highlights a "blueprint" which can be adapted and applied to other \acp{abm}.
 
@@
 }
+\textit{Ich habe den blueprint vollkommen entfernt. Er macht kein Sinn im MS.}
 
@@
 
-With this work we have layed out a "blueprint" to estimate the parameters of \acp{abm} using
+With this work we have layed out how to estimate the parameters of \acp{abm} using
 microscopic image time series.
@@
 }
+\textit{Alles zu blaupause ist wegradiert.}
 
@@
 design minimal \acp{abm} which can be fit to data.
-As such it formulates a blueprint for the parameter estimation of \acp{abm}.
 The developed methods are publicly available as a well-documented python package
@@
 of the fundamental spatial representation of our bacteria as a collection of vertices.
-In this way it highlights a modular blueprint that can be followed by researchers applied to models
-of rod-shaped bacteria which incorporate other and more complex behaviours.
+The developed functionalities are modular and can directly be reused by researchers to asses a wide
+range of rod-shaped bacterial models.
 Combining these methods, our workflow links experimental data to the parameter estimation of a
@@
 
-Our framework allow researchers to employ well-known methods such as profile-likelihood
+Our framework allows researchers to employ well-known methods such as profile-likelihood
 \textbf{TODO CITATION} and sensitivity-analysis.

### 2025-10-24 paper/main.tex
formatting; small comment

@@
         This approach is displayed in (D) where pixels of cells that are directly related have been
-        colored in gray.\\
+        colored in gray.
         (E) Comparison of cost functions, exponential regression (ER) and number of cells.
@@
 \textbf{[Wir müssen darüber reden, mir ist hier vieles unklar. Es könnte nützlich sein, vor der Application Sektion, eine Zusammenfassung zu setzen.]}
+\textit{Bis hierhin bin ich mit meiner Überarbeitung gekommen.}
 

### 2025-10-24 paper/main.tex
Revert "apply plos template"
This reverts commit d842d7f71b7724db3be3276dfbfa32451d5e5af9.

@@
-\documentclass[10pt,letterpaper]{article}
-
-\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
+\documentclass{article}
+
+\usepackage{arxiv}
 
@@
 \usepackage{nicefrac}       % compact symbols for 1/2, etc.
-% \usepackage{microtype}      % microtypography
+\usepackage{microtype}      % microtypography
 \usepackage{mathrsfs}
@@
 
-% Plos usepackages
-% amsmath and amssymb packages, useful for mathematical formulas and symbols
-\usepackage{amsmath,amssymb}
-
-% Use adjustwidth environment to exceed column width (see example table in text)
-\usepackage{changepage}
-
-% textcomp package and marvosym package for additional characters
-\usepackage{textcomp,marvosym}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-
-% Use nameref to cite supporting information files (see Supporting Information section for more info)
-\usepackage{nameref,hyperref}
-
-% line numbers
-\usepackage[right]{lineno}
-
-% ligatures disabled
-\usepackage[nopatch=eqnum]{microtype}
-\DisableLigatures[f]{encoding = *, family = * }
-
-% color can be used to apply background shading to table cells only
-\usepackage[table]{xcolor}
-
-% array package and thick rules for tables
-\usepackage{array}
-
-% create "+" rule type for thick vertical lines
-\newcolumntype{+}{!{\vrule width 2pt}}
 \title{Methods for Analyzing Models of\\ Rod-Shaped Bacteria}
@@
 
-\author{Jonas Pleyer and Christian Fleck}
+\author{
+    \href{https://orcid.org/0009-0001-0613-7978}{
+        \includegraphics[scale=0.06]{figures/orcid.pdf}
+        \hspace{1mm}Jonas Pleyer
+    }
+    \thanks{
+        \href{https://jonas.pleyer.org}{jonas.pleyer.org},
+        \href{https://cellular-raza.com}{cellular-raza.com}
+    }\\
+	Freiburg Center for Data-Analysis and Modeling\\
+	University of Freiburg\\
+	\texttt{jonas.pleyer@fdm.uni-freiburg.de} \\
+	%% examples of more authors
+	\And
+	\href{https://orcid.org/0000-0002-6371-4495}{
+        \includegraphics[scale=0.06]{figures/orcid.pdf}
+        \hspace{1mm}Christian Fleck
+    }\\
+	Freiburg Center for Data-Analysis and Modeling\\
+	University of Freiburg
+}
 
@@
 % Uncomment to override  the `A preprint' in the header
-% \renewcommand{\headeright}{Preprint}
+\renewcommand{\headeright}{Preprint}
 %\renewcommand{\undertitle}{Technical Report}
-% \renewcommand{\shorttitle}{Methods for Analyzing Models of Rod-Shaped Bacteria}
+\renewcommand{\shorttitle}{Methods for Analyzing Models of Rod-Shaped Bacteria}
 
@@
 
-% create "+" rule type for thick vertical lines
-\newcolumntype{+}{!{\vrule width 2pt}}
-
-% create \thickcline for thick horizontal lines of variable length
-\newlength\savedwidth
-\newcommand\thickcline[1]{%
-  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
-  \cline{#1}%
-  \noalign{\vskip\arrayrulewidth}%
-  \noalign{\global\arrayrulewidth\savedwidth}%
-}
-
-% \thickhline command for thick horizontal lines that span the table
-\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
-\hline
-\noalign{\global\arrayrulewidth\savedwidth}}
-
-% Remove comment for double spacing
-%\usepackage{setspace} 
-%\doublespacing
-
-% Text layout
-\raggedright
-\setlength{\parindent}{0.5cm}
-\textwidth 5.25in
-\textheight 8.75in
-
-% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
-% Captions will be left justified
-\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
-\renewcommand{\figurename}{Fig}
-
-% Remove brackets from numbering in List of References
+% MAKE TITLES IN THEOREMS BOLD
 \makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
+\def\th@plain{%
+  \thm@notefont{}% same as heading font
+  \itshape % body font
+}
+\def\th@definition{%
+  \thm@notefont{}% same as heading font
+  \normalfont % body font
+}
 \makeatother
 
-% Header and Footer with logo
-\usepackage{lastpage,fancyhdr,graphicx}
-\usepackage{epstopdf}
-%\pagestyle{myheadings}
-\pagestyle{fancy}
-\fancyhf{}
-%\setlength{\headheight}{27.023pt}
-%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
-\rfoot{\thepage/\pageref{LastPage}}
-\renewcommand{\headrulewidth}{0pt}
-\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
-\fancyheadoffset[L]{2.25in}
-\fancyfootoffset[L]{2.25in}
-\lfoot{\today}
-
-%% Include all user-defined macros below
-
-\newcommand{\lorem}{\textbf{LOREM}}
-\newcommand{\ipsum}{\textbf{IPSUM}}
-
-%% END MACROS SECTION
-
 \begin{document}
-\vspace*{0.2in}
-
-% Title must be 250 characters or less.
-\begin{flushleft}
-{\Large
-\textbf{Methods for analyzing models of rod-shaped bacteria}
-% Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
-}
-\newline
-% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
-\\
-\href{https://orcid.org/0009-0001-0613-7978}{
-    \includegraphics[scale=0.06]{figures/orcid.pdf}
-    \hspace{1mm}Jonas Pleyer\textsuperscript{1},
-}
-\href{https://orcid.org/0000-0002-6371-4495}{
-    \includegraphics[scale=0.06]{figures/orcid.pdf}
-    \hspace{1mm}Christian Fleck\textsuperscript{1}
-}
-\\
-\bigskip
-\textbf{1} Freiburg Center for Data-Analysis and Modeling, University of Freiburg, Freiburg Germany
-\\
-\bigskip
-
-\end{flushleft}
+\maketitle
 
 %###################################################################################################
-\section*{Abstract}
+\begin{abstract}
     \aclp{abm} have become popular tools in the study of complex cellular systems.
@@
     for estimating their parameters and performing identifiability analysis on them.
-Our approach provides a modular framework which can be adapted to a broad range of research
-questions regarding models of rod-shaped bacteria.
+    Our approach may serve as a blueprint for a modular framework which can be adapted to other
+    research questions.
     We use time series of microscopic images as input data and incorporate advanced effects as the
@@
     rod-shaped bacteria such as \textit{E.Coli} or \textit{B.Subtilis}.
+\end{abstract}
 
@@
 
-\bibliographystyle{plos2025}
+\bibliographystyle{IEEEtran}
 \bibliography{references}
@@
 
-\label{LastPage}
 \end{document}

### 2025-10-28 paper/main.tex
place all figures inline (no auto-placement)
add units for parameters and variables

@@
 
-\begin{table}
+\begin{table}[H]
     \centering
     \def\arraystretch{1.3}
-    \begin{tabular}{l c l c}
-        \textbf{Parameters} && \textbf{Variables}\\
+    \begin{tabular}{l c c l c c}
+        \textbf{Parameters} && \textbf{Units} & \textbf{Variables} && \textbf{Units}\\
         \toprule
@@
         \midrule
-        Spring tension & $\gamma$ & Position (vertices) & $\vec{x}_i$\\
-        Rigidity & $\eta$ & Velocity & $\partial_t\vec{x}_i$\\
-        Damping & $\lambda$ & Spring length & $l$ \\
+        Spring tension & $\gamma$ & $\unit{1\per\minute}$ & Position (vertices) & $\vec{x}_i$ & $\unit{\micro\metre}$\\
+        Rigidity & $\eta$ & $\unit{\micro\metre\per\minute^2}$ & Velocity & $\partial_t\vec{x}_i$ & $\unit{\micro\metre\per\minute}$\\
+        Damping & $\lambda$ & $\unit{\micro\metre}$ & Spring length & $l$ & $\unit{\micro\metre}$\\
         \multicolumn{4}{l}{\textbf{Interaction}}\\
         \midrule
-        Thickness & $R$\\
-        Strength & $\epsilon$\\
-        Exponent 1 & $n$\\
-        Exponent 2 & $m$\\
-        Bound & $\beta$\\
-        Cutoff & $\zeta$\\
+        Thickness & $R$ & $\unit{\micro\metre}$\\
+        Strength & $\epsilon$ & $\unit{\micro\metre^2\per\minute^2}$\\
+        Exponent 1 & $n$ & $1$\\
+        Exponent 2 & $m$ & $1$\\
+        Bound & $\beta$ & $\unit{\micro\metre^2\per\minute^2}$\\
+        Cutoff & $\zeta$ & $\unit{\micro\metre}$\\
         \multicolumn{4}{l}{\textbf{Growth \& Division}}\\
@@
 \label{subsec:parameter-estimation-individual-treatment}
-\begin{figure}
+\begin{figure}[H]
     \centering
@@
 %TODO {explain profiles}
-\begin{figure}
+\begin{figure}[H]
     \centering

### 2025-10-28 paper/main.tex
add note for christian
link to figure

@@
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
-    \end{tikzonimage}
+    \end{tikzonimage}\\
+    \textbf{@Christian should we remove images (A-C)? I do not think that they add too much
+    value. Maybe we should combine this figure with \ref{fig:position-extraction-algorithm}}\\
     \begin{tikzonimage}[width=0.5\textwidth]
@@
 Due to possibly differing division times, a cell which is present in the simulation might not yet
-have a corresponding cell in the experimental data at this time point yet (or vice-versa).
+have a corresponding cell in the experimental data at this time point yet (or vice-versa) (see
+Figure~\ref{fig:mask-difference-metric} (E)).
 In order to still be able to compare these datapoints, we require additional methods which can be

### 2025-10-28 paper/main.tex
restructure figure placement; link correctly in text

@@
     \centering
-    \begin{minipage}{0.5\textwidth}
-    \begin{tikzonimage}[width=0.48\textwidth]
+    \begin{tikzonimage}[width=0.24\textwidth]
         {docs/source/_static/fitting-methods/progressions-1.png}
@@
     \hspace{0.01\textwidth}%
-    \begin{tikzonimage}[width=0.48\textwidth]
+    \begin{tikzonimage}[width=0.24\textwidth]
         {docs/source/_static/fitting-methods/progressions-2.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
-    \end{tikzonimage}
-    \linebreak
-    \vspace{0.01\textwidth}
-    \begin{tikzonimage}[width=0.48\textwidth]
+    \end{tikzonimage}%
+    \hspace{0.01\textwidth}%
+    \begin{tikzonimage}[width=0.24\textwidth]
         {docs/source/_static/fitting-methods/progressions-3.png}
@@
     \hspace{0.01\textwidth}%
-    \begin{tikzonimage}[width=0.48\textwidth]
+    \begin{tikzonimage}[width=0.24\textwidth]
         {docs/source/_static/fitting-methods/progressions-4.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{D}};
-    \end{tikzonimage}
-    \end{minipage}%
-    \begin{minipage}{0.49\textwidth}
-        \begin{tikzonimage}[width=\textwidth]
-            {docs/source/_static/fitting-methods/penalty-time-flow.png}%
-            \node at (0, 0.975)[anchor=north west, draw, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{E}};
-        \end{tikzonimage}
+    \end{tikzonimage}\\
+    \begin{minipage}{\textwidth}
+        \begin{tikzonimage}[width=0.48\textwidth]
+            {docs/source/_static/fitting-methods/extract_positionsdivision-comparison.png}
+            \node at (0, 1)[anchor=north west, draw, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+        \end{tikzonimage}%
+        \hspace{0.01\textwidth}%
+        \begin{tikzonimage}[width=0.48\textwidth]
+            {docs/source/_static/fitting-methods/penalty-time-flow.png}
+            \node at (0, 0.975)[anchor=north west, draw, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{F}};
+        \end{tikzonimage}%
     \end{minipage}
@@
         colored in gray.
-        (E) Comparison of cost functions, exponential regression (ER) and number of cells.
+        (E) A schematic representation of the comparison problem involving multiple cells at
+        different generations.
+        An initial configuration produces two results by utilizing distinct methods (i.e. evoling
+        our computational model in time or assigning the data for next time point).
+        (F) Comparison of cost functions, exponential regression (ER) and number of cells.
         % Total cost between successive iteration steps $t_n,t_{n+1}$.
@@
 which is again an exponential function.
-Figure~\ref{fig:mask-difference-metric} (E) shows the total cost for this approach.
+Figure~\ref{fig:mask-difference-metric} (F) shows the total cost for this approach.
 We can see that by accounting for the parental relationship, we get the overall desired effect.

### 2025-10-28 paper/main.tex
rewrite initial applications text

@@
 \paragraph{Direct Comparison}
+\label{paragraph:direct-comparison}
 In the scenario where no cells divide during the observed interval, we can use our extraction
@@
 \textbf{[Wir müssen darüber reden, mir ist hier vieles unklar. Es könnte nützlich sein, vor der Application Sektion, eine Zusammenfassung zu setzen.]}
-\textit{Bis hierhin bin ich mit meiner Überarbeitung gekommen.}
 
@@
 \section{Application 1: Elastic deformations of \textit{E.Coli}}
-\label{section:application-1}
-
-We will use our previously developed computational model and apply it directly to the data provided
+
+To assess that our model can be applied to single cells and showcase the usefulness of the
+previously developed comparison techniques,
+will use our previously developed computational model and apply it directly to the data provided
 by the experiments of Amir et. al~\cite{Amir2014,Amir2014_2}.
-\textbf{[Es wäre gut zu sagen, was für ein Ziel du verfolgst. Warum machst du das?]}
-They studied how a single \textit{E.Coli}, which is fixed at the bottom, reacts to a horizontal
-fluid flow that exerts a force onto the bacteria.
-This setup is displayed in figure~\ref{fig:amir-bending-simulation} (A).
+They studied how a single \textit{E.Coli}, which is partially confined within a narrow tube, reacts
+to a horizontal fluid flow that exerts a force onto the bacterium.
+The experiment is particularly suited as a first attempt since it excludes cell-cell interactions
+and focusses on the behaviour of a single cell.
+It therefore reduces the total amount of parameters and allows for fast iteration times when
+performing parameter estimations.
+The experimental setup is displayed in figure~\ref{fig:amir-bending-simulation} (A).
 To model this configuration, the lowermost vertex of the rod is fixed at the base of the simulation
-domain.
-
-Furthermore, we introduce a force $\vec{F}_i = C |\vec{x}_i - \vec{x}_{i+1}|\sin(\alpha_i)\vec{e}_x$
-that acts similarly to Stoke's Law \textbf{TODO Citation} onto the normal component of each
+domain and vertices which are inside the tubulus are restricted to move only in the y-direction.
+
+\textbf{@Christian is this more clear now?}
+
+We model the force exerted by the fluid flow onto the individual vertices by
+
+\begin{equation}
+    \vec{F}_i = C |\vec{x}_i - \vec{x}_{i+1}|\sin(\alpha_i)\vec{e}_x
+\end{equation}
+
+which acts similarly to Stoke's Law \textbf{TODO Citation} onto the normal component of each
 rod-segment where $\alpha_i$ is the angle of the rod edge between $\vec{x}_i$ and $\vec{x}_{i+1}$
@@
 
-Since this model is void of any interactions or cell division, we are left with 5 parameters:
+Since this model is void of any cell-cell interactions or cell division, we are left with 5 parameters:
 Magnitude of the drag force $C$, growth rate $\alpha$, rigidity of the rod $\eta$, spring tension
-$\gamma$ and damping constant $\lambda$.
+$\gamma$ and damping constant $\lambda$ (see table~\ref{table:list-of-parameters}).
 Figure~\ref{fig:amir-bending-simulation} (B) shows a snapshot of the initial (orange) and final
@@
     \caption{
+        \textbf{TODO remove pixels in units}
+        \textbf{@Christian is it okay to change the units here to seconds from minutes? (compare
+        table above)}
         Subfigure (A) shows a snapshot of the experiments performed by \cite{Amir2014_2}.
-        Division of the bacteria is inhibited and they are trapped within a mother machine such that
-        a controlled flow can be exerted on them.
+        Division of the bacteria is inhibited and they are partially trapped within a mother machine
+        such that a controlled flow can be exerted on them.
         (B) A comparison of the initial and bent state show good visual agreement of our prediction
@@
 To show the usefulness of our previously developed methods
-(see section~\ref{section:parameter-estimation}), we will use them to estimate the parameters
-individual growing \textit{E.Coli} using a sequence of microscopic images.
+(see section~\ref{section:parameter-estimation}), we will use them to estimate the parameters of
+individually growing \textit{E.Coli} using a sequence of microscopic images.
 The data we use~\cite{https://doi.org/10.3203/iwf/k-129} (see
 section~\ref{section:data-code-availability}) has been automatically segmented with the
-omnipose~\cite{Cutler2022} algorithm by mean of which the necessary cell masks are obtained.
+omnipose~\cite{Cutler2022} algorithm to obtain the required cell masks.
 
@@
 
-We have already suspected (see table~\ref{table:simulation-aspects}) that in order to properly
-estimate parameters of our system, we need to treat every cell individually, thus assigning
-parameters on a single-cell level.
-This means that we increase the total number of parameters that need to be sampled within an
-optimization routine.
-Given $N_c$ cells and $N_p$ parameters per cell and $N_u$ individual parameters (where $N_u\leq N_p$), the total amount of parameters to be estimated is:
+We have already suspected (see Table~\ref{table:simulation-aspects}) that in order to properly
+estimate parameters of our system, we need to account for natural cell-variability, thus assigning
+parameters on a per-cell basis.
+We have previously presented a list of $N_p$ parameters (see Table~\ref{table:list-of-parameters})
+which will either be assigned individually or globally.
+We filter out a subset $N_g\leq N_p$ of global parameters, treating the remaining ones individually.
+Given $N_c$ cells, the total amount of parameters to be estimated is:
 \begin{equation}
-    N_{p,\text{total}} = N_u + N_c (N_p - N_u) = N_C N_p - (N_c - 1) N_u.
+    N_{p,\text{tot}} = N_g + N_c (N_p - N_g).
 \end{equation}
-It is clear, that we should avoid to needlessly assign every parameter individually since the volume
-of the parameter space that needs to be sampled, scales exponentially with the dimension ie. the
-number of parameters.
+It is clear, that we should avoid, to needlessly treat parameters non-globally since this increases
+the size of the parameter space considerably.
+The volume that needs to be sampled to optimize within a given parameter space scales exponentially
+with the dimension ie. the total number of parameters $N_{p,\text{tot}}$.
 \textbf{[Das ist leider komplett unverständlich.
@@
 Parameter per cell and individual parameters versteht auch niemand den Unterschied.]}
+\textit{Ich habe versucht, es klarer zu machen. Bitte dazu Feedback!}
 
@@
 in subfigure (C).
-There, we can see much more clearly the difference between considering individually assigned
-parameters and global ones.
-It is clear from this observation that in order for us to accurately estimate the parameters of this
-system, we need to treat the bacteria individually.
+There, we can see much more clearly the difference between considering assigning parameters per cell
+and global ones.
+These results show that in order to accurately estimate the parameters of this system, we need to
+account for this inherent heterogeneity and hence treat the growth parameters of the bacteria
+individually.
 \textbf{[Man ahnt was du meinst, doch leider ahnt man es nur. Vermeide Aussagen wie "it is clear".]}
+\textit{Jetzt klarer?}
 
@@
 \subsection{Without Division}
-In section we will focus on a time-interval which does not contain any division events and at which
-the initial lag-phase of our bacteria has already passed.
-\textbf{[Sprachlich wieder fragwürdig. "A time interval at which the initial lag-phase has passed" ist kein irgendwie korrekter englischer Satz.]}
-
-By doing so, we can utilize the simpler and more performant direct comparison scheme for our cost function.
-\textbf{[Hier weiss man nicht worüber du sprichst.]}
-
-Furthermore, we need not to estimate the lag-time parameters which improves performance of our
-optimization algorithm.
-\textbf{[Verweise auf die entsprechende Gleichung.]}
-\textbf{[Ich werde nun nicht mehr das Sprachliche kommentieren oder verbessern. Zuerst muss die gesamte Konzeption verbessert werden.]}
+In this section we will focus on a time-interval which does not contain any division events.
+By doing so, we can utilize the simpler and more performant direct comparison scheme (see
+\ref{paragraph:direct-comparison}) for our cost function.
+We have chosen said interval such that effects related to the lag phase can be neglected which
+allows us to sample a lower amount of paremeters thus improving runtime performance of our
+optimization scheme.
 
@@
 microscopic image time series.
-\textbf{
-    Es wird leider überhaupt nicht deutlich, was die Blaupause sein könnte.
-    Wenn man das MS liesst, gewinnt man den Eindruck, du machst dies und das, aber ein einheitliches Konzept fehlt.
-    Wir müssen zuerst eine Konzeption für das Paper entwickeln, bevor wir über weitere Details reden können.
-}
-\textit{Alles zu blaupause ist wegradiert.}
-
 Our approach is modularized such that its structure can be applied to a variety of problems.

### 2025-10-28 paper/main.tex
use new images

@@
     \begin{tikzonimage}[width=0.4\textwidth]
-        {docs/source/_static/09395645494836445480/raw_pv/000000400.png}
+        {docs/source/_static/11571737453049821261/raw_pv/000000400.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
@@
     \begin{tikzonimage}[width=0.4\textwidth]
-        {docs/source/_static/09395645494836445480/masks/000000400.png}
+        {docs/source/_static/11571737453049821261/masks/000000400.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};

### 2025-10-29 paper/main.tex
add keywords for title
extend general comment
add image for mie potential

@@
 
+\textbf{Discussed Keywords (Working Title)}
+\begin{itemize}
+    \item Parameter Estimation
+    \begin{itemize}
+        \item Mechanical Properties
+        \item Rod-shaped?
+        \item Single-cell
+    \end{itemize}
+    \item \ac{abm}
+    \item Methods
+\end{itemize}
 % keywords can be removed
@@
 }
-\textit{Ich habe den blueprint vollkommen entfernt. Er macht kein Sinn im MS.}
+\textit{Ich habe den blueprint vollkommen entfernt. Er macht kein Sinn im MS. Der Rest sollte noch
+angepasst werden.}
 
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{docs/source/_static/mechanics.png}
+    \begin{tikzonimage}[width=0.5\textwidth]
+        {docs/source/_static/mechanics.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.5\textwidth]
+        {docs/source/_static/mie-potential-shapes.pdf}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+    \end{tikzonimage}%
     \caption{
-        Example of an elonged bacterial rod represented by a collection of vertices which are
+        (A) Example of an elonged bacterial rod represented by a collection of vertices which are
         connected by springs.
@@
         equilibrium state of a straight rod with no curvature.
-    }
-    \label{fig:mechanics-bacterium}
+        (B) Potential curves of the Mie potential for various combinations of exponents $n,m$ with
+        upper bound $\beta=2.5U_0$ and cutoff $\zeta=3(R_1+R_2)$.
+        \textbf{TODO insert more (possibly schematic depictions of how the model works
+        (Interactions, Cycle))}
+    }
+    \label{fig:model-bacterium}
 \end{figure}

### 2025-10-29 paper/main.tex
correctly link to subfigure

@@
 The resulting force acts along the angle bisector which can be calculated from the edge vectors.
-Figure~\ref{fig:mechanics-bacterium} shows an elongated rod represented by $6$ vertices.
+Figure~\ref{fig:model-bacterium} (A) shows an elongated rod represented by $6$ vertices.
 The force acting on vertex $\vec{x}_i$ is given by
@@
 
-where $\eta_i$ is the rigidity at vertex $\vec{x}_i$ (see Figure~\ref{fig:mechanics-bacterium}).
+where $\eta_i$ is the rigidity at vertex $\vec{x}_i$ (see Figure~\ref{fig:model-bacterium}).
 The adjoining vertices $\vec{x}_{i-1},\vec{x}_{i+1}$ are subject to half of the same force in the
@@
 where $\theta$ is the heaviside step-function and the $\min$ term ensures the upper bound $\beta$.
+A collection of configurations are displayed in figure~\ref{fig:model-bacterium} (B).
 

### 2025-10-29 paper/main.tex
make rescaling of parameters more clear

@@
 
-where $\lambda$ a damping constant and the mass was set to $1$ to simplify the model.
+where $\lambda$ a damping constant.
+The equations have been rescaled by dividing with the mass, thus effectively eliminating this
+parameter and simplifying the model.
+All involved parameters and their units can be viewed in Table~\ref{table:list-of-parameters}.
 

### 2025-10-29 paper/main.tex
reformulate interaction part

@@
 
-Calculating the force between cells requires performing a double integral over the length of the
-cells.
+Calculating the force between cells requires performing a double integral over the two polygonal
+paths of the cells.
 To reduce the computational costs we use an approximative, simplified model.
@@
 for some specific $j$.
+This approach is especially justified for short-ranged interactions.
 The force is then calculated between the points $\vec{x}_i$ and $\vec{p}$ and acts on
@@
 \begin{align}
-    U(r) &= C\epsilon\left[ \left(\frac{\sigma}{r}\right)^n -
+    V(r) &= V_0C\left[ \left(\frac{\sigma}{r}\right)^n -
         \left(\frac{\sigma}{r}\right)^m\right]\\
@@
 
-The strength of the potential is given by $\epsilon$ and the thickness of our rods
+The strength of the potential is given by $V_0$ and the thickness of our rods
 is encoded in $\sigma = (m/n)^{1/(n-m)}(R_1+R_2)$ where $R_1,R_2$ are the thicknesses (radii) of the
@@
 We numerically regularize the potential by imposing an upper bound $\beta$ on the norm of the
-resulting force $\vec{F} = \vec{\nabla} V$.
+resulting force $\vec{F} = -\vec{\nabla} V$.
 Furthermore, we impose a cutoff parameter $\zeta$ which sets sets the force to zero when the
@@
 \begin{equation}
-    \vec{F}_{i,\text{external}} = \vec{F} \theta(r-\zeta) \min\left(1, \beta / |\vec{F}|\right)
+    \vec{F}_{i,\text{external}} = \left\{
+    \begin{array}{ll}
+        \beta\vec{F}/|\vec{F}| & |\vec{F}| >= \beta\\
+        \vec{0} & r > \zeta\\
+        \vec{F} & \text{else}
+    \end{array}\right.
+    % \vec{F} \theta(r-\zeta) \min\left(1, \beta / |\vec{F}|\right)
 \end{equation}
 
-where $\theta$ is the heaviside step-function and the $\min$ term ensures the upper bound $\beta$.
+\textbf{@Christian Can you find a more elegant notation than this equation?}\\
+where the first case ensures that the norm is bound and the second case ensures the cutoff.
 A collection of configurations are displayed in figure~\ref{fig:model-bacterium} (B).
@@
 Therefore, the Mie potential is a good choice for a simple, interpretable interaction potential with
-low amount of parameters that can be estimated by our data-driven approach.
+low amount of parameters that can be estimated.
 
@@
 In order to reuse this model, we also provide the option to assign growth rates by sampling from a
-normal distribution with user-defined parameters.
+normal distribution with user-defined parameters.\textbf{@Christian Supplement?}
 Every other intracellular parameter is copied identically to the daughter cells.
@@
         Thickness & $R$ & $\unit{\micro\metre}$\\
-        Strength & $\epsilon$ & $\unit{\micro\metre^2\per\minute^2}$\\
+        Strength & $V_0$ & $\unit{\micro\metre^2\per\minute^2}$\\
         Exponent 1 & $n$ & $1$\\

### 2025-10-29 paper/main.tex
add comment
combine figures; remove old subfigures

@@
 \subsection{Numerical Implementation}
+\textbf{TODO move this somewhere else (maybe parameter estimation)?}
 
@@
 the desired number of vertices $\vec{x}_i$.
-
-\begin{figure}[H]
-    \centering
-    \begin{tikzonimage}[width=0.36\textwidth]
-        {docs/source/_static/fitting-methods/algorithm/image001042.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.36\textwidth]
-        {docs/source/_static/fitting-methods/algorithm/mask-zoom.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.255\textwidth]
-        {docs/source/_static/fitting-methods/algorithm/interpolate-positions.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
-    \end{tikzonimage}
-    \caption{
-        (A) Input image.
-        (B) Result of the skeletonization algorithm for a single cell.
-        (C) Vertices $\vec{x}_i$ which have been interpolated from the skeletonized cell.
-        They represent the position of our cell-agents.
-    }
-    \label{fig:position-extraction-algorithm}
-\end{figure}
 
@@
     \centering
-    \begin{tikzonimage}[width=0.32\textwidth]
-        {docs/source/_static/fitting-methods/extract_positions-004000.png}
+    \begin{tikzonimage}[width=0.36\textwidth]
+        {docs/source/_static/fitting-methods/algorithm/image001042.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
     \end{tikzonimage}
-    \begin{tikzonimage}[width=0.32\textwidth]
-        {docs/source/_static/fitting-methods/extract_positions-006000.png}
+    \begin{tikzonimage}[width=0.36\textwidth]
+        {docs/source/_static/fitting-methods/algorithm/mask-zoom.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
     \end{tikzonimage}
-    \begin{tikzonimage}[width=0.32\textwidth]
-        {docs/source/_static/fitting-methods/extract_positions-008000.png}
+    \begin{tikzonimage}[width=0.255\textwidth]
+        {docs/source/_static/fitting-methods/algorithm/interpolate-positions.png}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
-    \end{tikzonimage}\\
-    \textbf{@Christian should we remove images (A-C)? I do not think that they add too much
-    value. Maybe we should combine this figure with \ref{fig:position-extraction-algorithm}}\\
+    \end{tikzonimage}
     \begin{tikzonimage}[width=0.5\textwidth]
@@
     \caption{
-        Subfigures (A-C) show rendered masks of a simulation where the extracted positions have been
-        drawn within the individual agents.
+        (A) Input image.
+        (B) Result of the skeletonization algorithm for a single cell.
+        (C) Vertices $\vec{x}_i$ which have been interpolated from the skeletonized cell.
+        They represent the position of our cell-agents.
         Subfigure (D) displayes the average rod length of the exact and extracted data.
@@
     \label{fig:benchmarking-extraction-algorithm}
+    \label{fig:position-extraction-algorithm}
 \end{figure}

### 2025-11-16 paper/main.tex
update todos

@@
 
+\begin{itemize}
+    \item Adapt introduction; \acp{abm} not focus of paper, but rather tool to do the job
+    \begin{itemize}
+        \item model division, growth of bacteria
+        \item \acp{abm} are good for this tool
+    \end{itemize}
+    \item Exchange $\sigma \rightarrow R_1+R_2$
+    \item Equation 8 make double norm $||\vec{F}||$
+    \item Remove auxiliary variable $\vec{F}$ only use $\vec{\nabla}V$
+    \item Table 2 in supplement? (probably not)
+    \item numerical implementation verschieben?
+    \item numerical implementation: waht do we need to simulate this?
+    \begin{itemize}
+        \item \acp{abm} are common tools to realize this
+        \item cellular\_raza is THE tool to do it
+    \end{itemize}
+    \item Methodology: single-cell parameter estimation as motivation for techniques
+    \item Which parameters are global which are individual? Discuss this in general; not in detail
+    \item depends on data, number of parameters (computational efficiency)
+    \item page 12; yes this is Discussion
+    \item page 12: maybe extend the paragraph wrt comparison in intervals
+\end{itemize}
+
 %###################################################################################################
 \section{Introduction}
-\textbf{
-    Mein Vorgehen: ich passe Formulierungen an, wenn ich im Prinzip mit der Aussage einverstanden
-    bin, aber den Wortlaut verbesserungswürdig finde.
-    Ich schreibe Kommentare, dort wo ich eher grundsätzliche Schwierigkeiten sehe.
-    Dort wo ich nichts geändert habe und keinen Kommentar gegeben habe, kann ich zu einem späteren
-    Zeitpunkt doch Änderungen vornehmen wollen.
-}
-
 \acp{abm} have proven to be powerfull tools in the study of complex cellular
@@
 
+\textbf{TODO adapt naming to appliations section}
+
 \begin{equation}
@@
 
+\begin{itemize}
+    \item go from simple to more complex: single-cell, many cells, with division
+    \item explain data in a generalized way; detailed explanations within subsections
+\end{itemize}
+
 %###################################################################################################
@@
 present.
-\textbf{[Ist mir nicht ganz klar]}
 
@@
 values of drag forces.
-\textbf{[Ist mir nicht ganz klar.]}
 
@@
 
+\textbf{TODO what is the parameter $s$? Explain more clearly.}\\
 To remove this non-identifiability we reduce the number of parameters by setting $\gamma=\gamma_0/s$
@@
 The corresponding plots can be seen in \ref{subsec:supplement-parameter-profiles-amir}.
+\textbf{TODO Make this more clear. Transient information would lead to more information. We only
+have initial and final snapshots.}
 
@@
     \caption{
-        \textbf{TODO remove pixels in units}
-        \textbf{@Christian is it okay to change the units here to seconds from minutes? (compare
-        table above)}
+        \textbf{TODO rework caption}
         Subfigure (A) shows a snapshot of the experiments performed by \cite{Amir2014_2}.
@@
 omnipose~\cite{Cutler2022} algorithm to obtain the required cell masks.
+
+\textbf{Possibly explain data more clearly}
+
+\begin{itemize}
+    \item show dataset (few snapshots)
+\end{itemize}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.5\textwidth]{figures/crm_fit-data-progression.png}
+    \caption{TODO}
+\end{figure}
 
@@
 Stichworte sind hier: Cell Variability, Heterogeneity, latent variables.
-Parameter per cell and individual parameters versteht auch niemand den Unterschied.]}
-\textit{Ich habe versucht, es klarer zu machen. Bitte dazu Feedback!}
+Parameter per cell and individual parameters versteht auch niemand den Unterschied.]}\\
+\textit{Ich habe versucht, es klarer zu machen. Bitte dazu Feedback!}\\
+\textbf{Es muss noch im Methoden Teil mehr erwähnt werden.}
 
@@
 We fit a delayed exponential function to them which is defined by
+
+\textbf{Explain fit parameter. Explain $t_0$. Give citation for exponential growth. Also link to
+equation in cell-cycle section}
 
@@
 % --------------------------------------------------------------------------------------------------
-\subsection{Without Division}
+\subsubsection{Without Division}
+
+\begin{itemize}
+    \item refer to which data?
+    \item refer to table of all parameters which are fitted
+    \item refer to equations which are fitted
+    \item "fit full mechanical model"
+    \item link to everything
+\end{itemize}
+
 In this section we will focus on a time-interval which does not contain any division events.

### 2025-11-16 paper/main.tex
add interaction schematic

@@
     \centering
-    \begin{tikzonimage}[width=0.5\textwidth]
-        {docs/source/_static/mechanics.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{A}};
-    \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.5\textwidth]
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {docs/source/_static/mechanics.pdf}
+        \node at (0, 1)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {docs/source/_static/interaction.pdf}
+        \node at (0, 1)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.3\textwidth]
         {docs/source/_static/mie-potential-shapes.pdf}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0, 1)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{C}};
     \end{tikzonimage}%

### 2025-11-16 paper/main.tex
add units
add rendering schematic; extend description of figure

@@
         \midrule
-        Growth rate & $\mu$\\
-        Spring length threshold & $l_\text{thresh}$\\
+        Growth rate & $\mu$ &$\unit{\per\minute}$\\
+        Spring length threshold & $l_\text{thresh}$ &$\unit{\micro\metre}$\\
         \bottomrule
@@
     \centering
-    \begin{tikzonimage}[width=0.4\textwidth]
+    \begin{tikzonimage}[width=0.32\textwidth]
+        {docs/source/_static/imaging-mesh.png}
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \end{tikzonimage}%
+    \hspace{0.01\textwidth}%
+    \begin{tikzonimage}[width=0.32\textwidth]
         {docs/source/_static/11571737453049821261/raw_pv/000000400.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
-    \end{tikzonimage}
-    \hspace{0.01\textwidth}
-    \begin{tikzonimage}[width=0.4\textwidth]
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+    \end{tikzonimage}%
+    \hspace{0.01\textwidth}%
+    \begin{tikzonimage}[width=0.32\textwidth]
         {docs/source/_static/11571737453049821261/masks/000000400.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
     \end{tikzonimage}
     \caption{
-        (A) shows the result from combining sphere and cylinder meshes to obtain the shape of a
+        (A) Cells are rendered by combining spheres and cylinders into a single mesh.
+        Spheres are highlighted in blue and cylinders in gray.
+        (B) shows the result from combining sphere and cylinder meshes to obtain the shape of a
         bacterial rod.
@@
         on the upper part of the meshes.
-        (B) Unique color values are assigned to each agent and light sources are removed.
+        (C) Unique color values are assigned to each agent and light sources are removed.
         Afterwards, the image is rendered by projecting along the z-axis.
+        Modifiers such as lighting or anti-aliasing are deactivated to ensure that each cell is made
+        up of exactly one RGB color.
     }

### 2025-11-16 paper/main.tex
modify section nesting
use new profile plots

@@
 
+\section{Applications}
+
 \begin{itemize}
@@
 %###################################################################################################
-\section{Application 1: Elastic deformations of \textit{E.Coli}}
+\subsection{Application 1: Elastic deformations of \textit{E.Coli}}
 
@@
         \begin{tikzonimage}[width=0.33\textwidth]
-            {figures/crm_amir/profiles-full/fit-comparison.pdf}%
-            \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+            {figures/crm_amir/profiles-reduced/fit-comparison.pdf}%
+            \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
         \end{tikzonimage}%
         \begin{tikzonimage}[width=0.33\textwidth]
-            {figures/crm_amir/profiles-full/drag force.pdf}%
-            \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+            {figures/crm_amir/profiles/drag_force.pdf}%
+            \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
         \end{tikzonimage}%
         \begin{tikzonimage}[width=0.33\textwidth]
-            {figures/crm_amir/profiles-full/growth rate.pdf}%
-            \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+            {figures/crm_amir/profiles/growth_rate.pdf}%
+            \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
         \end{tikzonimage}\\
         \begin{tikzonimage}[width=0.33\textwidth]
-            {figures/crm_amir/profiles-full/rod rigidity.pdf}%
-            \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+            {figures/crm_amir/profiles/rod_rigidity.pdf}%
+            \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
         \end{tikzonimage}%
         \begin{tikzonimage}[width=0.33\textwidth]
-            {figures/crm_amir/profiles-full/spring tension.pdf}%
-            \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
+            {figures/crm_amir/profiles/damping_to_spring_tension_ratio.pdf}%
+            \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
         \end{tikzonimage}%
         \begin{tikzonimage}[width=0.33\textwidth]
-            {figures/crm_amir/profiles-full/damping.pdf}%
-            \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
+            {figures/crm_amir/profiles/damping-spring_tension.pdf}%
+            \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
         \end{tikzonimage}%
@@
 %###################################################################################################
-\section{Application 2: \textit{E.Coli.} growing on Agar}
+\subsection{Application 2: \textit{E.Coli.} growing on Agar}
 \label{section:application-2}
@@
 % --------------------------------------------------------------------------------------------------
-\subsection{Individual Treatment}
+\subsubsection{The role of Heterogeneity}
 
@@
 % --------------------------------------------------------------------------------------------------
-\subsection{With Division}
+\subsubsection{With Division}
 

### 2025-11-16 paper/main.tex
rename rod length variable to match previous ref
change heterogeneity plot

@@
 \begin{equation}
-    x(t) =
+    l(t) =
     \left\{\begin{array}{ll}
-            x_0 \exp(\lambda (t-t_0)) & t\geq t_0\\
-            x_0 & \text{ else}
+            l_0 \exp(\lambda (t-t_0)) & t\geq t_0\\
+            l_0 & \text{ else}
     \end{array}\right.
@@
     \centering
-    \begin{tikzonimage}[width=0.33\textwidth]
+    \begin{tikzonimage}[width=0.4\textwidth]
         {figures/crm_estimate_params/IWF-Goettingen/rod-lengths-individual.pdf}%
@@
     \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.33\textwidth]
-        {figures/crm_estimate_params/IWF-Goettingen/rod-lengths-average.pdf}%
+    \begin{tikzonimage}[width=0.4\textwidth]
+        {figures/crm_estimate_params/IWF-Goettingen/growth-rates-distribution.pdf}%
         \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
-    \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.33\textwidth]
-        {figures/crm_estimate_params/IWF-Goettingen/parameter-distribution.pdf}%
-        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
     \end{tikzonimage}%

### 2025-11-16 paper/main.tex
remove sentence for diffusion
remove supplement subsection and link to it

@@
 interval.
-The corresponding plots can be seen in \ref{subsec:supplement-parameter-profiles-amir}.
+\textbf{Plots are not in supplement anymore}
+The corresponding plots can be seen in \textbf{TODO}.
 \textbf{TODO Make this more clear. Transient information would lead to more information. We only
@@
 individually for each cell.
-We removed any stochastic motion contributions by setting $D=0$.
 This leads to a total of $5+1\times 6=11$ variable parameters.
@@
 \label{sec:supplement-parameter-profiles}
-
-\subsubsection{Amir}
-\label{subsec:supplement-parameter-profiles-amir}
-
-The parameter ranges were chosen such that a suitable 
-Plots are chosen for visual ease; parameter ranges have been widened to ensure results are
-accurate (maybe just leave this in the appendix).
-
-\begin{figure}[H]
-    \centering
-    \begin{tikzonimage}[width=0.33\textwidth]
-        {figures/crm_amir/profiles-reduced/fit-comparison.pdf}%
-        \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
-    \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.33\textwidth]
-        {figures/crm_amir/profiles-reduced/drag force.pdf}%
-        \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
-    \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.33\textwidth]
-        {figures/crm_amir/profiles-reduced/growth rate.pdf}%
-        \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
-    \end{tikzonimage}\\
-    \begin{tikzonimage}[width=0.33\textwidth]
-        {figures/crm_amir/profiles-reduced/rod rigidity.pdf}%
-        \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
-    \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.33\textwidth]
-        {figures/crm_amir/profiles-reduced/rigidity to spring tension ratio.pdf}%
-        \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
-    \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.33\textwidth]
-        {figures/crm_amir/profiles-reduced/damping.pdf}%
-        \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
-    \end{tikzonimage}%
-    \caption{
-        \textbf{TODO}
-    }
-\end{figure}
 

### 2025-11-16 cr_mech_coli/crm_amir/main.py
add title; polish by default

@@ -340,6 +340,7 @@ def plot_results(
     )
     ax.set_xlabel("[µm]")
     ax.set_ylabel("[µm]")
+    ax.set_title("Fit Comparison")
     fig.savefig(output_dir / "fit-comparison.png")
     fig.savefig(output_dir / "fit-comparison.pdf")
     plt.close(fig)
@@ -612,7 +613,7 @@ def crm_amir_main():
     )
     parser.add_argument(
         "--skip-polish",
-        default=True,
+        default=False,
         action="store_true",
         help="Skips polishing the result of the differential evolution algorithm",
     )
@@ -630,7 +631,7 @@ def crm_amir_main():
     )
     parser.add_argument(
         "--skip-polish-profiles",
-        default=True,
+        default=False,
         action="store_false",
         help="See POLISH",
     )

### 2025-11-16 cr_mech_coli/crm_estimate_params/main.py
add new plot which shows overall distribution

@@ -3,7 +3,7 @@ import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib as mpl
 import cr_mech_coli as crm
-from cr_mech_coli.plotting import COLOR3, COLOR5
+from cr_mech_coli.plotting import COLOR2, COLOR3, COLOR4, COLOR5
 import scipy as sp
 import multiprocessing as mp
 from pathlib import Path
@@ -176,8 +176,74 @@ def estimate_growth_curves_individual(filenames, out_path, delay=None):
     fig.tight_layout()
     fig.savefig(out_path / "rod-lengths-individual.png")
     fig.savefig(out_path / "rod-lengths-individual.pdf")
-
     ax.cla()
+
+    growth_rates = [p[1] for p in parameters]
+    growth_rates_uncert = [p[1][1] ** 0.5 for p in covariances]
+
+    xmin = np.min(np.array(growth_rates) - 2 * np.array(growth_rates_uncert))
+    xmax = np.max(np.array(growth_rates) + 2 * np.array(growth_rates_uncert))
+
+    x = np.linspace(xmin, xmax, 200)
+    dx = (xmax - xmin) / len(x)
+
+    yfin = np.zeros(x.shape)
+    for n in range(len(growth_rates)):
+        mu = growth_rates[n]
+        sigma = growth_rates_uncert[n]
+        y = (
+            1
+            / np.sqrt(2 * np.pi * sigma**2)
+            * np.exp(-((x - mu) ** 2) / (2 * sigma**2))
+        )
+        yfin += y
+        ax.plot(x, y, color=COLOR2, alpha=0.6, linestyle="--", label="Individual Fits")
+    yfin /= len(growth_rates)
+
+    # Plot curve fillings for 20% intervals
+    z = np.array([np.sum(yfin[:i]) for i in range(len(yfin))]) * dx
+
+    ind0 = 0
+    for n, q in enumerate(np.arange(0, 1.2, 0.2)):
+        ind1 = np.argmin(q >= z)
+        if ind1 == 0:
+            ind1 = len(z)
+        ind2 = min(ind1 + 1, len(z))
+        ax.fill_between(
+            x[ind0:ind2],
+            np.zeros(ind2 - ind0),
+            yfin[ind0:ind2],
+            color=COLOR2 if n % 2 == 0 else COLOR4,
+            alpha=0.5,
+        )
+        ind0 = ind1
+
[…]

### 2025-11-16 cr_mech_coli/plotting.py
make minor axes optional

@@ -34,8 +34,11 @@ def set_mpl_rc_params():
     )
 
 
-def configure_ax(ax):
+def configure_ax(ax, minor=True):
     ax.grid(True, which="major", linestyle="-", linewidth=0.75, alpha=0.25)
     ax.minorticks_on()
-    ax.grid(True, which="minor", linestyle="-", linewidth=0.25, alpha=0.15)
+    if minor:
+        ax.grid(True, which="minor", linestyle="-", linewidth=0.25, alpha=0.15)
+    else:
+        ax.grid(False, which="minor")
     ax.set_axisbelow(True)

### 2025-11-16 paper/main.tex
restructure methods part
- make paragraphs into subsubsections
- shift figure inside of section

@@
 
-\paragraph{Direct Comparison}
+\subsubsection{Direct Comparison}
 \label{paragraph:direct-comparison}
@@
 
-\paragraph{Comparison across Generations}
+\subsubsection{Imaging-based Comparison Algorithm}
 The preceding algorithm is performant and effective but can not be applied in the event of a
@@
 
+\paragraph{Testing the Comparison Algorithms with Division Events}
+When utilizing a computational model (see section~\ref{sec:computational-model}), we can generate
+data and use the generated masks to provide a benchmark of our algorithm.
+To do this, we set out to calculate the cost between two successive iteration steps
+$t_n$ with $t_{n+1}$.
+We expect to obtain an exponential growth in the cost.
+This can be explained by the fact that the growth of the total area is exponential in time.
+Furthermore, the temporal area difference can be viewed as a derivative of the total area in time,
+which is again an exponential function.
+Figure~\ref{fig:mask-difference-metric} (F) shows the total cost for this approach.
+We can see that by accounting for the parental relationship, we get the overall desired effect.
+This method allows us to compare results between states that have a different number of cells.
+However, since this algorithm requires us to calculate a cell-mask for each comparison and then also
+process this mask as well as the comparator, it is much more computationally demanding.
+After all, for each comparison step, we need to run a full 3D visualization pipeline.
+In cases where the simpler approach of comparing positions of the agents is possible, this method
+should be preferred.
+\textbf{[Wir müssen darüber reden, mir ist hier vieles unklar. Es könnte nützlich sein, vor der Application Sektion, eine Zusammenfassung zu setzen.]}
+
 \begin{figure}[H]
@@
 
-\paragraph{Testing the Comparison Algorithm with Division Events}
-When utilizing a computational model (see section~\ref{sec:computational-model}), we can generate
-data and use the generated masks to provide a benchmark of our algorithm.
-To do this, we set out to calculate the cost between two successive iteration steps
-$t_n$ with $t_{n+1}$.
-We expect to obtain an exponential growth in the cost.
-This can be explained by the fact that the growth of the total area is exponential in time.
-Furthermore, the temporal area difference can be viewed as a derivative of the total area in time,
-which is again an exponential function.
-Figure~\ref{fig:mask-difference-metric} (F) shows the total cost for this approach.
-We can see that by accounting for the parental relationship, we get the overall desired effect.
-This method allows us to compare results between states that have a different number of cells.
-However, since this algorithm requires us to calculate a cell-mask for each comparison and then also
-process this mask as well as the comparator, it is much more computationally demanding.
-After all, for each comparison step, we need to run a full 3D visualization pipeline.
-In cases where the simpler approach of comparing positions of the agents is possible, this method
-should be preferred.
-\textbf{[Wir müssen darüber reden, mir ist hier vieles unklar. Es könnte nützlich sein, vor der Application Sektion, eine Zusammenfassung zu setzen.]}
-
 \section{Applications}
-
-\begin{itemize}
-    \item go from simple to more complex: single-cell, many cells, with division
-    \item explain data in a generalized way; detailed explanations within subsections
-\end{itemize}
 

### 2025-11-16 paper/main.tex
include data-set figure later

@@
 
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.5\textwidth]{figures/crm_fit-data-progression.png}
-    \caption{TODO}
-\end{figure}
-
 % --------------------------------------------------------------------------------------------------
@@
     \centering
-    \begin{tikzonimage}[width=0.4\textwidth]
+    \begin{tikzonimage}[width=0.28\textwidth]
+        {figures/crm_fit-data-progression.png}
+        \node at (0.02, 0.98)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \end{tikzonimage}%
+    \hspace{0.01\textwidth}%
+    \begin{tikzonimage}[width=0.35\textwidth]
         {figures/crm_estimate_params/IWF-Goettingen/rod-lengths-individual.pdf}%
-        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{A}};
-    \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.4\textwidth]
+        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.35\textwidth]
         {figures/crm_estimate_params/IWF-Goettingen/growth-rates-distribution.pdf}%
-        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
     \end{tikzonimage}%
     \caption{
+        \textbf{TODO Fix caption!}\\
         % TODO use same y-axis across the plots
-        (A) The subfigure shows the time-evolution of the rod-lengths for every individual
+        (B) The subfigure shows the time-evolution of the rod-lengths for every individual
         bacterium.
         These growth curves are fitted individually with delayed exponential function.
-        (B) The datapoints are created by averaging the rod-lengths shown in subfigure (A) and then
+        (C) The datapoints are created by averaging the rod-lengths shown in subfigure (A) and then
         fitted with an exponential function.

### 2025-11-16 paper/main.tex
write short introductory note for applications section

@@
 \section{Applications}
+
+In this section we apply the previously developed methods to the problem of parameter estimation
+of our model.
+We cover 3 distinct cases with increasing level of complexity.
+In the first study, we utilize results previously obtained by \cite{Amir2014,Amir2014_2} who
+investigated elastic deformations of \textit{E.Coli} that are partially trapped within a narrow
+tube and subject to a force resulting from a fluidic flow.
+This setup does not involve any cell-cell interactions but specifically requires a flexible rod.
+It is therefore well suited as a starting point to assess the usefulness of our methods and the
+ability of our model to describe these results.\\
+The last two cases consider time-series of microscopic images containing multiple cells (see
+subsection~\ref{section:application-2}) which will be used to perform a parameter estimation of our
+full mechanical model.
+The approach is split into a first part which considers a time interval that does not contain any
+division events and a second part which extends the initially chosen interval to include division
+events.
+In the first part, we utilize the direct comparison algorithm while the latter requires usage of the
+imaging-based comparison algorithm.
 

### 2025-11-18 paper/main.tex
rewrite application 1
- adjust to new simulation
- try to explain more clearly what is being done

@@
 
-\textbf{@Christian is this more clear now?}
-
 We model the force exerted by the fluid flow onto the individual vertices by
-
 \begin{equation}
-    \vec{F}_i = C |\vec{x}_i - \vec{x}_{i+1}|\sin(\alpha_i)\vec{e}_x
+    \vec{F}_{i,\text{external}} = C |\vec{x}_i - \vec{x}_{i+1}|(y-y_b)\sin(\beta_i)\vec{e}_x
 \end{equation}
-
-which acts similarly to Stoke's Law \textbf{TODO Citation} onto the normal component of each
-rod-segment where $\alpha_i$ is the angle of the rod edge between $\vec{x}_i$ and $\vec{x}_{i+1}$
-and the x-axis.
-We extract the position of the rod from the provided image and use it as initial value.
-Our simulation is split into two phases, one with an external force that initially bends the rod,
+which is proportional to the length of the rod segment projected along the y-axis.
+It acts similarly to Stoke's Law~\cite{Stokes1856} along the axis of the fluid flow (x-axis).
+Here $\beta_i$ is the angle of the rod edge between $\vec{x}_{i+1}-\vec{x}_i$ and the x-axis.
+Amir et. al~\cite{Amir2014_2} calculated the flow profile for this particular setup.
+Since the bacterium does not reach the peak of the velocity profile, we can estimate the slope via a
+linear dependence on the pertruding segment along the y-axis~\cite{Wolschin2021}.
+In particular, this enforces the correct boundary values the flow profile and the resulting force
+$\vec{F}_{i,\text{external}}|_{y=0}=0$.
+
+We extract the positions of the rod from the data images (see
+Figure~\ref{fig:amir-bending-simulation} (A)) and compare the simulated positions with the extracted
+ones using the direct comparison algorithm (see Subsection~\ref{subsec:direct-comparison}).
+Our simulation is split into two phases; one with an external force that initially bends the rod,
 and a second one during which the force is disabled and the cell returns to its original shape.
-We compare the simulated positions with the extracted ones and calculate the euclidean distance
-between them, which we use as the cost function to estimate the parameters of this model.
-
-Since this model is void of any cell-cell interactions or cell division, we are left with 5 parameters:
+We assume that these time intervals are identically long since the data from the movies does not
+allow for a finer resolution of the dynamics.
+
+Since this model is void of any cell-cell interactions or cell division, the dynamics of the cell
+can be described by a combination of growth and physical mechanics (see
+subsection~\ref{subsection:mechanical-model-mechanics}).
+We are thus left with 5 parameters:
 Magnitude of the drag force $C$, growth rate $\alpha$, rigidity of the rod $\eta$, spring tension
@@
 present.
-
-Note that each point represented in the subfigures~\ref{fig:amir-bending-simulation} (C-G) was
-optimized again, meaning that higher values of rigidity and spring tension can also lead to higher
-values of drag forces.
 

### 2025-11-18 src/crm_amir.rs
use tube width to calculate correct flow profile

@@ -26,6 +26,8 @@ short_default::default! {
         domain_size: f32 = 200.0,
         /// Size for which to block movement along additional coordinate
         block_size: f32 = 30.0,
+        /// Overall width of the tube
+        tube_width: f32 = 100.0,
         /// Drag force exerted by the flow
         drag_force: f32 = 0.1,
         /// Maximum simulation time
@@ -168,6 +170,7 @@ impl Cycle<FixedRod, f32> for FixedRod {
 struct MyDomain {
     domain: CartesianCuboidRods<f32, 3>,
     block_size: f32,
+    tube_width: f32,
 }
 
 impl Domain<FixedRod, MySubDomain> for MyDomain {
@@ -192,6 +195,7 @@ impl Domain<FixedRod, MySubDomain> for MyDomain {
             rng_seed,
         } = self.domain.decompose(n_subdomains, cells)?;
         let block_size = self.block_size;
+        let tube_width = self.tube_width;
         let index_subdomain_cells = index_subdomain_cells
             .into_iter()
             .map(|(index, subdomain, cells)| {
@@ -200,6 +204,7 @@ impl Domain<FixedRod, MySubDomain> for MyDomain {
                     MySubDomain {
                         subdomain,
                         block_size,
+                        tube_width,
                     },
                     cells,
                 )
@@ -221,6 +226,7 @@ struct MySubDomain {
     #[SortCells]
     subdomain: CartesianSubDomainRods<f32, 3>,
     block_size: f32,
+    tube_width: f32,
 }
 
 impl SubDomainForce<RodPos, RodPos, RodPos, f32> for MySubDomain {
@@ -240,7 +246,8 @@ impl SubDomainForce<RodPos, RodPos, RodPos, f32> for MySubDomain {
                 let dir = (p2 - p1).normalize();
                 let angle = dir.angle(&nalgebra::matrix![0.0, 1.0, 0.0]);
                 let y = ((p1[0] + p2[0]) / 2.0 - self.block_size).max(0.0);
-                let f = self.subdomain.gel_pressure * length * angle.sin() * y;
+                let r = self.tube_width;
+                let f = self.subdomain.gel_pressure * length * angle.sin() * (y * r - y.powi(2));
                 force.row_mut(n1)[2] -= f / 2.0;
                 force.row_mut(n2)[2] -= f / 2.0;
             });
@@ -339,6 +346,7 @@ fn run_sim(
             surface_friction_distance: f32::INFINITY,
         },
[…]

### 2025-11-18 paper/main.tex
rewrite amir subsection
- cover new force type
- explain observed profiles
- discuss model reduction

@@
 \begin{equation}
-    \vec{F}_{i,\text{external}} = C |\vec{x}_i - \vec{x}_{i+1}|(y-y_b)\sin(\beta_i)\vec{e}_x
+    \vec{F}_{i,\text{external}} = \frac{C}{2} |\vec{x}_i - \vec{x}_{i+1}|(yR - y^2)\sin(\beta_i)\vec{e}_x
 \end{equation}
 which is proportional to the length of the rod segment projected along the y-axis.
-It acts similarly to Stoke's Law~\cite{Stokes1856} along the axis of the fluid flow (x-axis).
+Similarly to Stoke's Law~\cite{Stokes1856}, it acts along the axis of the fluid flow (x-axis).
 Here $\beta_i$ is the angle of the rod edge between $\vec{x}_{i+1}-\vec{x}_i$ and the x-axis.
-Amir et. al~\cite{Amir2014_2} calculated the flow profile for this particular setup.
+The flow profile for this particular setup is proportional to $(2yR-y^2)$ where $y$ is the vertical
+displacement measured from the end of the tube~\cite{Wolschin2021}
+(see Figure~\ref{fig:amir-bending-simulation}) and $R$ is the total size of the chamber along the
+y-axis.
 Since the bacterium does not reach the peak of the velocity profile, we can estimate the slope via a
@@
 ones using the direct comparison algorithm (see Subsection~\ref{subsec:direct-comparison}).
-Our simulation is split into two phases; one with an external force that initially bends the rod,
-and a second one during which the force is disabled and the cell returns to its original shape.
-We assume that these time intervals are identically long since the data from the movies does not
-allow for a finer resolution of the dynamics.
-
-Since this model is void of any cell-cell interactions or cell division, the dynamics of the cell
+Our simulation is split into two phases; One with the external flow force that initially bends the
+rod and lasts $\SI{7}{\second}$ and a second phase during which the force is disabled and the cell
+returns to its original shape over a span of $\SI{17}{\second}$.
+
+Since this setup is void of any cell-cell interactions or cell division, the dynamics of the cell
 can be described by a combination of growth and physical mechanics (see
@@
 $\gamma$ and damping constant $\lambda$ (see table~\ref{table:list-of-parameters}).
-Figure~\ref{fig:amir-bending-simulation} (B) shows a snapshot of the initial (orange) and final
-(blue) position of the rod using the optimized parameters.
-While the parameters for the dragging force, growth rate and damping can be identified, the same
-can not be said for the rod rigidity and spring tension, which both are not identifiable.
-However this behaviour is expected since, the shape of the rod is almost exclusively straight, thus
-adding no further information except that a lower threshold of rigidity and spring tension should be
-present.
-
-Furthermore the used images only contain snapshots of a fully-bent or fully-relaxed rod without any
-intermediate steps.
-This effectively limits us to 2 datapoints and is likely the key factor that leads to the two local
-minima observed in the damping profile.
-This is further supported by the observation that the contributions to the cost function is evenly
-distributed between the bent and relaxed position of our system.
-
-\textbf{TODO what is the parameter $s$? Explain more clearly.}\\
-To remove this non-identifiability we reduce the number of parameters by setting $\gamma=\gamma_0/s$
-and $\eta=\eta_0 s$ for the rod rigidity and spring tension, resp.
-We then optimize for this parameter and pick $\gamma_0=\SI{15}{\micro\metre}$ with
-$\eta_0=\SI{150}{\micro\metre\per\second\squared}$ which corresponds to $60\%$ of the previously considered
-interval.
-\textbf{Plots are not in supplement anymore}
-The corresponding plots can be seen in \textbf{TODO}.
-\textbf{TODO Make this more clear. Transient information would lead to more information. We only
-have initial and final snapshots.}
+Figure~\ref{fig:amir-bending-simulation} (B) shows a snapshot of the initial (upright, light blue)
+and intermediate (bent, dark blue) position of the rod using the optimized parameters.
+The likelihood profiles displayed in figure~\ref{fig:amir-bending-simulation} (C-D, blue curves)
+show that each parameter is identifiable from the given data.
+However when it comes to the damping constant $\lambda$, its overall numerical value and steepness
+of the curve compared to the results for the other parameters are much lower.
+We can assess if the system could be described by a reduced model by setting $\lambda=0$, thus
+effectively removing any damping.
+The curves which are obtained in this way are displayed in orange and show that the system is also
+fully identifiable in its reduced version.
+
+It should be noted that the used images only contain snapshots of a fully-bent or fully-relaxed rod
+without any intermediate steps.
+This effectively limits us to effectively 2 datapoints and is likely the key factor that leads to
+the previously described phenomenon.
+The damping force is of dynamic nature, and takes no effect in the two steady states of the rod
+being either fully bent or fully relaxed.
+Gathering additional transient information would allow us to better estimate the damping parameter.
 

### 2025-11-18 paper/main.tex
add supplement section for amir details
- shows data and snapshots
- should be extended to include some formulae and explanations

@@
 $r\geq\xi$.
+
+%###################################################################################################
+\subsection{Experimental Setup of Amir et. al}
+
+\textbf{TODO explain some of the derivations for the flow profile etc.}
+
+\begin{figure}[H]
+    \centering
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {figures/crm_amir/progressions/step4-000010.png}
+        \node at (0.02,17pt)[anchor=north west, white, draw, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \end{tikzonimage}%
+    \hspace{0.01\textwidth}%
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {figures/crm_amir/progressions/step4-000017.png}
+        \node at (0.02,17pt)[anchor=north west, white, draw, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+    \end{tikzonimage}%
+    \hspace{0.01\textwidth}%
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {figures/crm_amir/progressions/step4-000034.png}
+        \node at (0.02,17pt)[anchor=north west, white, draw, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+    \end{tikzonimage}\\
+    \vspace{0.01\textwidth}%
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {figures/crm_amir/result-full/0000000000.png}
+        \node at (0.02, 0.98)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+    \end{tikzonimage}%
+    \hspace{0.01\textwidth}%
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {figures/crm_amir/result-full/0000000001.png}
+        \node at (0.02, 0.98)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+    \end{tikzonimage}%
+    \hspace{0.01\textwidth}%
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {figures/crm_amir/result-full/0000000002.png}
+        \node at (0.02, 0.98)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{F}};
+    \end{tikzonimage}%
+    \caption{
+        (A-C) Images by Amir et. al~\cite{Amir2014_2} which have been manually segmented.
+        The red line within the mask represents the extracted positions.
+        (D-F) Snapshots of the numerical simulation showing the rod in its initial, fully bent and
+        fully relaxed state.
+    }
+\end{figure}
 

### 2025-11-18 paper/main.tex
update todos

@@
 
+\textbf{TODO Talk about how to fit individual bacteria and why it is important.}
+
 % --------------------------------------------------------------------------------------------------
@@
 
+\textbf{Put this in Supplement and discuss it further. Still link to it here.}\\
 In a different approach, one might be tempted to utilize the previously constructed direct
@@
 
-\textbf{@Christian is this Supplement?}
-
-\textbf{[Wir müssen darüber reden, mir ist hier vieles unklar.]}
-
-\paragraph{Testing the Comparison Algorithms with Division Events}
+\subsubsection{Testing the Comparison Algorithms with Division Events}
 When utilizing a computational model (see section~\ref{sec:computational-model}), we can generate
@@
 
-\textbf{Possibly explain data more clearly}
-
-\begin{itemize}
-    \item show dataset (few snapshots)
-\end{itemize}
+\textbf{Explain data more clearly. Refer to figure.}
 
@@
         \textbf{TODO Fix caption!}\\
+        \textbf{Link to supplement for discussion of bar plot}\\
         % TODO use same y-axis across the plots
@@
     \caption{
+        \text{TODO include missing profiles here}
         Likelihood profiles of every estimated parameter.

### 2025-11-18 paper/main.tex
remove subsection from supplement
change label

@@
 % --------------------------------------------------------------------------------------------------
-\subsection{Parameter Estimation}
+\subsection{Parameter Estimation Methods}
 \label{section:parameter-estimation}
@@
 \subsubsection{Direct Comparison}
-\label{paragraph:direct-comparison}
+\label{subsec:direct-comparison}
 In the scenario where no cells divide during the observed interval, we can use our extraction
@@
 By doing so, we can utilize the simpler and more performant direct comparison scheme (see
-\ref{paragraph:direct-comparison}) for our cost function.
+\ref{subsec:direct-comparison}) for our cost function.
 We have chosen said interval such that effects related to the lag phase can be neglected which
@@
 \label{sec:supplement-parameter-profiles}
-
-\subsubsection{With Division}
 

### 2025-11-18 paper/main.tex
use new annotated plot
remove plot for void ratio; reintroduce profiles for every parameter

@@
         \begin{tikzonimage}[width=\textwidth]
-            {figures/amir-elastic-000032-wider.png}%
+            {figures/amir-elastic-000032-with-axes.png}%
             \node at (0.01, 0.01+17pt)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
@@
         \begin{tikzonimage}[width=0.33\textwidth]
-            {figures/crm_amir/profiles/damping_to_spring_tension_ratio.pdf}%
+            {figures/crm_amir/profiles/spring_tension.pdf}%
             \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
@@
         \begin{tikzonimage}[width=0.33\textwidth]
-            {figures/crm_amir/profiles/damping-spring_tension.pdf}%
+            {figures/crm_amir/profiles/damping.pdf}%
             \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};

### 2025-11-18 paper/main.tex
further improvements to amir section
- rewrite equations to make them more clear
- adapt caption of figure

@@
 We model the force exerted by the fluid flow onto the individual vertices by
-\begin{equation}
-    \vec{F}_{i,\text{external}} = \frac{C}{2} |\vec{x}_i - \vec{x}_{i+1}|(yR - y^2)\sin(\beta_i)\vec{e}_x
-\end{equation}
+\begin{align}
+    \vec{G}_i &= C
+    \underbrace{|\vec{x}_i - \vec{x}_{i+1}|\sin(\beta_i)}_\text{proj. rod segment length}
+    \underbrace{(yR - y^2)}_\text{flow contribution}\vec{e}_x\\
+    \vec{F}_{i,\text{external}} &= \frac{1}{2}(\vec{G}_{i-1}+\vec{G}_i)
+\end{align}
 which is proportional to the length of the rod segment projected along the y-axis.
@@
 Here $\beta_i$ is the angle of the rod edge between $\vec{x}_{i+1}-\vec{x}_i$ and the x-axis.
-The flow profile for this particular setup is proportional to $(2yR-y^2)$ where $y$ is the vertical
-displacement measured from the end of the tube~\cite{Wolschin2021}
-(see Figure~\ref{fig:amir-bending-simulation}) and $R$ is the total size of the chamber along the
-y-axis.
-Since the bacterium does not reach the peak of the velocity profile, we can estimate the slope via a
-linear dependence on the pertruding segment along the y-axis~\cite{Wolschin2021}.
+The flow profile for this particular setup is proportional~\cite{Wolschin2021} to $(2yR-y^2)$ where
+$y=0.5(\vec{x}_i+\vec{x}_{i+1})_y-y_b$ is the vertical displacement measured from the end of the
+tube $y_b$ (see Figure~\ref{fig:amir-bending-simulation}) and $R$ is the total size of the chamber
+along the y-axis.
 In particular, this enforces the correct boundary values the flow profile and the resulting force
 $\vec{F}_{i,\text{external}}|_{y=0}=0$.
+In addition we confine vertices who belong to segments that are partially within the tube
+$(\vec{x}_i)_y\leq y_b$ via $(\vec{x}_{i+1})_x=x_0$ and also fix the value for the first vertex
+$\vec{x}_0$ such that the bacterium is attached to the bottom.
 
@@
     \caption{
-        \textbf{TODO rework caption}
-        Subfigure (A) shows a snapshot of the experiments performed by \cite{Amir2014_2}.
-        Division of the bacteria is inhibited and they are partially trapped within a mother machine
-        such that a controlled flow can be exerted on them.
+        (A) Microscopic image containing a single bacterium which is partially confined in a narrow
+        tube on the bottom and free at the top where a controlled flow can be exerted on them.
+        Division of the bacteria is inhibited.
         (B) A comparison of the initial and bent state show good visual agreement of our prediction
         with the experimental data.
-        Subfigures (C-G) show likelihood profiles for the individual parameters.
+        (C-G) Likelihood profiles for the individual parameters before and after performing model
+        reduction.
     }

### 2025-11-18 paper/main.tex
make clear what chamber is; switch order of sentences

@@
 by the experiments of Amir et. al~\cite{Amir2014,Amir2014_2}.
+The experimental setup is displayed in figure~\ref{fig:amir-bending-simulation} (A).
 They studied how a single \textit{E.Coli}, which is partially confined within a narrow tube, reacts
@@
 performing parameter estimations.
-The experimental setup is displayed in figure~\ref{fig:amir-bending-simulation} (A).
 To model this configuration, the lowermost vertex of the rod is fixed at the base of the simulation
@@
 tube $y_b$ (see Figure~\ref{fig:amir-bending-simulation}) and $R$ is the total size of the chamber
-along the y-axis.
+along the y-axis which is hidden in Figure~\ref{fig:amir-bending-simulation} (A).
 In particular, this enforces the correct boundary values the flow profile and the resulting force

### 2025-11-19 cr_mech_coli/crm_amir/main.py
perform two model reductions; pick new parameters; plot 3rd model

@@ -12,7 +12,7 @@ import multiprocessing as mp
 import argparse
 import itertools
 
-from cr_mech_coli.plotting import COLOR2, COLOR3, COLOR5
+from cr_mech_coli.plotting import COLOR1, COLOR3, COLOR5
 
 GREEN_COLOR = np.array([95, 231, 76])
 
@@ -637,11 +637,11 @@ def crm_amir_main():
     positions_data, iterations_data = obtain_data("out/crm_amir/", n_vertices=20)
 
     # Define globals for all optimizations
-    rod_rigidity = (0.0, 20.0, 250, "[µm/s²]")
-    drag_force = (0.0000, 0.001, 0.1, "[1/s²µm]")
+    rod_rigidity = (0.0, 20.0, 450, "[µm/s²]")
+    drag_force = (0.0000, 0.0001, 0.001, "[1/s²µm²]")
     damping = (0.000, 0.1, 0.1, "[1/s]")
-    growth_rate = (0.0, 0.01, 0.05, "[1/s]")
-    spring_tension = (10, 30.0, 60.0, "[1/s²]")
+    growth_rate = (0.0, 0.01, 0.04, "[1/s]")
+    spring_tension = (10, 30.0, 200.0, "[1/s²]")
 
     # Define which parameters should be optimized
     x0_bounds = {
@@ -656,6 +656,7 @@ def crm_amir_main():
         positions_data,
         iterations_data,
         pyargs,
+        output_dir="out/crm_amir/result-1",
     )
 
     x0_bounds_reduced = {
@@ -665,16 +666,32 @@ def crm_amir_main():
         "growth_rate": growth_rate,
         "spring_tension": spring_tension,
     }
-    set_params = {
-        "damping": 0.0,
-    }
+    set_params = {"damping": params1["damping"]}
     pfin2, popt2, costs2, samples2, params2 = compare_with_data(
         x0_bounds_reduced,
         positions_data,
         iterations_data,
         pyargs,
         set_params=set_params,
-        output_dir="out/crm_amir/result-reduced/",
+        output_dir="out/crm_amir/result-2/",
+    )
+
+    x0_bounds_3 = {
+        "rod_rigidity": rod_rigidity,
+        "drag_force": drag_force,
+        "damping": damping,
+        # "growth_rate": growth_rate,
+        "spring_tension": spring_tension,
+    }
+    # set_params_3 = {"spring_tension": params2["spring_tension"]}
[…]

### 2025-11-19 paper/main.tex
write introductory note for 2nd application; link to previous chapters

@@
 \subsection{Position Extraction}
+\label{subsec:position-extraction-algorithm}
 
@@
 \subsubsection{Imaging-based Comparison Algorithm}
+\label{subsec:imaging-based-comparison}
 The preceding algorithm is performant and effective but can not be applied in the event of a
@@
 We cover 3 distinct cases with increasing level of complexity.
-In the first study, we utilize results previously obtained by \cite{Amir2014,Amir2014_2} who
-investigated elastic deformations of \textit{E.Coli} that are partially trapped within a narrow
+In the first study, we utilize results previously obtained by Amir et. al~\cite{Amir2014,Amir2014_2}
+who investigated elastic deformations of \textit{E.Coli} that are partially trapped within a narrow
 tube and subject to a force resulting from a fluidic flow.
@@
 events.
-In the first part, we utilize the direct comparison algorithm while the latter requires usage of the
-imaging-based comparison algorithm.
+In the first part, we cover a time interval which does not contain division events and thus utilize
+the direct comparison algorithm while the second case contains division events and in turn requires
+usage of the imaging-based comparison algorithm.
 
@@
 \label{section:application-2}
+
+In this section we study a collection of growing, dividing and physically interacting
+\textit{E.Coli}.
+We use a series of microscopic images and corresponding segmentation masks for the individual
+bacteria as the underlying dataset which is partially displayed in
+Figure~\ref{fig:application-2-data-overview}.
+With the previously developed extraction algorithm (see
+section~\ref{subsec:position-extraction-algorithm}) we obtain the positions of our agents at all
+time points.
+These can be used as initial values and later within our parameter estimation routine.
+We initialize our computational model (see section~\ref{sec:computational-model}) with the obtained
+positions and optimize the parameters in order to find a set which best describes the observed
+results.
+The first case considers a time interval without cell-division while the second case does contain
+division events.
+Therefore, the first approach will optimize the direct comparison cost function (see
+section~\ref{subsec:direct-comparison}) while the latter approach has to account for division events
+and thus requires the imaging-based method (see section~\ref{subsec:imaging-based-comparison}).
+These two scenarios present an opportunity to apply the previously developed methods within a
+realistic context and validate our approach.
+
 To show the usefulness of our previously developed methods
@@
 \textbf{Explain data more clearly. Refer to figure.}
+
+\begin{figure}[H]
+    \centering
+    \begin{tikzonimage}[width=0.24\textwidth]
+        {../data/crm_divide/0003/images/001044.png}%
+        \node at (0.02, 0.98)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \end{tikzonimage}%
+    \hspace{0.01\textwidth}%
+    \begin{tikzonimage}[width=0.24\textwidth]
+        {../data/crm_divide/0003/images/001054.png}%
+        \node at (0.02, 0.98)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+    \end{tikzonimage}%
+    \hspace{0.01\textwidth}%
+    \begin{tikzonimage}[width=0.24\textwidth]
+        {../data/crm_divide/0003/images/001063.png}%
+        \node at (0.02, 0.98)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+    \end{tikzonimage}%
+    \hspace{0.01\textwidth}%
+    \begin{tikzonimage}[width=0.24\textwidth]
+        {../data/crm_divide/0003/images/001074.png}%
+        \node at (0.02, 0.98)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+    \end{tikzonimage}%
+    \caption{TODO insert segmentation masks here}
+    \label{fig:application-2-data-overview}
+\end{figure}
 

### 2025-11-19 paper/main.tex
restructure introductory text

@@
 research~\cite{Ghaffarizadeh2018,Cooper2020}, Immuno-Oncology~\cite{Karolak2021} or the modeling
-of mechanical intercellular connections in plant cells \textbf{TODO CITATION}.
+of mechanical intercellular connections in plant cells~\cite{DeRybel2014}.
 By treating cells as individual entities, these tools allow researchers to think in terms of
@@
 often not designed to be flexible but rather built for specific purposes thus limiting their
-generalizability \textbf{TODO CITATION}.
-Typically researchers estimate the parameter of their model using methods such as Maximum Likelihood
+generalizability~\cite{Pleyer2025Flexibility}.
+
+Typically researchers estimate the parameters of their model using methods such as Maximum Likelihood
 or Maximum a Posteriori~\cite{Gbor2015,Banga2008,Ashyraliyev2009}, check for parameter
@@
 techniques~\cite{Kreutz2013,Raue2014,Simpson2025}.
-
-In the context of \acp{abm}, however, it is often only shown that the generated results lead to the
+However, in the context of \acp{abm} this practice is almost never followed and instead only shown
+that the generated results lead to the
 desired phenomenon aligning with the observed biological reality.
-But this process does not actually estimate the necessary parameters directly \textbf{TODO CITATION}.
+This lack of systematic analysis of the involved parameters leaves a large gap of explainability of
+the underlying models.
 Consequently, it reduces the confidence in the constructed models and raises concerns with respect
 to the validity of the obtained results.
-In this work, we address these challenges using a general-purpose mechanical model for rod-shaped
-bacteria as an example.
-Our approach provides various methods which can be combined to systematically estimate the
-parameters of our model. We deliberately distinguish between conceptually separate components within
-this framework.
-
-Mechanical properties of rod-shaped bacteria have been studied in great detail
+
+In this work, we showcase how a systematic analysis of parameters can be performed using a
+general-purpose mechanical model for rod-shaped bacteria.
+Rod-shaped Bacteria present an interesting target system since their physical form provides more
+surface information compared to other forms such as spherical cells.
+Their mechanical properties have been studied in great detail
 experimentally~\cite{Chatterjee1988,Takeuchi2005,IWAI2002}.
@@
 permanent plastic deformations~\cite{Amir2014_2}.
-Rod-shaped Bacteria are an interesting target system since their physical form provides more surface
-information compared to other forms such as spherical cells.
-In addition, parameters such as growth rates have been found to vary across multiple
-cells~\cite{Koutsoumanis2013} which highlights the need to incorporate this heterogeneity on a
-single-cell level.
+In addition, parameters such as growth rates have been found to vary within a given cell
+population~\cite{Koutsoumanis2013} which highlights the need to incorporate this heterogeneity when
+constructing models on the single-cell level.
 
@@
 
-\textbf{
-    Was mir hier nicht gefällt, ist, dass du von Parameter schätzen bei ABMs kommend ziemlich
-    detailliert über Mechanik redest.
-    Das passt nicht.
-    Man erwartet mehr Infos über den Blueprint und ist verwirrt.
-    Worum geht es hier?
-}
-\textit{Ich habe den blueprint vollkommen entfernt. Er macht kein Sinn im MS. Der Rest sollte noch
-angepasst werden.}
-
-This work addresses some of these limitations.
+This work addresses some of these previously listed limitations.
 We use microscopic image time series as inputs to obtain information about the spatial configuration
 of the cells.
-This information can be used as initial values for a numerical simulation.
+This information can be used as initial values for a numerical simulation and also serves as ground
+truth.
 By varying the parameters of the model and comparing the results to the microscopic data, we can
@@
 \textit{B.Subtilis} and \textit{E.Coli}.
+We deliberately distinguish between conceptually separate components within this framework to allow
+for downstream reusability.
+The developed computational methods can be combined to systematically estimate the parameters of our
+model and are readily available as a Python package.
 

### 2025-11-20 paper/main.tex
move table from supplement to applications section
completely reformulate introduction

@@
 \section{Introduction}
-\acp{abm} have proven to be powerfull tools in the study of complex cellular
-systems~\cite{Pleyer2023}.
-They have been applied in various scenarios, such as cancer
-research~\cite{Ghaffarizadeh2018,Cooper2020}, Immuno-Oncology~\cite{Karolak2021} or the modeling
-of mechanical intercellular connections in plant cells~\cite{DeRybel2014}.
+
+% \textbf{Restructure Introduction with these bullet points}
+% \begin{itemize}
+%     \item Problem 1: Parameter estimation of \acp{abm}: how to do at all?
+%     \item Problem 2: Many \acp{abm} have very many parameters
+%     \item Problem 3: No \ac{abm} framework with support for flexible rod-shaped bacteria
+%     \item Literature Gap 1: noone really does parameter estimation for \acp{abm}
+%     \item Literature Gap 2: also no methods do this
+%     \item Literture  Gap 3: Some literature on flexible rods and some models but often not with
+%         growth, reactions etc.
+%     \item Why matter 1: No explainability without values of parameters.
+%     \item Why matter 2: How can we be sure to actually model biological reality?
+%     \item Why matter 3: Bacteria especially elongated are fundamental. Understanding them is
+%         important.
+%     \item What we do 1: Construct new mechanical model for rod-shaped bacteria
+%     \item What we do 2: Develop methods to compare numerical results with data (for previous model)
+%     \item What we do 3: Apply these methods in 3 scenarios covering different aspects
+%     \item Why my contribution significant 1: Show that parameter estimation can be done
+%     \item Why my contribution significant 2: Show how to treat heterogeneity and that we can map 1:1
+%         of model to cells
+% \end{itemize}
+
+\acp{abm} are powerful toosl to study complex cellular systems~\cite{Pleyer2023}.
 By treating cells as individual entities, these tools allow researchers to think in terms of
@@
 reality.
-
-Despite their popularity, \acp{abm} face multiple challenges in practise.
-The construction of a fully functional \ac{abm} simulation from the ground up constitutes an
-intricate and laborious process, necessitating a comprehensive array of skills. It is therefore
-common practise to reuse existing simulation tools, which, in turn, brings along a number of
-other problems.
-Even if the existing tool is designed to be generalizable, it often comes with a pre-defined set
-of parameters that need to be specified.
-Furthermore, as we have discussed in our previous work~\cite{Pleyer2023}, these tools were most
-often not designed to be flexible but rather built for specific purposes thus limiting their
-generalizability~\cite{Pleyer2025Flexibility}.
-
-Typically researchers estimate the parameters of their model using methods such as Maximum Likelihood
-or Maximum a Posteriori~\cite{Gbor2015,Banga2008,Ashyraliyev2009}, check for parameter
-identifiability using methods like profile-likelihood or apply model reduction
-techniques~\cite{Kreutz2013,Raue2014,Simpson2025}.
-However, in the context of \acp{abm} this practice is almost never followed and instead only shown
-that the generated results lead to the
-desired phenomenon aligning with the observed biological reality.
-This lack of systematic analysis of the involved parameters leaves a large gap of explainability of
-the underlying models.
-Consequently, it reduces the confidence in the constructed models and raises concerns with respect
-to the validity of the obtained results.
-
-In this work, we showcase how a systematic analysis of parameters can be performed using a
-general-purpose mechanical model for rod-shaped bacteria.
-Rod-shaped Bacteria present an interesting target system since their physical form provides more
-surface information compared to other forms such as spherical cells.
-Their mechanical properties have been studied in great detail
-experimentally~\cite{Chatterjee1988,Takeuchi2005,IWAI2002}.
-It was found that the MreB protein is the driving force in maintaining their distinct elongated
-shape~\cite{Ursell2014}, which plays an important role in the dynamics of
-growth~\cite{Billaudeau2017}, division~\cite{Harry2001} and their collective
-behaviour~\cite{vanGestel2015,Balagam2015}.
-Furthermore, it has been discovered that the rod can be bent thus introducing temporary elastic or
-permanent plastic deformations~\cite{Amir2014_2}.
-In addition, parameters such as growth rates have been found to vary within a given cell
-population~\cite{Koutsoumanis2013} which highlights the need to incorporate this heterogeneity when
-constructing models on the single-cell level.
-
-Computational models for rod-shaped bacteria allow researchers to systematically analyze their
-properties and infer emergent phenomena~\cite{Cho2007,Jnsson2005}.
-Existing tools can describe growth, division and many effects such as colony formation and
-expansion~\cite{Matyjaszkiewicz2017,Doumic2020} but they are limited in describing the
-aforementioned plastic or elastic deformations of the rod. Furthermore, the literature on how to
-estimate the parameters of such models directly from data is very sparse.
-
-This work addresses some of these previously listed limitations.
-We use microscopic image time series as inputs to obtain information about the spatial configuration
-of the cells.
-This information can be used as initial values for a numerical simulation and also serves as ground
-truth.
-By varying the parameters of the model and comparing the results to the microscopic data, we can
-estimate and quantify the parameters of our model.
-Our approach takes a general stance which incorporates effects that are observed within
-\textit{B.Subtilis} and \textit{E.Coli}.
-We deliberately distinguish between conceptually separate components within this framework to allow
-for downstream reusability.
-The developed computational methods can be combined to systematically estimate the parameters of our
-model and are readily available as a Python package.
+Their applications range from cancer research~\cite{Ghaffarizadeh2018,Cooper2020} over
+Immuno-Oncology~\cite{Karolak2021} to the modeling of mechanical intercellular connections in plant
+cells~\cite{DeRybel2014} which demonstrates their ability to capture complex emergent multicellular
+behavior.
+However, despite their popularity \acp{abm} face multiple practical challenges.
+
+One major problem is how to effectively perform parameter estimations for \acp{abm}.
+Classical approaches in systems biology—such as maximum
+likelihood~\cite{Kreutz2013,Raue2014} or maximum a posteriori estimation, identifiability analysis
+and model reduction~\cite{Simpson2025,Gbor2015,Banga2008,Ashyraliyev2009}.
+While some attempts have been made to properly calibrate the \ac{abm} at hand and estimate its
+parameters~\cite{An2016,Lima2021,Dancik2010,Thiele2014} these methods are not widely applied and are
+generally not used to explicitly capture any possible heterogeneity within the cellular population.
+
+Many \acp{abm} contain a substantial amount of parameters which poses another major problem
+related to the tuning of the model and can limit it's biology interpretability.
+Therefore, many studies only show that a suitably chosen set of parameters produces the desired
+biological behavior and asses how a subset of parameters can effect their results.
+However, this can leave important questions unanswered:
+Which combination of parameters drive the observed phenomena?
+Can these parameters be identified given a set of datapoints?
+Is there a more simple, i.e. reduced model that can describe the same phenomena?
+If \acp{abm} are not able to answer these questions, their results will significantly lack in
+explanatory power and ability to link to the biological problem at hand.
+In addition, biolgical \ac{abm} are typically designed with specific systems in mind and come with a
+predefined set of parameters which need to be specified and are not easily adaptable.
+This circumstance leads to problems when aiming to find a minimal set of identifiable parameters
+that can describe the desired phenomena.
+
+Within our approach we will utilize a new computational model of flexible, elongated bacteria which
+is a mechanical setup that most \acp{abm} are not capable of describing.
+So far, only a few computational studies have considered flexible rods~\cite{Duman2018,Joshi2019}
+but their model can not be transferred directly to more complex biolgical systems which might
+involve intracellular reactions, extracellular processes, growth and division.
+This poses another problem in the computational study of these systems.
+
+The aforementioned problems present scientific significant gaps.
+First, no model can be explained or interpreted without knowing the roles and values of its
+parameters or how they are constrained.
+Furthermore, models which are not properly calibrated can not guarantee to accurately represent the
+underlying biolgical reality beyond numerical coincidence.
+Finally, elongated bacteria are fundamental biolgical organisms which can be seen as a precursor to
+more complex shapes such as spiral bacteria.
+Obtaining a mechanistic understanding of their mechanical properties is essential to understand
+emergent phenomana such as the formation of biofilms~\cite{Li2025} or bacterial
+swarming~\cite{Starru2007}.
+
+In this work, we address these challenges and aim to further close the previously identified gaps.
+We construct a novel mechanical model for flexible, elongated bacteria that captures growth,
+division, elastic deformations of the rod and mechanical interactions between cells.
+We develop computational methods which utilize segmented microscopic images to quantitatively
+compare numerical results with biolgical data on a single-cell level.
+With these methods we can estimate the parameters of our model in three scenarios, each probing
+different aspects of our model with increasing level of complexity.
+
+Our approach demonstrates that parameter estimation of \acp{abm} is indeed feasible, even for
+complex models and when incorporating cellular heterogeneity.
+The developed methods fill a methodological gap and present a novel way of fitting individual-based
+models to data which can be readily adapted to various other scenarios.
+We present a novel mechanical model for flexible elongated bacteria which seamlessly integrates into
+the \texttt{cellular\_raza} framework, enabling researchers to extend and adapt it in future
+studies.
+In total, these contributions push for a more rigorous, interpretable and biolgically grounded
+development process for \acp{abm} which provide deeper mechanistic insights into the nature of
+cellular behaviors.
 
@@
 \textbf{TODO CITATION}.
+
+\begin{table}[H]
+    \centering
+    \begin{tabular}{l c c c c c c}
+        \toprule
+        % \multicolumn{6}{Growth Rates}\\
+        Method & Cell $0$ & Cell $1$ & Cell $2$ & Cell $3$ & Cell $4$ & Cell $5$\\
+        \midrule
+        Fit   & 0.001405 & 0.001080 & 0.002317 & 0.001512 & 0.002048 & 0.001672\\
+        \ac{abm} Direct & 0.001152 & 0.001410 & 0.001876 & 0.001683 & 0.003610 & 0.001520\\
+        \ac{abm} Imaging\\
+        % 0.0014057073 & 0.0010807540 & 0.0023177727 & 0.0015120675 & 0.0020488179 & 0.0016726263\\
+        % direct & 0.0009785162 & 0.0012641432 & 0.0021701728 & 0.0015695889 & 0.0036255661 & 0.0018419186\\
+        \bottomrule
+    \end{tabular}
+    \caption{TODO Update and explain these values. And also add units!}
+\end{table}
 
@@
 \end{figure}
-
-\begin{table}[H]
-    \centering
-    \begin{tabularx}{\textwidth}{l c c c c c c}
-        \toprule
-        % \multicolumn{6}{Growth Rates}\\
-        Method & Cell $0$ & Cell $1$ & Cell $2$ & Cell $3$ & Cell $4$ & Cell $5$\\
-        \midrule
-        \ac{abm} & 0.0011527990 & 0.0014106040 & 0.0018761827 & 0.0016834959 & 0.0036106023 & 0.0015209642\\
-        Direct   & 0.0014057073 & 0.0010807540 & 0.0023177727 & 0.0015120675 & 0.0020488179 & 0.0016726263\\
-        % 0.0014057073 & 0.0010807540 & 0.0023177727 & 0.0015120675 & 0.0020488179 & 0.0016726263\\
-        % direct & 0.0009785162 & 0.0012641432 & 0.0021701728 & 0.0015695889 & 0.0036255661 & 0.0018419186\\
-        \bottomrule
-    \end{tabularx}
-    \caption{TODO}
-\end{table}
 

### 2025-11-21 paper/main.tex
formatting
use new acronyms for E.coli and B.subtilis

@@
 \newacro{ib}[IB]{Individual-based}
+\newacro{ecoli}[\textit{E. coli}]{\textit{Escherichia coli}}
+\newacro{bsubtilis}[\textit{B. subtilis}]{\textit{Bacillus subtilis}}
 
@@
     The developed model serves as a generalized foundation to describe multiple species of
-    rod-shaped bacteria such as \textit{E.Coli} or \textit{B.Subtilis}.
+    rod-shaped bacteria such as \acl{ecoli} or \acl{bsubtilis}.
 \end{abstract}
@@
 
-While some rod-shaped bacteria grow by inserting new material at the ends, \textit{E.Coli} and
-\textit{B.Subtilis} utilize a method where new material is inserted along the cylindrical part of
+While some rod-shaped bacteria grow by inserting new material at the ends, \ac{ecoli} and
+\ac{bsubtilis} utilize a method where new material is inserted along the cylindrical part of
 the rod. The details of the division event (i.e. Z-Ring formation) are very intricate and are heavily
@@
 In the first study, we utilize results previously obtained by Amir et. al~\cite{Amir2014,Amir2014_2}
-who investigated elastic deformations of \textit{E.Coli} that are partially trapped within a narrow
+who investigated elastic deformations of \ac{ecoli} that are partially trapped within a narrow
 tube and subject to a force resulting from a fluidic flow.
@@
 %###################################################################################################
-\subsection{Application 1: Elastic deformations of \textit{E.Coli}}
+\subsection{Application 1: Elastic deformations of \acl{ecoli}}
 
@@
 The experimental setup is displayed in figure~\ref{fig:amir-bending-simulation} (A).
-They studied how a single \textit{E.Coli}, which is partially confined within a narrow tube, reacts
+They studied how a single \ac{ecoli}, which is partially confined within a narrow tube, reacts
 to a horizontal fluid flow that exerts a force onto the bacterium.
@@
 %###################################################################################################
-\subsection{Application 2: \textit{E.Coli.} growing on Agar}
+\subsection{Application 2: \acl{ecoli} growing on Agar}
 \label{section:application-2}
@@
 In this section we study a collection of growing, dividing and physically interacting
-\textit{E.Coli}.
+\ac{ecoli}.
 We use a series of microscopic images and corresponding segmentation masks for the individual
@@
 (see section~\ref{section:parameter-estimation}), we will use them to estimate the parameters of
-individually growing \textit{E.Coli} using a sequence of microscopic images.
+individually growing \ac{ecoli} using a sequence of microscopic images.
 The data we use~\cite{https://doi.org/10.3203/iwf/k-129} (see
@@
 
+\newpage
 \appendix
@@
         (3) & Differentiation &
-            \textit{B.subtilis} is known to differentiate into matrix-producing and
+            \ac{bsubtilis} is known to differentiate into matrix-producing and
             surfactin-producing cells \cite{vanGestel2015,Lpez2010}.\\

### 2025-11-21 paper/main.tex
rewrite introduction; put more emphasis on model and biology

@@
 
-% \textbf{Restructure Introduction with these bullet points}
 % \begin{itemize}
-%     \item Problem 1: Parameter estimation of \acp{abm}: how to do at all?
-%     \item Problem 2: Many \acp{abm} have very many parameters
-%     \item Problem 3: No \ac{abm} framework with support for flexible rod-shaped bacteria
-%     \item Literature Gap 1: noone really does parameter estimation for \acp{abm}
-%     \item Literature Gap 2: also no methods do this
-%     \item Literture  Gap 3: Some literature on flexible rods and some models but often not with
-%         growth, reactions etc.
-%     \item Why matter 1: No explainability without values of parameters.
+%     \item Problem 1: How to model flexibel rods and validate these models?
+%     \begin{itemize}
+%         \item Flexible rods so far only covered in soft-matter studies; rarely applied in direct
+%             biological contexts
+%         \item \acp{abm} are \textit{the} tool to use for individual-based modeling (combine with
+%             reactions, cell cycle, division, differentiation, etc.)
+%     \end{itemize}
+%     \item Problem 2: How to estimate parameters for individual-based models of cells/bacteria
+%     \item Problem 3: How to incorporate division and heterogeneity in parameter estimation?
+% 
+%     \item Literature Gap 1: No \acp{abm} support flexible elongated bacteria
+%     \item Literature Gap 2: Noone really does parameter estimations
+%     \item Literature Gap 3: Cellular heterogeneity sometimes accounted for in simulation but not
+%         in parameter estimation
+% 
+%     \item Why matter 1: No explainability without parameters
 %     \item Why matter 2: How can we be sure to actually model biological reality?
@@
 %         important.
+% 
 %     \item What we do 1: Construct new mechanical model for rod-shaped bacteria
-%     \item What we do 2: Develop methods to compare numerical results with data (for previous model)
+%     \item What we do 2: Develop methods to compare numerical results with data
 %     \item What we do 3: Apply these methods in 3 scenarios covering different aspects
-%     \item Why my contribution significant 1: Show that parameter estimation can be done
-%     \item Why my contribution significant 2: Show how to treat heterogeneity and that we can map 1:1
+% 
+%     \item Why contribution significant 1: Show that parameter estimation can be done
+%     \item Why contribution significant 2: Show how to treat heterogeneity and that we can map 1:1
 %         of model to cells
+%     \item Why contribution significant 3: New model for flexible rods which integrates in
+%         \texttt{cellular\_raza} and can thus be readily reused by researchers
 % \end{itemize}
 
-\acp{abm} are powerful toosl to study complex cellular systems~\cite{Pleyer2023}.
-By treating cells as individual entities, these tools allow researchers to think in terms of
-cellular properties and processes thus allowing for a better understanding of the biological
-reality.
-Their applications range from cancer research~\cite{Ghaffarizadeh2018,Cooper2020} over
-Immuno-Oncology~\cite{Karolak2021} to the modeling of mechanical intercellular connections in plant
-cells~\cite{DeRybel2014} which demonstrates their ability to capture complex emergent multicellular
-behavior.
-However, despite their popularity \acp{abm} face multiple practical challenges.
-
-One major problem is how to effectively perform parameter estimations for \acp{abm}.
+Bacteria come in various shapes~\cite{Young2006,Young2007}, ranging from simple spheres to complex
+spirals and helices.
+Elongated bacteria form an important subclass that includes some of the most studied model
+organisms such as \ac{ecoli} or \ac{bsubtilis}.
+Their filamentous structure underlies distinct mechanical~\cite{Daniel2003} and physiological
+constraints~\cite{Jones2001}, thus influencing other cellular aspects such as
+growth~\cite{Chang2014} and division~\cite{Li2007}.
+It also facilitates processes such as swarming~\cite{Harshey1994} or biofilm
+formation~\cite{Duvernoy2018}.
+
+Computational methods can allow for a systematic analysis of the underlying biological reality.
+To this end, \acp{abm} provide a natural way of mapping cellular processes to computational
+routines.
+By treating cells as individual entities with their own state, these tools enable researchers to
+think in terms of cellular properties and processes, thereby allowing for a mechanistic
+understanding of the system.
+This bottom-up perspective has transformed the analysis of complex cellular
+systems~\cite{Pleyer2023} over a wide range of
+topics~\cite{Ghaffarizadeh2018,Cooper2020,Karolak2021,DeRybel2014}.
+\acp{abm} are generally capable of incorporating gene regulatory networks, cell cycle, division,
+differentiation, transport of extracellular compounds and many other processes which makes them the
+ideal tool to study the individual behavior of cells.
+
+Within our approach we develop a new computational model for flexible, elongated bacteria which
+is a mechanical setup that most \acp{abm} are not capable of describing.
+So far, only a few computational studies have considered flexible rods~\cite{Duman2018,Joshi2019},
+but their models cannot be transferred directly to more complex biological systems which might
+involve intracellular reactions, extracellular compounds, growth and division.
+Only a few \acp{abm} support even the less complex cylindrical
+shapes~\cite{Kang2014,breitwieser_biodynamo_2022,Gorochowski2012} but none are able to capture
+flexible filamentous bacteria.
+This poses a problem in the computational study of these systems.
+
+Although \acp{abm} seem destined to be applied to the modeling of individual elongated bacteria,
+this potential has not been fully realized yet.
+Another major problem is how to perform parameter estimations for \acp{abm}.
 Classical approaches in systems biology—such as maximum
@@
 generally not used to explicitly capture any possible heterogeneity within the cellular population.
-
-Many \acp{abm} contain a substantial amount of parameters which poses another major problem
-related to the tuning of the model and can limit it's biology interpretability.
-Therefore, many studies only show that a suitably chosen set of parameters produces the desired
-biological behavior and asses how a subset of parameters can effect their results.
 However, this can leave important questions unanswered:
@@
 Can these parameters be identified given a set of datapoints?
-Is there a more simple, i.e. reduced model that can describe the same phenomena?
+Is there a simpler, i.e. reduced model that can describe the same phenomena?
 If \acp{abm} are not able to answer these questions, their results will significantly lack in
-explanatory power and ability to link to the biological problem at hand.
-In addition, biolgical \ac{abm} are typically designed with specific systems in mind and come with a
-predefined set of parameters which need to be specified and are not easily adaptable.
-This circumstance leads to problems when aiming to find a minimal set of identifiable parameters
-that can describe the desired phenomena.
-
-Within our approach we will utilize a new computational model of flexible, elongated bacteria which
-is a mechanical setup that most \acp{abm} are not capable of describing.
-So far, only a few computational studies have considered flexible rods~\cite{Duman2018,Joshi2019}
-but their model can not be transferred directly to more complex biolgical systems which might
-involve intracellular reactions, extracellular processes, growth and division.
-This poses another problem in the computational study of these systems.
-
-The aforementioned problems present scientific significant gaps.
+explanatory power and ability to link to the specific problem at hand.
+
+The aforementioned problems present significant gaps in the scientific literature.
 First, no model can be explained or interpreted without knowing the roles and values of its
 parameters or how they are constrained.
-Furthermore, models which are not properly calibrated can not guarantee to accurately represent the
-underlying biolgical reality beyond numerical coincidence.
-Finally, elongated bacteria are fundamental biolgical organisms which can be seen as a precursor to
+Furthermore, models which are not properly calibrated cannot guarantee to accurately represent the
+underlying biological reality beyond numerical coincidence.
+Finally, elongated bacteria are fundamental organisms which can be seen as a precursor to
 more complex shapes such as spiral bacteria.
 Obtaining a mechanistic understanding of their mechanical properties is essential to understand
-emergent phenomana such as the formation of biofilms~\cite{Li2025} or bacterial
+emergent phenomena such as the formation of biofilms~\cite{Li2025} or bacterial
 swarming~\cite{Starru2007}.
@@
 We develop computational methods which utilize segmented microscopic images to quantitatively
-compare numerical results with biolgical data on a single-cell level.
+compare numerical results with biological data on a single-cell level.
 With these methods we can estimate the parameters of our model in three scenarios, each probing
-different aspects of our model with increasing level of complexity.
-
-Our approach demonstrates that parameter estimation of \acp{abm} is indeed feasible, even for
-complex models and when incorporating cellular heterogeneity.
+different aspects of our model with increasing levels of complexity.
+
+Our approach demonstrates that parameter estimation of \acp{abm} is indeed feasible, even in
+the case of complex models such as single-cell modeling of flexible elongated bacteria and when
+incorporating cellular heterogeneity.
 The developed methods fill a methodological gap and present a novel way of fitting individual-based

### 2025-11-21 paper/main.tex
clean up some bullet points
rewrite overview of package

@@
 This assumption alone is enough to enable a cascade of other components.
-Using the spatial representation, we provide visualization methods that create 3D models from the
-given vertices which can then either be rendered or projected to a 2D plane.
-The data extraction component determines the vertices from given cell masks which
-correspond to microscopic images.
+Using the spatial representation, we provide visualization methods that employ a 3D rendering
+engine and can produce cell masks or rendered images from a given configuration.
+The data extraction component determines the vertices from given segmented microscopic images.
 This allows us in particular to turn any sequence of images (movie) into a sequence of bacterial
-positions.
+positions and is a key feature of the parameter estimation methods.
 The dynamics of the vertices are determined by a computational model which is specific
 to the respective use-case and should be defined by the researcher.
-It merely needs to utilize the aforementioned spatial representation of bacterial rods as a
+It merely requires the aforementioned assumption of representing bacterial rods as a
 collection of vertices.
@@
 Notably, the visualization and data extraction components do not require any knowledge about the
-computational model, thus allowing us to reuse their functionality across multiple models.
-
-Constructing a computational model requires knowledge about the target system and thus allows us to
-interpret the results gathered by the provided data extraction methods.
-These results can then further be used to restrict or determine the parameters of our model.
+computational model, thus allowing us to reuse their functionality across differing models.
+
+By utilizing the provided data extraction methods, we can infer properties of the bacteria that aid
+us in constructing a suitable computational model.
+However, the construction of any mechanistic model also requires prior konwledge which can be used
+to inform our analysis methods.
 For example, by extracing the positions of bacteria across a time-series of microscopic images, we
@@
 effectively guiding our modeling steps.
+As we will show later, extracted data in the form of cellular positions or segmentation masks can
+also be used for comparing computational results with biological ground truth.
 
@@
 They allow researchers to obtain feedback quickly and improve iteration times when constructing new
-models. Furthermore, we utilize them in our more advanced parameter estimation methods (see
-section~\ref{section:parameter-estimation}) where they enable us to effectively compare
-two bacterial configurations across division events.
-In the future, we are planning to build on the current state and generate realistic microscopic
-images and time-series.
+models.
+Furthermore, we utilize them in our more advanced imaging-based comparison methods (see
+section~\ref{subsec:imaging-based-comparison}) where they enable us to effectively compare
+two bacterial configurations while accounting for division events.
+In the future, we are planning to extend the the rendering capabilities of our approach and
+generate realistic microscopic images and time-series.
 Together with the parameter estimation methods, they can then be used to provide large amounts of
@@
 
-\textbf{
-    Du musst den Bogen hinbekommen von dem allgemeinen Problem des Parameterschätzen bei ABMs zu dem
-    spezifischen Problem, welches du hier verhandelst.
-}
-
 \begin{figure}[H]
     \centering
-    % \includegraphics[width=0.8\textwidth]{overview.pdf}
     \includestandalone[width=0.8\textwidth]{figures/concept/overview}
     \caption{
-        The project can be divided into subcomponents (yellow, red) that interact with and depend on
+        The project can be divided into subcomponents that interact with and depend on
         each other.
-        Yellow blocks are provided by this work and red blocks require additional input such as
-        modeling assumptions or data.
-        Blue blocks indicate possible future extensions of this project.
         Every component implicitly builds on the spatial representation of rod-shaped bacteria as
-        vertices.
+        vertices (see section~\ref{subsec:spatial-representation}).
+        Red boxes highlight the key scientific drivers of this study while yellow boxes indicate
+        auxiliary methods.
+        The blue box "Data Generation" displays a extension of this approach which is under active
+        development.
     }
@@
 \label{sec:computational-model}
-
-@Christian
-\begin{itemize}
-    \item shortened everything
-    \item leave out parts which are not directly used by the model
-    \item move biological context to respective subsections to link more directly to model
-    \item remove duplicates and useless text
-    \item move heterogeneity considerations from table in main text
-\end{itemize}
 

### 2025-11-21 paper/main.tex
cleanup

@@
 
-%%% Add PDF metadata to help others organize their library
-%%% Once the PDF is generated, you can check the metadata with
-%%% $ pdfinfo template.pdf
-\hypersetup{
-pdftitle={% TODO
-},
-pdfsubject={q-bio.NC, q-bio.QM},
-pdfauthor={Jonas Pleyer, Christian Fleck},
-pdfkeywords={},
-}
-
-% Change numbering of equations
-% \numberwithin{equation}{section}
-
-% MAKE TITLES IN THEOREMS BOLD
-\makeatletter
-\def\th@plain{%
-  \thm@notefont{}% same as heading font
-  \itshape % body font
-}
-\def\th@definition{%
-  \thm@notefont{}% same as heading font
-  \normalfont % body font
-}
-\makeatother
-
 \begin{document}
@@
 
-\begin{itemize}
-    \item Adapt introduction; \acp{abm} not focus of paper, but rather tool to do the job
-    \begin{itemize}
-        \item model division, growth of bacteria
-        \item \acp{abm} are good for this tool
-    \end{itemize}
-    \item Exchange $\sigma \rightarrow R_1+R_2$
-    \item Equation 8 make double norm $||\vec{F}||$
-    \item Remove auxiliary variable $\vec{F}$ only use $\vec{\nabla}V$
-    \item Table 2 in supplement? (probably not)
-    \item numerical implementation verschieben?
-    \item numerical implementation: waht do we need to simulate this?
-    \begin{itemize}
-        \item \acp{abm} are common tools to realize this
-        \item cellular\_raza is THE tool to do it
-    \end{itemize}
-    \item Methodology: single-cell parameter estimation as motivation for techniques
-    \item Which parameters are global which are individual? Discuss this in general; not in detail
-    \item depends on data, number of parameters (computational efficiency)
-    \item page 12; yes this is Discussion
-    \item page 12: maybe extend the paragraph wrt comparison in intervals
-\end{itemize}
+% \begin{itemize}
+%     \item Adapt introduction; \acp{abm} not focus of paper, but rather tool to do the job
+%     \begin{itemize}
+%         \item model division, growth of bacteria
+%         \item \acp{abm} are good for this tool
+%     \end{itemize}
+%     \item Exchange $\sigma \rightarrow R_1+R_2$
+%     \item Equation 8 make double norm $||\vec{F}||$
+%     \item Remove auxiliary variable $\vec{F}$ only use $\vec{\nabla}V$
+%     \item Table 2 in supplement? (probably not)
+%     \item numerical implementation verschieben?
+%     \item numerical implementation: waht do we need to simulate this?
+%     \begin{itemize}
+%         \item \acp{abm} are common tools to realize this
+%         \item cellular\_raza is THE tool to do it
+%     \end{itemize}
+%     \item Methodology: single-cell parameter estimation as motivation for techniques
+%     \item Which parameters are global which are individual? Discuss this in general; not in detail
+%     \item depends on data, number of parameters (computational efficiency)
+%     \item page 12; yes this is Discussion
+%     \item page 12: maybe extend the paragraph wrt comparison in intervals
+% \end{itemize}
 

### 2025-11-21 src/config.rs
correctly convert to pydict

@@ -344,8 +344,8 @@ impl AgentSettings {
             (
                 "spring_length_threshold_setter",
                 spring_length_threshold_setter
-                    .clone()
-                    .into_pyobject_or_pyerr(py)?
+                    .borrow(py)
+                    .to_pydict(py)?
                     .into_any()
                     .unbind(),
             ),

### 2025-11-21 cr_mech_coli/simulation.py
export both setters

@@ -38,4 +38,6 @@ from .cr_mech_coli import (
     MiePotentialF32,
     PhysicalInteraction,
     StorageOption,
+    GrowthRateSetter,
+    SpringLengthThresholdSetter,
 )

### 2025-11-21 cr_mech_coli/crm_divide.py
set spring length threshold for next division to infinity; thus disabling any future divisions

@@ -291,6 +291,7 @@ def predict(
             growth_rate=growth_rate,
             growth_rate_setter={"g1": g1, "g2": g2},
             spring_length_threshold=spring_length_threshold,
+            spring_length_threshold_setter={"l1": np.inf, "l2": np.inf},
             neighbor_reduction=None,
         )
         for spring_length_threshold, pos, growth_rate, (g1, g2) in zip(

### 2025-11-21 paper/main.tex
cleanup

@@
 Table~\ref{table:simulation-aspects}.
-% For a collection of other known aspects which could be considered, we refer to the supplementary
-% section~\ref{section:discussion}, and~\ref{section:supplement-more-simulation-aspects}.
 
@@
 
-\textbf{TODO mention this in sections below}
-\begin{itemize}
-    \item plastic/elastic deformations
-    \item non-uniform thickness of the rod
-    \item polar interactions \cite{Duvernoy2018}
-    \item other dynamics
-\end{itemize}
-
-% \begin{enumerate}
-%     \item represent bacteria as polygon with vertices $\vec{x}_i$
-%     \item Assumes that these vertices are not fundamental but rather a discretization of the actual
-%         problem
-% \end{enumerate}
-
 \begin{figure}[H]
@@
         upper bound $\beta=2.5U_0$ and cutoff $\zeta=3(R_1+R_2)$.
-        \textbf{TODO insert more (possibly schematic depictions of how the model works
-        (Interactions, Cycle))}
     }
@@
 % but for sake of simplicity we chose to model the bacterium as a bending beam.
-% \textbf{[", thus utilizing the simplest solution which can still produce the desired effects." Es ist unklar was die "desired effects" sind. ]}\\
 Using our discretization $\vec{x}_i$, we connect the vertices by springs of length $l$.
@@
 % springs of lengths $l$.
-% \textbf{[Das klingt nach einer Wiederholung von Spatial Representation. Du erklärst ziemlich umständlich. Konzentriere dich auf das wesentliche und überlege, was du eigentlich sagen möchtest.]}\\
 We use the notation $\vec{w}_i:=\vec{x}_i-\vec{x}_{i-1}$ with which the resulting force acting
@@
 
-\textbf{@Christian Can you find a more elegant notation than this equation?}\\
 where the first case ensures that the norm is bound and the second case ensures the cutoff.
@@
 \cite{Takeuchi2005}.
-
-\textbf{TODO adapt naming to appliations section}
 
@@
 parameters within our model.
-In order to reuse this model, we also provide the option to assign growth rates by sampling from a
-normal distribution with user-defined parameters.\textbf{@Christian Supplement?}
 Every other intracellular parameter is copied identically to the daughter cells.
@@
 \subsection{Numerical Implementation}
-\textbf{TODO move this somewhere else (maybe parameter estimation)?}
-
 We implemented the described model in our recently developed \ac{abm} framework

### 2025-11-21 paper/main.tex
rewrite short subsection on numerical implementation

@@
 \texttt{cellular\_raza}~\cite{Pleyer2025}.
-By design \texttt{cellular\_raza} does not assume any particular cellular representation, thus makes
-it flexible and easy to adapt the simulation to particular needs.
-Its design allows us to experiment with different model implementations while retaining fast
-execution times.
-\textbf{[Der Zusammenhang zwischen der Flexibilität und Performance ist unklar.]}
-\textit{Oft muss man das eine für das andere opfern. Hier nicht/kaum.}
-
-Furthermore, it is an \ac{abm} which is able to natively track cells through the course of a
-numerical simulation and also keeps track of mother/daughter relationships.
-\textbf{[Das scheint hier trivial zu sein.]}
-\textit{Machen aber viele (zB. PhysiCell) einfach nicht. (Mother-Daughter relation)}
+By design \texttt{cellular\_raza} does not assume any particular cellular representation, which
+makes it flexible and easy to adapt the simulation to particular needs.
+Its design allows us to experiment with different model implementations and facilitates the
+explorative capabilities of researchers.
+It is furthermore able to seamlessly integrate into Python code-bases thus allowing us to utilize
+methods from other popular software libraries such as \texttt{numpy} and \texttt{scipy}.
+Furthermore, it is an \ac{abm} which is able to natively track mother-daughter relations of cells
+over the course of a numerical simulation which is an integral component of our imaging-based
+comparison algorithm (see section~\ref{subsec:imaging-based-comparison}).
 

### 2025-11-21 paper/main.tex
comment out supplementary material on additional simulation aspects

@@
 %###################################################################################################
-\subsection{More Simulation Aspects (TODO)}
-\label{section:supplement-more-simulation-aspects}
-%TODO insert this into the discussion later
-\begin{table}[H]
-    \centering
-    \def\arraystretch{1.3}
-    \begin{tabularx}{\textwidth}{c l X}
-        &\textbf{Aspect} & \textbf{Description}\\
-        \toprule
-        &\textbf{(C) Cellular}\\
-        \midrule
-        (1) & Rod-Shaped Mechanics &
-            Rod-shaped bacteria are flexible rods which are able to freely move around (off-lattice
-            approach) \cite{Takeuchi2005,Ursell2014,Amir2014_2}.\\
-        (2) & Growth &
-            Cells grow exponentially by inserting new material either along the circular part of the
-            rod or at the tip~\cite{Robert2014,Takeuchi2005}.\\
-        (3) & Differentiation &
-            \ac{bsubtilis} is known to differentiate into matrix-producing and
-            surfactin-producing cells \cite{vanGestel2015,Lpez2010}.\\
-        (4) & Division &
-            The formation and continuation of van Gogh bundles is driven by cell division
-            \cite{vanGestel2015}.
-            Bacteria can form multilayers during their growth phase \cite{Duvernoy2018}.\\
-        (5) & Variable Parameters &
-            Parameters for individual cells are not fixed values but rather taken from a
-            distribution \cite{Koutsoumanis2013}.\\
-        &\textbf{(CC) Cell-Cell Interactions}\\
-        \midrule
-        (6) & Adhesion &
-            Bacteria adhere to each other at longer distances and attach when in close contact
-            \cite{Verwey1947,Trejo2013}.
-            The interaction between rods can be polarized \cite{Duvernoy2018}.\\
-        (7) & Friction &
-            Friction between cells \cite{Grant2014} can be asymmetrical \cite{Doumic2020}.\\
-        &\textbf{(DC) Domain-Cell Interactions}\\
-        \midrule
-        (8) & External Forces &
-            Bacteria tend to stick to surfaces \cite{vanLoosdrecht1989}.
-            The extracellular gel exerts a force onto the cells \cite{Grant2014}.\\
-        (9) & Extracellular Reactions &
-            Bacteria can take up nutrients or possible secrete/take up signalling molecules
-            \cite{Li2025}.\\
-        \bottomrule
-    \end{tabularx}
-    \label{table:simulation-aspects-supplement}
-    \caption{
-        This table extends table~\ref{table:simulation-aspects} with additional observations.
-        It has to be noted that this table includes more specific behaviour that might not be
-        generalizable across all types of rod-shaped bacteria.
-    }
-\end{table}
-
-\begin{itemize}
-    \item new class of aspects: (DC) Domain-Cell Interactions
-    \item List which aspects are here additionally
-    \item discuss which of them might need specific models and can not be paired with others
-    \item which can be bundled together without problems?
-\end{itemize}
+\subsection{How to use obtained heterogenous results}
+In order to reuse this model, we also provide the option to assign growth rates by sampling from a
+normal distribution with user-defined parameters.
+
+%###################################################################################################
+% \subsection{More Simulation Aspects (TODO)}
+% \label{section:supplement-more-simulation-aspects}
+% %TODO insert this into the discussion later
+% \begin{table}[H]
+%     \centering
+%     \def\arraystretch{1.3}
+%     \begin{tabularx}{\textwidth}{c l X}
+%         &\textbf{Aspect} & \textbf{Description}\\
+%         \toprule
+%         &\textbf{(C) Cellular}\\
+%         \midrule
+%         (1) & Rod-Shaped Mechanics &
+%             Rod-shaped bacteria are flexible rods which are able to freely move around (off-lattice
+%             approach) \cite{Takeuchi2005,Ursell2014,Amir2014_2}.\\
+%         (2) & Growth &
+%             Cells grow exponentially by inserting new material either along the circular part of the
+%             rod or at the tip~\cite{Robert2014,Takeuchi2005}.\\
+%         (3) & Differentiation &
+%             \ac{bsubtilis} is known to differentiate into matrix-producing and
+%             surfactin-producing cells \cite{vanGestel2015,Lpez2010}.\\
+%         (4) & Division &
+%             The formation and continuation of van Gogh bundles is driven by cell division
+%             \cite{vanGestel2015}.
+%             Bacteria can form multilayers during their growth phase \cite{Duvernoy2018}.\\
+%         (5) & Variable Parameters &
+%             Parameters for individual cells are not fixed values but rather taken from a
+%             distribution \cite{Koutsoumanis2013}.\\
+%         &\textbf{(CC) Cell-Cell Interactions}\\
+%         \midrule
+%         (6) & Adhesion &
+%             Bacteria adhere to each other at longer distances and attach when in close contact
+%             \cite{Verwey1947,Trejo2013}.
+%             The interaction between rods can be polarized \cite{Duvernoy2018}.\\
+%         (7) & Friction &
+%             Friction between cells \cite{Grant2014} can be asymmetrical \cite{Doumic2020}.\\
+%         &\textbf{(DC) Domain-Cell Interactions}\\
+%         \midrule
+%         (8) & External Forces &
+%             Bacteria tend to stick to surfaces \cite{vanLoosdrecht1989}.
+%             The extracellular gel exerts a force onto the cells \cite{Grant2014}.\\
+%         (9) & Extracellular Reactions &
+%             Bacteria can take up nutrients or possible secrete/take up signalling molecules
+%             \cite{Li2025}.\\
+%         \bottomrule
+%     \end{tabularx}
+%     \label{table:simulation-aspects-supplement}
+%     \caption{
+%         This table extends table~\ref{table:simulation-aspects} with additional observations.
+%         It has to be noted that this table includes more specific behaviour that might not be
+%         generalizable across all types of rod-shaped bacteria.
+%     }
+% \end{table}
+% 
+% \begin{itemize}
+%     \item new class of aspects: (DC) Domain-Cell Interactions
+%     \item List which aspects are here additionally
+%     \item discuss which of them might need specific models and can not be paired with others
+%     \item which can be bundled together without problems?
+% \end{itemize}
 

### 2025-11-21 src/cell_container.rs
fix test

@@ -415,6 +415,10 @@ fn cell_container_de_serialize() {
                 std: 0.0,
             },
             spring_length_threshold: 2.0,
+            spring_length_threshold_setter: SpringLengthThresholdSetter::NormalDistr {
+                mean: 2.0,
+                std: 0.0,
+            },
             neighbor_reduction: None,
         })
         .collect();

### 2025-11-21 paper/main.tex
reference figure in bullet points
move text upfron to methods section and discuss in more detail how number of parameters grows

@@
 
-\textbf{TODO Talk about how to fit individual bacteria and why it is important.}
+The developed methods are independent of how we assign parameters within our model.
+This means we can opt to incorporate intracellular variability by specifying parameters on a
+per-cell basis.
+Or we can employ shared parameters between all agents for processes which can be expected to behave
+identical within margin of error.
+Furthermore, mixed approaches where some parameters are assigned globally and others individually
+can be realized as well.
+
+For the presented model a list of parameters can be seen in table~\ref{table:list-of-parameters} but
+the following calculations are equally true for any model with $N_p$ number of parameters.
+Consider the case where we filter out a subset $N_g\leq N_p$ of parameters which will be treated
+globally, thus with equal numerical value for every cell.
+This means that $N_p - N_g$ parameters will be assigned individually for each of $N_c$ number of
+cells.
+The total amount of parameters in the system is then given by
+\begin{equation}
+    N_\text{tot} = N_g + N_c (N_p - N_g).
+\end{equation}
+It is clear, that we should avoid, to needlessly treat parameters non-globally since this increases
+the size of the parameter space considerably.
+The volume of parameter space which needs to be sampled by optimization schemes scales exponentially
+with the dimension ie. the total number of parameters $N_\text{tot}$.
+Notice that the presented approach reflects our current work but can easily be adapted to other
+scenarios where different strategies of parameters assignment are desired, e.g. when considering
+two distinct species of cells which should be displayed by 2 parameters in total regardless of the
+number of cells.
 
@@
 parameters on a per-cell basis.
-We have previously presented a list of $N_p$ parameters (see Table~\ref{table:list-of-parameters})
-which will either be assigned individually or globally.
-We filter out a subset $N_g\leq N_p$ of global parameters, treating the remaining ones individually.
-Given $N_c$ cells, the total amount of parameters to be estimated is:
-\begin{equation}
-    N_{p,\text{tot}} = N_g + N_c (N_p - N_g).
-\end{equation}
-It is clear, that we should avoid, to needlessly treat parameters non-globally since this increases
-the size of the parameter space considerably.
-The volume that needs to be sampled to optimize within a given parameter space scales exponentially
-with the dimension ie. the total number of parameters $N_{p,\text{tot}}$.
-\textbf{[Das ist leider komplett unverständlich.
-Ich glaube, ich weiss schon, was du sagen möchtest, aber das muss umformuliert werden.
-Stichworte sind hier: Cell Variability, Heterogeneity, latent variables.
-Parameter per cell and individual parameters versteht auch niemand den Unterschied.]}\\
-\textit{Ich habe versucht, es klarer zu machen. Bitte dazu Feedback!}\\
-\textbf{Es muss noch im Methoden Teil mehr erwähnt werden.}
 
@@
 \begin{itemize}
-    \item Quadratic Scaling comes from more dense simulation
+    \item Quadratic Scaling (see figure~\ref{fig:timings-crm_divide}) comes from more dense
+        simulation
     \item Overall number of agents which can be simulated is satisfactory for current purposes

### 2025-11-21 paper/main.tex
extend description of visualization

@@
 A render of this result can be seen in figure~\ref{fig:progression-image-generation}.
-Since we use a 3D rendering engine for our visualization, the resulting image can have
+Since we use a 3D rendering engine for our visualization, the resulting image can display
 overlaps of bacteria which are placed at higher or lower values along the z-direction.
-To generate masks, we assign unique colours to every mesh of the respective cellular agent, ensuring
-no two agents will use identical colours. \textbf{TODO semantic segmentation Citation}
+To generate masks, we use a pixel-based (in-image) encoding scheme~\cite{JianboShi2000}, thus
+assigning unique colours to every mesh of the respective cellular agent, ensuring no two agents
+will use identical colours.
 We then use a parallel projection and remove any existing light sources in order to obtain a solid

### 2025-11-21 paper/main.tex
move around some stuff

@@
 \section{Methodology}
+\label{sec:methodology}
 These sections introduce methods which can be combined in order to estimate the parameters of our
@@
 \label{subsec:imaging-based-comparison}
-The preceding algorithm is performant and effective but can not be applied in the event of a
-division process.
-Due to possibly differing division times, a cell which is present in the simulation might not yet
-have a corresponding cell in the experimental data at this time point yet (or vice-versa) (see
-Figure~\ref{fig:mask-difference-metric} (E)).
-In order to still be able to compare these datapoints, we require additional methods which can be
-used for comparisons across division events.
-\textbf{[Es bleibt unklar, was du hier verhandeln möchtest.]}
-\textit{Jetzt klarer?}
-
+The preceding algorithm is performant and effective but can not be applied in the case when a
+division event splits apart the bacterium since this would involve the comparison of two sets of
+vertices with only one.
+This configuration is displayed in figure~\ref{fig:mask-difference-metric} (E) where two distinct
+bacterial configurations originate from the same initial state but contain different amounts of
+bacteria.
+This mismatch can happen, e.g. due to non-matching division times.
+Thus a cell which is present in the simulation might not yet have a corresponding cell in the
+experimental data at this time point yet (or vice-versa).
+To explain why the previous developed direct comparison algorithm can not be reused on confined
+intervals which do not include division events and for further details, we refer the interested
+reader to the supplementary
+material~\ref{sec:supplement-imaging-based-comparison-algorithm-details}.
+
+In order to be able to compare the differing cellular configurations, we require additional
+methods which can be used in time intervals containing cellular divison events.
 To this end we use the numerical results of the \ac{abm} simulation and our visualization
 algorithms to generate cell-masks (see section~\ref{subsection:visualization}).
-Each pixel has a unique color value which is directly related to one particular cell or background
-(black).
-We compare the generated masks to the experimentally obtained ones pixel by pixel.
-Whenever two colors are not matching, a penalty of $1$ is assigned.
+Since each pixel has a unique color value which is directly related to one particular cell or
+background (black), we compare the generated masks to the experimentally obtained ones pixel by
+pixel.
+Whenever two colors are not matching, a penalty is assigned for each pixel.
 The total cost is the sum of all penalties.
@@
 
-Future extensions should introduce a time-dependent parent penalty $p=p(t)$ which grows with the age
-of the daughter cells.
-Furthermore, the developed algorithm relies on the comparison of two images and can as such make use
-of existing machine-learning workflows that construct cost functions for such
-scenarios~\cite{Dice1945}.\\
-\textbf{@Christian are these two sentences Discussion?}
-
 Although this optimization scheme was driven by technical challenges, it is also biologically
@@
 cells begins and the one of the mother ends.
-
-\textbf{Put this in Supplement and discuss it further. Still link to it here.}\\
-In a different approach, one might be tempted to utilize the previously constructed direct
-comparison algorithm for time intervals that do not contain any division events.
-But this attempt would quickly lead to multiple problems.
-It will not be able to deal with the uncertainties of the timing of the division event.
-As the number of investigated cells grows, more division events will occurr even in a fixed
-amount of time thus greatly decreasing or completely vanishing the effective interval at which no
-division event is observed.
-The latter situation is equivalent to two microscopic images as input data between which a division
-event has occured.
 
@@
 \section{Discussion}
+
+\textbf{TODO decide if we mention these}
+\begin{itemize}
+    \item plastic/elastic deformations
+    \item non-uniform thickness of the rod
+    \item polar interactions \cite{Duvernoy2018}
+    \item other dynamics
+\end{itemize}
+
+Future extensions should introduce a time-dependent parent penalty $p=p(t)$ which grows with the age
+of the daughter cells.
+Furthermore, the developed algorithm relies on the comparison of two images and can as such make use
+of existing machine-learning workflows that construct cost functions for such
+scenarios~\cite{Dice1945}.\\
+\textbf{@Christian are these two sentences Discussion?}
 
@@
 \end{enumerate}
+
+%###################################################################################################
+\subsection{Details for the Imaging-Based Comparison Algorithm}
+\label{sec:supplement-imaging-based-comparison-algorithm-details}
+
+In a different approach, one might be tempted to utilize the previously constructed direct
+comparison algorithm for time intervals that do not contain any division events.
+But this attempt would quickly lead to multiple problems.
+It will not be able to deal with the uncertainties of the timing of the division event.
+As the number of investigated cells grows, more division events will occurr even in a fixed
+amount of time thus greatly decreasing or completely vanishing the effective interval at which no
+division event is observed.
+The latter situation is equivalent to two microscopic images as input data between which a division
+event has occured.
 

### 2025-11-21 paper/main.tex
cleanup, fix typo, enhance description of how to assign parameters

@@
 Any point with exactly one neighboring point must be and endpoint.
-If we encounter more or less than two points which fit this criterion, the routine can not commence
+If we encounter more or less than two points which fit this criterion, the routine cannot commence
 and is aborted.
@@
 In doing so, we hope to identify all parameters, given relevant experimental data.
-Parameters may be defined as global ones, which are identical across all cells or individual per
-cell.
+As we have explained in the preceding section~\ref{sec:methodology}, parameters may be specified as
+global ones, which are identical across all cells or individual per cell which allows us to capture
+cell-cell variability.
 We utilize the preceding algorithms to extract data from cell-masks, simulate from an initial state
@@
 tools~\cite{Cutler2022,Stringer2020,Hardo2022} or manually annotating microscopic image series.
-
-\textbf{Du solltest hier erwähnen, dass du die Parameter per Bakterium fittest und nicht globale Parameter.}
-\textit{Tue ich aber nicht grundsätzlich. Es gibt globale und individuelle parameter; also ein "mixed approach". Ich habe aber einen Hinweis eingefügt.}
 

### 2025-11-22 paper/main.tex
add to description of imaging-based comparison

@@
 
-Although this optimization scheme was driven by technical challenges, it is also biologically
-motivated.
-In the infenitisemal time point after the division event the resulting two daughter cells will cover
-precisely the space that the mother had taken up.
+The developed optimization scheme was mainly driven by technical challenges, but it is also
+biologically motivated.
+In the short time period after the division event, the two resulting daughter cells will cover
+the space that the mother had taken up.
 Often times, the exact time at which the fission event has commenced can not be determined
@@
 cells begins and the one of the mother ends.
+These uncertainties which are also common problems for segmentation methods, further highlight the
+importance of this comparison scheme.
 

### 2025-11-22 paper/main.tex
add link to supplement
add short introduction for methodological work

@@
 comparison algorithm (see section~\ref{subsec:imaging-based-comparison}).
+Further details can be found in the supplement (see section~\ref{section:supplement-performance})
 
@@
 computational model.
-It has to be noted that they only depend on the spatial representation described in
+It is split into 3 parts, each covering distinct purposes.
+We begin by explaining the visualization process which will later be used within our imaging-based
+comparison algorithm (see section~\ref{subsec:imaging-based-comparison}) and for testing purposes.
+Afterwards, we introduce a position extraction algorithm which extracts cellular information in the
+form of collections of vertices from segmented microscopic images.
+Finally, we will provide two distinct comparison algorithms which can be used to compare numerically
+calculated results with experimental data.
+We will test and benchmark the constructed algorithms within the coming sections.
+
+It should be noted that these methods only depend on the spatial representation described in
 section~\ref{subsec:spatial-representation}.
-Consequently, these methods can be applied to a broader class of mechanical models that may exhibit
+Consequently, these methods can be applied to a broader class of mechanical models that can exhibit
 varied cellular behaviours.
-
-The developed methods are independent of how we assign parameters within our model.
+Furthermore, the developed methods are independent of how we assign parameters within our model.
 This means we can opt to incorporate intracellular variability by specifying parameters on a

### 2025-11-22 paper/main.tex
better motivate method section

@@
 
-For the presented model a list of parameters can be seen in table~\ref{table:list-of-parameters} but
-the following calculations are equally true for any model with $N_p$ number of parameters.
-Consider the case where we filter out a subset $N_g\leq N_p$ of parameters which will be treated
+In order to properly estimate the parameters of our model on the single-cell level, we need to be
+able to take into account any heterogeneity that may be present.
+For the particular model chosen here, a list of parameters can be seen in
+table~\ref{table:list-of-parameters} but the following calculations are equally true for any
+single-cell model with $N_p$ number of parameters.
+Consider the case where we identify a subset $N_g\leq N_p$ of parameters which will be treated
 globally, thus with equal numerical value for every cell.
@@
 cells.
-The total amount of parameters in the system is then given by
+The total amount of parameters $N_\text{tot}$ in the system is then given by
 \begin{equation}
@@
 \end{equation}
+where $N_c$ is the number of unique cells which we encounter over the time interval of interest.
 It is clear, that we should avoid, to needlessly treat parameters non-globally since this increases
@@
 scenarios where different strategies of parameters assignment are desired, e.g. when considering
-two distinct species of cells which should be displayed by 2 parameters in total regardless of the
+two distinct species of cells which should be described by 2 parameters in total regardless of the
 number of cells.
@@
 To do this, we set out to calculate the cost between two successive iteration steps
-$t_n$ with $t_{n+1}$.
+$t_n$ and $t_{n+1}$.
 We expect to obtain an exponential growth in the cost.
 This can be explained by the fact that the growth of the total area is exponential in time.
-Furthermore, the temporal area difference can be viewed as a derivative of the total area in time,
-which is again an exponential function.
+Furthermore, the temporal area difference can be viewed as a derivative of the total area with
+respect to time, which is again an exponential function.
 Figure~\ref{fig:mask-difference-metric} (F) shows the total cost for this approach.

### 2025-11-22 paper/main.tex
adjust awkward formulation
rephrase introduction to methods again

@@
 Afterwards, we introduce a position extraction algorithm which extracts cellular information in the
-form of collections of vertices from segmented microscopic images.
+form of polygons from segmented microscopic images.
 Finally, we will provide two distinct comparison algorithms which can be used to compare numerically
-calculated results with experimental data.
-We will test and benchmark the constructed algorithms within the coming sections.
-
-It should be noted that these methods only depend on the spatial representation described in
+obtained results with experimental data.
+We will test and benchmark the constructed algorithms within the respective sections.
+As highlighted before, the developed methods only depend on the spatial representation described in
 section~\ref{subsec:spatial-representation}.
@@
 varied cellular behaviours.
-Furthermore, the developed methods are independent of how we assign parameters within our model.
-This means we can opt to incorporate intracellular variability by specifying parameters on a
-per-cell basis.
-Or we can employ shared parameters between all agents for processes which can be expected to behave
-identical within margin of error.
-Furthermore, mixed approaches where some parameters are assigned globally and others individually
-can be realized as well.
 
@@
 able to take into account any heterogeneity that may be present.
-For the particular model chosen here, a list of parameters can be seen in
-table~\ref{table:list-of-parameters} but the following calculations are equally true for any
-single-cell model with $N_p$ number of parameters.
-Consider the case where we identify a subset $N_g\leq N_p$ of parameters which will be treated
-globally, thus with equal numerical value for every cell.
-This means that $N_p - N_g$ parameters will be assigned individually for each of $N_c$ number of
-cells.
+Therefore, the developed methods are independent of how we assign parameters within our model.
+This means we can opt to incorporate cell-cell variability by specifying parameters on a
+per-cell basis.
+Or we can employ shared parameters with identical numerical values across all agents for properties
+which are expected to be identical across the population.
+Furthermore, mixed approaches where some parameters are assigned globally and others are chosen
+individually can be realized as well.
+
+For our particular model presented in section~\ref{sec:computational-model}, a list of parameters
+is presented in table~\ref{table:list-of-parameters} but the following calculations are equally
+true for any single-cell model.
+We assume that each cell contains $N_p$ number of parameters that need to be assigned numerical
+values.
+We identify a subset $N_g\leq N_p$ of parameters which will be treated globally, thus with equal
+numerical value for every cell.
+This means that $N_p - N_g$ parameters will be assigned individually for each of number of cells.
 The total amount of parameters $N_\text{tot}$ in the system is then given by
@@
 Notice that the presented approach reflects our current work but can easily be adapted to other
-scenarios where different strategies of parameters assignment are desired, e.g. when considering
-two distinct species of cells which should be described by 2 parameters in total regardless of the
-number of cells.
+scenarios where different strategies of parameters assignment are desired.
+This can be usefull e.g. when considering two distinct species where the value of a particular
+parameter only depends on the species and is thus neither identical across all cells, nor
+individual per cell.
 

### 2025-11-22 paper/main.tex
make visualization more clear

@@
 \label{subsection:visualization}
-To visualize the spatial configuration of the results we employ pyvista, which provides high-level
-wrappers around the popular VTK library~\cite{vtkBook,Sullivan2019}.
+To visualize the spatial configuration of the results we employ \texttt{pyvista}, which provides
+high-level wrappers around the popular VTK library~\cite{vtkBook,Sullivan2019}.
 The initial phase of the visualization process entails the generation of three-dimensional meshes
-for each bacterium as an intermediate representation.
+for each bacterium as an intermediate representation (see
+Figure~\ref{fig:progression-image-generation}).
 We insert spheres matching the thickness of the rod $R$ at each vertex $\vec{x}_i$ and connect
-adjacent vertices $\vec{x}_i,\vec{x}_{i+1}$ with a cylinder, also of radius $r$ which is positioned
+adjacent vertices $\vec{x}_i,\vec{x}_{i+1}$ with a cylinder, of radius $r$ which is positioned
 such that its core is identical to the connecting line segment between the two vertices.
 Finally, we combine the meshes of all of these objects to obtain a single mesh of one bacterium.
-
-A render of this result can be seen in figure~\ref{fig:progression-image-generation}.
+A render including smoothened meshes and external light sources of this configuration can be seen in
+Figure~\ref{fig:progression-image-generation} (B).
+
 Since we use a 3D rendering engine for our visualization, the resulting image can display
@@
 assigning unique colours to every mesh of the respective cellular agent, ensuring no two agents
-will use identical colours.
-We then use a parallel projection and remove any existing light sources in order to obtain a solid
-color for every cell-agent.
-The resulting image is shown in subfigure (B).
+will use identical colours and colors are unique across the duration of the simulation.
+We then use a parallel projection and remove any existing lighting effects in order to obtain an
+image solely comprised of solid colors.
+The resulting image is shown in Figure~\ref{fig:progression-image-generation} (C).
 

### 2025-11-22 paper/main.tex
add citation for pyvista

@@
 \label{subsection:visualization}
-To visualize the spatial configuration of the results we employ \texttt{pyvista}, which provides
+To visualize the spatial configuration of the results we employ
+\texttt{pyvista}~\cite{sullivan2019pyvista}, which provides
 high-level wrappers around the popular VTK library~\cite{vtkBook,Sullivan2019}.

### 2025-11-22 paper/main.tex
rewrite introduction for extraction algorithm

@@
         Spheres are highlighted in blue and cylinders in gray.
-        (B) shows the result from combining sphere and cylinder meshes to obtain the shape of a
-        bacterial rod.
-        This render contains lighting coming from 2 light sources which thus produces a small glow
-        on the upper part of the meshes.
-        (C) Unique color values are assigned to each agent and light sources are removed.
-        Afterwards, the image is rendered by projecting along the z-axis.
-        Modifiers such as lighting or anti-aliasing are deactivated to ensure that each cell is made
-        up of exactly one RGB color.
+        (B) A result from combining sphere and cylinder meshes to obtain the shape of a bacterium.
+        (C) Unique color values are assigned to each agent and additional lighting effects are
+        removed which ensures that the final image only contains one singular color per cell and
+        black background.
+        Finally, the image is rendered by projecting along the z-axis.
     }
@@
 
-This section will explain how we can determine the collection of vertices $\vec{x}_i$ introduced in
-subsection~\ref{subsection:mechanical-model-mechanics} from microscopic images
-(Figure~\ref{fig:position-extraction-algorithm} (A)).
-The extracted vertices for multiple time points will be used by our parameter estimation methods as
-input data.
-The algorithms which we devised requires the input of cell-masks which utilize a unique color for
-every cell in the image.
-
-Such masks can be generated by cell-segmentation tools or obtained by annottating microscopic images
-by hand (Figure~\ref{fig:position-extraction-algorithm} (B)).
-Afterwards, we proceed with the following 5 steps.
+This section explains how we can transform a segmented microscopic image such that we obtain a
+collection of vertices (see Section~\ref{subsection:mechanical-model-mechanics}) that can be used
+to initialize our agents and to perform parameter estimation.
+There exists a plethora of tools to perform segmentation of
+images~\cite{Stringer2020,Cutler2022,Hardo2022}
+Figure~\ref{fig:position-extraction-algorithm} (A) shows a typical microscopic image.
+Our algorithm requires that images are prepared using a pixel-based (in-image) encoding scheme, as
+we explained in section~\ref{subsection:visualization}.
+
+Having defined the input type of our algorithm, we proceed with the following 5 steps:
+\textbf{TODO possibly restructure this as an algorithm instead of list}
 \begin{enumerate}

### 2025-11-22 plots/cell_mesh.py
save image as pdf as well

@@ -164,6 +164,7 @@ def plot_3d_grid(points, radius):
         add_mesh(cylinder)
 
     plotter.screenshot(filename="docs/source/_static/imaging-mesh.png")
+    plotter.save_graphic(filename="docs/source/_static/imaging-mesh.pdf")
 
 
 if __name__ == "__main__":

### 2025-11-23 plots/fitting_progression.py
also save images as pdf; adapt to new render function

@@ -1,8 +1,16 @@
 import cr_mech_coli as crm
 import cv2 as cv
 from pathlib import Path
+from PIL import Image
 
 from fitting_extract_positions import create_simulation_result
+
+
+def store_image(mask, path, name):
+    pil_img = Image.fromarray(mask).convert("RGB")
+    pil_img.save(str(path / name) + ".pdf")
+    cv.imwrite(filename=str(path / name) + ".png", img=mask)
+
 
 if __name__ == "__main__":
     config, cell_container = create_simulation_result(8)
@@ -13,16 +21,20 @@ if __name__ == "__main__":
     i1 = iterations[25]
     i2 = iterations[35]
 
+    color_to_cell = cell_container.color_to_cell
+    parent_map = cell_container.parent_map
+
     rs = crm.RenderSettings()
     mask1 = crm.render_mask(all_cells[i1], colors, config.domain_size, rs)
     mask2 = crm.render_mask(all_cells[i2], colors, config.domain_size, rs)
     mask3 = crm.area_diff_mask(mask1, mask2)
-    mask4 = crm.parents_diff_mask(mask1, mask2, cell_container, 0.5)
+    mask4 = crm.parents_diff_mask(mask1, mask2, color_to_cell, parent_map, 0.5)
 
     # Save first mask
     path = Path("docs/source/_static/fitting-methods/")
     path.mkdir(parents=True, exist_ok=True)
-    cv.imwrite(filename=str(path / "progressions-1.png"), img=mask1)
-    cv.imwrite(filename=str(path / "progressions-2.png"), img=mask2)
-    cv.imwrite(filename=str(path / "progressions-3.png"), img=mask3 * 255.0)
-    cv.imwrite(filename=str(path / "progressions-4.png"), img=mask4 * 255.0)
+
+    store_image(mask1, path, "progressions-1")
+    store_image(mask2, path, "progressions-2")
+    store_image(mask3 * 255.0, path, "progressions-3")
+    store_image(mask4 * 255.0, path, "progressions-4")

### 2025-11-23 plots/fitting_extract_positions.py
also store as pdfs; update images

@@ -7,6 +7,7 @@ import matplotlib as mpl
 from tqdm import tqdm
 import argparse
 import time
+from PIL import Image
 
 
 def calculate_lengths_distances(
@@ -66,10 +67,10 @@ def create_simulation_result(n_vertices: int, rng_seed: int = 3):
         domain_size=np.array([200, 200]),
     )
     config.storage_options = [crm.StorageOption.Memory]
-    config.show_progressbar = True
+    config.progressbar = ""
     agent_settings = crm.AgentSettings(
         growth_rate=0.012,
-        growth_rate_distr=(0.012, 0.002),
+        growth_rate_setter={"mean": 0.012, "std": 0.002},
     )
     agent_settings.mechanics.rigidity = 8.0
     config.domain_height = 0.2
@@ -174,9 +175,11 @@ if __name__ == "__main__":
 
             path = Path("docs/source/_static/fitting-methods/")
             cv.imwrite(
-                filename=str(path / "extract_positions-{:06}.png".format(iteration)),
+                filename=str(path / f"extract_positions-{iteration:06}.png"),
                 img=mask,
             )
+            pil_img = Image.fromarray(mask)
+            pil_img.save(str(path / f"extract_positions-{iteration:06}.pdf"))
 
     ccs = cell_container.serialize()
     arglist = tqdm(

### 2025-11-23 plots/position_extraction_algorithm.py
also save as pdf; update figures; adapt to new extraction method

@@ -83,13 +83,18 @@ if __name__ == "__main__":
 
     fig.tight_layout()
     fig.savefig(
+        "docs/source/_static/fitting-methods/algorithm/mask-zoom.pdf",
+        bbox_inches="tight",
+        pad_inches=0,
+    )
+    fig.savefig(
         "docs/source/_static/fitting-methods/algorithm/mask-zoom.png",
         bbox_inches="tight",
         pad_inches=0,
     )
     plt.close(fig)
 
-    pos, _, _ = crm.extract_positions(mask)
+    pos = crm.extract_positions(mask)[0]
     submask = np.copy(mask[xmin:xmax, ymin:ymax, :])
 
     fig, ax = plt.subplots()
@@ -116,6 +121,11 @@ if __name__ == "__main__":
         ax.plot(q1 + dx / 2, q2 + dy / 2, color="white", marker="+", markersize=15)
 
     fig.tight_layout()
+    fig.savefig(
+        "docs/source/_static/fitting-methods/algorithm/interpolate-positions.pdf",
+        bbox_inches="tight",
+        pad_inches=0,
+    )
     fig.savefig(
         "docs/source/_static/fitting-methods/algorithm/interpolate-positions.png",
         bbox_inches="tight",

### 2025-11-23 paper/main.tex
save image as pdf as well
use pdfs where possible

@@
     \begin{tikzonimage}[width=0.32\textwidth]
-        {docs/source/_static/imaging-mesh.png}
+        {docs/source/_static/imaging-mesh.pdf}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{A}};
@@
     \begin{tikzonimage}[width=0.36\textwidth]
-        {docs/source/_static/fitting-methods/algorithm/mask-zoom.png}
+        {docs/source/_static/fitting-methods/algorithm/mask-zoom.pdf}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
@@
     \begin{tikzonimage}[width=0.255\textwidth]
-        {docs/source/_static/fitting-methods/algorithm/interpolate-positions.png}
+        {docs/source/_static/fitting-methods/algorithm/interpolate-positions.pdf}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
@@
     \begin{tikzonimage}[width=0.5\textwidth]
-        {docs/source/_static/fitting-methods/displacement-calculations.png}%
+        {docs/source/_static/fitting-methods/displacement-calculations.pdf}%
         \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
@@
     \begin{tikzonimage}[width=0.5\textwidth]
-        {docs/source/_static/fitting-methods/displacement-distribution.png}
+        {docs/source/_static/fitting-methods/displacement-distribution.pdf}
         \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
@@
 should be preferred.
-\textbf{[Wir müssen darüber reden, mir ist hier vieles unklar. Es könnte nützlich sein, vor der Application Sektion, eine Zusammenfassung zu setzen.]}
 
@@
     \begin{tikzonimage}[width=0.24\textwidth]
-        {docs/source/_static/fitting-methods/progressions-1.png}
+        {docs/source/_static/fitting-methods/progressions-1.pdf}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
@@
     \begin{tikzonimage}[width=0.24\textwidth]
-        {docs/source/_static/fitting-methods/progressions-2.png}
+        {docs/source/_static/fitting-methods/progressions-2.pdf}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
@@
     \begin{tikzonimage}[width=0.24\textwidth]
-        {docs/source/_static/fitting-methods/progressions-3.png}
+        {docs/source/_static/fitting-methods/progressions-3.pdf}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
@@
     \begin{tikzonimage}[width=0.24\textwidth]
-        {docs/source/_static/fitting-methods/progressions-4.png}
+        {docs/source/_static/fitting-methods/progressions-4.pdf}
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{D}};
@@
         \begin{tikzonimage}[width=0.48\textwidth]
-            {docs/source/_static/fitting-methods/extract_positionsdivision-comparison.png}
+            {docs/source/_static/fitting-methods/extract_positionsdivision-comparison.pdf}
             \node at (0, 1)[anchor=north west, draw, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{E}};
@@
         \begin{tikzonimage}[width=0.48\textwidth]
-            {docs/source/_static/fitting-methods/penalty-time-flow.png}
+            {docs/source/_static/fitting-methods/penalty-time-flow.pdf}
             \node at (0, 0.975)[anchor=north west, draw, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{F}};

### 2025-11-23 paper/main.tex
migrate to plos template

@@
-\documentclass{article}
-
-\usepackage{arxiv}
+\documentclass[10pt,letterpaper]{article}
+\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
 
@@
 \usepackage{acronym}
+\usepackage[right]{lineno}
 \usepackage{listings}
@@
 
-\author{
-    \href{https://orcid.org/0009-0001-0613-7978}{
-        \includegraphics[scale=0.06]{figures/orcid.pdf}
-        \hspace{1mm}Jonas Pleyer
-    }
-    \thanks{
-        \href{https://jonas.pleyer.org}{jonas.pleyer.org},
-        \href{https://cellular-raza.com}{cellular-raza.com}
-    }\\
-	Freiburg Center for Data-Analysis and Modeling\\
-	University of Freiburg\\
-	\texttt{jonas.pleyer@fdm.uni-freiburg.de} \\
-	%% examples of more authors
-	\And
-	\href{https://orcid.org/0000-0002-6371-4495}{
-        \includegraphics[scale=0.06]{figures/orcid.pdf}
-        \hspace{1mm}Christian Fleck
-    }\\
-	Freiburg Center for Data-Analysis and Modeling\\
-	University of Freiburg
-}
-
 \newacro{abm}[ABM]{Agent-Based Model}
@@
 \newacro{bsubtilis}[\textit{B. subtilis}]{\textit{Bacillus subtilis}}
-
-% Define new units
-\DeclareSIUnit\pixel{pix}
-
-% Uncomment to remove the date
-%\date{}
-
-% Uncomment to override  the `A preprint' in the header
-\renewcommand{\headeright}{Preprint}
-%\renewcommand{\undertitle}{Technical Report}
-\renewcommand{\shorttitle}{Methods for Analyzing Models of Rod-Shaped Bacteria}
-
-\usepackage{enumitem}
-\setlist{nolistsep}
+%
+% % Uncomment to remove the date
+% %\date{}
+%
+% % Uncomment to override  the `A preprint' in the header
+% \renewcommand{\headeright}{Preprint}
+% %\renewcommand{\undertitle}{Technical Report}
+% \renewcommand{\shorttitle}{Methods for Analyzing Models of Rod-Shaped Bacteria}
+%
+% \usepackage{enumitem}
+% \setlist{nolistsep}
+
+% create "+" rule type for thick vertical lines
+\newcolumntype{+}{!{\vrule width 2pt}}
+
+% create \thickcline for thick horizontal lines of variable length
+\newlength\savedwidth
+\newcommand\thickcline[1]{%
+  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+  \cline{#1}%
+  \noalign{\vskip\arrayrulewidth}%
+  \noalign{\global\arrayrulewidth\savedwidth}%
+}
+
+% \thickhline command for thick horizontal lines that span the table
+\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+\hline
+\noalign{\global\arrayrulewidth\savedwidth}}
+
+% Remove comment for double spacing
+%\usepackage{setspace}
+%\doublespacing
+
+% Text layout
+\raggedright
+\setlength{\parindent}{0.5cm}
+\textwidth 5.25in
+\textheight 8.75in
+
+% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
+% Captions will be left justified
+\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
+\renewcommand{\figurename}{Fig}
+
+% Please use the included `plos2025.bst` as your BibTeX style
+\bibliographystyle{plos2025}
+
+% Remove brackets from numbering in List of References
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
+
+% Header and Footer with logo
+\usepackage{lastpage,fancyhdr,graphicx}
+\usepackage{epstopdf}
+%\pagestyle{myheadings}
+\pagestyle{fancy}
+\fancyhf{}
+%\setlength{\headheight}{27.023pt}
+%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+\rfoot{\thepage/\pageref{LastPage}}
+\renewcommand{\headrulewidth}{0pt}
+\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
+\fancyheadoffset[L]{2.25in}
+\fancyfootoffset[L]{2.25in}
+\lfoot{\today}
+
+%% Include all user-defined macros below
+
+\newcommand{\lorem}{\textbf{LOREM}}
+\newcommand{\ipsum}{\textbf{IPSUM}}
+
+%% END MACROS SECTION
 
 \begin{document}
-\maketitle
+\begin{flushleft}
+{\Large
+\textbf{Methods for analyzing models of elongated bacteria}
+% Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+}
+\newline
+% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
+\\
+Jonas Pleyer\textsuperscript{1}, Christian Fleck\textsuperscript{1}\\
+\bigskip
+\textbf{1} Freiburg Center for Data-Analysis and Modeling, University of Freiburg, Freiburg Germany
+\\
+\bigskip
 
@@
 
-\bibliographystyle{IEEEtran}
+% \bibliographystyle{IEEEtran}
 \bibliography{references}

### 2025-11-23 paper/main.tex
fix some bib entries

@@
 individually growing \ac{ecoli} using a sequence of microscopic images.
-The data we use~\cite{https://doi.org/10.3203/iwf/k-129} (see
-section~\ref{section:data-code-availability}) has been automatically segmented with the
-omnipose~\cite{Cutler2022} algorithm to obtain the required cell masks.
+The data we use~\cite{Trude1982} (see section~\ref{section:data-code-availability}) has been
+automatically segmented with the omnipose~\cite{Cutler2022} algorithm to obtain the required cell
+masks.
 
@@
 Additional biological effects include polar interactions~\cite{Duvernoy2018},
-differentiation~\cite{vanGestel2015}, extracellular reactions~\cite{Li2025} (see also
-table~\ref{table:simulation-aspects-supplement}).
+differentiation~\cite{vanGestel2015}, extracellular reactions~\cite{Li2025}.
 When it comes to the designed algorithms, we have already discussed that the performance of
@@
 The time series of microscopic images which has been studied in section~\ref{section:application-2}
-was extracted from individual frames of a video~\cite{https://doi.org/10.3203/iwf/k-129} that
+was extracted from individual frames of a video~\cite{Trude1982} that
 showcases the exponential growth of bacteria.

### 2025-11-27 cr_mech_coli/crm_amir/main.py
change color values and sample range

@@ -12,7 +12,7 @@ import multiprocessing as mp
 import argparse
 import itertools
 
-from cr_mech_coli.plotting import COLOR1, COLOR3, COLOR5
+from cr_mech_coli.plotting import COLOR2, COLOR3, COLOR5
 
 GREEN_COLOR = np.array([95, 231, 76])
 
@@ -682,7 +682,7 @@ def crm_amir_main():
 
     # Define globals for all optimizations
     rod_rigidity = (50, 120.0, 400, "[µm/s²]")
-    drag_force = (0.0, 0.0001, 0.0003, "[1/s²µm²]")
+    drag_force = (0.00003, 0.0001, 0.0003, "[1/s²µm²]")
     damping = (0.000, 0.1, 0.3, "[1/s]")
     growth_rate = (0.002, 0.01, 0.02, "[1/s]")
     spring_tension = (10, 30.0, 160.0, "[1/s²]")
@@ -760,7 +760,7 @@ def crm_amir_main():
             pfin1,
             x0_bounds,
             ax,
-            color=COLOR3,
+            color=COLOR2,
             label="Full",
             displacement_error=displacement_error,
         )
@@ -776,7 +776,7 @@ def crm_amir_main():
                 pfin2,
                 x0_bounds_reduced,
                 ax,
-                color=COLOR5,
+                color=COLOR3,
                 label="$\\lambda=0$",
                 displacement_error=displacement_error,
             )
@@ -792,7 +792,7 @@ def crm_amir_main():
                 pfin3,
                 x0_bounds_3,
                 ax,
-                color=COLOR1,
+                color=COLOR5,
                 label="$\\gamma=\\gamma_\\text{opt}$",
                 displacement_error=displacement_error,
             )

### 2025-11-27 paper/main.tex
rewrite list as algorithm

@@
 \usepackage{url}            % simple URL typesetting
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 \usepackage{booktabs}       % professional-quality tables
@@
 \usepackage{mathrsfs}
+\usepackage{mathtools}
 \usepackage{onimage}
@@
 
-Having defined the input type of our algorithm, we proceed with the following 5 steps:
-\textbf{TODO possibly restructure this as an algorithm instead of list}
-\begin{enumerate}
-    \item Obtain submasks for each cell individually.
-    \item Skeletonize submask with algorithm from Lee~\cite{Lee1994}
-    \item Determine endpoints $\vec{q}_0,\vec{q}_1$.
-        Assert that there are only 2 endpoints.
-    \item Sort points such that they form a polygon.
-    \item Interpolate vertices $\vec{x}_i$ from obtained polygon.
-\end{enumerate}
-
+% Having defined the input type of our algorithm, we proceed with the following 5 steps:
+
+\begin{algorithm}
+    \caption{Position Extraction Algorithm}
+    \label{algo:extraction-algorithm}
+    \begin{algorithmic}[1]
+        \Function{Extraction}{\texttt{mask}, \texttt{n\_vertices}}
+        \State $\texttt{positions} \gets [\hspace{6pt}]$
+        \Comment{Initialzie empty array}
+        \For{\texttt{color $\in$ colors(mask)}}
+            \State $\texttt{submask} \gets \texttt{mask}$% \Comment{Obtain submask for each cell}
+            \For{$\texttt{pixel}\in\texttt{submask}$
+                \Comment{Only retain \texttt{background} and \texttt{color}}
+                \If{$\texttt{pixel}\neq\texttt{color}$}}
+                    \State $\texttt{pixel}\gets\texttt{background}$
+                    \Comment{Background is $(0,0,0)$ in \texttt{RGB}}
+                \EndIf
+            \EndFor
+            \State $\mathrlap{\texttt{skeleton}}\hphantom{\texttt{verticess}} \gets \texttt{skeletonization}_\text{Lee}(\texttt{submask})$
+            \Comment{Method by Lee~\cite{Lee1994}}
+            \State $\mathrlap{\texttt{ends}}\hphantom{\texttt{verticess}} \gets \texttt{endpoints}(\texttt{skeleton})$
+            \If{$\#\texttt{ends}\neq2$}
+                \State \textit{Routine Fails}
+            \EndIf
+            \State $\mathrlap{\texttt{polygon}}\hphantom{\texttt{verticess}} \gets \texttt{sort\_points}(\texttt{skeleton},\texttt{ends})$
+            \Comment{Between endpoints}
+            \State $\mathrlap{\texttt{vertices}}\hphantom{\texttt{verticess}} \gets \texttt{interpolate\_}(\texttt{polygon}, \texttt{n\_vertices})$
+            \State $\texttt{append}(\texttt{positions}, \texttt{vertices})$
+        \EndFor
+        \State \Return \texttt{positions}
+        \EndFunction
+    \end{algorithmic}
+\end{algorithm}
+
+Algorithm~\ref{algo:extraction-algorithm} shows the performed steps which allow us to obtain
+vertices for individual bacteria.
+It requires a segmented image and a desired number of vertices to obtain as inputs.
 In the first step, the image will be split into a collection of subimages that only contain

### 2025-11-27 paper/main.tex
adapt abstract to Plos template

@@
 
-%###################################################################################################
-\begin{abstract}
-    \aclp{abm} have become popular tools in the study of complex cellular systems.
-    However, despite their popularity, it is often difficult to systematically estimate their
-    parameters and perform advanced analyses, e.g., using the profile-likelihood.
-    In this work, we study the mechanical properties of rod-shaped bacteria and provide techniques
-    for estimating their parameters and performing identifiability analysis on them.
-    Our approach may serve as a blueprint for a modular framework which can be adapted to other
-    research questions.
-    We use time series of microscopic images as input data and incorporate advanced effects as the
-    elasticity of the rod.
-    The developed model serves as a generalized foundation to describe multiple species of
-    rod-shaped bacteria such as \acl{ecoli} or \acl{bsubtilis}.
-\end{abstract}
+\end{flushleft}
 
@@
 % \keywords{}
+
+%###################################################################################################
+\section*{Abstract}
+Elongated bacteria such as \acl{ecoli} and \acl{bsubtilis} possess distinct mechanical
+and physiological properties that shape their growth, division, and collective behaviors.
+Despite the importance of these organisms, existing computational frameworks rarely capture their
+flexible filamentous morphology, limiting the ability to study their mechanics in a systematic and
+quantitatively grounded way.
+Agent-based models are well suited for representing individual cells and their interactions, yet
+their application to flexible rod-shaped bacteria has been hindered by the absence of appropriate
+mechanical formulations and by the lack of practical methods for estimating model parameters from
+experimental data.
+In this work, we introduce a computational model that represents elongated bacteria as deformable
+rods capable of growth, elastic bending, and mechanical interactions.
+We develop a set of imaging-based and geometry-based comparison methods that extract cell shape and
+position information from segmented microscopy data and quantitatively compare it with model
+simulations at the single-cell level.
+These methods enable the estimation of model parameters across multiple experimental scenarios of
+increasing complexity, including cases that exhibit substantial cell-to-cell variability.
+Our results demonstrate that parameter estimation for complex agent-based models of flexible
+bacteria is feasible and can yield biologically interpretable constraints on cellular mechanics.
+The modeling and analysis framework is modular, generalizable, and readily adaptable to a broad
+range of rod-shaped species.
+By integrating our mechanical model into an extensible computational platform, we provide tools that
+support more rigorous model construction, systematic calibration, and mechanistic interpretation.
+This work establishes a foundation for quantitatively studying the mechanical behavior of
+filamentous bacteria and paves the way for future applications ranging from biofilm formation to
+multicellular
+patterning.
+
+\section*{Author summary}
+
+\linenumbers
 

### 2025-11-27 paper/main.tex
reorganize subsections; only mention heterogeneity in applications as paragraph

@@
 can also calculate their lengths at these time-points and thus determine their rate of growth and
-type of growth (see section~\ref{subsec:parameter-estimation-individual-treatment}), thus
+type of growth (see section~\ref{paragraph:parameter-estimation-individual-treatment}), thus
 effectively guiding our modeling steps.
@@
 
-We have already suspected (see Table~\ref{table:simulation-aspects}) that in order to properly
+\paragraph{Cell-Cell Variability within the Dataset}
+\label{paragraph:parameter-estimation-individual-treatment}
+We have already discussed (see Section~\ref{sec:role-of-heterogeneity}) that in order to properly
 estimate parameters of our system, we need to account for natural cell-variability, thus assigning
-parameters on a per-cell basis.
+selected parameters on a per-cell basis.
 
@@
 
-\label{subsec:parameter-estimation-individual-treatment}
 \begin{figure}[H]

### 2025-11-27 paper/main.tex
move heterogeneity into separate subsection at end of methods

@@
 
-In order to properly estimate the parameters of our model on the single-cell level, we need to be
-able to take into account any heterogeneity that may be present.
-Therefore, the developed methods are independent of how we assign parameters within our model.
-This means we can opt to incorporate cell-cell variability by specifying parameters on a
-per-cell basis.
-Or we can employ shared parameters with identical numerical values across all agents for properties
-which are expected to be identical across the population.
-Furthermore, mixed approaches where some parameters are assigned globally and others are chosen
-individually can be realized as well.
-
-For our particular model presented in section~\ref{sec:computational-model}, a list of parameters
-is presented in table~\ref{table:list-of-parameters} but the following calculations are equally
-true for any single-cell model.
-We assume that each cell contains $N_p$ number of parameters that need to be assigned numerical
-values.
-We identify a subset $N_g\leq N_p$ of parameters which will be treated globally, thus with equal
-numerical value for every cell.
-This means that $N_p - N_g$ parameters will be assigned individually for each of number of cells.
-The total amount of parameters $N_\text{tot}$ in the system is then given by
-\begin{equation}
-    N_\text{tot} = N_g + N_c (N_p - N_g).
-\end{equation}
-where $N_c$ is the number of unique cells which we encounter over the time interval of interest.
-It is clear, that we should avoid, to needlessly treat parameters non-globally since this increases
-the size of the parameter space considerably.
-The volume of parameter space which needs to be sampled by optimization schemes scales exponentially
-with the dimension ie. the total number of parameters $N_\text{tot}$.
-Notice that the presented approach reflects our current work but can easily be adapted to other
-scenarios where different strategies of parameters assignment are desired.
-This can be usefull e.g. when considering two distinct species where the value of a particular
-parameter only depends on the species and is thus neither identical across all cells, nor
-individual per cell.
-
 % --------------------------------------------------------------------------------------------------
@@
 
+% --------------------------------------------------------------------------------------------------
+\subsection{The challenge of Cellular Heterogeneity}
+\label{sec:role-of-heterogeneity}
+In order to properly estimate the parameters of our model on the single-cell level, we need to be
+able to take into account any heterogeneity that may be present.
+Therefore, the developed methods are independent of how we assign parameters within our model.
+This means we can opt to incorporate cell-cell variability by specifying parameters on a
+per-cell basis.
+Or we can employ shared parameters with identical numerical values across all agents for properties
+which are expected to be identical across the population.
+Furthermore, mixed approaches where some parameters are assigned globally and others are chosen
+individually can be realized as well.
+
+For our particular model presented in section~\ref{sec:computational-model}, a list of parameters
+is presented in table~\ref{table:list-of-parameters} but the following calculations are equally
+true for any single-cell model.
+We assume that each cell contains $N_p$ number of parameters that need to be assigned numerical
+values.
+We identify a subset $N_g\leq N_p$ of parameters which will be treated globally, thus with equal
+numerical value for every cell.
+This means that $N_p - N_g$ parameters will be assigned individually for each of number of cells.
+The total amount of parameters $N_\text{tot}$ in the system is then given by
+\begin{equation}
+    N_\text{tot} = N_g + N_c (N_p - N_g).
+\end{equation}
+where $N_c$ is the number of unique cells which we encounter over the time interval of interest.
+This value includes cells which have divided or died during the inspected time interval.
+
+It is clear, that we should avoid, to needlessly treat parameters non-globally since this increases
+the size of the parameter space considerably.
+The volume of parameter space which needs to be sampled by optimization schemes scales exponentially
+with the dimension ie. the total number of parameters $N_\text{tot}$.
+Notice that the presented approach reflects our current work but can be adapted easily to other
+scenarios where different strategies of parameter assignment are desired.
+This can be usefull e.g. when considering two distinct species where the value of a particular
+parameter only depends on the species and is thus neither identical across all cells, nor
+individual per cell.
+
+%###################################################################################################
 \section{Applications}
@@
 
-%###################################################################################################
+% --------------------------------------------------------------------------------------------------
 \subsection{Application 1: Elastic deformations of \acl{ecoli}}

### 2025-11-27 paper/main.tex
link to supplement

@@
 In contrast, a popular alternative is the Morse potential~\cite{Morse1929}, which is less
-customizable due to the lower amount of parameters and more restrictive mathematical formulation.
+customizable due to the lower amount of parameters and more restrictive mathematical formulation
+(see section~\ref{sec:supplement-morse-potential}).
 Therefore, the Mie potential is a good choice for a simple, interpretable interaction potential with
@@
 \subsection{Alternative Interaction Potential - Morse Potential}
+\label{sec:supplement-morse-potential}
 The morse potential~\cite{Morse1929} is a common example of numerically inexpensive interaction

### 2025-11-27 paper/main.tex
more details to comparison algorithms

@@
 
+\paragraph{Determining the underlying Error Model}
+
 \begin{figure}[H]
@@
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
-    \end{tikzonimage}
+    \end{tikzonimage}%
+    \hspace{0.01\textwidth}%
     \begin{tikzonimage}[width=0.36\textwidth]
@@
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
-    \end{tikzonimage}
+    \end{tikzonimage}%
+    \hspace{0.01\textwidth}%
     \begin{tikzonimage}[width=0.255\textwidth]
@@
         \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
-    \end{tikzonimage}
-    \begin{tikzonimage}[width=0.5\textwidth]
+    \end{tikzonimage}\\
+    \vspace{0.005\textwidth}%
+    \begin{tikzonimage}[width=0.195\textwidth]
+        {docs/source/_static/fitting-methods/extract_positions-001800.pdf}
+        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+    \end{tikzonimage}%
+    \hspace{0.005\textwidth}%
+    \begin{tikzonimage}[width=0.195\textwidth]
+        {docs/source/_static/fitting-methods/extract_positions-003800.pdf}
+        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+    \end{tikzonimage}%
+    \hspace{0.005\textwidth}%
+    \begin{tikzonimage}[width=0.195\textwidth]
+        {docs/source/_static/fitting-methods/extract_positions-005800.pdf}
+        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{F}};
+    \end{tikzonimage}%
+    \hspace{0.005\textwidth}%
+    \begin{tikzonimage}[width=0.195\textwidth]
+        {docs/source/_static/fitting-methods/extract_positions-007800.pdf}
+        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{G}};
+    \end{tikzonimage}%
+    \hspace{0.005\textwidth}%
+    \begin{tikzonimage}[width=0.195\textwidth]
+        {docs/source/_static/fitting-methods/extract_positions-009800.pdf}
+        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{H}};
+    \end{tikzonimage}\\
+    \hspace{0.005\textwidth}%
+    \begin{tikzonimage}[width=0.33\textwidth]
         {docs/source/_static/fitting-methods/displacement-calculations.pdf}%
-        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
-    \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.5\textwidth]
+        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{I}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.33\textwidth]
         {docs/source/_static/fitting-methods/displacement-distribution.pdf}
-        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
-    \end{tikzonimage}
+        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{J}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.33\textwidth]
+        {docs/source/_static/fitting-methods/displacement-fit-over-time.pdf}
+        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{K}};
+    \end{tikzonimage}%
     \caption{
-        (A) Input image.
+        (A) Microscopic image used as input for the extraction algorithm.
         (B) Result of the skeletonization algorithm for a single cell.
         (C) Vertices $\vec{x}_i$ which have been interpolated from the skeletonized cell.
-        They represent the position of our cell-agents.
-        Subfigure (D) displayes the average rod length of the exact and extracted data.
+        They can be directly used as the position of our cell-agents.
+        (D-H) Snapshots of simulated data used for benchmarking.
+        (I) Average rod length of the exact and extracted data.
         The light blue region is the mean distance between vertices of extracted and exact
         positions which captures the inherent uncertainty of our method.
-        Subfigure (E) displays all of these distances on a logarithmic scale.
-        The colorscale represents earlier (darker) and later (brighter) temporal states of our
+        (J) Displacement of individual vertices parallel and orthogonal to the rod.
+        The exact value of the vertex is always in the center.
+        Brighter colors indicate higher densities.
+        The range of the plot is restricted to a $3\sigma$ interval, to enhance visualization and
+        exclude outliers.
+        (K) Mean $\mu_{p,o}$ and variance $\sigma_{p,o}$ of fitted normal distribution for
+        displacements parallel and orthogonal to the rod displayed over the duration of the
         simulation.
@@
 $r\geq\xi$.
+
+%###################################################################################################
+\subsection{Details for the Imaging-Based Comparison Algorithm}
+\label{sec:supplement-imaging-based-comparison-algorithm-details}
+
+In a different approach, one might be tempted to utilize the previously constructed direct
+comparison algorithm for time intervals that do not contain any division events.
+But this attempt would quickly lead to multiple problems.
+It will not be able to deal with the uncertainties of the timing of the division event.
+As the number of investigated cells grows, more division events will occurr even in a fixed
+amount of time thus greatly decreasing or completely vanishing the effective interval at which no
+division event is observed.
+The latter situation is equivalent to two microscopic images as input data between which a division
+event has occured.
+
+\begin{figure}
+    \centering
+    \begin{tikzonimage}[width=0.5\textwidth]
+        {docs/source/_static/fitting-methods/displacement-distr-x.pdf}
+        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.5\textwidth]
+        {docs/source/_static/fitting-methods/displacement-distr-y.pdf}
+        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
+    \end{tikzonimage}\\
+    \begin{tikzonimage}[width=0.5\textwidth]
+        {docs/source/_static/fitting-methods/displacement-fit-over-time-covariance.pdf}
+        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{H}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.5\textwidth]
+        {docs/source/_static/fitting-methods/displacement-distribution.pdf}
+        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{I}};
+    \end{tikzonimage}
+    \caption{TODO}
+\end{figure}
 

### 2025-11-27 paper/main.tex
remove root from direct-comparison method
combine all figures into one large panel for amir

@@
     L_{\text{LS}}(\vec{x}, \vec{y})
-        = \sqrt{\sum\limits_{i,j,k}|\nu_{i,j,k} - \tilde{\nu}_{i,j,k}|^2}
+        = \sum\limits_{i,j,k}|\nu_{i,j,k} - \tilde{\nu}_{i,j,k}|^2
 \end{equation}
@@
     \centering
-    \begin{minipage}{0.38\textwidth}
-        \begin{tikzonimage}[width=\textwidth]
-            {figures/amir-elastic-000032-with-axes.png}%
-            \node at (0.01, 0.01+17pt)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
-        \end{tikzonimage}%
-    \end{minipage}%
-    \begin{minipage}{0.62\textwidth}
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {figures/crm_amir/progressions/step4-000010.png}
+        \node at (0.02,17pt)[anchor=north west, white, draw, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \end{tikzonimage}%
+    \hspace{0.01\textwidth}%
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {figures/crm_amir/progressions/step4-000017.png}
+        \node at (0.02,17pt)[anchor=north west, white, draw, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+    \end{tikzonimage}%
+    \hspace{0.01\textwidth}%
+    \begin{tikzonimage}[width=0.3\textwidth]
+        {figures/crm_amir/progressions/step4-000034.png}
+            \node at (0.01, 0.01+17pt)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+    \end{tikzonimage}\\
     \begin{tikzonimage}[width=0.33\textwidth]
@@
         {figures/crm_amir/profiles/drag_force.pdf}%
-            \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+        \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
     \end{tikzonimage}%
@@
         {figures/crm_amir/profiles/growth_rate.pdf}%
-            \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+        \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
     \end{tikzonimage}\\
@@
         {figures/crm_amir/profiles/rod_rigidity.pdf}%
-            \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+        \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
     \end{tikzonimage}%
@@
         {figures/crm_amir/profiles/spring_tension.pdf}%
-            \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
+        \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
     \end{tikzonimage}%
@@
         {figures/crm_amir/profiles/damping.pdf}%
-            \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
-        \end{tikzonimage}%
-    \end{minipage}
+        \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{H}};
+    \end{tikzonimage}%
     \caption{
+        \textbf{TODO fix caption numbers and add missing stuff}
         (A) Microscopic image containing a single bacterium which is partially confined in a narrow
@@
         with the experimental data.
-        (C-G) Likelihood profiles for the individual parameters before and after performing model
-        reduction.
+        (C-G) Likelihood profiles for the individual parameters before and after performing two
+        successive model reductions.
+        The first step removes any damping effects by settings $\lambda=0$ while the second fixes
+        the spring tension to its calculated optimum $\gamma=\gamma_\text{opt}$.
     }

### 2025-11-27 paper/main.tex
show data alongside heterogeneic growth rates for application 2

@@
     \centering
-    \begin{tikzonimage}[width=0.24\textwidth]
-        {../data/crm_divide/0003/images/001044.png}%
-        \node at (0.02, 0.98)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+    \begin{tikzonimage}[width=0.49\textwidth]
+        {figures/data/crm_fit/001044.png}%
+        \node at (0.02, 0.98)[anchor=north west, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
     \end{tikzonimage}%
     \hspace{0.01\textwidth}%
-    \begin{tikzonimage}[width=0.24\textwidth]
-        {../data/crm_divide/0003/images/001054.png}%
-        \node at (0.02, 0.98)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+    \begin{tikzonimage}[width=0.49\textwidth]
+        {figures/data/crm_fit/001044-markers.pdf}%
+        \node at (0.02, 0.98)[anchor=north west, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+    \end{tikzonimage}\\
+    \begin{tikzonimage}[width=0.28\textwidth]
+        {figures/crm_fit-data-progression.png}
+        \node at (0.02, 0.98)[anchor=north west, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
     \end{tikzonimage}%
     \hspace{0.01\textwidth}%
-    \begin{tikzonimage}[width=0.24\textwidth]
-        {../data/crm_divide/0003/images/001063.png}%
-        \node at (0.02, 0.98)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
-    \end{tikzonimage}%
-    \hspace{0.01\textwidth}%
-    \begin{tikzonimage}[width=0.24\textwidth]
-        {../data/crm_divide/0003/images/001074.png}%
-        \node at (0.02, 0.98)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{D}};
-    \end{tikzonimage}%
-    \caption{TODO insert segmentation masks here}
+    \begin{tikzonimage}[width=0.35\textwidth]
+        {figures/crm_estimate_params/IWF-Goettingen/rod-lengths-individual.pdf}%
+        \node at (0, 1)[anchor=north west, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.35\textwidth]
+        {figures/crm_estimate_params/IWF-Goettingen/growth-rates-bar-plot.pdf}%
+        \node at (0, 1)[anchor=north west, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+    \end{tikzonimage}%
+    \caption{
+        \textbf{TODO also insert vertical in (B) line to indicate when direct comparison algorithm
+        starts.}
+        \textbf{Link to supplement for discussion of bar plot}\\
+        (A) Snapshot of used microscopic images.
+        (B) Segmented microscopic image.
+        (C) Selected time points of the utilized time series showcasing the spatial configurations
+        of the investigated bacteria.
+        A complete list of snapshots is included in the supplementary material (see
+        section~\ref{sec:supplement-data-details}).
+        (D) Growth curves of individual bacteria including a discrete delay before exponential
+        growth is experienced.
+        (E) Distribution of growth parameters shows cell-cell variability of the analyzed sample.
+        The obtained results are further discussed in the supplementary material section~\ref{}.
+    }
     \label{fig:application-2-data-overview}

### 2025-11-27 paper/main.tex
fixup! reorganize subsections; only mention heterogeneity in applications as paragraph
fixup! show data alongside heterogeneic growth rates for application 2

@@
 
-% --------------------------------------------------------------------------------------------------
-\subsubsection{The role of Heterogeneity}
-
 \paragraph{Cell-Cell Variability within the Dataset}
@@
 However, we will now present data that indicate that an individual treatment is in fact necessary.
-Figure~\ref{fig:estimated-growth-rates} shows the rod-lengths of individual bacteria which have been
+Figure~\ref{fig:application-2-data-overview} shows the rod-lengths of individual bacteria which have been
 calculated from the extracted positions.
@@
 Fitting this function to the individual growth curves of the bacteria and their mean growth curve,
-both results in good agreements (Figure~\ref{fig:estimated-growth-rates} A,B).
+both results in good agreements (Figure~\ref{fig:application-2-data-overview} A,B).
 However, the latter exhibits large amounts of uncertainty within the data which is also illustrated
@@
 % \end{enumerate}
-
-\begin{figure}[H]
-    \centering
-    \begin{tikzonimage}[width=0.28\textwidth]
-        {figures/crm_fit-data-progression.png}
-        \node at (0.02, 0.98)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
-    \end{tikzonimage}%
-    \hspace{0.01\textwidth}%
-    \begin{tikzonimage}[width=0.35\textwidth]
-        {figures/crm_estimate_params/IWF-Goettingen/rod-lengths-individual.pdf}%
-        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
-    \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.35\textwidth]
-        {figures/crm_estimate_params/IWF-Goettingen/growth-rates-distribution.pdf}%
-        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
-    \end{tikzonimage}%
-    \caption{
-        \textbf{TODO Fix caption!}\\
-        \textbf{Link to supplement for discussion of bar plot}\\
-        % TODO use same y-axis across the plots
-        (B) The subfigure shows the time-evolution of the rod-lengths for every individual
-        bacterium.
-        These growth curves are fitted individually with delayed exponential function.
-        (C) The datapoints are created by averaging the rod-lengths shown in subfigure (A) and then
-        fitted with an exponential function.
-        The uncertainty band is the standard deviation of the averaged datapoints.
-        (C) Blue dots show the combination of estimated growth rates and delays for individual
-        bacteria while the orange dot corresponds to the average.
-        Ellipses around the datapoints are taken from the uncertainty of the fits.
-        Time is given in units of frames of the movie from which the data have been extracted.
-    }
-    \label{fig:estimated-growth-rates}
-\end{figure}
 

### 2025-11-27 paper/main.tex
add author contributions
reorganize supplementary sections in order

@@
 
+\section{Author Contributions}
+\textbf{Conceptualization:} Jonas Pleyer\\
+\textbf{Investigation:} Jonas Pleyer\\
+\textbf{Methodology} Jonas Pleyer\\
+\textbf{Software:} Jonas Pleyer\\
+\textbf{Supervision:} Christian Fleck\\
+\textbf{Writing – original draft:} Jonas Pleyer\\
+\textbf{Writing – review \& editing:} Jonas Pleyer \& Christian Fleck
+
 %###################################################################################################
@@
 \section{Supplementary Material}
-
-%###################################################################################################
-\subsection{Data Preparation (TODO)}
-\begin{enumerate}
-    \item Have masks of data; reference image segmantation tools; or do it by hand
-    \item ensure that colors are matching for bacteria between time-steps
-    \item clean borders of images if needed
-    \item know domain size; otherwise use pixels
-\end{enumerate}
-
-%###################################################################################################
-\subsection{Details for the Imaging-Based Comparison Algorithm}
-\label{sec:supplement-imaging-based-comparison-algorithm-details}
-
-In a different approach, one might be tempted to utilize the previously constructed direct
-comparison algorithm for time intervals that do not contain any division events.
-But this attempt would quickly lead to multiple problems.
-It will not be able to deal with the uncertainties of the timing of the division event.
-As the number of investigated cells grows, more division events will occurr even in a fixed
-amount of time thus greatly decreasing or completely vanishing the effective interval at which no
-division event is observed.
-The latter situation is equivalent to two microscopic images as input data between which a division
-event has occured.
-
-%###################################################################################################
-\subsection{Parameter Estimation Results (TODO)}
-\label{sec:supplement-parameter-profiles}
-
-%TODO {explain profiles}
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/growth-rate-0.pdf}%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/growth-rate-1.pdf}%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/growth-rate-2.pdf}\\%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/growth-rate-3.pdf}%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/growth-rate-4.pdf}%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/growth-rate-5.pdf}
-    \caption{Parameter estimation of growth rates for individual cells}%TODO extend caption}
-    \label{fig:parameter-estimates-supplement-growth-rates}
-\end{figure}
 
@@
 %###################################################################################################
-\subsection{How to use obtained heterogenous results}
+\subsection{Analysis \& Usage of heterogeneous Parameter Value (TODO)}
 In order to reuse this model, we also provide the option to assign growth rates by sampling from a
 normal distribution with user-defined parameters.
+
+%###################################################################################################
+\subsection{Application Data \& Preparation (TODO)}
+\label{sec:supplement-data-details}
+
+\begin{figure}[H]
+    \centering
+    % TODO include snapshots of utilized data
+    \caption{TODO}
+\end{figure}
+
+\begin{enumerate}
+    \item Have masks of data; reference image segmantation tools; or do it by hand
+    \item ensure that colors are matching for bacteria between time-steps
+    \item clean borders of images if needed
+    \item know domain size; otherwise use pixels
+\end{enumerate}
+
+%###################################################################################################
+\subsection{Parameter Estimation Results (TODO)}
+\label{sec:supplement-parameter-profiles}
+
+%TODO {explain profiles}
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.33\textwidth]
+        {figures/crm_fit/profiles/growth-rate-0.pdf}%
+    \includegraphics[width=0.33\textwidth]
+        {figures/crm_fit/profiles/growth-rate-1.pdf}%
+    \includegraphics[width=0.33\textwidth]
+        {figures/crm_fit/profiles/growth-rate-2.pdf}\\%
+    \includegraphics[width=0.33\textwidth]
+        {figures/crm_fit/profiles/growth-rate-3.pdf}%
+    \includegraphics[width=0.33\textwidth]
+        {figures/crm_fit/profiles/growth-rate-4.pdf}%
+    \includegraphics[width=0.33\textwidth]
+        {figures/crm_fit/profiles/growth-rate-5.pdf}
+    \caption{Parameter estimation of growth rates for individual cells}%TODO extend caption}
+    \label{fig:parameter-estimates-supplement-growth-rates}
+\end{figure}
 

### 2025-11-27 paper/main.tex
add idea for discussion
rephrase slightly

@@
     \item other dynamics
+    \item advantage of imaging-based comparison algo: no need to specify number of vertices; more
+        naturally aligned with biolgical reality (no vertices)
 \end{itemize}
@@
 stiffness $\lambda$ which controls the rate of change as one travels along the radial axis.
-We employ a spatial cutoff which sets the interaction strength to $0$ when exceeding a distance
-$r\geq\xi$.
+As done with our approach towards the Mie potential, we are free to assign a spatial cutoff $\xi$
+to the range of the interaction potential.
 

### 2025-11-28 paper/main.tex
do not draw boxes; add new figures

@@
         {docs/source/_static/mechanics.pdf}
-        \node at (0, 1)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+        \node at (0, 1)[anchor=north west, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{A}};
     \end{tikzonimage}%
@@
         {docs/source/_static/interaction.pdf}
-        \node at (0, 1)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0, 1)[anchor=north west, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{B}};
     \end{tikzonimage}%
@@
         {docs/source/_static/mie-potential-shapes.pdf}
-        \node at (0, 1)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+        \node at (0, 1)[anchor=north west, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{C}};
     \end{tikzonimage}%
@@
         {docs/source/_static/imaging-mesh.pdf}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+        \node at (0.025, 0.975)[anchor=north west, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{A}};
     \end{tikzonimage}%
@@
         {docs/source/_static/11571737453049821261/raw_pv/000000400.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0.025, 0.975)[anchor=north west, rectangle, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
     \end{tikzonimage}%
@@
         {docs/source/_static/11571737453049821261/masks/000000400.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+        \node at (0.025, 0.975)[anchor=north west, rectangle, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
     \end{tikzonimage}
@@
         {docs/source/_static/fitting-methods/algorithm/image001042.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+        \node at (0.025, 0.975)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/algorithm/mask-zoom.pdf}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0.025, 0.975)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/algorithm/interpolate-positions.pdf}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+        \node at (0.025, 0.975)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
     \end{tikzonimage}\\
-    \vspace{0.005\textwidth}%
-    \begin{tikzonimage}[width=0.195\textwidth]
+    \vspace{0.004\textwidth}%
+    \begin{tikzonimage}[width=0.1625\textwidth]
         {docs/source/_static/fitting-methods/extract_positions-001800.pdf}
-        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{D}};
-    \end{tikzonimage}%
-    \hspace{0.005\textwidth}%
-    \begin{tikzonimage}[width=0.195\textwidth]
+        \node at (0.03, 0.99)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+    \end{tikzonimage}%
+    \hspace{0.004\textwidth}%
+    \begin{tikzonimage}[width=0.1625\textwidth]
         {docs/source/_static/fitting-methods/extract_positions-003800.pdf}
-        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{E}};
-    \end{tikzonimage}%
-    \hspace{0.005\textwidth}%
-    \begin{tikzonimage}[width=0.195\textwidth]
+        \node at (0.03, 0.99)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+    \end{tikzonimage}%
+    \hspace{0.004\textwidth}%
+    \begin{tikzonimage}[width=0.1625\textwidth]
         {docs/source/_static/fitting-methods/extract_positions-005800.pdf}
-        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{F}};
-    \end{tikzonimage}%
-    \hspace{0.005\textwidth}%
-    \begin{tikzonimage}[width=0.195\textwidth]
+        \node at (0.03, 0.99)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{F}};
+    \end{tikzonimage}%
+    \hspace{0.004\textwidth}%
+    \begin{tikzonimage}[width=0.1625\textwidth]
         {docs/source/_static/fitting-methods/extract_positions-007800.pdf}
-        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{G}};
-    \end{tikzonimage}%
-    \hspace{0.005\textwidth}%
-    \begin{tikzonimage}[width=0.195\textwidth]
+        \node at (0.03, 0.99)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{G}};
+    \end{tikzonimage}%
+    \hspace{0.004\textwidth}%
+    \begin{tikzonimage}[width=0.1625\textwidth]
         {docs/source/_static/fitting-methods/extract_positions-009800.pdf}
-        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{H}};
+        \node at (0.03, 0.99)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{H}};
+    \end{tikzonimage}%
+    \hspace{0.004\textwidth}%
+    \begin{tikzonimage}[width=0.1625\textwidth]
+        {docs/source/_static/fitting-methods/extract_positions-007800-cropped.pdf}%
+        \node at (0, 1)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{I}};
     \end{tikzonimage}\\
-    \hspace{0.005\textwidth}%
     \begin{tikzonimage}[width=0.33\textwidth]
         {docs/source/_static/fitting-methods/displacement-calculations.pdf}%
-        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{I}};
+        \node at (0, 1)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{J}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/displacement-distribution.pdf}
-        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{J}};
+        \node at (0, 1)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{K}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/displacement-fit-over-time.pdf}
-        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{K}};
+        \node at (0, 1)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{L}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/progressions-1.pdf}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+        \node at (0.025, 0.975)[anchor=north west, rectangle, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/progressions-2.pdf}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0.025, 0.975)[anchor=north west, rectangle, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/progressions-3.pdf}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+        \node at (0.025, 0.975)[anchor=north west, rectangle, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/progressions-4.pdf}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+        \node at (0.025, 0.975)[anchor=north west, rectangle, white, minimum width=15pt, minimum height=15pt]{\textbf{D}};
     \end{tikzonimage}\\
@@
             {docs/source/_static/fitting-methods/extract_positionsdivision-comparison.pdf}
-            \node at (0, 1)[anchor=north west, draw, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+            \node at (0, 1)[anchor=north west, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{E}};
         \end{tikzonimage}%
@@
             {docs/source/_static/fitting-methods/penalty-time-flow.pdf}
-            \node at (0, 0.975)[anchor=north west, draw, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{F}};
+            \node at (0, 0.975)[anchor=north west, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{F}};
         \end{tikzonimage}%
@@
         {figures/crm_amir/progressions/step4-000010.png}
-        \node at (0.02,17pt)[anchor=north west, white, draw, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+        \node at (0.02,17pt)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
     \end{tikzonimage}%
@@
         {figures/crm_amir/progressions/step4-000017.png}
-        \node at (0.02,17pt)[anchor=north west, white, draw, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0.02,17pt)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
     \end{tikzonimage}%
@@
         {figures/crm_amir/progressions/step4-000034.png}
-            \node at (0.01, 0.01+17pt)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+            \node at (0.01, 0.01+17pt)[anchor=north west, rectangle, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
     \end{tikzonimage}\\
@@
         {figures/data/crm_fit/001044.png}%
-        \node at (0.02, 0.98)[anchor=north west, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+        \node at (0.02, 0.98)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
     \end{tikzonimage}%
@@
         {figures/data/crm_fit/001044-markers.pdf}%
-        \node at (0.02, 0.98)[anchor=north west, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0.02, 0.98)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
     \end{tikzonimage}\\
+    \vspace{0.01\textwidth}%
     \begin{tikzonimage}[width=0.28\textwidth]
         {figures/crm_fit-data-progression.png}
-        \node at (0.02, 0.98)[anchor=north west, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+        \node at (0.02, 0.98)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
     \end{tikzonimage}%
@@
         {figures/crm_estimate_params/IWF-Goettingen/rod-lengths-individual.pdf}%
-        \node at (0, 1)[anchor=north west, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+        \node at (0, 1)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
     \end{tikzonimage}%
@@
         {figures/crm_estimate_params/IWF-Goettingen/growth-rates-bar-plot.pdf}%
-        \node at (0, 1)[anchor=north west, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+        \node at (0, 1)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
     \end{tikzonimage}%
@@
         {figures/crm_fit/snapshots/predicted-003889.png}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+        \node at (0.025, 0.975)[anchor=north west, rectangle, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
     \end{tikzonimage}%
@@
         {figures/crm_fit/celldiffs/diff-003889.png}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0.025, 0.975)[anchor=north west, rectangle, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
     \end{tikzonimage}%
@@
         {figures/crm_fit/profiles/damping.pdf}%
-        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+        \node at (0, 1.1)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{A}};
     \end{tikzonimage}%
@@
         {figures/crm_fit/profiles/exponent-m.pdf}%
-        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0, 1.1)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
     \end{tikzonimage}%
@@
         {figures/crm_fit/profiles/exponent-n.pdf}%
-        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+        \node at (0, 1.1)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
     \end{tikzonimage}%
@@
         {figures/crm_fit/profiles/strength.pdf}%
-        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+        \node at (0, 1.1)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
     \end{tikzonimage}\\
@@
         {figures/crm_fit/profiles/radius.pdf}%
-        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+        \node at (0, 1.1)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
     \end{tikzonimage}%
@@
         {figures/crm_fit/profiles/growth-rate-0.pdf}%
-        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
+        \node at (0, 1.1)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
     \end{tikzonimage}%
@@
         {figures/crm_fit/profiles/growth-rate-1.pdf}%
-        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
+        \node at (0, 1.1)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
     \end{tikzonimage}%
@@
         {figures/crm_fit/profiles/growth-rate-2.pdf}%
-        \node at (0, 1.1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{H}};
+        \node at (0, 1.1)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{H}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/masks_diff/000000.png}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+        \node at (0.025, 0.975)[anchor=north west, rectangle, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/masks_diff/000010.png}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0.025, 0.975)[anchor=north west, rectangle, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/masks_diff/000020.png}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+        \node at (0.025, 0.975)[anchor=north west, rectangle, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/profiles/profile-000000.png}%
-        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+        \node at (0.15, 0.85)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{A}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/profiles/profile-000001.png}%
-        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0.15, 0.85)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/profiles/profile-000002.png}%
-        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+        \node at (0.15, 0.85)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/profiles/profile-000003.png}%
-        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+        \node at (0.15, 0.85)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
     \end{tikzonimage}\\
@@
         {figures/crm_divide/profiles/profile-000004.png}%
-        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+        \node at (0.15, 0.85)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/profiles/profile-000005.png}%
-        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
+        \node at (0.15, 0.85)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/profiles/profile-000006.png}%
-        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
+        \node at (0.15, 0.85)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/profiles/profile-000007.png}%
-        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{H}};
+        \node at (0.15, 0.85)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{H}};
     \end{tikzonimage}\\
@@
         {figures/crm_divide/profiles/profile-000008.png}%
-        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{I}};
+        \node at (0.15, 0.85)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{I}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/profiles/profile-000009.png}%
-        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{J}};
+        \node at (0.15, 0.85)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{J}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/profiles/profile-000010.png}%
-        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{K}};
+        \node at (0.15, 0.85)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{K}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/profiles/profile-000011.png}%
-        \node at (0.15, 0.85)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{L}};
+        \node at (0.15, 0.85)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{L}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/displacement-distr-x.pdf}
-        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
+        \node at (0.03, 0.99)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/displacement-distr-y.pdf}
-        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
+        \node at (0.03, 0.99)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
     \end{tikzonimage}\\
@@
         {docs/source/_static/fitting-methods/displacement-fit-over-time-covariance.pdf}
-        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{H}};
+        \node at (0.03, 0.99)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{H}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/displacement-distribution.pdf}
-        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{I}};
+        \node at (0.03, 0.99)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{I}};
     \end{tikzonimage}
@@
         {figures/crm_amir/progressions/step4-000034.png}
-        \node at (0.02,17pt)[anchor=north west, white, draw, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+        \node at (0.02,17pt)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
     \end{tikzonimage}\\
@@
         {figures/crm_amir/result-full/0000000000.png}
-        \node at (0.02, 0.98)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+        \node at (0.02, 0.98)[anchor=north west, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{D}};
     \end{tikzonimage}%
@@
         {figures/crm_amir/result-full/0000000001.png}
-        \node at (0.02, 0.98)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+        \node at (0.02, 0.98)[anchor=north west, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{E}};
     \end{tikzonimage}%
@@
         {figures/crm_amir/result-full/0000000002.png}
-        \node at (0.02, 0.98)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{F}};
+        \node at (0.02, 0.98)[anchor=north west, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{F}};
     \end{tikzonimage}%

### 2025-11-28 paper/main.tex
rewrite text for benchmarking extraction algorithm

@@
 
-In order to determine if our extraction algorithm provides reliable results, we set out to compare
-extracted data with known real data.
-Our methodology makes use of the fact that we have already set up our model and solving scheme and
-are able to produce cell masks from these numerical results.
-In doing so, we have exact knowledge about the position of our agents since we can simply obtain
-this information from our computational model.
-But on the other hand we can also try to extract these positions from the generated cell-masks.
-By comparing the differences between them, we get a good estimate of how well our extraction
-algorithm can determine the true position of the bacteria.
-
-In figure~\ref{fig:benchmarking-extraction-algorithm} (A-C) we see snapshots of a time series of
-masks.
-This particular simulation includes division events between images (B) and (C).
-For all images (A-C), we have drawn the exact position of the agents with white lines and indicated
-the extracted positions with crosses.
-They align very well visually.
-It may be hard to spot with the bare eye, but the agreement between extracted and exact positions is
-visually worse at the endpoints of the rods.
-To quantify this disparity, we calculate the distance between these individual vertices.
-Figure~\ref{fig:benchmarking-extraction-algorithm} (D) shows the estimated length of the rods from
-the extracted and exact vertices which is surrounded by the average vertex distance.
-Division events lead to steps in our rod-length function since the bacterial length is halved.
-We can clearly see that the agreement is very good in the beginning but right after every division
-event, the uncertainty becomes large.
-To illustrate this, we can see that in figure~\ref{fig:benchmarking-extraction-algorithm} (C) some
-of the rods are partially overlapping which hides parts of the individual cell-masks and thus
-changes the extraction result.
-We also plotted the distribution of individual vertex distances in
-figure~\ref{fig:benchmarking-extraction-algorithm} (E).
-Darker areas in the plot correspond to distances at time-points earlier in the simulatin time.
-We can see that brighter areas tend to shift towards the right, thus increasing the uncertainty of
-the extraction.
-However, the overall distribution shows small values of distances for a large number of vertices and
-is not dominated by outliers.
-
-\textbf{
-    Die Verständlichkeit und die Formulierungen hier bedürfen noch der Verbesserung.
-    Damit ich aber schneller durchkomme, verschieben wir das auf später.
-}
+To validate the reliability of the extraction algorithm
+(Algorithm~\ref{algo:extraction-algorithm}), we benchmarked it using synthetic data with known
+ground truth.
+This reference data in the form of vertices can be obtained directly from our computational model.
+Furthermore as we have explained in section~\ref{subsection:visualization}, we generate cell masks
+from a given set of vertices and thickness of the rod which will serve as inputs for our extraction
+algorithm.
+By comparing the extracted positions with known ground truth, we will obtain quantitative measures
+of the extraction accuracy.
+
+Figure~\ref{fig:benchmarking-extraction-algorithm} (D-H) shows snapshots of the synthetically
+generated cell masks.
+The initial configuration consists of 4 bacteria which grow and ultimately divide thus expanding
+the filled up area.
+Meanwhile figure~\ref{fig:benchmarking-extraction-algorithm} (I) shows a zoomed-in version of
+snapshot (G), highlighting the direct comparison of the exact (dark cross) and extracted vertices.
+Visual comparison of this time series indicates strong alignment between exact and extracted
+results.
+Upon closer inspection with figure~\ref{fig:benchmarking-extraction-algorithm} (I), we can see
+that small differences start to manifest, especially at the tip of the rods in more crowded regions.
+
+In order to quantitatively analyze this disparity we use both data to calculate the average rod
+length of the bacteria (Figure~\ref{fig:benchmarking-extraction-algorithm} (J)) over the course of
+the simulated time interval.
+The results show good agreement with larger discrepancies directly after division events.
+In addition figure~\ref{fig:benchmarking-extraction-algorithm} (K) displays individual displacements
+of exact and extracted vertices which have been reoriented with respect to the direction of the rod.
+Values closer to the center represent better agreement.
+The distribution of these displacements can be well described by a normal distribution with mean
+$\mu$ and variance $\sigma$.
+These parameters can be obtained by fitting the distribution to the displacements at every time
+point which is displayed in Figure~\ref{fig:benchmarking-extraction-algorithm}.
+Here we can also see that the overall variance in the displacements increases directly after
+division events.
+However, the center of the distributions remain concentrated near zero without significant outliers
+which demonstrates the robustness of our method in these densely packed scenarios.
+Further details are given in the supplementary
+materialsec:supsec:supplement-imaging-based-comparison-algorithm-details
+(section~\ref{sec:supplement-imaging-based-comparison-algorithm-details}).
 

### 2025-11-28 paper/main.tex
exchange order
fixup! rewrite text for benchmarking extraction algorithm

@@
 which demonstrates the robustness of our method in these densely packed scenarios.
-Further details are given in the supplementary
-materialsec:supsec:supplement-imaging-based-comparison-algorithm-details
+Further details are given in the supplementary materials
 (section~\ref{sec:supplement-imaging-based-comparison-algorithm-details}).
@@
 \textbf{Writing – original draft:} Jonas Pleyer\\
-\textbf{Writing – review \& editing:} Jonas Pleyer \& Christian Fleck
+\textbf{Writing – review \& editing:} Christian Fleck \& Jonas Pleyer
 

### 2025-11-28 paper/main.tex
update text for extraction algorithm

@@
 \label{subsection:visualization}
+Our visualization routine relies on just the spatial representation in the form of vertices and the
+thickness of the rod which is assumed to be uniform along the rod.
 To visualize the spatial configuration of the results we employ
@@
 generated cell masks.
-The initial configuration consists of 4 bacteria which grow and ultimately divide thus expanding
-the filled up area.
+The initial configuration consists of four bacteria which grow and divide thus expanding the
+occupied area.
 Meanwhile figure~\ref{fig:benchmarking-extraction-algorithm} (I) shows a zoomed-in version of
-snapshot (G), highlighting the direct comparison of the exact (dark cross) and extracted vertices.
+snapshot (G), highlighting the direct comparison of the exact (white cross) and extracted vertices.
 Visual comparison of this time series indicates strong alignment between exact and extracted
@@
 These parameters can be obtained by fitting the distribution to the displacements at every time
-point which is displayed in Figure~\ref{fig:benchmarking-extraction-algorithm}.
-Here we can also see that the overall variance in the displacements increases directly after
-division events.
-However, the center of the distributions remain concentrated near zero without significant outliers
-which demonstrates the robustness of our method in these densely packed scenarios.
+point.
+The parameters resulting from this fit are displayed in
+Figure~\ref{fig:benchmarking-extraction-algorithm} (L), where the overall variance (filled area) in
+the displacements significantly increases directly after division events.
+However, the centers of the distributions (solid curves) remain concentrated near zero without
+significant outliers which demonstrates the robustness of our method in these densely packed
+scenarios.
+These observations provide us with enough information to construct an error model for the extraction
+algorithm.
 Further details are given in the supplementary materials
 (section~\ref{sec:supplement-imaging-based-comparison-algorithm-details}).
-
-\paragraph{Determining the underlying Error Model}
 
@@
     \begin{tikzonimage}[width=0.1625\textwidth]
-        {docs/source/_static/fitting-methods/extract_positions-007800-cropped.pdf}%
+        {docs/source/_static/fitting-methods/extract_positions-009800-cropped.pdf}%
         \node at (0, 1)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{I}};
@@
         (D-H) Snapshots of simulated data used for benchmarking.
-        (I) Average rod length of the exact and extracted data.
+        (I) Zoomed-in version of (H).
+        (J) Average rod length of the exact and extracted data.
         The light blue region is the mean distance between vertices of extracted and exact
         positions which captures the inherent uncertainty of our method.
-        (J) Displacement of individual vertices parallel and orthogonal to the rod.
+        (K) Displacement of individual vertices parallel and orthogonal to the rod.
         The exact value of the vertex is always in the center.
@@
         exclude outliers.
-        (K) Mean $\mu_{p,o}$ and variance $\sigma_{p,o}$ of fitted normal distribution for
-        displacements parallel and orthogonal to the rod displayed over the duration of the
-        simulation.
+        (L) Mean $\mu_{p,o}$ (curve) and variance $\sigma_{p,o}$ (filled space) of fitted normal
+        distribution for displacements parallel and orthogonal to the rod displayed over the
+        duration of the simulation.
     }

### 2025-11-28 plots/fitting_extract_positions.py
update plots

@@ -175,6 +175,15 @@ if __name__ == "__main__":
                 p2 = crm.convert_pixel_to_position(
                     p0, config.domain_size, mask.shape[:2]
                 )
+                for pi in p1[:, ::-1]:
+                    mask = cv.drawMarker(
+                        mask,
+                        pi,
+                        (250, 250, 250),
+                        cv.MARKER_CROSS,
+                        14,
+                        2,
+                    )
                 mask = cv.polylines(
                     mask,
                     [p0[:, ::-1]],
@@ -182,13 +191,13 @@ if __name__ == "__main__":
                     color=(250, 250, 250),
                     thickness=1,
                 )
-                for pi in p1[:, ::-1]:
+                for qi in p0[:, ::-1]:
                     mask = cv.drawMarker(
                         mask,
-                        pi,
-                        (50, 50, 50),
+                        qi,
+                        (42, 112, 232),
                         cv.MARKER_TILTED_CROSS,
-                        14,
+                        9,
                         2,
                     )
 
@@ -311,6 +320,9 @@ if __name__ == "__main__":
             ax.set_yscale("log")
             ax.set_xlim(-1.1 * dx, 1.1 * dx)
 
+            ax.set_ylabel("Count")
+            ax.set_xlabel("Displacement [µm]")
+
             fig.savefig(OPATH / f"displacement-distr-{name}.png")
             fig.savefig(OPATH / f"displacement-distr-{name}.pdf")
             plt.close(fig)
@@ -386,7 +398,7 @@ if __name__ == "__main__":
         ax.plot(
             t,
             covs[:, 0, 1] / covs[:, 0, 0],
-            color=crm.plotting.COLOR4,
+            color=crm.plotting.COLOR3,
             label="$\\sigma_{01}/\\sigma_{00}$",
         )
         ax.set_xlabel("Time [min]")

### 2025-11-29 paper/main.tex
change plotting of overlays

@@
         {docs/source/_static/mechanics.pdf}
-        \node at (0, 1)[anchor=north west, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+        \node at (0, 1)[anchor=north west]{\textbf{A}};
     \end{tikzonimage}%
@@
         {docs/source/_static/interaction.pdf}
-        \node at (0, 1)[anchor=north west, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0, 1)[anchor=north west]{\textbf{B}};
     \end{tikzonimage}%
@@
         {docs/source/_static/mie-potential-shapes.pdf}
-        \node at (0, 1)[anchor=north west, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+        \node at (0, 1)[anchor=north west]{\textbf{C}};
     \end{tikzonimage}%
@@
         {docs/source/_static/imaging-mesh.pdf}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+        \node at (0.025, 0.975)[anchor=north west]{\textbf{A}};
     \end{tikzonimage}%
@@
         {docs/source/_static/11571737453049821261/raw_pv/000000400.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0.025, 0.975)[anchor=north west, white]{\textbf{B}};
     \end{tikzonimage}%
@@
         {docs/source/_static/11571737453049821261/masks/000000400.png}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+        \node at (0.025, 0.975)[anchor=north west, white]{\textbf{C}};
     \end{tikzonimage}
@@
         {docs/source/_static/fitting-methods/algorithm/image001042.png}
-        \node at (0.025, 0.975)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+        \node at (0.025, 0.975)[anchor=north west, white]{\textbf{A}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/algorithm/mask-zoom.pdf}
-        \node at (0.025, 0.975)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0.025, 0.975)[anchor=north west, white]{\textbf{B}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/algorithm/interpolate-positions.pdf}
-        \node at (0.025, 0.975)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+        \node at (0.025, 0.975)[anchor=north west, white]{\textbf{C}};
     \end{tikzonimage}\\
@@
         {docs/source/_static/fitting-methods/extract_positions-001800.pdf}
-        \node at (0.03, 0.99)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+        \node at (0.03, 0.99)[anchor=north west, white]{\textbf{D}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/extract_positions-003800.pdf}
-        \node at (0.03, 0.99)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+        \node at (0.03, 0.99)[anchor=north west, white]{\textbf{E}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/extract_positions-005800.pdf}
-        \node at (0.03, 0.99)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{F}};
+        \node at (0.03, 0.99)[anchor=north west, white]{\textbf{F}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/extract_positions-007800.pdf}
-        \node at (0.03, 0.99)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{G}};
+        \node at (0.03, 0.99)[anchor=north west, white]{\textbf{G}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/extract_positions-009800.pdf}
-        \node at (0.03, 0.99)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{H}};
+        \node at (0.03, 0.99)[anchor=north west, white]{\textbf{H}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/extract_positions-009800-cropped.pdf}%
-        \node at (0, 1)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{I}};
+        \node at (0, 1)[anchor=north west, white]{\textbf{I}};
     \end{tikzonimage}\\
@@
         {docs/source/_static/fitting-methods/displacement-calculations.pdf}%
-        \node at (0, 1)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{J}};
+        \node at (0, 1)[anchor=north west, black]{\textbf{J}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/displacement-distribution.pdf}
-        \node at (0, 1)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{K}};
+        \node at (0, 1)[anchor=north west, black]{\textbf{K}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/displacement-fit-over-time.pdf}
-        \node at (0, 1)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{L}};
+        \node at (0, 1)[anchor=north west, black]{\textbf{L}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/progressions-1.pdf}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+        \node at (0.025, 0.975)[anchor=north west, white]{\textbf{A}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/progressions-2.pdf}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0.025, 0.975)[anchor=north west, white]{\textbf{B}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/progressions-3.pdf}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+        \node at (0.025, 0.975)[anchor=north west, white]{\textbf{C}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/progressions-4.pdf}
-        \node at (0.025, 0.975)[anchor=north west, rectangle, white, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+        \node at (0.025, 0.975)[anchor=north west, white]{\textbf{D}};
     \end{tikzonimage}\\
@@
             {docs/source/_static/fitting-methods/extract_positionsdivision-comparison.pdf}
-            \node at (0, 1)[anchor=north west, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+            \node at (0, 1)[anchor=north west]{\textbf{E}};
         \end{tikzonimage}%
@@
             {docs/source/_static/fitting-methods/penalty-time-flow.pdf}
-            \node at (0, 0.975)[anchor=north west, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{F}};
+            \node at (0, 0.975)[anchor=north west]{\textbf{F}};
         \end{tikzonimage}%
@@
         {figures/crm_amir/progressions/step4-000010.png}
-        \node at (0.02,17pt)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+        \node at (0.02,17pt)[anchor=north west, white]{\textbf{A}};
     \end{tikzonimage}%
@@
         {figures/crm_amir/progressions/step4-000017.png}
-        \node at (0.02,17pt)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0.02,17pt)[anchor=north west, white]{\textbf{B}};
     \end{tikzonimage}%
@@
         {figures/crm_amir/progressions/step4-000034.png}
-            \node at (0.01, 0.01+17pt)[anchor=north west, rectangle, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+            \node at (0.01, 0.01+17pt)[anchor=north west, white]{\textbf{C}};
     \end{tikzonimage}\\
@@
         {figures/crm_amir/profiles-reduced/fit-comparison.pdf}%
-        \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0.13, 0.88)[anchor=north west, black]{\textbf{B}};
     \end{tikzonimage}%
@@
         {figures/crm_amir/profiles/drag_force.pdf}%
-        \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+        \node at (0.13, 0.88)[anchor=north west, black]{\textbf{D}};
     \end{tikzonimage}%
@@
         {figures/crm_amir/profiles/growth_rate.pdf}%
-        \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+        \node at (0.13, 0.88)[anchor=north west, black]{\textbf{E}};
     \end{tikzonimage}\\
@@
         {figures/crm_amir/profiles/rod_rigidity.pdf}%
-        \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
+        \node at (0.13, 0.88)[anchor=north west, black]{\textbf{F}};
     \end{tikzonimage}%
@@
         {figures/crm_amir/profiles/spring_tension.pdf}%
-        \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
+        \node at (0.13, 0.88)[anchor=north west, black]{\textbf{G}};
     \end{tikzonimage}%
@@
         {figures/crm_amir/profiles/damping.pdf}%
-        \node at (0.13, 0.88)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{H}};
+        \node at (0.13, 0.88)[anchor=north west, black]{\textbf{H}};
     \end{tikzonimage}%
@@
         {figures/data/crm_fit/001044.png}%
-        \node at (0.02, 0.98)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+        \node at (0.02, 0.98)[anchor=north west, white]{\textbf{A}};
     \end{tikzonimage}%
@@
         {figures/data/crm_fit/001044-markers.pdf}%
-        \node at (0.02, 0.98)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0.02, 0.98)[anchor=north west, white]{\textbf{B}};
     \end{tikzonimage}\\
@@
         {figures/crm_fit-data-progression.png}
-        \node at (0.02, 0.98)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+        \node at (0.02, 0.98)[anchor=north west, white]{\textbf{C}};
     \end{tikzonimage}%
@@
         {figures/crm_estimate_params/IWF-Goettingen/rod-lengths-individual.pdf}%
-        \node at (0, 1)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+        \node at (0, 1)[anchor=north west, black]{\textbf{D}};
     \end{tikzonimage}%
@@
         {figures/crm_estimate_params/IWF-Goettingen/growth-rates-bar-plot.pdf}%
-        \node at (0, 1)[anchor=north west, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+        \node at (0, 1)[anchor=north west, black]{\textbf{E}};
     \end{tikzonimage}%
@@
         {figures/crm_fit/snapshots/predicted-003889.png}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+        \node at (0.025, 0.975)[anchor=north west, white]{\textbf{A}};
     \end{tikzonimage}%
@@
         {figures/crm_fit/celldiffs/diff-003889.png}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0.025, 0.975)[anchor=north west, white]{\textbf{B}};
     \end{tikzonimage}%
@@
         {figures/crm_fit/profiles/damping.pdf}%
-        \node at (0, 1.1)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+        \node at (0, 1.1)[anchor=north west, black]{\textbf{A}};
     \end{tikzonimage}%
@@
         {figures/crm_fit/profiles/exponent-m.pdf}%
-        \node at (0, 1.1)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0, 1.1)[anchor=north west, black]{\textbf{B}};
     \end{tikzonimage}%
@@
         {figures/crm_fit/profiles/exponent-n.pdf}%
-        \node at (0, 1.1)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+        \node at (0, 1.1)[anchor=north west, black]{\textbf{C}};
     \end{tikzonimage}%
@@
         {figures/crm_fit/profiles/strength.pdf}%
-        \node at (0, 1.1)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+        \node at (0, 1.1)[anchor=north west, black]{\textbf{D}};
     \end{tikzonimage}\\
@@
         {figures/crm_fit/profiles/radius.pdf}%
-        \node at (0, 1.1)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+        \node at (0, 1.1)[anchor=north west, black]{\textbf{E}};
     \end{tikzonimage}%
@@
         {figures/crm_fit/profiles/growth-rate-0.pdf}%
-        \node at (0, 1.1)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
+        \node at (0, 1.1)[anchor=north west, black]{\textbf{F}};
     \end{tikzonimage}%
@@
         {figures/crm_fit/profiles/growth-rate-1.pdf}%
-        \node at (0, 1.1)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
+        \node at (0, 1.1)[anchor=north west, black]{\textbf{G}};
     \end{tikzonimage}%
@@
         {figures/crm_fit/profiles/growth-rate-2.pdf}%
-        \node at (0, 1.1)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{H}};
+        \node at (0, 1.1)[anchor=north west, black]{\textbf{H}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/masks_diff/000000.png}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+        \node at (0.025, 0.975)[anchor=north west, white]{\textbf{A}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/masks_diff/000010.png}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0.025, 0.975)[anchor=north west, white]{\textbf{B}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/masks_diff/000020.png}%
-        \node at (0.025, 0.975)[anchor=north west, rectangle, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+        \node at (0.025, 0.975)[anchor=north west, white]{\textbf{C}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/profiles/profile-000000.png}%
-        \node at (0.15, 0.85)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{A}};
+        \node at (0.15, 0.85)[anchor=north west, black]{\textbf{A}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/profiles/profile-000001.png}%
-        \node at (0.15, 0.85)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
+        \node at (0.15, 0.85)[anchor=north west, black]{\textbf{B}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/profiles/profile-000002.png}%
-        \node at (0.15, 0.85)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+        \node at (0.15, 0.85)[anchor=north west, black]{\textbf{C}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/profiles/profile-000003.png}%
-        \node at (0.15, 0.85)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+        \node at (0.15, 0.85)[anchor=north west, black]{\textbf{D}};
     \end{tikzonimage}\\
@@
         {figures/crm_divide/profiles/profile-000004.png}%
-        \node at (0.15, 0.85)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+        \node at (0.15, 0.85)[anchor=north west, black]{\textbf{E}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/profiles/profile-000005.png}%
-        \node at (0.15, 0.85)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
+        \node at (0.15, 0.85)[anchor=north west, black]{\textbf{F}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/profiles/profile-000006.png}%
-        \node at (0.15, 0.85)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
+        \node at (0.15, 0.85)[anchor=north west, black]{\textbf{G}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/profiles/profile-000007.png}%
-        \node at (0.15, 0.85)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{H}};
+        \node at (0.15, 0.85)[anchor=north west, black]{\textbf{H}};
     \end{tikzonimage}\\
@@
         {figures/crm_divide/profiles/profile-000008.png}%
-        \node at (0.15, 0.85)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{I}};
+        \node at (0.15, 0.85)[anchor=north west, black]{\textbf{I}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/profiles/profile-000009.png}%
-        \node at (0.15, 0.85)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{J}};
+        \node at (0.15, 0.85)[anchor=north west, black]{\textbf{J}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/profiles/profile-000010.png}%
-        \node at (0.15, 0.85)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{K}};
+        \node at (0.15, 0.85)[anchor=north west, black]{\textbf{K}};
     \end{tikzonimage}%
@@
         {figures/crm_divide/profiles/profile-000011.png}%
-        \node at (0.15, 0.85)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{L}};
+        \node at (0.15, 0.85)[anchor=north west, black]{\textbf{L}};
     \end{tikzonimage}%
@@
     \begin{tikzonimage}[width=0.5\textwidth]
+        {docs/source/_static/fitting-methods/displacement-distribution.pdf}
+        \node at (0, 1)[anchor=north west, black]{\textbf{A}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.5\textwidth]
+        {docs/source/_static/fitting-methods/displacement-fit-over-time-covariance.pdf}
+        \node at (0, 1)[anchor=north west, black]{\textbf{B}};
+    \end{tikzonimage}\\
+    \begin{tikzonimage}[width=0.5\textwidth]
         {docs/source/_static/fitting-methods/displacement-distr-x.pdf}
-        \node at (0.03, 0.99)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
+        \node at (0, 1)[anchor=north west, black]{\textbf{C}};
     \end{tikzonimage}%
@@
         {docs/source/_static/fitting-methods/displacement-distr-y.pdf}
-        \node at (0.03, 0.99)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
-    \end{tikzonimage}\\
-    \begin{tikzonimage}[width=0.5\textwidth]
-        {docs/source/_static/fitting-methods/displacement-fit-over-time-covariance.pdf}
-        \node at (0.03, 0.99)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{H}};
-    \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.5\textwidth]
-        {docs/source/_static/fitting-methods/displacement-distribution.pdf}
-        \node at (0.03, 0.99)[anchor=north west, rectangle, black, minimum width=15pt, minimum height=15pt]{\textbf{I}};
+        \node at (0, 1)[anchor=north west, black]{\textbf{D}};
     \end{tikzonimage}
@@
         {figures/crm_amir/progressions/step4-000034.png}
-        \node at (0.02,17pt)[anchor=north west, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
+        \node at (0.02,17pt)[anchor=north west, white]{\textbf{C}};
     \end{tikzonimage}\\
@@
         {figures/crm_amir/result-full/0000000000.png}
-        \node at (0.02, 0.98)[anchor=north west, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{D}};
+        \node at (0.02, 0.98)[anchor=north west]{\textbf{D}};
     \end{tikzonimage}%
@@
         {figures/crm_amir/result-full/0000000001.png}
-        \node at (0.02, 0.98)[anchor=north west, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{E}};
+        \node at (0.02, 0.98)[anchor=north west]{\textbf{E}};
     \end{tikzonimage}%
@@
         {figures/crm_amir/result-full/0000000002.png}
-        \node at (0.02, 0.98)[anchor=north west, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{F}};
+        \node at (0.02, 0.98)[anchor=north west]{\textbf{F}};
     \end{tikzonimage}%

### 2025-11-29 paper/main.tex
include time point in comparison

@@
 Each cell contains a collection of vertices.
-Given $M$ cells with $N$ vertices each and dimension $d$, the total information can be stored in an
-array $\nu_{i,j,k}$ of shape $(M,N,2)$.
+Given $N_c$ cells with $N_v$ vertices in $d$ dimensions at $N_t$ time points, the total information
+can be stored in an array $\nu_{i,j,k,l}$ of shape $(N_c,N_v,d,N_t)$.
 Comparing the results of a simulation $\nu$ with experimental values $\tilde{\nu}$ is thus only a
@@
     L_{\text{LS}}(\vec{x}, \vec{y})
-        = \sum\limits_{i,j,k}|\nu_{i,j,k} - \tilde{\nu}_{i,j,k}|^2
+        = \sum\limits_{i,j,k,l}|\nu_{i,j,k,l} - \tilde{\nu}_{i,j,k,l}|^2
 \end{equation}
 
-which minimizes the distance between the corresponding vertices.
+which minimizes the distance between the corresponding vertices at each time point.
 One important detail is that the ordering of vertices must be identical in the numerical and

### 2025-11-29 paper/main.tex
link to supplement; fill supplement; reorder some parts

@@
 reader to the supplementary
-material~\ref{sec:supplement-imaging-based-comparison-algorithm-details}.
+material~\ref{sec:supplement-comparison-algorithms}.
 
@@
 
+\subsection{Further Details of the Extraction Algorithm}
+\label{sec:supplement-extraction-algorithm}
+
+The main text stated that the distribution of displacements which has been split up into parallel
+and orthogonal components along the rod can be represented by a multivariate normal distribution.
+We first describe in more detail how these distributions are obtained.
+Consider the case where we have a set of exactly known vertices $\nu_{j,k}$ and a set of extracted
+vertices $\nu_{j,k}'$.
+We now calculate the displacements simply via
+\begin{equation}
+    \Delta \nu_{j,k} = \nu_{j,k} - \nu_{j,k}'.
+\end{equation}
+In order to calculate their positions with respect to the direction of the rod, we assign a
+direction $d_{j,k}$ at each vertex of the known exact values.
+At the tips of the rod this is done by assigning the direction of the closest edge while vertices in
+the middle are calculated using a central difference approach.
+\begin{align}
+    d_{0,k} &= \frac{(\nu_{1,k}-\nu_{0,k})}{||\nu_1 - \nu_0||}\\
+    d_{N_v,k} &= \frac{(\nu_{N_v,k}-\nu_{N_v-1,k})}{||\nu_{N_v} - \nu_{N_v-1}||}\\
+    d_{j,k} &= \frac{(\nu_{j+1,k}-\nu_{j-1,k})}{||\nu_{j+1} - \nu_{j-1}||}
+\end{align}
+We can use these directions at each vertex to assign rotation matrices
+\begin{equation}
+    M_{j,l}^k = \begin{pmatrix}
+        d_{j,0} & d_{j,1}\\
+        -d_{j,1} & d_{j,0}
+    \end{pmatrix}
+\end{equation}
+at each vertex and apply them to the displacements via
+\begin{equation}
+    (\Delta\nu_{j,k})_\text{directed} = \sum\limits_l M_{j,l}^k \Delta\nu_{j,l}.
+\end{equation}
+The resulting points are displayed in Figure~\ref{fig:supplement-extraction-algorithm} (A) which is
+identical to the one in the main text.
+
+The main text claims that this distribution is well described by a multimodal normal distribution
+with diagonal covariance matrix.
+Figure~\ref{fig:supplement-extraction-algorithm} (C,D) show slices of the full distribution along
+the parallel and orthogonal axis.
+Darker colors indicate earlier stages of the simulation while lighter colors indicate later stages.
+Visually, we can already see that the distribution is quite narrow initially but widens with
+increased simulation time.
+To quantify this behaviour, we fit a multimodal normal distribution at each time point to the
+distribution of displacements.
+The main text has already discussed results for the variances $\sigma_{00}$ and $\sigma_{11}$.
+Figure~\ref{fig:supplement-extraction-algorithm} (B) shows the trajectory of the covariance
+$\sigma_{01}=\sigma_{10}$ which has been normalized to the value of $\sigma_{00}$.
+Its small relative value justifies the statement in the main text and shows no correlation between
+parallel and orthogonal displacement values.
+
+\begin{itemize}
+    \item for low resolution pictures this can be a problem
+    \item in our benchmarking it is not important since we chose large resolution where such
+        artifacts can not occur
+\end{itemize}
+
+\begin{figure}[H]
+    \centering
+    \begin{tikzonimage}[width=0.5\textwidth]
+        {docs/source/_static/fitting-methods/displacement-distribution.pdf}
+        \node at (0, 1)[anchor=north west, black]{\textbf{A}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.5\textwidth]
+        {docs/source/_static/fitting-methods/displacement-fit-over-time-covariance.pdf}
+        \node at (0, 1)[anchor=north west, black]{\textbf{B}};
+    \end{tikzonimage}\\
+    \begin{tikzonimage}[width=0.5\textwidth]
+        {docs/source/_static/fitting-methods/displacement-distr-x.pdf}
+        \node at (0, 1)[anchor=north west, black]{\textbf{C}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.5\textwidth]
+        {docs/source/_static/fitting-methods/displacement-distr-y.pdf}
+        \node at (0, 1)[anchor=north west, black]{\textbf{D}};
+    \end{tikzonimage}
+    \caption{TODO}
+    \label{fig:supplement-extraction-algorithm}
+\end{figure}
+
 %###################################################################################################
 \subsection{Details for the Imaging-Based Comparison Algorithm}
-\label{sec:supplement-imaging-based-comparison-algorithm-details}
+\label{sec:supplement-comparison-algorithms}
 
@@
 event has occured.
-
-\begin{figure}
-    \centering
-    \begin{tikzonimage}[width=0.5\textwidth]
-        {docs/source/_static/fitting-methods/displacement-distribution.pdf}
-        \node at (0, 1)[anchor=north west, black]{\textbf{A}};
-    \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.5\textwidth]
-        {docs/source/_static/fitting-methods/displacement-fit-over-time-covariance.pdf}
-        \node at (0, 1)[anchor=north west, black]{\textbf{B}};
-    \end{tikzonimage}\\
-    \begin{tikzonimage}[width=0.5\textwidth]
-        {docs/source/_static/fitting-methods/displacement-distr-x.pdf}
-        \node at (0, 1)[anchor=north west, black]{\textbf{C}};
-    \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.5\textwidth]
-        {docs/source/_static/fitting-methods/displacement-distr-y.pdf}
-        \node at (0, 1)[anchor=north west, black]{\textbf{D}};
-    \end{tikzonimage}
-    \caption{TODO}
-\end{figure}
 

### 2025-11-29 paper/main.tex
say that this is a multivariate distribution

@@
 Values closer to the center represent better agreement.
-The distribution of these displacements can be well described by a normal distribution with mean
-$\mu$ and variance $\sigma$.
+The distribution of these displacements can be well described by a multivariate normal
+distribution with mean $\mu$ and variances $\sigma_{ij}$.
 These parameters can be obtained by fitting the distribution to the displacements at every time
@@
 algorithm.
-Further details are given in the supplementary materials
-(section~\ref{sec:supplement-imaging-based-comparison-algorithm-details}).
+Further details such as information about the covariances are given in the supplementary materials
+(section~\ref{sec:supplement-extraction-algorithm}).
 
@@
 
-The main text claims that this distribution is well described by a multimodal normal distribution
+The main text claims that this distribution is well described by a multivariate normal distribution
 with diagonal covariance matrix.
@@
 increased simulation time.
-To quantify this behaviour, we fit a multimodal normal distribution at each time point to the
+To quantify this behaviour, we fit a multivariate normal distribution at each time point to the
 distribution of displacements.

### 2025-11-29 paper/main.tex
finish supplementary section for extraction algorithm

@@
 
-\begin{itemize}
-    \item for low resolution pictures this can be a problem
-    \item in our benchmarking it is not important since we chose large resolution where such
-        artifacts can not occur
-\end{itemize}
+Another detail is the resolution of the utilized image.
+Situations of high resolution generally present no concerns but images with low resolutions where a
+single bacterium is only made up of a few pixels in its diameter can present problems.
+Firstly, the skeletonization algorithm will not be able to accurately determine the middle of the
+rod due to the discreteness of the pixel-grid.
+Secondly, the same discreteness presents a minimal error threshold which should probably always be
+considered.
+Within our performed benchmarks, we did not run into such situations since we chose large enough
+resolutions.
+In general, we do not anticipate that these scenarios are common for most practical applications.
 
@@
     \end{tikzonimage}
-    \caption{TODO}
+    \caption{
+        (A) Distribution of displacements between exact and extracted vertices parallel and
+        orthogonal to the rod.
+        (B) Covariance $\sigma_{01}$ normalized by $\sigma_{00}$ over the course of the simulation.
+        (C,D) Distributions of displacements sliced parallel (C) and orthogonal (D) to the rod.
+        Darker colors indicate earlier stages of the simulation while brighter colors indicate later
+        stages.
+    }
     \label{fig:supplement-extraction-algorithm}

### 2025-11-29 paper/main.tex
rename comparison metrics

@@
 \newacro{bsubtilis}[\textit{B. subtilis}]{\textit{Bacillus subtilis}}
+\newacro{ls}[LS]{Least Squares}
 %
@@
 models.
-Furthermore, we utilize them in our more advanced imaging-based comparison methods (see
-section~\ref{subsec:imaging-based-comparison}) where they enable us to effectively compare
+Furthermore, we utilize them in our image-space metric (see
+section~\ref{subsec:image-space-metric}) where they enable us to effectively compare
 two bacterial configurations while accounting for division events.
@@
 Furthermore, it is an \ac{abm} which is able to natively track mother-daughter relations of cells
-over the course of a numerical simulation which is an integral component of our imaging-based
-comparison algorithm (see section~\ref{subsec:imaging-based-comparison}).
+over the course of a numerical simulation which is an integral component of our image-space metric
+comparison algorithm (see section~\ref{subsec:image-space-metric}).
 Further details can be found in the supplement (see section~\ref{section:supplement-performance})
@@
 It is split into 3 parts, each covering distinct purposes.
-We begin by explaining the visualization process which will later be used within our imaging-based
-comparison algorithm (see section~\ref{subsec:imaging-based-comparison}) and for testing purposes.
+We begin by explaining the visualization process which will later be used within our image-space
+metric (see section~\ref{subsec:image-space-metric}) and for testing purposes.
 Afterwards, we introduce a position extraction algorithm which extracts cellular information in the
@@
 
-\subsubsection{Direct Comparison}
-\label{subsec:direct-comparison}
+\subsubsection{Direct Skeleton-Space Metric}
+\label{subsec:skeleton-metric}
 In the scenario where no cells divide during the observed interval, we can use our extraction
-algorithm to obtain the position of the cells and then compare them directly to the positions
-calculated from the computational results.
-Each cell contains a collection of vertices.
+algorithm to obtain the position of the cells and then compare them directly to the cellular
+skeletons obtained from the computational results.
 Given $N_c$ cells with $N_v$ vertices in $d$ dimensions at $N_t$ time points, the total information
@@
 Comparing the results of a simulation $\nu$ with experimental values $\tilde{\nu}$ is thus only a
-matter of comparing two of these arrays within a suitable cost function~\cite{Wang2020}.
-In our approach we utilize a least squares approach given by
+matter of comparing two of these arrays via a suitable cost function~\cite{Wang2020}.
+In our approach we utilize a \ac{ls} approach given by
 
@@
 experimental data at this time point yet (or vice-versa).
-To explain why the previous developed direct comparison algorithm can not be reused on confined
+To explain why the direct skeleton-space metric can not be reused on confined
 intervals which do not include division events and for further details, we refer the interested
-reader to the supplementary
-material~\ref{sec:supplement-comparison-algorithms}.
+reader to the supplementary material~\ref{sec:supplement-comparison-algorithms}.
 
@@
 the direct comparison algorithm while the second case contains division events and in turn requires
-usage of the imaging-based comparison algorithm.
+usage of the image-space metric.
 
@@
 Figure~\ref{fig:amir-bending-simulation} (A)) and compare the simulated positions with the extracted
-ones using the direct comparison algorithm (see Subsection~\ref{subsec:direct-comparison}).
+ones using the direct comparison algorithm (see Subsection~\ref{subsec:skeleton-metric}).
 Our simulation is split into two phases; One with the external flow force that initially bends the
@@
 Therefore, the first approach will optimize the direct comparison cost function (see
-section~\ref{subsec:direct-comparison}) while the latter approach has to account for division events
-and thus requires the imaging-based method (see section~\ref{subsec:imaging-based-comparison}).
+section~\ref{subsec:skeleton-metric}) while the latter approach has to account for division events
+and thus requires the image-space metric method (see section~\ref{subsec:image-space-metric}).
 These two scenarios present an opportunity to apply the previously developed methods within a
@@
 By doing so, we can utilize the simpler and more performant direct comparison scheme (see
-\ref{subsec:direct-comparison}) for our cost function.
+\ref{subsec:skeleton-metric}) for our cost function.
 We have chosen said interval such that effects related to the lag phase can be neglected which
@@
         (B) Differences between extracted and predicted positions.
-        Note that the imaging-based comparison algorithm was not used to optimize this system but is
+        Note that the image-space metric was not used to optimize this system but is
         chosen here as a tool to aid visualization of the prediction difference.
@@
 In the next step, we extend our previously used time-series with successive images including
-cell division events which means that we now have to use the imaging-based comparison algorithm (see
+cell division events which means that we now have to use the image-space metric (see
 section~\ref{section:parameter-estimation}).
@@
 
-The imaging-based comparison algorithm is computationally much more expensive compared to the
-simpler position-comparison algorithm.
+The image-space metric is computationally much more expensive compared to the direct method.
 There are 4 main steps which are executed within any single optimization iteration that contribute
@@
 generating masks from computational results is the limiting factor for the performance of the
-imaging-based comparison algorithm.
+image-space metric.
 This problem could be targeted by rewriting existing code in C++ using native VTK~\cite{vtkBook}
@@
 %###################################################################################################
-\subsection{Details for the Imaging-Based Comparison Algorithm}
+\subsection{Details for the Image-Space Metric}
 \label{sec:supplement-comparison-algorithms}

### 2025-11-29 paper/main.tex
restructure acknowledgements

@@
 
-\section{Author Contributions}
+%###################################################################################################
+\section{Acknowledgements}
+\subsection{Author Contributions}
 \textbf{Conceptualization:} Jonas Pleyer\\
@@
 
-%###################################################################################################
+\subsection{Declaration of AI Use}
+The authors acknowledge the use of Gemini (Google) for assistance with language editing and
+proofreading.
+In particular, no generative aspect was used and each suggested change was carefully reviewed and
+utilized as additional feedback source.
+Every text was written and audited by the authors which is why they take full responsibility for the
+contents presented in this publication.
+
 \section{Data and Code Availability}

### 2025-11-29 paper/main.tex
rewrite imaging-based metric subsection

--- 2025-11-29 paper/main.tex
@@
 
-\subsubsection{Imaging-based Comparison Algorithm}
-\label{subsec:imaging-based-comparison}
+\subsubsection{Imaging-Space Metric}
+\label{subsec:image-space-metric}
+\paragraph{Motivation}
 The preceding algorithm is performant and effective but can not be applied in the case when a
-division event splits apart the bacterium since this would involve the comparison of two sets of
-vertices with only one.
-This configuration is displayed in figure~\ref{fig:mask-difference-metric} (E) where two distinct
-bacterial configurations originate from the same initial state but contain different amounts of
-bacteria.
-This mismatch can happen, e.g. due to non-matching division times.
+division event splits apart the bacterium since this would require comparing sets of unequal
+cardinality.
+This situation is displayed in figure~\ref{fig:mask-difference-metric} (E) where two distinct
+bacterial configurations originate from the same initial state but contain varying numbers of
+bacteria at the next time step.
+Such a mismatch can happen, e.g. due to non-matching division times of model and data.
 Thus a cell which is present in the simulation might not yet have a corresponding cell in the
 experimental data at this time point yet (or vice-versa).
-To explain why the direct skeleton-space metric can not be reused on confined
-intervals which do not include division events and for further details, we refer the interested
-reader to the supplementary material~\ref{sec:supplement-comparison-algorithms}.
-
-In order to be able to compare the differing cellular configurations, we require additional
-methods which can be used in time intervals containing cellular divison events.
-To this end we use the numerical results of the \ac{abm} simulation and our visualization
-algorithms to generate cell-masks (see section~\ref{subsection:visualization}).
-Since each pixel has a unique color value which is directly related to one particular cell or
-background (black), we compare the generated masks to the experimentally obtained ones pixel by
-pixel.
-Whenever two colors are not matching, a penalty is assigned for each pixel.
+
+In addition to these technical challenges come practical ones.
+Within a short time period after the division event, the two resulting daughter cells will cover
+the space that the mother had taken up.
+However, the exact time point for the fission event can not be determined accurately in practice,
+thus leading to natural uncertainties with respect to the question when the lives of the daughter
+cells begin and the one of the mother ends.
+Such uncertainties which are also common problems for segmentation methods, further highlight the
+importance of a purely imaging-based comparison scheme.
+
+These thoughts explain why the direct metric can not be simply reused on confined intervals which
+do not include division events:
+It does not account for the inaccuracy of the division timing.
+Furthermore, finding such confined intervals becomes increasingly difficult for larger numbers of
+bacteria where division events occur much more frequently.
+This fundamental limitation will be circumvented by the imaging-based method which we propose.
+
+\paragraph{The Metric}
+We develop an approach which can be understood as a metric in image space where we quantify
+differences in the visual footprint of two bacterial configurations.
+Within the realm of image segmentation many metrics such as the Jaccard Index~\cite{Murphy1996},
+Dice-Sørensen Coefficient~\cite{Dice1945}, Precision and Recall~\cite{Takaya2015} or F1
+Score~\cite{Taha2015} are actively used to deal with similar questions.
+However, these statistical methods are all based upon the notion of identifying matching and
+non-matching regions.
+It is therefore first necessary to define a criterion for this.
+
+% To compare the differing cellular configurations, we require additional
+% methods which can be used in time intervals containing cellular divison events.
+We can use the previously introduced \nameref{subsection:visualization} methods to obtain cell-masks
+from the numerical results of the \ac{abm} simulation.
+Since we utilize an in-image encoding scheme, each pixel has a unique color value which is directly
+related to one particular cell or background (black).
+We can now compare the two masks pixel by pixel.
+Whenever two colors are not matching, a penalty is assigned for each point in the pixel-grid.
 The total cost is the sum of all penalties.
-% This algorithm is an implementation of a signed area measure.
-It is important to take into account the parental relationship between the cells, by checking if
-either one of the colors corresponding to the compared colors is a daughter of the other cell.
-Figure~\ref{fig:mask-difference-metric} (A,B) shows two successive states of a numerical simulation
+With this approach it is important to take into account the parental relationship between the cells,
+by checking if either one of the colors corresponding to the compared cells is a daughter of the
+other cell.
+Figure~\ref{fig:mask-difference-metric} (A,B) shows two states of a numerical simulation
 with varying amounts of cells.
-We can see that some cells have divided while others have only grown without completing a division
-step.
+We can see that some cells have divided while others have only extended in the same time interval.
 The resulting penalties without accounting for the parental relationships are displayed in
 Figure~\ref{fig:mask-difference-metric} (C).
-Black indicates that the pixels are matching and no penalty is assigned while a white pixel
+Black indicates that the pixels are matching and no penalty is assigned while white
 indicates that the original pixels are not matching, thus adding to the overall cost.
@@
 account and highlights regions where compared pixels belong to mother-daughter cells in gray.
-We assign a penalty $0\leq p \leq 1$ which is a hyperparameter for this optimization algorithm.
+We assign a specific parent penalty $0\leq p\leq 1$ value for these points which should be seen as
+an additional hyperparameter for this optimization algorithm.
 % It is possible to assign a custom value for the parental cost value.
 Settings its value to $0$ results in a contiuous cost function in time.
-
-The developed optimization scheme was mainly driven by technical challenges, but it is also
-biologically motivated.
-In the short time period after the division event, the two resulting daughter cells will cover
-the space that the mother had taken up.
-Often times, the exact time at which the fission event has commenced can not be determined
-accurately, thus leading to uncertainties with respect to the question when the life of the daughter
-cells begins and the one of the mother ends.
-These uncertainties which are also common problems for segmentation methods, further highlight the
-importance of this comparison scheme.
+In the future, more complex choices for this parent penalty might be feasible such as
+time-dependence $p(t-t_\text{division})$.
+
+While our primary concern is to compare numerically obtained results with experimental data,
+this approach allows us to compare all spatial configurations which can be represented
+by segmented microscopic images irrespective of how these are obtained.
+On a higher level, this means that our method is agnostic to the underlying cellular representation.
+This means that other models who even utilize very distinct cellular representations but are able to
+render segmented images can reuse our method.
 
 \subsubsection{Testing the Comparison Algorithms with Division Events}
-When utilizing a computational model (see section~\ref{sec:computational-model}), we can generate
+When utilizing a computational model (see \ref{sec:computational-model}), we can generate
 data and use the generated masks to provide a benchmark of our algorithm.
@@
 
-%###################################################################################################
+\subsection{Declaration of AI Use}
+The authors acknowledge the use of Gemini (Google) for assistance with language editing and
+proofreading.
+In particular, no generative aspect was used and each suggested change was carefully reviewed and
+utilized only as additional feedback source.
+All contents of the paper are written and audited by the authors which is why they take full
+responsibility for the contents presented in this publication.
+
 \section{Data and Code Availability}

### 2025-11-29 paper/main.tex
restructure appendix to fit plots template

--- 2025-11-29 paper/main.tex
@@
 
+\newpage
 %###################################################################################################
-\section{Acknowledgements}
-\subsection{Author Contributions}
-\textbf{Conceptualization:} Jonas Pleyer\\
-\textbf{Investigation:} Jonas Pleyer\\
-\textbf{Methodology} Jonas Pleyer\\
-\textbf{Software:} Jonas Pleyer\\
-\textbf{Supervision:} Christian Fleck\\
-\textbf{Writing – original draft:} Jonas Pleyer\\
-\textbf{Writing – review \& editing:} Christian Fleck \& Jonas Pleyer
-
-\subsection{Declaration of AI Use}
-The authors acknowledge the use of Gemini (Google) for assistance with language editing and
-proofreading.
-In particular, no generative aspect was used and each suggested change was carefully reviewed and
-utilized only as additional feedback source.
-All contents of the paper are written and audited by the authors which is why they take full
-responsibility for the contents presented in this publication.
-
-\section{Data and Code Availability}
-\label{section:data-code-availability}
-All data and code which has been used in this study is available at the github repository
-\url{https://github.com/jonaspleyer/cr\_mech\_coli}.
-It includes detailed descriptions of how to obtain the used data and bring it into a usable format.
-The package which bundles the developed funtionality is published at
-\url{https://pypi.org/project/cr-mech-coli/} from which it is readily available as a python
-dependency.\\
-The time series of microscopic images which has been studied in section~\ref{section:application-2}
-was extracted from individual frames of a video~\cite{Trude1982} that
-showcases the exponential growth of bacteria.
-
-% \bibliographystyle{IEEEtran}
-\bibliography{references}
-
-\newpage
-\appendix
+% \addcontentsline{toc}{chapter}{Supporting Information}
+% \section*{Supplementary Material}
+% \renewcommand{\thesection}{S\arabic{section}}
+% \renewcommand{\thesubsection}{S\arabic{subsection}}
+
+\section*{Supporting Information}
+
 %###################################################################################################
-\renewcommand{\thesection}{}
-\renewcommand{\thesubsection}{S\arabic{subsection}}
-\section{Supplementary Material}
-
-%###################################################################################################
-\subsection{Alternative Interaction Potential - Morse Potential}
+\paragraph*{Alternative Interaction Potential - Morse Potential}
 \label{sec:supplement-morse-potential}
@@
 
-\subsection{Further Details of the Extraction Algorithm}
+\paragraph*{Further Details of the Extraction Algorithm}
 \label{sec:supplement-extraction-algorithm}
@@
 %###################################################################################################
-\subsection{Analysis \& Usage of heterogeneous Parameter Value (TODO)}
+\paragraph*{Analysis \& Usage of heterogeneous Parameter Value (TODO)}
 In order to reuse this model, we also provide the option to assign growth rates by sampling from a
@@
 %###################################################################################################
-\subsection{Application Data \& Preparation (TODO)}
+\paragraph*{Application Data \& Preparation (TODO)}
 \label{sec:supplement-data-details}
@@
 %###################################################################################################
-\subsection{Parameter Estimation Results (TODO)}
+\paragraph*{Parameter Estimation Results (TODO)}
 \label{sec:supplement-parameter-profiles}
@@
 %###################################################################################################
-\subsection{Implementation \& Performance (TODO)}
+\paragraph*{Implementation \& Performance (TODO)}
 \label{section:supplement-performance}
@@
 
+%###################################################################################################
+\section*{Acknowledgements}
+\paragraph{Author Contributions}
+\textbf{Conceptualization:} Jonas Pleyer\\
+\textbf{Investigation:} Jonas Pleyer\\
+\textbf{Methodology} Jonas Pleyer\\
+\textbf{Software:} Jonas Pleyer\\
+\textbf{Supervision:} Christian Fleck\\
+\textbf{Writing – original draft:} Jonas Pleyer\\
+\textbf{Writing – review \& editing:} Christian Fleck \& Jonas Pleyer
+
+\paragraph*{Declaration of AI Use}
+The authors acknowledge the use of Gemini (Google) for assistance with language editing and
+proofreading.
+In particular, no generative aspect was used and each suggested change was carefully reviewed and
+utilized only as additional feedback source.
+All contents of the paper are written and audited by the authors which is why they take full
+responsibility for the contents presented in this publication.
+
+\section*{Data and Code Availability}
+\label{section:data-code-availability}
+All data and code which has been used in this study is available at the github repository
+\url{https://github.com/jonaspleyer/cr\_mech\_coli}.
+It includes detailed descriptions of how to obtain the used data and bring it into a usable format.
+The package which bundles the developed funtionality is published at
+\url{https://pypi.org/project/cr-mech-coli/} from which it is readily available as a python
+dependency.\\
+The time series of microscopic images which has been studied in section~\ref{section:application-2}
+was extracted from individual frames of a video~\cite{Trude1982} that
+showcases the exponential growth of bacteria.
+
+\nolinenumbers
+\newpage
+\bibliography{references}
+
 \end{document}

### 2025-11-29 paper/main.tex
remove supplement part of imaging-based metric

@@
 %###################################################################################################
-\subsection{Details for the Image-Space Metric}
-\label{sec:supplement-comparison-algorithms}
-
-In a different approach, one might be tempted to utilize the previously constructed direct
-comparison algorithm for time intervals that do not contain any division events.
-But this attempt would quickly lead to multiple problems.
-It will not be able to deal with the uncertainties of the timing of the division event.
-As the number of investigated cells grows, more division events will occurr even in a fixed
-amount of time thus greatly decreasing or completely vanishing the effective interval at which no
-division event is observed.
-The latter situation is equivalent to two microscopic images as input data between which a division
-event has occured.
-
-%###################################################################################################
-\subsection{Experimental Setup of Amir et. al}
+\paragraph*{Experimental Setup of Amir et. al}
 

### 2025-11-29 paper/main.tex
use nameref instead of ref

@@
 can also calculate their lengths at these time-points and thus determine their rate of growth and
-type of growth (see section~\ref{paragraph:parameter-estimation-individual-treatment}), thus
+type of growth (see \nameref{paragraph:parameter-estimation-individual-treatment}), thus
 effectively guiding our modeling steps.
@@
         Every component implicitly builds on the spatial representation of rod-shaped bacteria as
-        vertices (see section~\ref{subsec:spatial-representation}).
+        vertices (see \nameref{subsec:spatial-representation}).
         Red boxes highlight the key scientific drivers of this study while yellow boxes indicate
@@
         In order for our methods to work, we are really only required to satisfy the assumption of
-        our spatial representation (see section~\ref{subsec:spatial-representation}) and thus, these
+        our spatial representation (see \nameref{subsec:spatial-representation}) and thus, these
         assumptions can also be loosened if required.
@@
 customizable due to the lower amount of parameters and more restrictive mathematical formulation
-(see section~\ref{sec:supplement-morse-potential}).
+(see \nameref{sec:supplement-morse-potential}).
 Therefore, the Mie potential is a good choice for a simple, interpretable interaction potential with
@@
 over the course of a numerical simulation which is an integral component of our image-space metric
-comparison algorithm (see section~\ref{subsec:image-space-metric}).
-Further details can be found in the supplement (see section~\ref{section:supplement-performance})
+comparison algorithm (see \nameref{subsec:image-space-metric}).
+Further details can be found in the supplement (see \nameref{section:supplement-performance})
 
@@
 We begin by explaining the visualization process which will later be used within our image-space
-metric (see section~\ref{subsec:image-space-metric}) and for testing purposes.
+metric (see \nameref{subsec:image-space-metric}) and for testing purposes.
 Afterwards, we introduce a position extraction algorithm which extracts cellular information in the
@@
 This section explains how we can transform a segmented microscopic image such that we obtain a
-collection of vertices (see Section~\ref{subsection:mechanical-model-mechanics}) that can be used
+collection of vertices (see \nameref{subsection:mechanical-model-mechanics}) that can be used
 to initialize our agents and to perform parameter estimation.
@@
 Further details such as information about the covariances are given in the supplementary materials
-(section~\ref{sec:supplement-extraction-algorithm}).
+(see \nameref{sec:supplement-extraction-algorithm}).
 
@@
         A complete list of snapshots is included in the supplementary material (see
-        section~\ref{sec:supplement-data-details}).
+        \nameref{sec:supplement-data-details}).
         (D) Growth curves of individual bacteria including a discrete delay before exponential
@@
 masks ($14.49(19)\%$), mask rendering for the numerical results ($69.2(1.4)\%$) and finally
-comparison of the masks ($0.68(5)\%$) (see section \ref{section:supplement-performance}).
+comparison of the masks ($0.68(5)\%$) (see section \nameref{section:supplement-performance}).
 The second step alters the masks provided by the data for every individual numerical simulation.

### 2025-11-29 paper/main.tex
update labels

@@
 \subsubsection{With Division}
+\label{sec:application-1-with-division}
 
@@
 simulation, using the built-in parallelization methods (see Supplementary Material,
-Figure~\ref{fig:performance}).
+Figure~\ref{fig:timings-crm_divide}).
 

### 2025-11-29 paper/main.tex
split figure; extend caption

@@
     \label{fig:parameter-estimates-supplement-growth-rates}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{figures/crm_divide/time-evolution.pdf}%
+    \caption{TODO}
 \end{figure}
@@
     \centering
-    \includegraphics[width=0.49\textwidth]
+    \begin{tikzonimage}[width=0.5\textwidth]
         {docs/source/_static/performance/computation-time-with-initial-agents.pdf}
-    \includegraphics[width=0.49\textwidth]{docs/source/_static/performance/flamegraph.pdf}
-    \caption{TODO}
-    \label{fig:performance}
-\end{figure}
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.5\textwidth]{figures/crm_divide/timings.pdf}%
-    \includegraphics[width=0.5\textwidth]{figures/crm_divide/time-evolution.pdf}%
-    \caption{TODO}
+        \node at (0, 1)[anchor=north west]{\textbf{A}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.5\textwidth]
+        {figures/crm_divide/timings.pdf}
+        \node at (0, 1)[anchor=north west]{\textbf{B}};
+    \end{tikzonimage}\\
+    \begin{tikzonimage}[width=\textwidth]
+        {docs/source/_static/performance/flamegraph.pdf}
+        \node at (0.02, 0.98)[anchor=north west]{\textbf{C}};
+    \end{tikzonimage}\\
+    \caption{
+        (A) Wall time of the performed simulation with increasing number of final agents inside the
+        simulation domain.
+        (B) Distribution of computation time for the division script utilized in
+        section~\ref{sec:application-1-with-division}.
+        (C) Flamegraph of a typical simulation showcasing frequency of function calls for
+        parallelized execution of simulations.
+    }
     \label{fig:timings-crm_divide}

### 2025-11-29 paper/main.tex
fix some typos; add equation for cost function

@@
 \paragraph{Motivation}
-The preceding algorithm is performant and effective but can not be applied in the case when a
+The preceding algorithm is performant and effective but cannot be applied in the case when a
 division event splits apart the bacterium since this would require comparing sets of unequal
@@
 the space that the mother had taken up.
-However, the exact time point for the fission event can not be determined accurately in practice,
+However, the exact time point for the fission event cannot be determined accurately in practice,
 thus leading to natural uncertainties with respect to the question when the lives of the daughter
@@
 
-These thoughts explain why the direct metric can not be simply reused on confined intervals which
+Consequently the direct metric cannot be simply reused on confined intervals which
 do not include division events:
@@
 bacteria where division events occur much more frequently.
-This fundamental limitation will be circumvented by the imaging-based method which we propose.
+Our proposed imaging-based method circumvents this fundamental limitation.
 
@@
 We develop an approach which can be understood as a metric in image space where we quantify
-differences in the visual footprint of two bacterial configurations.
+differences in the segmentation profile of two bacterial configurations.
 Within the realm of image segmentation many metrics such as the Jaccard Index~\cite{Murphy1996},
@@
 
-% To compare the differing cellular configurations, we require additional
-% methods which can be used in time intervals containing cellular divison events.
 We can use the previously introduced \nameref{subsection:visualization} methods to obtain cell-masks
@@
 We can now compare the two masks pixel by pixel.
-Whenever two colors are not matching, a penalty is assigned for each point in the pixel-grid.
-The total cost is the sum of all penalties.
+Whenever two colors are not matching, a penalty of $p\in[0,1]$ is assigned for each point in the
+pixel-grid.
 With this approach it is important to take into account the parental relationship between the cells,
@@
 Figure~\ref{fig:mask-difference-metric} (A,B) shows two states of a numerical simulation
-with varying amounts of cells.
+with varying numbers of cells.
 We can see that some cells have divided while others have only extended in the same time interval.
@@
 Figure~\ref{fig:mask-difference-metric} (C).
-Black indicates that the pixels are matching and no penalty is assigned while white
-indicates that the original pixels are not matching, thus adding to the overall cost.
+% Black indicates that the pixels are matching and no penalty is assigned while white
+% indicates that the original pixels are not matching, thus adding to the overall cost.
 In contrast, Figure~\ref{fig:mask-difference-metric} (D) takes the parental relationship into
 account and highlights regions where compared pixels belong to mother-daughter cells in gray.
-We assign a specific parent penalty $0\leq p\leq 1$ value for these points which should be seen as
+We assign a specific parent penalty $0\leq p_p\leq 1$ value for these points which should be seen as
 an additional hyperparameter for this optimization algorithm.
 % It is possible to assign a custom value for the parental cost value.
-Settings its value to $0$ results in a contiuous cost function in time.
-In the future, more complex choices for this parent penalty might be feasible such as
-time-dependence $p(t-t_\text{division})$.
+Setting its value to $0$ results in a continuous cost function in time.
+% In the future, more complex choices for this parent penalty might be feasible such as
+% time-dependence $p(t-t_\text{division})$.
+The total cost between two images $I^1$ and $I^2$ is the sum of all individual penalties
+\begin{align*}
+    L_I(I^1,I^2) &= \sum\limits_{i,j} p(I^1_{ij}, I^2_{ij})\\
+    p(i^1,i^2) &= \begin{cases*}
+        p_p & \text{if $i^1$ is daughter/mother of $i^2$}\\
+        1 & \text{if } $i^1\neq i^2$\\
+        0 & \text{else}
+    \end{cases*}
+\end{align*}
+where the sum is over all grid points $i,j$ of the two compared images and a penalty of $p=1$ is
+assigned for non-matching pixel values whose corresponding cells are not related by a parental
+relationship.
 
@@
 On a higher level, this means that our method is agnostic to the underlying cellular representation.
-This means that other models who even utilize very distinct cellular representations but are able to
-render segmented images can reuse our method.
+This means that other models who potentially utilize very distinct cellular representations but are
+able to render segmented images can reuse our method.
 

### 2025-11-29 paper/main.tex
rewrite test of image-space metric

@@
 
-\subsubsection{Testing the Comparison Algorithms with Division Events}
-When utilizing a computational model (see \ref{sec:computational-model}), we can generate
-data and use the generated masks to provide a benchmark of our algorithm.
-To do this, we set out to calculate the cost between two successive iteration steps
-$t_n$ and $t_{n+1}$.
-We expect to obtain an exponential growth in the cost.
-This can be explained by the fact that the growth of the total area is exponential in time.
-Furthermore, the temporal area difference can be viewed as a derivative of the total area with
-respect to time, which is again an exponential function.
-Figure~\ref{fig:mask-difference-metric} (F) shows the total cost for this approach.
-We can see that by accounting for the parental relationship, we get the overall desired effect.
-This method allows us to compare results between states that have a different number of cells.
-However, since this algorithm requires us to calculate a cell-mask for each comparison and then also
-process this mask as well as the comparator, it is much more computationally demanding.
-After all, for each comparison step, we need to run a full 3D visualization pipeline.
-In cases where the simpler approach of comparing positions of the agents is possible, this method
-should be preferred.
+\subsubsection{Benchmarking the Image-Space Metric under Division Events}
+To benchmark the stability and robustness of the developed metric under division events, we compute
+the cost between successive simulation steps $t_n$ and $t_{n+1}$ and compare the obtained values to
+expected behaviour.
+We utilize our previously defined computational model (see \nameref{sec:computational-model}), to
+generate data which includes division events.
+Since the total bacterial area $A_{t_n}$ grows exponentially in time, the differences between total
+area of successive time steps $A_{t_{n+1}}-A_{t_n}$, which can be viewed as the area's derivative in
+time, is expected to follow this trend.
+
+Figure~\ref{fig:mask-difference-metric} (F) confirms this behaviour, showcasing exponential scaling
+of the difference with larger deviations initially and better overall goodness of fit in the later
+stages of the simulation.
+In the beginning, division events introduced by individual bacteria introduce larger deviations in
+relation to the total area compared to later time points.
+Every division event introduces a spike in the cost function which is stabilized when accounting
+for parental relationships as explained earlier.
+This shows that the metric is able to interpret growth and division as a continuous biological
+process rather than an instantaneous topological break.
+
+Although this metric allows us to investigate previously unaccessible cellular dynamics, the
+algorithm is computationally expensive.
+Unlike the direct approach, it requires the rendering of segmentation masks for each time point in
+question which is currently handled by a full 3D visualization pipeline.
+Consequently, this method should be reserved for cases where the direct method is not applicable,
+e.g. when dealing with division events.
 

### 2025-11-29 paper/main.tex
rewrite figure caption

@@
     \caption{
-        The figure shows two masks (A,B) which will be compared against each other.
-        Image (B) contains more cells which are a result of cellular division events that have not
-        yet occurred in image (A).
-        (C) "Naive" comparison of cell masks, which identifies pixels of non-matching colors and
-        assigns a cost by counting these pixels.
-        This simple approach does not take into consideration the parental relationship.
-        This approach is displayed in (D) where pixels of cells that are directly related have been
-        colored in gray.
-        (E) A schematic representation of the comparison problem involving multiple cells at
-        different generations.
-        An initial configuration produces two results by utilizing distinct methods (i.e. evoling
-        our computational model in time or assigning the data for next time point).
-        (F) Comparison of cost functions, exponential regression (ER) and number of cells.
-        % Total cost between successive iteration steps $t_n,t_{n+1}$.
-        The overall shape of the cost is expected to have the slope of an exponential function.
-        We can clearly see that division events introduce jumps within this relationship.
-        In the case where we account for the parental relationship, these jumps are mitigated.
-        The cost value was set to $0$ for pixel-comparisons which belong to cells that are in a
-        parental relationship.
+        (A,B) Sequential simulatin masks before and after division event.
+        (C) Naive pixel-wise comparison, not accounting for parental relationships.
+        White regions indicate mismatches, black indicates matches.
+        (D) Lineage-aware comparison. Parental relationships have been marked in gray, resulting in
+        a modified penalty.
+        (E) Schematic of topological difference between model and data due to division.
+        (F) Evolution of cost functions applied to successive time steps $t_n$ and $t_{n+1}$,
+        exponential regression (ER) and total number of cells.
+        The naive metric (solid line) spikes at division events while the lineage-aware metric
+        (dashed line) eliminates discontinuities, following the expected exponential scaling (ER).
+        The exponential regression is fitted to the lineage-aware data.
     }

### 2025-11-29 paper/main.tex
behaviour -> behavior; amount -> number

@@
 
-The mechanical behaviour of rod-shaped bacteria is governed by a plethora of effects.
+The mechanical behavior of rod-shaped bacteria is governed by a plethora of effects.
 In this work, we aim to establish a model which is general enough such that it can be reused as the
@@
     \caption{
-        This table lists aspects of cellular behaviour which we consider in our computational model.
+        This table lists aspects of cellular behavior which we consider in our computational model.
         It is split into cellular aspects (C) which concern only an individual bacterium and
@@
 In contrast, a popular alternative is the Morse potential~\cite{Morse1929}, which is less
-customizable due to the lower amount of parameters and more restrictive mathematical formulation
+customizable due to the lower number of parameters and more restrictive mathematical formulation
 (see \nameref{sec:supplement-morse-potential}).
 Therefore, the Mie potential is a good choice for a simple, interpretable interaction potential with
-low amount of parameters that can be estimated.
+low number of parameters that can be estimated.
 
@@
 Consequently, these methods can be applied to a broader class of mechanical models that can exhibit
-varied cellular behaviours.
+varied cellular behaviors.
 
@@
 the cost between successive simulation steps $t_n$ and $t_{n+1}$ and compare the obtained values to
-expected behaviour.
+expected behavior.
 We utilize our previously defined computational model (see \nameref{sec:computational-model}), to
@@
 
-Figure~\ref{fig:mask-difference-metric} (F) confirms this behaviour, showcasing exponential scaling
+Figure~\ref{fig:mask-difference-metric} (F) confirms this behavior, showcasing exponential scaling
 of the difference with larger deviations initially and better overall goodness of fit in the later
@@
 The experiment is particularly suited as a first attempt since it excludes cell-cell interactions
-and focusses on the behaviour of a single cell.
-It therefore reduces the total amount of parameters and allows for fast iteration times when
+and focusses on the behavior of a single cell.
+It therefore reduces the total number of parameters and allows for fast iteration times when
 performing parameter estimations.
@@
 We have chosen said interval such that effects related to the lag phase can be neglected which
-allows us to sample a lower amount of paremeters thus improving runtime performance of our
+allows us to sample a lower number of paremeters thus improving runtime performance of our
 optimization scheme.
@@
         We show only one of the 6 individually-chosen growth rate as an example.
-        The remaining plots show similar behaviour although with varying values and can be found in
+        The remaining plots show similar behavior although with varying values and can be found in
         the supplementary material in section~\ref{sec:supplement-parameter-profiles}.
@@
 if and when a cell divides.
-Such behaviour can be attributed to the missing images around the division events of the cells which
+Such behavior can be attributed to the missing images around the division events of the cells which
 were not able to be segmented.
@@
 This is true for any parameter, not just the growth rates, since our approach can be generalized to
-other individual cellular behaviours.
+other individual cellular behaviors.
 
@@
 tecnhique.
-Another open question is how to efficiently scale up this operation in order to treat large amounts
+Another open question is how to efficiently scale up this operation in order to treat large numbers
 of cells such that distributions of parameters themselves can be estimated.
@@
 increased simulation time.
-To quantify this behaviour, we fit a multivariate normal distribution at each time point to the
+To quantify this behavior, we fit a multivariate normal distribution at each time point to the
 distribution of displacements.
@@
 %         This table extends table~\ref{table:simulation-aspects} with additional observations.
-%         It has to be noted that this table includes more specific behaviour that might not be
+%         It has to be noted that this table includes more specific behavior that might not be
 %         generalizable across all types of rod-shaped bacteria.

### 2025-11-29 paper/main.tex
rewrite cellular heterogeneity challenge section

@@
 \label{sec:role-of-heterogeneity}
-In order to properly estimate the parameters of our model on the single-cell level, we need to be
-able to take into account any heterogeneity that may be present.
-Therefore, the developed methods are independent of how we assign parameters within our model.
-This means we can opt to incorporate cell-cell variability by specifying parameters on a
-per-cell basis.
-Or we can employ shared parameters with identical numerical values across all agents for properties
-which are expected to be identical across the population.
-Furthermore, mixed approaches where some parameters are assigned globally and others are chosen
-individually can be realized as well.
-
-For our particular model presented in section~\ref{sec:computational-model}, a list of parameters
-is presented in table~\ref{table:list-of-parameters} but the following calculations are equally
-true for any single-cell model.
-We assume that each cell contains $N_p$ number of parameters that need to be assigned numerical
-values.
-We identify a subset $N_g\leq N_p$ of parameters which will be treated globally, thus with equal
-numerical value for every cell.
-This means that $N_p - N_g$ parameters will be assigned individually for each of number of cells.
-The total amount of parameters $N_\text{tot}$ in the system is then given by
+To capture cell-cell heterogeneity, the parameter estimation framework needs to support both global
+(population-wide) and local (cell-specific) parameters.
+In practice it is often desirable to utilize a mixed approach in which parameters corresponding to
+shared behavior are treated globally and others are specific per cell.
+
+Let $N_p$ be the total number of parameters of a single cell.
+We identify a subset $N_g\leq N_p$ of global parameters for shared behavior and $N_l=N_p-N_g$ local
+parameters assigned individually.
+The total dimension of the parameter space is given by
 \begin{equation}
-    N_\text{tot} = N_g + N_c (N_p - N_g).
+    N_\text{tot} = N_g + N_c (N_p - N_g)
 \end{equation}
-where $N_c$ is the number of unique cells which we encounter over the time interval of interest.
-This value includes cells which have divided or died during the inspected time interval.
-
-It is clear, that we should avoid, to needlessly treat parameters non-globally since this increases
-the size of the parameter space considerably.
-The volume of parameter space which needs to be sampled by optimization schemes scales exponentially
-with the dimension ie. the total number of parameters $N_\text{tot}$.
-Notice that the presented approach reflects our current work but can be adapted easily to other
-scenarios where different strategies of parameter assignment are desired.
-This can be usefull e.g. when considering two distinct species where the value of a particular
-parameter only depends on the species and is thus neither identical across all cells, nor
-individual per cell.
+where $N_c$ is the cumulative number of unique cell identifiers existing within the simulated time
+interval.
+This value includes cells which have divided or died during the inspected time interval or may only
+be present for a subset of the total time span.
+
+The volume of the search space scales exponentially with $N_\text{tot}$ and therefore also with the
+number of local parameters $N_l$ or number of cells $N_c$ respectively.
+Keeping the number of individually assigned parameters $N_l=N_p-N_g$ minimal is crucial to retain
+computational tractability.
+The presented approach reflects our current work but can be adapted easily for other approaches,
+i.e. assigning distinct parameters only depending on the species in a multi-species biofilm.
+Our developed methods are independent of these choices, allowing for flexible assignment strategies.
 

### 2025-11-30 paper/main.tex
rename section for imaging-metric
rename subsection

@@
 
-\subsubsection{Direct Skeleton-Space Metric}
-\label{subsec:skeleton-metric}
+\subsubsection{Direct Position-Space Metric}
+\label{subsec:direct-metric}
 In the scenario where no cells divide during the observed interval, we can use our extraction
@@
 
-\subsubsection{Imaging-Space Metric}
+\subsubsection{Lineage-aware Imaging-Space Metric}
 \label{subsec:image-space-metric}
@@
 Figure~\ref{fig:amir-bending-simulation} (A)) and compare the simulated positions with the extracted
-ones using the direct comparison algorithm (see Subsection~\ref{subsec:skeleton-metric}).
+ones using the direct comparison algorithm (see Subsection~\ref{subsec:direct-metric}).
 Our simulation is split into two phases; One with the external flow force that initially bends the
@@
 Therefore, the first approach will optimize the direct comparison cost function (see
-section~\ref{subsec:skeleton-metric}) while the latter approach has to account for division events
+section~\ref{subsec:direct-metric}) while the latter approach has to account for division events
 and thus requires the image-space metric method (see section~\ref{subsec:image-space-metric}).
@@
 By doing so, we can utilize the simpler and more performant direct comparison scheme (see
-\ref{subsec:skeleton-metric}) for our cost function.
+\ref{subsec:direct-metric}) for our cost function.
 We have chosen said interval such that effects related to the lag phase can be neglected which

### 2025-11-30 paper/main.tex
rewrite introduction to applications

@@
 
-In this section we apply the previously developed methods to the problem of parameter estimation
-of our model.
-We cover 3 distinct cases with increasing level of complexity.
-In the first study, we utilize results previously obtained by Amir et. al~\cite{Amir2014,Amir2014_2}
-who investigated elastic deformations of \ac{ecoli} that are partially trapped within a narrow
-tube and subject to a force resulting from a fluidic flow.
-This setup does not involve any cell-cell interactions but specifically requires a flexible rod.
-It is therefore well suited as a starting point to assess the usefulness of our methods and the
-ability of our model to describe these results.\\
-The last two cases consider time-series of microscopic images containing multiple cells (see
-subsection~\ref{section:application-2}) which will be used to perform a parameter estimation of our
-full mechanical model.
-The approach is split into a first part which considers a time interval that does not contain any
-division events and a second part which extends the initially chosen interval to include division
-events.
-In the first part, we cover a time interval which does not contain division events and thus utilize
-the direct comparison algorithm while the second case contains division events and in turn requires
-usage of the image-space metric.
+In this section we apply the previously developed methods and model to three case-studies, each
+increasing in complexity.
+We set out to estimate the parameters of every configuration.
+
+First, we fit our model to data of elastically bending bacteria confined in a microfluidic
+apparatus which was previously published by Amir et. al~\cite{Amir2014,Amir2014_2}.
+This scenario isolates the mechanics of the flexible rod and is void of any complex cell-cell
+interactions.
+It is therefore well suited as a starting point in assessing the fidelity of our approach.
+
+Subsequently, we extend our scope by incorporating multiple bacteria with cell-cell interactions
+(see subsection~\ref{section:application-2}).
+Initially, we consider a time interval that does not contain division events and are thus able to
+employ the direct metric (see \nameref{subsec:direct-metric}).
+Finally, we apply the image-space metric (see \nameref{subsec:image-space-metric}) to a time-series
+containing multiple division events, demonstrating the feasibility of our approach in such
+scenarios.
 

### 2025-11-30 paper/main.tex
change name of subsection again
rewrite introduction to applications again

@@
 
-\subsubsection{Lineage-aware Imaging-Space Metric}
+\subsubsection{Lineage-aware Image-Space Metric}
 \label{subsec:image-space-metric}
@@
 increasing in complexity.
-We set out to estimate the parameters of every configuration.
-
-First, we fit our model to data of elastically bending bacteria confined in a microfluidic
-apparatus which was previously published by Amir et. al~\cite{Amir2014,Amir2014_2}.
-This scenario isolates the mechanics of the flexible rod and is void of any complex cell-cell
+We estimate the parameters for every configuration.
+
+First, we fit our model to data of elastically bending bacteria published by Amir et.
+al~\cite{Amir2014,Amir2014_2}.
+This scenario isolates the mechanics of the flexible rod and excludes any complex cell-cell
 interactions.
@@
 Subsequently, we extend our scope by incorporating multiple bacteria with cell-cell interactions
-(see subsection~\ref{section:application-2}).
-Initially, we consider a time interval that does not contain division events and are thus able to
-employ the direct metric (see \nameref{subsec:direct-metric}).
-Finally, we apply the image-space metric (see \nameref{subsec:image-space-metric}) to a time-series
-containing multiple division events, demonstrating the feasibility of our approach in such
-scenarios.
+(see \nameref{section:application-2}).
+Initially, we consider a time interval that does not contain division events and enabling us to
+employ the direct Position-space metric (Section~\ref{subsec:direct-metric}).
+Finally, we apply the lineage-aware Image-space metric (Section~\ref{subsec:image-space-metric})
+to a time-series containing multiple division events, demonstrating the feasibility of our approach
+in such scenarios.
 

### 2025-11-30 paper/main.tex
fix acronyms
combine introduction and overview section

@@
 \newacro{ib}[IB]{Individual-based}
-\newacro{ecoli}[\textit{E. coli}]{\textit{Escherichia coli}}
-\newacro{bsubtilis}[\textit{B. subtilis}]{\textit{Bacillus subtilis}}
+\newacro{ecoli}[\textit{E.coli}]{\textit{Escherichia coli}}
+\newacro{bsubtilis}[\textit{B.subtilis}]{\textit{Bacillus subtilis}}
 \newacro{ls}[LS]{Least Squares}
@@
 
-Within our approach we develop a new computational model for flexible, elongated bacteria which
-is a mechanical setup that most \acp{abm} are not capable of describing.
+We develop a new computational model for flexible, elongated bacteria which
+is a mechanical feature that is absent in most \acp{abm}.
 So far, only a few computational studies have considered flexible rods~\cite{Duman2018,Joshi2019},
@@
 Another major problem is how to perform parameter estimations for \acp{abm}.
-Classical approaches in systems biology—such as maximum
+Classical approaches in systems biology include techniques such as maximum
 likelihood~\cite{Kreutz2013,Raue2014} or maximum a posteriori estimation, identifiability analysis
@@
 
-In this work, we address these challenges and aim to further close the previously identified gaps.
-We construct a novel mechanical model for flexible, elongated bacteria that captures growth,
+In this work, we address these challenges with the aim to further close the previously identified
+gaps.
+We provide a framework that links mechanistic modeling of elongated bacteria with quantitative
+analysis methods (Figure~\ref{fig:flowchart-project-structure}).
+First, we construct a novel mechanical model for flexible, elongated bacteria that captures growth,
 division, elastic deformations of the rod and mechanical interactions between cells.
-We develop computational methods which utilize segmented microscopic images to quantitatively
-compare numerical results with biological data on a single-cell level.
+Second, we develop computational methods which utilize segmented microscopic images to quantitatively
+compare numerical results with biological data on the single-cell level.
 With these methods we can estimate the parameters of our model in three scenarios, each probing
 different aspects of our model with increasing levels of complexity.
+
+A key feature of this framework is its modularity.
+While most subcomponents (see Figure~\ref{fig:flowchart-project-structure}) rely on the spatial
+representation of bacteria as discretized vertices, others i.e. the
+\nameref{subsec:image-space-metric} operate purely on segmentation masks, making them fully
+model-agnostic.
+This enables researchers to freely reuse our developed methods across a wide range of models for
+elongated organisms.
 
@@
 models to data which can be readily adapted to various other scenarios.
-We present a novel mechanical model for flexible elongated bacteria which seamlessly integrates into
-the \texttt{cellular\_raza} framework, enabling researchers to extend and adapt it in future
-studies.
-In total, these contributions push for a more rigorous, interpretable and biolgically grounded
+Our novel mechanical model for flexible elongated bacteria integrates seamlessly into the
+\texttt{cellular\_raza} framework, enabling researchers to extend and adapt it in future studies.
+In total, these contributions push for a more rigorous, interpretable and biologically grounded
 development process for \acp{abm} which provide deeper mechanistic insights into the nature of
@@
 %###################################################################################################
-\section{Overview}
-This work consists of many subcomponents which can be combined in order to achieve various tasks.
-Figure~\ref{fig:flowchart-project-structure} shows a high-level overview of all components.
-We assume that bacterial rods can be represented by a colleciton of vertices (see
-section~\ref{subsec:spatial-representation}).
-This assumption alone is enough to enable a cascade of other components.
-Using the spatial representation, we provide visualization methods that employ a 3D rendering
-engine and can produce cell masks or rendered images from a given configuration.
-The data extraction component determines the vertices from given segmented microscopic images.
-This allows us in particular to turn any sequence of images (movie) into a sequence of bacterial
-positions and is a key feature of the parameter estimation methods.
-The dynamics of the vertices are determined by a computational model which is specific
-to the respective use-case and should be defined by the researcher.
-It merely requires the aforementioned assumption of representing bacterial rods as a
-collection of vertices.
-This work will provide one example of how such a model can be defined.
-Notably, the visualization and data extraction components do not require any knowledge about the
-computational model, thus allowing us to reuse their functionality across differing models.
-
-By utilizing the provided data extraction methods, we can infer properties of the bacteria that aid
-us in constructing a suitable computational model.
-However, the construction of any mechanistic model also requires prior konwledge which can be used
-to inform our analysis methods.
-For example, by extracing the positions of bacteria across a time-series of microscopic images, we
-can also calculate their lengths at these time-points and thus determine their rate of growth and
-type of growth (see \nameref{paragraph:parameter-estimation-individual-treatment}), thus
-effectively guiding our modeling steps.
-As we will show later, extracted data in the form of cellular positions or segmentation masks can
-also be used for comparing computational results with biological ground truth.
-
-Visualization methods are an integral part of any modeling endevaour.
-They allow researchers to obtain feedback quickly and improve iteration times when constructing new
-models.
-Furthermore, we utilize them in our image-space metric (see
-section~\ref{subsec:image-space-metric}) where they enable us to effectively compare
-two bacterial configurations while accounting for division events.
-In the future, we are planning to extend the the rendering capabilities of our approach and
-generate realistic microscopic images and time-series.
-Together with the parameter estimation methods, they can then be used to provide large amounts of
-mechanistic data for the training of cell-segmentation and cell-tracking machine-learning
-algorithms.
+% \section{Overview}
+% This work consists of many subcomponents which can be combined in order to achieve various tasks.
+% Figure~\ref{fig:flowchart-project-structure} shows a high-level overview of all components.
+% We assume that bacterial rods can be represented by a colleciton of vertices (see
+% section~\ref{subsec:spatial-representation}).
+% This assumption alone is enough to enable a cascade of other components.
+% Using the spatial representation, we provide visualization methods that employ a 3D rendering
+% engine and can produce cell masks or rendered images from a given configuration.
+% The data extraction component determines the vertices from given segmented microscopic images.
+% This allows us in particular to turn any sequence of images (movie) into a sequence of bacterial
+% positions and is a key feature of the parameter estimation methods.
+% The dynamics of the vertices are determined by a computational model which is specific
+% to the respective use-case and should be defined by the researcher.
+% It merely requires the aforementioned assumption of representing bacterial rods as a
+% collection of vertices.
+% This work will provide one example of how such a model can be defined.
+% Notably, the visualization and data extraction components do not require any knowledge about the
+% computational model, thus allowing us to reuse their functionality across differing models.
+%
+% By utilizing the provided data extraction methods, we can infer properties of the bacteria that aid
+% us in constructing a suitable computational model.
+% However, the construction of any mechanistic model also requires prior konwledge which can be used
+% to inform our analysis methods.
+% For example, by extracing the positions of bacteria across a time-series of microscopic images, we
+% can also calculate their lengths at these time-points and thus determine their rate of growth and
+% type of growth (see \nameref{paragraph:parameter-estimation-individual-treatment}), thus
+% effectively guiding our modeling steps.
+% As we will show later, extracted data in the form of cellular positions or segmentation masks can
+% also be used for comparing computational results with biological ground truth.
+%
+% Visualization methods are an integral part of any modeling endevaour.
+% They allow researchers to obtain feedback quickly and improve iteration times when constructing new
+% models.
+% Furthermore, we utilize them in our image-space metric (see
+% section~\ref{subsec:image-space-metric}) where they enable us to effectively compare
+% two bacterial configurations while accounting for division events.
+% In the future, we are planning to extend the the rendering capabilities of our approach and
+% generate realistic microscopic images and time-series.
+% Together with the parameter estimation methods, they can then be used to provide large amounts of
+% mechanistic data for the training of cell-segmentation and cell-tracking machine-learning
+% algorithms.
 

### 2025-12-02 paper/main.tex
rework figure caption

@@
     \caption{
-        The project can be divided into subcomponents that interact with and depend on
-        each other.
-        Every component implicitly builds on the spatial representation of rod-shaped bacteria as
-        vertices (see \nameref{subsec:spatial-representation}).
-        Red boxes highlight the key scientific drivers of this study while yellow boxes indicate
-        auxiliary methods.
-        The blue box "Data Generation" displays a extension of this approach which is under active
-        development.
+        Modular framework for parameter estimation of elongated bacteria.
+        The individual components rely on a unified spatial representation (left) which discretizes
+        bacteria into vertices (see \nameref{subsec:spatial-representation}).
+        Building on top are visualization methods (top), a novel computational model for flexible
+        elongated bacteria (middle) and data extraction routines (botom).
+        Analysis methods (bottom, right) utilize extraction methods and can be informed by the model
+        or used to guide modeling choices.
+        These components can be combined to estimate the parameters of the developed computational
+        model which can then be used in generating synthetic data (top, right).
+        Colors indicate the componet type; core scientific drivers (red), auxiliary methods (yellow)
+        and extensions (blue).
     }

### 2025-12-02 paper/main.tex
rework again

--- 2025-11-30 paper/main.tex
@@
     \caption{
-        The project can be divided into subcomponents that interact with and depend on
-        each other.
-        Every component implicitly builds on the spatial representation of rod-shaped bacteria as
-        vertices (see \nameref{subsec:spatial-representation}).
-        Red boxes highlight the key scientific drivers of this study while yellow boxes indicate
-        auxiliary methods.
-        The blue box "Data Generation" displays a extension of this approach which is under active
-        development.
+        Modular framework for parameter estimation of elongated bacteria.
+        The framework relies on a unified spatial representation (left) which discretizes
+        bacteria into vertices.
+        Building on top are visualization methods (top), a novel computational model for flexible
+        elongated bacteria (middle) and data extraction routines (bottom).
+        Analysis methods (bottom right) utilize extraction methods and can be informed by the model
+        or used to guide modeling choices.
+        The parameter estimation module integrates the preceding ones and allows us to calibrate the
+        model which can then be used in generating synthetic data (top right).
+        Colors indicate the component type; core scientific drivers (red), auxiliary methods (yellow)
+        and extensions (blue).
     }

### 2025-12-02 paper/main.tex
update text with likelihood stuff

@@
 can be stored in an array $\nu_{i,j,k,l}$ of shape $(N_c,N_v,d,N_t)$.
-Comparing the results of a simulation $\nu$ with experimental values $\tilde{\nu}$ is thus only a
-matter of comparing two of these arrays via a suitable cost function~\cite{Wang2020}.
+Comparing the results of a simulation $\nu$ with parameters $\theta$ with experimental values
+$\tilde{\nu}$ is thus only a matter of comparing two of these arrays via a suitable cost
+function~\cite{Wang2020}.
 In our approach we utilize a \ac{ls} approach given by
-
 \begin{equation}
-    L_{\text{LS}}(\vec{x}, \vec{y})
-        = \sum\limits_{i,j,k,l}|\nu_{i,j,k,l} - \tilde{\nu}_{i,j,k,l}|^2
+    L_{\text{LS}}(\theta|\vec{x}, \vec{y})
+        = \sum\limits_{ijkl}|\nu_{i,j,k,l} - \tilde{\nu}_{ijkl}|^2
 \end{equation}
-
 which minimizes the distance between the corresponding vertices at each time point.
@@
 experimental values.
+
+% This metric can be used to compare the model output to datapoints obtained from microscopic images.
+A common technique to investigate identifiability questions and provide insights into confidence
+levels of the obtained results is the
+\textit{profile-likelihood}~\cite{Murphy2000,Kreutz2013,Tnsing2023}.
+It is able to provide confidence intervals which are invariant under parameter
+transformations~\cite{Meeker1995}.
+It builds on the so-called \textit{log-likelihood} which propagates uncertainties $\sigma_{ijkl}$ in
+the datapoints.
+Under the assumption of a constant, known uncertainty, i.e. $\sigma_{ijkl}=\sigma$, we can rewrite
+the log-likelihood in terms of $L_\text{LS}$ via
+\begin{align}
+    L(\theta) &= \sum\limits_{ijkl}\left(\left(\frac{\nu_{ijkl}-\tilde{\nu}_{ijkl}}{\sigma_{ijkl}}\right)^2
+    + 2 \log(\sigma_{ijkl})
+    +\text{const}\right)\\
+    &= \frac{L_\text{LS}(\theta)}{\sigma^2} + \text{const}
+\end{align}
+where we absorbed the terms $\log(\sigma)$ in the sum into the constant term.
+With this simplification, it is important to pick a value for $\sigma$ such that the real
+uncertainty is overestimated.
+Furthermore, the preceding section (\nameref{subsec:position-extraction-algorithm}) has shown that
+the error distribution of the underlying data follows a normal distribution which justifies the core
+assumption of this framework.
+We conclude that minimizing, the least squares value $L_\text{LS}$ under constant $\sigma$ is
+equivalent to minimizing the \textit{log-likelihood} $L$ which is again equivalent to optimizing the
+likelihood that the chosen parameters represent our model.
+
+The \textit{profile likelihood} $PL(\theta_i)$ scans a single parameter $\theta_i$ along its axis
+while reoptimizing the remaining ones.
+We can obtain confidence intervals via
+\begin{equation}
+    PL(\theta_i) - L(\theta) \leq \Delta_\alpha
+\end{equation}
+where the value $\Delta_\alpha$ is calculated from the $(1-\alpha)$ quantile of a $\chi^2$
+distribution with one degree of freedom~\cite{Tnsing2023}.
 

### 2025-12-03 paper/main.tex
add more details for likelihood approach

@@
 We conclude that minimizing, the least squares value $L_\text{LS}$ under constant $\sigma$ is
-equivalent to minimizing the \textit{log-likelihood} $L$ which is again equivalent to optimizing the
+equivalent to minimizing the \textit{log-likelihood} $L$ which is again equivalent to maximizing the
 likelihood that the chosen parameters represent our model.
 
-The \textit{profile likelihood} $PL(\theta_i)$ scans a single parameter $\theta_i$ along its axis
-while reoptimizing the remaining ones.
+The \textit{profile likelihood} $PL(\theta_i)$ makes use of these formalisms and scans a single
+parameter $\theta_i$ along its axis while reoptimizing the remaining ones.
 We can obtain confidence intervals via
@@
     PL(\theta_i) - L(\theta) \leq \Delta_\alpha
+    \label{eq:profile-likelihood-confidence-interval}
 \end{equation}
@@
 distribution with one degree of freedom~\cite{Tnsing2023}.
+This means that every parameter $\theta_i$ which satisfies
+relation~\eqref{eq:profile-likelihood-confidence-interval} has confidence level $(1-\alpha)$.
+In this work we will use confidence intervals $1-\alpha=68\%,90\%,95\%$.
 

### 2025-12-04 paper/main.tex
correct units

@@
         \midrule
-        Spring tension & $\gamma$ & $\unit{1\per\minute}$ & Position (vertices) & $\vec{x}_i$ & $\unit{\micro\metre}$\\
+        Spring tension & $\gamma$ & $\unit{\per\minute}$ & Position (vertices) & $\vec{x}_i$ & $\unit{\micro\metre}$\\
         Rigidity & $\eta$ & $\unit{\micro\metre\per\minute^2}$ & Velocity & $\partial_t\vec{x}_i$ & $\unit{\micro\metre\per\minute}$\\
-        Damping & $\lambda$ & $\unit{\micro\metre}$ & Spring length & $l$ & $\unit{\micro\metre}$\\
+        Damping & $\lambda$ & $\unit{\per\minute}$ & Spring length & $l$ & $\unit{\micro\metre}$\\
         \multicolumn{4}{l}{\textbf{Interaction}}\\

### 2025-12-04 paper/main.tex
rewrite introductory phrase of application 2

@@
     \caption{
-        \textbf{TODO fix caption numbers and add missing stuff}
-        (A) Microscopic image containing a single bacterium which is partially confined in a narrow
-        tube on the bottom and free at the top where a controlled flow can be exerted on them.
-        Division of the bacteria is inhibited.
-        (B) A comparison of the initial and bent state show good visual agreement of our prediction
-        with the experimental data.
-        (C-G) Likelihood profiles for the individual parameters before and after performing two
-        successive model reductions.
-        The first step removes any damping effects by settings $\lambda=0$ while the second fixes
-        the spring tension to its calculated optimum $\gamma=\gamma_\text{opt}$.
+        (A-C) Manually segmented experimental snapshots of by Amir et al.~\cite{Amir2014_2}.
+        The red line within the green mask shows the extracted position.
+        (A-C) Snapshots of bending experiment by Amir et al.~\cite{Amir2014_2} with overlayed
+        manual segmentation mask (bright green) and extracted position (red).
+        The sequence shows the rod initially (A), under flow (B) and after relaxation (C).
+        (D) Comparison of simulated and experimental rod in bent and relaxed state.
+        (D-H) Profile-likelihoods for the estimated parameters.
+        Brigth blue curves represent the full model, dark blue a reduced model without damping
+        $\lambda=0$ and orange curves when fixing the auxiliary spring tension parameter
+        $\gamma=\gamma_\text{opt}$.
+        With these reductions, every parameter is identifiable.
     }
@@
 \ac{ecoli}.
-We use a series of microscopic images and corresponding segmentation masks for the individual
-bacteria as the underlying dataset which is partially displayed in
-Figure~\ref{fig:application-2-data-overview}.
-With the previously developed extraction algorithm (see
-section~\ref{subsec:position-extraction-algorithm}) we obtain the positions of our agents at all
-time points.
-These can be used as initial values and later within our parameter estimation routine.
-We initialize our computational model (see section~\ref{sec:computational-model}) with the obtained
-positions and optimize the parameters in order to find a set which best describes the observed
-results.
-The first case considers a time interval without cell-division while the second case does contain
-division events.
-Therefore, the first approach will optimize the direct comparison cost function (see
-section~\ref{subsec:direct-metric}) while the latter approach has to account for division events
-and thus requires the image-space metric method (see section~\ref{subsec:image-space-metric}).
-These two scenarios present an opportunity to apply the previously developed methods within a
-realistic context and validate our approach.
-
-To show the usefulness of our previously developed methods
-(see section~\ref{section:parameter-estimation}), we will use them to estimate the parameters of
-individually growing \ac{ecoli} using a sequence of microscopic images.
-The data we use~\cite{Trude1982} (see section~\ref{section:data-code-availability}) has been
-automatically segmented with the omnipose~\cite{Cutler2022} algorithm to obtain the required cell
-masks.
-
-\textbf{Explain data more clearly. Refer to figure.}
+We consider two scenarios to apply the previously developed metrics within a realistic context and
+validate our approach.
+Our dataset~\cite{Trude1982} consists of of a time-series of phase-contrast images (see
+Figure~\ref{fig:application-2-data-overview} (A)).
+We segmented them using omnipose~\cite{Cutler2022} (see Figure~\ref{fig:application-2-data-overview}
+(B)) and extracted their trajectories with the previously described pipeline
+(Section~\ref{sec:supplement-extraction-algorithm}).
+The model is initialized with the extracted position of the first snapshot and parameters are
+optimized to minimize the respective metrics.
+
+For our first case-study, we select 7 individual images spanning over an interval of
+$\SI{20}{\minute}$ without any division events.
+Figure~\ref{fig:application-2-data-overview} (C) displays the initial (left) an intermediate
+(middle) and final state of this interval.
+From these images we can already see that the cellular growth is the driving factor behind the
+dynamics of the system.
+We estimate the parameters of this systems via the direct position-space metric
+(see Section~\ref{subsec:direct-metric}).
+
+In the second case, we extend the previously used interval to 13 images over $\SI{85}{\minute}$
+therefore including division events.
+This change requires us to use the lineage-aware image-space metric (see
+Section~\ref{subsec:image-space-metric}).
 

### 2025-12-04 paper/main.tex
rework figure caption

@@
     \caption{
-        \textbf{TODO also insert vertical in (B) line to indicate when direct comparison algorithm
-        starts.}
-        \textbf{Link to supplement for discussion of bar plot}\\
-        (A) Snapshot of used microscopic images.
-        (B) Segmented microscopic image.
-        (C) Selected time points of the utilized time series showcasing the spatial configurations
-        of the investigated bacteria.
+        \textbf{TODO adjust units in Plots (D-E)}
+        (A) Snapshot of initial microscopic images.
+        (B) Segmentation result of (A).
+        (C) Initial (left), intermediate (middle) and final (right) microscopic snapshots.
         A complete list of snapshots is included in the supplementary material (see
         \nameref{sec:supplement-data-details}).
-        (D) Growth curves of individual bacteria including a discrete delay before exponential
-        growth is experienced.
+        (D) Analysis of growth for the chosen time interval.
         (E) Distribution of growth parameters shows cell-cell variability of the analyzed sample.
-        The obtained results are further discussed in the supplementary material section~\ref{}.
+        The obtained results are further discussed in the supplementary material
+        section~\ref{sec:supplement-reuse-estimated-parameters}.
     }

### 2025-12-04 paper/main.tex
rewrite beginning of applications and "without division"

@@
 We have already discussed (see Section~\ref{sec:role-of-heterogeneity}) that in order to properly
-estimate parameters of our system, we need to account for natural cell-variability, thus assigning
-selected parameters on a per-cell basis.
-
-However, we will now present data that indicate that an individual treatment is in fact necessary.
-Figure~\ref{fig:application-2-data-overview} shows the rod-lengths of individual bacteria which have been
-calculated from the extracted positions.
-We fit a delayed exponential function to them which is defined by
-
-\textbf{Explain fit parameter. Explain $t_0$. Give citation for exponential growth. Also link to
-equation in cell-cycle section}
-
+estimate the parameters of our system, we need to account for natural cell-variability, thus
+assigning optimmized parameters on a per-cell basis.
+A preliminary analysis of the chosen dataset further highlights this necessity.
+Figure~\ref{fig:application-2-data-overview} (D) shows the length of the rods as calculated via our
+extraction algorithm.
+To quantify this heterogeneity, we fit the individual curves with an exponential function
 \begin{equation}
-    l(t) =
-    \left\{\begin{array}{ll}
-            l_0 \exp(\lambda (t-t_0)) & t\geq t_0\\
-            l_0 & \text{ else}
-    \end{array}\right.
+    l(t) = l_0 \exp(\lambda t).
 \end{equation}
-
-where $\lambda$ is the growth rate and $t_0$ the delay.
-Fitting this function to the individual growth curves of the bacteria and their mean growth curve,
-both results in good agreements (Figure~\ref{fig:application-2-data-overview} A,B).
-However, the latter exhibits large amounts of uncertainty within the data which is also illustrated
-in subfigure (C).
-There, we can see much more clearly the difference between considering assigning parameters per cell
-and global ones.
+where $l_0$ is the starting length of the rod and $\lambda$ is the growth rate.
+Figure~\ref{fig:application-2-data-overview} (E) shows the distribution of the obtained growth rates.
+We can see, the distinct phenotypic properties of the cells result in large deviations of the
+individual growth rates.
 These results show that in order to accurately estimate the parameters of this system, we need to
@@
 individually.
-\textbf{[Man ahnt was du meinst, doch leider ahnt man es nur. Vermeide Aussagen wie "it is clear".]}
-\textit{Jetzt klarer?}
-
-% \begin{enumerate}
-%     \item motivate individual treatment of cells
-%     \item pro: get at distributions of parameters; even more than growth rates
-%     \item pro: probably good idea to do this in this case, since more or less required (see figure)
-%     \item con: computationally more demanding
-%     \item can we improve the optimization algorithm to somehow "know" that some parameters are for
-%         different cells and thus do not "interact" as other parameters with each other?
-%         ie. they are part of the same subspace.
-% \end{enumerate}
 
@@
 
-\begin{itemize}
-    \item refer to which data?
-    \item refer to table of all parameters which are fitted
-    \item refer to equations which are fitted
-    \item "fit full mechanical model"
-    \item link to everything
-\end{itemize}
-
-In this section we will focus on a time-interval which does not contain any division events.
-By doing so, we can utilize the simpler and more performant direct comparison scheme (see
-\ref{subsec:direct-metric}) for our cost function.
-We have chosen said interval such that effects related to the lag phase can be neglected which
-allows us to sample a lower number of paremeters thus improving runtime performance of our
-optimization scheme.
-
-% For this dataset, we assumed that all parameters which specify the interaction potential are
-% identical across all agents but growth rates of the cells are individual.
-We chose to optimize for the damping of agents' movement within the medium, 3 interaction parameters
-of the Mie potential $n,m,V_0$ a single radius $r$ for all agents and growth rates which were chosen
-individually for each cell.
-This leads to a total of $5+1\times 6=11$ variable parameters.
-As highlighted before, in order to model non-flexible rods, we set the rigidity and spring tension
-of the rod to a high, fixed value.
-To estimate these parameters we used the differential evolution~\cite{Storn1997} global optimization
-algorithm which was initially seeded with a sample generated by the latin-hypercube~\cite{McKay1979}
-method.
+We first resturct our analysis to a time interval prior to the first division event which allows us
+to use the computationally more efficient direct position-space metric
+(Section~\ref{subsec:image-space-metric}).
+
+Utilizing our full mechanical model (Section~\ref{sec:computational-model}), we optimized a total
+of $11$ parameters: $5$ global parameters governing mechanics and interactions (damping
+$\lambda$, radius $R$, and parameters $n,m,V_0$ of the mie potential) together with $6$
+cell-specific growth rates $\mu_i$.
+\ac{ecoli} are short and typically straight, allowing us to fix the rigidity $\eta$ to a high value.
+Furthermore, the spring tension $\gamma$ serves as an auxiliary parameter which is also assumed
+constant.
+
+We employed a two-staged optimization strategy.
+The parameter space was first explored using the global Differential Evolution
+algorithm~\cite{Storn1997} which was initially seeded with Latin-Hypercube sampling~\cite{McKay1979}.
 Finally, we performed a local minimization using the L-BFGS-B algorithm~\cite{Liu1989}.
@@
 Figure~\ref{fig:parameter-estimates-single-step} shows profile-likelihoods for each optimized
-parameter with respect to the optimization results.
+parameter.
+\textbf{TODO adapt these last sentences to new results}
 We can see that all parameters can be estimated by our optimization approach and that no

### 2025-12-04 paper/main.tex
also remember to adjust figure for new results

@@
     \caption{
-        We calculated the likelihood profiles for every optimized parameter of our model.
-        All plots show how the value of the cost function varies when fixing a single parameter but
-        still optimizing for the remaining ones.
+        (A-H) Likelihood profiles of optimized parameters.
+        \textbf{TODO adapt this description to new results}
         Every plot indicates that a minimum was achieved, thus every parameter could be identified.
@@
         The remaining plots show similar behavior although with varying values and can be found in
-        the supplementary material in section~\ref{sec:supplement-parameter-profiles}.
+        the supplementary material (Section~\ref{sec:supplement-parameter-profiles}).
     }

### 2025-12-04 paper/main.tex
insert some of christians suggestions

@@
 
-We develop a new computational model for flexible, elongated bacteria which
+We develop a new agent-based computational model for flexible, elongated bacteria which
 is a mechanical feature that is absent in most \acp{abm}.
@@
 If \acp{abm} are not able to answer these questions, their results will significantly lack in
-explanatory power and ability to link to the specific problem at hand.
+explanatory power and their ability to link to the specific problem at hand.
 
@@
 parameters or how they are constrained.
-Furthermore, models which are not properly calibrated cannot guarantee to accurately represent the
-underlying biological reality beyond numerical coincidence.
+Furthermore, models which are not properly calibrated, i.e. by a data-based approach or via
+domain-specific knowledge, cannot guarantee to accurately represent the underlying biological reality.
 Finally, elongated bacteria are fundamental organisms which can be seen as a precursor to
@@
 models to data which can be readily adapted to various other scenarios.
-Our novel mechanical model for flexible elongated bacteria integrates seamlessly into the
-\texttt{cellular\_raza} framework, enabling researchers to extend and adapt it in future studies.
+Our novel mechanical model for flexible elongated bacteria integrates seamlessly into our previously
+developed agent-based modeling framework \texttt{cellular\_raza}~\cite{Pleyer2025}.
+It is purposefully designed to provide a flexible platform without any particular assumptions on the
+cellular representation or the nature of their processes.
+This allows researchers to freely adapt the underlying cellular model to their needs and enables
+further usage of the presented approach in future studies.
 In total, these contributions push for a more rigorous, interpretable and biologically grounded

### 2025-12-04 paper/main.tex
add some changes suggested by christian

@@
 
-We model the force exerted by the fluid flow onto the individual vertices by
+The flow profile for this setup is proportional~\cite{Wolschin2021} to $(2y_iy_\text{max}-y_i^2)$
+where $y_i$ is the vertical displacement measured from the end of the tube and $y_\text{max}$
+is the total size of the champer along the y-axis which is hidden in
+Figure~\ref{fig:amir-bending-simulation} (A).
+For every segment that connects two vertices $\vec{x}_i,\vec{x}_{i+1}$, we evaluate the flow profile
+at $y_i=0.5(\vec{x}_i+\vec{x}_{i+1})_y$.
+In particular, this enforces the correct boundary values the flow profile and the resulting force
+$\vec{F}_{i,\text{external}}|_{y=0}=0$.
+We therefore model the force exerted by the fluid flow onto the individual vertices by
 \begin{align}
@@
     \underbrace{|\vec{x}_i - \vec{x}_{i+1}|\sin(\beta_i)}_\text{proj. rod segment length}
-    \underbrace{(yR - y^2)}_\text{flow contribution}\vec{e}_x\\
+    \underbrace{(y_iy_\text{max} - y_i^2)}_\text{flow contribution}\vec{e}_x\\
     \vec{F}_{i,\text{external}} &= \frac{1}{2}(\vec{G}_{i-1}+\vec{G}_i)
@@
 Here $\beta_i$ is the angle of the rod edge between $\vec{x}_{i+1}-\vec{x}_i$ and the x-axis.
-The flow profile for this particular setup is proportional~\cite{Wolschin2021} to $(2yR-y^2)$ where
-$y=0.5(\vec{x}_i+\vec{x}_{i+1})_y-y_b$ is the vertical displacement measured from the end of the
-tube $y_b$ (see Figure~\ref{fig:amir-bending-simulation}) and $R$ is the total size of the chamber
-along the y-axis which is hidden in Figure~\ref{fig:amir-bending-simulation} (A).
-In particular, this enforces the correct boundary values the flow profile and the resulting force
-$\vec{F}_{i,\text{external}}|_{y=0}=0$.
-In addition we confine vertices who belong to segments that are partially within the tube
-$(\vec{x}_i)_y\leq y_b$ via $(\vec{x}_{i+1})_x=x_0$ and also fix the value for the first vertex
+In addition we confine vertices who belong to segments that are within the tube
+via $(\vec{x}_{i+1})_x=x_0$ and fix the value for the first vertex
 $\vec{x}_0$ such that the bacterium is attached to the bottom.

### 2025-12-04 paper/main.tex
add explanations for confidence intervals
minor style adjustments

@@
         (D-H) Profile-likelihoods for the estimated parameters.
+        Bands of blue and yellow indicate confidence intervals of $68\%,90\%$ and $95\%$.
         Brigth blue curves represent the full model, dark blue a reduced model without damping
@@
 Figure~\ref{fig:application-2-data-overview} (A)).
-We segmented them using omnipose~\cite{Cutler2022} (see Figure~\ref{fig:application-2-data-overview}
-(B)) and extracted their trajectories with the previously described pipeline
-(Section~\ref{sec:supplement-extraction-algorithm}).
+We segmented them using the omnipose toolkit~\cite{Cutler2022} (see
+Figure~\ref{fig:application-2-data-overview} (B)) and extracted their trajectories with the
+previously described pipeline (Section~\ref{sec:supplement-extraction-algorithm}).
 The model is initialized with the extracted position of the first snapshot and parameters are
@@
         The obtained results are further discussed in the supplementary material
-        section~\ref{sec:supplement-reuse-estimated-parameters}.
+        (Section~\ref{sec:supplement-reuse-estimated-parameters}).
     }
@@
         (A-H) Likelihood profiles of optimized parameters.
+        Bands of blue and yellow represent confidence intervals of $68\%,90\%$ and $95\%$.
         \textbf{TODO adapt this description to new results}

### 2025-12-05 paper/main.tex
fixup! add some changes suggested by christian
new unit for pixel
formatting

@@
 \makeatother
+
+\DeclareSIUnit\pixel{pixel}
 
@@
 between two vertices if given by
-
 \begin{equation}
@@
 \end{equation}
-
 The value of the spring constants determines the elasticity with respect to elongation of the cell.
@@
 The force acting on vertex $\vec{x}_i$ is given by
-
 \begin{equation}
@@
 \end{equation}
-
 where $\eta_i$ is the rigidity at vertex $\vec{x}_i$ (see Figure~\ref{fig:model-bacterium}).
@@
 The total force is the sum of interal and external forces.
-
 \begin{equation}
@@
 \end{equation}
-
 and are integrated via damped newtonian dynamics
-
 \begin{equation}
@@
 \end{equation}
-
 where $\lambda$ a damping constant.
@@
 Figure~\ref{fig:amir-bending-simulation} (A).
+In particular, this enforces the correct boundary values the flow profile and the resulting force
+$\vec{F}_{i,\text{external}}|_{y=0}=0$.
 For every segment that connects two vertices $\vec{x}_i,\vec{x}_{i+1}$, we evaluate the flow profile
 at $y_i=0.5(\vec{x}_i+\vec{x}_{i+1})_y$.
-In particular, this enforces the correct boundary values the flow profile and the resulting force
-$\vec{F}_{i,\text{external}}|_{y=0}=0$.
 We therefore model the force exerted by the fluid flow onto the individual vertices by
@@
 
-\begin{figure}[H]
-    \centering
-    % TODO include snapshots of utilized data
-    \caption{TODO}
-\end{figure}
+\subparagraph*{Application 1}
+
+\subparagraph*{Application 2}
+\begin{itemize}
+    \item Data from IWF Goettingen \cite{Trude1982}
+    \item $\SI{15}{\pixel\per\micro\metre}$
+    \item $8$ images are equivalent to $\SI{20}{\minute}$
+\end{itemize}
 

### 2025-12-05 paper/main.tex
rewrite amir part to match figure

@@
 and intermediate (bent, dark blue) position of the rod using the optimized parameters.
-The likelihood profiles displayed in figure~\ref{fig:amir-bending-simulation} (C-D, blue curves)
-show that each parameter is identifiable from the given data.
-However when it comes to the damping constant $\lambda$, its overall numerical value and steepness
-of the curve compared to the results for the other parameters are much lower.
+The likelihood profiles displayed in figure~\ref{fig:amir-bending-simulation} (D-H, bright blue
+curves) show that each parameter sits at minimum of the \texttt{profile-likelihood}.
+However, Figure~\ref{fig:amir-bending-simulation} (E-G) exhibit multimodal minima which can not be
+distinguished.
+Furthermore, the damping constant $\lambda$ can not be distinguished from $\lambda=0$ within the
+lowest confidence interval of $68\%$.
 We can assess if the system could be described by a reduced model by setting $\lambda=0$, thus
 effectively removing any damping.
-The curves which are obtained in this way are displayed in orange and show that the system is also
-fully identifiable in its reduced version.
+The resulting curves (dark blue) still exhibit the multimodal minima but a further reduction of
+fixing the spring tension $\gamma$ produces a (orange curves) fully identifiable system.
+As we have explained earlier (Section~\ref{sec:computational-model}), $\gamma$ can be viewed as an
+auxiliary parameter.
 
@@
 This effectively limits us to effectively 2 datapoints and is likely the key factor that leads to
-the previously described phenomenon.
-The damping force is of dynamic nature, and takes no effect in the two steady states of the rod
-being either fully bent or fully relaxed.
-Gathering additional transient information would allow us to better estimate the damping parameter.
+the non-identifiability of the damping constant $\lambda$, since this force takes no effect in the
+two steady states (fully bent or fully relaxed).
+Gathering additional transient information would allow us to better estimate this parameter.
 

### 2025-12-05 src/crm_fit/settings.rs
fix test

@@ -906,6 +906,7 @@ mod test {
                         n_vertices: 8.try_into().unwrap(),
                         n_saves: 0,
                         displacement_error: 0.5,
+                        error_cost: default_error_cost(),
                     },
                 )?,
                 parameters: Py::new(
@@ -962,6 +963,7 @@ n_voxels=[1, 1]
 rng_seed=0
 cutoff=20.0
 n_vertices=8
+displacement_error=0.5
 
 [parameters]
 radius = { min = 3.0, max=6.0, initial=4.5, individual=true }
@@ -980,8 +982,8 @@ bound = 8.0
 seed = 0
 tol = 1e-3
 
-[other]
-progressbar = false
+[others]
+# progressbar = false
 "
             .to_string();
             Ok((settings1, toml_string))

### 2025-12-05 paper/main.tex
cleanup

@@
     \caption{
-        \textbf{TODO adjust units in Plots (D-E)}
         (A) Snapshot of initial microscopic images.
@@
 
-We first resturct our analysis to a time interval prior to the first division event which allows us
+We first restrict our analysis to a time interval prior to the first division event which allows us
 to use the computationally more efficient direct position-space metric
@@
 
-Our framework allows researchers to employ well-known methods such as profile-likelihood
-\textbf{TODO CITATION} and sensitivity-analysis.
+Our framework allows researchers to employ well-known methods such as profile-likelihood and
+sensitivity-analysis.
 The code is publicly available as the python package
@@
 microscopic data with single-cell agent-based modeling techniques and offers a wide range of
-opportunities for future quantitative studies in cellular biolgoy.
+opportunities for future quantitative studies in cellular biology.
 
@@
 \paragraph*{Analysis \& Usage of heterogeneous Parameter Value (TODO)}
+\label{sec:supplement-reuse-estimated-parameters}
 In order to reuse this model, we also provide the option to assign growth rates by sampling from a
@@
 %###################################################################################################
-\paragraph*{Application Data \& Preparation (TODO)}
+\paragraph*{Application Data \& Preparation}
 \label{sec:supplement-data-details}

### 2025-12-05 paper/main.tex
rewrite section with division

@@
 
-In the next step, we extend our previously used time-series with successive images including
-cell division events which means that we now have to use the image-space metric (see
-section~\ref{section:parameter-estimation}).
-A practical problem in extending the time series is that the segmentation algorithm was not able to
-properly segment the individual bacteria around the division event which meant that we had to omit
-parts of the snapshots, leaving us with only incomplete data.
+In the next step, we extended our previously used time-series with successive images including
+cell division events which requires us to use the Lineage-Aware Image-Space Metric
+(section~\ref{section:parameter-estimation}).
+A practical challenge in extending the time series arose during data preparation: the segmentation
+algorithm was not able to properly label the individual bacteria for timepoints close to division
+events which meant that we had to omit a subset of snapshots, leaving us with incomplete temporal
+data.
+This problem is common for existing segmentation algorithms and may also be encountered in future
+studies.
+We therefore explicitly include this gap to assess the fidelity of our approach under these
+practical challenges.
 
@@
 % daughter cells will carry.
-We optimize for the same parameters as done before.
-In addition to assigning growth rates individually, we also need to consider division lengths at
+In addition to the previously used parameters, we also need to consider division lengths at
 which the bacteria split and the new individual growth rates of the daughter cells as additional
 parameters.
-Since our simulation involves the division of 4 cells into 8 new ones, in principle, this leads
-to a total of $8$ new growth rates and $4$ division lengths.
+Since our simulation involves the division of 4 cells into 8 new ones, this amounts to a total of $8$
+new growth rates and $4$ division lengths.
 % But to decrease the size of the parameter space, we assume that the growth rates of the daughter
@@
 To optimize this system, we used a similar optimization approach by again seeding with samples
-generated via latin-hypercube~\cite{McKay1979}, then applying the differential evolution
+generated via Latin-Hypercube~\cite{McKay1979}, then applying the differential evolution
 algorithm~\cite{Storn1997} and finally a local minimization with Nelder-Mead~\cite{Gao2010}.
-We used the value $0.5$ for the parent penalty, which aims to provide a smooth transition between
+We used the value $p=0.5$ for the parent penalty, which aims to provide a smooth transition between
 not accounting for them and assigning full weighting to this difference.
@@
 
-We can again calculate likelihood profiles for every optimized parameter.
-The results displayed in Figure~\ref{fig:likelihood-profiles-comparison-with-division} show that
-some curves exhibit a plateau close to the calculated optimum of the estimated parameter.
+% We can again calculate likelihood profiles for every optimized parameter.
+Since further mathematical analysis is required to assess how to utilize the Lineage-Aware
+Image-Space metric for maximum-likelihood approaches, we simply display the change of the cost
+function when fixing a single parameter.
+While only confidence bands for likelihood profiles as presented in the previous sections could
+paint a clearer picture with regards to the identifiability of the parameters, we can still
+identify some trends.
+Figure~\ref{fig:likelihood-profiles-comparison-with-division} shows that some curves (F,G,I,L)
+exhibit a plateau close to the calculated optimum of the estimated parameter.
 This concerns the division lengths and some growth rates of the mother bacteria which both determine
@@
 were not able to be segmented.
+This presents a restriction on the lower and upper bounds of the parameter value.
+However for some parameters (H,J,K) the lower region of the plots contains a slight gradient,
+indicating that our mechanistic model is able to provide some level of information about the missing
+datapoints.
+It therefore further underlines the accuracy and predictive power of our methods.
 \textbf{TODO check this with results}
-It can be argued that this leads to a non-identifiability for the first division length while the
-remaining parameters are still identifiable although the effect is also notable.
-These results highlight a reassuring finding, namely that the method of profile-likelihood is able
-to properly display this lack of data.
-It therefore further underlines the accuracy and predictive power of our methods.
 
@@
 to the total wall time.
-These steps are prediction via the computational model ($15.63(21)\%$), mask adjustment of the data
-masks ($14.49(19)\%$), mask rendering for the numerical results ($69.2(1.4)\%$) and finally
-comparison of the masks ($0.68(5)\%$) (see section \nameref{section:supplement-performance}).
-The second step alters the masks provided by the data for every individual numerical simulation.
-This is required since we need to ensure that the colors of the data-masks match the colors of the
-masks produced by our computational model.
-To this end, we need to construct a family tree of the data-bacteria and compare with the
-\textit{in-silico} family tree.
+Rendering of the masks dominates the workload ($69.2(1.4)\%$), followed by the model prediction
+($15.63(21)\%$) and mask adjustment ($14.49(19)\%$) while the actual pixel-wise compariosn is
+negligible ($0.68(5)\%$) (see also Section \ref{section:supplement-performance}).
+The mask-adjustment step is necessary for every individual numerical simulation to ensure that masks
+obtained from the data are color-comparable to the ones produced by our computational model.
+To this end, we need to construct a set of family trees (forest) of the data-bacteria and compare
+with the \textit{in-silico} family trees.
 We can see that the generation of masks from the numerical results takes up the majority of the work.

### 2025-12-05 paper/main.tex
move table to front and discuss it

@@
 
+\textbf{TODO write text to compare the numerical values. This needs updated input from results.}
+Table~\ref{tab:comparison-growth-rates} shows the obtained values for the growth rates via the
+different techniques.
+Although numerical values differ between the individual settings, each estimation technique was able
+to capture the same trends.
+Cell $0,1,3$ tend to be the fastest growing cells while $2,4,5$ are slower.
+Furthermore, we can see that daughter cells tend to grow quicker compared to mother cells,
+indicating the their parents were still adapting to the new environment.
+
+\begin{table}[H]
+    \centering
+    \begin{tabular}{l c c c c c c}
+        \multicolumn{4}{l}{\textbf{Growth Parameters for different Methods}}\\
+        \toprule
+        % \multicolumn{6}{Growth Rates}\\
+        Method & Cell $0$ & Cell $1$ & Cell $2$ & Cell $3$ & Cell $4$ & Cell $5$\\
+        \midrule
+        Fit             & 0.00703 & 0.00540 & 0.01159 & 0.00756 & 0.01024 & 0.00836\\
+        \ac{abm} Direct & 0.00671 & 0.00746 & 0.01138 & 0.00848 & 0.02005 & 0.00952\\
+        \ac{abm} Imaging& 0.00104 & 0.00146 & 0.00206 & 0.00286 & 0.00190 & 0.00171\\
+        & \textbf{Daughters}\\
+        \midrule
+        & Cell $0$ & Cell $1$ & Cell $2$ & Cell $3$ & Cell $4$ & Cell $5$\\
+        & 0.00240 & 0.00494 & 0.00096 & 0.00363\\
+        & 0.00490 & 0.00226 & 0.00448 & 0.00048\\
+        % 0.0014057073 & 0.0010807540 & 0.0023177727 & 0.0015120675 & 0.0020488179 & 0.0016726263\\
+        % direct & 0.0009785162 & 0.0012641432 & 0.0021701728 & 0.0015695889 & 0.0036255661 & 0.0018419186\\
+        \bottomrule
+    \end{tabular}
+    \caption{
+        \textbf{TODO maybe report uncertainties?}
+        \textbf{TODO Update and explain these values. And also add units!}\\
+        \textbf{TODO FIX NUMERICAL VALUES}\\
+        \textbf{TODO check ordering}
+    }
+    \label{tab:comparison-growth-rates}
+\end{table}
+
 %###################################################################################################
@@
 \textbf{TODO CITATION}.
-
-\begin{table}[H]
-    \centering
-    \begin{tabular}{l c c c c c c}
-        \toprule
-        % \multicolumn{6}{Growth Rates}\\
-        Method & Cell $0$ & Cell $1$ & Cell $2$ & Cell $3$ & Cell $4$ & Cell $5$\\
-        \midrule
-        Fit   & 0.001405 & 0.001080 & 0.002317 & 0.001512 & 0.002048 & 0.001672\\
-        \ac{abm} Direct & 0.001152 & 0.001410 & 0.001876 & 0.001683 & 0.003610 & 0.001520\\
-        \ac{abm} Imaging\\
-        % 0.0014057073 & 0.0010807540 & 0.0023177727 & 0.0015120675 & 0.0020488179 & 0.0016726263\\
-        % direct & 0.0009785162 & 0.0012641432 & 0.0021701728 & 0.0015695889 & 0.0036255661 & 0.0018419186\\
-        \bottomrule
-    \end{tabular}
-    \caption{TODO Update and explain these values. And also add units!}
-\end{table}
 

### 2025-12-05 paper/main.tex
fixup! move table to front and discuss it
add one more idea

@@
     \caption{
-        \textbf{TODO maybe report uncertainties?}
-        \textbf{TODO Update and explain these values. And also add units!}\\
-        \textbf{TODO FIX NUMERICAL VALUES}\\
-        \textbf{TODO check ordering}
+        \textbf{TODO maybe report uncertainties?;
+        Update and explain these values. And also add units!;
+        FIX NUMERICAL VALUES;
+        check ordering of values}
     }
@@
     \item other dynamics
-    \item advantage of imaging-based comparison algo: no need to specify number of vertices; more
-        naturally aligned with biolgical reality (no vertices)
+    \item advantage of image-space metric: no need to specify number of vertices; more
+        naturally aligned with biological reality (no vertices); keywords: Topology-Restricted vs.
+        Topology-Agnostic
 \end{itemize}

### 2025-12-05 src/crm_fit/optimize.rs
add option to specify mutation parameter

@@ -232,6 +232,7 @@ res = sp.optimize.differential_evolution(
     polish=optimization.polish,
     rng=optimization.seed,
     callback=callback,
+    mutation=optimization.mutation,
 )
 "#
                 ),

### 2025-12-05 src/crm_fit/settings.rs

@@ -317,6 +317,10 @@ pub struct DifferentialEvolution {
     /// Recombination value of the differential evolution algorithm
     #[serde(default = "default_recombination")]
     pub recombination: f32,
+    /// Mutation variable of the differential evolution algorithm
+    #[approx(epsilon_map = |x| (x, x))]
+    #[serde(default = "default_mutation")]
+    pub mutation: (f32, f32),
     /// Determines if the final result should be polished
     #[approx(equal)]
     #[serde(default = "default_polish")]
@@ -383,6 +387,10 @@ pub(crate) const fn default_pop_size() -> usize {
 
 pub(crate) const fn default_recombination() -> f32 {
     0.3
+}
+
+pub(crate) const fn default_mutation() -> (f32, f32) {
+    (0.5, 1.5)
 }
 
 pub(crate) const fn default_polish() -> bool {
@@ -949,6 +957,7 @@ mod test {
                         max_iter: default_max_iter(),
                         pop_size: default_pop_size(),
                         recombination: default_recombination(),
+                        mutation: default_mutation(),
                         polish: default_polish(),
                     }),
                 )?,

### 2025-12-05 cr_mech_coli/imaging.py
allow pixel_per_micron to have two components for image

@@ -55,9 +55,9 @@ class RenderSettings:
     Others are used by :mod:`cv2`.
     """
 
-    pixel_per_micron: float | np.float32 = (
-        12.8  #: Determines Resolution of the generated image
-    )
+    pixel_per_micron: (
+        float | np.float32 | tuple[float, float] | tuple[np.float32, np.float32]
+    ) = 12.8  #: Determines Resolution of the generated image
     diffuse: float | np.float32 = 0.5  #: Value Between 0 and 1.
     ambient: float | np.float32 = 0.5  #: Value between 0 and 1.
     specular: float | np.float32 = 0.5  #: Value between 0 and 1.
@@ -152,11 +152,16 @@ def render_pv_image(
     if render_distance is None:
         render_distance = max(domain_size)
 
+    try:
+        ppm = (render_settings.pixel_per_micron[0], render_settings.pixel_per_micron[1])
+    except:
+        ppm = (render_settings.pixel_per_micron, render_settings.pixel_per_micron)
+
     plotter = pv.Plotter(
         off_screen=True,
         window_size=[
-            int(np.round(domain_size[0] * render_settings.pixel_per_micron)),
-            int(np.round(domain_size[1] * render_settings.pixel_per_micron)),
+            int(np.round(domain_size[0] * ppm[0])),
+            int(np.round(domain_size[1] * ppm[1])),
         ],
     )
 

### 2025-12-05 cr_mech_coli/crm_estimate_params/main.py
store obtained growth parameters

@@ -283,6 +283,8 @@ def estimate_growth_curves_individual(
     ax.set_xlabel("Growth Rate [1/min]")
     ax.set_ylabel("Count")
 
+    np.savetxt(out_path / "growth_rates.csv", growth_rates, delimiter=",")
+
     yticks = ax.get_yticks()
     yticks = list(filter(lambda x: int(x) == x, yticks))
     ax.set_yticks(yticks, minor=False)

### 2025-12-05 cr_mech_coli/crm_fit/main.py
add option to use global pre-optimization; add arguments for that

@@ -200,6 +200,18 @@ def crm_fit_main():
         default="Nelder-Mead",
         help="Method to be used to minimize cost function in profiles",
     )
+    parser.add_argument(
+        "--profiles-pre-global",
+        default=False,
+        action="store_true",
+        help="Perform a pre-optimization with the differential_evolution algorithm before locally minimizing.",
+    )
+    parser.add_argument(
+        "--profiles-pre-maxiter",
+        type=int,
+        default=20,
+        help="Number of iterations for global pre-optimization",
+    )
     pyargs = parser.parse_args()
     if pyargs.workers == -1:
         pyargs.workers = mp.cpu_count()

### 2025-12-05 cr_mech_coli/crm_fit/plotting.py

@@ -41,16 +41,31 @@ def optimize_around_single_param(opt_args):
 
     bounds = [(b_low[i], b_upp[i]) for i in range(len(b_low))]
 
+    x0 = params_opt
+    if pyargs.profiles_pre_global:
+        res = sp.optimize.differential_evolution(
+            prediction_optimize_helper,
+            x0=params_opt,
+            args=(param_single, n, *args),
+            bounds=bounds,
+            disp=True,
+            maxiter=pyargs.profiles_pre_maxiter,
+            polish=False,
+            workers=1,
+            popsize=5,
+            mutation=(0.5, 1.5),
+        )
+        x0 = res.x
+
     res = sp.optimize.minimize(
         prediction_optimize_helper,
-        x0=params_opt,
+        x0=x0,
         args=(param_single, n, *args),
         bounds=bounds,
         method=pyargs.profiles_method,
         options={
             "disp": True,
             "maxiter": pyargs.profiles_maxiter,
-            "maxfev": pyargs.profiles_maxiter,
         },
     )
     return res.fun

### 2025-12-05 cr_mech_coli/crm_fit/plotting.py
make limits smaller

@@ -191,7 +191,7 @@ def plot_profile(
     xmax = x[nxmax]
     dx = xmax - xmin
 
-    ax.set_xlim(xmin - 0.2 * dx, xmax + 0.2 * dx)
+    ax.set_xlim(xmin - 0.05 * dx, xmax + 0.05 * dx)
 
     plt.savefig(f"{odir / name.lower().replace(' ', '-')}.png")
     plt.savefig(f"{odir / name.lower().replace(' ', '-')}.pdf")

### 2025-12-11 paper/main.tex
change order of authors
add reminders
fixup! rename folder

@@
     \begin{tikzonimage}[width=0.49\textwidth]
-        {figures/crm_fit/snapshots/predicted-003889.png}%
+        {figures/crm_fit/mie/snapshots/predicted-003889.png}%
         \node at (0.025, 0.975)[anchor=north west, white]{\textbf{A}};
@@
     \begin{tikzonimage}[width=0.49\textwidth]
-        {figures/crm_fit/celldiffs/diff-003889.png}%
+        {figures/crm_fit/mie/celldiffs/diff-003889.png}%
         \node at (0.025, 0.975)[anchor=north west, white]{\textbf{B}};
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_fit/profiles/damping.pdf}%
+        {figures/crm_fit/mie/profiles/damping.pdf}%
         \node at (0, 1.1)[anchor=north west, black]{\textbf{A}};
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_fit/profiles/exponent-m.pdf}%
+        {figures/crm_fit/mie/profiles/exponent-m.pdf}%
         \node at (0, 1.1)[anchor=north west, black]{\textbf{B}};
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_fit/profiles/exponent-n.pdf}%
+        {figures/crm_fit/mie/profiles/exponent-n.pdf}%
         \node at (0, 1.1)[anchor=north west, black]{\textbf{C}};
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_fit/profiles/strength.pdf}%
+        {figures/crm_fit/mie/profiles/strength.pdf}%
         \node at (0, 1.1)[anchor=north west, black]{\textbf{D}};
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_fit/profiles/radius.pdf}%
+        {figures/crm_fit/mie/profiles/radius.pdf}%
         \node at (0, 1.1)[anchor=north west, black]{\textbf{E}};
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_fit/profiles/growth-rate-0.pdf}%
+        {figures/crm_fit/mie/profiles/growth-rate-0.pdf}%
         \node at (0, 1.1)[anchor=north west, black]{\textbf{F}};
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_fit/profiles/growth-rate-1.pdf}%
+        {figures/crm_fit/mie/profiles/growth-rate-1.pdf}%
         \node at (0, 1.1)[anchor=north west, black]{\textbf{G}};
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_fit/profiles/growth-rate-2.pdf}%
+        {figures/crm_fit/mie/profiles/growth-rate-2.pdf}%
         \node at (0, 1.1)[anchor=north west, black]{\textbf{H}};
@@
         division lengths.
+        \textbf{TODO add profiles for daughter cells}
         \textbf{TODO possibly extend this description?}
@@
         \ac{abm} Direct & 0.00671 & 0.00746 & 0.01138 & 0.00848 & 0.02005 & 0.00952\\
+        \textbf{Imaging values will be adjusted!}\\
         \ac{abm} Imaging& 0.00104 & 0.00146 & 0.00206 & 0.00286 & 0.00190 & 0.00171\\
@@
     \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/growth-rate-0.pdf}%
+        {figures/crm_fit/mie/profiles/growth-rate-0.pdf}%
     \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/growth-rate-1.pdf}%
+        {figures/crm_fit/mie/profiles/growth-rate-1.pdf}%
     \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/growth-rate-2.pdf}\\%
+        {figures/crm_fit/mie/profiles/growth-rate-2.pdf}\\%
     \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/growth-rate-3.pdf}%
+        {figures/crm_fit/mie/profiles/growth-rate-3.pdf}%
     \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/growth-rate-4.pdf}%
+        {figures/crm_fit/mie/profiles/growth-rate-4.pdf}%
     \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/profiles/growth-rate-5.pdf}
+        {figures/crm_fit/mie/profiles/growth-rate-5.pdf}
     \caption{Parameter estimation of growth rates for individual cells}%TODO extend caption}
@@
 \textbf{Writing – original draft:} Jonas Pleyer\\
-\textbf{Writing – review \& editing:} Christian Fleck \& Jonas Pleyer
+\textbf{Writing – review \& editing:} Jonas Pleyer \& Christian Fleck
 

### 2025-12-11 cr_mech_coli/crm_fit/__init__.py
use separate function to plot optimization-progression

@@ -42,5 +42,5 @@
 """
 
 from .plotting import *
-from .main import crm_fit_main
+from .main import crm_fit_main, plot_optimization_progression
 from cr_mech_coli.crm_fit.crm_fit_rs import *

### 2025-12-11 cr_mech_coli/crm_fit/main.py

@@ -119,6 +119,28 @@ def transform_input_mask(
         return res
 
 
+def plot_optimization_progression(evals, out):
+    fig, ax = plt.subplots(figsize=(8, 8))
+    crm.plotting.configure_ax(ax)
+    evals = np.sort(evals)[::-1]
+    ax.plot(
+        np.arange(len(evals)),
+        evals,
+        label="Cost Function",
+        color=crm.plotting.COLOR3,
+    )
+    ax.legend(
+        loc="upper center",
+        bbox_to_anchor=(0.5, 1.1),
+        ncol=1,
+        frameon=False,
+    )
+    ax.set_xlabel("Iterations")
+    fig.savefig(out / "optimization-progression.png")
+    fig.savefig(out / "optimization-progression.pdf")
+    plt.close(fig)
+
+
 def crm_fit_main():
     parser = argparse.ArgumentParser(
         description="Fits the Bacterial Rods model to a system of cells."
@@ -312,28 +334,7 @@ def crm_fit_main():
 
     interval = time.time()
 
-    def plot_optimization_progression(evals):
-        fig, ax = plt.subplots(figsize=(8, 8))
-        crm.plotting.configure_ax(ax)
-        evals = np.sort(evals)[::-1]
-        ax.plot(
-            np.arange(len(evals)),
-            evals,
-            label="Cost Function",
-            color=crm.plotting.COLOR3,
-        )
-        ax.legend(
-            loc="upper center",
-            bbox_to_anchor=(0.5, 1.1),
-            ncol=1,
-            frameon=False,
-        )
-        ax.set_xlabel("Iterations")
-        fig.savefig(out / "optimization-progression.png")
-        fig.savefig(out / "optimization-progression.pdf")
-        plt.close(fig)
-
-    plot_optimization_progression(optimization_result.evals)
+    plot_optimization_progression(optimization_result.evals, out)
 
     displacement_error = settings.constants.displacement_error
 

### 2025-12-11 cr_mech_coli/crm_divide.py
plot optimization-progression for crm_divide

@@ -825,6 +825,13 @@ def run_optimizer(
     args,
     pyargs,
 ):
+    evals = []
+
+    def callback(intermediate_result):
+        fun = intermediate_result.fun
+        global evals
+        evals.append(float(fun))
+
     # Try loading data
     if iteration is not None:
         result = np.loadtxt(output_dir / "optimize_result.csv")
@@ -847,6 +854,7 @@ def run_optimizer(
             polish=False,
             init="latinhypercube",
             strategy="best1bin",
+            callback=callback,
         )
         if not pyargs.skip_polish:
             res = sp.optimize.minimize(
@@ -865,7 +873,7 @@ def run_optimizer(
         final_cost = res.fun
         np.savetxt(output_dir / "optimize_result.csv", [*final_parameters, final_cost])
 
-    return final_parameters, final_cost
+    return final_parameters, final_cost, evals
 
 
 def plot_snapshots(
@@ -1080,7 +1088,7 @@ def crm_divide_main():
         parent_penalty,
     )
 
-    final_parameters, final_cost = run_optimizer(
+    final_parameters, final_cost, evals = run_optimizer(
         x0,
         bounds,
         output_dir,
@@ -1088,6 +1096,8 @@ def crm_divide_main():
         args,
         pyargs,
     )
+
+    crm_fit.plot_optimization_progression(evals, output_dir)
 
     if (
         not pyargs.skip_snapshots

### 2025-12-11 cr_mech_coli/crm_fit/main.py
plot distribution of growth rates

@@ -86,6 +86,13 @@ def estimate_growth_rates(iterations, lengths, settings, out_path):
 
     fig.savefig(out_path / "estimated-growth-rates.png")
     fig.savefig(out_path / "estimated-growth-rates.pdf")
+
+    ax.cla()
+    crm.plotting.configure_ax(ax)
+    ax.hist(growth_rates)
+    fig.savefig(out_path / "estimated-growth-rates-distribution.png")
+    fig.savefig(out_path / "estimated-growth-rates-distribution.pdf")
+
     return growth_rates, growth_rates_err
 
 

### 2025-12-11 cr_mech_coli/crm_fit/plotting.py
modular plotting of morse potential

@@ -288,6 +288,57 @@ def __plot_mie_potential(
     plt.close(fig)
 
 
+def plot_morse_potential(
+    x,
+    radius,
+    potential_stiffness,
+    strength,
+    cutoff,
+    fig_ax,
+    ls,
+    label=None,
+):
+    def morse_potential(x, r, potential_stiffness, cutoff):
+        t = 1 - np.exp(-potential_stiffness * (x - r))
+        c = x <= cutoff
+        n_last = np.argmax(1 - c)
+        y = strength * t**2
+        return y * c + (1 - c) * y[n_last], n_last
+
+    if fig_ax is None:
+        crm.plotting.set_mpl_rc_params()
+        fig, ax = plt.subplots(figsize=(8, 8))
+        crm.plotting.configure_ax(ax)
+    else:
+        fig, ax = fig_ax
+
+    y, n_y_bound = morse_potential(x, radius, potential_stiffness, cutoff)
+    ax.set_xlabel("Distance [µm]")
+    ax.set_ylabel("Interaction Strength [µm^2/min^2]")
+
+    # ax.plot(x, y, linestyle=ls, color=crm.plotting.COLOR2)
+    if label is None:
+        label = f"ω={potential_stiffness:4.2f}"
+    ax.plot(
+        x,
+        y,
+        label=label,
+        linestyle=ls,
+        color=crm.plotting.COLOR3,
+    )
+
+    ylower = np.min(y)
+    yupper = np.max(y)
+    dy = yupper - ylower
+    ax.set_ylim(ylower - 0.05 * dy, yupper + 0.05 * dy)
+
+    # n_y_bound = len(y) - np.argmax(y[::-1] > 0)
+    yfinmax = y[n_y_bound]
+    ax.vlines(cutoff, ylower - dy, yfinmax, color=crm.plotting.COLOR5)
+
+    return fig, ax, y
+
+
 def __plot_morse_potential(
     settings: Settings,
     optimization_result: OptimizationResult,
@@ -310,40 +361,15 @@ def __plot_morse_potential(
     )
     cutoff = settings.constants.cutoff
 
-    def morse_potential(x, r, potential_stiffness, cutoff):
-        t = 1 - np.exp(-potential_stiffness * (x - r))
-        c = x <= cutoff
-        n_last = np.argmax(1 - c)
-        y = strength * t**2
[…]

### 2025-12-11 plots/mie-potential.py
add plotting of morse potential
- slightly adjust variables

@@ -1,12 +1,13 @@
 import numpy as np
+import matplotlib.pyplot as plt
 
 from cr_mech_coli import crm_fit
 
 if __name__ == "__main__":
     bound = 6.0
-    cutoff = 3.0
+    cutoff = 3.5
 
-    x = np.linspace(0.01, 3.3, 100)
+    x = np.linspace(0.01, 3.9, 200)
 
     data = [
         (2, 1, ":"),
@@ -21,13 +22,37 @@ if __name__ == "__main__":
         fig_ax = (f[0], f[1])
     fig, ax = fig_ax
 
-    ax.set_xlim(0, 3.3)
+    ax.set_xlim(0, np.max(x))
     ax.set_ylim(-2.5, 3)
     ax.legend()
     ax.set_title("Mie Potential")
 
     ax.set_xlabel("Distance [R$_1$+R$_2$]")
-    ax.set_ylabel("Interaction Strength [U$_0$]")
+    ax.set_ylabel("Interaction Strength [V$_0$]")
 
     fig.savefig("docs/source/_static/mie-potential-shapes.png")
     fig.savefig("docs/source/_static/mie-potential-shapes.pdf")
+
+    plt.close(fig)
+
+    data2 = [(1.0, ":"), (1.5, "-."), (3.0, "--")]
+    fig_ax = None
+    for sti, ls in data2:
+        label = f"ω={sti:3.1f}/(R$_1$+R$_2$)"
+        f = crm_fit.plot_morse_potential(x, 1, sti, 1, cutoff, fig_ax, ls, label=label)
+        fig_ax = (f[0], f[1])
+    fig, ax = fig_ax
+
+    ax.set_xlim(0, np.max(x))
+    xmin = 0
+    xmax = 5
+    dx = 0.05 * (xmax - xmin)
+    ax.set_ylim(xmin - dx, xmax + dx)
+    ax.legend()
+    ax.set_title("Morse Potential")
+
+    ax.set_xlabel("Distance [R$_1$+R$_2$]")
+    ax.set_ylabel("Interaction Strength [V$_0$]")
+
+    fig.savefig("docs/source/_static/morse-potential-shapes.png")
+    fig.savefig("docs/source/_static/morse-potential-shapes.pdf")

### 2025-12-11 paper/main.tex
also plot morse potential and adjust figure caption

@@
     \centering
-    \begin{tikzonimage}[width=0.3\textwidth]
+    \begin{tikzonimage}[width=0.5\textwidth]
         {docs/source/_static/mechanics.pdf}
@@
     \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.3\textwidth]
+    \begin{tikzonimage}[width=0.5\textwidth]
         {docs/source/_static/interaction.pdf}
         \node at (0, 1)[anchor=north west]{\textbf{B}};
-    \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.3\textwidth]
+    \end{tikzonimage}\\
+    \begin{tikzonimage}[width=0.5\textwidth]
         {docs/source/_static/mie-potential-shapes.pdf}
         \node at (0, 1)[anchor=north west]{\textbf{C}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.5\textwidth]
+        {docs/source/_static/morse-potential-shapes.pdf}
+        \node at (0, 1)[anchor=north west]{\textbf{D}};
     \end{tikzonimage}%
@@
         equilibrium state of a straight rod with no curvature.
-        (B) Potential curves of the Mie potential for various combinations of exponents $n,m$ with
-        upper bound $\beta=2.5U_0$ and cutoff $\zeta=3(R_1+R_2)$.
+        (B) Schematic of calculating the physical interaction between two bacteria.
+        Vertices on the upper agent are interacting with nearest points on the polygon by the lower
+        agent.
+        (C) Potential curves of the Mie potential for various combinations of exponents $n,m$ with
+        upper bound $\beta=2.5U_0$ and cutoff $\zeta=3.5(R_1+R_2)$.
+        (D) Potential curves of the Morse potential with various stiffnesses $\omega$ and cutoff
+        $\zeta=3.5(R_1+R_2)$.
     }

### 2025-12-11 paper/main.tex
adjust interaction section; add morse potential; restructure slightly

@@
 % --------------------------------------------------------------------------------------------------
-\subsection{Physical Interactions - Mie Potential}
+\subsection{Physical Interactions}
+
+The physical interaction between the bacteria plays a crucial role for the overall dynamics of the
+system.
+The Mie potential~\cite{Mie1903} is capable of describing a wide range of interaction
+types from softer ones that permit some level of overlap to rigid ones with steep ascends for
+distance valuels corresponding to overlaps.
+In contrast, a popular alternative is the Morse potential~\cite{Morse1929}, which is less
+customizable due to the lower number of parameters and more restrictive mathematical formulation
+but may therefore be advantagous in situations of sparse data.
+These potentials are a good choice for a simple, interpretable interaction potential with
+low number of parameters that can be estimated.
+
+\subsubsection{Mie Potential}
 
@@
 
-The physical interaction between the bacteria plays a crucial role for the overall dynamics of the
-system.
-By varying its parameters, the Mie potential is capable of describing a wide range of interaction
-types from softer ones that permit some level of overlap to rigid ones with steep ascends for
-distance valuels that correspond to overlaps.
-In contrast, a popular alternative is the Morse potential~\cite{Morse1929}, which is less
-customizable due to the lower number of parameters and more restrictive mathematical formulation
-(see \nameref{sec:supplement-morse-potential}).
-Therefore, the Mie potential is a good choice for a simple, interpretable interaction potential with
-low number of parameters that can be estimated.
+%###################################################################################################
+\subsubsection{Morse Potential}
+\label{subsec:morse-potential}
+The morse potential~\cite{Morse1929} is a common example of numerically inexpensive interaction
+potential.
+It was originally brought up to describe the interaction of diatomic molecules but has since found
+use in other areas~\cite{Breitwieser2021}.
+It contains a repulsive and attractive part and is bound from above which makes it particularly
+suitable for numerical simulations since unbound potentials can lead to large numerical deviations
+thus compromising the prediction of the model.
+The potential is of the form
+\begin{equation}
+    V(r) = V_0\left(1-e^{-\omega(r-(R_1+R_2))}\right)^2
+\end{equation}
+and has 3 parameters.
+The overall potential strength $V_0$, the radii of the particles $R_1,R_2$ and the potential
+stiffness $\lambda$ which controls the rate of change as one travels along the radial axis.
+We also assign a spatial cutoff $\xi$ to the range of the interaction potential.
+Figure~\ref{fig:model-bacterium} (D) shows various configurations of a morse potential.
 
@@
 \section*{Supporting Information}
-
-%###################################################################################################
-\paragraph*{Alternative Interaction Potential - Morse Potential}
-\label{sec:supplement-morse-potential}
-The morse potential~\cite{Morse1929} is a common example of numerically inexpensive interaction
-potential.
-It was originally designed to describe the interaction of diatomic molecules but has since found use
-in other areas~\cite{Breitwieser2021}.
-It contains a repulsive and attractive part and is bound from above which makes it particularly
-suitable for numerical simulations since unbound potentials can lead to large numerical deviations
-thus compromising the prediction of the model.
-The potential is of the form
-\begin{equation}
-    V(r) = V_0\left(1-e^{-\lambda(r-(R_1+R_2))}\right)^2
-\end{equation}
-and has 3 parameters.
-The overall potential strength $V_0$, the radii of the particles $R_1,R_2$ and the potential
-stiffness $\lambda$ which controls the rate of change as one travels along the radial axis.
-As done with our approach towards the Mie potential, we are free to assign a spatial cutoff $\xi$
-to the range of the interaction potential.
 

### 2025-12-11 paper/main.tex
adjust table for coming additional parameter values

@@
     \centering
+    \textbf{Growth Parameters obtained by various Methods}
     \begin{tabular}{l c c c c c c}
-        \multicolumn{4}{l}{\textbf{Growth Parameters for different Methods}}\\
+        % \multicolumn{7}{l}{}\\
+        % \multicolumn{6}{Growth Rates}\\
+        & Cell $0$ & Cell $1$ & Cell $2$ & Cell $3$ & Cell $4$ & Cell $5$\\
         \toprule
-        % \multicolumn{6}{Growth Rates}\\
-        Method & Cell $0$ & Cell $1$ & Cell $2$ & Cell $3$ & Cell $4$ & Cell $5$\\
+        Exponential Fit & 0.00703 & 0.00540 & 0.01159 & 0.00756 & 0.01024 & 0.00836\\
+        % \multicolumn{7}{l}{\textbf{Direct Position-Space Metric}}\\
         \midrule
-        Fit             & 0.00703 & 0.00540 & 0.01159 & 0.00756 & 0.01024 & 0.00836\\
-        \ac{abm} Direct & 0.00671 & 0.00746 & 0.01138 & 0.00848 & 0.02005 & 0.00952\\
-        \textbf{Imaging values will be adjusted!}\\
-        \ac{abm} Imaging& 0.00104 & 0.00146 & 0.00206 & 0.00286 & 0.00190 & 0.00171\\
-        & \textbf{Daughters}\\
+        Direct \ac{abm} Mie    & 0.00671 & 0.00746 & 0.01138 & 0.00848 & 0.02005 & 0.00952\\
+        Direct \ac{abm} Morse  & \textbf{TODO}\\
+        % \multicolumn{7}{l}{\textbf{Lineage-Aware Image-Space Metric (TODO)}}\\
         \midrule
-        & Cell $0$ & Cell $1$ & Cell $2$ & Cell $3$ & Cell $4$ & Cell $5$\\
-        & 0.00240 & 0.00494 & 0.00096 & 0.00363\\
-        & 0.00490 & 0.00226 & 0.00448 & 0.00048\\
+        Imaging \ac{abm} Mie & 0.00104 & 0.00146 & 0.00206 & 0.00286 & 0.00190 & 0.00171\\
+        % & \textbf{Daughters}\\
+        % \midrule
+        % & Cell $0$ & Cell $1$ & Cell $2$ & Cell $3$ & Cell $4$ & Cell $5$\\
+        Daughter 0 & 0.00240 & 0.00494 & 0.00096 & 0.00363 & - & -\\
+        Daughter 1 & 0.00490 & 0.00226 & 0.00448 & 0.00048 & - & -\\
         % 0.0014057073 & 0.0010807540 & 0.0023177727 & 0.0015120675 & 0.0020488179 & 0.0016726263\\

### 2025-12-16 cr_mech_coli/crm_fit/plotting.py
add optional label argument

@@ -82,6 +82,7 @@ def plot_profile_from_data(
     displacement_error,
     ls_color=crm.plotting.COLOR3,
     fill=True,
+    label=None,
 ):
     # Extend x and y by values from final_params and final cost
     x = np.append(x, p_fixed)
@@ -140,6 +141,7 @@ def plot_profile_from_data(
         y,
         color=ls_color,  # crm.plotting.COLOR3,
         linestyle="--",
+        label=label,
     )
 
     upper = np.min([4 * thresh_prev, 1.05 * np.max([np.max(y), thresh_prev])])
@@ -207,7 +209,7 @@ def plot_profile(
 
     savename = name.strip().lower().replace(" ", "-")
     try:
-        x = np.load(odir / f"profile-{savename}-params")
+        x = np.load(odir / f"profile-{savename}-params.npy")
         y = np.loadtxt(odir / f"profile-{savename}")
     except:
         x = np.linspace(bound_lower, bound_upper, pyargs.profiles_samples)

### 2025-12-16 paper/main.tex
reference figures at new path

@@
     \begin{tikzonimage}[width=0.49\textwidth]
-        {figures/crm_fit/mie/snapshots/predicted-003889.png}%
+        {figures/crm_fit/mie_all/snapshots/predicted-003889.png}%
         \node at (0.025, 0.975)[anchor=north west, white]{\textbf{A}};
@@
     \begin{tikzonimage}[width=0.49\textwidth]
-        {figures/crm_fit/mie/celldiffs/diff-003889.png}%
+        {figures/crm_fit/mie_all/celldiffs/diff-003889.png}%
         \node at (0.025, 0.975)[anchor=north west, white]{\textbf{B}};
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_fit/mie/profiles/damping.pdf}%
+        {figures/crm_fit/mie_all/profiles/damping.pdf}%
         \node at (0, 1.1)[anchor=north west, black]{\textbf{A}};
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_fit/mie/profiles/exponent-m.pdf}%
+        {figures/crm_fit/mie_all/profiles/exponent-m.pdf}%
         \node at (0, 1.1)[anchor=north west, black]{\textbf{B}};
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_fit/mie/profiles/exponent-n.pdf}%
+        {figures/crm_fit/mie_all/profiles/exponent-n.pdf}%
         \node at (0, 1.1)[anchor=north west, black]{\textbf{C}};
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_fit/mie/profiles/strength.pdf}%
+        {figures/crm_fit/mie_all/profiles/strength.pdf}%
         \node at (0, 1.1)[anchor=north west, black]{\textbf{D}};
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_fit/mie/profiles/radius.pdf}%
+        {figures/crm_fit/mie_all/profiles/radius.pdf}%
         \node at (0, 1.1)[anchor=north west, black]{\textbf{E}};
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_fit/mie/profiles/growth-rate-0.pdf}%
+        {figures/crm_fit/mie_all/profiles/growth-rate-0.pdf}%
         \node at (0, 1.1)[anchor=north west, black]{\textbf{F}};
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_fit/mie/profiles/growth-rate-1.pdf}%
+        {figures/crm_fit/mie_all/profiles/growth-rate-1.pdf}%
         \node at (0, 1.1)[anchor=north west, black]{\textbf{G}};
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_fit/mie/profiles/growth-rate-2.pdf}%
+        {figures/crm_fit/mie_all/profiles/growth-rate-2.pdf}%
         \node at (0, 1.1)[anchor=north west, black]{\textbf{H}};
@@
     \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/mie/profiles/growth-rate-0.pdf}%
+        {figures/crm_fit/mie_all/profiles/growth-rate-0.pdf}%
     \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/mie/profiles/growth-rate-1.pdf}%
+        {figures/crm_fit/mie_all/profiles/growth-rate-1.pdf}%
     \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/mie/profiles/growth-rate-2.pdf}\\%
+        {figures/crm_fit/mie_all/profiles/growth-rate-2.pdf}\\%
     \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/mie/profiles/growth-rate-3.pdf}%
+        {figures/crm_fit/mie_all/profiles/growth-rate-3.pdf}%
     \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/mie/profiles/growth-rate-4.pdf}%
+        {figures/crm_fit/mie_all/profiles/growth-rate-4.pdf}%
     \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/mie/profiles/growth-rate-5.pdf}
+        {figures/crm_fit/mie_all/profiles/growth-rate-5.pdf}
     \caption{Parameter estimation of growth rates for individual cells}%TODO extend caption}

### 2025-12-16 paper/figures/crm_fit/plot_profiles.py
extend comparison of profiles script

@@ -7,14 +7,11 @@ import cr_mech_coli as crm
 from cr_mech_coli import crm_fit
 
 
-def plot_potential_single(n, name, short, units, path, ax, color=crm.plotting.COLOR3):
+def plot_potential_single(
+    n, name, short, units, path, ax, label, color=crm.plotting.COLOR3
+):
     savename = name.strip().lower().replace(" ", "-")
     y = np.loadtxt(path / f"profiles/profile-{savename}")
-
-    settings = crm_fit.Settings.from_toml(path / "settings.toml")
-    optimization_result = crm_fit.OptimizationResult.load_from_file(
-        path / "final_params.toml"
-    )
 
     settings = crm_fit.Settings.from_toml(path / "settings.toml")
     optim_res = crm_fit.OptimizationResult.load_from_file(path / "final_params.toml")
@@ -26,10 +23,12 @@ def plot_potential_single(n, name, short, units, path, ax, color=crm.plotting.CO
     bound_lower = infos.bounds_lower[n]
     bound_upper = infos.bounds_upper[n]
 
+    param_path = path / f"profiles/profile-{savename}-params"
     try:
-        x = np.load(path / f"profiles/profile-{savename}-params")
+        x = np.load(param_path)
     except:
         x = np.linspace(bound_lower, bound_upper, len(y))
+        np.save(x, param_path)
 
     crm_fit.plot_profile_from_data(
         x,
@@ -43,20 +42,21 @@ def plot_potential_single(n, name, short, units, path, ax, color=crm.plotting.CO
         displacement_error,
         ls_color=color,
         fill=False,
+        label=label,
     )
 
     return ax, x, (y - final_cost) / displacement_error**2
 
 
-def plot_single_profile_combination(name, short, units, n1_p1, n2_p2, odir):
+def plot_single_profile_combination(name, short, units, n1_p1_l1, n2_p2_l2, odir):
     crm.plotting.set_mpl_rc_params()
     fig, ax = plt.subplots(figsize=(8, 8))
     crm.plotting.configure_ax(ax)
 
     x1 = []
     y1 = []
-    if type(n1_p1) is tuple:
-        n1, p1 = n1_p1
+    if type(n1_p1_l1) is tuple:
+        n1, p1, label1 = n1_p1_l1
         res1 = plot_potential_single(
             n1,
             name,
@@ -64,6 +64,7 @@ def plot_single_profile_combination(name, short, units, n1_p1, n2_p2, odir):
             units,
[…]

### 2025-12-20 cr_mech_coli/crm_divide.py
adjust all masks (not only the later ones)

@@ -93,7 +93,7 @@ def adjust_masks(
     # Tranform the data masks with the above mapping
     # WARNING MAGIC NUMBERS
     masks_data_new = [np.array(m) for m in masks_data]
-    for m in masks_data_new[8:]:
+    for m in masks_data_new:
         colors = list(sorted(np.unique(m)))[1:]
         for c in colors:
             m[m == c] = align_mask_data_color[c]

### 2025-12-20 cr_mech_coli/crm_divide.py
use 15 pixel per micron as conversion unit

@@ -384,7 +384,7 @@ def objective_function(
             container.get_cells_at_iteration(iter),
             cell_to_color,
             settings.constants.domain_size,
-            render_settings=crm.RenderSettings(pixel_per_micron=1),
+            render_settings=crm.RenderSettings(pixel_per_micron=15),
         )
         for iter in tqdm(
             iterations_simulation if return_all else iters_filtered,

### 2025-12-20 cr_mech_coli/crm_divide.py
pick new parameters to visualizae mask adjustment

@@ -494,20 +494,20 @@ def preprocessing(n_masks=None):
 def plot_mask_adjustment(
     output_dir, masks_data, positions_all, settings, iterations_data
 ):
-    radius = 0.5
-    strength = 0.5
-    en = 0.50215733
-    em = 0.21933548
-    damping = 1.0
+    radius = 0.4782565
+    strength = 0.02
+    en = 2.1887856
+    em = 2.2355218
+    damping = 2.0
     growth_rates = [
-        0.016,
-        0.016,
-        0.016,
-        0.016,
-        0.016,
-        0.016,
+        0.005995107,
+        0.0068584173,
+        0.0070885874,
+        0.009034319,
+        0.007861354,
+        0.008311217,
     ]
-    spring_length_thresholds = [15] * 4
+    spring_length_thresholds = [0.8, 0.8, 0.7, 0.7]
     new_growth_rates = [0.001] * 8
     x0 = [
         radius,

### 2025-12-20 cr_mech_coli/crm_divide.py
pick new optimization parameters for crm_divide

@@ -1030,34 +1030,34 @@ def crm_divide_main():
         if pyargs.only_mask_adjustment:
             exit()
 
-    radius = 8.059267
-    strength = 10.584545
-    en = 0.50215733
-    em = 0.21933548
-    damping = 2.5799131
+    radius = 0.4782565
+    strength = 0.584545
+    en = 2.1887856
+    em = 2.2355218
+    damping = 1.0
     growth_rates = [
-        0.001152799,
-        0.001410604,
-        0.0018761827,
-        0.0016834959,
-        0.0036106023,
-        0.0015209642,
+        0.005995107,
+        0.0068584173,
+        0.0070885874,
+        0.009034319,
+        0.007861354,
+        0.008311217,
     ]
     spring_length_thresholds = [
-        9.405841188088112759e00,
-        1.103879179742345329e01,
-        9.277040994277374608e00,
-        7.600778468661159692e00,
+        0.8,
+        0.8,
+        0.8,
+        0.8,
     ]
     new_growth_rates = [
-        2.186928453188154847e-03,
-        2.186928453188154847e-03,
-        4.834987621055546192e-04,
-        4.834987621055546192e-04,
-        1.669060863394238470e-03,
-        1.669060863394238470e-03,
-        2.300937616285135622e-03,
-        2.300937616285135622e-03,
+        0.016,
+        0.016,
+        0.016,
+        0.016,
+        0.016,
+        0.016,
+        0.016,
+        0.016,
     ]
     x0 = [
         radius,
@@ -1071,15 +1071,15 @@ def crm_divide_main():
     ]
     bounds = (
         # Radius, Strength, en, em, Damping
-        [(0.05, 1.0), (0.0, 60.0), (0.0, 12.5), (0.1, 4.0), (0.0, 30.0)]
+        [(0.003, 1.0), (0.0, 0.6), (0.0, 10.0), (0.0, 10.0), (0.0, 60.0)]
         # Growth rates
-        + [(0.0000, 0.010)] * 6
+        + [(0.0000, 0.1)] * 6
         # Spring length thresholds
-        + [(4, 15)] * 4
+        + [(0.4, 1.2)] * 4
         # new growth rates
-        + [(0.0000, 0.010)] * 8
+        + [(0.0000, 0.1)] * 8
     )
-    parent_penalty = 0.5
+    parent_penalty = 1 / 3
     args = (
         positions_all,
         settings,

### 2025-12-20 cr_mech_coli/crm_fit/plotting.py
use filter for when optimizer was not successful

@@ -67,7 +67,7 @@ def optimize_around_single_param(opt_args):
             "maxiter": pyargs.profiles_maxiter,
         },
     )
-    return res.fun
+    return res.fun, res.success
 
 
 def fill_confidence_levels(x, y, ax, fill=True, thresholds=[0.68, 0.90, 0.95]):
@@ -185,7 +185,9 @@ def calculate_profile(n, x, name, n_workers, optimization_result, infos, args, p
         desc=f"Profile: {name}",
         max_workers=n_workers,
     )
-    return y
+    y_values = [yi[0] for yi in y]
+    filt = [yi[1] for yi in y]
+    return np.array(y_values), np.array(filt)
 
 
 def plot_profile(
@@ -222,11 +224,12 @@ def plot_profile(
         y = np.loadtxt(odir / f"profile-{savename}")
     except:
         x = np.linspace(bound_lower, bound_upper, pyargs.profiles_samples)
-        y = calculate_profile(
+        y, filt = calculate_profile(
             n, x, name, n_workers, optimization_result, infos, args, pyargs
         )
         np.savetxt(odir / f"profile-{savename}", y)
         np.save(odir / f"profile-{savename}-params", x)
+        np.save(odir / f"profile-{savename}-filter", filt)
 
     final_params = optimization_result.params
     final_cost = optimization_result.cost

### 2025-12-20 paper/main.tex
use new figures; update caption; add optimization progression to supplementary material

--- 2025-12-11 paper/main.tex
@@
     \begin{tikzonimage}[width=0.49\textwidth]
-        {figures/crm_fit/mie/snapshots/predicted-003889.png}%
+        {figures/crm_fit/morse_partial/snapshots/predicted-006667.png}%
         \node at (0.025, 0.975)[anchor=north west, white]{\textbf{A}};
@@
     \begin{tikzonimage}[width=0.49\textwidth]
-        {figures/crm_fit/mie/celldiffs/diff-003889.png}%
+        {figures/crm_fit/morse_partial/celldiffs/diff-006667.png}%
         \node at (0.025, 0.975)[anchor=north west, white]{\textbf{B}};
@@
     \caption{
+        Predicted results by the reduced model with Morse-type interaction.
         (A) Final microscopic image of the time series with the predicted positions of the cells
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_fit/mie/profiles/damping.pdf}%
+        {figures/crm_fit/profiles/profile-damping.pdf}%
         \node at (0, 1.1)[anchor=north west, black]{\textbf{A}};
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_fit/mie/profiles/exponent-m.pdf}%
+        {figures/crm_fit/profiles/profile-potential-stiffness.pdf}%
         \node at (0, 1.1)[anchor=north west, black]{\textbf{B}};
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_fit/mie/profiles/exponent-n.pdf}%
+        {figures/crm_fit/profiles/profile-exponent-n.pdf}%
         \node at (0, 1.1)[anchor=north west, black]{\textbf{C}};
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_fit/mie/profiles/strength.pdf}%
+        {figures/crm_fit/profiles/profile-exponent-m.pdf}%
         \node at (0, 1.1)[anchor=north west, black]{\textbf{D}};
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_fit/mie/profiles/radius.pdf}%
+        {figures/crm_fit/profiles/profile-strength.pdf}%
         \node at (0, 1.1)[anchor=north west, black]{\textbf{E}};
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_fit/mie/profiles/growth-rate-0.pdf}%
+        {figures/crm_fit/profiles/profile-radius.pdf}%
         \node at (0, 1.1)[anchor=north west, black]{\textbf{F}};
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_fit/mie/profiles/growth-rate-1.pdf}%
+        {figures/crm_fit/profiles/profile-growth-rate-0.pdf}%
         \node at (0, 1.1)[anchor=north west, black]{\textbf{G}};
@@
     \begin{tikzonimage}[width=0.25\textwidth]
-        {figures/crm_fit/mie/profiles/growth-rate-2.pdf}%
+        {figures/crm_fit/profiles/profile-growth-rate-1.pdf}%
         \node at (0, 1.1)[anchor=north west, black]{\textbf{H}};
+    \end{tikzonimage}\\
+    \begin{tikzonimage}[width=0.25\textwidth]
+        {figures/crm_fit/profiles/profile-growth-rate-2.pdf}%
+        \node at (0, 1.1)[anchor=north west, black]{\textbf{I}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.25\textwidth]
+        {figures/crm_fit/profiles/profile-growth-rate-3.pdf}%
+        \node at (0, 1.1)[anchor=north west, black]{\textbf{J}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.25\textwidth]
+        {figures/crm_fit/profiles/profile-growth-rate-4.pdf}%
+        \node at (0, 1.1)[anchor=north west, black]{\textbf{K}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.25\textwidth]
+        {figures/crm_fit/profiles/profile-growth-rate-5.pdf}%
+        \node at (0, 1.1)[anchor=north west, black]{\textbf{L}};
     \end{tikzonimage}%
     \caption{
-        (A-H) Likelihood profiles of optimized parameters.
+        (A-L) Likelihood profiles for 4 scenarios: Full model with Morse/Mie interaction and with a
+        reduced model with Morse/Mie interaction, setting the damping constant
+        $\lambda=\SI{1}{\per\minute}$.
         Bands of blue and yellow represent confidence intervals of $68\%,90\%$ and $95\%$.
-        \textbf{TODO adapt this description to new results}
-        Every plot indicates that a minimum was achieved, thus every parameter could be identified.
-        We show only one of the 6 individually-chosen growth rate as an example.
-        The remaining plots show similar behavior although with varying values and can be found in
-        the supplementary material (Section~\ref{sec:supplement-parameter-profiles}).
+        (A) Damping constant before model reduction.
+        (B) Potential Stiffness $\omega$ of the Morse potential.
+        (C-D) Exponents $n,m$ of the Mie potential.
+        (E) Strength $V_0$ of the interaction potential.
+        The x-axis is split in the middle to display a narrow and wider range for lower and higher
+        parameter values.
+        (F) Radius $R$ (thickness of the rod) of the potential.
+        (G-L) Individual growth rates $\mu$ of the 6 bacterial agents.
     }
@@
 
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.5\textwidth]{figures/crm_fit/optimization-progression.pdf}
+    \caption{TODO}
+\end{figure}
+
 %TODO {explain profiles}
@@
     \centering
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/mie/profiles/growth-rate-0.pdf}%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/mie/profiles/growth-rate-1.pdf}%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/mie/profiles/growth-rate-2.pdf}\\%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/mie/profiles/growth-rate-3.pdf}%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/mie/profiles/growth-rate-4.pdf}%
-    \includegraphics[width=0.33\textwidth]
-        {figures/crm_fit/mie/profiles/growth-rate-5.pdf}
+    % \includegraphics[width=0.33\textwidth]
+    %     {figures/crm_fit/morse_profiles/profile-growth-rate-0.pdf}%
+    % \includegraphics[width=0.33\textwidth]
+    %     {figures/crm_fit/morse_profiles/profile-growth-rate-1.pdf}%
+    % \includegraphics[width=0.33\textwidth]
+    %     {figures/crm_fit/morse_profiles/profile-growth-rate-2.pdf}\\%
+    % \includegraphics[width=0.33\textwidth]
+    %     {figures/crm_fit/morse_profiles/profile-growth-rate-3.pdf}%
+    % \includegraphics[width=0.33\textwidth]
+    %     {figures/crm_fit/morse_profiles/profile-growth-rate-4.pdf}%
+    % \includegraphics[width=0.33\textwidth]
+    %     {figures/crm_fit/morse_profiles/profile-growth-rate-5.pdf}
     \caption{Parameter estimation of growth rates for individual cells}%TODO extend caption}

### 2025-12-20 paper/main.tex
adjust notation
clarify that this is a phenomenological approach

@@
 system.
+We follow a phenomenological approach thus aiming to fit our model to given datapoints and determine
+the type of interaction in this way.
+In order to accurately depict the physical size of the rod, our interactions will be repulsive for
+short ranges $r<R_1+R_2$ and attractive at longer ranges.
 The Mie potential~\cite{Mie1903} is capable of describing a wide range of interaction
@@
 \end{align}
-where we absorbed the terms $\log(\sigma)$ in the sum into the constant term.
+where we absorbed the terms $2\log(\sigma_{ijkl})$ in the sum into the constant term.
 With this simplification, it is important to pick a value for $\sigma$ such that the real

### 2025-12-23 paper/main.tex
update table with new numerical values and updated format

@@
     \centering
-    \textbf{Growth Parameters obtained by various Methods}
+    % \textbf{Growth Parameters obtained by various Methods}
     \begin{tabular}{l c c c c c c}
         % \multicolumn{7}{l}{}\\
-        % \multicolumn{6}{Growth Rates}\\
-        & Cell $0$ & Cell $1$ & Cell $2$ & Cell $3$ & Cell $4$ & Cell $5$\\
+        & \multicolumn{6}{c}{Growth Rates $[\mu]=\SI{}{\per\minute}$}\\
+        \textbf{Data Extraction} & Cell $0$ & Cell $1$ & Cell $2$ & Cell $3$ & Cell $4$ & Cell $5$\\
         \toprule
-        Exponential Fit & 0.00703 & 0.00540 & 0.01159 & 0.00756 & 0.01024 & 0.00836\\
-        % \multicolumn{7}{l}{\textbf{Direct Position-Space Metric}}\\
+        % \multicolumn{7}{l}{\textbf{Data Extraction}}\\
+        % \midrule
+        Exponential Fit
+            & 0.0070
+            & 0.0054
+            & 0.0116
+            & 0.0076
+            & 0.0102
+            & 0.0084\\
+
+        \multicolumn{7}{l}{\textbf{Direct Position-Space Metric}}\\
         \midrule
-        Direct \ac{abm} Mie    & 0.00671 & 0.00746 & 0.01138 & 0.00848 & 0.02005 & 0.00952\\
-        Direct \ac{abm} Morse  & \textbf{TODO}\\
-        % \multicolumn{7}{l}{\textbf{Lineage-Aware Image-Space Metric (TODO)}}\\
+        Mie
+            & 0.0079
+            & 0.0090
+            & 0.0131
+            & 0.0103
+            & 0.0215
+            & 0.0097\\
+
+        Mie $\lambda=\SI{1}{\per\minute}$
+            & 0.0060
+            & 0.0069
+            & 0.0109
+            & 0.0080
+            & 0.0186
+            & 0.0083\\
+
+        % 0.00671 & 0.00746 & 0.01138 & 0.00848 & 0.02005 & 0.00952\\
+        Morse
+            & 0.0070
+            & 0.0078
+            & 0.0118
+            & 0.0089
+            & 0.0203
+            & 0.0094\\
+
+        Morse $\lambda=\SI{1}{\per\minute}$
+            & 0.0060
+            & 0.0070
+            & 0.0113
+            & 0.0078
+            & 0.0191
+            & 0.0093\\
+
+        \multicolumn{7}{l}{\textbf{Lineage-Aware Image-Space Metric (TODO)}}\\
         \midrule

### 2025-12-26 paper/figures/crm_fit/plot_profiles.py
correct order of arguments

@@ -141,7 +141,7 @@ def plot_all_profiles_combined(*args, odir, bounds={}):
 
     infos_combined = {}
     for (p, label, _), (info, kwargs) in zip(args, infos):
-        for n, (name, short, units) in enumerate(info.parameter_infos):
+        for n, (name, units, short) in enumerate(info.parameter_infos):
             if name in infos_combined:
                 infos_combined[name].append((n, p, label, name, short, units, kwargs))
             else:

### 2025-12-26 paper/main.tex
fixup! also extract data from pixel based approach; results in subfolders

@@
     \begin{tikzonimage}[width=0.35\textwidth]
-        {figures/crm_estimate_params/IWF-Goettingen/rod-lengths-individual.pdf}%
+        {figures/crm_estimate_params/IWF-Goettingen/positions/rod-lengths-individual.pdf}%
         \node at (0, 1)[anchor=north west, black]{\textbf{D}};
@@
     \begin{tikzonimage}[width=0.35\textwidth]
-        {figures/crm_estimate_params/IWF-Goettingen/growth-rates-bar-plot.pdf}%
+        {figures/crm_estimate_params/IWF-Goettingen/positions/growth-rates-bar-plot.pdf}%
         \node at (0, 1)[anchor=north west, black]{\textbf{E}};

### 2025-12-26 paper/main.tex
add discussion of parameter estimation results

@@
 parameter.
-\textbf{TODO adapt these last sentences to new results}
-We can see that all parameters can be estimated by our optimization approach and that no
-non-identifiability~\cite{Raue2009} is present.
-These plots also show that the individual growth rates of the bacteria can be estimated effectively.
+We compare four distinct models, covering Morse and Mie-type interactions with fixed damping
+constant $\lambda=\SI{1}{\per\minute}$.
+Subplot (A) displays the profile-likelihood for the damping constant $\lambda$ and exhibits a wide
+range of permissible parameter values which motivates the approach to treat this parameter as a
+constant thus reducing our overall model.
+The parameters which characterize the Morse-type interaction are the potential stiffness (B),
+strength (E) and radius (F).
+While the potential stiffness (B) and radius (F) can be estimated well with the given data, the
+potential strength (E) contains a practical unidentifiability~\cite{Raue2009} which is lifted when
+fixing the damping constant $\lambda=\SI{1}{\per\minute}$.
+In contrast to the Morse potential, the Mie potential uses two parameters $n,m$ (C,D) to determine
+the shape of the interaction potential which allows it to describe a larger class of curves.
+However, as subplots (A,C,D,E) show, neither the full Mie-type interaction nor a reduced model with
+$\lambda=\SI{1}{\per\minute}$ allow for the estimation of the exponents $n,m$.
+Although a minimum could be determined and the steepness of the enclosing profile towards the larger
+values of exponents $n,m$ increases for the reduced model, the confidence intervals clearly show a
+practical non-identifiability.
+Since the reduction of the model improves the situation we are hopeful that an increased data set
+could be suitable to estimate these parameters.
+Notably, the two interaction types predict differing radii which are still within the $68\%$
+confidence interval of each other.
+The growth rates (G-L) can all be estimated for any interaction type but the non-reduced Mie-type
+interaction shows wider confidence intervals, thus greater uncertainty in the predicted parameter
+value.
 

### 2025-12-26 paper/main.tex
reformat keywords
rewrite abstract of paper

@@
 
-\textbf{Discussed Keywords (Working Title)}
-\begin{itemize}
-    \item Parameter Estimation
-    \begin{itemize}
-        \item Mechanical Properties
-        \item Rod-shaped?
-        \item Single-cell
-    \end{itemize}
-    \item \ac{abm}
-    \item Methods
-\end{itemize}
+\textbf{Discussed Keywords (Working Title)}\\
+Parameter Estimation (Mechanical Properties, Rod-shaped, Single-cell), \ac{abm}, Methods
 % keywords can be removed
@@
 rods capable of growth, elastic bending, and mechanical interactions.
-We develop a set of imaging-based and geometry-based comparison methods that extract cell shape and
-position information from segmented microscopy data and quantitatively compare it with model
-simulations at the single-cell level.
-These methods enable the estimation of model parameters across multiple experimental scenarios of
-increasing complexity, including cases that exhibit substantial cell-to-cell variability.
-Our results demonstrate that parameter estimation for complex agent-based models of flexible
-bacteria is feasible and can yield biologically interpretable constraints on cellular mechanics.
-The modeling and analysis framework is modular, generalizable, and readily adaptable to a broad
-range of rod-shaped species.
-By integrating our mechanical model into an extensible computational platform, we provide tools that
-support more rigorous model construction, systematic calibration, and mechanistic interpretation.
-This work establishes a foundation for quantitatively studying the mechanical behavior of
-filamentous bacteria and paves the way for future applications ranging from biofilm formation to
-multicellular
-patterning.
+We develop a novel set of imaging- and geometry-based methods that extract cell shape and positional
+information from segmented microscopy data and quantitatively compare it with model simulations at
+the single-cell level.
+These methods allow for a detailed analysis of the underlying dynamics of the model.
+They allow us to apply maximum-likelihood and profile-likelihood estimation techniques to determine
+the parameters of the underlying model.
+We apply them across multiple experimental scenarios of increasing complexity, including cases that
+exhibit substantial cell-cell variability and show how these methods can be used to perform model
+validation and reduction.
+Our results demonstrate that direct parameter estimation with cellular heterogeneity for complex
+agent-based models of bacteria is feasible and can yield biologically interpretable constraints on
+cellular parameters.
+The developed are readiliy available as a modular Python package backed by a high-performing backend
+written in Rust and integrate into our previously devleped agent-based simulation framework
+\texttt{cellular\_raza}.
 

### 2025-12-26 paper/main.tex
add and fix citations

@@
 Only a few \acp{abm} support even the less complex cylindrical
-shapes~\cite{Kang2014,breitwieser_biodynamo_2022,Gorochowski2012} but none are able to capture
+shapes~\cite{Kang2014,Breitwieser2021,Gorochowski2012} but none are able to capture
 flexible filamentous bacteria.
@@
 distance valuels corresponding to overlaps.
-In contrast, a popular alternative is the Morse potential~\cite{Morse1929}, which is less
+In contrast, an alternative which has been used by other \acp{abm} is the Morse
+potential~\cite{Morse1929,Kang2014,Breitwieser2021}, which is less
 customizable due to the lower number of parameters and more restrictive mathematical formulation

### 2025-12-27 cr_mech_coli/crm_multilayer/main.py
add initial structure for arguments and subcommands

@@ -236,8 +236,27 @@ def plot_colony_height(ml_config: MultilayerConfig, container):
 
 
 def crm_multilayer_main():
+    parser = argparse.ArgumentParser(
+        prog="crm_multilayer",
+        description="Run Simulations to analyze Multilayer-behaviour of Rod-Shaped Bacteria.",
+    )
+    subparsers = parser.add_subparsers()
+    parse_run = subparsers.add_parser(
+        "run", help="Run simulation for specified parameters"
+    )
+    parse_plot = subparsers.add_parser("plot", help="Perform plotting actions")
+
+    parse_plot.add_argument("colony-height", action="store_true", default=False)
+    # parse_plot.add_argument("", action="store_true", default=False)
+
+    # parser.add_argument("--plot-snapshots", action="store_true")
+    # parser.add_argument("--seeds", nargs="+", default=[0, 1, 2, 3], type=int)
+    pyargs = parser.parse_args()
+    # pyargs.seeds = [int(n) for n in pyargs.seeds]
+
     # ml_config = MultilayerConfig.load_from_toml_file(Path(file_path))
     ml_config = produce_ml_config()
+    ml_config.config.progressbar = "Run Simulation"
 
     container = load_or_compute(ml_config)
 

### 2025-12-27 cr_mech_coli/crm_multilayer/plotting.py

@@ -45,15 +45,6 @@ def render_image_helper(args):
 
 
 def plot_colony_height_over_time():
-    # parser = argparse.ArgumentParser(
-    #     prog="crm_multilayer",
-    #     description="Run Simulations to analyze Multilayer-behaviour of Rod-Shaped Bacteria.",
-    # )
-    # parser.add_argument("--plot-snapshots", action="store_true")
-    # parser.add_argument("--seeds", nargs="+", default=[0, 1, 2, 3], type=int)
-    # pyargs = parser.parse_args()
-    # pyargs.seeds = [int(n) for n in pyargs.seeds]
-
     ml_config = produce_ml_config()
 
     def create_new_ml_configs(ml_config, seeds):

### 2025-12-27 cr_mech_coli/crm_multilayer/runner.py
refactor generation of MultilayerConfig

@@ -4,53 +4,63 @@ import numpy as np
 from glob import glob
 
 from cr_mech_coli.crm_multilayer import MultilayerConfig
+from cr_mech_coli import GrowthRateSetter
 
 
 def produce_ml_config() -> MultilayerConfig:
-    # Create many Multilayer-Configs
+    """
+    Produces a :class:`MultilayerConfig` with default parameters.
+    """
     ml_config = crm.crm_multilayer.MultilayerConfig()
+
+    # TIME SETTINGS
     ml_config.config.dt = 0.05
-    # ml_config.config.t_max = 20
-    # ml_config.config.n_saves = int(
-    #     np.ceil(ml_config.config.t_max / (ml_config.config.dt * 100))
-    # )
-    ml_config.config.domain_height = 20.0
-    # ml_config.config.domain_size = (1600, 1600)
-    # ml_config.dx = (700, 700)
-    ml_config.config.n_voxels = (10, 10)
-    ml_config.config.gel_pressure = 0.05
-    ml_config.config.n_threads = 1
-
-    ml_config.config.surface_friction = 0.3
-    ml_config.config.surface_friction_distance = (
-        ml_config.agent_settings.interaction.radius / 10
-    )
-
-    ml_config.agent_settings.neighbor_reduction = (200, 0.5)
-
     ml_config.config.t_max = 1800
     ml_config.config.dt = 0.025
     ml_config.config.n_saves = 19
-    ml_config.config.domain_size = (400, 400)
-    ml_config.config.rng_seed = 0
-    ml_config.dx = (100, 100)
-    ml_config.agent_settings.growth_rate = 0.005
-    ml_config.agent_settings.growth_rate_setter = GrowthRateSetter.NormalDistr(
-        0.005, 0.001
-    )
-    ml_config.agent_settings.mechanics.damping = 0.02
-    ml_config.agent_settings.mechanics.diffusion_constant = 0.03
-    ml_config.agent_settings.mechanics.rigidity = 1.0
-    ml_config.agent_settings.mechanics.spring_tension = 0.3
-    ml_config.agent_settings.interaction.strength = 0.02
-    ml_config.agent_settings.spring_length_threshold = 20.0
-    ml_config.config.progressbar = ""
 
+    # SOLVER SETTINGS
+    ml_config.config.n_threads = 1
+    ml_config.config.rng_seed = 0
+    ml_config.config.progressbar = None
[…]
